### 2021-02-01 paper/mesa.tex
Added template and content for paper

\documentclass[twocolumn,showpacs,preprintnumbers,nofootinbib,prd,
superscriptaddress,10pt]{revtex4-1}

\usepackage{amsmath,amssymb}
\usepackage{amsfonts}
\usepackage[normalem]{ulem}
\usepackage{textcomp}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{bm}
\usepackage{afterpage}
\usepackage{graphicx}
\usepackage{psfrag}
\usepackage{mathtools}
%\usepackage{tensor}
\usepackage{layouts}
%\usepackage{DejaVuSans}
\usepackage{epstopdf}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{ragged2e}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{siunitx}
	\sisetup{output-decimal-marker={.}}
\newtheorem{theorem}{Theorem}

	%some math symbols
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\DeclareMathOperator{\sign}{sign}
\renewcommand{\d}[1]{\ensuremath{\operatorname{d}\!{#1}}}
%argmin and argmax
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% comments command
\newcommand{\an}[1]{{\textcolor{orange}{\texttt{AN: #1}} }}
\newcommand{\pr}[1]{{\textcolor{cyan}{\texttt{PR: #1}} }}
\newcommand{\fm}[1]{{\textcolor{blue}{\texttt{FM: #1}} }}
\newcommand{\sschmidt}[1]{{\textcolor{red}{\texttt{SS: #1}} }}
\newcommand{\BS}[1]{{\textcolor{green}{\texttt{SB: #1}} }}
\newcommand{\oldnewtxt}[2]{\sout{#1}\textcolor{red}{#2}}

\begin{document}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT
\begin{abstract}
Burg method of Maximum Entropy Spectral Analysis (MESA) provides a powerful tool to perform spectral estimation of a time-series. The method is a generalization of Jaynes maximum entropy principle and provides the estimate for the spectrum in terms of the coefficients of some autoregressive order p process (AR(p)). The recursive solution requires a method for the estimate of the order $p$ for the process. Several methods are characterized on synthetic noise having a known power spectral density (PSD) via the computation of the percentage error and then compared.

\end{abstract}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE
	\title{Maximum Entropy Spectral Analysis}
	\author{Alessandro \surname{Martini}}
		\email{martini.alessandr@gmail.com}
        \affiliation{Dipartimento di Fisica  Universit\`a di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}        
        %

	\maketitle

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BODY

\section{Maximum Entropy Principle}

Jaynes Maximum Entropy principle (MAXENT) \cite{JaynesArticle}\cite{jaynes2003ptl} represents one of the most important results in Bayesian Statistics: it provides a way to uniquely assign probabilities to a phenomenon in a way that best represent our state of knowledge, while at the same time it is non committal with unavailable information. It is formulated as a variational problem in which Shannon's entropy \cite{Shannon}
\begin{equation}\label{eq:shannonEntropy}
H[p_1, \dots, p_N] = - \sum_{i = 1}^{N}p_i\log(p_i)
\end{equation}
or, in the continuous case
\begin{equation}
H[p(x)] = - \int dx p(x)\log p(x)
\end{equation}
is the functional to be maximized. The constraints on the variational problem are given by the prior knowledge $I$ and are usually represented as fixed values for various momenta of the probability distribution.  The most interesting result of MAXENT is obtained when a constraint on both mean and variance is considered. Consider for simplicity a vector $\vec x(t)$ with components in $\mathbb{R}^N$
\begin{equation*}
    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
\end{equation*}
We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals. If we put constraints in both mean $\vec \mu(t)$ and covariance matrix $C(t_i, t_j)$, MAXENT allow us to assign for $\vec x(t)$ a multivariate normal distribution \cite{gregory_2005}: 
\begin{equation}
    p\left(\vec x(t)\vert I\right) = 
    \frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp^{\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right)}. 
\end{equation}
With MAXENT, the often unjustified assumptions for the data to be normally distributed finds an elegant explanation. It is not only the central limit distribution for finite variance processes, it is also the distribution that maximizes the entropy and, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process. In some sense, we can interpret the central limit theorem as the natural 'statistical' evolution toward a configuration that maximizes entropy. 
Thanks to Burg, maximum entropy method has been generalized to compute the most unbiased estimate of the spectrum for a realization of some stochastic process. 
Since it contains the same information of a power spectral density, high resolution computation of a power spectrum from a single time series $\vec x(t)$ of length $N$ is a fundamental task in signal analysis.  \\
Since we are dealing with a finite number of samples $N$, it is not possible to have an accurate computation of the autocorrelation function for every time interval. Also, the error $\sigma$ in the estimate increases with the time interval $k$ as $\sigma \sim \frac{1}{\sqrt{N - k}}$, so that only a few values for the autocorrelation function can actually be computed. \\ 
The problem we want to address is how to compute a PSD from the partial knowledge of the autocorrelation function. \\
The standard approach of peridiograms \cite{Lomb} \cite{Scargle} consists in taking the Fourier Transform of the sample autocorrelation, written as \begin{equation}
    \rho = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\}.
\end{equation}
The sequence $W$ is a window function that can be chosen in several different way, each method having pros and cons for the final estimate of the power spectral density. \\ 
The choice of a window function is arbitrary and made by trial and error, until a satisfactory compromise between variance and resolution of the estimate of power spectral density  is reached. A high frequency resolution implies high variance and vice-versa. Another problem,is that the result depends on this choice, and this fact may raise the question of which we should consider as the "real" power spectral density. \\  
Another problem with this approach is the requirement for the window to be $0$ outside the interval: we are arbitrarily assuming $\rho_j = 0$ for $j > M$, and modifying the estimate if a non-rectangular window is chosen. Making assumptions on unavailable data and modifying the ones we have at our disposal does not seem a completely fair approach. \\
MAXENT gives us the possibility to develop a completely new way for the estimation of PSDs, without any a priori assumptions on the unavailable data. Our requirements (the constraints for the variational problem) are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation. \\
To do so, there is a problem we have to solve: our definition of entropy (equation \ref{eq:entropy}) depends on the probability distribution, not the power spectrum. We need to find a way to change the definition's domain for the entropy to formulate the variational problem in terms of the power spectral density $S(f)$ alone. A way out of this problem is considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA}. One can show that the 'entropy gain' obtained filtering white noise is: 
\begin{equation}\label{eq:EntropyGain}
    \Delta H = \int_{-\infty}^{\infty}\log S(f) df , 
\end{equation}
so that we can reformulate the variational problem requiring $S(f)$ to maximize the entropy gain. In real examples we have no access to all frequencies but only on those between minus and plus the Nyquist frequency, defined from the sampling interval $\Delta t$ as 
\begin{equation}
    Ny = \frac{1}{2 \Delta t}.
\end{equation}
All our quantities are well defined in this range only, so that we are not going to maximize \ref{eq:EntropyGain}, but
\begin{equation}\label{eq:EntropyGain2}
    \Delta H = \int_{- Ny}^{Ny}\log S(f) df.
\end{equation}
Defining $\bar r_n$ as the sample autocorrelation, our constraints will be
\begin{equation}\label{eq:MaxConstraint}
    \int_{-Ny}^{Ny} S(f) e^{\imath 2 \pi f n \Delta t} df = \bar r_{n}.
\end{equation}
This approach on power spectral density estimate was developed by Burg \cite{burg1975maximum} and it is a way out from the problems raised in periodograms estimate. It requires no assumptions on the unavailable estimates for $\bar r_n$ and is consistent with the sample autocorrelation function by construction. The hard task is now to show that this problem has a solution. Remarkably, it admits an analytical solution in a closed expression for $S(f)$, that we now work out.

\newpage 
a
\newpage

\section{Introduction}
The Maximum Entropy principle (MAXENT) is one of the most important results in Bayesian Statistics. It provides a way to uniquely assign probabilities to a phenomenon in a way that best represent our state of knowledge, while at the same time it is non committal with unavailable information. Its domain of application turned out to be wider than expected. In fact, thanks to Burg \cite{burg1975maximum}, this method has also been applied to perform high quality computation of power spectral densities of time series. After a short introduction to Jayne's MAXENT, we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form. Since its natural domain of application are time series, we will also provide a brief explanation of the basic concepts in stochastic processes that are needed to fully understand the method and the cases in which it can be correctly applied. 
\section{Maximum Entropy}
One of the main problems in statistics is the assignment of a probability for an event. There have been several proposals from both 'objective' and 'subjective' approaches to probability and statistic. The former assert that concept of probability is intrinsic to the event itself and can be computed as the 'percentage of occurrences of an attribute in a given ensemble':
speaking of probability only makes sense when an ensemble is considered. 
For the latter approach instead, probability is a representation of the state of knowledge of the observer about the event itself: the higher the probability, the higher our belief that the event will occur and vice versa. 
It appears to be clear that for the Bayesian, information will play a key role in the a priori assignment of the probability for an event, while for the frequentist prior assignment does not make sense at all, since the whole information is contained in the experiment. \\
So, assigning probability given some prior information is a pure Bayesian problem.

\subsection{Indifference principle}
The indifference principle (IP) is one of the easiest principle that can be used to assign a probability to some events. 
Consider a set of independent, mutually exclusive hypotheses: 
\begin{equation}\nonumber
    \left\{H_i\right\} \quad i = 1 \dots N
\end{equation}
so that they are exhaustive: 
\begin{equation}\nonumber
    \sum_{i = 1}^{N} P(H_i \vert I) = 1.
\end{equation}
If there is no reason to prefer one hypothesis with respect to another, the IP states that we should assign the same probability for every event: 
\begin{equation}
    P(H_i\vert I) = \frac{1}{N}. 
\end{equation}
This can be easily generalized to the continuous case. Consider, for simplicity, a one dimensional case, and suppose we have the additional information: $a < x < b$.\\ 
In the continuous case, the probability $p(x\vert I)$ is meant as the probability for the event to assume values in a given subset $\left[x, x +dx\right]$. Requiring events to be mutually exclusive is requiring their interception to be the empty set, while the exhaustiveness condition is the normalization condition:
\begin{equation}
    \nonumber
    \int_a^b p(x\vert I)dx = 1.
\end{equation}
In this case, the IP reduces to the computation of the normalization constant: 
\begin{equation}
    p(x\vert I) = \frac{1}{b-a}
\end{equation}
and allow us to easily introduce uniform priors, the simplest prior distribution for Bayesian computation. The only hypothesis is for our events to be mutually exclusive and exhaustive. In this case, since there is no evidence for a given hypothesis to be more plausible than another, according to IP we have to assign the same probability for every event. \\
If that is not the case, how can we generalize IP principle for a larger amount of information?\\ 
Jaynes\cite{JaynesArticle}\cite{jaynes2003ptl} formulated an exhaustive answer as a variational problem and rely to the concept of information entropy: 
the distribution that reflects knowledge being maximally non committal to unknown information is uniquely determined maximizing Shannon's entropy\cite{Shannon} using information as constraints. \\ 
This is now known as the 'Maximum Entropy Principle' (MAXENT)

\subsection{Maximum Entropy Principle}
To introduce MAXENT as a variational problem, we will workout some simple examples to understand how information's entropy and 'uncertainty' are related. \\ 
The word 'information' will not be used to indicate our knowledge on the system under study: this is what we will call 'evidence'. Information will be intended as in communication theory: it specifies the quantity of details needed to provide a full description of the system under study. \\ 
If we consider a perfectly sinusoidal signal, knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information entropy content . \\ 
Even less information is required if we are studying a system whose outcome is certain (has probability $p = 1$).
If the quantity of information can be considered to be $I = 0$, a communication is not even needed.  
Shannon \cite{Shannon} proposed the quantity
\begin{equation}\label{eq:information}
    I_i = \log_2 \frac{1}{p(x_i)}
\end{equation}
to represent the quantity of information brought by an outcome $x_i$ with probability $p_i$. It has the nice property of being an additive quantity and it is monotonically decreasing as a function of $p \in [0, 1]$. The more uncertain the outcome, the higher the information it brings. \\ 
What happens if we consider not one outcome, but a larger set of different events? What is  the 'most uncertain' outcome?  Since we are intending probability as a representation of the state of knowledge, one of the first requirement for our assignment is the agreement with common sense. \\ 
Consider an experiment based on the occurrence of one of two different outcomes $E_1, E_2$ with given probabilities $P_1$ and $P_2$. What are the values that make our outcome the most uncertain? \\ 
If $P_1$ and $P_2$ are largely different, like $P_1 = 0.999$ and $P_2 = 0.001$, we have enough evidence to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome. \\ 
The most unpredictable experiment is the one in which 
\begin{equation}\nonumber
    P_1 = P_2 = \frac{1}{2}:
\end{equation}
this describes a situation of 'maximum ignorance'; we have no evidence at all and we have no instruments to predict what the outcome of such an experiment will be.  \\ 
This can be generalized in the case of $N$ events. Assigning same probability
\begin{equation*}
    P_1 = \dots = P_N = \frac{1}{N}
\end{equation*}
for each outcome, i.e. appealing to the indifference principle, is equivalent to the requirement for our experiment to be the most unpredictable. 
All we know is that one over $N$ events is going to happen. With this evidence, we cannot expect one outcome to occur over the others. 
To report the results of such an experiment, we need to communicate every single outcome: it 'carries' high information. 
This description also makes clearer the role played by evidence: it has to be used in probability assignment in order to make the result 'more certain' or 'less informative'. It is clear from the point of view of communication theory: the higher the evidence on a system, the lower the information I need to send to describe it. \\ 
It is now clear that any assignment different from the 'most uncertain' is violating the hypothesis for it to being non committal to non-available evidence. We can only assign a more certain probability, i.e. less informative, if we have evidence for it. If we have none, the result has to be the most unpredictable, i.e. the most informative. 
Assigning probabilities requires maximization of uncertainty using evidence as constraint. This looks like a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized. $H$ has to be a 'generalization' of information defined in equation (\ref{eq:information}) to be something like the 'expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.  \\ 
Some properties for $H$ can be worked out easily. 
From previous discussion, given a fixed number $N$ of possible outcomes, we want $H$ to be maximized by the choice $p_1 = \dots = p_N$. \\ 
Consider now the opposite situation, in which probability is assigned appealing to IP and we have a different number of possible outcomes: 
\begin{itemize}
\centering
    \item $P_1 = P_2 = \frac{1}{2}$
    \item $P_1 = P_2 = P_2 = \frac{1}{3}$
    \item $P_1 = \dots = P_4 = \frac{1}{4}$
\end{itemize}
Intuitively, the most unpredictable experiment is the one with larger number of outcomes: probability of getting it wrong increases as $\frac{N -1}{N}$. So when equal probabilities are assigned, we want
\begin{equation}
    \nonumber
    H\left[\frac{1}{N}, \dots, \frac{1}{N}\right]
\end{equation}
to increase monotonically with N. \\ 
We also want our function to be continuous in its parameters: any small change in probability has to slightly affect uncertainty. \\
Last is composition property. Consider three events with probability $p_1, p_2, p_3$ such that $p_2 + p_3 = 1 - p_1 = q$. We require 
\begin{equation}
    H\left(p_1, p_2, p_3\right) = H(p_1, q) + qH\left(\frac{p_2}{q}, \frac{p_3}{q}\right),
\end{equation}
so that uncertainty is additive. \\ 
Shannon \cite{Shannon} showed that the only functional form satisfying our assumptions is given by:
\begin{equation}\label{eq:entropy}
    H[p_1, \dots, p_N] = - \sum_{i = 1}^N p_i\log{p_i},
\end{equation}
or, in the continuous case:
\begin{equation}
    H[p(x)] = - \int p(x)\ln p(x) dx,
\end{equation}
and $H$ is what we will call entropy. This result is very interesting, if we compare it with Eq.~(\ref{eq:information}), we see that the entropy is defined as the 'average information'. The probability assignments made maximizing it with respect to $p$, is somehow requiring  the 'expected value' of uncertainty to be a maximum. Such an assignment is in agreement with our hypothesis of being 'non committal' with unavailable evidence, since \ref{eq:entropy} is all we need to assign probabilities uniquely. Evidence can be added in form of constraints in the maximization. \\ It is easy to verify that this expression for entropy verifies the aforementioned properties. 
Assigning the same value for each probability $p_i$, it is straightforward to verify $H$ increases with N: 
\begin{equation}
    -\sum_{i = 1}^N \frac{1}{N}\log{\frac{1}{N}} = - \log\frac{1}{N} = \log N.
\end{equation}
We want to show that, in case in which our only constraint is exhaustiveness \begin{equation}\nonumber 
    \sum_{i = 1}^{N}p_i = 1,
\end{equation}
MAXENT reduces to indifference principle as it should be. The variational problem is defined as
\begin{equation}\label{eq:entropyNormConst}
    \delta\left[H - \lambda\sum_{i = 1}^N (p_i - 1)\right] = 0 
\end{equation}
with
\begin{equation}
    \delta H\left[p_1, \dots, p_N\right] = \sum_{i = 1}^N\frac{\partial H}{\partial p_i}\delta p_i = - \sum_{i = 1}^N\left(\log p_i + 1\right),
\end{equation}
so that equation \ref{eq:entropyNormConst} reduce to
\begin{equation}\nonumber 
    \sum_{i = 1}^N\left[\log p_i + \lambda + 1 \right]\delta p_i = 0\,.
\end{equation}
The solution of the previous equation is easily worked out requiring every term to be zero: 
\begin{equation}\nonumber 
    p_i = e^{-\lambda - 1}.
\end{equation}
Applying the normalization condition, it is immediate to show that
\begin{equation}
    p_i = \frac{1}{N}.
\end{equation}
MAXENT reduces to the IP when only the normalization constraint is used. What happens if more than one constraint is used? \\
For simplicity, we will work through the next examples in the continuous case. Suppose the mean of the distribution is known $\int xP(x) dx = \bar{x}$. The variational problem can be rewritten as: 
\begin{equation}
    \int \left[\log p(x) + 1 + \lambda + \mu x \right]\delta p(x) = 0
\end{equation}
and its solution in term of $p_i$ is: 
\begin{equation}
    p(x) = e^{-1 - \lambda - \mu x} 
\end{equation}
and constraints: 
\begin{equation} \nonumber
    \int p(x) dx  = 1 \text { and }
    \int x p(x) dx = \bar x.
\end{equation}
For simplicity, we consider $x$ to be a variable ranging in the interval $(-\infty,+\infty)$, but this can be easily computed for every other interval. The normalization conditions implies 
\begin{equation}\nonumber 
   1 =  \int p(x \vert I) dx = e^{-1-\lambda}\int e^{-\mu x} dx, 
\end{equation}
so that: 
\begin{equation}\nonumber 
    e^{-1 -\lambda} = \mu.
\end{equation}
In the continuous case MAXENT assigns for $x$ an exponential distribution
\begin{equation}
    p(x \vert I) = \mu e^{-\mu x}. \nonumber 
\end{equation}
It is clear from the functional form that the Lagrange multiplier has to be $\mu = \frac{1}{\bar x}$ to satisfy the second constraint. \\ 
The most interesting and useful result of MAXENT is obtained when a constraint on the variance is also considered. Take a vector  $\vec x(t)$ with components in $\mathbb{R}^N$
\begin{equation*}
    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
\end{equation*}
We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals. If we put constraints in both mean $\vec \mu(t)$ and covariance matrix $C_(t_i, t_j)$, MAXENT allow us to assign for $\vec x(t)$ a multivariate normal distribution \cite{gregory_2005}: 
\begin{equation}
    p\left(\vec x(t)\vert I\right) = 
    \frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right). 
\end{equation}
With MAXENT, the often unjustified assumptions for the data to be normally distributed finds an elegant explanation. It is not only the central limit distribution for finite variance processes, it is also the distribution that maximizes the entropy and, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process. In some sense, we can interpret the central limit theorem as the natural 'statistical' evolution toward a configuration that maximizes entropy.

\section{Time Series}
Before we introduce MAXENT applications to spectral analysis, we provide a short review on time series and some concepts that will be used in following chapters. \\ 
Consider a stochastic process, i.e. any process with some non deterministic component that takes values in some space $\bar X = \left\{x_1 \right\}$. A time series X of the process, i.e. a realization, is a set of subsequent values assumed by the variable at different times, i.e.
\begin{equation}
    X = \left\{x_1, x_2, \dots, x_N \right\}; \qquad \text{ with } x_i = x(t_i)
\end{equation}
For our goal is important to understand how a single time series can be used to infer properties on the stochastic process itself. A full characterization for the stochastic process can only be provided if all the joint probability distributions for the process are known
\begin{equation}\nonumber 
    p(x_1, t_1 | I); \quad \dots \quad ;  p(x_1, t_1; \dots x_n, t_n | I);  \quad \dots  
\end{equation}
or, equivalently, all momenta for the distribution are provided: 
\begin{align}
    \nonumber &\mu(t_1) = \int  x_1p(x_1, t_1 \vert I)dx_1  \\
    \nonumber 
    & R(t_1, t_2) = \int  x_1x_2p(x_1, t_1;x_2, t_2\vert I) dx_1dx_2\\
    \nonumber 
    \vdots \\ \nonumber
    &\mu_n(t_1, \dots, t_n) = \int  x_1\dots x_np(x_1; \dots; x_n \vert I) dx_1dx_2\dots dx_n\\ 
    \nonumber
    &\vdots
\end{align}
Since both descriptions require $N \to \infty$, they are impossible to achieve and very often an estimate of just the first two moments is given, hence only a partial characterization can be made. \\ 
The first momentum is known as the mean for the process, while second moment is known as 'Autocorrelation', and gives information on the deterministic dependence of a process from its previous values. It is a central quantity in time series analysis and is related to the autocovariance function via: 
\begin{equation}\nonumber 
    C(t_1, t_2) = R(t_1, t_2) - \mu(t_1)\mu(t_2).
\end{equation} \\
Study of time series analysis is often a hard task, but there are some special cases that simplify the analysis: stationary and autoregressive processes.  
\subsection{Stationary processes}
\subsubsection{Stationarity conditions}
There are several definitions of stationarity for a process \cite{Mantegna2007Introduction}. A stochastic process is said to be 'strictly stationary' if the joint distribution for the process: 
\begin{equation}
p(x_1,t_1; \dots; x_n,t_n \vert I) = p(x_1,t_1 + \Delta t; \dots; x_n, t_n +\Delta t\vert I)
\end{equation}
is constant in time whatever the order $n$ and time lag $\Delta t$. This implies that its moments are also constant in time. In this case, the time series consists in independent identically distributed terms. Such a process is rarely encountered and more general definitions are needed for real case studies.  \\ 
On of the most important is the 'wide sense stationary process', whose constraints are: 
\begin{equation}
    \mu_1(t) = \mu; \quad R(t_1, t_2) = R(t_2 - t_1) = R(\tau).
\end{equation}
Such a process has constant mean and autocorrelation that only depends on time lag between the considered values. \\
The mean does not bring high information about the process, and can always be reabsorbed in the process defining 
\begin{equation}
    \nonumber
    y(t) = x(t) - \mu.
\end{equation}
y(t) is a stationary process with 0 mean but same statistical properties of $x(t)$. $R(\tau)$ brings information on the amount of dependence of the time series on its previous values: if some process repeats similarly after $\bar t$ seconds, autocorrelation will have a secondary maximum at $\tau = \bar t$. Also, if the function is periodic of period $T$, the autocorrelation will have same periodicity. 
In general, autocorrelation represents most of the information we can obtain from the study of a stationary stochastic process, and it is rather fundamental. Its computation is even more important in the case of a Gaussian process. For such a process it is the only moment required to fully characterize it. 
One of the most important results regarding the autocorrelation function, is the Wiener-Khinchin theorem. It ensures that, for a wide sense stationary process, the frequency domain counterpart of the autocorrelation function is the power spectral density:
\begin{equation}
    R(\tau) = \int  S(\omega) e^{-\imath \omega \tau} d\omega,
\end{equation}
Autocorrelation and power spectral density bring the same information about the process. It is fundamental to develop methods to accurately estimate the power spectral density in such a way that it is consistent with the autocorrelation function.
\subsubsection{Gaussian noise}
MAXENT is a powerful tool to study noise. If one considers a wide sense stationary process  $\vec n(t) = \left(n_1, n_2, \dots, n_k \right)$ with 0 mean and known autocorrelation matrix, one can assign uniquely a normal distribution: 
\begin{equation}\label{eq:noisePDF}
    p\left(\vec n(t)\vert I\right) = \frac{1}{\left(2\pi\det R \right)^{\frac{k}{2}}}\exp\left(-\frac{1}{2}\sum_{i, j}n_i R^{-1}_{ij}n_j\right)\,.
\end{equation}
In the case of a wide sense stationary process, being $R_{ij} = R(t_j - t_i)$ dependent on time steps only, the autocorrelation matrix is a Toeplitz Matrix (see Appendix \ref{app:Toeplitz}) and can be diagonalized, in the limit of large number of samples, by a Discrete Fourier Transform: 
\begin{equation}
    S = F R F^{-1}. 
\end{equation}
For the Wiener-Khinchin theorem, $S$ is the power spectral density for the noise itself. Defining
\begin{equation}
    \tilde n(\omega) = F\left[n(t) \right],
\end{equation}
with $F$ the Fourier transform, the probability distribution (\ref{eq:noisePDF}) for noise in frequency domain (FD) becomes  
\begin{equation}
    p(\tilde n(f) \vert I) = \frac{1}{\left(2 \pi \det S \right)} \exp \left(-\frac{1}{2}\sum_{i} \frac{\tilde n_i \tilde n_i}{S_i} \right).
\end{equation}
Power spectral density plays a central role in noise study: it is the covariance matrix for noise in frequency domain. Also, in large N domain, frequency noise is serially uncorrelated. \\
There is a very special case of Gaussian Noise that we have to mention: white noise. White noise can be represented as Gaussian noise with autocorrelation function
\begin{equation}
    R_{ij} = \sigma^2 \delta_{ij},
\end{equation}
 it is a strictly stationary process whose autocorrelation function is 0 for any time lag different from 0, and its realization is made of independent identically distributed samples.
 Going in the Fourier Domain
\begin{equation}
S_{ij} = \sigma^2 F_{ik}\delta_{kl}F^{-1}_{lj} = \sigma^2 \delta_{ij} = R_{ij}, 
\end{equation}
we show that power spectral density for white noise is constant over frequency.

\subsection{Autoregressive process}\label{sec:ARprocess}
The description of a process is simplified if we are dealing with some autoregressive process of order p, or simply an AR(p) process. For an AR(p) process, each value depends linearly from its previous values
\begin{equation}
    x_i = \sum_{i = 1}^{p}a_{i}x_{i - p} + \nu_i, 
\end{equation}
with $a_i$ some constant coefficients and $\nu_i$ white noise. \\
An AR(1) process is known as a Markov process. Complete knowledge on AR(p) can be achieved with the computation of $p + 1$ momenta (or conditional distributions) only. We prove it for a Markov process.\\  
Consider the $n'th$ order joint probability distribution
\begin{equation}\label{eq:NthOrderPDF}
    p(x_1, t_1; x_2, t_2; \dots; x_n, t_n | I),
\end{equation}
from the definition of conditional probability
\begin{equation}\nonumber 
p(x,y \vert I) = p(x \vert y I) p(y \vert I)
\end{equation}
one can rewrite \ref{eq:NthOrderPDF} as 
\begin{equation} \nonumber 
    p(x_n, t_n; \dots; x_1, t_1 \vert I) = p(x_n, t_n \vert x_{n-1}, t_{n-1}; \dots x_1, t_1 I) p(x_{n-1}, t_{n-1} \dots x_1, t_1 \vert I). 
\end{equation}
Since under Markov assumption, conditional $pdf$ only depends on the previous value for the variable 
\begin{equation} \nonumber 
    p(x_n, t_n \vert x_{n-1}, t_{n-1}; \dots, x_{1}, t_1 I) = p(x_n, t_n \vert x_{n-1}. t_{n-1} I) 
\end{equation}
reiterating the decomposition of the joint probability, we find that
\begin{equation}
\left(\ref{eq:NthOrderPDF} \right)=  p(x_n, t_n \vert x_{n-1}, t_{n-1}, I )p(x_{n-1}, t_{n-1}\vert x_{n-2}, t_{n-2}, I) \dots p(x_1, t_1\vert  I),
\end{equation}
so that for every order $n$, the probability distribution can be reduced to the product of $n$ second order probability distributions. \\
The knowledge of the distribution for the variable and of the second order conditional distribution fully characterize a Markov process.\\
The same reasoning can be repeated for any order $p$. For an AR(p) process the $n$-th order joint probability distribution ($n > p$) can be rewritten as a product of the $p$-th order conditional distributions, so that they completely characterize the process.

\subsection{ARMA process and Autoregressive stationary process}
There are two fundamental results for an autoregressive process that are strictly related with Burg's Algorithm (see section \ref{sec:MESA}) for spectral analysis, and they are all about autoregressive stationary processes and their generalization, ARMA(p,q) processes. 
\subsubsection{ARMA process}
We will only need to define ARMA(p,q) and their relation with AR(p) processes without stepping in many details. These concepts are needed to give a stronger background to our future results.  \\
The generalization of the Autoregressive Process, the $ARMA(p,q)$ process, is defined by the following relation: 
\begin{equation}
    x_t = \sum_{i = 1}^p \alpha_i x_{t - i} + \sum_{i = 1}^q \beta_i \nu_{t - i} + \nu_t,
\end{equation}
again, $\nu$ is white noise and both $\alpha$ and $\beta$ are constants. It is evident that an autoregressive process is a special case of the more general $ARMA(p,q)$ process, since  $Ar(p) = ARMA(p,0)$. We will not enter in the details of the description of $ARMA(p,q)$ processes, but they are important for two main reasons. \\
First, $ARMA$ is a parsimonious way to characterize an $AR$ process. It is, in fact, possible to prove that an $ARMA(p,q)$ process, with both $p, q$ finite, is equivalent to some $AR(\infty)$ process \cite{Haykin}. \\
The second and even more important result about these processes is Wold's Theorem\cite{Wold}: 
\begin{theorem}
Every real valued stationary stochastic process $P$ can be written as an $ARMA(p,q)$ process. 
\end{theorem}
This theorem makes clear the central role played by the ARMA representation: it allow us to fully characterize any stochastic process if its coefficients are known. This is usually an hard task, but, since we know the $AR / ARMA$ equivalence, we can find an approximate description of the process in terms of an $AR$ process, all is needed is a way to estimate the autoregressive coefficients. \\ 
\subsubsection{Autoregressive stationary process}
The last important ingredient is the condition of stationarity for AR processes. The conditions of stationarity for a stochastic process are well known. Defining the 'lag' operator $B$ such that 
\begin{equation}
    \nonumber 
    B x_{i} = x_{i - 1},
\end{equation}
one can rewrite the autoregressive process as 
\begin{equation} \nonumber 
    x_i - \sum_{k = 1}^p a_k B^k x_i = \nu_i, 
\end{equation}
or, defining $b_0 = 1$ and $b_k = - a_k$
\begin{equation}
    \left(\sum_{k = 0}^p b_k B^{k}\right) x_i = \nu_i.
\end{equation}
The polynomial on the left side is known as the 'characteristic polynomial' of the autoregressive process $\phi (B) = \sum_{k = 0}^p b_k B^k$. 
The condition of stationarity for autoregressive processes is
\begin{theorem}
An autoregressive process of order $p$ is stationary if all the roots $\beta_i$ of its characteristic polynomials lay outside the unit circle, i.e. $\vert\beta_i\vert > 1$.
\end{theorem}
Consider for simplicity a Markov process that only takes real values
\begin{equation}\nonumber 
    x_t - a_1 x_{t- 1} = \nu_t; \quad \text{ with } a_i \in \mathbb{R}
\end{equation}
whose characteristic polynomial is
\begin{equation}
    \nonumber 
    \phi(B) = 1 - a_1 B .
\end{equation}
Let $\beta$ be the root of the polynomial $\phi(\beta) = 0$. 
If $\vert \beta \vert > 1$ ($\vert \beta \vert < 1$) $a_1$ has to be $\vert a_1 \vert < 1$ ($\vert a_1 \vert > 1$). It is easy to see that if condition of stationarity is not met, the process explodes and $x_t$ will quickly diverge. \\ 
These few concepts are fundamental in stochastic processes and necessary for what follows.
\section{Maximum Entropy Spectral Analysis}\label{sec:MESA}
From our previous discussion on time series, the importance of the autocorrelation function should be clear. Since it contains the same information of a power spectral density, high resolution computation of a power spectrum from a single time series $\vec x(t)$ of length $N$ is a fundamental task in signal analysis.  \\
Since we are dealing with a finite number of samples $N$, it is not possible to have an accurate computation of the autocorrelation function for every time interval. Also, the error $\sigma$ in the estimate increases with the time interval $k$ as $\sigma \sim \frac{1}{\sqrt{N - k}}$, so that only a few values for the autocorrelation function can actually be computed. \\ 
The problem we want to address is how to compute a PSD from the partial knowledge of the autocorrelation function. \\
The standard approach of peridiograms \cite{Lomb} \cite{Scargle} consists in taking the Fourier Transform of the sample autocorrelation, written as \begin{equation}
    \rho = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\}.
\end{equation}
The sequence $W$ is a window function that can be chosen in several different way, each method having pros and cons for the final estimate of the power spectral density. \\ 
The choice of a window function is arbitrary and made by trial and error, until a satisfactory compromise between variance and resolution of the estimate of power spectral density  is reached. A high frequency resolution implies high variance and vice-versa. Another problem,is that the result depends on this choice, and this fact may raise the question of which we should consider as the "real" power spectral density. \\  
Another problem with this approach is the requirement for the window to be $0$ outside the interval: we are arbitrarily assuming $\rho_j = 0$ for $j > M$, and modifying the estimate if a non-rectangular window is chosen. Making assumptions on unavailable data and modifying the ones we have at our disposal does not seem a completely fair approach. \\
MAXENT gives us the possibility to develop a completely new way for the estimation of PSDs, without any a priori assumptions on the unavailable data. Our requirements (the constraints for the variational problem) are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation. \\
To do so, there is a problem we have to solve: our definition of entropy (equation \ref{eq:entropy}) depends on the probability distribution, not the power spectrum. We need to find a way to change the definition's domain for the entropy to formulate the variational problem in terms of the power spectral density $S(f)$ alone. A way out of this problem is considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA}. One can show that the 'entropy gain' obtained filtering white noise is: 
\begin{equation}\label{eq:EntropyGain}
    \Delta H = \int_{-\infty}^{\infty}\log S(f) df , 
\end{equation}
so that we can reformulate the variational problem requiring $S(f)$ to maximize the entropy gain. In real examples we have no access to all frequencies but only on those between minus and plus the Nyquist frequency, defined from the sampling interval $\Delta t$ as 
\begin{equation}
    Ny = \frac{1}{2 \Delta t}.
\end{equation}
All our quantities are well defined in this range only, so that we are not going to maximize \ref{eq:EntropyGain}, but
\begin{equation}\label{eq:EntropyGain2}
    \Delta H = \int_{- Ny}^{Ny}\log S(f) df.
\end{equation}
Defining $\bar r_n$ as the sample autocorrelation, our constraints will be
\begin{equation}\label{eq:MaxConstraint}
    \int_{-Ny}^{Ny} S(f) e^{\imath 2 \pi f n \Delta t} df = \bar r_{n}.
\end{equation}
This approach on power spectral density estimate was developed by Burg \cite{burg1975maximum} and it is a way out from the problems raised in periodograms estimate. It requires no assumptions on the unavailable estimates for $\bar r_n$ and is consistent with the sample autocorrelation function by construction. The hard task is now to show that this problem has a solution. Remarkably, it admits an analytical solution in a closed expression for $S(f)$, that we now work out. 
\subsection{MESA solution}
We will follow the solution proposed by Burg in his PhD thesis \cite{burg1975maximum}. Instead of directly solving the variational problem for the power spectral density, we will rewrite it in terms of the constraints (\ref{eq:MaxConstraint}) of the problem. $S(f)$ is defined as the Fourier Transform of the autocorrelation function, as: 
\begin{equation}
    S(f) = \frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{- \imath 2 \pi n \Delta t}.
\end{equation}
Substituting in \ref{eq:EntropyGain2}, the variational problem reduces to the maximization of: 
\begin{equation}
    \int_{-Ny}^{Ny} 
    \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{-\imath 2 \pi f n \Delta t}
    \right).
\end{equation}
As we said before, not all the values for the autocorrelation are known. Generally, we have access to a finite number of values for the autocorrelation function, ranging in a symmetric interval $n \in [-N, N]$. 
In the previous equation, since the summation index is ranging between $-\infty$ and $+\infty$, we are correctly assuming that also the unknown values have to be used in the computation of the power spectral density. But, since they are not known, we have to require the maximization of entropy to be independent of these values:
\begin{equation}\nonumber 
    \frac{\delta H}{\delta \bar r_s} = 0 \text{ for } \vert s \vert > N. 
\end{equation}
Using the explicit form for $H$, this becomes
\begin{equation}
      \frac{\delta H}{\delta \bar r_s} = \frac{1}{2Ny}\int_{-Ny}^{Ny}S(f)^{-1}e^{-\imath 2 \pi f s \Delta t} = \lambda_s = 0 \text{ for } \vert s \vert > N. 
\end{equation}
We see that $S(f)$ can be expressed in terms of the $\lambda s$ via a Fourier Series
\begin{equation}\label{eq:PSDconstraint}
    S(f)^{-1} = \sum_{s = -N}^N \lambda_s e^{-\imath 2 \pi f s \Delta t},
\end{equation}
This is a non-linear closed expression for the power spectral density. The requirement for $S(f)$ to be positive definite implies that: 
\begin{equation}
    \nonumber 
    \lambda_{-s} = \lambda_s^* 
\end{equation}
With this property, and defining $z = e^{-\imath 2 \pi f \Delta t}$, $S(f)^{-1}$ can be rewritten as a polynomial in $z$ and $z^*$
\begin{equation}
    \label{eq:zExp}
    S(f)^{-1} = \lambda_0 + \sum_{s = i}^N \lambda_s z^s + \sum_{s = 1}^N \lambda^*_s z^{-s}.
\end{equation}
Let $z_0$ be a root for the polynomial
\begin{equation}
    \nonumber
    \lambda_0 + \sum_{s = 1}^N \lambda_s z_0^s + \sum_{s = 1}^N \lambda^*_s z_0^{-s} = 0,
\end{equation}
it is easy to show that $(z_0^*)^{-1}$ is also a root for the equation: 
\begin{align}
    \nonumber 
    &\lambda_0 + \sum_{s = 1}^{N} \lambda s \left(\frac{1}{z_0^s} \right)^* + \sum_{s = 1}^N \lambda^*_s \left(\frac{1 }{z_0^{-s}}\right)^* = \\ \nonumber 
    &\lambda_0 + \sum_{s = 1}^N \lambda_s (z_0^{-s})^* + \sum_{s=1}^N \left(\lambda_s (z_0^s)\right)^* = \\ \nonumber 
    &\left(\lambda_0 + \sum_{s=1}^N \lambda_s z_0^s + \sum_{s = 1}^N \lambda^*_s z_0^{-s}\right)^* = 0 
\end{align}
This means that for every pole $z_0$ laying outside (inside) the unit circle, there is another pole inside (outside) the unit circle.  So, if there are no roots on the unit circle, $N$ roots lay outside and $N$ inside of it. \\ 
These properties allow us to rewrite the Fourier expansion (\ref{eq:zExp}) as \cite{1975STIN...7714318B}
\begin{equation}\nonumber 
    S(f)^{-1} = \frac{1}{P_N \Delta t} \left(1 + \sum_{s}^N a_s z^s\right)\left(1 + \sum_s^N  a^*_s z^{-s}\right),
\end{equation}
or, defining $a_0 = 1$,
\begin{equation}\label{eq:MESApsd}
    S(f) = \frac{P_N \Delta_t}{\left(\sum_{s=0}^N a_s z^z\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}. 
\end{equation}
The MESA expression for the power spectral density admits a non-linear, all pole representation when evaluated in $z = e^{-\imath 2 \pi f \Delta t}$. All the roots of the first polynomial can be chosen to lay outside the unit circle, so that all the roots of the second polynomial will lay inside of it \cite{1975STIN...7714318B}. 
If one is able to compute the values of the coefficients $a_i$ the PSD is uniquely determined from the previous equation. \\ 
To compute the $a_i$-s, we substitute in equation
(\ref{eq:MaxConstraint}) with the above representation for $S(f)$ in terms of the $a_i$ coefficients. Changing  variable from $f$ to $z(f) = e^{-\imath 2 \pi f \Delta t}$, the integral (\ref{eq:MaxConstraint}) becomes a contour integral in the unit circle centered at the origin of the complex plane: 
\begin{equation}
   \frac{P_N}{2 \pi \imath} \oint _{\mathbb S^1}\frac{z^{-s - 1}}{\sum_{n = 0}^N a_n z^n \sum_{n = 0}^N a^*_n z^{-n}} = \bar r_s. 
\end{equation}
Sending $s \to s - r$, multiplying by $a^*_s$ and summing over all possible values for $s$ the equation becomes 
\begin{align} \nonumber 
    \sum_{s = 0}^N a_s \bar r_{s - r} &= \frac{P_N}{2 \pi \imath}\oint \frac{z^{r - 1} \sum_{s = 0}^N a^*s z^{-s}}{\sum_{s = 0}^N a_s z^s \sum_{s = 0}^N a^*s z^{-s}} dz\\
    & = \frac{P_N}{2 \pi \imath}\oint \frac{z^{r -1}}{\sum_{s = 0}^N a_s z^s}dz\label{eq:ErFilter}
\end{align}
Since all the zeros of the denominator lay outside the unit circle, the contour integral is 0. The only exception is for $r = 0$ with a residual in the origin of the axis. For $r = 0$ the integral can be computed via the Cauchy integral formula and (\ref{eq:ErFilter}) becomes: 
\begin{align}\label{eq:errorFilter1}
    \sum_{s = 0}^N a_s \bar r_{r - s} &= P_N \quad \text{ if } r = 0 \\ \label{eq:errorFilter2}
    \sum_{s = 0}^N a_s \bar r_{r - s} & = 0 \qquad \text{ if } r \neq 0.
\end{align}
This kind of equations are solved by the use of the Levinson recursion \cite{doi:10.1002/sapm1946251261}
\subsection{Levinson Recursion}
The equations (\ref{eq:errorFilter1}) and (\ref{eq:errorFilter2}) can be rewritten as a Matrix equation of the form
\begin{equation}\label{eq:predictionErrorFilter}
    \begin{pmatrix}
    \bar r_0 &  \bar r_{-1} & \dots & \bar r_{-N}\\
    \bar r_{1} & \bar r_ 0 & \dots & \bar r_{1 - N} \\ 
    \bar r_{2} & \bar r_{1} & \dots & \bar r_{2 - N}
    \\
    \vdots & \vdots & \ddots & \vdots  \\
    \bar r_{N} & \bar r_{N - 1} & \dots  &\bar r_0
    \end{pmatrix}
    \begin{pmatrix}
    1 \\   a_1 \\   a_ 2 \\  \vdots \\   a_N
    \end{pmatrix} = 
    \begin{pmatrix}
    P_N \\   0 \\  0 \\ \vdots  \\   0
    \end{pmatrix}
\end{equation}
know as the prediction error filter equation. It is a matrix equation involing the Toeplitz Matrix $R_{ij}$ and the vector $\vec a$, known as the forward prediction error filter. The approach for solving the equations is the Levinson - Durbin recursion. \\
Our goal is to obtain the $N$-th order equation (\ref{eq:predictionErrorFilter}) from the $N-1$-th order equation: 
\begin{equation}\label{eq:Forward}
    \begin{pmatrix}
    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N}\\
    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} \\ 
    \vdots & \vdots & \ddots & \vdots \\ 
    \bar r_{N - 1} & \bar r_{N -2} & \dots & \bar r_0
    \end{pmatrix} 
    \begin{pmatrix}
    1  \\   b_1  \\  \vdots \\   b_{N - 1}
    \end{pmatrix} =
    \begin{pmatrix}
    P_{N -1} \\  0 \\ \vdots \\ 0
    \end{pmatrix} .
\end{equation}
This equation is not sufficient to solve the recursion. Since the autocorrelation matrix is self-adjoint, from equations (\ref{eq:errorFilter1}) and (\ref{eq:errorFilter2}) and noting that the first and the last row of the autocorrelation matrix are respectively the complex conjugate reverse, is easy to show that
\begin{equation}\label{eq:Backward}
     \begin{pmatrix}
    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N}\\
    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} \\ 
    \vdots & \vdots & \ddots & \vdots \\ 
    \bar r_{N - 1} & \bar r_{N -2} & \dots & \bar r_0
    \end{pmatrix} 
    \begin{pmatrix}
    b^*_{N-1} \\ b^*_{N- 2} \\\vdots \\ 1
    \end{pmatrix} 
    = 
    \begin{pmatrix}
    0 \\ 0 \\ \vdots \\ P_{N-1}
    \end{pmatrix} 
\end{equation}
holds. The vector obtained reflecting b and taking its complex conjugate is known as ``backward prediction error". \\ 
Stepping from order $N - 1$ to order $N$, the autocorrelation matrix grows from size $(N,N)$ to size $(N + 1, N + 1)$, this implies that a new component $b_n$ for the prediction error filter is needed. If we take it to be $b_N = 0$, this choice does not modify the first $N$ component of the right side of equation (\ref{eq:Forward}) (or the last $N$ components of (\ref{eq:Backward})), i.e.:

\begin{equation}\label{eq:OrderN}
    \begin{pmatrix}
    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N} & \bar r_{- N}\\
    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} & \bar r_{1 - N}\\ 
    \vdots & \vdots & \ddots & \vdots & \vdots \\ 
    \bar r_{N -1} & \bar r_{N-2} & \dots & \bar r_0 & \bar r_{-1}
    \\
    \bar r_{N } & \bar r_{N -1} & \dots & \bar r_1 & r_0 
    \end{pmatrix} 
    \begin{pmatrix}
    1  \\   b_1  \\  \vdots \\   b_{N - 1} \\ 0 
    \end{pmatrix} =
    \begin{pmatrix}
    P_{N -1} \\  0 \\ \vdots \\ 0 \\ \Delta_N
    \end{pmatrix} .
\end{equation}
with 
\begin{equation}
    \Delta_N = \sum_{n=0}^N \bar r_{N - n}b_n.
\end{equation}
Adding the previous expansion for both forward and backward equations, one obtains 
\begin{equation}\nonumber 
        \begin{pmatrix}
    \bar r_0 & \bar r_{-1} & \dots & \bar r_{- N}\\
    \bar r_{1} & \bar r_0 & \dots & \bar r_{1 - N}\\ 
    \vdots & \vdots & \ddots & \vdots \\ 
    \bar r_{N -1} & \bar r_{N-2} & \dots &  \bar r_{-1}
    \\
    \bar r_{N } & \bar r_{N -1} & \dots  & r_0 
    \end{pmatrix} 
    \begin{pmatrix}
    1  \\   b_1  \\  \vdots \\   b_{N - 1} \\ 0 
    \end{pmatrix} 
    + c_N 
    \begin{pmatrix}
    0 \\ b^*_{N -1} \\ \vdots  \\ b_1 \\ 1 
    \end{pmatrix}
    =
    \begin{pmatrix}
    P_{N -1} \\  0 \\ \vdots \\ 0 \\ \Delta_N
    \end{pmatrix} .
    + c_N
    \begin{pmatrix}
    \Delta^*_N \\ 0 \\\vdots \\ 0 \\ P_{N-1}
    \end{pmatrix}.
\end{equation}
With the reflection coefficient $c_N$ to be computed. Requiring the previous equation to be equivalent to \ref{eq:predictionErrorFilter}, one obtains
\begin{equation}\nonumber 
\begin{cases}
    P_{N - 1} + c_N\Delta^*_N = P_N \\
    c_N P_{N - 1} + \Delta_N = 0
\end{cases}
\end{equation}
which can be easily solved for $c_N$ and $P_N$ as 
\begin{equation}
    c_N = -\frac{\Delta_N}{P_{N-1}}; \quad \text{ and } P_N = P_{N - 1}\left(1 - \vert c_N \vert ^2 \right). 
\end{equation}
Having computed the value for the reflection coefficient, the $N$-th order forward prediction error filter is found to be 
\begin{equation}
    a_i = b_i + c_N b^*_{N - i}.
\end{equation}
To obtain the solution for the $N$-th order, one starts $\bar r_0 = P_0$ and recursively computes the new values until the desired order is computed. 
This problem is often solved numerically with Burg's Algorithm \cite{Vos}.
A python implementation of Standard Burg's Method can be found in Appendix \ref{app:Burg} 
\section{The Autoregressive process analogy}
The application of MESA is not limited to spectral estimates, but it also provides a bridge between spectral analysis and the analysis
of autoregressive processes.\cite{doi:10.1029/RG013i001p00183}.
Consider an autoregressive stationary process of order $p$: 
\begin{equation*}
    x_t - b_1 x_{t-1} - b_2 x_{t-2} \dots b_n x_{t - n} = \nu_t, 
\end{equation*}
taking the z transform of the previous: 
\begin{align}
    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t
\end{align}
and calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
in the $z$ domain, the process takes the form:
\begin{equation}
    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
\end{equation}
Since we assumed stationarity for the process, $\tilde{x}(z)$ is analytic both on and inside the unit circle. Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
\begin{equation}\label{eq:ARspectrum}
    S(f) = \vert \tilde x(z)\vert ^ 2 = 
    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}.
\end{equation}
The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$. \\ 
Comparing (\ref{eq:ARspectrum}) with (\ref{eq:MESApsd}), identifying $b_1 = - a_i$ and $P_N \Delta t= \sigma ^ 2$, the expression for the power spectral density obtained with MESA is formally equivalent to the power spectral density of some stationary autoregressive process. \\ 
Having computed the prediction error filter one can easily obtain an estimate of the $AR$ coefficients for the process just changing sign to the prediction error filter $a_i$.
Taken a time series, MESA allows us to compute the power spectrum and to find an effective description in terms of an autoregressive process. 
Wold's theorem ensures that this (approximate) description for the phenomenon is legitimate when the process is found to be stationary. So, MESA defines a bridge between spectral estimation and the study of autoregressive processes, an unexpected and fundamental result. \\
Another very remarkable process for this estimate, is that the linear predictor obtained solving the error filter equation, is the best linear predictor, i.e. it minimizes error, and is equivalent to a least square fit for the coefficients. \cite{VDBos}  \\ 
To prove that it minimizes error, we will follow the original proof provided by Burg \cite{burg1975maximum}. Consider a general linear predictor $-c_i$: 
\begin{equation}
    x_{t} - \sum_{i = 1}^p (-c_i)x_{t - i} = \sum_{i = 0}^p c_i x_{t - i} = e'_t, 
\end{equation}
Denoting with $E\left[\cdot\right]$ the expectation value of some variable, the mean square error is given by: 
\begin{equation}
E\left[\sum_{i = 0}^p c_i x_{t-i}\sum_{j = 0}^p c^*_j x^*_{t - j}\right] = \sum_{i,j} c^*_j E\left[x^*_{t-j}x_{t-i} \right] c_j = \sum_{i, j}c^*_j R(j - i) c_i > 0,
\end{equation}
which, being $R$ positive definite and defined by the multiplication of a complex number times its complex conjugate, is always positive definite. 
Let $-\vec{b}$ be the prediction error filter obtained with MESA. Multiplying the left hand side of \ref{eq:predictionErrorFilter} for the complex conjugate of the forward prediction error, 
\begin{equation}\nonumber 
    \sum_{i,j}b^*_jR(j-i)b_i = P_N
\end{equation}
is obtained . \\ 
The (positive) quantity  
\begin{equation}\label{eq:linearPredictor}
    \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i),
\end{equation}
using Einstein notation on repeated indices and recall that $c_0 = b_0 = 1$, can be rewritten as 
\begin{align}
    & c^*_j R(j - i) c_i - c^*_j R(j - i) b_i - b^*_j R(j - i) c_i + b^*_j R(j - i)  b_i =\\ \nonumber 
    &c^*_j R(j - i) c_i - c^*_0 P_N - P_N c_0 + P_N =\\ \nonumber
    & c^*_j R(j - i) c_i - P_N.
\end{align}
Moving $P_N$ on the left hand side, it becomes 
\begin{equation}\label{eq:bestLinearPredictor}
    \sum_{i, j}c^*_j R(j-i)c_i = \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i) + P_N.
\end{equation}
Being all the quantities positive definite, the left hand side is minimized by the choice $c_i = b_i$. Also, known $P_N$, the previous equation gives an estimate on the additional mean square error associated at choices that differ from the prediction error filter.

\section{Choosing an autoregressive order}\label{sec:optimizers} 
Before being able to apply MESA, we need to address the question of the best length $M$ of the filter that provides the optimal estimate of the power spectrum. If the autocorrelation matrix was perfectly know, the answer would be simply $M = N$. \\ 
In real cases, such a knowledge is not available and one has to estimate the autocorrelation function from the data. The accuracy for the estimate decreases as the time lag $k$ increases as $\sim \left(\sqrt{N - k}\right)^{-1}$. \\ 
The accuracy of the estimate should then increase as the length of the autocorrelation function increases, but, from a certain point on, inaccuracy in the estimate for $R$ becomes so large that the estimate for the spectrum is negatively affected by this fact. A compromise between the two behaviour is proposed to be \cite{doi:10.1190/1.1440902}
\begin{equation}\label{eq:MMAx}
M_{max} = 2N / \ln{(2N)}
\end{equation}
Still, this is not a satisfactory answer since it is an upper limit for the estimate of the length $M$.
Various approaches to optimize the order $M$ have been proposed in literature; in the following we present them.
\begin{itemize}
\item \textbf{Final prediction Error (FPE)} 
The first criterion is due to Akaike (1970) \cite{Akaike1970StatisticalPI}. He proposed that M has to be chosen as the length that reduces the error when the filter is used as a predictor: 
\begin{equation}
    FPE: \min_{M}E\left((x_t - \hat x_t) ^ 2\right)
\end{equation}
with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.  This criterion is equivalent to the choice of $M$ such that 
\begin{equation}
    \min_{M}P_{m} \frac{N + m + 1}{N - m - 1}
\end{equation}
with $P_M$ defined by the prediction error filter equation. 
\item \textbf{Criterion Autoregressive Transfer function (CAT)}
This second criterion is due to Parzen \cite{bhansali1986}, and requires $M$ to be chosen as the length $m$ the minimizes the quantity 
\begin{equation}
    \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
\end{equation}
It has the interesting property that the distance between the original filter and the estimate is minimum and asymptotically unbiased.

\item \textbf{Optimum Bayes Decision rule (OBD)} 
The last criterion we will consider is the OBD (Rao, 1982)\cite{doi:10.1029/WR018i004p01097}. With the technique of asymptotic integration, $M$ is proposed to be the length that maximizes the posterior distribution, condition met selecting $m$ that minimize 
   \begin{equation}
        (N - m - 2) \log(P_m) + m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2.
    \end{equation}
\end{itemize}
Once an optimizer is chosen to select the best recursion order $M$, we then proceed as follows:  we solve the Levinson recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached. Then, the order $M$ is selected to be the one that minimize the chosen optimizer value. \\
The following chapter will be devoted to the study of the statistical properties of the previous optimizers to understand which choice provides the best quality in the reproduction of some known power spectral densities. 
\section{Conclusions}
In this chapter, we investigated several aspects of the maximum entropy approach to spectral analysis. We saw how MAXENT was originally introduced and how it solves the problem of probability assignments given some information about the phenomenon under study. We have shown that the best assignment we can make using no more assumptions than those given by the available evidence, is the one obtained maximizing Shannon's Entropy \ref{eq:entropy}. \\ 
The domain of application of MAXENT is even wider. With a proper redefinition and analogy with filters, we have been able to rewrite the variational problem in terms of the power spectral density and the constraints in terms of the sample autocorrelation. This allowed us to define MESA to compute power spectral densities of a large range of spectra, obtaining an analytical representation for the power spectral density, that result to be defined for every frequency comprised between plus or minus Nyquist frequency. \\
The analogy with autoregressive processes can be taken even further, and MESA is found to provide an estimate of the best linear predictor for a stationary autoregressive process. With the support of Wold's theorem, we can state that MESA defines a bridge between the study of spectral densities and analysis of stationary processes. \\
Even more remarkable is the fact that, in principle, this approach requires no assumption at all about the process, and all the information is directly extracted from the dataset. 
But, from a practical point of view, there is one a priori choice that one has to make: the optimizer to better approximate the length of the autoregressive process. This problem arises because we do not have a full knowledge of the autocorrelation function, but just an estimate whose error increases with the order of the time window. We briefly introduced three different optimizers, FPE, CAT and OBD, and in the next chapter we will study the properties of each of them.

\section{Analysis of optimizers - Introduction}
To apply Burg's method, it is a fundamental task to understand what is the best choice for estimating the order for the autoregressive stochastic process. Before applying it to real datasets, we will study main properties of the three different optimizer, introduce in section \ref{sec:optimizers} on distinct, simulated, realizations of some AR(p) processes. We will generate different Gauss noise time series with a given known power spectral density (PSD) to understand how each method performs in different situation. The analysis will concern the quality, to be defined later, of the reconstruction of the spectrum itself. We will consider two different kinds of spectra: i) a Gaussian spectrum with a given mean and variance; ii) an Advanced LIGO design sensitivity theoretical spectral curve.  As a part of our investigation, we will consider both ensemble-averaged, indicated with $\bar S$ and non averaged quantities. \\ 
In either cases, we will quantify the agreement between the reconstructed quantities and the expected ones.\\
A short description of the principle of the algorithm used to construct the time series is also given.

\section{Gaussian noise} 
This section is devoted to the analysis of the optimizations methods, applied on pure Gaussian noise process $\vec n = (n_1, n_2, \dots, n_N),\text{ with } n_i = n(t_i)$, assuming our process to be a wide-sense stationary process, i.e. autocorrelation matrix is a Toeplitz matrix depending only on time intervals.\\ 
The spectrum chosen for our first analysis is a Gaussian distribution
\begin{equation}
    \nonumber
    S(f) = \frac{1}{\sqrt{2 \pi} \sigma}e^{-\frac{(f - \mu)^2}{2 \sigma ^2}}
\end{equation}
with mean $\mu =2.5 \quad a.u.$ and standard deviation $\sigma =  0.5 \quad  a.u.$. 
To generate noise with a power spectrum that is known, we have to first understand what is the role played by the PSD. \\ 
Via MAXENT principle, we can state that noise in normally distributed with 0 mean and some covariance matrix $C_{ij} = C(t_i, t_j)$, i.e.
\begin{equation}
    \nonumber
    p(\vec{n}(t) \vert I) \propto \exp{\left\{-\frac{1}{2}n_i C^{-1}_{ij}n_j\right\}},
\end{equation}
\begin{figure}
    \centering
        \includegraphics[scale = 0.43]{Images/Noise and PSD/NormalPSDautocorr.pdf}
        \caption{A priori spectrum (left) and autocorrelation (right)}
        \label{fig:autocorr}
\end{figure}and, since process has zero mean, autocovariance and autocorrelation matrices are identical. \\ 
Using the properties of Toeplitz matrices (see appendix \ref{app:Toeplitz}) and the relation between autocorrelation and power spectral density, it is immediate to verify that, in the limit of large number of time samples, the distribution of noise in frequency domain is given by:
\begin{equation}
    p(\tilde{n}(f_i) \vert I) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}},
\end{equation}
so that the role of PSD appears to be clear: it is the diagonal covariance matrix for the noise distribution in Fourier Domain: \\
\begin{equation}
    \nonumber
    <\tilde n_i \tilde n_j > = S(f_i)\delta_{ij}.
\end{equation}
With this information, it is easy to generate any time series with power spectral density of any shape.
Choosing a priori values for sampling interval $dt$ and total time of observation $T$, recalling Nyquist theory of sampling, we generate arrays of given length N / 2 in frequency domain from 0 to the Nyquist Frequency $f_{Ny} = \frac{1}{2 dt}$. We then convert noise to the time domain using an Inverse Fast Fourier Transform.\cite{oliphant2006guide}\cite{van2011numpy} \\ Basic idea for fast algorithm used to construct noise can be found in ~\cite{noiseGen}. We worked it out in a slightly different way, the imaginary component is not obtained sampling a random complex phase from a uniform distribution. We choose to sample both real and imaginary part from two identical normal distribution distributing variance in a proper way between real and imaginary components. 
\section{Test on Time Series}
It is now possible to test the three methods on a given time series. We will concentrate on a power spectral density that is a normal distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$.
We will apply Burg method with each optimizer on a large amount of simulated dataset, obtaining a variety of estimate for PSD. We will study statistical properties for both single reproductions and ensemble average. In many cases, such as gravitational waves, one cannot observe multiple time series realizations for a fixed process, so that single reproduction studies are probably more useful. Still, study of ensemble average is important to highlight some interesting properties. In this way we can fully characterize the performance of each optimizer 
We will attentively study the role of the filter length estimate M in determining the accuracy of the reconstruction. We will consider every optimizer separately and compare result only at the end of this section. We will then study in depth the Advanced LIGO design sensitivity power spectrum \cite{Ligo}
\subsection{Normal Power Spectral Density}
\begin{figure}
    \centering
        \includegraphics[scale = .43]{Images/NormalPSD/NormalNoise.pdf}
    \caption{Noise in both time and frequency domain. Frequency noise behaviour is representative of the chosen spectrum, having its maximum amplitude around $f = 2.5 au$}
    \label{fig:noise}
\end{figure}
Fig \ref{fig:noise} shows a realization of the gaussian process in the frequency (left panel) and time domain (right panel).   \\ 
To study noise with normal-shaped PSD, we generate 1000 datasets of 3000 points each. This will provide a number of samples that is large enough to perform statistical analysis. For every dataset we will reproduce its spectrum with every optimization method. We will show that the quality of reconstruction is very different in different parts of the spectrum: the minimum of the error will be achieved around the peaks, increasing fastly in the tails
 \\ \\ 
\subsubsection{Final Prediction Error (FPE)}
The estimate of the recursive order obtained with FPE on the 1000 realization of noise, is found to lay between 30 and 70. Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum: a logarithmic plot of the mean of all the spectra versus the true spectrum is reported in \textbf{Figure \ref{fig:FPEmean}} with 90\% credible regions. It is evident that, despite not too large error bars, the tail of the normal distribution is not comprised in the interval of the estimation. 
\begin{figure}[H]
    \centering
    \includegraphics[scale = .4]{Images/NormalPSD/FPEpsdAndResiduals.pdf}
    \caption{Ensemble Average for FPE reconstruction with 90\% Credibility Regions (above) and percentage error (below)}
    \label{fig:FPEmean}
\end{figure}
To assess the magnitude of the disagreement, we compute the relative error between the mean spectrum $\hat{S}$ and the true spectrum $S$ for every frequency bin
\begin{equation}
    \nonumber
    r(f) = \frac{|\hat{S}(f) - S(f)|}{S(f)}\,.
\end{equation}
We found this to be a better choice than residuals since relative errors are not dependent on the magnitude of the data. Second Fig \ref{fig:FPEmean} (below) shows that the largest error in the reconstruction of the spectrum is in the tails, while in the bulk of the distribution we have a good approximation of the true function. 
For the entire spectrum, the mean relative error $\Bar{r}$ is 
\begin{equation}
    \nonumber
    \bar{r}_{\bar S, FPE} = 0.025.
\end{equation}
FPE performs well in reconstructing the spectrum even with relatively short time series, with a deviation from the true spectrum around 2\%, but it is evident that it is not able to capture all the features of the spectrum, since the reconstruction in the tails is inaccurate. If we compute the average percentage error in three standard deviations from the mean we gain an order of magnitude estimate of the accuracy of the reconstruction, obtaining: 
\begin{equation}
    \nonumber
    \bar{r}_{\bar S, FPE}^{\mu \pm 3\sigma} = 0.005. 
\end{equation}
and we can state the the method is highly performant in this frequency window. 
Since it is not always possible to work with more then one realization for a process, as in observational experiments, it is important to understand how this method behaves when we apply it on a single time series. In that case, the mean error and its standard deviation are: 
\begin{equation}
    \nonumber
    \bar r_{FPE} = 0.16; \quad \sigma_{r} = 0.16\,.
\end{equation}
The average error of a reconstruction based on a single time series is bigger than the one obtained considering the average of all the reconstructions.
Standard deviation of the errors assume small values and 90\% credibility regions (Cr.) in \ref{fig:FPEmean} are tight. This means that if we randomly take two of the reconstructed spectra, we expect their differences to be statistically small. These facts are evidence for FPE to be a relatively stable optimizer. 
We already said that the estimate for filter's length lays in a well defined range of values. The quality of the reconstruction, as noted in Chapter 1, depends on the order of the reconstructed AR process. 
To investigate this idea further, we plot the value of the optimizer for each recursion and the residuals as a function of the estimate for the order of the autoregressive process (\textbf{Figure \ref{fig:FPEErrorOrder}}),

\begin{figure}[H]
    \noindent
    \centering
    \includegraphics[scale = .3]{Images/NormalPSD/FPEoptimizer.pdf}
    \caption{Stability of FPE, that results in a cluster of values in a small region (left), is due to the typical behaviour of the optimizer as a function of the AR order (right).}
    \label{fig:FPEErrorOrder}
\end{figure}
Foremost evidence for the stability of this optimizer is found in its values as a function of the AR order. Figure \ref{fig:FPEErrorOrder} shows that Values tend to decrease monotonically until a sharp and well defined minimum is reached, to start growing again after this area. This is representative of the typical behaviour of this optimizer as a function of the length of the filter, and explains why the estimates for AR orders are confined in a small region. This plot also shows that most of the values collapsed in the region of AR orders comprised between 30 an 50. There are only few values outside this interval, and get fewer as order's estimate increases. In this region, there is no strong evidence of some correlation between the estimate for M and its associated relative error. It is important to notice that the estimate associated with the higher values of M (around $M > 60$) exhibit the highest 'mean' relative error. \\ 
We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length M is clustered in a region where there is no dependence of error from M. 
\subsubsection{Optimum Bayes Decision Rule (OBD)}
The second optimizer in consideration is OBD.  (\textbf{Figure \ref{fig:OBD3000}}) shows a good agreement between the average over the reconstructions and true spectrum.
\begin{figure}[H]
    \centering
    \includegraphics[scale = .45]{Images/NormalPSD/OBDpsdAndResiduals.pdf}
    \caption{Reconstruction of the spectrum from 1000 simulations with OBD optimizer. Plot is mean of all the results with 90\% credibility regions (above) and percentage error (below)}
    \label{fig:OBD3000}
\end{figure}
The same behaviour of FPE is observed, a good quality reconstruction at the center of the distribution with quite small Credibility regions (Cr.) at a $90$\% credibility level.
Qualitatively they behave similarly but quantitatively the disagreement of this method is found to be even larger than the one obtained with FPE, as can be noticed in the plot of relative errors in figure \ref{fig:OBD3000} (below) where at the extremes the inaccuracy is even larger than 300\% with respect to the real value of the spectrum. For the overall spectrum and within three standard deviation from the mean, we find that the reconstructed spectrum has a deviation of
\begin{equation}
    \nonumber
    r_{\bar S, OBD} = 0.15, \qquad r^{\mu \pm 3\sigma}_{\bar S, OBD} = 0.018. 
\end{equation}
When applied to single realizations of the time series, we find
\begin{equation}
    \nonumber
    \bar{r}_{OBD} = 0.237, \qquad \sigma_{r, OBD} = 0.0296.
\end{equation}
It is interesting to notice that overall relative error of the average spectrum is of the same order of magnitude of the average single reconstruction for the spectrum. Averaging over a 1000 realisations does not substantially improve our results. As a function of the filter length, the values taken by the OBD cluster in a small region, indicating the stability of the reconstructed process order $M$.

\begin{figure}[H]
    \centering
    \includegraphics[scale = .3]{Images/NormalPSD/OBDoptimizer.pdf}
    \caption{Stability of OBD, that results in a cluster of values in a small region (left), is due to the typical behaviour of the optimizer as a function of the AR order (right)}
       \label{fig:OBDErrorOrder}
\end{figure}we see that again our results are clustered in a very small region, due to the behaviour of the optimizer that converge in a small area before growing again. The estimate of the length of the filter is around $p = 20 \div 30$.Note the difference compared to FPE (see \ref{fig:FPEErrorOrder}); for OBD the mean relative error depends on $M$ as
\begin{equation}
    \hat r \sim M ^ {-2.016 \pm 0.006}.
\end{equation}
In summary, OBD shows greater error wherever in the spectrum, due to the fact that it choose filters that are too short and in a region where quality drastically change when changing M. In this case we can say that we have a stable estimate for M, but due to this dependence, the estimate is not as stable as FPE's. 
\subsubsection{Criterion Autoregressive Transfer function (CAT)}
This last criterion differs substantially from the two other optimizers. Figure \ref{fig:CATmean} (above) shows a good agreement between the estimate of the spectrum for every frequency
\begin{figure}[H]
    \centering
    \includegraphics[scale = .43]{Images/CAT/CATnormal.pdf}
    \caption{Mean spectrum for CAT optimizer, 1000 simulations for 3000 points datasets}
    \label{fig:CATmean}
\end{figure} and also tails are well captured by the average spectrum. However, the variance of the reconstructed spectrum is quite large, much larger than for FPE and OBD. In general, the relative error is nearly constant, $\sim 10\%$ over all frequencies, see  Fig. \ref{fig:CATmean} (below). The residuals, averaged over all frequencies, are: 
\begin{equation}
    \nonumber
    \bar r_{\bar S, CAT} = 0.016 \qquad \bar r^{\mu \pm \sigma}_{\bar S, CAT} = 0.017.
\end{equation}
the overall reconstruction is thus quite accurate, with deviations from the simulated spectrum of the order of $1\% \div 2\%$.\\
The picture drastically changes in the case of individual time series realizations. The average residuals are, in fact
\begin{equation}
    \bar r_{CAT} = 0.42 \qquad \sigma_{CAT} = .44\,.
\end{equation}
The error is indeed $\sim 42\%$, hence much larger than in the case of FPE and OBD. Both large error bars and the large standard deviation are indications of the instability of this optimizer. This is exemplified by the variety of different spectra that can be obtained considering different time series.

\begin{figure}[H]
    \centering
    \includegraphics[scale = .43]{Images/CAT/CAToptimizer1.pdf}
    \caption{CAT is not a stable optimizer since its optimizer do not converge to any specific region of the AR estimate order}
    \label{fig:CAToptimizer}

\end{figure}
The reason for this behaviour, can be found in \textbf{Figure (\ref{fig:CAToptimizer})}. CAT does not converge to any specific value for the order of the autoregressive process. In turn, this results in estimates for the order that span a large range.  Figure \ref{fig:CAToptimizer} also shows a strong dependence of the error on the estimated length of the filter. The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like the big excursion in the size of datas of a normal distribution, but this length is also responsible of an augmented variance in the estimate, introducing spurious peaks in the reproduction (Figure \ref{fig:GaussMedian})

Averaging, the noise in each spectrum is cancelled. It is interesting that the cancelling of errors in averaging is property of gaussian noise (the process is gaussian, after all). This implies, in a sense, that each estimate of the spectrum is independent of any other, as suggested by the huge variance in the residuals. This instability is not a good property for an estimation, so that this estimator might be found to be optimal for repeatable experiments and not for single observation of some process. Another approach to this method will be proposed in section \ref{sec:ACAT}

\begin{figure}
    \centering
    \includegraphics[scale = .45]{Images/NormalPSD/Gaussmedian.pdf}
    \caption{Reconstructed power spectral density with the three method. The reported reconstruction is the one that shows the median percentage error (overall)}
    \label{fig:GaussMedian}
\end{figure}
\subsection{Summary and Comparison}
\begin{figure}
    \centering
    \includegraphics[scale = .3]{Images/NormalPSD/Normal Residuals.pdf}
    \caption{Percentage error in the reconstruction for every optimizer, as a function of the frequency. Lower panel restricts the domain to three standard deviation to increase the definition}
    \label{fig:NormalEnsembleResiduals}
\end{figure}
In our analysis, FPE and OBD optimizers are found to behave similarly while CAT shows completely different properties. Studying the average spectrum is not sufficient to choose which method behaves better. \\ 
The bad behaviour in the reconstruction of the tails from OBD and FPE might be due to the fact that very small values require longer filters. 
In fact when we consider the average for all the spectra we note that the shorter the filter the larger the error in the tails: OBD, which select the shortest filters, arrives to a $300 \%$ error at the extrema of the power spectral density, FPE which select intermediate order has got a $85\%$ error in same window, while CAT, that selects a wide range of long filters, has got a constant accuracy everywhere in the spectrum.  \\ 
Figure \ref{fig:NormalEnsembleResiduals} shows that CAT is the method that minimizes errors in the tails (upper plot), but if one restrict to three standard deviations from the mean, FPE is found to reach best accuracy (lower plot). 
The frequency averaged percentage error is reflecting this behaviour. Considering the whole reproduction, error's average and standard deviation are found to be: 
\begin{align*}
\centering
&r_{\bar S, FPE} = 0.025; \quad \sigma_{\bar S, FPE} = 0.083;  \\
&r_{\bar S, OBD} = 0.15;\quad \sigma_{\bar S, OBD} = 0.42 ; \\ 
&r_{\bar S, CAT} = 0.015; \quad \sigma_{\bar S, CAT} = 0.012
\end{align*}
In is interesting to see that the average for FPE and OBD is strongly affected from the high value of the error in the tails of the distribution. If one consider the three standard deviations from the mean, the value of the error substantially decreases, 
\begin{align*}
\centering
&r_{\bar S, FPE} = 0.005; \quad \sigma_{\bar S, FPE} = 0.004;  \\
&r_{\bar S, OBD} = 0.018;\quad \sigma_{\bar S, OBD} = 0.019 ; \\ 
&r_{\bar S, CAT} = 0.015; \quad \sigma_{\bar S, CAT} = 0.011
\end{align*}
showing that the overall percentage error is just a first indicator for the quality of the reproduction, but is not sufficient to perform a full characterization. \\
It is evident that tails do not affect CAT  average error and standard deviation at all, so that it's accuracy is approximately constant everywhere in the spectrum. \\ \ 
Since both FPE and CAT shows pros and cons, at this level, there is no reason to choose between them.\\ 
Analyzing the resolution of the single reconstruction, figure \ref{fig:optcomparison} clearly shows that FPE minimize overall relative error giving a stable result.\\ 
CAT instead shows a larger error and do not converge to any specific value. Almost every reproduction obtained with CAT is worst with respect to those obtained with FPE, but their mean is more precise. \\ 
It's instability is probably the cause for this behaviour. Since it does not show a strong convergence to some order with respect to others, very different lengths are selected.
Longer lengths are able to capture a lot of details but are more likely to introduce noise. In other words, long filters reduce bias but increment the variance of the result. 
Taking the average, for central limit theorem the error decreases as $\sim \frac{1}{\sqrt{N}}$. In this way we compensate noise and obtain a spectrum with small bias and small fluctuations.
Considering the average, CAT reconstruct precisely and with good accuracy every part of the spectrum, even if in "regular" parts best accuracy is reached from FPE. For a single reproduction, CAT is very noisy and unstable, while FPE performs the better. \ref{fig:optcomparison}
\begin{figure}
    \centering
    \includegraphics[scale = .3]{Images/NormalPSD/NormalPSDcomparison.pdf}
    \caption{Relative error as function of the length of the filter.}
    \label{fig:optcomparison}
\end{figure}
\begin{figure}
    \hfill
    \centering
    \includegraphics[scale = .45]{Images/NormalPSD/ordersComparison.pdf}
    \caption{Histogram of estimate orders for each method}
    \label{fig:ordersCompairson}

\end{figure}
definitely shows that FPE gives the best results: it is a stable method that choose filters in the clustered area associated with minimum error, while both OBD and CAT result lies outside this area and show bigger errors. Two other interesting plots that stress this fact are reported in figure \ref{fig:ordersCompairson} and \ref{fig:residualsComparison}
\begin{figure}
    \centering
    \includegraphics[scale = .46]{Images/NormalPSD/ResidualsComparison.pdf}
    \caption{Histogram of residuals for each method}
    \label{fig:residualsComparison}
\end{figure}
where the histograms for the errors and for order's estimate are reported. These graphs provide stronger evidence for our previous statements. As already noticed, FPE is very stable and provide the smallest 'overall relative error'. \\
CAT histogram shows a very interesting behaviour. The histogram for the order show two different peaks,  a small one associated to 'short filters', and a second high peak at $m = M$. This means that CAT is most likely to not converge at all at some value for the regression order, and it is more likely selects a length that is comparable with the largest available.  If short filters are selected, the overall percentage error will be low, respectively represented by the first peak in the distribution of the errors, while if long filters are selected, the error is maximized.

So, even if in some cases CAT might be more accurate when taking the average over several spectra, FPE guarantees that minimum error will be reached.

\subsection{Analysis of Ligo Spectrum}
\begin{figure}
    \centering
    \includegraphics[scale = .45]{Images/LIGOsimulate/LigoPSDoriginal.pdf}
    \caption{Original power spectral density from Ligo}
    \label{fig:LigoPSDOriginal}
\end{figure}
In this subsection we concentrate again on the reconstruction of a specific, known, power spectrum.
We analyze the properties of reconstruction of this second spectrum whose shape is the Advanced LIGO design sensitivity theoretical spectral curve (Figure \ref{fig:LigoPSDOriginal}). 
For this analysis we will study properties only for arrays of 40960 points. We consider a sampling rate of $2048 samples / second$ that means we can reconstruct a Nyquist frequency of $1024 Hz$. We chose a longer dataset than before because algorithm works with a 1 dimensional interpolation of spectrum, and with shorter datasets the interpolated spectrum is not consistent with the original spectrum and does not capture the height of the peaks. Shorter choices would result in noise extracted from a different power spectrum than the one in figure \ref{fig:LigoPSDOriginal}. \\ 
We will study statistical properties from samples obtained with 500 simulations and compare the obtained results for each optimizer. 
We start considering statistical properties of mean spectrum obtained averaging over all the simulations and conclude with an analysis for the statistics of the reproduction of single spectrum. 
\subsubsection{Properties for the average spectrum}
Here we report mean properties of the reproductions of the spectrum. The spectrum we want to reconstruct is reported in Figure \ref{fig:ligospectrum},
\begin{figure}
    \centering

        \includegraphics[scale = 0.45]{Images/LIGOsimulate/PSD.pdf}
        \caption{A priori spectrum from Ligo database}
        \label{fig:ligospectrum}
\end{figure}
shows the original power spectral density and the ensemble average of every method. For the mean, the relative errors are: 
\begin{equation}
    \bar r_{\bar S, FPE} = 0.0426\%, \qquad \bar r_{\bar S, OBD} = 0.129\%, \qquad \bar r_{\bar S, CAT} = 0.0211\%
\end{equation}
Again, if one only check overall accuracy for every method, CAT is outperforming. In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them. To check this behaviour we take a look at the two peaks separately. For the first one a log plot of the reconstructed peaks and their relative errors are shown in Figure \ref{fig:ligo1peak} 
\begin{figure}[h]
        \includegraphics[scale = 0.45]{Images/LIGOsimulate/1stPeakComparison.pdf}
        \caption{1st peak of the spectrum and it reconstruction with every optimizer (left) and associated percentage error (right)}
        \label{fig:ligo1peak}
\end{figure}
and the results reinforce our previous statement. The behaviour of the error is very similar for CAT and FPE, showing similar shapes but very different sizes, while CAT is giving a different and way more precise result. Both FPE and OBD are not able to capture the width of the peak and also show larger errors, while CAT has a good behaviour in reconstructing width's also. 
The reconstruction for very sharp peaks considering the average over all the spectra is found to be strongly inaccurate for both FPE and OBD, while a good accuracy is reached with CAT method. \\ 
Considering the estimate for the position of the peak, whose true value is found to be $16.85 \quad a.u. $, within numerical accuracy both FPE and CAT find its exact position, while OBD shift the peak left at $16.7 a.u.$, with a $0.9\%$ inaccuracy.\\ 
Second peak results are analogous, plotting the peak and the errors value around it
\begin{figure}[h]
    \centering
        \includegraphics[scale = 0.45]{Images/LIGOsimulate/2ndPeakComparison.pdf}
        \caption{2nd peak of the spectrum, Relative error around 2nd peak of the spectrum}
        \label{fig:ligo2peak}
\end{figure}
we find again that FPE and OBD are performing similarly, the magnitude of the error is maximum for OBD and minimum for CAT. \\ 
In this case even CAT shows large errors in some points, but they are reasonable in a large part of it, still, it remains the best reconstruction.  For this second peak, all the methods find maximum position to be at $438.05 \quad a.u.$, which is the true position of the maximum.
Looking at Figure \ref{fig:ligo1peak} at \ref{fig:ligo2peak}, it appears to be clear that the least of the accuracy is not on the peak, but immediately around, so that one of the biggest deal is the reconstruction of peak's width, that only CAT is able to capture with high definition. \\ 
Even if FPE shows larger error in the peaks than CAT, when the average over all the frequencies is considered, their percentage errors are of the same order of magnitude. This means that outside the peaks FPE is more accurate than CAT, i.e. CAT is still giving noisy results.\\ 
As for the normal distributed noise, mean spectrum is not sufficient to declare what is the best optimizer to be used, since most of experiment are purely observational and mean spectrum over different realizations of the same stochastic process cannot be computed. So, it is important to also understand what is their behaviour for single realizations of the process.
\subsubsection{Statistical properties of single realizations} 
The errors for the single realizations behave very differently from the errors obtained when we considered the mean spectrum. In computing the mean of the overall errors for each single spectrum, the results we obtain are the following:
\begin{equation}
    \bar{r}_{FPE} = 0.13 \pm 0.01; \qquad \bar{r}_{OBD} = 0.17 \pm .02; \qquad 
    \bar{r}_{CAT} = 0.4 \pm 0.1
\end{equation}
where the error is just the standard deviation of the samples.
In this case we find a different result: FPE shows the best results and order's estimate is associated with least errors. Also, having found a smaller standard deviation of the datas, we can state that FPE and OBD are more stable optimizers then CAT. In fact, their properties as a function of filter's length are almost identical to those obtained in the study of normal spectrum. If we again plot the value of the errors as a function of the estimate of the order (Figure \ref{fig:LigoOrderError})
\begin{figure}
    \centering
    \includegraphics[scale = .43]{Images/LIGOsimulate/orderVSerror.pdf}
    \caption{Plot of relative error as a function of the estimate for filter's length}
    \label{fig:LigoOrderError}
\end{figure}
it is immediate to see that, again, FPE results lies in the are of filter's length associated with a minimum of relative errors, OBD lies in a clustered window with a higher error and CAT is not convergent, but most of the results are outside the minimum error area. So, we arrive at the same conclusion we had when we considered the normal specturm. OBD has no general properties that make it a preferable optimizer then FPE or CAT while, what to choose beneath the last two is based on the problem we are dealing with: if we have one single realization for the process, we are mostly sure that FPE would catch the area associated with the best resolution possible, while CAT would result in a spurious result with, in general, a not so good resolution. But, in case we have several realization for the same process, this last property of CAT, together with its instability, result to be his luck and, taking the average errors tend to compensate noise in the result, originating a very well defined spectrum that show same resolution almost everywhere, no matter how small the values are (as for normal PSD case) or how sharp peaks can be, the result has got generally good accuracy, also if it is clear that we have to pay this fact with very large error bars. So a choice between these methods is a matter of convenience and interest for each specific problem. \\
\begin{figure}
    \centering
    \includegraphics[scale = .43]{Images/LIGOsimulate/LigoHists.pdf}
    \caption{Histograms showing the estimate for the autoregressive order and overall percentage error obtained with all the three optimizers.}
    \label{fig:my_label}
\end{figure}

\section{Another approach to CAT} \label{sec:ACAT}
We stated that, if one consider a single realization for some stochastic process the choice for the optimizer should rely on FPE because of its stability and better single estimate, while CAT, when applied on a single dataset, does not assure a good resolution for our estimate. \\
It was found that CAT shows its best properties when considering the average over a big number of reconstructions, where noisy results tend to compensate each other reaching a very good quality reconstruction even for not too long arrays (as in the case of Gaussian PSD where we considered simulations of 3000 points). This led us to propose another approach to CAT: instead of applying Burg's method on the whole data set, we can divide it in smaller arrays of data and apply it to each one of them taking their mean at the end of the computation. In this way, accuracy can be made incredibly higher than the one obtained with the standard approach. \\ 
Unluckly, it is very hard to investigate how one should divide the dataset to reach minimum error. If one consider a known power spectrum, the problem could be solved numerically, but unuckly in real life experiments we do not know what true spectrum is, and such an approach is non-sense. \\ 
The non linear form for the estimator together with the finite-order recursion make it hard or even impossible to find general properties for the estimate, and so we will not investigate many of that questions that could arise.\\
For examples, what is the best choice for the length of the sub data-set? If we choose too short arrays, we will mediate over a lot of different spectra, decreasing variance, but it is probable that we will not be able to capture all the features of the spectrum. Too long subsets, instead, would result in too few reconstructed spectra, so that taking their average would not reduce variance in a significant way. It seems a very hard task to answer this question uniquely, and maybe such an answer does not even exist in a general case. One prescription on the length of the sub arrays may concern some consideration about frequencies: if we want to accurately reconstruct low frequency behaviour, one has to make long time observations, so for those cases longer datasets are required.  
Since taking overlapping arrays does not seems to affect negatively single estimates, this could be a good prescription to maximize the number of reconstructions for a given length. \\
Stressing to its maximum the hypothesis that such a procedure is harmless, one could try any overlapping sequence to maximize the number of reconstructions. To study this approach, since we want every reconstruction to be more than slightly different from each other, we chose an overlapping that is half the length for the sub arrays. 
We test this new approach on LIGO dataset on the same noise simulations as before. We chose sub segments of length L = 3000, 5000, 7000 and 10000.. \\
In figure \ref{fig:alternativeCAT} we compare the result associated with with this new approach for the lengths written above. For each of the 500 reproductions, we chose the one associated with the median overall error. 
\begin{figure}
    \centering
    \includegraphics[scale = .45]{Images/AlternativeCAT/AlternativeCAT.pdf}
    \caption{Alternative CAT approach to Ligo Spectrum for different lengths of the sub arrays}
    \label{fig:alternativeCAT}
\end{figure}
As one can expect from CAT behaviour results obtained with larger arrays are associated with increasing noise. Considering the reconstruction for single peaks, it is found that accuracy is certainly dependent from the length, but it seems to give the best result for $L = 7000$, and not for $L = 10000$ as we were expecting. This difference can be understood if one considers the number of realizations over which we take our average: for arrays of $7000$ points we are taking the average over 12 different realizations, while for arrays of $10000$ points we are taking our average on eight different realizations only. This is probably the reason why this higher length is not associated with higher resolution. 
We clearly see that the reconstruction achieved using $L = 7000$ is the best in capturing the width of the peaks and their position, but none of the reconstructions is able to accurately reproduce the second peak. 
We want to compare these results with the standard approach, and with both the results obtained with FPE and with CAT.

\subsection{Alternative CAT vs FPE}
The first comparison we make between FPE and alternative CAT regards the distribution of the overall errors. 
\begin{figure}
    \centering
    \includegraphics[scale = .45]{Images/AlternativeCAT/histACatVSFpe.pdf}
    \caption{Histogram of the percentage error as a function of the length of the sub arrays. The blue panel shows the histogram obtained from FPE}
    \label{fig:aCATvsFPEhist}
\end{figure}
Figure \ref{fig:aCATvsFPEhist} shows the distribution of the percentage error as a function of the length of the segments of the data. The support of the histogram is, for every length, shorter than the support the one expect to be obtained with CAT (y-axis of figure \ref{fig:LigoOrderError}), so that this approach is found to be more stable than standard CAT. \\ 
Shorter arrays are found to minimize the error, as could be expected from the fact that the variance is reduced from this choice, but non of the chosen lengths reaches the overall accuracy of FPE. \\ 
From the previous histograms, the most meaningful comparison of FPE with alternative CAT is obtained using as preferable length the reconstruction '3000'. \\ 
The mean and standard deviations of the percentage error are, in the two cases 
\begin{align} 
    \bar &r_{FPE} = 0.13; \qquad 
    \sigma_{FPE} = 0.01 \\ 
    \bar &r_{aCAT} = 0.158; \quad 
    \sigma_{aCAT} = 0.007
\end{align}
With this alternative approach, we have been able to obtain reconstructions for the spectra whose accuracy is almost comparable with the one obtained with FPE. But the comparison of the overall error is just a rough indicator of the accuracy in the reconstruction, for that reason we compare the median error reconstruction of both methods. 
\begin{figure}
    \centering
    \includegraphics[scale = .45]{Images/AlternativeCAT/FPEaCATwhole.pdf}
    \caption{Comparison of FPE method with alternative cat with length of 3000 points (above panel) and their percentage error (below panel)}
    \label{fig:aCATvsFPEwhole}
\end{figure}
Figure \ref{fig:aCATvsFPEwhole} shows both the reconstruction obtained with FPE and alternative CAT (above panel) and the percentage error for each frequency (below panel). It is evident that the two result are very similar. Alternative CAT provides a sightly better accuracy in the reproduction for the first peak. But, for the rest of the reconstruction FPE has to be preferred. Regarding the second peak, FPE accuracy is largely better. Alternative CAT flattens the peak and its maximum is not localized, while the reconstruction of FPE is representative of true position of the peak. Also, in the rest of the spectrum the result of FPE still shows less noise. \\ 
From thie brief comparison, we can conclude that alternative CAT approach shouldn't be used with "short" arrays (short with respect to the total length N of the time series), since FPE provides a better estimate. \\ 
Since for short arrays the best quality is still achieved by the use of FPE, a second interesting comparison is the one between FPE with the alternative CAT estimate on longer arrays. From figure \ref{fig:alternativeCAT}, it is evident that the best compromise between noise and accuracy in the reproduction is obtained by the arrays of 7000 points. It is clear that, outside peaks, the smaller error is achieved by FPE, so that te interesting part will be the expected gain in the definition of the peaks.
\begin{figure}
    \centering
    \includegraphics[scale = .43]{Images/AlternativeCAT/FPEaCAT7000.pdf}
    \caption{Comparison of the alternative CAT with 7000 points subsegments of data with FPE}
    \label{fig:aCAT7000FPE}
\end{figure}
Figure \ref{fig:aCAT7000FPE} shows the reconstruction obtained with FPE (red line) and the one obtained with the alternative CAT approach, for sub-segments of 7000 data and a 50\% overlapping. This result is more interesting than the previous. As expected, the noise in the reproduction is larger, as can also be noticed looking at the error in the below panel, but the introduced noise is a reasonable quantity.  Concerning the peaks, the first peaks reconstruction has gained a large accuracy in the reproduction, for both its width and position. The second peak is reconstructed in a comparable way for both the methods. FPE captures in a better way the position of the peak, while alternative CAT gains in the reproduction of the width and general shape. \\ 
Concluding, this new approach might be a valid alternative to FPE if 'not-too-short' sub segments of data are chosen. In this way, introducing a reasonable amount of noise, we are able to capture some details with more accuracy.
But if we deal with shorter datasets, like 3000 points, without any a priori knowledge on the shape of the spectrum, is this alternative method convenient? In chapter 4 we will see that this method is the best to characterize brown noise spectrum even with dataset of $1000$ points only, but such an investigation would require a long analysis and would deviate too much from our main interests. 
Since FPE has minimum requirement of external input we will prefer this standard approach, even if this alternative approach seems to have its merits. 
\subsection{Alternative VS Standard CAT}
We now compare the performance of this alternative approach with the standard approach with CAT optimizer. 
\begin{figure}
    \centering
    \includegraphics[scale = .45]{Images/AlternativeCAT/AlternativeVSstandard.pdf}
    \caption{Spectral reconstruction obtained with standard CAT approach and alternative approach}
    \label{fig:alternativevsstandard}
\end{figure}
Figure \ref{fig:alternativevsstandard} show that the alternative approach provides a reconstruction that presents a largely decreased quantity of noise. This decreased level of noise has as counterpart a decreased accuracy in the reconstruction of the width of both the peaks. But, even with this loss, the estimate is still very representative of the obtained spectrum, and might be a good price to pay to reduce the variance of the estimate.
\begin{figure}
    \centering
    \includegraphics[scale = .45]{Images/AlternativeCAT/CATvsALTerrors.pdf}
    \caption{Percentage error of spectral estimate with CAT (red) and alternative CAT (green) }
    \label{fig:CATALTerrors}
\end{figure}
Figure \ref{fig:CATALTerrors} shows that loss in the definition due to the averaging over the subset is concentrated in a very small number of points of the estimate, and  negligible with respect to the gain obtained in the whole spectrum. \\ 
In this case, the reduction of noise is important because we selected the spectrum associated with the median overall error, highly influenced from noise. If CAT is 'lucky enough' to select filters whose length is comparable with FPE, probably such an achievement wouldn't be reached, but we expect the alternative estimate won't show any big loss. So, since this alternative approach gives better result for long filters and comparable results for shorter filter, being also more stable, is probably to be preferred to the standard approach with CAT as optimizer.
Clearly, such an approach has a great disadvantage: we lose all the information regarding the prediction error filter. So, if we want to characterize the process, FPE is with no doubt the best choice. While, if we are only interested in reconstructing the power spectral density, alternative CAT's approach might be worth an attempt. 
\section{Comparison of MESA with Welch Method}
We now compare the performance of MESA algorithm with estimate obtained with Welch method \cite{Welch1967}. Welch introduced his method in 1967, and it is based of Fast Fourier Transform. 
Estimate is obtained averaging over modified peridiograms computed on subset of the original dataset. Here, as in the alternative approach to CAT, one has to choose an opportune length for the arrays and their overlapping. Else, a window is required to emulate period of the Fourier Transform, and with this choice one is obliged to assume data to be zero outside this window. The shape of the reconstruction is strongly dependent from window's choice, so, in general, Welch method is not a stable estimate of the spectra, since it strongly depends on this arbitrary choice. We compare the reconstruction obtained with the two methods in figure \ref{fig:WelchVSfpe}

\begin{figure}[H]
    \centering
    \includegraphics[scale = .3]{Images/WelchVSfpe.pdf}
    \caption{Comparison of the reconstructed spectrum obtained with FPE (red line) and Welch's method (green line)}
    \label{fig:WelchVSfpe}
\end{figure}

Graphically both methods are found to perform similarly. Since MESA is analytical, graph obtained with this method is smoother than one obtained with Welch. In fact, Welch spectrum is obtained performing a FFT and so its output length is related to the length of the subsets. Computing percentage errors, we obtain
\begin{equation}
    \nonumber
    \bar r_{MESA} = 0.11; \qquad \bar r_{Welch} = 0.26
\end{equation}
so that better overall accuracy is reached with MESA. A plot of both the errors of FPE and Welch method is reported in figure \ref{fig:WelchError}

\begin{figure}
    \centering
    \includegraphics[scale = .3]{Images/WelchErrors.pdf}
    \caption{Comparison of percentage error of reconstruction with MESA and Welch method}
    \label{fig:WelchError}
\end{figure}It is evident that greater error for Welch is due to two main reasons. First reason is the reconstruction of the second peak. While for the first peak accuracy of the two methods is comparable, regarding the second peak Welch method shows a huge disagreement on the peak, with a value of percentage error near 4000\%. Clearly, being one in more than 1000 points, its own contribute to overall error is not sufficient to justify the different accuracy of the two methods. To explain this we have to look at noise, in fact, it is evident that in the other parts of the spectrum Welch method shows an error that is almost everywhere bigger than the one obtained with MESA. Taking care of noise justify the big difference of the accuracy between the two methods. 
We can conclude that, even if Welch method is a fine reconstruction for the spectrum, MESA has to be preferred for several reasons: firstly, accuracy is mostly everywhere better than the one obtained with Welch method and secondly MESA is more stable, does not require to work with arbitrary choices that significantly change results and even more important MESA does not require a modification of data, being maximally non committal to not available data, while Welch require them to be zero. So we can conclude that MESA is a better approach for both the quality of the reconstruction and scientific reasons. 
\section{Conclusions}
In this chapter we studied the properties for the reconstruction of power spectral density using maximum entropy principle. We tried to understand what is the best criterion to infer the recursion order and that provides the better estimate of the power spectral density. We studied all the optimizers applying them to a normal shaped spectrum Ligo's. FPE and OBD optimizers are found to be very similar optimizers, for both quality of the reconstruction and stability of the method itself, but OBD shows everywhere a larger error. Figures {\ref{fig:optcomparison}} and \ref{fig:LigoOrderError} clearly explain the reason for this lower quality: percentage error is dependent on the estimate for filter's length, and the range where filter's  chosen from OBD lays is associated with higher error with respect to those reached by FPE, where errors seems to be independent from the choice for the autoregressive order and reach their minimum values. So, this allowed us to exclude OBD from the race for the best optimizer. \\
The last comparison was between CAT and FPE. If one only considers the regions of estimated lengths, CAT estimates lay in a region associated with larger errors and it is also found to be a very unstable optimizer, since it does not show any convergence property and it is more likely to choose an order comparable with maximum length available (figure \ref{fig:ordersCompairson}). This behaviour led us to choose FPE as the best method to infer what is the best recursion order that allow us to reproduce the spectrum. But an analysis of the spectrum obtained as an average over all the simulations showed important and very interesting features of CAT: it is able to reproduce the spectrum with the same accuracy everywhere, also where FPE is found to fail. This is probably due to the high variability of the results reproduced from CAT. \\
This led us to propose a different approach to MESA when CAT is used. We apply it to different subsets (of fixed length and overlapping) of the original dataset, taking the average over all the reconstructions at the end. This approach is found to have a better reproduction than the standard approach (figure (\ref{fig:alternativevsstandard})), providing an estimate that reduces noise with a reasonable loss in terms of details.  Unlucky this require to introduce some arbitrary parameters for the estimate, such as the length of the smaller data extracted from the original dataset and their overlapping. A lot of interesting questions arises in the application of this method but, since their investigation is not our purpose, further investigation is postponed.   
At the end we compared the performance of MESA with Welch method and found that MESA has to be preferred for several reasons, firstly because it result in a better quality reconstruction for the spectrum, secondly because it requires a smaller number of 'arbitrary' assumptions, making it a more stable estimator. Thirdly, it requires no assumption on not available data, while Welch method requires the choice of a window that forces the data outside the window to be zero. So, MESA is a better performing estimate with less commitment with data and has to be preferred to Welch method.

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END & BIB

        \begin{acknowledgments}

          R.~G. acknowledges support from the Deutsche Forschungsgemeinschaft
          (DFG) under Grant No. 406116891 within the Research Training Group
          RTG 2522/1. 
          M.~B. and S.~B. acknowledge support by the EU H2020 under ERC Starting
          Grant, no.~BinGraSp-714626.  
          M.~B.~ acknowledges support from the Deutsche Forschungsgemeinschaft
          (DFG) under Grant No. 406116891 within the Research Training Group
          RTG 2522/1. 
          %

          %
          This research has made use of data, software and/or web tools obtained 
          from the Gravitational Wave Open Science Center (https://www.gw-openscience.org), 
          a service of LIGO Laboratory, the LIGO Scientific Collaboration and the 
          Virgo Collaboration. LIGO is funded by the U.S. National Science Foundation. 
          Virgo is funded by the French Centre National de Recherche Scientifique (CNRS), 
          the Italian Istituto Nazionale della Fisica Nucleare (INFN) and the 
          Dutch Nikhef, with contributions by Polish and Hungarian institutes.
        \end{acknowledgments}

	\bibliography{Bibliography.bib}
	\bibliographystyle{ieeetr}

\end{document}

### 2021-02-01 memspectrum.py
Some changes

@@ -362,7 +362,7 @@ class MESA(object):
               optimisation_method = "FPE",
               method              = "Fast",
               regularisation      = 1.0,
-              early_stop          = False):
+              early_stop          = True):
         """
         Computes the power spectral density of the attribute data for the class
         using standard Burg method recursive and a Faster (but less stable) version. 
@@ -389,7 +389,7 @@ class MESA(object):
             Default is 1, which means no regularisation 
                                        
         early_stop: 'boolean'          
-            Default is True. Breaks the iteration if there is no  no new global 
+            Default is True. Breaks the iteration if there is no new global 
             maximum after 200 iterations. 
             Recommended for every optimisation method but CAT.
 

### 2021-02-02 examples/sunspots.py
refactored the code to include several checks

@@ -13,13 +13,14 @@ if __name__ == "__main__":
     datafile = "data/zuerich-monthly-sunspot-numbers-.tsv"
     _, data = np.genfromtxt(datafile, unpack = True)
     T = len(data)
+    print([d for d in data])
     dt = 1.
     srate = 1./dt
     N = data.shape[0]
     f = np.fft.fftfreq(N, d=dt)
     t = np.arange(0,T,step=dt)
-    M = MESA(data)
-    P, ak, _ = M.solve(method = "Fast", optimisation_method = "FPE", m = int(2*N/(2*np.log(N))))
+    M = MESA()
+    P, ak, _ = M.solve(data, method = "Fast", optimisation_method = "FPE", m = int(2*N/(2*np.log(N))))
     print("p = {0}".format(len(ak)))
     PSD, _     = M.spectrum(dt)[:N//2]
 

### 2021-02-02 examples/synthetic_noise.py

@@ -44,8 +44,8 @@ if __name__ == "__main__":
                           asd = False)
     N = time_series.shape[0]
     f = np.fft.fftfreq(N, d=dt)
-    M = MESA(time_series)
-    P, ak, _ = M.solve(method = "Fast", optimisation_method = "FPE", m = int(2*N/(2*np.log(N))))
+    M = MESA()
+    P, ak, _ = M.solve(time_series, method = "Fast", optimisation_method = "FPE", m = int(2*N/(2*np.log(N))))
     print("p = {0}".format(len(ak)))
     PSD     = M.spectrum(dt,f)
 
@@ -57,10 +57,10 @@ if __name__ == "__main__":
     ax.set_xlabel("frequency (Hz)")
     ax.set_ylabel("PSD (Hz$^{-1}$)")
     
-    M = MESA(time_series[:int(0.9*N)])
-    P, ak, _ = M.solve(method = "Fast", optimisation_method = "FPE", m = int(2*N/(2*np.log(N))))
+    M = MESA()
+    P, ak, _ = M.solve(time_series[:int(0.9*N)],method = "Fast", optimisation_method = "FPE", m = int(2*N/(2*np.log(N))))
     Np = 3
-    prediction = M.forecast(int(0.1*N), Np)
+    prediction = M.forecast(time_series[:int(0.9*N)], int(0.1*N), Np)
     l, h = np.percentile(prediction,[5,95],axis=0)
     fig = plt.figure(2)
     ax  = fig.add_subplot(111)

### 2021-02-02 memspectrum.py

@@ -179,7 +179,7 @@ class MESA(object):
     init: data: `np.ndarray` shape (N,)
     solve(): 
     """
-    def __init__(self, data):
+    def __init__(self, *args, **kwargs):
         """ 
         Class that implements Burg method to estimate power spectral densitiy of
         time series. 
@@ -188,9 +188,7 @@ class MESA(object):
         ----------
         data: 'np.ndarray'       
             Time series with power spectral density to be computed 
-        """ 
-        self.data = data
-        self.N    = len(self.data)
+        """
         self.P = None
         self.a_k = None #If a_k and P are None, the model is not already fitted
         self.optimization = None
@@ -209,9 +207,8 @@ class MESA(object):
         if self.P is None or self.a_k is None:
             raise RuntimeError("PSD analysis is not performed yet: unable to save the model. You should call solve() before saving to file") 
         
-        to_save = np.concatenate([[self.P], self.a_k, self.optimization, self.data])
-        header = "(1,{},{},{})".format(len(self.a_k), len(self.optimization), len(self.data))
-        print(header)
+        to_save = np.concatenate([[self.P], [self.N], self.a_k])
+        header = "(1,1,{})".format(len(self.a_k))
         np.savetxt(filename, to_save, header = header)
         return
         
@@ -235,13 +232,11 @@ class MESA(object):
         shapes = eval(first_line.translate({ord('#'): None, ord(' '): None}))
             #checking for the header
         if not isinstance(shapes, tuple):
-            if len(shapes) != 4 or not np.all([isinstance(s, int) for s in shapes]):
+            if len(shapes) != 3 or not np.all([isinstance(s, int) for s in shapes]):
                 raise ValueError("Wrong format for the header: unable to load the model")
 
             #assigning values
-        self.P, self.a_k, self.optimization, self.data = np.split(data, np.cumsum(shapes)[:-1])
-
-        self.N = shapes[3]
+        self.P, self.N, self.a_k = np.split(data, np.cumsum(shapes)[:-1])
         
         return
         
@@ -316,6 +311,7 @@ class MESA(object):
             
         frequencies: 'np.ndarray'        
             (positive) frequencies to evaluate the spectrum at (shape (N,))
+            must be equally spaced
 
         
 
@@ -339,26 +335,24 @@ class MESA(object):
         """
         f_ny = .5 / dt 
         
-        if isinstance(frequencies, np.ndarray): 
-            df = np.min(np.abs(np.diff(frequencies))) * 0.9
[…]

### 2021-02-02 scripts/analyse.py

@@ -10,7 +10,7 @@ import matplotlib.pyplot as plt
 from optparse import OptionParser
 
 if __name__ == "__main__":
-    parser = OptionParser(usage)
+    parser = OptionParser()
     parser.add_option('--data', default=None, type='string', help='input data file')
     parser.add_option('--srate', default=None, type='float', help='sampling rate')
     parser.add_option('-T', default=None, type='float', help='duration of the data')
@@ -21,10 +21,10 @@ if __name__ == "__main__":
     
     srate = opts.srate
     dt = 1./srate
-    T  = opts.T
     datafile = opts.data
-    data = np.loadtxt(datafile)
-    
+    d = np.genfromtxt(datafile, delimiter=',', names=True)
+    data = d['CloseLast'][::-1]
+    T = len(data)
     N = data.shape[0]
     f = np.fft.fftfreq(N, d=dt)
     t = np.arange(0,T,step=dt)
@@ -37,6 +37,8 @@ if __name__ == "__main__":
     print ("Time spent MESA: {0} s".format(elapsed))
     start = time.perf_counter()
     PSD    = M.spectrum(dt,f)
+    print(PSD)
+    
     elapsed = time.perf_counter()
     elapsed = elapsed - start
     print ("Time spent PSD: {0} s".format(elapsed))
@@ -44,8 +46,20 @@ if __name__ == "__main__":
     fig = plt.figure(1)
     ax  = fig.add_subplot(111)
     ax.loglog(f[:N//2], M.spectrum(dt,f)[:N//2],'-k')
-    ax.set_xlim(1,srate/2.)
+#    ax.set_xlim(1,srate/2.)
     ax.set_xlabel("frequency (Hz)")
     ax.set_ylabel("PSD (Hz$^{-1}$)")
+    
+    
+    t = np.arange(T-1,T+199,step=dt)
+    prediction = M.forecast(200, 1000)
+    l, m, h = np.percentile(prediction,[5,50,95],axis=0)
+    fig = plt.figure(2)
+    ax  = fig.add_subplot(111)
+    ax.plot(np.arange(0,T,step=dt),data,linewidth=1.5,color='r',zorder=2)
+    ax.plot(t,prediction.T,linewidth=0.5, color='k', zorder = 0)
+    ax.plot(t,m,linewidth=1.5, color='k', zorder=3)
+    ax.fill_between(t,l,h,facecolor='turquoise',alpha=0.5, zorder = 1)
[…]

### 2021-02-04 examples/check_forecast.py
Fix load function, examples and package

@@ -1,42 +1,0 @@
-try:
-	import sys
-	sys.path.insert(0,'..')
-	import memspectrum
-except:
-	import memspectrum
-
-import matplotlib.pyplot as plt
-import numpy as np
-
-import mlgw.GW_generator as gen
-t = np.linspace(-10.,0.1,4000)
-g = gen.GW_generator()
-h_p, h_c = g.get_WF([20,10,0.1,-0.5],t)
-
-data = h_p[-3000:-2000]*1e20
-#data = np.sin(10.*t)
-
-#data = np.random.normal(0,1,10000)*np.random.poisson(10,10000) #this is a sanity check that it works
-
-plt.figure()
-plt.plot(data)
-#plt.show()
-
-m = memspectrum.MESA(data)
-
-m.solve(method = "Fast", optimisation_method = "FPE")
-
-np.random.seed(0)
-forecast = m.forecast(1500,1, include_data = True)
-
-print(len(m.a_k))
-plt.figure()
-plt.plot(np.squeeze(forecast[0,:]), label = 'vec')
-plt.legend()
-
-
-plt.figure()
-plt.plot(*m.spectrum(np.diff(t)[0]))
-
-
-plt.show()

### 2021-02-04 examples/sound_MESA.py

@@ -15,7 +15,7 @@ except:
     from memspectrum import *
     import memspectrum.GenerateTimeSeries as GenerateTimeSeries
 
-t = 4. #seconds of data
+t = 14. #seconds of data
 
     #loading data and preparing input to MESA
 rate, data = scipy.io.wavfile.read("data/waterfall_noise.wav") #loading waterfall noise file
@@ -25,8 +25,8 @@ dt = 1./rate
 times_out = np.linspace(0., len(data_MESA)*dt, len(data_MESA))
 
     #computing PSD with MESA
-M = MESA(data_MESA)
-P, ak, opt= M.solve(method = "Fast", optimisation_method = "FPE", m = int(2*len(data_MESA)/(2*np.log(len(data_MESA)))))
+M = MESA()
+P, ak, opt= M.solve(data_MESA, method = "Fast", optimisation_method = "FPE", m = int(2*len(data_MESA)/(2*np.log(len(data_MESA)))))
 
     #evaluating the spectrum
 N_points = 100000
@@ -44,10 +44,10 @@ scipy.io.wavfile.write("data/simulated_noise.wav", rate, to_save.real.astype(np.
     #simulated noise is loaded and psd is estimated again
 rate, data = scipy.io.wavfile.read("data/simulated_noise.wav") #loading waterfall noise file
 data_MESA = data[:int(t*rate),0].astype(np.float64)
-data_MESA = data_MESA+1j*0.
+data_MESA = data_MESA
 dt = 1./rate
-M = MESA(data_MESA)
-P, ak, opt= M.solve(method = "Fast", optimisation_method = "FPE", m = int(2*len(data_MESA)/(2*np.log(len(data_MESA)))))
+M = MESA()
+P, ak, opt= M.solve(data_MESA, method = "Fast", optimisation_method = "FPE", m = int(2*len(data_MESA)/(2*np.log(len(data_MESA)))))
 PSD_sim =  M.spectrum(dt,f_PSD) #PSD computed on simulated data
 
 print("All done: if you like to listen to the output file, type \"aplay data/simulated_noise.wav\" ")
@@ -73,6 +73,19 @@ plt.plot(f_PSD, PSD_sim.real, label = "simulated")
 plt.yscale('log')
 plt.xlabel("frequency (Hz)")
 plt.legend()
+
[…]

### 2021-02-05 paper/mesa.tex
Selected piece of theory
Changes in paper and bibliography

@@
 \newtheorem{theorem}{Theorem}
+\usepackage{float}
 
@@
 \section{Maximum Entropy Principle} 
-
-Jaynes Maximum Entropy principle (MAXENT) \cite{JaynesArticle}\cite{jaynes2003ptl} represents one of the most important results in Bayesian Statistics: it provides a way to uniquely assign probabilities to a phenomenon in a way that best represent our state of knowledge, while at the same time it is non committal with unavailable information.
-It is formulated as a variational problem in which Shannon's entropy \cite{Shannon}
-\begin{equation}\label{eq:shannonEntropy}
-H[p_1, \dots, p_N] = - \sum_{i = 1}^{N}p_i\log(p_i)
-\end{equation}
-or, in the continuous case
-\begin{equation}
-H[p(x)] = - \int dx p(x)\log p(x)
-\end{equation}
-is the functional to be maximized.
-The constraints on the variational problem are given by the prior knowledge $I$ and are usually represented as fixed values for various momenta of the probability distribution.
-The most interesting result of MAXENT is obtained when a constraint on both mean and variance is considered.
-Consider for simplicity a vector $\vec x(t)$ with components in $\mathbb{R}^N$
-\begin{equation*}
-    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
-\end{equation*}
-We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals.
-If we put constraints in both mean $\vec \mu(t)$ and covariance matrix $C(t_i, t_j)$, MAXENT allow us to assign for $\vec x(t)$ a multivariate normal distribution \cite{gregory_2005}:
-\begin{equation}
-    p\left(\vec x(t)\vert I\right) = 
-    \frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp^{\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right)}. 
-\end{equation}
-With MAXENT, the often unjustified assumptions for the data to be normally distributed finds an elegant explanation.
-It is not only the central limit distribution for finite variance processes, it is also the distribution that maximizes the entropy and, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process.
-In some sense, we can interpret the central limit theorem as the natural 'statistical' evolution toward a configuration that maximizes entropy.
-Thanks to Burg, maximum entropy method has been generalized to compute the most unbiased estimate of the spectrum for a realization of some stochastic process. 
-Since it contains the same information of a power spectral density, high resolution computation of a power spectrum from a single time series $\vec x(t)$ of length $N$ is a fundamental task in signal analysis.
-\\
-Since we are dealing with a finite number of samples $N$, it is not possible to have an accurate computation of the autocorrelation function for every time interval.
-Also, the error $\sigma$ in the estimate increases with the time interval $k$ as $\sigma \sim \frac{1}{\sqrt{N - k}}$, so that only a few values for the autocorrelation function can actually be computed.
-\\
-The problem we want to address is how to compute a PSD from the partial knowledge of the autocorrelation function. \\
-The standard approach of peridiograms \cite{Lomb} \cite{Scargle} consists in taking the Fourier Transform of the sample autocorrelation, written as \begin{equation}
-    \rho = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\}.
-\end{equation}
-The sequence $W$ is a window function that can be chosen in several different way, each method having pros and cons for the final estimate of the power spectral density. \\ 
-The choice of a window function is arbitrary and made by trial and error, until a satisfactory compromise between variance and resolution of the estimate of power spectral density  is reached.
-A high frequency resolution implies high variance and vice-versa.
-Another problem,is that the result depends on this choice, and this fact may raise the question of which we should consider as the "real" power spectral density.
-\\
-Another problem with this approach is the requirement for the window to be $0$ outside the interval: we are arbitrarily assuming $\rho_j = 0$ for $j > M$, and modifying the estimate if a non-rectangular window is chosen.
-Making assumptions on unavailable data and modifying the ones we have at our disposal does not seem a completely fair approach.
-\\
-MAXENT gives us the possibility to develop a completely new way for the estimation of PSDs, without any a priori assumptions on the unavailable data.
-Our requirements (the constraints for the variational problem) are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation.
-\\
-To do so, there is a problem we have to solve: our definition of entropy (equation \ref{eq:entropy}) depends on the probability distribution, not the power spectrum.
-We need to find a way to change the definition's domain for the entropy to formulate the variational problem in terms of the power spectral density $S(f)$ alone.
-A way out of this problem is considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA}.
-One can show that the 'entropy gain' obtained filtering white noise is:
-\begin{equation}\label{eq:EntropyGain}
-    \Delta H = \int_{-\infty}^{\infty}\log S(f) df , 
-\end{equation}
-so that we can reformulate the variational problem requiring $S(f)$ to maximize the entropy gain.
-In real examples we have no access to all frequencies but only on those between minus and plus the Nyquist frequency, defined from the sampling interval $\Delta t$ as
-\begin{equation}
-    Ny = \frac{1}{2 \Delta t}.
-\end{equation}
-All our quantities are well defined in this range only, so that we are not going to maximize \ref{eq:EntropyGain}, but
-\begin{equation}\label{eq:EntropyGain2}
-    \Delta H = \int_{- Ny}^{Ny}\log S(f) df.
-\end{equation}
-Defining $\bar r_n$ as the sample autocorrelation, our constraints will be
-\begin{equation}\label{eq:MaxConstraint}
-    \int_{-Ny}^{Ny} S(f) e^{\imath 2 \pi f n \Delta t} df = \bar r_{n}.
-\end{equation}
-This approach on power spectral density estimate was developed by Burg \cite{burg1975maximum} and it is a way out from the problems raised in periodograms estimate.
-It requires no assumptions on the unavailable estimates for $\bar r_n$ and is consistent with the sample autocorrelation function by construction.
-The hard task is now to show that this problem has a solution.
-Remarkably, it admits an analytical solution in a closed expression for $S(f)$, that we now work out.
-
-\newpage 
-a
-\newpage
-
-\section{Introduction}
 The Maximum Entropy principle (MAXENT) is one of the most important results in Bayesian Statistics.
@@
 After a short introduction to Jayne's MAXENT, we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form.
-Since its natural domain of application are time series, we will also provide a brief explanation of the basic concepts in stochastic processes that are needed to fully understand the method and the cases in which it can be correctly applied.
-\section{Maximum Entropy}
-One of the main problems in statistics is the assignment of a probability for an event.
-There have been several proposals from both 'objective' and 'subjective' approaches to probability and statistic.
-The former assert that concept of probability is intrinsic to the event itself and can be computed as the 'percentage of occurrences of an attribute in a given ensemble':
-speaking of probability only makes sense when an ensemble is considered. 
-For the latter approach instead, probability is a representation of the state of knowledge of the observer about the event itself: the higher the probability, the higher our belief that the event will occur and vice versa.
-It appears to be clear that for the Bayesian, information will play a key role in the a priori assignment of the probability for an event, while for the frequentist prior assignment does not make sense at all, since the whole information is contained in the experiment.
-\\
-So, assigning probability given some prior information is a pure Bayesian problem.
-
-\subsection{Indifference principle}
-The indifference principle (IP) is one of the easiest principle that can be used to assign a probability to some events. 
-Consider a set of independent, mutually exclusive hypotheses: 
-\begin{equation}\nonumber
-    \left\{H_i\right\} \quad i = 1 \dots N
-\end{equation}
-so that they are exhaustive: 
-\begin{equation}\nonumber
-    \sum_{i = 1}^{N} P(H_i \vert I) = 1.
-\end{equation}
-If there is no reason to prefer one hypothesis with respect to another, the IP states that we should assign the same probability for every event: 
-\begin{equation}
-    P(H_i\vert I) = \frac{1}{N}. 
-\end{equation}
-This can be easily generalized to the continuous case. Consider, for simplicity, a one dimensional case, and suppose we have the additional information: $a < x < b$.\\ 
-In the continuous case, the probability $p(x\vert I)$ is meant as the probability for the event to assume values in a given subset $\left[x, x +dx\right]$.
-Requiring events to be mutually exclusive is requiring their interception to be the empty set, while the exhaustiveness condition is the normalization condition:
-\begin{equation}
-    \nonumber
-    \int_a^b p(x\vert I)dx = 1.
-\end{equation}
-In this case, the IP reduces to the computation of the normalization constant: 
-\begin{equation}
-    p(x\vert I) = \frac{1}{b-a}
-\end{equation}
-and allow us to easily introduce uniform priors, the simplest prior distribution for Bayesian computation.
-The only hypothesis is for our events to be mutually exclusive and exhaustive.
-In this case, since there is no evidence for a given hypothesis to be more plausible than another, according to IP we have to assign the same probability for every event.
-\\
-If that is not the case, how can we generalize IP principle for a larger amount of information?\\ 
-Jaynes\cite{JaynesArticle}\cite{jaynes2003ptl} formulated an exhaustive answer as a variational problem and rely to the concept of information entropy: 
-the distribution that reflects knowledge being maximally non committal to unknown information is uniquely determined maximizing Shannon's entropy\cite{Shannon} using information as constraints. \\ 
-This is now known as the 'Maximum Entropy Principle' (MAXENT)
-
-\subsection{Maximum Entropy Principle}
 To introduce MAXENT as a variational problem, we will workout some simple examples to understand how information's entropy and 'uncertainty' are related. \\ 
@@
 \end{equation*}
-for each outcome, i.e. appealing to the indifference principle, is equivalent to the requirement for our experiment to be the most unpredictable. 
+for each outcome, i.e. appealing to the indifference principle (IP), is equivalent to the requirement for our experiment to be the most unpredictable. 
 All we know is that one over $N$ events is going to happen. With this evidence, we cannot expect one outcome to occur over the others. 
@@
 From previous discussion, given a fixed number $N$ of possible outcomes, we want $H$ to be maximized by the choice $p_1 = \dots = p_N$. \\ 
-Consider now the opposite situation, in which probability is assigned appealing to IP and we have a different number of possible outcomes: 
-\begin{itemize}
-\centering
-    \item $P_1 = P_2 = \frac{1}{2}$
-    \item $P_1 = P_2 = P_2 = \frac{1}{3}$
-    \item $P_1 = \dots = P_4 = \frac{1}{4}$
-\end{itemize}
-Intuitively, the most unpredictable experiment is the one with larger number of outcomes: probability of getting it wrong increases as $\frac{N -1}{N}$.
+Consider now the opposite situation, in which probability is assigned appealing to IP and we have a different number of possible outcomes.
+Intuitively, the most unpredictable experiment is the one with larger number of outcomes: probability of "getting it wrong" increases as $\frac{N -1}{N}$.
 So when equal probabilities are assigned, we want
@@
 
-\section{Time Series}
-Before we introduce MAXENT applications to spectral analysis, we provide a short review on time series and some concepts that will be used in following chapters. \\ 
-Consider a stochastic process, i.e.
-any process with some non deterministic component that takes values in some space $\bar X = \left\{x_1 \right\}$.
-A time series X of the process, i.e.
-a realization, is a set of subsequent values assumed by the variable at different times, i.e.
-\begin{equation}
-    X = \left\{x_1, x_2, \dots, x_N \right\}; \qquad \text{ with } x_i = x(t_i)
-\end{equation}
-For our goal is important to understand how a single time series can be used to infer properties on the stochastic process itself.
-A full characterization for the stochastic process can only be provided if all the joint probability distributions for the process are known
-\begin{equation}\nonumber 
-    p(x_1, t_1 | I); \quad \dots \quad ;  p(x_1, t_1; \dots x_n, t_n | I);  \quad \dots  
-\end{equation}
-or, equivalently, all momenta for the distribution are provided: 
-\begin{align}
-    \nonumber &\mu(t_1) = \int  x_1p(x_1, t_1 \vert I)dx_1  \\
-    \nonumber 
-    & R(t_1, t_2) = \int  x_1x_2p(x_1, t_1;x_2, t_2\vert I) dx_1dx_2\\
-    \nonumber 
-    \vdots \\ \nonumber
-    &\mu_n(t_1, \dots, t_n) = \int  x_1\dots x_np(x_1; \dots; x_n \vert I) dx_1dx_2\dots dx_n\\ 
-    \nonumber
-    &\vdots
-\end{align}
-Since both descriptions require $N \to \infty$, they are impossible to achieve and very often an estimate of just the first two moments is given, hence only a partial characterization can be made. \\ 
-The first momentum is known as the mean for the process, while second moment is known as 'Autocorrelation', and gives information on the deterministic dependence of a process from its previous values.
-It is a central quantity in time series analysis and is related to the autocovariance function via:
-\begin{equation}\nonumber 
-    C(t_1, t_2) = R(t_1, t_2) - \mu(t_1)\mu(t_2).
-\end{equation} \\
-Study of time series analysis is often a hard task, but there are some special cases that simplify the analysis: stationary and autoregressive processes.  
-\subsection{Stationary processes}
-\subsubsection{Stationarity conditions}
-There are several definitions of stationarity for a process \cite{Mantegna2007Introduction}. A stochastic process is said to be 'strictly stationary' if the joint distribution for the process: 
-\begin{equation}
-p(x_1,t_1; \dots; x_n,t_n \vert I) = p(x_1,t_1 + \Delta t; \dots; x_n, t_n +\Delta t\vert I)
-\end{equation}
-is constant in time whatever the order $n$ and time lag $\Delta t$.
-This implies that its moments are also constant in time.
-In this case, the time series consists in independent identically distributed terms.
-Such a process is rarely encountered and more general definitions are needed for real case studies.
-\\
-On of the most important is the 'wide sense stationary process', whose constraints are: 
-\begin{equation}
-    \mu_1(t) = \mu; \quad R(t_1, t_2) = R(t_2 - t_1) = R(\tau).
-\end{equation}
-Such a process has constant mean and autocorrelation that only depends on time lag between the considered values. \\
-The mean does not bring high information about the process, and can always be reabsorbed in the process defining 
-\begin{equation}
-    \nonumber
-    y(t) = x(t) - \mu.
-\end{equation}
-y(t) is a stationary process with 0 mean but same statistical properties of $x(t)$.
-$R(\tau)$ brings information on the amount of dependence of the time series on its previous values: if some process repeats similarly after $\bar t$ seconds, autocorrelation will have a secondary maximum at $\tau = \bar t$.
-Also, if the function is periodic of period $T$, the autocorrelation will have same periodicity.
-In general, autocorrelation represents most of the information we can obtain from the study of a stationary stochastic process, and it is rather fundamental.
-Its computation is even more important in the case of a Gaussian process.
-For such a process it is the only moment required to fully characterize it.
-One of the most important results regarding the autocorrelation function, is the Wiener-Khinchin theorem.
-It ensures that, for a wide sense stationary process, the frequency domain counterpart of the autocorrelation function is the power spectral density:
-\begin{equation}
-    R(\tau) = \int  S(\omega) e^{-\imath \omega \tau} d\omega,
-\end{equation}
-Autocorrelation and power spectral density bring the same information about the process.
-It is fundamental to develop methods to accurately estimate the power spectral density in such a way that it is consistent with the autocorrelation function.
-\subsubsection{Gaussian noise}
-MAXENT is a powerful tool to study noise.
-If one considers a wide sense stationary process  $\vec n(t) = \left(n_1, n_2, \dots, n_k \right)$ with 0 mean and known autocorrelation matrix, one can assign uniquely a normal distribution:
-\begin{equation}\label{eq:noisePDF}
-    p\left(\vec n(t)\vert I\right) = \frac{1}{\left(2\pi\det R \right)^{\frac{k}{2}}}\exp\left(-\frac{1}{2}\sum_{i, j}n_i R^{-1}_{ij}n_j\right)\,.
-\end{equation}
-In the case of a wide sense stationary process, being $R_{ij} = R(t_j - t_i)$ dependent on time steps only, the autocorrelation matrix is a Toeplitz Matrix (see Appendix \ref{app:Toeplitz}) and can be diagonalized, in the limit of large number of samples, by a Discrete Fourier Transform:
-\begin{equation}
-    S = F R F^{-1}. 
-\end{equation}
-For the Wiener-Khinchin theorem, $S$ is the power spectral density for the noise itself. Defining
-\begin{equation}
-    \tilde n(\omega) = F\left[n(t) \right],
-\end{equation}
-with $F$ the Fourier transform, the probability distribution (\ref{eq:noisePDF}) for noise in frequency domain (FD) becomes  
-\begin{equation}
-    p(\tilde n(f) \vert I) = \frac{1}{\left(2 \pi \det S \right)} \exp \left(-\frac{1}{2}\sum_{i} \frac{\tilde n_i \tilde n_i}{S_i} \right).
-\end{equation}
-Power spectral density plays a central role in noise study: it is the covariance matrix for noise in frequency domain. Also, in large N domain, frequency noise is serially uncorrelated. \\
-There is a very special case of Gaussian Noise that we have to mention: white noise. White noise can be represented as Gaussian noise with autocorrelation function
-\begin{equation}
-    R_{ij} = \sigma^2 \delta_{ij},
-\end{equation}
- it is a strictly stationary process whose autocorrelation function is 0 for any time lag different from 0, and its realization is made of independent identically distributed samples.
- Going in the Fourier Domain
-\begin{equation}
-S_{ij} = \sigma^2 F_{ik}\delta_{kl}F^{-1}_{lj} = \sigma^2 \delta_{ij} = R_{ij}, 
-\end{equation}
-we show that power spectral density for white noise is constant over frequency.
-
-\subsection{Autoregressive process}\label{sec:ARprocess}
-The description of a process is simplified if we are dealing with some autoregressive process of order p, or simply an AR(p) process.
-For an AR(p) process, each value depends linearly from its previous values
-\begin{equation}
-    x_i = \sum_{i = 1}^{p}a_{i}x_{i - p} + \nu_i, 
-\end{equation}
-with $a_i$ some constant coefficients and $\nu_i$ white noise. \\
-An AR(1) process is known as a Markov process.
-Complete knowledge on AR(p) can be achieved with the computation of $p + 1$ momenta (or conditional distributions) only.
-We prove it for a Markov process.\\
-Consider the $n'th$ order joint probability distribution
-\begin{equation}\label{eq:NthOrderPDF}
-    p(x_1, t_1; x_2, t_2; \dots; x_n, t_n | I),
-\end{equation}
-from the definition of conditional probability
-\begin{equation}\nonumber 
-p(x,y \vert I) = p(x \vert y I) p(y \vert I)
-\end{equation}
-one can rewrite \ref{eq:NthOrderPDF} as 
-\begin{equation} \nonumber 
-    p(x_n, t_n; \dots; x_1, t_1 \vert I) = p(x_n, t_n \vert x_{n-1}, t_{n-1}; \dots x_1, t_1 I) p(x_{n-1}, t_{n-1} \dots x_1, t_1 \vert I). 
-\end{equation}
-Since under Markov assumption, conditional $pdf$ only depends on the previous value for the variable 
-\begin{equation} \nonumber 
-    p(x_n, t_n \vert x_{n-1}, t_{n-1}; \dots, x_{1}, t_1 I) = p(x_n, t_n \vert x_{n-1}. t_{n-1} I) 
-\end{equation}
-reiterating the decomposition of the joint probability, we find that
-\begin{equation}
-\left(\ref{eq:NthOrderPDF} \right)=  p(x_n, t_n \vert x_{n-1}, t_{n-1}, I )p(x_{n-1}, t_{n-1}\vert x_{n-2}, t_{n-2}, I) \dots p(x_1, t_1\vert  I),
-\end{equation}
-so that for every order $n$, the probability distribution can be reduced to the product of $n$ second order probability distributions. \\
-The knowledge of the distribution for the variable and of the second order conditional distribution fully characterize a Markov process.\\
-The same reasoning can be repeated for any order $p$.
-For an AR(p) process the $n$-th order joint probability distribution ($n > p$) can be rewritten as a product of the $p$-th order conditional distributions, so that they completely characterize the process.
-
-\subsection{ARMA process and Autoregressive stationary process}
-There are two fundamental results for an autoregressive process that are strictly related with Burg's Algorithm (see section \ref{sec:MESA}) for spectral analysis, and they are all about autoregressive stationary processes and their generalization, ARMA(p,q) processes.
-\subsubsection{ARMA process}
-We will only need to define ARMA(p,q) and their relation with AR(p) processes without stepping in many details. These concepts are needed to give a stronger background to our future results.  \\
-The generalization of the Autoregressive Process, the $ARMA(p,q)$ process, is defined by the following relation: 
-\begin{equation}
-    x_t = \sum_{i = 1}^p \alpha_i x_{t - i} + \sum_{i = 1}^q \beta_i \nu_{t - i} + \nu_t,
-\end{equation}
-again, $\nu$ is white noise and both $\alpha$ and $\beta$ are constants.
-It is evident that an autoregressive process is a special case of the more general $ARMA(p,q)$ process, since  $Ar(p) = ARMA(p,0)$.
-We will not enter in the details of the description of $ARMA(p,q)$ processes, but they are important for two main reasons.
-\\
-First, $ARMA$ is a parsimonious way to characterize an $AR$ process.
-It is, in fact, possible to prove that an $ARMA(p,q)$ process, with both $p, q$ finite, is equivalent to some $AR(\infty)$ process \cite{Haykin}.
-\\
-The second and even more important result about these processes is Wold's Theorem\cite{Wold}: 
-\begin{theorem}
-Every real valued stationary stochastic process $P$ can be written as an $ARMA(p,q)$ process. 
-\end{theorem}
-This theorem makes clear the central role played by the ARMA representation: it allow us to fully characterize any stochastic process if its coefficients are known.
-This is usually an hard task, but, since we know the $AR / ARMA$ equivalence, we can find an approximate description of the process in terms of an $AR$ process, all is needed is a way to estimate the autoregressive coefficients.
-\\
-\subsubsection{Autoregressive stationary process}
-The last important ingredient is the condition of stationarity for AR processes. The conditions of stationarity for a stochastic process are well known. Defining the 'lag' operator $B$ such that 
-\begin{equation}
-    \nonumber 
-    B x_{i} = x_{i - 1},
-\end{equation}
-one can rewrite the autoregressive process as 
-\begin{equation} \nonumber 
-    x_i - \sum_{k = 1}^p a_k B^k x_i = \nu_i, 
-\end{equation}
-or, defining $b_0 = 1$ and $b_k = - a_k$
-\begin{equation}
-    \left(\sum_{k = 0}^p b_k B^{k}\right) x_i = \nu_i.
-\end{equation}
-The polynomial on the left side is known as the 'characteristic polynomial' of the autoregressive process $\phi (B) = \sum_{k = 0}^p b_k B^k$. 
-The condition of stationarity for autoregressive processes is
-\begin{theorem}
-An autoregressive process of order $p$ is stationary if all the roots $\beta_i$ of its characteristic polynomials lay outside the unit circle, i.e. $\vert\beta_i\vert > 1$.
-\end{theorem}
-Consider for simplicity a Markov process that only takes real values
-\begin{equation}\nonumber 
-    x_t - a_1 x_{t- 1} = \nu_t; \quad \text{ with } a_i \in \mathbb{R}
-\end{equation}
-whose characteristic polynomial is
-\begin{equation}
-    \nonumber 
-    \phi(B) = 1 - a_1 B .
-\end{equation}
-Let $\beta$ be the root of the polynomial $\phi(\beta) = 0$. 
-If $\vert \beta \vert > 1$ ($\vert \beta \vert < 1$) $a_1$ has to be $\vert a_1 \vert < 1$ ($\vert a_1 \vert > 1$).
-It is easy to see that if condition of stationarity is not met, the process explodes and $x_t$ will quickly diverge.
-\\
-These few concepts are fundamental in stochastic processes and necessary for what follows.
-\section{Maximum Entropy Spectral Analysis}\label{sec:MESA}
-From our previous discussion on time series, the importance of the autocorrelation function should be clear.
-Since it contains the same information of a power spectral density, high resolution computation of a power spectrum from a single time series $\vec x(t)$ of length $N$ is a fundamental task in signal analysis.
-\\
+\section{Maximum Entropy Spectral Analysis} 
+The computation of the power spectral density is an important component in time series analysis and characterization.
+Also, if the process is at least wide sense stationary, PSD is related to the autocorrelation function via a Fourier Transform.
 Since we are dealing with a finite number of samples $N$, it is not possible to have an accurate computation of the autocorrelation function for every time interval.
@@
 A high frequency resolution implies high variance and vice-versa.
-Another problem,is that the result depends on this choice, and this fact may raise the question of which we should consider as the "real" power spectral density.
+Another problem is that the result is very unstable with respect to the window's choice, and this fact may raise the question of which we should consider as the "real" power spectral density.
 \\
@@
 The hard task is now to show that this problem has a solution.
-Remarkably, it admits an analytical solution in a closed expression for $S(f)$, that we now work out.
+Remarkably, it admits an analytical solution in a closed expression for $S(f)$
+\begin{equation}\label{eq:MESApsd}
+    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^z\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}, 
+\end{equation}
+with $\Delta t$ the sampling interval $a_0$ = 1 and the coefficients $P_N$ and $a_i (i \neq 0)$ to be determined. The explicit solution can be found in Appendix REF
+
+\section{Autoregressive Process Analogy} 
+The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the analysis
+of autoregressive processes.\cite{doi:10.1029/RG013i001p00183}.
+Consider an autoregressive stationary process of order $p$ (AR(p)): 
+\begin{equation*}
+    x_t - b_1 x_{t-1} - b_2 x_{t-2} \dots b_p x_{t - p} = \nu_t, 
+\end{equation*}
+taking the z transform of the previous: 
+\begin{align}
+    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t
+\end{align}
+and calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
+in the $z$ domain, the process takes the form:
+\begin{equation}
+    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
+\end{equation}
+Since we assumed stationarity for the process, $\tilde{x}(z)$ is analytic both on and inside the unit circle.
+Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
+\begin{equation}\label{eq:ARspectrum}
+    S(f) = \vert \tilde x(z)\vert ^ 2 = 
+    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}.
+\end{equation}
+The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$. \\ 
+Comparing (\ref{eq:ARspectrum}) with (\ref{eq:MESApsd}), identifying $b_1 = - a_i$ and $P_N \Delta t= \sigma ^ 2$, the expression for the power spectral density obtained with MESA is formally equivalent to the power spectral density of some stationary autoregressive process.
+\\
+Having computed the prediction error filter one can easily obtain an estimate of the $AR$ coefficients for the process just changing sign to the prediction error filter $a_i$.
+Taken a time series, MESA allows us to compute the power spectrum and to find an effective description in terms of an autoregressive process. 
+Wold's theorem ensures that this (approximate) description for the phenomenon is legitimate when the process is found to be stationary.
+So, MESA defines a bridge between spectral estimation and the study of autoregressive processes, an unexpected and fundamental result.
+\\
+Another very remarkable process for this estimate, is that the linear predictor obtained solving the error filter equation, is the best linear predictor, i.e.
+it minimizes error, and is equivalent to a least square fit for the coefficients.
+\cite{VDBos}  \\
+To prove that it minimizes error, we will follow the original proof provided by Burg \cite{burg1975maximum}. Consider a general linear predictor $-c_i$: 
+\begin{equation}
+    x_{t} - \sum_{i = 1}^p (-c_i)x_{t - i} = \sum_{i = 0}^p c_i x_{t - i} = e'_t, 
+\end{equation}
+Denoting with $E\left[\cdot\right]$ the expectation value of some variable, the mean square error is given by: 
+\begin{equation}
+E\left[\sum_{i = 0}^p c_i x_{t-i}\sum_{j = 0}^p c^*_j x^*_{t - j}\right] = \sum_{i,j} c^*_j E\left[x^*_{t-j}x_{t-i} \right] c_j = \sum_{i, j}c^*_j R(j - i) c_i > 0,
+\end{equation}
+which, being $R$ positive definite and defined by the multiplication of a complex number times its complex conjugate, is always positive definite. 
+Let $-\vec{b}$ be the prediction error filter obtained with MESA. Multiplying the left hand side of \ref{eq:predictionErrorFilter} for the complex conjugate of the forward prediction error, 
+\begin{equation}\nonumber 
+    \sum_{i,j}b^*_jR(j-i)b_i = P_N
+\end{equation}
+is obtained . \\ 
+The (positive) quantity  
+\begin{equation}\label{eq:linearPredictor}
+    \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i),
+\end{equation}
+using Einstein notation on repeated indices and recall that $c_0 = b_0 = 1$, can be rewritten as 
+\begin{align}
+    & c^*_j R(j - i) c_i - c^*_j R(j - i) b_i - b^*_j R(j - i) c_i + b^*_j R(j - i)  b_i =\\ \nonumber 
+    &c^*_j R(j - i) c_i - c^*_0 P_N - P_N c_0 + P_N =\\ \nonumber
+    & c^*_j R(j - i) c_i - P_N.
+\end{align}
+Moving $P_N$ on the left hand side, it becomes 
+\begin{equation}\label{eq:bestLinearPredictor}
+    \sum_{i, j}c^*_j R(j-i)c_i = \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i) + P_N.
+\end{equation}
+Being all the quantities positive definite, the left hand side is minimized by the choice $c_i = b_i$.
+Also, known $P_N$, the previous equation gives an estimate on the additional mean square error associated at choices that differ from the prediction error filter.
+
+\section{Forecasting}
+
+\section{Choosing an autoregressive order}\label{sec:optimizers} 
+Before being able to apply MESA, we need to address the question of the best length $M$ of the filter that provides the optimal estimate of the power spectrum.
+If the autocorrelation matrix was perfectly know, the answer would be simply $M = N$.
+\\
+In real cases, such a knowledge is not available and one has to estimate the autocorrelation function from the data.
+The accuracy for the estimate decreases as the time lag $k$ increases as $\sim \left(\sqrt{N - k}\right)^{-1}$.
+\\
+The accuracy of the estimate should then increase as the length of the autocorrelation function increases, but, from a certain point on, inaccuracy in the estimate for $R$ becomes so large that the estimate for the spectrum is negatively affected by this fact.
+A compromise between the two behaviour is proposed to be \cite{doi:10.1190/1.1440902}
+\begin{equation}\label{eq:MMAx}
+M_{max} = 2N / \ln{(2N)}
+\end{equation}
+Still, this is not a satisfactory answer since it is an upper limit for the estimate of the length $M$.
+Various approaches to optimize the order $M$ have been proposed in literature; in the following we present them.
+\begin{itemize}
+\item \textbf{Final prediction Error (FPE)} 
+The first criterion is due to Akaike (1970) \cite{Akaike1970StatisticalPI}. He proposed that M has to be chosen as the length that reduces the error when the filter is used as a predictor: 
+\begin{equation}
+    FPE: \min_{M}E\left((x_t - \hat x_t) ^ 2\right)
+\end{equation}
+with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.  This criterion is equivalent to the choice of $M$ such that 
+\begin{equation}
+    \min_{M}P_{m} \frac{N + m + 1}{N - m - 1}
+\end{equation}
+with $P_M$ defined by the prediction error filter equation. 
+\item \textbf{Criterion Autoregressive Transfer function (CAT)}
+This second criterion is due to Parzen \cite{bhansali1986}, and requires $M$ to be chosen as the length $m$ the minimizes the quantity 
+\begin{equation}
+    \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
+\end{equation}
+It has the interesting property that the distance between the original filter and the estimate is minimum and asymptotically unbiased.
+
+\item \textbf{Optimum Bayes Decision rule (OBD)} 
+The last criterion we will consider is the OBD (Rao, 1982)\cite{doi:10.1029/WR018i004p01097}.
+With the technique of asymptotic integration, $M$ is proposed to be the length that maximizes the posterior distribution, condition met selecting $m$ that minimize
+   \begin{equation}
+        (N - m - 2) \log(P_m) + m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2.
+    \end{equation}
+\end{itemize}
+Once an optimizer is chosen to select the best recursion order $M$, we then proceed as follows:  we solve the Levinson recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached.
+Then, the order $M$ is selected to be the one that minimize the chosen optimizer value.
+\\
+PARLARE QUI DELL'EARLY STOP? 
+The following chapter will be devoted to the study of the statistical properties of the previous optimizers to understand which choice provides the best quality in the reproduction of some known power spectral densities.
+
+\section{Analysis of the optimizers}
+
+\section{Example}
+
+\section{Comparison with Welch method}
+
+\section{Appendix 1: MESA solution}
 \subsection{MESA solution}
@@
     \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{-\imath 2 \pi f n \Delta t} 
-    \right).
+    \right) df.
 \end{equation}
@@
 \begin{equation}
-      \frac{\delta H}{\delta \bar r_s} = \frac{1}{2Ny}\int_{-Ny}^{Ny}S(f)^{-1}e^{-\imath 2 \pi f s \Delta t} = \lambda_s = 0 \text{ for } \vert s \vert > N. 
+      \frac{\delta H}{\delta \bar r_s} = \frac{1}{2Ny}\int_{-Ny}^{Ny} S(f)^{-1}e^{-\imath 2 \pi f s \Delta t } df = \lambda_s = 0 \text{ for } \vert s \vert > N. 
 \end{equation}
@@
 \begin{equation}\label{eq:MESApsd}
-    S(f) = \frac{P_N \Delta_t}{\left(\sum_{s=0}^N a_s z^z\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}. 
+    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^z\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}. 
 \end{equation}
@@
 A python implementation of Standard Burg's Method can be found in Appendix \ref{app:Burg}
+
+\newpage 
+a
+\newpage
+
+\section{Introduction}
+The Maximum Entropy principle (MAXENT) is one of the most important results in Bayesian Statistics.
+It provides a way to uniquely assign probabilities to a phenomenon in a way that best represent our state of knowledge, while at the same time it is non committal with unavailable information.
+Its domain of application turned out to be wider than expected.
+In fact, thanks to Burg \cite{burg1975maximum}, this method has also been applied to perform high quality computation of power spectral densities of time series.
+After a short introduction to Jayne's MAXENT, we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form.
+Since its natural domain of application are time series, we will also provide a brief explanation of the basic concepts in stochastic processes that are needed to fully understand the method and the cases in which it can be correctly applied.
+\section{Maximum Entropy}
+One of the main problems in statistics is the assignment of a probability for an event.
+There have been several proposals from both 'objective' and 'subjective' approaches to probability and statistic.
+The former assert that concept of probability is intrinsic to the event itself and can be computed as the 'percentage of occurrences of an attribute in a given ensemble':
+speaking of probability only makes sense when an ensemble is considered. 
+For the latter approach instead, probability is a representation of the state of knowledge of the observer about the event itself: the higher the probability, the higher our belief that the event will occur and vice versa.
+It appears to be clear that for the Bayesian, information will play a key role in the a priori assignment of the probability for an event, while for the frequentist prior assignment does not make sense at all, since the whole information is contained in the experiment.
+\\
+So, assigning probability given some prior information is a pure Bayesian problem.
+
+\subsection{Indifference principle}
+The indifference principle (IP) is one of the easiest principle that can be used to assign a probability to some events. 
+Consider a set of independent, mutually exclusive hypotheses: 
+\begin{equation}\nonumber
+    \left\{H_i\right\} \quad i = 1 \dots N
+\end{equation}
+so that they are exhaustive: 
+\begin{equation}\nonumber
+    \sum_{i = 1}^{N} P(H_i \vert I) = 1.
+\end{equation}
+If there is no reason to prefer one hypothesis with respect to another, the IP states that we should assign the same probability for every event: 
+\begin{equation}
+    P(H_i\vert I) = \frac{1}{N}. 
+\end{equation}
+This can be easily generalized to the continuous case. Consider, for simplicity, a one dimensional case, and suppose we have the additional information: $a < x < b$.\\ 
+In the continuous case, the probability $p(x\vert I)$ is meant as the probability for the event to assume values in a given subset $\left[x, x +dx\right]$.
+Requiring events to be mutually exclusive is requiring their interception to be the empty set, while the exhaustiveness condition is the normalization condition:
+\begin{equation}
+    \nonumber
+    \int_a^b p(x\vert I)dx = 1.
+\end{equation}
+In this case, the IP reduces to the computation of the normalization constant: 
+\begin{equation}
+    p(x\vert I) = \frac{1}{b-a}
+\end{equation}
+and allow us to easily introduce uniform priors, the simplest prior distribution for Bayesian computation.
+The only hypothesis is for our events to be mutually exclusive and exhaustive.
+In this case, since there is no evidence for a given hypothesis to be more plausible than another, according to IP we have to assign the same probability for every event.
+\\
+If that is not the case, how can we generalize IP principle for a larger amount of information?\\ 
+Jaynes\cite{JaynesArticle}\cite{jaynes2003ptl} formulated an exhaustive answer as a variational problem and rely to the concept of information entropy: 
+the distribution that reflects knowledge being maximally non committal to unknown information is uniquely determined maximizing Shannon's entropy\cite{Shannon} using information as constraints. \\ 
+This is now known as the 'Maximum Entropy Principle' (MAXENT)
+
+\subsection{Maximum Entropy Principle}
+To introduce MAXENT as a variational problem, we will workout some simple examples to understand how information's entropy and 'uncertainty' are related. \\ 
+The word 'information' will not be used to indicate our knowledge on the system under study: this is what we will call 'evidence'.
+Information will be intended as in communication theory: it specifies the quantity of details needed to provide a full description of the system under study.
+\\
+If we consider a perfectly sinusoidal signal, knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information entropy content . \\ 
+Even less information is required if we are studying a system whose outcome is certain (has probability $p = 1$).
+If the quantity of information can be considered to be $I = 0$, a communication is not even needed.  
+Shannon \cite{Shannon} proposed the quantity
+\begin{equation}\label{eq:information}
+    I_i = \log_2 \frac{1}{p(x_i)}
+\end{equation}
+to represent the quantity of information brought by an outcome $x_i$ with probability $p_i$.
+It has the nice property of being an additive quantity and it is monotonically decreasing as a function of $p \in [0, 1]$.
+The more uncertain the outcome, the higher the information it brings.
+\\
+What happens if we consider not one outcome, but a larger set of different events?
+What is  the 'most uncertain' outcome?
+Since we are intending probability as a representation of the state of knowledge, one of the first requirement for our assignment is the agreement with common sense.
+\\
+Consider an experiment based on the occurrence of one of two different outcomes $E_1, E_2$ with given probabilities $P_1$ and $P_2$. What are the values that make our outcome the most uncertain? \\ 
+If $P_1$ and $P_2$ are largely different, like $P_1 = 0.999$ and $P_2 = 0.001$, we have enough evidence to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome.
+\\
+The most unpredictable experiment is the one in which 
+\begin{equation}\nonumber
+    P_1 = P_2 = \frac{1}{2}:
+\end{equation}
+this describes a situation of 'maximum ignorance'; we have no evidence at all and we have no instruments to predict what the outcome of such an experiment will be.  \\ 
+This can be generalized in the case of $N$ events. Assigning same probability
+\begin{equation*}
+    P_1 = \dots = P_N = \frac{1}{N}
+\end{equation*}
+for each outcome, i.e. appealing to the indifference principle, is equivalent to the requirement for our experiment to be the most unpredictable. 
+All we know is that one over $N$ events is going to happen. With this evidence, we cannot expect one outcome to occur over the others. 
+To report the results of such an experiment, we need to communicate every single outcome: it 'carries' high information. 
+This description also makes clearer the role played by evidence: it has to be used in probability assignment in order to make the result 'more certain' or 'less informative'.
+It is clear from the point of view of communication theory: the higher the evidence on a system, the lower the information I need to send to describe it.
+\\
+It is now clear that any assignment different from the 'most uncertain' is violating the hypothesis for it to being non committal to non-available evidence.
+We can only assign a more certain probability, i.e.
+less informative, if we have evidence for it.
+If we have none, the result has to be the most unpredictable, i.e.
+the most informative.
+Assigning probabilities requires maximization of uncertainty using evidence as constraint.
+This looks like a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized.
+$H$ has to be a 'generalization' of information defined in equation (\ref{eq:information}) to be something like the 'expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
+\\
+Some properties for $H$ can be worked out easily. 
+From previous discussion, given a fixed number $N$ of possible outcomes, we want $H$ to be maximized by the choice $p_1 = \dots = p_N$. \\ 
+Consider now the opposite situation, in which probability is assigned appealing to IP and we have a different number of possible outcomes: 
+\begin{itemize}
+\centering
+    \item $P_1 = P_2 = \frac{1}{2}$
+    \item $P_1 = P_2 = P_2 = \frac{1}{3}$
+    \item $P_1 = \dots = P_4 = \frac{1}{4}$
+\end{itemize}
+Intuitively, the most unpredictable experiment is the one with larger number of outcomes: probability of getting it wrong increases as $\frac{N -1}{N}$.
+So when equal probabilities are assigned, we want
+\begin{equation}
+    \nonumber
+    H\left[\frac{1}{N}, \dots, \frac{1}{N}\right]
+\end{equation}
+to increase monotonically with N. \\ 
+We also want our function to be continuous in its parameters: any small change in probability has to slightly affect uncertainty. \\
+Last is composition property. Consider three events with probability $p_1, p_2, p_3$ such that $p_2 + p_3 = 1 - p_1 = q$. We require 
+\begin{equation}
+    H\left(p_1, p_2, p_3\right) = H(p_1, q) + qH\left(\frac{p_2}{q}, \frac{p_3}{q}\right),
+\end{equation}
+so that uncertainty is additive. \\ 
+Shannon \cite{Shannon} showed that the only functional form satisfying our assumptions is given by:
+\begin{equation}\label{eq:entropy}
+    H[p_1, \dots, p_N] = - \sum_{i = 1}^N p_i\log{p_i},
+\end{equation}
+or, in the continuous case:
+\begin{equation}
+    H[p(x)] = - \int p(x)\ln p(x) dx,
+\end{equation}
+and $H$ is what we will call entropy.
+This result is very interesting, if we compare it with Eq.~(\ref{eq:information}), we see that the entropy is defined as the 'average information'.
+The probability assignments made maximizing it with respect to $p$, is somehow requiring  the 'expected value' of uncertainty to be a maximum.
+Such an assignment is in agreement with our hypothesis of being 'non committal' with unavailable evidence, since \ref{eq:entropy} is all we need to assign probabilities uniquely.
+Evidence can be added in form of constraints in the maximization.
+\\ It is easy to verify that this expression for entropy verifies the aforementioned properties.
+Assigning the same value for each probability $p_i$, it is straightforward to verify $H$ increases with N: 
+\begin{equation}
+    -\sum_{i = 1}^N \frac{1}{N}\log{\frac{1}{N}} = - \log\frac{1}{N} = \log N.
+\end{equation}
+We want to show that, in case in which our only constraint is exhaustiveness \begin{equation}\nonumber 
+    \sum_{i = 1}^{N}p_i = 1,
+\end{equation}
+MAXENT reduces to indifference principle as it should be. The variational problem is defined as
+\begin{equation}\label{eq:entropyNormConst}
+    \delta\left[H - \lambda\sum_{i = 1}^N (p_i - 1)\right] = 0 
+\end{equation}
+with
+\begin{equation}
+    \delta H\left[p_1, \dots, p_N\right] = \sum_{i = 1}^N\frac{\partial H}{\partial p_i}\delta p_i = - \sum_{i = 1}^N\left(\log p_i + 1\right),
+\end{equation}
+so that equation \ref{eq:entropyNormConst} reduce to
+\begin{equation}\nonumber 
+    \sum_{i = 1}^N\left[\log p_i + \lambda + 1 \right]\delta p_i = 0\,.
+\end{equation}
+The solution of the previous equation is easily worked out requiring every term to be zero: 
+\begin{equation}\nonumber 
+    p_i = e^{-\lambda - 1}.
+\end{equation}
+Applying the normalization condition, it is immediate to show that
+\begin{equation}
+    p_i = \frac{1}{N}.
+\end{equation}
+MAXENT reduces to the IP when only the normalization constraint is used. What happens if more than one constraint is used? \\
+For simplicity, we will work through the next examples in the continuous case. Suppose the mean of the distribution is known $\int xP(x) dx = \bar{x}$. The variational problem can be rewritten as: 
+\begin{equation}
+    \int \left[\log p(x) + 1 + \lambda + \mu x \right]\delta p(x) = 0
+\end{equation}
+and its solution in term of $p_i$ is: 
+\begin{equation}
+    p(x) = e^{-1 - \lambda - \mu x} 
+\end{equation}
+and constraints: 
+\begin{equation} \nonumber
+    \int p(x) dx  = 1 \text { and }
+    \int x p(x) dx = \bar x.
+\end{equation}
+For simplicity, we consider $x$ to be a variable ranging in the interval $(-\infty,+\infty)$, but this can be easily computed for every other interval. The normalization conditions implies 
+\begin{equation}\nonumber 
+   1 =  \int p(x \vert I) dx = e^{-1-\lambda}\int e^{-\mu x} dx, 
+\end{equation}
+so that: 
+\begin{equation}\nonumber 
+    e^{-1 -\lambda} = \mu.
+\end{equation}
+In the continuous case MAXENT assigns for $x$ an exponential distribution
+\begin{equation}
+    p(x \vert I) = \mu e^{-\mu x}. \nonumber 
+\end{equation}
+It is clear from the functional form that the Lagrange multiplier has to be $\mu = \frac{1}{\bar x}$ to satisfy the second constraint. \\ 
+The most interesting and useful result of MAXENT is obtained when a constraint on the variance is also considered. Take a vector  $\vec x(t)$ with components in $\mathbb{R}^N$
+\begin{equation*}
+    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
+\end{equation*}
+We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals.
+If we put constraints in both mean $\vec \mu(t)$ and covariance matrix $C_(t_i, t_j)$, MAXENT allow us to assign for $\vec x(t)$ a multivariate normal distribution \cite{gregory_2005}:
+\begin{equation}
+    p\left(\vec x(t)\vert I\right) = 
+    \frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right). 
+\end{equation}
+With MAXENT, the often unjustified assumptions for the data to be normally distributed finds an elegant explanation.
+It is not only the central limit distribution for finite variance processes, it is also the distribution that maximizes the entropy and, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process.
+In some sense, we can interpret the central limit theorem as the natural 'statistical' evolution toward a configuration that maximizes entropy.
+
+\section{Time Series}
+Before we introduce MAXENT applications to spectral analysis, we provide a short review on time series and some concepts that will be used in following chapters. \\ 
+Consider a stochastic process, i.e.
+any process with some non deterministic component that takes values in some space $\bar X = \left\{x_1 \right\}$.
+A time series X of the process, i.e.
+a realization, is a set of subsequent values assumed by the variable at different times, i.e.
+\begin{equation}
+    X = \left\{x_1, x_2, \dots, x_N \right\}; \qquad \text{ with } x_i = x(t_i)
+\end{equation}
+For our goal is important to understand how a single time series can be used to infer properties on the stochastic process itself.
+A full characterization for the stochastic process can only be provided if all the joint probability distributions for the process are known
+\begin{equation}\nonumber 
+    p(x_1, t_1 | I); \quad \dots \quad ;  p(x_1, t_1; \dots x_n, t_n | I);  \quad \dots  
+\end{equation}
+or, equivalently, all momenta for the distribution are provided: 
+\begin{align}
+    \nonumber &\mu(t_1) = \int  x_1p(x_1, t_1 \vert I)dx_1  \\
+    \nonumber 
+    & R(t_1, t_2) = \int  x_1x_2p(x_1, t_1;x_2, t_2\vert I) dx_1dx_2\\
+    \nonumber 
+    \vdots \\ \nonumber
+    &\mu_n(t_1, \dots, t_n) = \int  x_1\dots x_np(x_1; \dots; x_n \vert I) dx_1dx_2\dots dx_n\\ 
+    \nonumber
+    &\vdots
+\end{align}
+Since both descriptions require $N \to \infty$, they are impossible to achieve and very often an estimate of just the first two moments is given, hence only a partial characterization can be made. \\ 
+The first momentum is known as the mean for the process, while second moment is known as 'Autocorrelation', and gives information on the deterministic dependence of a process from its previous values.
+It is a central quantity in time series analysis and is related to the autocovariance function via:
+\begin{equation}\nonumber 
+    C(t_1, t_2) = R(t_1, t_2) - \mu(t_1)\mu(t_2).
+\end{equation} \\
+Study of time series analysis is often a hard task, but there are some special cases that simplify the analysis: stationary and autoregressive processes.  
+\subsection{Stationary processes}
+\subsubsection{Stationarity conditions}
+There are several definitions of stationarity for a process \cite{Mantegna2007Introduction}. A stochastic process is said to be 'strictly stationary' if the joint distribution for the process: 
+\begin{equation}
+p(x_1,t_1; \dots; x_n,t_n \vert I) = p(x_1,t_1 + \Delta t; \dots; x_n, t_n +\Delta t\vert I)
+\end{equation}
+is constant in time whatever the order $n$ and time lag $\Delta t$.
+This implies that its moments are also constant in time.
+In this case, the time series consists in independent identically distributed terms.
+Such a process is rarely encountered and more general definitions are needed for real case studies.
+\\
+On of the most important is the 'wide sense stationary process', whose constraints are: 
+\begin{equation}
+    \mu_1(t) = \mu; \quad R(t_1, t_2) = R(t_2 - t_1) = R(\tau).
+\end{equation}
+Such a process has constant mean and autocorrelation that only depends on time lag between the considered values. \\
+The mean does not bring high information about the process, and can always be reabsorbed in the process defining 
+\begin{equation}
+    \nonumber
+    y(t) = x(t) - \mu.
+\end{equation}
+y(t) is a stationary process with 0 mean but same statistical properties of $x(t)$.
+$R(\tau)$ brings information on the amount of dependence of the time series on its previous values: if some process repeats similarly after $\bar t$ seconds, autocorrelation will have a secondary maximum at $\tau = \bar t$.
+Also, if the function is periodic of period $T$, the autocorrelation will have same periodicity.
+In general, autocorrelation represents most of the information we can obtain from the study of a stationary stochastic process, and it is rather fundamental.
+Its computation is even more important in the case of a Gaussian process.
+For such a process it is the only moment required to fully characterize it.
+One of the most important results regarding the autocorrelation function, is the Wiener-Khinchin theorem.
+It ensures that, for a wide sense stationary process, the frequency domain counterpart of the autocorrelation function is the power spectral density:
+\begin{equation}
+    R(\tau) = \int  S(\omega) e^{-\imath \omega \tau} d\omega,
+\end{equation}
+Autocorrelation and power spectral density bring the same information about the process.
+It is fundamental to develop methods to accurately estimate the power spectral density in such a way that it is consistent with the autocorrelation function.
+\subsubsection{Gaussian noise}
+MAXENT is a powerful tool to study noise.
+If one considers a wide sense stationary process  $\vec n(t) = \left(n_1, n_2, \dots, n_k \right)$ with 0 mean and known autocorrelation matrix, one can assign uniquely a normal distribution:
+\begin{equation}\label{eq:noisePDF}
+    p\left(\vec n(t)\vert I\right) = \frac{1}{\left(2\pi\det R \right)^{\frac{k}{2}}}\exp\left(-\frac{1}{2}\sum_{i, j}n_i R^{-1}_{ij}n_j\right)\,.
+\end{equation}
+In the case of a wide sense stationary process, being $R_{ij} = R(t_j - t_i)$ dependent on time steps only, the autocorrelation matrix is a Toeplitz Matrix (see Appendix \ref{app:Toeplitz}) and can be diagonalized, in the limit of large number of samples, by a Discrete Fourier Transform:
+\begin{equation}
+    S = F R F^{-1}. 
+\end{equation}
+For the Wiener-Khinchin theorem, $S$ is the power spectral density for the noise itself. Defining
+\begin{equation}
+    \tilde n(\omega) = F\left[n(t) \right],
+\end{equation}
+with $F$ the Fourier transform, the probability distribution (\ref{eq:noisePDF}) for noise in frequency domain (FD) becomes  
+\begin{equation}
+    p(\tilde n(f) \vert I) = \frac{1}{\left(2 \pi \det S \right)} \exp \left(-\frac{1}{2}\sum_{i} \frac{\tilde n_i \tilde n_i}{S_i} \right).
+\end{equation}
+Power spectral density plays a central role in noise study: it is the covariance matrix for noise in frequency domain. Also, in large N domain, frequency noise is serially uncorrelated. \\
+There is a very special case of Gaussian Noise that we have to mention: white noise. White noise can be represented as Gaussian noise with autocorrelation function
+\begin{equation}
+    R_{ij} = \sigma^2 \delta_{ij},
+\end{equation}
+ it is a strictly stationary process whose autocorrelation function is 0 for any time lag different from 0, and its realization is made of independent identically distributed samples.
+ Going in the Fourier Domain
+\begin{equation}
+S_{ij} = \sigma^2 F_{ik}\delta_{kl}F^{-1}_{lj} = \sigma^2 \delta_{ij} = R_{ij}, 
+\end{equation}
+we show that power spectral density for white noise is constant over frequency.
+
+\subsection{Autoregressive process}\label{sec:ARprocess}
+The description of a process is simplified if we are dealing with some autoregressive process of order p, or simply an AR(p) process.
+For an AR(p) process, each value depends linearly from its previous values
+\begin{equation}
+    x_i = \sum_{i = 1}^{p}a_{i}x_{i - p} + \nu_i, 
+\end{equation}
+with $a_i$ some constant coefficients and $\nu_i$ white noise. \\
+An AR(1) process is known as a Markov process.
+Complete knowledge on AR(p) can be achieved with the computation of $p + 1$ momenta (or conditional distributions) only.
+We prove it for a Markov process.\\
+Consider the $n'th$ order joint probability distribution
+\begin{equation}\label{eq:NthOrderPDF}
+    p(x_1, t_1; x_2, t_2; \dots; x_n, t_n | I),
+\end{equation}
+from the definition of conditional probability
+\begin{equation}\nonumber 
+p(x,y \vert I) = p(x \vert y I) p(y \vert I)
+\end{equation}
+one can rewrite \ref{eq:NthOrderPDF} as 
+\begin{equation} \nonumber 
+    p(x_n, t_n; \dots; x_1, t_1 \vert I) = p(x_n, t_n \vert x_{n-1}, t_{n-1}; \dots x_1, t_1 I) p(x_{n-1}, t_{n-1} \dots x_1, t_1 \vert I). 
+\end{equation}
+Since under Markov assumption, conditional $pdf$ only depends on the previous value for the variable 
+\begin{equation} \nonumber 
+    p(x_n, t_n \vert x_{n-1}, t_{n-1}; \dots, x_{1}, t_1 I) = p(x_n, t_n \vert x_{n-1}. t_{n-1} I) 
+\end{equation}
+reiterating the decomposition of the joint probability, we find that
+\begin{equation}
+\left(\ref{eq:NthOrderPDF} \right)=  p(x_n, t_n \vert x_{n-1}, t_{n-1}, I )p(x_{n-1}, t_{n-1}\vert x_{n-2}, t_{n-2}, I) \dots p(x_1, t_1\vert  I),
+\end{equation}
+so that for every order $n$, the probability distribution can be reduced to the product of $n$ second order probability distributions. \\
+The knowledge of the distribution for the variable and of the second order conditional distribution fully characterize a Markov process.\\
+The same reasoning can be repeated for any order $p$.
+For an AR(p) process the $n$-th order joint probability distribution ($n > p$) can be rewritten as a product of the $p$-th order conditional distributions, so that they completely characterize the process.
+
+\subsection{ARMA process and Autoregressive stationary process}
+There are two fundamental results for an autoregressive process that are strictly related with Burg's Algorithm (see section \ref{sec:MESA}) for spectral analysis, and they are all about autoregressive stationary processes and their generalization, ARMA(p,q) processes.
+\subsubsection{ARMA process}
+We will only need to define ARMA(p,q) and their relation with AR(p) processes without stepping in many details. These concepts are needed to give a stronger background to our future results.  \\
+The generalization of the Autoregressive Process, the $ARMA(p,q)$ process, is defined by the following relation: 
+\begin{equation}
+    x_t = \sum_{i = 1}^p \alpha_i x_{t - i} + \sum_{i = 1}^q \beta_i \nu_{t - i} + \nu_t,
+\end{equation}
+again, $\nu$ is white noise and both $\alpha$ and $\beta$ are constants.
+It is evident that an autoregressive process is a special case of the more general $ARMA(p,q)$ process, since  $Ar(p) = ARMA(p,0)$.
+We will not enter in the details of the description of $ARMA(p,q)$ processes, but they are important for two main reasons.
+\\
+First, $ARMA$ is a parsimonious way to characterize an $AR$ process.
+It is, in fact, possible to prove that an $ARMA(p,q)$ process, with both $p, q$ finite, is equivalent to some $AR(\infty)$ process \cite{Haykin}.
+\\
+The second and even more important result about these processes is Wold's Theorem\cite{Wold}: 
+\begin{theorem}
+Every real valued stationary stochastic process $P$ can be written as an $ARMA(p,q)$ process. 
+\end{theorem}
+This theorem makes clear the central role played by the ARMA representation: it allow us to fully characterize any stochastic process if its coefficients are known.
+This is usually an hard task, but, since we know the $AR / ARMA$ equivalence, we can find an approximate description of the process in terms of an $AR$ process, all is needed is a way to estimate the autoregressive coefficients.
+\\
+\subsubsection{Autoregressive stationary process}
+The last important ingredient is the condition of stationarity for AR processes. The conditions of stationarity for a stochastic process are well known. Defining the 'lag' operator $B$ such that 
+\begin{equation}
+    \nonumber 
+    B x_{i} = x_{i - 1},
+\end{equation}
+one can rewrite the autoregressive process as 
+\begin{equation} \nonumber 
+    x_i - \sum_{k = 1}^p a_k B^k x_i = \nu_i, 
+\end{equation}
+or, defining $b_0 = 1$ and $b_k = - a_k$
+\begin{equation}
+    \left(\sum_{k = 0}^p b_k B^{k}\right) x_i = \nu_i.
+\end{equation}
+The polynomial on the left side is known as the 'characteristic polynomial' of the autoregressive process $\phi (B) = \sum_{k = 0}^p b_k B^k$. 
+The condition of stationarity for autoregressive processes is
+\begin{theorem}
+An autoregressive process of order $p$ is stationary if all the roots $\beta_i$ of its characteristic polynomials lay outside the unit circle, i.e. $\vert\beta_i\vert > 1$.
+\end{theorem}
+Consider for simplicity a Markov process that only takes real values
+\begin{equation}\nonumber 
+    x_t - a_1 x_{t- 1} = \nu_t; \quad \text{ with } a_i \in \mathbb{R}
+\end{equation}
+whose characteristic polynomial is
+\begin{equation}
+    \nonumber 
+    \phi(B) = 1 - a_1 B .
+\end{equation}
+Let $\beta$ be the root of the polynomial $\phi(\beta) = 0$. 
+If $\vert \beta \vert > 1$ ($\vert \beta \vert < 1$) $a_1$ has to be $\vert a_1 \vert < 1$ ($\vert a_1 \vert > 1$).
+It is easy to see that if condition of stationarity is not met, the process explodes and $x_t$ will quickly diverge.
+\\
+These few concepts are fundamental in stochastic processes and necessary for what follows.
+\section{Maximum Entropy Spectral Analysis}\label{sec:MESA}
+From our previous discussion on time series, the importance of the autocorrelation function should be clear.
+Since it contains the same information of a power spectral density, high resolution computation of a power spectrum from a single time series $\vec x(t)$ of length $N$ is a fundamental task in signal analysis.
+\\
+Since we are dealing with a finite number of samples $N$, it is not possible to have an accurate computation of the autocorrelation function for every time interval.
+Also, the error $\sigma$ in the estimate increases with the time interval $k$ as $\sigma \sim \frac{1}{\sqrt{N - k}}$, so that only a few values for the autocorrelation function can actually be computed.
+\\
+The problem we want to address is how to compute a PSD from the partial knowledge of the autocorrelation function. \\
+The standard approach of peridiograms \cite{Lomb} \cite{Scargle} consists in taking the Fourier Transform of the sample autocorrelation, written as \begin{equation}
+    \rho = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\}.
+\end{equation}
+The sequence $W$ is a window function that can be chosen in several different way, each method having pros and cons for the final estimate of the power spectral density. \\ 
+The choice of a window function is arbitrary and made by trial and error, until a satisfactory compromise between variance and resolution of the estimate of power spectral density  is reached.
+A high frequency resolution implies high variance and vice-versa.
+Another problem is that the result depends on this choice, and this fact may raise the question of which we should consider as the "real" power spectral density.
+\\
+Another problem with this approach is the requirement for the window to be $0$ outside the interval: we are arbitrarily assuming $\rho_j = 0$ for $j > M$, and modifying the estimate if a non-rectangular window is chosen.
+Making assumptions on unavailable data and modifying the ones we have at our disposal does not seem a completely fair approach.
+\\
+MAXENT gives us the possibility to develop a completely new way for the estimation of PSDs, without any a priori assumptions on the unavailable data.
+Our requirements (the constraints for the variational problem) are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation.
+\\
+To do so, there is a problem we have to solve: our definition of entropy (equation \ref{eq:entropy}) depends on the probability distribution, not the power spectrum.
+We need to find a way to change the definition's domain for the entropy to formulate the variational problem in terms of the power spectral density $S(f)$ alone.
+A way out of this problem is considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA}.
+One can show that the 'entropy gain' obtained filtering white noise is:
+\begin{equation}\label{eq:EntropyGain}
+    \Delta H = \int_{-\infty}^{\infty}\log S(f) df , 
+\end{equation}
+so that we can reformulate the variational problem requiring $S(f)$ to maximize the entropy gain.
+In real examples we have no access to all frequencies but only on those between minus and plus the Nyquist frequency, defined from the sampling interval $\Delta t$ as
+\begin{equation}
+    Ny = \frac{1}{2 \Delta t}.
+\end{equation}
+All our quantities are well defined in this range only, so that we are not going to maximize \ref{eq:EntropyGain}, but
+\begin{equation}\label{eq:EntropyGain2}
+    \Delta H = \int_{- Ny}^{Ny}\log S(f) df.
+\end{equation}
+Defining $\bar r_n$ as the sample autocorrelation, our constraints will be
+\begin{equation}\label{eq:MaxConstraint}
+    \int_{-Ny}^{Ny} S(f) e^{\imath 2 \pi f n \Delta t} df = \bar r_{n}.
+\end{equation}
+This approach on power spectral density estimate was developed by Burg \cite{burg1975maximum} and it is a way out from the problems raised in periodograms estimate.
+It requires no assumptions on the unavailable estimates for $\bar r_n$ and is consistent with the sample autocorrelation function by construction.
+The hard task is now to show that this problem has a solution.
+Remarkably, it admits an analytical solution in a closed expression for $S(f)$, that we now work out.
+\subsection{MESA solution}
+We will follow the solution proposed by Burg in his PhD thesis \cite{burg1975maximum}.
+Instead of directly solving the variational problem for the power spectral density, we will rewrite it in terms of the constraints (\ref{eq:MaxConstraint}) of the problem.
+$S(f)$ is defined as the Fourier Transform of the autocorrelation function, as:
+\begin{equation}
+    S(f) = \frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{- \imath 2 \pi n \Delta t}.
+\end{equation}
+Substituting in \ref{eq:EntropyGain2}, the variational problem reduces to the maximization of: 
+\begin{equation}
+    \int_{-Ny}^{Ny} 
+    \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{-\imath 2 \pi f n \Delta t}
+    \right).
+\end{equation}
+As we said before, not all the values for the autocorrelation are known.
+Generally, we have access to a finite number of values for the autocorrelation function, ranging in a symmetric interval $n \in [-N, N]$.
+In the previous equation, since the summation index is ranging between $-\infty$ and $+\infty$, we are correctly assuming that also the unknown values have to be used in the computation of the power spectral density.
+But, since they are not known, we have to require the maximization of entropy to be independent of these values:
+\begin{equation}\nonumber 
+    \frac{\delta H}{\delta \bar r_s} = 0 \text{ for } \vert s \vert > N. 
+\end{equation}
+Using the explicit form for $H$, this becomes
+\begin{equation}
+      \frac{\delta H}{\delta \bar r_s} = \frac{1}{2Ny}\int_{-Ny}^{Ny}S(f)^{-1}e^{-\imath 2 \pi f s \Delta t} = \lambda_s = 0 \text{ for } \vert s \vert > N. 
+\end{equation}
+We see that $S(f)$ can be expressed in terms of the $\lambda s$ via a Fourier Series
+\begin{equation}\label{eq:PSDconstraint}
+    S(f)^{-1} = \sum_{s = -N}^N \lambda_s e^{-\imath 2 \pi f s \Delta t},
+\end{equation}
+This is a non-linear closed expression for the power spectral density. The requirement for $S(f)$ to be positive definite implies that: 
+\begin{equation}
+    \nonumber 
+    \lambda_{-s} = \lambda_s^* 
+\end{equation}
+With this property, and defining $z = e^{-\imath 2 \pi f \Delta t}$, $S(f)^{-1}$ can be rewritten as a polynomial in $z$ and $z^*$
+\begin{equation}
+    \label{eq:zExp}
+    S(f)^{-1} = \lambda_0 + \sum_{s = i}^N \lambda_s z^s + \sum_{s = 1}^N \lambda^*_s z^{-s}.
+\end{equation}
+Let $z_0$ be a root for the polynomial
+\begin{equation}
+    \nonumber
+    \lambda_0 + \sum_{s = 1}^N \lambda_s z_0^s + \sum_{s = 1}^N \lambda^*_s z_0^{-s} = 0,
+\end{equation}
+it is easy to show that $(z_0^*)^{-1}$ is also a root for the equation: 
+\begin{align}
+    \nonumber 
+    &\lambda_0 + \sum_{s = 1}^{N} \lambda s \left(\frac{1}{z_0^s} \right)^* + \sum_{s = 1}^N \lambda^*_s \left(\frac{1 }{z_0^{-s}}\right)^* = \\ \nonumber 
+    &\lambda_0 + \sum_{s = 1}^N \lambda_s (z_0^{-s})^* + \sum_{s=1}^N \left(\lambda_s (z_0^s)\right)^* = \\ \nonumber 
+    &\left(\lambda_0 + \sum_{s=1}^N \lambda_s z_0^s + \sum_{s = 1}^N \lambda^*_s z_0^{-s}\right)^* = 0 
+\end{align}
+This means that for every pole $z_0$ laying outside (inside) the unit circle, there is another pole inside (outside) the unit circle.
+So, if there are no roots on the unit circle, $N$ roots lay outside and $N$ inside of it.
+\\
+These properties allow us to rewrite the Fourier expansion (\ref{eq:zExp}) as \cite{1975STIN...7714318B}
+\begin{equation}\nonumber 
+    S(f)^{-1} = \frac{1}{P_N \Delta t} \left(1 + \sum_{s}^N a_s z^s\right)\left(1 + \sum_s^N  a^*_s z^{-s}\right),
+\end{equation}
+or, defining $a_0 = 1$,
+\begin{equation}\label{eq:MESApsd}
+    S(f) = \frac{P_N \Delta_t}{\left(\sum_{s=0}^N a_s z^z\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}. 
+\end{equation}
+The MESA expression for the power spectral density admits a non-linear, all pole representation when evaluated in $z = e^{-\imath 2 \pi f \Delta t}$.
+All the roots of the first polynomial can be chosen to lay outside the unit circle, so that all the roots of the second polynomial will lay inside of it \cite{1975STIN...7714318B}.
+If one is able to compute the values of the coefficients $a_i$ the PSD is uniquely determined from the previous equation. \\ 
+To compute the $a_i$-s, we substitute in equation
+(\ref{eq:MaxConstraint}) with the above representation for $S(f)$ in terms of the $a_i$ coefficients.
+Changing  variable from $f$ to $z(f) = e^{-\imath 2 \pi f \Delta t}$, the integral (\ref{eq:MaxConstraint}) becomes a contour integral in the unit circle centered at the origin of the complex plane:
+\begin{equation}
+   \frac{P_N}{2 \pi \imath} \oint _{\mathbb S^1}\frac{z^{-s - 1}}{\sum_{n = 0}^N a_n z^n \sum_{n = 0}^N a^*_n z^{-n}} = \bar r_s. 
+\end{equation}
+Sending $s \to s - r$, multiplying by $a^*_s$ and summing over all possible values for $s$ the equation becomes 
+\begin{align} \nonumber 
+    \sum_{s = 0}^N a_s \bar r_{s - r} &= \frac{P_N}{2 \pi \imath}\oint \frac{z^{r - 1} \sum_{s = 0}^N a^*s z^{-s}}{\sum_{s = 0}^N a_s z^s \sum_{s = 0}^N a^*s z^{-s}} dz\\
+    & = \frac{P_N}{2 \pi \imath}\oint \frac{z^{r -1}}{\sum_{s = 0}^N a_s z^s}dz\label{eq:ErFilter}
+\end{align}
+Since all the zeros of the denominator lay outside the unit circle, the contour integral is 0.
+The only exception is for $r = 0$ with a residual in the origin of the axis.
+For $r = 0$ the integral can be computed via the Cauchy integral formula and (\ref{eq:ErFilter}) becomes:
+\begin{align}\label{eq:errorFilter1}
+    \sum_{s = 0}^N a_s \bar r_{r - s} &= P_N \quad \text{ if } r = 0 \\ \label{eq:errorFilter2}
+    \sum_{s = 0}^N a_s \bar r_{r - s} & = 0 \qquad \text{ if } r \neq 0.
+\end{align}
+This kind of equations are solved by the use of the Levinson recursion \cite{doi:10.1002/sapm1946251261}
+\subsection{Levinson Recursion}
+The equations (\ref{eq:errorFilter1}) and (\ref{eq:errorFilter2}) can be rewritten as a Matrix equation of the form
+\begin{equation}\label{eq:predictionErrorFilter}
+    \begin{pmatrix}
+    \bar r_0 &  \bar r_{-1} & \dots & \bar r_{-N}\\
+    \bar r_{1} & \bar r_ 0 & \dots & \bar r_{1 - N} \\ 
+    \bar r_{2} & \bar r_{1} & \dots & \bar r_{2 - N}
+    \\
+    \vdots & \vdots & \ddots & \vdots  \\
+    \bar r_{N} & \bar r_{N - 1} & \dots  &\bar r_0
+    \end{pmatrix}
+    \begin{pmatrix}
+    1 \\   a_1 \\   a_ 2 \\  \vdots \\   a_N
+    \end{pmatrix} = 
+    \begin{pmatrix}
+    P_N \\   0 \\  0 \\ \vdots  \\   0
+    \end{pmatrix}
+\end{equation}
+know as the prediction error filter equation.
+It is a matrix equation involing the Toeplitz Matrix $R_{ij}$ and the vector $\vec a$, known as the forward prediction error filter.
+The approach for solving the equations is the Levinson - Durbin recursion.
+\\
+Our goal is to obtain the $N$-th order equation (\ref{eq:predictionErrorFilter}) from the $N-1$-th order equation: 
+\begin{equation}\label{eq:Forward}
+    \begin{pmatrix}
+    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N}\\
+    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} \\ 
+    \vdots & \vdots & \ddots & \vdots \\ 
+    \bar r_{N - 1} & \bar r_{N -2} & \dots & \bar r_0
+    \end{pmatrix} 
+    \begin{pmatrix}
+    1  \\   b_1  \\  \vdots \\   b_{N - 1}
+    \end{pmatrix} =
+    \begin{pmatrix}
+    P_{N -1} \\  0 \\ \vdots \\ 0
+    \end{pmatrix} .
+\end{equation}
+This equation is not sufficient to solve the recursion.
+Since the autocorrelation matrix is self-adjoint, from equations (\ref{eq:errorFilter1}) and (\ref{eq:errorFilter2}) and noting that the first and the last row of the autocorrelation matrix are respectively the complex conjugate reverse, is easy to show that
+\begin{equation}\label{eq:Backward}
+     \begin{pmatrix}
+    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N}\\
+    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} \\ 
+    \vdots & \vdots & \ddots & \vdots \\ 
+    \bar r_{N - 1} & \bar r_{N -2} & \dots & \bar r_0
+    \end{pmatrix} 
+    \begin{pmatrix}
+    b^*_{N-1} \\ b^*_{N- 2} \\\vdots \\ 1
+    \end{pmatrix} 
+    = 
+    \begin{pmatrix}
+    0 \\ 0 \\ \vdots \\ P_{N-1}
+    \end{pmatrix} 
+\end{equation}
+holds. The vector obtained reflecting b and taking its complex conjugate is known as ``backward prediction error". \\ 
+Stepping from order $N - 1$ to order $N$, the autocorrelation matrix grows from size $(N,N)$ to size $(N + 1, N + 1)$, this implies that a new component $b_n$ for the prediction error filter is needed.
+If we take it to be $b_N = 0$, this choice does not modify the first $N$ component of the right side of equation (\ref{eq:Forward}) (or the last $N$ components of (\ref{eq:Backward})), i.e.:
+
+\begin{equation}\label{eq:OrderN}
+    \begin{pmatrix}
+    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N} & \bar r_{- N}\\
+    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} & \bar r_{1 - N}\\ 
+    \vdots & \vdots & \ddots & \vdots & \vdots \\ 
+    \bar r_{N -1} & \bar r_{N-2} & \dots & \bar r_0 & \bar r_{-1}
+    \\
+    \bar r_{N } & \bar r_{N -1} & \dots & \bar r_1 & r_0 
+    \end{pmatrix} 
+    \begin{pmatrix}
+    1  \\   b_1  \\  \vdots \\   b_{N - 1} \\ 0 
+    \end{pmatrix} =
+    \begin{pmatrix}
+    P_{N -1} \\  0 \\ \vdots \\ 0 \\ \Delta_N
+    \end{pmatrix} .
+\end{equation}
+with 
+\begin{equation}
+    \Delta_N = \sum_{n=0}^N \bar r_{N - n}b_n.
+\end{equation}
+Adding the previous expansion for both forward and backward equations, one obtains 
+\begin{equation}\nonumber 
+        \begin{pmatrix}
+    \bar r_0 & \bar r_{-1} & \dots & \bar r_{- N}\\
+    \bar r_{1} & \bar r_0 & \dots & \bar r_{1 - N}\\ 
+    \vdots & \vdots & \ddots & \vdots \\ 
+    \bar r_{N -1} & \bar r_{N-2} & \dots &  \bar r_{-1}
+    \\
+    \bar r_{N } & \bar r_{N -1} & \dots  & r_0 
+    \end{pmatrix} 
+    \begin{pmatrix}
+    1  \\   b_1  \\  \vdots \\   b_{N - 1} \\ 0 
+    \end{pmatrix} 
+    + c_N 
+    \begin{pmatrix}
+    0 \\ b^*_{N -1} \\ \vdots  \\ b_1 \\ 1 
+    \end{pmatrix}
+    =
+    \begin{pmatrix}
+    P_{N -1} \\  0 \\ \vdots \\ 0 \\ \Delta_N
+    \end{pmatrix} .
+    + c_N
+    \begin{pmatrix}
+    \Delta^*_N \\ 0 \\\vdots \\ 0 \\ P_{N-1}
+    \end{pmatrix}.
+\end{equation}
+With the reflection coefficient $c_N$ to be computed. Requiring the previous equation to be equivalent to \ref{eq:predictionErrorFilter}, one obtains
+\begin{equation}\nonumber 
+\begin{cases}
+    P_{N - 1} + c_N\Delta^*_N = P_N \\
+    c_N P_{N - 1} + \Delta_N = 0
+\end{cases}
+\end{equation}
+which can be easily solved for $c_N$ and $P_N$ as 
+\begin{equation}
+    c_N = -\frac{\Delta_N}{P_{N-1}}; \quad \text{ and } P_N = P_{N - 1}\left(1 - \vert c_N \vert ^2 \right). 
+\end{equation}
+Having computed the value for the reflection coefficient, the $N$-th order forward prediction error filter is found to be 
+\begin{equation}
+    a_i = b_i + c_N b^*_{N - i}.
+\end{equation}
+To obtain the solution for the $N$-th order, one starts $\bar r_0 = P_0$ and recursively computes the new values until the desired order is computed. 
+This problem is often solved numerically with Burg's Algorithm \cite{Vos}.
+A python implementation of Standard Burg's Method can be found in Appendix \ref{app:Burg} 
 \section{The Autoregressive process analogy}
-The application of MESA is not limited to spectral estimates, but it also provides a bridge between spectral analysis and the analysis
+The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the analysis
 of autoregressive processes.\cite{doi:10.1029/RG013i001p00183}.
@@
 
+\chapter{Solar Spots Analysis}
+\section{Introduction}
+Having studied in detail MESA and understood what optimizer provides a better reproduction of spectral densities, in this chapter we provide a first and simple application of the method on the number of Sun spots as a function of time.
+MESA is a natural resolution to the
+problems raised and studied by Yule and Walker, since the autoregressive
+component naturally arises from the data and does not require any major work.   \\
+Such a study is interesting for two reasons: it is an application on a simple dataset where most of the variables can be easily taken under
+control. Secondly, the study of Sun spots allow us to apply MESA for both
+spectral analysis and time series forecasting. 
+At first, we estimate the PSD of the time series on the whole dataset.
+Secondly, we will estimate the filter on a reduced dataset and use the result to forecast the time series.
+We will then compare the estimate with the unused, observed, data.
+\section{Analysis of Sun Spots}
+First attempts to give a description of the periodical behaviour of solar
+activity is due to Yule and Walker \cite{doi:10.1098/rsta.1927.0007}\cite{1931RSPSA.131..518W}, who first proposed that the time series of solar spots does not have to be regarded as some
+perfect sinusoidal signal with a mere superposition of fluctuations, but is
+better described if a deterministic autoregressive component is considered
+in the description. A description with some deterministic dependence
+on the previous values of the series, allowed to reconstruct with higher precision the
+time dependence of the amplitude of the process. 
+In their description the choice for autoregressive order
+has to be made with different trials. It seems to be clear that, at least
+from this point of view, after a method for selecting the autoregressive 
+order has been chosen, MESA provides a natural solution to this study. 
+\begin{figure}
+    \centering
+    \includegraphics[scale = .45]{Images/SunSpots/SunSpotsNumbers.pdf}
+    \caption{Monthly number of solar spots.}
+    \label{fig:sunSpotsPlot}
+\end{figure}
+Fig \ref{fig:sunSpotsPlot} shows the monthly time series of Zurich Sunspots Number in a time period from $1749$ until $1983$. The periodicity in the time series is eye catching. \\
+We are going to estimate the time series spectrum as a means of capturing periodicity in our dataset, MESA is applied with the FPE chosen as optimizer.  
+\begin{figure}
+    \centering
+    \includegraphics[scale = .45]{Images/SunSpots/SunSpotsPSD&ACR.pdf}
+    \caption{Top panel: power spectral density for Zurich Monthly Number of Sun spots. Bottom panel: reconstructed autocorrelation function.}
+    \label{fig:sunSpotsPSD&ACR}
+\end{figure}
+Figure \ref{fig:sunSpotsPSD&ACR} (bottom panel) shows the the autocorrelation function.
+It is a slowly decaying function of time whose peaks are located at multiples of 121 months (around 10 years), reproducing the periodicity of the sample.
+A slowly decaying function is a sign of the process to be long range.
+Figure \ref{fig:sunSpotsPSD&ACR} (top panel) shows the frequency estimate of the Power Spectral Density obtained with MESA.
+The 0 frequency component is different from zero, indicating that the mean of the process is different from zero.
+There is evidence of some well defined periodicity at some frequency $f_0$, superimposed to some $f^{-\alpha}$ noise, both $f_0$ and $\alpha$ to be determined.
+MESA is also able to capture secondary peaks, at frequencies that are multiple of the fundamental frequency.
+\\
+To study it, we consider the power spectral density as a superimposition of colored noise and a Cauchy resonance.
+A fit of PSD background, see the top panel in Figure \ref{fig:SunSpotsFit}, shows that the decreasing component of the power spectral density can be actually treated as colored noise of the form:
+\begin{equation}
+    f_{noise} = (1.1 \pm .2) f^{-1.74 \pm 0.03}
+\end{equation}
+\begin{figure}
+    \centering
+    \includegraphics[scale = .45]{Images/SunSpots/SunSpotsFit.pdf}
+    \caption{Power spectral density against fitting of background with colored noise (top) and with a Cauchy distribution (bottom).}
+    \label{fig:SunSpotsFit}
+\end{figure}
+
+In this way, the non periodic component can be eliminated from the spectrum and the peak can be fit.
+Even if not perfectly symmetrical, we chose a Cauchy distribution to fit the peak (Figure \ref{fig:SunSpotsFit}, below), parameterized as
+\begin{equation}
+    C(f) = \frac{\hat{h} }{1 +\left( \frac{f - \hat f_0}{\hat \gamma}\right)^2}.
+\end{equation}
+The fit provided for the parameters the following values
+\begin{equation}
+    \nonumber 
+    \hat h = (3.35 \pm 0.02)\cdot 10^5; \quad 
+    \hat f_0 = (8.104 \pm 0.003) \cdot 10^{-3} \mbox{months}^{-1}; \quad 
+    \hat \gamma = (5.49 \pm 0.05) \cdot 10^{-4} \mbox{months}^{-1}
+\end{equation}
+and the periodicity in Sun spots is found to be 
+\begin{equation*}
+    10.28 \pm 0.05 y_{ears}.
+\end{equation*}
+It is interesting to understand the meaning of the parameters by taking an inverse Fourier Transform, the autocorrelation correspondent to a pure Cauchy shaped PSD is:
+\begin{equation}
+    K \gamma e^{-\gamma \vert t \vert } e^{-\imath f_0 t}.
+\end{equation}
+$f_0$ is the natural frequency of the oscillating component for the power spectrum, while $\gamma$ is the inverse of a lifetime $\gamma = \frac{1}{\tau}$ and can be taken as a typical time scale for the autocorrelation of the periodic component, but this is not true for the whole dataset since its power spectral density is, at first approximation, the sum of the two components.
+Since they do not provide additional information, secondary peaks (Figure \ref{fig:sunSpotsPSD&ACR}) have been neglected.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale = .47]{Images/SunSpots/SunSpotsTotalFit.pdf}
+    \caption{Fits to the spectrum considered with two components, one $f$ noise component and one Cauchy component for the resonance.}
+    \label{fig:PSDendFit}
+\end{figure}
+Figure \ref{fig:PSDendFit} shows the fit obtained adding the $f$ noise component with a Cauchy resonance.
+The result is accurate enough to describe the main component of the power spectral density: having neglected secondary peaks does not affect the computation of the quantity of interest.
+
+\section{Time Series Forcasting}
+With MESA we have been able to compute a smooth and high quality power spectral density for the time series of Sun spots, estimating the period of solar activity to be $  10.28 \pm 0.05 y_{ears}$.
+This is just a first result of the application of MESA.
+On the other hand, the computation of the prediction error filter allows us to predict the future values for our observations as well as their credibility regions.
+Since we want to compare the results of the forecast with real data, we apply MESA again on our dataset, excluding the 200 last data points.
+We will then compare the excised data with the predictions of MESA.
+Having computed the prediction error filter from the data, eliminating the first value (which is one) and changing the sign to the remaining, we have an estimate of the autoregressive coefficients.
+In this way, The process is estimated to be an $AR(96)$ process.
+\begin{figure}
+    \centering
+    \includegraphics[scale = .47]{Images/SunSpots/SunSpotsARcoefficients.pdf}
+    \caption{Estimate of the autoregressive coefficients for Sun Spots time series.}
+    \label{fig:SunSpotsFilter}
+\end{figure}
+Figure \ref{fig:SunSpotsFilter} shows estimate for the autoregressive coefficient of the time series. 
+The behaviour of the coefficient as a function of the length is characteristic, the dependence is different from zero and positive for only few values, and becomes a noisy fluctuation around $0$ of both positive and negative components after few observations.
+The standard deviation $\sqrt{P_N}$ of the superimposed white noise is estimated to be $\sqrt{P_N} = 14.77$.
+Having computed the filter and variance for the $AR$ predictor, new values can be inferred by
+\begin{equation}
+    x_t = \sum_{i = 1}^{96}a_i x_{t-i} + \nu_t \text{ with } \nu_t \sim \mathbf{N}(0, \sqrt{P_N}).
+\end{equation}
+In every step of the prediction, the deterministic component has to be summed to the white noise component.
+Since the process is linearly dependent on its previous values, this random component will propagate in the time series.
+For this reason, in forecasting it makes no sense to consider a realization only: even if a single prediction is an admissible realization for the process, taken alone it has no statistical meaning.
+\\
+For this reason, to have a fair comparison between simulated and observed realization, we will simulate 3000 series of 200 points each. \\ 
+Since random error propagates in the time series, we expect our prediction to be a consistent estimate of the observed time series for a number of points shorter than filter's length.
+In this interval, in fact, both simulation and real data are still dependent on the observed data.
+For predictions of length comparable with AR order or longer, our prediction is based on simulated data only and can be largely different from the observed series.
+\begin{figure}
+    \centering
+    \includegraphics[scale = .47]{Images/SunSpots/SunSpotsForecast.pdf}
+    \caption{Forecasting on Sun spots time series using MESA linear predictor. Observed time series is compared with ensemble mean and 95\% Cr.}
+    \label{fig:SunSpotsForecast}
+\end{figure}
+Figure \ref{fig:SunSpotsForecast} shows the numerical prediction of MESA (red dashed line), computed as a mean over the ensemble, and the future values assumed by the series (black solid line) for 200 months, credibility regions at 95\% are also reported.
+\\
+The vertical line at an abscissa of $96$ provides a visual separation between a time interval shorter and longer than the autoregressive order for the process.
+Is evident that this line separates two different regimes for the estimate.
+At earlier times, our inference is almost everywhere consistent with the observed data, and they are everywhere comprised in the $95\%$ regions.
+\\
+At later times, the quality of inference is degraded, because the inference in this region is no more dependent on observed data.
+Even if the credible regions become larger there are some values laying outside them.
+Secondly, the observed data are no more consistent with the estimate obtained from the mean.
+\\
+The only information we can achieve in this area is about the increasing / decreasing trend of the series.
+This happens because this trend is determined by the deterministic component of the process.
+\\
+So, we conclude that a proper inference with MESA can be obtained for a time period shorter than the length of the filter, in that case a period of $70\sim 80 months$, i.e.
+around $6$ years, which is a remarkable result.
+\\
+It is important to stress a fact: to obtain all the results, we assumed no knowledge on the process itself.
+To study power spectral density, gain quantitative information and make high quality forecast, we did not assume any physical model.
+All we obtained was directly contained in the data themselves and MESA extracted as much information as possible.
+This shows how the PSD - Autorgressive duality is powerful.
+With the support of Wold's theorem, given a stationary process, MESA is a tool that allow to find an approximate description of its ARMA(p,q) dual without making any assumption on the process.
+
+\section{Conclusions}
+In this section we applied MESA to a first simple dataset: the number of Sun spots.
+Even if simple, this dataset makes clear some important features of the method.
+\\ Firstly, we have been able to estimate accurately the power spectral density of the process.
+We have seen that, at least at the leading order, it can be treated as the sum of a colored noise component and a periodic component, described by a Cauchy resonance.
+The f noise component has an exponential decay with exponent $\alpha = -1.74 \pm 0.03 $ .
+The periodic component is associated to a frequency
+\begin{equation}
+    \nonumber 
+    \hat f_0 = \left(8.104 \pm 0.003 \right) \cdot 10^{-3}. months^{-1}
+\end{equation}
+Secondly, we worked out with the autoregressive duality computing its AR coefficient from the prediction error filter.
+The AR process is found to be well approximated by an $AR(96)$ process.
+Having computed the coefficients, we showed how MESA can be used to  forecast over a time period shorter than the AR order $p$.
+This is because in that range, our estimate depends on both observed and simulated data.
+With no knowledge at all on the physics behind the process, we have been able to perform accurate inference and predictions with the only application of MESA.
+This is a very remarkable result and shows how MAXENT provide a powerful tool to make inference without superimposing any model to the dataset.
+
+\newpage 
+\chapter{Sinusoidal Signals}
+\section{Introduction}
+The estimation of parameters describing theoretical models is one of the key point in statistical Analysis.
+Considering some dataset and model, dependent on a set of parameters $\vec \theta$, the objective is to find the values for $\vec \theta$ that better describe the observation.
+\\
+Such an estimate can only be performed if noise is correctly characterized, otherwise the inference might be imprecise.
+Since data come out as a superposition of signal and noise, the characterization of noise is often made 'off-source', i.e.
+studied in a time window of proper length that precedes the stretch of data containing the signal.
+Such an analysis assumes the noise to be stationary for a period of time and is seldom justified.
+In this chapter, we explore a novel idea that allows to perform an "on-source" characterization of noise by the use of MESA.
+After a short explanation of the basic idea, we will apply this new method to a toy problem of sinusoidal signal superimposed with colored noise.
+\section{Bayesian inference}
+\subsection{Bayes Theorem}
+Bayes theorem is the fundamental tool for Bayesian inference. \\ 
+Let $\vec \theta \in \mathbb{R}^N$ be the set of parameters we want to infer.
+The objective in Bayesian Analysis is the computation of the posterior distribution for the parameters given some data $D$, prior information $I$ and model $M$.
+\\
+Starting from Bayes theorem, the goal is to determine the distribution $P(\vec \theta \vert D M I)$ such that 
+\begin{equation}\label{eq:Bayes}
+    P(\vec \theta \vert D M I) = \frac{p(D\vert \vec \theta M I)}{p(D \vert M I )} p(\vec \theta \vert M I) 
+\end{equation}
+holds. \\ 
+The posterior on the left hand side indicates how the parameters are distributed according to the model, prior information and the observation of some data $D$. \\ 
+On the right hand side $p(\vec \theta \vert M I)$ is the prior distribution for the parameters given only model and prior information, while $p(D \vert \vec \theta M I)$ is the Likelihood.
+The quantity in the denominator $p(D\vert M I)$ is a purely Bayesian quantity called evidence.
+For sake of simplicity, since we are not going to use it explicitly, being computed as
+\begin{equation} 
+    p(D \vert M I) = \int d\vec \theta p(D \vert \vec \theta M I) p(\vec \theta \vert M I),
+\end{equation}
+we can consider it as a normalization factor. \\ It is clear that, to perform Bayesian inference,
+we need to correctly determine the previous quantities. \\ 
+Concerning the prior distribution, MAXENT provides a way to assign them, but there are several more methods and a huge literature about it \cite{jaynes2003ptl}. \\ 
+\subsubsection{Assigning Likelihood}
+Our main interest is the determination of the likelihood from the information and a Model $M$. \\
+The Model $M$ defines functional relations between physical quantities, dependent on the set of parameters $\vec \theta$.
+Given some value $x$ for a variable, the model $M$ allows us to determine the value of a causally related variable $z$, via
+\begin{equation}
+    \nonumber 
+    \vec z = M(\vec x \vert \vec \theta).
+\end{equation}
+The previous relations are our physical laws that we want to test and the determination of the parameters is a fundamental task to assert the validity of some Model or theory.
+In general, we do not observes pairs of values $(z_i, x_i)$, but we have to deal with some error that can mask the deterministic relation between the two.
+Out data will be "couples" of points of the form
+\begin{equation}
+    D = (y_i, x_i),
+\end{equation}
+where $y_i$ is determined from the model $M$ and some error $e_i$, via 
+\begin{equation}
+    y_i = z_i + e_i .
+\end{equation}
+In general, each of the previous quantities, the observation $y_i$, the prediction from the model $z_i$ and the error $y_i$ will have their own probability distributions.
+The Likelihood can be obtained marginalizing the joint probability distribution over $z_i$ and $e_i$, assuming them to be independent:
+\begin{align}\nonumber 
+    p_y(y_i \vert M \vec \theta I) &= \int de_i dz_i p(y_i, z_i, e_i \vert M \vec \theta I) = \int de_i dz_i p_y(y_i \vert z_i e_i M \vec \theta I) p_z(z_i \vert M \vec \theta I) p_e(e_i \vert M \vec \theta I)
+\end{align}
+Since $y_i$ is uniquely determined from the model prediction $z_i$ and the error $e_i$, its conditional distribution is 
+\begin{equation}\nonumber 
+    p(y_i \vert z_i e_i M \vec \theta I) = \delta(y_i - z_i - e_i),
+\end{equation}
+so that 
+\begin{align}\nonumber 
+    p_y(y_i \vert M \vec \theta I) &= \int de_i dz_i \delta(y_i - z_i - e_i) p_z(z_i \vert M \vec \theta I) p_e(e_i \vert M \vec \theta I )\\ \nonumber 
+    &= \int dz_i p_z(z_i \vert M \vec \theta I) p_e((y_i - z_i) \vert M \vec \theta I) 
+\end{align}
+that is the convolution between the distribution of the model and the distribution for the error.
+For a deterministic model, such that $z_i = m(x_i \vec \theta)$ is uniquely determined, the previous can be furhter simplified
+\begin{align}
+    p(y_i \vert M \vec \theta I) &= \int dz_i \delta\left(z_i - m(x_i \vert \vec \theta)\right) p_e\left((y_i - z_i) \vert M \vec \theta I\right) \\ \nonumber 
+    & = p_e\left((y_i - m(x_i\vert \vec \theta)) \vert M \vec \theta I\right)
+\end{align}
+and for such a model the likelihood is uniquely determined from the distribution of the errors.
+For independent data $y_i$ and errors $e_i$, given a set of observations, the Likelihood can be rewritten as
+\begin{equation}
+    p(\vec y \vert M \vec \theta I) = \prod_{i}p_e\left(y_i - m(x_i \vert \vec \theta) \vert M \vec \theta I \right).
+\end{equation}
+Given some deterministic model $M$, the Likelihood is uniquely determined by the distribution of the errors,
+so that a full characterization of the error is a necessary step to compute the posterior for the distribution. \\ 
+\newpage
+\section{Power Spectral density Computation}
+From the previous discussion, the role of power spectral density in parameter estimation appears to be clear: Parameter estimation can only be done if an estimate for the variance of the noise is given.
+\\
+We briefly review how this is done for signals and propose a new 'on-source' method based on MAXENT. 
+\subsection{Pure noise signals} 
+Computation of power spectral density for pure noise signals of the form
+\begin{equation}
+    y(t) = n(t)
+\end{equation}
+is the easiest task, since our observation is coincident with noise only and we don't have to extract it from the data.
+Also if simple, this is the very start to study even more complex data that are a superposition of signal and noise.
+The most traditional approach is based on Bayes theorem.
+We can consider noise to be normally distributed with 0 mean and with some variance to be determined. To infer on the posterior distribution, a prior distribution for $\sigma$ have to be assigned
+\begin{equation}
+    \nonumber
+    p(\sigma \vert D I) = \frac{p(D \vert \sigma D I)}{p(D \vert I)}p(\sigma \vert I).
+\end{equation}
+Firstly, we need to rewrite the previous in Frequency domain with discrete time sampling.
+Under these conditions, the so-called Whittle Likelihood \cite{Rover} provides a discretized version of a normal Likelihood in frequency Domain:
+\begin{equation}
+    \label{eq:Whittle}
+    -\frac{1}{2}\log L(\tilde n(f_i) \vert \vec S) = - \sum_{j = - N /2}^{N / 2}\left(\log(S(f_j) +  \frac{\frac{\Delta_t}{N}\vert \tilde \tilde n \vert ^2}{S(f_j)} \right), 
+\end{equation}
+being $\tilde n$ the Fourier Transform of noise.
+In this domain, a prior on Power Spectral Density has to be assigned also.
+In this way, the posterior on the power spectral density can be uniquely determined as
+\begin{equation}
+    p(S(f) \vert D I ) = p(\tilde n \vert S I) p(S(f) \vert I) 
+\end{equation}
+This approach allows us to determine a posterior for the power spectral density and make an estimate with error bars. 
+This easy example can become very complex in real cases, since the number of parameters can be arbitrarily large and a priori assignment for each of them is needed. \\
+But, neglecting technicalities, such an approach is still possible. \\ 
+It is interesting to see how MESA enters in this case. Having a realization of pure noise, the determination of power spectral density simply reduces to the application of MESA to the time series.
+The disadvantage is that, with respect to Bayesian inference, we will only have an estimate for the power spectral density and no other quantities.
+The advantage is that in complicated cases, this estimate is model independent, and no assumption on noise behaviour are needed.
+Such a simple examples makes clear what are the main differences between MESA and model based approaches, both having their pro and cons depending on the case of study.
+\subsection{Off Source Estimate}
+Even if pure noise signals are not in general our main goal, they provide a useful way to perform parameter estimation in case of data that are a superposition of both signal and noise:
+\begin{equation}
+\nonumber 
+d(t) = z(t) + e(t).  
+\end{equation}
+In cases like this, the main objective is to compute the Posterior for the parameters $\vec \theta$ so that an estimate on their value $\vec \theta_0$ based on the observed Data can be performed. \\ 
+Equation (\ref{eq:Bayes}) only formally solve this problem: to apply it, assigning likelihood is a necessary step and so posterior can be computed only after an estimate for power spectral density and noise characterization is needed.
+\\
+One of the most common techniques for solving this problem is the "off source noise characterization" \cite{offsource}.
+With off source studies, noise characterization is not made in the same time window where signal also lives. If signal duration is $T$, noise is studied in the $T$ seconds before signal detection. \\ 
+In this way, our problem can be divided in two subsequent steps, the first one being the characterization of a pure noise signal and the second one being the estimate of parameters via Bayes theorem, that can be made with proper algorithms like MCMC.
+\\
+Off source characterization is an efficient and fast way to characterize noise and is very proficient when noise is stationary. 
+But in the case of non stationary noise such an approach shouldn't be applied since it could bring to misleading and wrong results.
+
+\subsection{On-source Estimate}
+Our objective is to propose an 'On-source' computation to characterize noise in the same time window where signal lives, to perform a better estimate of parameters in case of non stationary noise.
+To avoid making assumptions on power spectral density shape, we will use MESA for its estimate and computation.
+This will allow us to reduce at a minimum the number of assumptions needed to perform the estimate.
+\\
+We will explain our idea referring to Markov Chain Monte Carlo (MCMC) algorithm. 
+\subsubsection{The Idea in a nutshell}
+The conceptual simplicity of this class of algorithms is ideal to explain the idea of on source estimate without introducing blinding technicalities.  \\ 
+Given some observations $D(t)$ described with some model $M(\vec \theta)$, MCMC algorithm allow to sample parameters space to find the maximum of posterior distribution for the parameters $\vec \theta$.
+\\
+Let N be the number of recursion for the algorithm, and suppose Likelihood and prior are assigned. The structure of MCMC for posterior sampling is the following:
+
+\begin{enumerate}
+    \item Set $N = N_0$; 
+    \item Set $i = 0$; 
+    \item Randomly samples $\vec \theta_i$ values from prior; 
+    \item Compute value for Posterior on samples $\vec \theta_i$
+    \begin{equation*}
+        P_i = P(\vec \theta = \vec \theta_i \vert D M) = p(D \vert \vec \theta_i M) * p(\vec \theta_i \vert M) 
+    \end{equation*}
+    \item Update parameters $\theta_i$
+    \begin{equation}\nonumber 
+        \vec \theta_{k} = \vec \theta_i + \vec \epsilon_i \qquad \text{with: } \vec \epsilon_i \sim D(0, \vec \sigma) 
+    \end{equation}
+    \item Compute $P_k$
+    \item If: 
+        \begin{enumerate}
+            \item $P_k \geq P_i$: Set $\vec \theta_{i+1} = \vec \theta_{k}$
+            \item $P_k < P_i$, Set: \\ 
+            $\vec \theta_{i + 1}= \theta_k$ with probability $\frac{P_k}{P_i}$ or \\ $\vec \theta_{i+1} = \theta_i$ with probability $1 - \frac{P_k}{P_i}$
+        \end{enumerate}
+    \item if $i = N_0$: end
+    \item increment $i \rightarrow i + 1$ 
+    \item back to step 5. 
+\end{enumerate}
+With such an algorithm one can produce an arbitrary number of samples representative of the peak for the posterior distribution. \\ 
+If the power spectral density is given, it has to be used to compute the posterior value. What happens if we don't know it? \\ 
+The simplest approach is considering it as a parameter and infer it simply sampling a larger posterior distribution of the kind: 
+\begin{equation}\nonumber 
+    P(S_1, \dots, S_{N}, \theta_1, \dots, \theta_N \vert D M I). 
+\end{equation}
+This would require the introduction of a prior distribution for every component of the PSD.
+Conceptually it is a practicable approach, but for large data set of $N$ points one should infer $N / 2$ values for the power spectral density, and complexity would increase accordingly.
+\\
+Our objective is to propose a different approach based on MAXENT. At first, we note that at step $(6)$ the computation of the Posterior implies that we are computing the quantity: 
+\begin{equation}\nonumber 
+    d(t) - z_i(t),  
+\end{equation}
+with $z_i$ the model computed on the current values $\vec \theta_i$ for the parameters. 
+Collaterally, an estimate of the noise $n_i(t) = d(t) - z_i(t)$ is provided as a function of the current value for the parameters. \\ 
+Having a sample of possible realization for noise, the basic idea is to compute $z_i$ power spectral density with MESA and use this estimate in the evaluation of the posterior.  \\ 
+The modification needed in the algorithm to insert the computation of PSD is very simple, and can be resumed as follows: 
+\begin{enumerate}
+    \item Set $N = N_0$; 
+    \item Set $i = 0$; 
+    \item Randomly samples $\vec \theta_i$ values from prior; 
+    \item Compute noise as $n_i = d - z_i$
+    \item Compute power spectral density $S_i$ applying MESA on $n_i$
+    \item Compute value for Posterior on samples $\vec \theta_i$ and $S_i$
+    \begin{equation*}
+        P_i = P(\vec \theta = \vec \theta_i \vert D S_i M) = p(D \vert \vec \theta_i S_i M) * p(\vec \theta_i \vert M) 
+    \end{equation*}
+    \item Update parameters $\theta_i$
+    \begin{equation}\nonumber 
+        \vec \theta_{k} = \vec \theta_i + \vec \epsilon_i \qquad \text{with: } \vec \epsilon_i \sim D(0, \vec \sigma) 
+    \end{equation}
+    \item Compute noise as $n_k = d - z_k$
+    \item Compute power spectral density $S_k$ applying MESA on $n_k$
+    \item Compute $P_k$
+    \item If: 
+        \begin{enumerate}
+            \item $P_k \geq P_i$: Set $\vec \theta_{i+1} = \vec \theta_{k}$
+            \item $P_k < P_i$, Set: \\ 
+            $\vec \theta_{i + 1}= \theta_k$ with probability $\frac{P_k}{P_i}$ or \\ $\vec \theta_{i+1} = \theta_i$ with probability $1 - \frac{P_k}{P_i}$
+        \end{enumerate}
+    \item if $i = N_0$: end
+    \item increment $i \rightarrow i + 1$ 
+    \item back to step 5. 
+\end{enumerate}
+This method has many great advantages. \\
+Firstly, it doesn't require any assumption on prior for power spectral density and so minimizes the number of assumptions needed. \\ 
+Secondly, we end up having a large number of samples for power spectral density also, and its posterior can be reconstructed. \\ 
+The most important implication is that we can marginalize the posterior for the parameters with respect to the power spectral density, obtaining a conservative - model independent estimate of their distribution.
+\\
+One objection could be raised at this point. 
+We saw that MESA spectrum is analytical when stationary noise is considered, since it has no poles on the unitary circles where PSD is actually evaluated.
+One could state that we are trying to abandon the too restrictive hypothesis for noise to be stationary, studying it with some method that requires is stationarity.
+\\
+To answer this objection some specifications are needed, in fact, this statement is correct.
+The assumption of stationarity is a simplifying assumption that allow us to study noise with very powerful tools.
+Our objective is not to completely relax this assumption, but to make it less restrictive.
+With off-source analysis, we are assuming noise to be "approximately stationary"  for a time period that is (at least) twice the duration of the signal.
+\\
+On source analysis requires signal to be approximately stationary in the same time window where signal lives, that is a less stringent hypothesis.
+This also clarify the domain of application of this on-source method.
+If the typical time scale of significant changes of noise is larger than the duration of the signal, on-source parameters estimate with MESA might provide an efficient solution to the problem.
+If typical time scale are shorter or comparable with signal duration, this method cannot ensure to provide a correct estimate.
+
+\section{Sinusoidal Signal} 
+As proof-of-principle we perform on-source study with the above method on two different time series: a sinusoidal signal with the superimposition of brown noise and a sinusoidal signal with a drift in the frequency, superimposed to noise generated from Ligo Power Spectral Density.
+For both its stability and fast computation, FPE is the ideal optimizer to perform this kind of analysis.
+Off-source estimate of the data will also be provided using Welch Method and Alternative CAT.
+Since noise is stationary, we expect the on-source method to provide an equivalent estimate on parameters.
+\subsection{Brown Noise}
+As first step, we generate the sinusoidal signal 
+\begin{equation}
+    a_1 \sin(2\pi f t) + a_2 \cos(2\pi f t)
+\end{equation}
+with values for amplitude and frequency parameters chosen to be
+\begin{equation}\nonumber 
+    a_1 = 5, a_2 = 7, f = 7.3. 
+\end{equation}
+The signal is superimposed with Brownian noise, whose PSD is 
+\begin{equation}
+    S(f) = \frac{3}{f^2 + 0.01}
+\end{equation}
+whit 0.01 inserted to avoid the divergence at f= 0.
+Figure (\ref{fig:BrownData}) represents the dataset we will analyze.
+First half of the data are a pure noise signal, while the second half a the superimposition of noise and sinusoidal signal with a signal to noise ratio (SNR) $\sim 37\%$
+
+\begin{figure}
+    \centering
+    \includegraphics[scale = 0.40]{Images/BrownData/BrownData.pdf}
+    \caption{Dataset under analysis.
+First half of the data are pure noise signal d(t) = n(t), needed to perform off-source characterization of noise, while the second half is the superposition of noise and sinusoidal signal d(t) = n(t) + z(t), that will be used to perform parameter estimation with both off-source/on-source methods}
+    \label{fig:BrownData}
+\end{figure}
+This dataset is composed of 2000 points.
+The whole allow us to perform off-source analysis computing the power spectral density off noise in the first 1000 and parameter estimation with the second 1000.
+This second half will be the only one needed to perform on-source  analysis.
+The results of parameters estimation and PSD reconstruction will be then compared.
+\subsubsection{Off-source analysis}
+As first step to perform off-source analysis, the reconstruction of the power spectral density is needed. Such an estimate is obtained with both Welch Method and Alternative CAT method. 
+\begin{figure}
+    \centering
+    \includegraphics[scale = .40]{Images/BrownData/OffSource_PSDcomparison.pdf}
+    \caption{Comparison of true power spectral density with estimate obtained with Welch method (green) and Alternative CAT (red).
+Alternative CAT estimate is made on sub arrays of 100 points with a 50\% superimposition}
+    \label{fig:brownPSD_offSource}
+\end{figure}
+Figure \ref{fig:brownPSD_offSource} compares the estimate of the power spectral density obtained with Welch Method (green) and Alternative Cat (Red).
+Both of the estimates are representative for the power spectral density, but aCAT minimizes the variance of the result.
+The parameter estimation is obtained with the MCMC sampling of a Nested Sampling
+Algorithm \cite{Skilling..735..395S} \cite{john_veitch_2020_4109277}. 
+Figures (\ref{fig:BrownCATposterior})shows the posterior for the parameters computed on alternative CAT (above) and Welch  (below) power spectral densities 
+\begin{figure}
+    \centering
+    \includegraphics[scale = .47]{Images/BrownData/BrownOffPosteriors.pdf}
+    \caption{Posterior distributions for the parameters obtained using alternative CAT (above) and Welch (below).}
+    \label{fig:BrownCATposterior}
+\end{figure}
+The estimate for the parameters with Welch's
+is 
+\begin{equation}\nonumber 
+    \hat a_{1; Welch} = 4.86 \pm 0.04 ; \quad \hat a_{2;  Welch} = 7.10 \pm 0.05 \quad \hat f_{Welch} = 7.30000 \pm 4 \cdot 10^{-5}
+\end{equation}
+The frequency estimate is unbiased and with a small error.
+The estimate for the second amplitude $a2$ lays in one standard deviation from the mean, while for the 1st amplitude parameter the real value is found to lay in 3.5 standard deviation.
+The results obtained with alternative CAT are more accurate
+\begin{equation}
+    \hat a_{1; aCAT} = 4.94 \pm 0.07 ;\quad 
+    \hat a_{2; aCAT} = 6.96 \pm 0.06;\quad 
+    \hat f_{aCAT} =  7.30000 \pm 0.00005
+\end{equation}
+and a correct estimate for the parameters, since their true values are always comprised in one standard deviation from the mean.
+In similar cases where noise is stationary, the off-source approach can give results that are consistent with the real values for a signal.
+\subsubsection{On-Source Analysis}
+On source estimate is performed directly on the dataset, so in the second half part of the data in figure \ref{fig:BrownData}, and the posterior sampling is obtained via the same Nested sampling algorithm used before.
+\begin{figure}
+    \centering
+    \includegraphics[scale = .47]{Images/BrownData/BrownOnPosteriors.pdf}
+    \caption{Posterior distribution for the parameters obtained as on-source estimate}
+    \label{fig:brownOnPosterior}
+\end{figure}
+Figure (\ref{fig:brownOnPosterior}) shows the posterior distribution for the parameters obtained with the on source estimate.
+With respect to the posteriors obtained with the off-source estimate, the domain is larger and so is the error in the estimate.
+the larger support for the posterior in the on source case is due to the fact that the power spectral density is not 'a priori' known, but evaluated step by step with the parameters.
+The uncertainty on the posterior affect the parameter estimation, resulting in larger posterior distributions.
+The estimate for the parameters obtained with on-source analysis is
+\begin{equation}\label{brownONestimate}
+    \hat a_1 = 5.0 \pm 0.2 ; 
+    \quad \hat a_2 = 7.0 \pm 0.4;
+    \quad f = 7.3000 \pm 0.0001 
+\end{equation}
+Remarkably, the estimate results to be unbiased for every parameter, concerning the standard deviation it result to be one order of magnitude larger than that obtained with the off-source estimate.\newline
+\begin{figure}
+    \centering
+    \includegraphics[scale = .43]{Images/BrownData/On_Source_PSDposterior.pdf}
+    \caption{On-source reconstructed posterior with 90\% credibility regions. FPE estimate on real error is also reported. }
+    \label{fig:brownOnSourcePSD}
+\end{figure}Figure (\ref{fig:brownOnSourcePSD}) show the on-source estimate for the power spectral density obtain with $90\%$ Credbility regions.
+FPE estimate on the real Error is also reported.
+The estimate is computed taking the median over all the reconstruction obtained in the Monte Carlo.
+Median is chosen over the mean to be more representative for its greater stability.
+\\
+The true power spectral density is not consistent with the median of the ensemble, neither comprised in the $90\%$ Credibility.
+But it also has to be noticed that the median on-source estimate is almost coincident with the FPE estimate for the power spectral density, that is computed on the real error.
+This fact allow us to state that the off-set in the estimate is due to MESA and not to on-source estimate.
+The fact that the estimate of PSD on real error is coincident with our estimate is rather a good result, and the estimate of noise stress this fact.
+\begin{figure}
+    \centering
+    \includegraphics[scale = .43]{Images/BrownData/OnSource_ReconstructedNoise.pdf}
+    \caption{'On-source' estimate of noise}
+    \label{fig:brownNoise}
+\end{figure}
+Figure (\ref{fig:brownNoise}) compares the real error (black) with its estimate (red crosses).
+This estimate is computed taking the median of the errors over all the possible models reconstructed with MCMC algorithm.
+The agreement is impressive and gives even more evidence that the disagreement between real PSD and the reconstruction is due to MESA only.
+This is easily understood comparing MESA analytical solution with the functional form we used to generate noise.
+Neglecting constants, we are requiring
+\begin{equation}\nonumber 
+    \nonumber 
+    \biggr \vert\sum_{i = 0}^p a_i e^{- \imath 2 \pi f i \Delta t}\biggr \vert ^2 \sim f^2. 
+\end{equation}
+The expansion on the left hand side could be consistent with $f$ if long orders are selected.
+In general $f^{-2}$ noise is brownian motion spectrum, that is an $AR(1)$ non-stationary process.
+Under these assumptions we are working outside the limits of MESA, and so we are not ensured that we will obtain a consistent estimate.
+In fact, a check on the autoregressive coefficient shows that the filter obtained with FPE is of order 3, too short to be an accurate approximation.
+It interesting to note that the estimate obtained with alternative CAT method in the previous section, shown in figure \ref{fig:brownPSD_offSource}, is a way more accurate estimate for the power spectral density.
+Reminding that CAT chooses long filters, it is more able to capture the power low behaviour of the power spectral density, and the variance of the estimate is reduced taking the average over different realizations.
+All these facts underline that the non consistent estimate for the power spectral density is not to be considered as a failure of our proposal for on-source estimates.
+\newpage 
+\section{Ligo Noise}
+The same analysis will be repeated on a sinusoidal signal with a linear drift $f'$ in the frequencies 
+\begin{equation}
+    z(t) = a_1 \sin\left(2 \pi (f + f' t) t\right) + a_2 \cos\left(2\pi(f + f' t)t \right), 
+\end{equation}
+whose parameters are valued 
+\begin{equation}\nonumber 
+    a_1 = 10^{-21}; \quad a_2 = 5\cdot 10^{-22}; \quad f = 128, \quad f' = 1.4 
+\end{equation}
+the sampling interval and observation time are respectively 
+\begin{equation*}
+    dt = \frac{1}{2048}; \quad T = 8,
+\end{equation*}
+both parameters and sampling time expressed in arbitrary units. 
+The signal is superimposed with some noise randomly generated from the Ligo Power Spectral Density studied in chapter 2. 
+\begin{figure}
+    \centering
+    \includegraphics[scale = .43]{Images/LigoOnSource/ObservedData.pdf}
+    \caption{Simulated dataset used to perform the analysis}
+    \label{fig:LigoData}
+\end{figure}
+Figure \ref{fig:LigoData} shows the dataset used for both 'off-source' and 'on-source' analysis.
+As a main difference from the previous case, the amplitude for noise is much greater than the amplitude of the signal.
+\\
+The first 4 seconds of observation are pure noise, while the second are a superimposition of the deterministic signal $z(t)$ and noise $n(t)$.
+Again, the first 4 seconds will be used to study the power spectral density with the off-source estimate, while the last 4 seconds to perform parameter estimation.
+For these last for second SNR is reduced to $\sim 10^{-3}$.
+\subsection{Off-Source Analysis}
+The off source analysis is performed characterizing noise with the alternative CAT approach and Welch approach.
+The reconstruction of the power spectral densities with the above methods is reported in figure \ref{fig:offPSDLigo}.
+\begin{figure}
+    \centering
+    \includegraphics[scale = .43]{Images/LigoOnSource/offPSDcomparison.pdf}
+    \caption{Comparison of the estimate of the power spectral density obtained with Alternative CAT (red line) and Welch method (green line)}
+    \label{fig:offPSDLigo}
+\end{figure}
+Both peaks are recognized but not perfectly characterized, especially the first peak's height is underestimated.
+Still, alternative CAT's estimate shows improved variance with respect to Welch's method.
+A comparison of the posterior for the parameters obtained with both methods is shown in figure (\ref{fig:posteriorLigoComparison}).
+\begin{figure}
+    \centering
+    \includegraphics[scale = .46]{Images/LigoOnSource/PosteriorsComparison.pdf}
+    \caption{Posteriors for the parameters of the model for both Alternative CAT (above) and Welch (Below)}
+    \label{fig:posteriorLigoComparison}
+\end{figure}
+Alternative CAT results are peaked around the true results, while the posteriors obtained with Welch's method are slightly biased.
+The estimate for the parameters obtained with alternative CAT are found to be
+\begin{align*}
+\centering
+    \hat a_1 = (1.005 \pm 0.005) \cdot 10^{-21}; \quad 
+    &\hat a_2 = (4.99 \pm 0.09) \cdot 10^{-22}; \\
+    \hat f = 128.000 \pm 0.002; \quad 
+    &\hat f' = 1.3998 \pm 0.0004, 
+\end{align*}
+so that the true values are always found to lay in one standard deviation from the mean. As can be seen from the posteriors, the estimate for the parameters obtained with Welch's method 
+\begin{align*}
+\centering
+    \hat a_1 = (0.999 \pm 0.003) \cdot 10^{-21}; \quad 
+    &\hat a_2 = (5.29 \pm 0.05) \cdot 10^{-22}; \\
+    \hat f = 127.998 \pm 0.001; \quad 
+    &\hat f' = 1.4004 \pm 0.0002
+\end{align*}
+gives less accurate results.
+The only parameters that lays in one standard deviation from the mean is the first amplitude parameter $a_1$.
+Both drift and frequency are correctly estimated in 2 standard deviations, that is a reasonable result, while for the second amplitude parameter (a2) the estimate is outside 3 standard deviations.
+The big variance of Welch's method in power spectral density estimate is
+naturally affecting the estimate on the parameters, but still providing reasonable results, but the averaged MESA based results of CAT are way more accurate.
+
+\subsection{On-Source Analysis}
+We now provide the result of the characterization of noise and the parameter estimate using the proposed on-source estimate.
+This case is of particular interest because signal noise ratio (SNR) is around $\sim 10^{-3}$, so that we are trying to characterize the dominant component of noise and separate it from a relatively weak signal.
+Also in the study of brownian noise, even if SNR incredibly decreased, the result in the estimate is very accurate. 
+\begin{figure}
+    \centering
+    \includegraphics[scale = .43]{Images/LigoOnSource/PArametersPosteriors.pdf}
+    \caption{Posterior distribution for the parameters obtained with the on-source estimate}
+    \label{fig:LigoOnPosteriors}
+\end{figure}
+Figure \ref{fig:LigoOnPosteriors} shows the posterior distribution computed for each parameter using the proposal method.
+Again the uncertainty on the estimate of the power spectral density propagates in the estimate for the parameters increasing the domain of the posterior and resulting in larger error bars.
+The estimate for each parameter is found to be
+\begin{align}\label{LigoEstimate} 
+    \hat a_1 = (1.00 \pm 0.01) \cdot 10^{-21}; \quad 
+    &\hat f = 128.003 \pm, 0.004 \\ \nonumber 
+    \hat a_2 = (4.7 \pm 0.2) \cdot 10^{-22} ; \quad 
+    &\hat f' = 1.3993 \pm 0.0009
+\end{align}
+all the estimate but the second amplitude are very accurate and the true value lays inside the error defined by one standard deviation.
+For the second amplitude parameter the result is comprised in two standard deviation.
+\\
+So, even for a weak signal we have been able to accurately resolve the deterministic signal described by the sinusoidal model. \\ 
+But the added value of this new approach is the characterization of noise and the computation of its power spectral density. 
+\begin{figure}
+    \centering
+    \includegraphics[scale = .43]{Images/LigoOnSource/PSDestimate.pdf}
+    \caption{Estimate for the power spectral density obtained with the on-source analysis.}
+    \label{fig:LOnSourcePSD}
+\end{figure}
+Figure \ref{fig:LOnSourcePSD} (above) shows the estimate for the power spectral density obtained with the on-source estimate, computed with FPE optimizer.
+The result of the estimate is accurate and representative of the real power spectral density.
+The credibility regions for the estimate are tight.
+This is probably due to the stability of FPE.
+We expect that for slightly small changes in the parameters the estimate for the power spectral density won't be largely affected.
+Since frequency and drift estimate are narrow, this is reflected on the posterior obtained for the power spectral density.
+This result is clearly related to the estimate and capability of the algorithm to discriminate between noise and signal.
+\begin{figure}
+    \centering
+    \includegraphics[scale = .43] {Images/LigoOnSource/ErrorEstimate.pdf}
+    \caption{Comparison of the real noise (black dashed line) with the estimate obtained with the on-source analysis (red solid line)}
+    \label{fig:LErrorEstim}
+\end{figure}
+Figure \ref{fig:LErrorEstim} shows the real error, superimposed to the data, and its estimate obtained with the on-source algorithm.
+This is computed as the ensemble median over all the reconstructed time series for the noise.
+The accuracy and precision in the reconstruction are incredibly high.
+The percentage error in the reconstruction of noise is found to be $r = 0.78 \%$, with a standard deviation $\sigma_r = 62.4\%$.
+The credibility regions are very tight, their average width is found to be $(3 \pm 1) \cdot 10^{-23}$, so that they are very narrow with respect to the values of noise time series, even though 100\% of the real values are found to lay between them.
+This explains why the accuracy in the reproduction is in one standard deviation around
+$0.2 \% \div 2 \%$ and also the very tight credibility regions in the estimate for the power spectral density. \\ 
+\begin{figure}
+    \centering
+    \includegraphics[scale = .43]{Images/LigoOnSource/NoiseEstimateBisect.pdf}
+    \caption{Noise estimate VS real noise plot. The red dashed line represents the bisector of the axis}
+    \label{fig:Bisector}
+\end{figure}
+To study the accuracy in the reproduction, the curve obtained in a "real error VS error estimate" is fitted with a straight line to see if the result is consistent with the bisector of the axis.
+Figure \ref{fig:Bisector} shows both the plot of the estimated error VS the real error and the bisector of the axis.
+The fit with a line 
+\begin{equation}\nonumber 
+    y = m x + q,
+\end{equation}
+gives as estimate for slope and intercept the following values 
+\begin{equation}
+    \hat m = 1.00000 \pm 0.00001; \quad \hat q = (0.02 \pm 1) \cdot 10^{-25},
+\end{equation}
+their functional relation is perfectly coincident with the bisector of the axis with an incredibly high precision, so that the small differences in the reproduction are probably statistically meaningless.
+With the on-source study parameters can be estimate with great accuracy, just renouncing to some precision in the result.
+Even for small values of SNR, this technique allow to distinguish between noise and signal in the data, reconstructing both in a very accurate way, just with the specification of the model for the signal.
+
+\newpage 
+\section{Conclusions}
+In this chapter we briefly discussed the parameter estimate and the role played by the computation of power spectral density for noise to achieve this goal.
+Since likelihood distribution is uniquely determined from distribution of noise, a method to characterize this distribution is necessary.
+The standard off-source analysis requires to assume that noise is stationary for a time length that is at least twice the time of the observation of the signal.
+That assumption can lead to inaccurate parameter estimate when dealing with real noise, so that less restrictive one is needed.
+\\ For this reason, we proposed a new method to perform parameter estimation, introducing MAXENT computation of power spectral density in a MCMC algorithm.
+This means that only the strain of data where the signal live is needed.
+This new approach has been applied to two different 'toy models'.
+The first consists in a sinusoid with constant frequency, superimposed to some colored noise, with a SNR around $36 \%$.
+\\
+In the second example we complicated the model by the introduction of a linear drift in the frequencies and superimposed with noise generated to be consistent with the Ligo power spectral density.
+In this second case, the signal noise ratio is reduced to $\sim 10^{-3}$.
+Comparing with the standard 'off-source' analysis, this new approach enlarges the error in the estimate of the parameters but lead to very accurate results: all the true values for the parameters have been recovered in no more than two standard deviations from the mean (eq: \ref{brownONestimate} and \ref{LigoEstimate}).
+The real improvement regards the characterization of noise.
+Just using the strain of data where the signal actually lives, this approach has the capability to discriminate with huge accuracy which is the deterministic component of the data and which is simply noise, also for very low value of SN ratio.
+The estimate of noise (Figure \ref{fig:brownNoise} and \ref{fig:LErrorEstim}) is extremely accurate and precise: even if the credibility regions are very sharp, they always comprise the real values.
+Even the reconstruction of the power spectral density is extremely accurate, and most of the differences between the real power spectral density are attributable to MESA.
+\\
+Even if the aforementioned results are based on simple toy-models, the obtained results provide enough evidence that this method could refined to try an implementation on different, real, contexts.
+
+\newpage 
+\chapter*{Conclusions and future developments}
+In this thesis we introduced the maximum entropy principle (MAXENT) with some of its most relevant applications to the study of stationary time series.
+In chapter one MAXENT has been introduced and reformulated, according to the original work of Burg, to be applied in the estimate of power spectral densities.
+This method is known as 'Maximum Entropy Spectral Analysis' (MESA).
+\\
+Burg's approach admits a closed expression for the estimator of the power spectral density  in terms of some coefficients known as 'forward prediction error coefficients' (equation \ref{eq:MESApsd}). 
+These $p$ coefficients provide a link between the spectral estimate and the characterization of the observed stochastic process.
+In fact, simply changing their sign, these new coefficient are a 'linear predictor' for the time series (equation \ref{eq:ARspectrum}).
+This means that they allow us to express the observed time series in terms of some autoregressive process of order $p$ (AR(p)).
+\\
+Even more remarkably, the coefficients obtained with MESA are found to minimize the error in the prediction (equation \ref{eq:bestLinearPredictor}), i.e.
+their computation is equivalent to a least square fitting for the time series with an $AR(p)$ model.
+It is evident that importance of MESA goes far beyond his original domain of application: with no assumption on the physical model behind the observed data,  it accurately reconstructs the spectrum and characterizes the process..\\
+Burg's algorithm is recursive and, for dataset whose autocorrelation function is not perfectly know, requires some criterion to estimate the recursive order.
+Different criteria to address this problem can be found in literature, and in chapter two we characterized three of them: Final Prediction Error (FPE), Criterion on Auoregressive Transfert function (CAT) and Optimum Bayes Decision rule (OBD).
+The study considered the accuracy in the reproduction of two different spectra: a normal distribution and the a simplified spectrum of noise for the LIGO experiment (figure \ref{fig:LigoPSDOriginal}).
+\\
+The error in the reproduction resulted to be strongly dependent on the estimate of the recursive order $p$ (Figures \ref{fig:optcomparison} and \ref{fig:LigoOrderError}), and in both cases the estimate for $p$ associated with the smallest error is obtained with FPE.
+OBD lays in a region of shorter filter and larger error .
+The reconstructions obtained with these two methods are similar, but OBD provides estimates with a larger bias.
+CAT, instead, shows a completely different behaviour.
+It selects larger orders than FPE,  but does not converge to any specific order.
+These long arrays are associated to a decreased bias in the estimate but with higher variance: it introduces a large number of spurious peaks in the reproduction.
+\\
+For its peculiar properties, CAT's estimates might be improved averaging on various estimates obtained from subsets of the original dataset.
+The result will depend on both the length of the subsets and their overlapping.
+Short arrays (and so shorter filters) provide more biased results but less variance, while the opposite holds for longer arrays.
+For 'short' arrays the estimate is similar to the results obtained with FPE, but with lower accuracy.
+For longer arrays, it provides a similar result to the standard CAT but with reduced noise, so that this alternative approach makes sense if relatively ``long" arrays are chosen.
+In chapter 4 this method is found to be way more accurate than $FPE$ (Figures (\ref{fig:brownPSD_offSource}) and (\ref{fig:brownOnSourcePSD})) in the reconstruction of power spectral density for Brown Noise.
+\\
+In every application of this alternative method, the overlapping between the segments of data was fixed to $50\%$, so that a future perspective could be the characterization of the method for different lengths and overlapping.
+\\
+This modified approach has several cons.
+MESA is a powerful approach also because it is free of prior assumptions.
+The introduction of this alternative approach require some a priori choices on length and overlapping of sub segments of data that are hard to justify, we can only appeal to some 'expected behaviour' of the estimate with no general arguments supporting this choice.
+Also, it increases the complexity of the algorithm.
+But the worst loss regards the link with autoregressive processes: taking the average, every information on the autoregressive coefficients is completely lost.
+\\
+For both its accuracy and its stability, FPE has been chosen as the best optimizer to improve the estimate on the recursive order.\\ 
+There are two other aspects that would deserve some studies. Firstly, for both spectra we investigated the estimate for fixed length of the time series.
+Since $AR(p)$ processes are, for $p$ finite, just an approximation for the more general $ARMA(p,q)$ process, we expect the estimate for $p$ to be dependent on the length $N$ of the input time series.
+It would be interesting to see how and if the length of the estimate of recursive order is dependent on the original length of the time series $N$.
+\\
+Secondly, since our interest was in characterizing noise, we studied the three optimizer focusing on power spectral density only.
+To achieve a full characterization of the criteria, one should generate an $AR(p)$ process with known coefficients and compare them with the estimates for length and values obtained with the three methods.\\
+Finally, having characterized the three optimizer, in chapter 3 we applied MESA on a very simple time series: sun spots number.
+We obtained a description of the time series in terms of a AR(96) process.
+Fitting the spectrum (figure \ref{fig:sunSpotsPSD&ACR}) with some colored noise for the background and with a Cauchy resonance for the peak, we found a periodicity in the series of $10.28 \pm 0.05 years$.
+Knowing the forward prediction error, we have also been able to accurately predict the future values of the time series for $\sim 80$ months (figure \ref{fig:SunSpotsForecast}).
+\\
+The last chapter developed a new proposal to perform 'on-source' bayesian parameter estimation, inserting MESA computation in a Markov Chain Monte Carlo algorithm (MCMC).
+With this approach, a model for the power spectral density is constructed "in real time" with the sampling for the parameters.
+In this way, with a relatively small number of assumptions, posterior samples are obtained for both the parameters of the model and the power spectral density.
+This proposal is convenient because, given some observed data $d(t) = z(t) + n(t)$, it allow us to separate the deterministic signal component $z(t)$ from the noise $n(t)$ directly 'on' the data.
+This approach allow us to relax the restrictive hypothesis on the stationarity for noise, that are usually required when noise is characterized 'off-source'.
+\\
+This approach has been tested on a sinusoidal signal superimposed to brown noise and noise generated with the same LIGO Spectrum studied in the second chapter.
+We obtained very promising results.
+Concerning the estimate of parameters, we have great accuracy in the estimate (equations \ref{brownONestimate} and \ref{LigoEstimate}), with the loss of one order of magnitude in precision.
+This lost is probably related to the uncertainty in the estimate of the power spectral density.
+The most remarkable result is about the noise reconstruction from the data: the algorithm has been able to almost perfectly distinguish and separate noise and signal from the data (figures \ref{fig:brownNoise} and \ref{fig:LErrorEstim}).
+A linear least square fit of the estimated noise as a function of the real noise is found to be perfectly coincident with the bisector of the axis, up to a precision of $10^{-5}$, so that the differences between the real values for noise and its estimate are negligible from a statistical point of view.
+\\
+From the beginning, MESA have been applied to reproduce the simplified power spectral density of LIGO noise.
+The reason behind this choice concerns our future perspectives: the very next step will be the application of the new on-source method to public data of LIGO-Virgo collaboration, that will be a test bench for our analysis: the noise in the detector is not perfectly stationary as the one we generated in our simulation.
+But, since this analysis is also performed 'off-source', considering the results obtained with our toy models, it seems very reasonable to believe that the on-source method can be applied on gravitational waves data with optimum results.
+We are confident that we can be able to characterize almost exactly the noise component in the apparatus at the time the signal was caught.
+\\
+The complete absence of assumptions on noise, makes this method a worth trying approach to several process.
+Special candidates are observations whose noise component cannot be completely understood and modeled.
+A simple example, reminiscent of our sinusoidal model, is the transit of exoplanets around a star.
+The partial eclipse due to the transit of the exoplanet in front of the star, will result in an observed luminosity that is periodic in time.
+This oscillatory, deterministic behaviour, will be superimposed to noise from various sources, such as stellar activity, the presence of the interstellar medium and many others.
+If one wants to isolate the phenomenon itself, not caring to understand what is every component of noise, MESA on-source estimate might be a good approach.
+Another, king size, example is LISA.
+For such an apparatus, it is impossible to provide a satisfactory model for noise: every signal coming from the galaxy,  such as stochastic background, binaries etc..
+is a source for noise, and it is impossible to characterize them all.
+Also, noise is highly non stationary so that off-source analysis can lead to very misleading results.
+Having modeled what the "signal" is, one might attempt to use our on-source estimate to discriminate between signal and noise, giving up on characterizing every source of noise and focusing on reconstructing the overall noise only.
+Clearly, the aforementioned examples require a lot of technical work to do, and MESA on-source analysis is just "the core" of way more complex analysis framework.
+For the exoplanet complications may arise from non-uniform sampling interval and from the fact that data can be recorded for 6 months only.
+For LISA a problem might be the understanding of what has to be considered a signal and what is noise.
+MESA on source estimate seems to be a promising approach to separate noise and signal in a variety of cases, but is just the core for most  applications: every implementation requires some ad hoc analysis and characterizations.
+
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END & BIB

### 2021-02-08 examples/mem_welch_comparison.py
Added comparison between MESA and Welch

@@ -1,0 +1,99 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Feb  4 09:56:33 2021
+
+@author: alessandro
+"""
+
+from memspectrum import MESA
+from memspectrum import GenerateTimeSeries
+from scipy.signal import welch, tukey
+import scipy.stats 
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+def welch_estimate(noise, N, dt, overlap_fraction = .5):
+    sampling_rate = 1 / dt
+    segment_duration = N * dt #lunghezza sottoarray
+    padding = 0.4 / segment_duration 
+    window_function  = tukey(int(sampling_rate*segment_duration), padding)
+    wf, ws = welch(noise, fs  = 1 / dt,
+                              window = window_function,
+                              nperseg         = N,
+                              noverlap        = int(overlap_fraction*N),
+                              return_onesided = False,
+                              scaling         = 'density',
+                              nfft = noise.size) #dopo N // 2 frequ negative
+    return wf, ws 
+
+
+def relative_error(real, estimate):
+    return np.abs(real - estimate) / real 
+
+
+if __name__ == '__main__': 
+    
+    #Generate the Time Series 
+    f_ny = 5 
+    dt = 1 / (2 * f_ny)
+    N = 8192
+    T = N * dt 
+    frequency = np.arange(0, N // 2 + 1) / T 
+    bnoise_psd = 1 / (frequency ** 2 + 1)
+    normal_psd = scipy.stats.norm.pdf(frequency, 2.5, .5) 
+    number_of_segments = 5 #Number of segments for Welch method
+    
+    #Generate Time series 
+    time, normal_series, frequency, normal_frequency_series, normal_psd =\
+        GenerateTimeSeries.generate_data(frequency, normal_psd, T, 1 / dt)
+        
+    
+    time, bNoise_series, frequency, bnoise_frequency_series, bnoise_psd =\
+        GenerateTimeSeries.generate_data(frequency, bnoise_psd, T, 1 / dt)
+    
+    #Estimate the spectrum for normal PSD 
+    M = MESA(normal_series)
[…]

### 2021-02-08 paper/mesa.tex
Change in paper

@@
 		\email{martini.alessandr@gmail.com}
-        \affiliation{Dipartimento di Fisica  Universit\`a di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}        
+        \affiliation{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}        
         %
@@
 	\maketitle
+	\tableofcontents
 
@@
 
-\section{Maximum Entropy Principle} 
+\section{Introduction}
+\blindtext
+
+\section{Theoretical foundations}
+\subsection{Maximum Entropy Principle} 
 The Maximum Entropy principle (MAXENT) is one of the most important results in Bayesian Statistics.
@@
 
-\section{Maximum Entropy Spectral Analysis} 
+\subsection{Maximum Entropy Spectral Analysis} 
 The computation of the power spectral density is an important component in time series analysis and characterization.
@@
 
-\section{Autoregressive Process Analogy} 
+\subsection{Autoregressive Process Analogy} 
 The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the analysis
@@
 
-\section{Forecasting}
-
-\section{Choosing an autoregressive order}\label{sec:optimizers} 
+\subsection{Forecasting}
+
+\section{Validation of the model}
+\subsection{Choosing an autoregressive order}\label{sec:optimizers} 
 Before being able to apply MESA, we need to address the question of the best length $M$ of the filter that provides the optimal estimate of the power spectrum.
@@
 
-\section{Analysis of the optimizers}
-
-\section{Example}
-
-\section{Comparison with Welch method}
-
-\section{Appendix 1: MESA solution}
+\subsection{Analysis of the optimizers}
+
+\subsection{Comparison with Welch method}
+
+\section{Examples}
+
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END & BIB
+
+\appendix
+\section{MESA solution}
 \subsection{MESA solution}
@@
 
-\newpage 
-a
-\newpage
+	\bibliography{Bibliography.bib}
+	\bibliographystyle{ieeetr}
+
+\end{document}
+
+%%%%%%%%%%%%%% BELOW THE OLD PART... USEFUL TO TAKE INSPIRATION FROM
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 MESA on source estimate seems to be a promising approach to separate noise and signal in a variety of cases, but is just the core for most  applications: every implementation requires some ad hoc analysis and characterizations.
-
-	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END & BIB
-
-        \begin{acknowledgments}
-
-          R.~G. acknowledges support from the Deutsche Forschungsgemeinschaft
-          (DFG) under Grant No. 406116891 within the Research Training Group
-          RTG 2522/1. 
-          M.~B. and S.~B. acknowledge support by the EU H2020 under ERC Starting
-          Grant, no.~BinGraSp-714626.  
-          M.~B.~ acknowledges support from the Deutsche Forschungsgemeinschaft
-          (DFG) under Grant No. 406116891 within the Research Training Group
-          RTG 2522/1. 
-          %
-
-          %
-          This research has made use of data, software and/or web tools obtained 
-          from the Gravitational Wave Open Science Center (https://www.gw-openscience.org), 
-          a service of LIGO Laboratory, the LIGO Scientific Collaboration and the 
-          Virgo Collaboration. LIGO is funded by the U.S. National Science Foundation. 
-          Virgo is funded by the French Centre National de Recherche Scientifique (CNRS), 
-          the Italian Istituto Nazionale della Fisica Nucleare (INFN) and the 
-          Dutch Nikhef, with contributions by Polish and Hungarian institutes.
-        \end{acknowledgments}
-
-	\bibliography{Bibliography.bib}
-	\bibliographystyle{ieeetr}
-
-\end{document}

### 2021-02-08 paper/scripts/utils.py

@@ -1,342 +1,0 @@
-from __future__ import division
-import numpy as np
-import math
-import lal
-
-G = lal.G_SI
-c = lal.C_SI
-
-"""
-    BBH final-state fitting formulas from https://arxiv.org/abs/1611.00332
-    (c) 2016-2017 Xisco Jimenez-Forteza, David Keitel, Sascha Husa, Mark Hannam, Sebastian Khan, Michael Puerrer
-    also included in LALInference under GPL at
-    https://versions.ligo.org/cgit/lalsuite/tree/lalinference/python/lalinference/imrtgr/nrutils.py
-"""
-
-def fits_setup(m1, m2, chi1, chi2):
-    """
-    common setup function for UIB final-state and luminosity fit functions
-    """
-
-    # Vectorize the function if arrays are provided as input
-    m1   = np.vectorize(float)(np.array(m1))
-    m2   = np.vectorize(float)(np.array(m2))
-    chi1 = np.vectorize(float)(np.array(chi1))
-    chi2 = np.vectorize(float)(np.array(chi2))
-
-    if np.any(m1<0):
-      raise ValueError("m1 must not be negative")
-    if np.any(m2<0):
-      raise ValueError("m2 must not be negative")
-
-    if np.any(abs(chi1)>1):
-      raise ValueError("chi1 has to be in [-1, 1]")
-    if np.any(abs(chi2)>1):
-      raise ValueError("chi2 has to be in [-1, 1]")
-
-    # binary masses
-    m    = m1+m2
-    if np.any(m<=0):
-      raise ValueError("m1+m2 must be positive")
-    msq  = m*m
-    m1sq = m1*m1
-    m2sq = m2*m2
-
-    # symmetric mass ratio
-    eta  = m1*m2/msq
-    if np.any(eta>0.25):
-      print("Truncating eta from above to 0.25. This should only be necessary in some rounding corner cases, but better check your m1 and m2 inputs...")
-      eta = np.minimum(eta,0.25)
-    if np.any(eta<0.0):
-      print("Truncating negative eta to 0.0. This should only be necessary in some rounding corner cases, but better check your m1 and m2 inputs...")
-      eta = np.maximum(eta,0.0)
-    eta2 = eta*eta
[…]

### 2021-02-11 paper/scripts/plot_posterior_summary.py
removed files

@@ -1,146 +1,0 @@
-import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
-from scipy.stats import gaussian_kde
-import os
-from cosmolisa.likelihood import find_redshift
-import cosmolisa.cosmology as cs
-import h5py
-
-def init_plotting():
-    plt.rcParams['figure.figsize'] = (2*3.4, 3.4)
-    plt.rcParams['font.size'] = 11
-    plt.rcParams['font.family'] = 'Times New Roman'
-    plt.rcParams['font.sans-serif'] = ['Bitstream Vera Sans']
-    plt.rcParams['axes.labelsize'] = plt.rcParams['font.size']
-    plt.rcParams['axes.titlesize'] = plt.rcParams['font.size']
-    plt.rcParams['legend.fontsize'] = 10
-    plt.rcParams['xtick.labelsize'] = 9
-    plt.rcParams['ytick.labelsize'] = 9
-    plt.rcParams['xtick.major.size'] = 3
-    plt.rcParams['xtick.minor.size'] = 3
-    plt.rcParams['xtick.major.width'] = 1
-    plt.rcParams['xtick.minor.width'] = 1
-    plt.rcParams['ytick.major.size'] = 3
-    plt.rcParams['ytick.minor.size'] = 3
-    plt.rcParams['ytick.major.width'] = 1
-    plt.rcParams['ytick.minor.width'] = 1
-    plt.rcParams['legend.frameon'] = False
-    plt.rcParams['legend.loc'] = 'center left'
-    plt.rcParams['axes.linewidth'] = 1
-    plt.rcParams['contour.negative_linestyle'] = 'solid'
-    plt.gca().spines['right'].set_color('none')
-    plt.gca().spines['top'].set_color('none')
-    plt.gca().xaxis.set_ticks_position('bottom')
-    plt.gca().yaxis.set_ticks_position('left')
-
-def twod_kde(x,y):
-    X, Y = np.mgrid[x.min()*0.9:x.max()*1.1:100j, y.min()*0.9:y.max()*1.1:100j]
-    positions = np.vstack([X.ravel(), Y.ravel()])
-    values = np.vstack([x, y])
-    kernel = gaussian_kde(values)
-    return X, Y, np.reshape(kernel(positions).T, X.shape)
-
-def FindHeightForLevel(inArr, adLevels):
-    """
-    compute the height of a two D function
[…]

### 2021-02-11 paper/scripts/plot_posterior_summary_new.py

@@ -1,200 +1,0 @@
-import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
-from scipy.stats import gaussian_kde
-import os
-from cosmolisa.likelihood import find_redshift
-import cosmolisa.cosmology as cs
-import h5py
-import matplotlib.font_manager
-
-def init_plotting():
-    plt.rcParams['figure.figsize'] = (3*3.4, 3.4)
-    plt.rcParams['font.size'] = 11
-    plt.rcParams['font.family'] = 'Times New Roman'
-    plt.rcParams['font.sans-serif'] = ['Bitstream Vera Sans']
-    plt.rcParams['axes.labelsize'] = plt.rcParams['font.size']
-    plt.rcParams['axes.titlesize'] = plt.rcParams['font.size']
-    plt.rcParams['legend.fontsize'] = 10
-    plt.rcParams['xtick.labelsize'] = 9
-    plt.rcParams['ytick.labelsize'] = 9
-    plt.rcParams['xtick.major.size'] = 3
-    plt.rcParams['xtick.minor.size'] = 3
-    plt.rcParams['xtick.major.width'] = 1
-    plt.rcParams['xtick.minor.width'] = 1
-    plt.rcParams['ytick.major.size'] = 3
-    plt.rcParams['ytick.minor.size'] = 3
-    plt.rcParams['ytick.major.width'] = 1
-    plt.rcParams['ytick.minor.width'] = 1
-    plt.rcParams['legend.frameon'] = False
-    plt.rcParams['legend.loc'] = 'center left'
-    plt.rcParams['axes.linewidth'] = 1
-    plt.rcParams['contour.negative_linestyle'] = 'solid'
-    plt.gca().spines['right'].set_color('none')
-    plt.gca().spines['top'].set_color('none')
-    plt.gca().xaxis.set_ticks_position('bottom')
-    plt.gca().yaxis.set_ticks_position('left')
-    plt.rcParams['text.usetex'] = True
-
-
-def twod_kde(x,y):
-    X, Y = np.mgrid[x.min()*0.9:x.max()*1.1:100j, y.min()*0.9:y.max()*1.1:100j]
-    positions = np.vstack([X.ravel(), Y.ravel()])
-    values = np.vstack([x, y])
-    kernel = gaussian_kde(values)
-    return X, Y, np.reshape(kernel(positions).T, X.shape)
-
[…]

### 2021-02-11 scripts/welch.py
Added welch script

@@ -1,0 +1,45 @@
+# -*- coding: utf-8 -*-
+from scipy.signal import tukey, welch
+
+def psd(strain,
+        sampling_rate,
+        segment_duration,
+        window_function  = None,
+        overlap_fraction = 0.5):
+    """
+    Computes the power spectral density of a time series using the Welch method
+   
+    :param strain: input time series
+    :type strain: array
+    :param sampling_rate: sampling rate of the input strain (Hz)
+    :type sampling_rate: float
+    :param segment_duration: length of each segment (in seconds) for the application of the Welch method
+    :type segment_duration: float
+    :param window_function: window type
+    :type window_function: string, tuple or array-like
+    :param overlap_fraction: overlap between segments (as a fraction of the segment duration)
+    :type overlap_fraction: float
+    
+    :return: frequencies, corresponding power spectral density
+    :rtype: array, array
+    """
+    
+    # if no window has been passed
+    # create a default Tukey window
+    if window_function is None:
+        padding = 0.4/segment_duration
+        window_function  = tukey(int(sampling_rate*segment_duration),padding)
+    
+    # number of samples per segments
+    N = int(segment_duration*sampling_rate)
+    
+    # compute the PSD and its frequencies
+    frequencies, psd = welch(strain,
+                             fs              = sampling_rate,
+                             window          = window_function,
+                             nperseg         = N,
+                             noverlap        = int(overlap_fraction*N),
+                             return_onesided = True,
+                             scaling         = 'density')
+                             
+    return frequencies, psd

### 2021-02-11 scripts/init_plotting.py
added plotting initialisation function

@@ -1,0 +1,35 @@
+import matplotlib.pyplot as plt
+
+def init_plotting():
+    """
+    Initialises matplotlib *rcParams*
+    
+    :return: no return
+    """
+    
+    # plotting options
+    plt.rcParams['figure.figsize']             = (3.4, 3.4)
+    plt.rcParams['font.size']                  = 11
+    plt.rcParams['font.family']                = 'DejaVu Sans'
+    plt.rcParams['font.sans-serif']            = ['Bitstream Vera Sans']
+    plt.rcParams['axes.labelsize']             = plt.rcParams['font.size']
+    plt.rcParams['axes.titlesize']             = plt.rcParams['font.size']
+    plt.rcParams['legend.fontsize']            = 10
+    plt.rcParams['xtick.labelsize']            = 9
+    plt.rcParams['ytick.labelsize']            = 9
+    plt.rcParams['xtick.major.size']           = 3
+    plt.rcParams['xtick.minor.size']           = 3
+    plt.rcParams['xtick.major.width']          = 1
+    plt.rcParams['xtick.minor.width']          = 1
+    plt.rcParams['ytick.major.size']           = 3
+    plt.rcParams['ytick.minor.size']           = 3
+    plt.rcParams['ytick.major.width']          = 1
+    plt.rcParams['ytick.minor.width']          = 1
+    plt.rcParams['legend.frameon']             = False
+    plt.rcParams['legend.loc']                 = 'center left'
+    plt.rcParams['axes.linewidth']             = 1
+    plt.rcParams['contour.negative_linestyle'] = 'solid'
+    plt.gca().spines['right'].set_color('none')
+    plt.gca().spines['top'].set_color('none')
+    plt.gca().xaxis.set_ticks_position('bottom')
+    plt.gca().yaxis.set_ticks_position('left')

### 2021-02-11 memspectrum.py
Bug fixed, N to self.N

@@ -423,7 +423,7 @@ class MESA(object):
         self.regularisation = regularisation
         self.early_stop = early_stop
         if m is None:
-            self.mmax = int(2*N/np.log(2.*N))
+            self.mmax = int(2*self.N/np.log(2.*self.N))
         else:
             self.mmax = m
            

### 2021-02-11 memspectrum.py
Set nan warnings

@@ -11,6 +11,7 @@ Algorithm for the computation of the power spectral density.
 import sys
 import numpy as np
 import warnings
+warnings.simplefilter('once', RuntimeWarning)
 from scipy.signal import correlate
 
 #############DEBUG LINE PROFILING
@@ -340,6 +341,7 @@ class MESA(object):
         """
         f_ny = .5 / dt 
         
+        self.N = len(self.a_k)
         spec, f_spec = self._spectrum(dt, self.N)
         
         if frequencies is None:
@@ -347,7 +349,7 @@ class MESA(object):
         
         elif isinstance(frequencies, np.ndarray):
             if np.max(frequencies) > f_ny *1.01: 
-                warnings.warn("Some of the required frequencies are higher than the Nyquist frequency ({} Hz): a zero PSD is returned for f>Nyquist".format(f_ny))
+                warnings.warn("Some of the required frequencies are higher than the Nyquist frequency ({} Hz): a zero PSD is returned for f>Nyquist".format(f_ny), UserWarning)
                 
             f_interp = np.interp(frequencies, f_spec[:int(self.N/2+0.5)], spec.real[:int(self.N/2+0.5)], left = 0., right = 0.)
         
@@ -355,6 +357,7 @@ class MESA(object):
         else:
             raise ValueError("Type of frequencies not understood: expected to be None or np.ndarray but given {} insted".format(type(frequencies)))
 
+        return
         
     def solve(self,
               data,
@@ -409,7 +412,6 @@ class MESA(object):
             (Shape (N,))   
 
         """
-        
         data = np.array(data)
         data = np.squeeze(data)
         
@@ -419,11 +421,11 @@ class MESA(object):
             raise ValueError("Type of data should np.ndarray: given {} instead. ".format(type(data)))
         
         self.data = data
-        self.N    = len(data)
+        self.N  = len(data)
         self.regularisation = regularisation
         self.early_stop = early_stop
         if m is None:
-            self.mmax = int(2*N/np.log(2.*N))
+            self.mmax = int(2*self.N/np.log(2.*self.N))
         else:
             self.mmax = m
            
@@ -489,7 +491,7 @@ class MESA(object):
         for i in range(self.mmax):
             #Update prediction error filter and reflection error coefficient
             k, new_a = self._updateCoefficients(a[-1], g)
-            #Update variables. Check paper for indeces at j-th loop.
[…]

### 2021-02-11 GenerateTimeSeries.py
Added comparison script

@@ -53,8 +53,7 @@ def generate_data(f,
     # f, psd = np.loadtxt(psd_file, unpack=True)
     if asd is True : psd *= psd
     # generate an interpolant for the PSD
-    #psd_int = interp1d(f, psd, bounds_error=False, fill_value= 'extrapolate')
-    psd_int = interp1d(f, psd, bounds_error=False, fill_value= 0.) #proposed change
+    psd_int = interp1d(f, psd, bounds_error=False, fill_value= 0.)
     df      = 1 / T
     N       = int(sampling_rate * T)
     times   = np.linspace(0, T, N) 

### 2021-02-11 scripts/compare_LVC_data.py

@@ -1,0 +1,97 @@
+import sys
+sys.path.insert(0,'..')
+from memspectrum import MESA
+from welch import *
+import numpy as np
+import matplotlib.pyplot as plt
+from init_plotting import init_plotting
+
+plot = True
+compute = True
+generate_fake_data = True
+fake_data = False
+
+	#folder to save the plot at
+save_folder = '../paper/Images/comparison_LVC_data/'
+true_PSD_file = 'GWTC1_GW150914_PSDs.dat'
+#save_folder = 'comparison_LVC_data' #if on the cluster
+srate = 4096.
+
+if generate_fake_data:
+	PSD = np.loadtxt(true_PSD_file)
+	freqs, PSD = PSD[:,0], PSD[:,1]
+	import GenerateTimeSeries
+	times, time_series, frequencies, frequency_series, psd_int =GenerateTimeSeries.generate_data(freqs, PSD, 32., srate, np.min(freqs), np.max(freqs))
+	np.savetxt('fake_data_4KHZ-32.txt', time_series)
+	
+if generate_fake_data:
+	data = np.loadtxt("fake_data_4KHZ-32.txt") #for fake data
+else:
+	#data = np.loadtxt("../examples/data/V-V1_GWOSC_4KHZ_R1-1186741846-32.txt")
+	data = np.loadtxt("../../GWAnomalyDetection/maxent/H-H1_GWOSC_16KHZ_R1-1126259447-32.txt.gz")
+
+
+T_list = [1,10,30] #list of times to make the comparison at
+
+if compute:
+	for i, T in enumerate(T_list):
+		data_T = data[: int(srate*T)]
+
+		freqs, PSD_Welch = psd(data_T, srate, T,
+			window_function  = None,
+			overlap_fraction = 0.5,
+			nfft = None,
+			return_onesided = True)
+		M = MESA()
+		M.solve(data_T, early_stop = True)
+		PSD_MESA = M.spectrum(1./srate, freqs)
+		
+		np.savetxt("plot_data/plot_{}_{}.txt".format(T, fake_data), np.column_stack([freqs, PSD_MESA, PSD_Welch]))
+		
+		#plt.loglog(freqs, PSD_Welch)
+		#plt.loglog(freqs, PSD_MESA)
+		#plt.show()
+	
+	
[…]

### 2021-02-11 scripts/init_plotting.py

@@ -29,7 +29,9 @@ def init_plotting():
     plt.rcParams['legend.loc']                 = 'center left'
     plt.rcParams['axes.linewidth']             = 1
     plt.rcParams['contour.negative_linestyle'] = 'solid'
+    fig = plt.figure()
     plt.gca().spines['right'].set_color('none')
     plt.gca().spines['top'].set_color('none')
     plt.gca().xaxis.set_ticks_position('bottom')
     plt.gca().yaxis.set_ticks_position('left')
+    return fig

### 2021-02-17 GenerateTimeSeries.py
edits to compare script

@@ -53,7 +53,7 @@ def generate_data(f,
     # f, psd = np.loadtxt(psd_file, unpack=True)
     if asd is True : psd *= psd
     # generate an interpolant for the PSD
-    psd_int = interp1d(f, psd, bounds_error=False, fill_value= 0.)
+    psd_int = interp1d(f, psd, bounds_error=False, fill_value='extrapolate')
     df      = 1 / T
     N       = int(sampling_rate * T)
     times   = np.linspace(0, T, N) 
@@ -73,7 +73,7 @@ def generate_data(f,
     frequency_series = sigma * (np.random.normal(0, 1, len(sigma)) + 1j * np.random.normal(0, 1, len(sigma)))
       
     # inverse FFT to return the TD strain
-    time_series = np.fft.irfft(frequency_series) * df * N 
+    time_series = np.fft.irfft(frequency_series, n=N) * df * N
     return times, time_series, frequencies, frequency_series, psd_int(frequencies)
 
 

### 2021-02-17 scripts/compare_LVC_data.py

@@ -6,11 +6,11 @@ import numpy as np
 import matplotlib.pyplot as plt
 from init_plotting import init_plotting
 import pandas as pd
-
+from scipy.interpolate import interp1d
 
 plot = True
-compute = False
-generate_fake_data = False
+compute = True
+generate_fake_data = True
 use_fake_data = True
 
 	#folder to save the plot at
@@ -19,14 +19,14 @@ true_PSD_file = 'GWTC1_GW150914_PSDs.dat'
 #save_folder = 'comparison_LVC_data' #if on the cluster
 srate = 4096.
 
-T_list = [1,5, 10,100, 1000] #list of times to make the comparison at
-seglen = [512, 1024, 2048, 8192, 32768]
+T_list = [1, 5, 10,100]#, 1000] #list of times to make the comparison at
+seglen = [512, 1024, 2048, 8192]#, 32768]
 
 if generate_fake_data:
 	PSD = np.loadtxt(true_PSD_file)
 	freqs, PSD = PSD[:,0], PSD[:,1]
 	import GenerateTimeSeries
-	times, time_series, frequencies, frequency_series, psd_int =GenerateTimeSeries.generate_data(freqs, PSD, 1002., srate, np.min(freqs), np.max(freqs))
+	times, time_series, frequencies, frequency_series, psd_int = GenerateTimeSeries.generate_data(freqs, PSD, 1002., srate)#, np.min(freqs), np.max(freqs))
 	np.savetxt('fake_data_4KHZ-1000.txt', time_series)
 	
 if compute:
@@ -40,20 +40,20 @@ if compute:
 		data = pd.read_csv("H-H1_GWOSC_4KHZ_R1-1247614487-4096.txt.gz", skiprows = 3).to_numpy()
 
 	data = np.squeeze(data)
-	print("Loaded data: shape {}; srate {}; length {}s".format(data.shape,srate, len(data)/srate))
+	print("Loaded data: shape {}; srate {}; length {}s".format(data.shape, srate, len(data)/srate))
 
 	for i, T in enumerate(T_list):
 		print("Batch length: {}s".format(T))
 		data_T = data[:int(srate*T)]
 		
 		M = MESA()
-		M.solve(data_T, early_stop = True, method = 'Standard')
[…]

### 2021-02-17 scripts/compare_LVC_data.py
Change comparison script

@@ -10,7 +10,7 @@ from scipy.interpolate import interp1d
 
 plot = True
 compute = True
-generate_fake_data = True
+generate_fake_data = False
 use_fake_data = True
 
 	#folder to save the plot at
@@ -20,7 +20,7 @@ true_PSD_file = 'GWTC1_GW150914_PSDs.dat'
 srate = 4096.
 
 T_list = [1, 5, 10,100]#, 1000] #list of times to make the comparison at
-seglen = [512, 1024, 2048, 8192]#, 32768]
+seglen = [512, 1024, 2048, 8192, 32768]
 
 if generate_fake_data:
 	PSD = np.loadtxt(true_PSD_file)
@@ -48,20 +48,16 @@ if compute:
 		
 		M = MESA()
 		M.solve(data_T, early_stop = True, method = 'Fast')
-
+		print("\tDone MESA")
 		freqs, PSD_Welch = psd(data_T, srate, seglen[i]/float(srate),
 			window_function  = None,
 			overlap_fraction = 0.5,
 			nfft = None,
 			return_onesided = False)
-		
+		print("\tDone Welch")
 		PSD_MESA = M.spectrum(1./srate, freqs)
 		
 		np.savetxt("plot_data/plot_{}_{}.txt".format(T, use_fake_data), np.column_stack([freqs, PSD_MESA, PSD_Welch]))
-		
-		#plt.loglog(freqs, PSD_Welch)
-		#plt.loglog(freqs, PSD_MESA)
-		#plt.show()
 	
 	
 if plot:
@@ -69,15 +65,17 @@ if plot:
 		true_PSD = np.loadtxt(true_PSD_file)
 	for i, T in enumerate(T_list):
 		PSDs = np.loadtxt("plot_data/plot_{}_{}.txt".format(T, use_fake_data))
-		freqs, PSD_MESA, PSD_Welch = PSDs[:,0], PSDs[:,1], PSDs[:,2]
+		N = PSDs.shape[0]
+		freqs, PSD_MESA, PSD_Welch = PSDs[:int(N/2),0], PSDs[:int(N/2),1], PSDs[:int(N/2),2]
 		
 		fig = init_plotting()
 		ax = fig.gca()
 		ax.set_title(r"$T = {}s$".format(T))
[…]

### 2021-02-17 GenerateTimeSeries.py
Super small fix in GenerateTimeSeries

@@ -15,10 +15,9 @@ def generate_data(f,
                   sampling_rate = 1.,
                   fmin = None,
                   fmax = None,
-                  zero_noise = False,
                   asd = False):
     """
-    Generate a time series with a given power spectraul density 
+    Generate a time series with a given power spectral density 
 
     Parameters
     ----------
@@ -77,5 +76,3 @@ def generate_data(f,
     return times, time_series, frequencies, frequency_series, psd_int(frequencies)
 
 
-
-    

### 2021-02-17 examples/mem_welch_comparison.py
New comparison between MESA and Welch

@@ -1,99 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Thu Feb  4 09:56:33 2021
-
-@author: alessandro
-"""
-
-from memspectrum import MESA
-from memspectrum import GenerateTimeSeries
-from scipy.signal import welch, tukey
-import scipy.stats 
-import numpy as np
-import matplotlib.pyplot as plt
-
-
-def welch_estimate(noise, N, dt, overlap_fraction = .5):
-    sampling_rate = 1 / dt
-    segment_duration = N * dt #lunghezza sottoarray
-    padding = 0.4 / segment_duration 
-    window_function  = tukey(int(sampling_rate*segment_duration), padding)
-    wf, ws = welch(noise, fs  = 1 / dt,
-                              window = window_function,
-                              nperseg         = N,
-                              noverlap        = int(overlap_fraction*N),
-                              return_onesided = False,
-                              scaling         = 'density',
-                              nfft = noise.size) #dopo N // 2 frequ negative
-    return wf, ws 
-
-
-def relative_error(real, estimate):
-    return np.abs(real - estimate) / real 
-
-
-if __name__ == '__main__': 
-    
-    #Generate the Time Series 
-    f_ny = 5 
-    dt = 1 / (2 * f_ny)
-    N = 81920 * 5 
-    T = N * dt 
-    frequency = np.arange(0, N // 2 + 1) / T 
-    bnoise_psd = 1 / (frequency ** 2 + 1)
-    normal_psd = scipy.stats.norm.pdf(frequency, 2.5, .5) 
-    number_of_segments = 10 #Number of segments for Welch method
-    
-    #Generate Time series 
-    time, normal_series, frequency, normal_frequency_series, normal_psd =\
-        GenerateTimeSeries.generate_data(frequency, normal_psd, T, 1 / dt)
-        
-    
-    time, bNoise_series, frequency, bnoise_frequency_series, bnoise_psd =\
-        GenerateTimeSeries.generate_data(frequency, bnoise_psd, T, 1 / dt)
-    
-    #Estimate the spectrum for normal PSD 
[…]

### 2021-02-17 scripts/mem_welch_comparison.py

@@ -1,0 +1,58 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Feb  4 09:56:33 2021
+
+@author: alessandro
+"""
+
+import sys
+sys.path.insert(0,'..')
+from memspectrum import MESA
+import welch
+import numpy as np
+import matplotlib.pyplot as plt
+from init_plotting import init_plotting
+from GenerateTimeSeries import generate_data
+from scipy.interpolate import interp1d 
+
+
+
+def compute_spectra(time_series, dt):
+    welch_frequency, welch_spectrum = [], []
+    MESA_frequency, MESA_spectrum = [], [] 
+    for i in range(len(time_series)): 
+        None 
+
+def relative_error(real, estimate):
+    return np.abs(real - estimate) / real 
+
+
+
+if __name__ == '__main__': 
+    #Sampling variables 
+    dt = 1. / 4096
+    times = np.array([1, 5, 10, 100, 1000])
+    N = times // dt 
+    segment_duration = np.array([521, 1024, 2048, 8192, 32768]) 
+    Ny = 0.5 / dt 
+    
+    #Generate normal PSD and import Ligo PSD 
+    wnoise_frequency = np.arange(0, N[0] // 2 + 1) / times[0]
+    wnoise_spectrum = np.repeat(1, wnoise_frequency.size) * dt 
+    ligo_frequency, ligo_spectrum = np.loadtxt('LIGO-P1200087-v18-AdV_DESIGN_psd.dat', unpack=True)     
+    ligo_interp = interp1d(ligo_frequency, ligo_spectrum)
+    #Lists containing all simulated arrays, frequencies and psds 
+    print('Generating white noise series')
+    Wtime, Wtime_series, Wfrequency, Wfrequency_series, Wpsd = generate_data(wnoise_frequency,
+                                                                        wnoise_spectrum,
+                                                                        times[-1],
+                                                                        1/ dt)
+    print('Generating Ligo noise series')
+    Ltime, Ltime_series, Lfrequency, Lfrequency_series, Lpsd = generate_data(ligo_frequency,
+                                                                             ligo_spectrum,
+                                                                             times[-1],
+                                                                             1 / dt)
+    
+    
+ 

### 2021-02-17 paper/mesa.tex
changes in paper

@@
 % comments command
-\newcommand{\an}[1]{{\textcolor{orange}{\texttt{AN: #1}} }}
-\newcommand{\pr}[1]{{\textcolor{cyan}{\texttt{PR: #1}} }}
-\newcommand{\fm}[1]{{\textcolor{blue}{\texttt{FM: #1}} }}
+\newcommand{\wdp}[1]{{\textcolor{green}{\texttt{SS: #1}} }}
+\newcommand{\amartini}[1]{{\textcolor{blue}{\texttt{SS: #1}} }}
 \newcommand{\sschmidt}[1]{{\textcolor{red}{\texttt{SS: #1}} }}
-\newcommand{\BS}[1]{{\textcolor{green}{\texttt{SB: #1}} }}
-\newcommand{\oldnewtxt}[2]{\sout{#1}\textcolor{red}{#2}}
 
@@
 Burg method of Maximum Entropy Spectral Analysis (MESA) provides a powerful tool to perform spectral estimation of a time-series.
-The method is a generalization of Jaynes maximum entropy principle and provides the estimate for the spectrum in terms of the coefficients of some autoregressive order p process (AR(p)).
-The recursive solution requires a method for the estimate of the order $p$ for the process.
-Several methods are characterized on synthetic noise having a known power spectral density (PSD) via the computation of the percentage error and then compared.
+The method is a generalization of Jaynes maximum entropy principle and provides the estimate for the spectrum in terms of the coefficients of some autoregressive process AR(p) or order $p$.
+A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
+We provide a ready to use implementation of the algorithm in the form of a python package `memspectrum`.
+We characterize our code performing a power spectral density analysis on some syntethic generated data (with known power spectral density) and we compare different criteria for stopping the recursion.
+Furthermore, we compare the performance of our code with the traditional Welch algorithm, using both syntethic data and real gravitational waves data from the LIGO-Virgo collaboration.
+We find that, when compared to the Welch's method, the Burg's method provides a power spectral density (PSD) estimation with a systematically lower variance.
+This is particularly manifest in the case of a little number of datapoints, making Burg method most suitable to work in this regime.
+%Several methods are characterized on synthetic noise having a known power spectral density (PSD) via the computation of the percentage error and then compared.
 
@@
         \affiliation{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}        
-        %
+	\author{Stefano \surname{Schmidt}}
+		\email{s.schmidt@uu.nl}
+        \affiliation{add me}      
+	\author{Walter \surname{Del Pozzo}}
+		\email{walter.delpozzo@unipi.it}
+        \affiliation{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}
 
@@
 \section{Introduction}
-\blindtext
+
+The study of the properties of a noisy process is a crucial task in many fields of physics and mathematical physics [Here we should make some examples...].
+A particularly interesting class of stochastic processes are the so-called \textit{wide-sense} stationary processes.
+These are stochastic processes that display an invariance of their statistical properties, such as their two-point autocovariance function, with respect to time translation.
+If $x(t)$ is a wide-sense stationary process, it is completely determined by the knowledge of the auto-covariance function
+\begin{equation}
+	R(\tau) = \mathbf{E}[x_t \cdot x_{t+\tau}]
+\end{equation}
+or equivalently by the knowledge of their \emph{power spectral density} (PSD) $S(f)$. Thanks to the Wiener-Khinchin theorem the two are related by a Fourier transform: 
+\begin{equation}
+	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau R(\tau) e^{-i 2 \pi f \tau}\,.
+\end{equation}
+In some literature, the PSD is introduced as  
+\begin{equation}\label{eq:psd-f-definition}
+	\mathbf{E}[\tilde{x}(f) \cdot \tilde{x}(f^\prime)] = S(f) \delta(f-f^\prime)
+\end{equation}
+without highlighting its connection with the time structure of the process itself, thus masking some important properties that will explore further in what follows.
+The latter definition in (\ref{eq:psd-f-definition}) gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating for an unknown process.
+%Under such assumption, it can be shown [cite something] that the problem of characterizing the noise properties is greatly simplified and its properties are summarized by a single \textit{autocorrelation function}, which describes how an observation at a given time correlates with another observation a later time.
+More formally, the properties of a zero mean time series $x_t$ are described by the function:
+%\begin{equation}
+%	R(\tau) = \mathbf{E}[x_t \cdot x_{t+\tau}]
+%\end{equation}
+%where the average is computed on different times $t$.
+%Thus, a stationary noise time series is completely characterized by the autocorellation function.
+%\par
+%More often, the study is performed in frequency domain: indeed, for stationary noise the components in frequency $\tilde{x}(f)$ of a stationary time series are conditionally independent from each other.
+It can be shown that:
+%\begin{equation}
+%	\mathbf{E}[\tilde{x}(f) \cdot \tilde{x}(f^\prime)] = S(f) \delta(f-f^\prime)
+%\end{equation}
+%This equation defines the function \textit{power spectral density} (PSD) $S(f)$.
+%The PSD has a straightforward interpretation: it measures how much signal ``power" is located in each frequency bin.
+Furthermore, it is not difficult to show that it is the Fourier transform of the autocorrelation function:
+%\begin{equation}
+%	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau R(\tau) e^{-i 2 \pi f \tau}
+%\end{equation}
+%A power spectral density fully describes a stationary noisy time series and this shows how important is in the study of noise.
+%\par
+%Thus, in many real application, the knowledge of the PSD is crucial for the understanding of the process and indeed algorithms for PSD computation are a standard part of every software package for computational physics.
+An ubiquitous method for such computation is due to Welch \cite{Welch} and it based on the method of periodograms.
+Basically, the PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many windowed batches and averaging the squared moduli of their Fourier transforms.
+The standard approach of periodograms \cite{Lomb,Scargle} consists in taking the Fourier Transform of the sample autocorrelation, written as \begin{equation}
+    \rho = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\}.
+\end{equation}
+The sequence $W$ is a window function that can be chosen in several different way, each choice presenting advantages and disadvantages for the final estimate of the PSD. \\ 
+The choice of a window function is arbitrary and typically is made by trial and error, until a satisfactory compromise between variance and resolution of the estimate of PSD is reached.
+A high frequency resolution implies high variance and vice-versa.
+Moreover, the final result depends on the choice of window function, hence begging the question of what the ``real" PSD is.
+\\
+Another problem with this approach is the requirement for the window to be $0$ outside the interval: we are arbitrarily assuming $\rho_j = 0$ for $j > M$, and modifying the estimate if a non-rectangular window is chosen.
+Making assumptions on unobserved data and modifying the ones we have at our disposal introduces information about the process that is not really there.
+Moreover, this method requires a number of arbitrary choices to be made, e.g.
+the number of time slices, the kind of Window function, the overlap between consecutive slices, etc...
+All the knobs must be tuned by hand and their choice can dramatically affect the PSD estimation.
+\par
+An appealing alternative approach, based on the Maximum Entropy principle \cite{Jaynes}, has been put forward by Burg \cite{burg1975maximum}.
+Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the arbitrary parameters, as it provides an iterative closed form expression for the spectrum of a stochastic time series.
+Furthermore, it embeds the PSD estimation problem into an elegant theoretical framework and makes minimal assumptions on the nature of the data.
+Lastly and most importantly, it provides a robust link between spectral density estimation and the field of autoregressive processes.
+This provide a natural and simple machinery to forecast a time series, thus predicting past observations based on previous ones.
+In this paper we discuss the details of the Maximum entropy principle, its application to Power Spectral Density estimation with Burg algorithm and the link between Burg's algorithm and autoregressive process.
+Furthermore, we present and describe a new code, `memspectrum`, that provides a robust python implementation of the algorithm.
+We provide a thorough assessment of the performance of our code and we validate our results performing a number of tests.
+We compare our results with those of spectral analysis carried out with the standard Welch method.
+\par
+Our paper is organized as follows: blablabla...
 
 \section{Theoretical foundations}
-\subsection{Maximum Entropy Principle} 
-The Maximum Entropy principle (MAXENT) is one of the most important results in Bayesian Statistics.
+The Maximum Entropy principle (MAXENT) is among the most important results in Bayesian Statistics.
 It provides a way to uniquely assign probabilities to a phenomenon in a way that best represent our state of knowledge, while at the same time it is non committal with unavailable information.
@@
 In fact, thanks to Burg \cite{burg1975maximum}, this method has also been applied to perform high quality computation of power spectral densities of time series.
-After a short introduction to Jayne's MAXENT, we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form.
-To introduce MAXENT as a variational problem, we will workout some simple examples to understand how information's entropy and 'uncertainty' are related. \\ 
-The word 'information' will not be used to indicate our knowledge on the system under study: this is what we will call 'evidence'.
+
+After a short introduction to Jayne's MAXENT (sec.~\ref{sec:MAXENT}), we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form (sec.~\ref{sec:MESA}).
+Next, we will discuss an interesting link between Burg's method and autoregressive processes (sec.~\ref{sec:autoregr}) and in sec.~\ref{sec:forecasting} we will use such link to introduce a method for forecasting a time series.
+
+\subsection{Maximum Entropy Principle} \label{sec:MAXENT}
+\sschmidt{This section is yet to prune}
+To introduce MAXENT as a variational problem, we will workout some simple examples to understand how information's entropy and `uncertainty' are related\footnote{
+The word `information' will not be used to indicate our knowledge on the system under study: this is what we will call `evidence'.
 Information will be intended as in communication theory: it specifies the quantity of details needed to provide a full description of the system under study.
-\\
-If we consider a perfectly sinusoidal signal, knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information entropy content . \\ 
+}.
+
+If we consider a perfectly sinusoidal signal, knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information entropy content.
 Even less information is required if we are studying a system whose outcome is certain (has probability $p = 1$).
-If the quantity of information can be considered to be $I = 0$, a communication is not even needed.  
+If the amount of information $I$ can be considered to zero, a communication is not even needed.  
 Shannon \cite{Shannon} proposed the quantity
 \begin{equation}\label{eq:information}
-    I_i = \log_2 \frac{1}{p(x_i)}
-\end{equation}
-to represent the quantity of information brought by an outcome $x_i$ with probability $p_i$.
+    I = \log_2 \frac{1}{p(x)}
+\end{equation}
+to represent the quantity of information brought by an outcome $x$ with probability $p(x)$.
 It has the nice property of being an additive quantity and it is monotonically decreasing as a function of $p \in [0, 1]$.
 The more uncertain the outcome, the higher the information it brings.
-\\
-What happens if we consider not one outcome, but a larger set of different events?
-What is  the 'most uncertain' outcome?
+\par
+What happens if we consider a larger set of different events?
+What is  the `most uncertain' outcome?
 Since we are intending probability as a representation of the state of knowledge, one of the first requirement for our assignment is the agreement with common sense.
-\\
-Consider an experiment based on the occurrence of one of two different outcomes $E_1, E_2$ with given probabilities $P_1$ and $P_2$. What are the values that make our outcome the most uncertain? \\ 
+Consider an experiment based on the occurrence of one of two different outcomes $E_1, E_2$ with given probabilities $P_1$ and $P_2$. What are the values that make our outcome the most uncertain?
 If $P_1$ and $P_2$ are largely different, like $P_1 = 0.999$ and $P_2 = 0.001$, we have enough evidence to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome.
-\\
 The most unpredictable experiment is the one in which 
@@
 \end{equation}
-this describes a situation of 'maximum ignorance'; we have no evidence at all and we have no instruments to predict what the outcome of such an experiment will be.  \\ 
+this describes a situation of `maximum ignorance'; we have no evidence at all and we have no instruments to predict what the outcome of such an experiment will be.  \\ 
 This can be generalized in the case of $N$ events. Assigning same probability
@@
 It is clear from the point of view of communication theory: the higher the evidence on a system, the lower the information I need to send to describe it.
-\\
+\par
 It is now clear that any assignment different from the 'most uncertain' is violating the hypothesis for it to being non committal to non-available evidence.
@@
 $H$ has to be a 'generalization' of information defined in equation (\ref{eq:information}) to be something like the 'expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
-\\
+
 Some properties for $H$ can be worked out easily. 
-From previous discussion, given a fixed number $N$ of possible outcomes, we want $H$ to be maximized by the choice $p_1 = \dots = p_N$. \\ 
+From previous discussion, given a fixed number $N$ of possible outcomes, we want $H$ to be maximized by the choice $p_1 = \dots = p_N$.
 Consider now the opposite situation, in which probability is assigned appealing to IP and we have a different number of possible outcomes.
@@
 \end{equation}
-to increase monotonically with N. \\ 
+to increase monotonically with N.
+
 We also want our function to be continuous in its parameters: any small change in probability has to slightly affect uncertainty. \\
@@
 \end{equation}
-so that uncertainty is additive. \\ 
+so that uncertainty is additive.
+
 Shannon \cite{Shannon} showed that the only functional form satisfying our assumptions is given by:
@@
 and $H$ is what we will call entropy.
-This result is very interesting, if we compare it with Eq.~(\ref{eq:information}), we see that the entropy is defined as the 'average information'.
+
+This interpretation is straigthforward: if we compare it with Eq.~(\ref{eq:information}), we see that the entropy is defined as the 'average information'.
 The probability assignments made maximizing it with respect to $p$, is somehow requiring  the 'expected value' of uncertainty to be a maximum.
-Such an assignment is in agreement with our hypothesis of being 'non committal' with unavailable evidence, since \ref{eq:entropy} is all we need to assign probabilities uniquely.
+Such an assignment is in agreement with our hypothesis of being `non committal' with unavailable evidence, since \ref{eq:entropy} is all we need to assign probabilities uniquely.
 Evidence can be added in form of constraints in the maximization.
-\\ It is easy to verify that this expression for entropy verifies the aforementioned properties.
-Assigning the same value for each probability $p_i$, it is straightforward to verify $H$ increases with N: 
-\begin{equation}
-    -\sum_{i = 1}^N \frac{1}{N}\log{\frac{1}{N}} = - \log\frac{1}{N} = \log N.
-\end{equation}
-We want to show that, in case in which our only constraint is exhaustiveness \begin{equation}\nonumber 
-    \sum_{i = 1}^{N}p_i = 1,
-\end{equation}
+
+%It is easy to verify that this expression for entropy verifies the aforementioned properties. 
+%Assigning the same value for each probability $p_i$, it is straightforward to verify $H$ increases with N: 
+%\begin{equation}
+%    -\sum_{i = 1}^N \frac{1}{N}\log{\frac{1}{N}} = - \log\frac{1}{N} = \log N.
+%\end{equation}
+
+To show how the maximization of entropy, supplemented with additional evidence, can provide a unique probability assignment, we will work out three examples.
+In each of them, we will add more evindence (i.e. constraints on the optimization problem) and we will obtain different results.
+
+First of all, we want to show that, in case in which our only constraint is exhaustiveness
+$\sum_{i = 1}^{N}p_i = 1 $,
 MAXENT reduces to indifference principle as it should be. The variational problem is defined as
@@
 \end{equation}
-MAXENT reduces to the IP when only the normalization constraint is used. What happens if more than one constraint is used? \\
-For simplicity, we will work through the next examples in the continuous case. Suppose the mean of the distribution is known $\int xP(x) dx = \bar{x}$. The variational problem can be rewritten as: 
+MAXENT reduces to the IP when only the normalization constraint is used.
+
+In the second example, we will add the constraint that the distribution has a known mean value: $\int \d{x} x p(x) = \bar{x}$. For simplicity we will work in the continuous case.
+The variational problem can be rewritten as: 
 \begin{equation}
@@
 \end{equation}
-and its solution in term of $p_i$ is: 
+with constraints: 
+\begin{equation*}
+    \int \d{x} \; p(x)  = 1 \text{  and  }
+    \int \d{x} \; x p(x) = \bar x.
+\end{equation*}
+For simplicity, we consider $x$ to be a variable ranging in the interval $(-\infty,+\infty)$, but this can be easily computed for every other interval.
+
+The general solution for $p(x)$ is:
 \begin{equation}
@@
 \end{equation}
-and constraints: 
-\begin{equation} \nonumber
-    \int p(x) dx  = 1 \text { and }
-    \int x p(x) dx = \bar x.
-\end{equation}
-For simplicity, we consider $x$ to be a variable ranging in the interval $(-\infty,+\infty)$, but this can be easily computed for every other interval. The normalization conditions implies 
-\begin{equation}\nonumber 
-   1 =  \int p(x \vert I) dx = e^{-1-\lambda}\int e^{-\mu x} dx, 
+where $\lambda$ and $\mu$ are two Lagrange multipliers
+The normalization conditions implies 
+\begin{equation}\nonumber 
+   1 =  \int \d{x} \; p(x) = e^{-1-\lambda}\int e^{-\mu x} dx, 
 \end{equation}
@@
 \end{equation}
-In the continuous case MAXENT assigns for $x$ an exponential distribution
-\begin{equation}
-    p(x \vert I) = \mu e^{-\mu x}. \nonumber 
-\end{equation}
-It is clear from the functional form that the Lagrange multiplier has to be $\mu = \frac{1}{\bar x}$ to satisfy the second constraint. \\ 
-The most interesting and useful result of MAXENT is obtained when a constraint on the variance is also considered. Take a vector  $\vec x(t)$ with components in $\mathbb{R}^N$
-\begin{equation*}
-    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
-\end{equation*}
-We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals.
+The fixed average condition provides a value for $\mu$ in terms of $\bar{x}$:
+\begin{equation}
+	\mu = \frac{1}{\bar{x}}
+\end{equation}
+The solution is then an exponential distribution:
+\begin{equation}
+    p(x) = \frac{1}{\bar{x}} \exp\left(-\frac{x}{\bar{x}}\right). \nonumber 
+\end{equation}
+
+Lastly, and more interestingly, we also consider a constraint on the variance. For this reason, we set:
+\begin{align*}
+    \int \d{x} \; p(x)  = 1 \text{  and  }
+	\int \d{x} \; x p(x) = \bar{x} \\
+	\text{  and  }
+	\int \d{x} \; (x-\bar{x})^2 p(x)  = \sigma^2
+\end{align*}
+Performing the same computation as above, we obtain a well known result:
+\begin{equation}
+	p(x) = \frac{1}{\sigma\sqrt{2\pi}} e^{-\frac{1}{2}\left(\frac{x-\bar{x}}{\sigma}\right)^2}
+\end{equation}
+
+%Take a vector  $\vec x(t)$ with components in $\mathbb{R}^N$
+%\begin{equation*}
+%    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
+%\end{equation*}
+%We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals.
 If we put constraints in both mean $\vec \mu(t)$ and covariance matrix $C_(t_i, t_j)$, MAXENT allow us to assign for $\vec x(t)$ a multivariate normal distribution \cite{gregory_2005}:
-\begin{equation}
-    p\left(\vec x(t)\vert I\right) = 
-    \frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right). 
-\end{equation}
-With MAXENT, the often unjustified assumptions for the data to be normally distributed finds an elegant explanation.
-It is not only the central limit distribution for finite variance processes, it is also the distribution that maximizes the entropy and, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process.
-In some sense, we can interpret the central limit theorem as the natural 'statistical' evolution toward a configuration that maximizes entropy.
-
-\subsection{Maximum Entropy Spectral Analysis} 
-The computation of the power spectral density is an important component in time series analysis and characterization.
-Also, if the process is at least wide sense stationary, PSD is related to the autocorrelation function via a Fourier Transform.
-Since we are dealing with a finite number of samples $N$, it is not possible to have an accurate computation of the autocorrelation function for every time interval.
-Also, the error $\sigma$ in the estimate increases with the time interval $k$ as $\sigma \sim \frac{1}{\sqrt{N - k}}$, so that only a few values for the autocorrelation function can actually be computed.
-\\
-The problem we want to address is how to compute a PSD from the partial knowledge of the autocorrelation function. \\
-The standard approach of peridiograms \cite{Lomb} \cite{Scargle} consists in taking the Fourier Transform of the sample autocorrelation, written as \begin{equation}
-    \rho = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\}.
-\end{equation}
-The sequence $W$ is a window function that can be chosen in several different way, each method having pros and cons for the final estimate of the power spectral density. \\ 
-The choice of a window function is arbitrary and made by trial and error, until a satisfactory compromise between variance and resolution of the estimate of power spectral density  is reached.
-A high frequency resolution implies high variance and vice-versa.
-Another problem is that the result is very unstable with respect to the window's choice, and this fact may raise the question of which we should consider as the "real" power spectral density.
-\\
-Another problem with this approach is the requirement for the window to be $0$ outside the interval: we are arbitrarily assuming $\rho_j = 0$ for $j > M$, and modifying the estimate if a non-rectangular window is chosen.
-Making assumptions on unavailable data and modifying the ones we have at our disposal does not seem a completely fair approach.
-\\
-MAXENT gives us the possibility to develop a completely new way for the estimation of PSDs, without any a priori assumptions on the unavailable data.
-Our requirements (the constraints for the variational problem) are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation.
-\\
-To do so, there is a problem we have to solve: our definition of entropy (equation \ref{eq:entropy}) depends on the probability distribution, not the power spectrum.
-We need to find a way to change the definition's domain for the entropy to formulate the variational problem in terms of the power spectral density $S(f)$ alone.
-A way out of this problem is considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA}.
+%\begin{align}
+%    p&\left(\vec x(t)\vert I\right) = \nonumber \\
+%    &\frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right). 
+%\end{align}
+Thus MAXENT provide a deeper insight into the Gaussian distribution.
+It is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy.
+For this reason, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process.
+In some sense, we can interpret the central limit theorem as the natural `statistical' evolution toward a configuration that maximizes entropy.
+
+\subsection{Maximum Entropy Spectral Analysis} \label{sec:MESA}
+The computation of the power spectral density is an important component in stationary time series analysis and characterization.
+In principle, if the autocorrelation was known exactly (i.e. at every time $\tau$), the computation of the PSD would be just a single Fourier transform.
+However, in every realistic setting, we are dealing with a finite number of samples $N$ and it is not possible such precise computation.
+Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases with $k$ as $\sigma \sim \frac{1}{\sqrt{N - k}}$, so that only a few values for the autocorrelation function can actually be computed reliably.
+This bring us the core of our problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function?
+The Maximum Entropy Principle can guide use in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method.
+The latter assumes that the unknown values of the autocorrelation are $0$.
+Clearly, this assumption is unjustified and MAXENT is a good way to drop this assumption.}.
+\par
+As in the previous examples, we need to set up a variational problem where the entropy eq.~\ref{eq:entropy} is maximized with some problem-specific constraints. 
+In our case, they are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation (wherever an estimate of this is available).
+\par
+Before doing so, there is a technical problem to solve: our definition of entropy (eq.~\ref{eq:entropy}) depends on a probability distribution, not on the power spectrum.
+We then need to to formulate the variational problem in terms of the power spectral density $S(f)$ alone.
+A way out of this is considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA}.
 One can show that the 'entropy gain' obtained filtering white noise is:
@@
 \end{equation}
-so that we can reformulate the variational problem requiring $S(f)$ to maximize the entropy gain.
+so that we can reformulate the variational problem requiring $S(f)$ to maximize the entropy gain \sschmidt{Non ho capito questa parte...:(}.
 In real examples we have no access to all frequencies but only on those between minus and plus the Nyquist frequency, defined from the sampling interval $\Delta t$ as
@@
 \end{equation}
-Defining $\bar r_n$ as the sample autocorrelation, our constraints will be
+
+Now the the objective to maximize is well defined, we can write down the constraints of our problem.
+Defining $\bar r_n$ as the sample autocorrelation, we require that the variational solution $S(f)$ for the PSD matches the empirical autocorrelation. This means:
 \begin{equation}\label{eq:MaxConstraint}
@@
 It requires no assumptions on the unavailable estimates for $\bar r_n$ and is consistent with the sample autocorrelation function by construction.
-The hard task is now to show that this problem has a solution.
-Remarkably, it admits an analytical solution in a closed expression for $S(f)$
+Remarkably, it admits an analytical solution in a closed expression for $S(f)$.
+
 \begin{equation}\label{eq:MESApsd}
@@
 \end{equation}
-with $\Delta t$ the sampling interval $a_0$ = 1 and the coefficients $P_N$ and $a_i (i \neq 0)$ to be determined. The explicit solution can be found in Appendix REF
-
-\subsection{Autoregressive Process Analogy} 
-The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the analysis
+with $\Delta t$ the sampling interval $a_0$ = 1 and the coefficients $P_N$ and $a_i (i \neq 0)$ to be determined.
+The details of the derivation and the actual values of the coefficients $a_i$ can be found in appendix \ref{sec:MESA_solution}.
+
+\subsection{Autoregressive Process Analogy} \label{sec:autoregr}
+The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the study
 of autoregressive processes.\cite{doi:10.1029/RG013i001p00183}.
-Consider an autoregressive stationary process of order $p$ (AR(p)): 
-\begin{equation*}
-    x_t - b_1 x_{t-1} - b_2 x_{t-2} \dots b_p x_{t - p} = \nu_t, 
-\end{equation*}
-taking the z transform of the previous: 
+An autoregressive stationary process of order $p$ (AR(p)) is a time series whose values satisfy the following expression: 
+\begin{equation} \label{eq:AR_p}
+    x_t - b_1 x_{t-1} - b_2 x_{t-2} \dots b_p x_{t - p} = \nu_t
+\end{equation}
+where $b_1, \ldots, b_p$ are real coefficients and $\nu_t$ is white noise with a given variance $\sigma^2$.
+Thus, an AR(p) process models the dependence on every point on the $p$ past observation, thus being potentially able to model complex autocorrelation between data.
+\par
+To show the analogy, we compute the PSD of such process and we show that it is formally equivalent to the PSD obtained in eq.~\ref{eq:MESApsd}.
+We start taking the z \sschmidt{What is the z transform} transform of the previous eq.~\ref{eq:AR_p}: 
 \begin{align}
-    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t
+    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t.
 \end{align}
-and calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
+Calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
 in the $z$ domain, the process takes the form:
@@
 \end{equation}
-Since we assumed stationarity for the process, $\tilde{x}(z)$ is analytic both on and inside the unit circle.
+Since we assumed a stationary process, $\tilde{x}(z)$ is analytic both on and inside the unit circle.
 Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
@@
 \end{equation}
-The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$. \\ 
-Comparing (\ref{eq:ARspectrum}) with (\ref{eq:MESApsd}), identifying $b_1 = - a_i$ and $P_N \Delta t= \sigma ^ 2$, the expression for the power spectral density obtained with MESA is formally equivalent to the power spectral density of some stationary autoregressive process.
-\\
+The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.
+Equation~\ref{eq:ARspectrum} and eq.~\ref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
+This shows that the MAXENT estimation of the PSD computes the coefficients of the autoregressive process that is able to model the data faithfully.
+
+\sschmidt{I don't understand the next of this section...}
 Having computed the prediction error filter one can easily obtain an estimate of the $AR$ coefficients for the process just changing sign to the prediction error filter $a_i$.
@@
 
-\subsection{Forecasting}
+\subsection{Forecasting} \label{sec:forecasting}
+The link between MESA and autoregressive process allow us to forecast from a given time series, thus providing plausible future observation for the data.
+Indeed, an autoregressive process of order $p$ models the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation:
+\begin{align}\label{eq:p_forecast}
+	p&(x_t|x_{t-1}, \ldots , x_{t-p}) \nonumber\\
+	&= \frac{1}{\sigma\sqrt{2\pi}} \exp\left[-\frac{1}{2} \left(\frac{x_t - \sum_{i = 1}^p b_i x_{t-i}}{\sigma}\right)^2\right].
+\end{align}
+The interpretation of eq.~\ref{eq:p_forecast} is straightforward: $x_t$ follows a Guassian distribution with a fixed variance an a mean value $m_t = \sum_{i = 1}^p b_i x_{t-i}$ computed from past observations.
+Equation~\ref{eq:p_forecast} provides then a well defined probability framework for predicting future observations: this is a very useful feature of the method, which does not have an equivalent in any other spectral analysis computation.
 
@@
 The following chapter will be devoted to the study of the statistical properties of the previous optimizers to understand which choice provides the best quality in the reproduction of some known power spectral densities.
-
+\section{Analysis of the optimizers}
+\section{Analysis of optimizers - Introduction}
+To apply Burg's method, it is a fundamental task to understand what is the best choice for estimating the order for the autoregressive stochastic process.
+Before applying it to real datasets, we will study main properties of the three different optimizer, introduce in section \ref{sec:optimizers} on distinct, simulated, realizations of some AR(p) processes.
+We will generate different Gauss noise time series with a given known power spectral density (PSD) to understand how each method performs in different situation.
+The analysis will concern the quality, to be defined later, of the reconstruction of the spectrum itself.
+We will consider two different kinds of spectra: i) a Gaussian spectrum with a given mean and variance; ii) an Advanced LIGO design sensitivity theoretical spectral curve.
+As a part of our investigation, we will consider both ensemble-averaged, indicated with $\bar S$ and non averaged quantities.
+\\
+In either cases, we will quantify the agreement between the reconstructed quantities and the expected ones.\\
+A short description of the principle of the algorithm used to construct the time series is also given.
+
+\section{Gaussian noise} 
+This section is devoted to the analysis of the optimizations methods, applied on pure Gaussian noise process $\vec n = (n_1, n_2, \dots, n_N),\text{ with } n_i = n(t_i)$, assuming our process to be a wide-sense stationary process, i.e.
+autocorrelation matrix is a Toeplitz matrix depending only on time intervals.\\
+The spectrum chosen for our first analysis is a Gaussian distribution
+\begin{equation}
+    \nonumber
+    S(f) = \frac{1}{\sqrt{2 \pi} \sigma}e^{-\frac{(f - \mu)^2}{2 \sigma ^2}}
+\end{equation}
+with mean $\mu =2.5 \quad a.u.$ and standard deviation $\sigma =  0.5 \quad  a.u.$. 
+To generate noise with a power spectrum that is known, we have to first understand what is the role played by the PSD. \\ 
+Via MAXENT principle, we can state that noise in normally distributed with 0 mean and some covariance matrix $C_{ij} = C(t_i, t_j)$, i.e.
+\begin{equation}
+    \nonumber
+    p(\vec{n}(t) \vert I) \propto \exp{\left\{-\frac{1}{2}n_i C^{-1}_{ij}n_j\right\}},
+\end{equation}
+\begin{figure}
+    \centering
+        \includegraphics[width = \linewidth]{Images/Noise and PSD/NormalPSDautocorr.pdf}
+        \caption{A priori spectrum (left) and autocorrelation (right)}
+        \label{fig:autocorr}
+\end{figure}and, since process has zero mean, autocovariance and autocorrelation matrices are identical. \\ 
+Using the properties of Toeplitz matrices (see appendix \ref{app:Toeplitz}) and the relation between autocorrelation and power spectral density, it is immediate to verify that, in the limit of large number of time samples, the distribution of noise in frequency domain is given by:
+\begin{equation}
+    p(\tilde{n}(f_i) \vert I) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}},
+\end{equation}
+so that the role of PSD appears to be clear: it is the diagonal covariance matrix for the noise distribution in Fourier Domain: \\
+\begin{equation}
+    \nonumber
+    <\tilde n_i \tilde n_j > = S(f_i)\delta_{ij}.
+\end{equation}
+With this information, it is easy to generate any time series with power spectral density of any shape.
+Choosing a priori values for sampling interval $dt$ and total time of observation $T$, recalling Nyquist theory of sampling, we generate arrays of given length N / 2 in frequency domain from 0 to the Nyquist Frequency $f_{Ny} = \frac{1}{2 dt}$.
+We then convert noise to the time domain using an Inverse Fast Fourier Transform.\cite{oliphant2006guide}\cite{van2011numpy} \\ Basic idea for fast algorithm used to construct noise can be found in ~\cite{noiseGen}.
+We worked it out in a slightly different way, the imaginary component is not obtained sampling a random complex phase from a uniform distribution.
+We choose to sample both real and imaginary part from two identical normal distribution distributing variance in a proper way between real and imaginary components.
+\section{Test on Time Series}
+It is now possible to test the three methods on a given time series.
+We will concentrate on a power spectral density that is a normal distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$.
+We will apply Burg method with each optimizer on a large amount of simulated dataset, obtaining a variety of estimate for PSD.
+We will study statistical properties for both single reproductions and ensemble average.
+In many cases, such as gravitational waves, one cannot observe multiple time series realizations for a fixed process, so that single reproduction studies are probably more useful.
+Still, study of ensemble average is important to highlight some interesting properties.
+In this way we can fully characterize the performance of each optimizer
+We will attentively study the role of the filter length estimate M in determining the accuracy of the reconstruction.
+We will consider every optimizer separately and compare result only at the end of this section.
+We will then study in depth the Advanced LIGO design sensitivity power spectrum \cite{Ligo}
+\subsection{Normal Power Spectral Density}
+\begin{figure}
+    \centering
+        \includegraphics[width = \linewidth]{Images/NormalPSD/NormalNoise.pdf}
+    \caption{Noise in both time and frequency domain. Frequency noise behaviour is representative of the chosen spectrum, having its maximum amplitude around $f = 2.5 au$}
+    \label{fig:noise}
+\end{figure}
+Fig \ref{fig:noise} shows a realization of the gaussian process in the frequency (left panel) and time domain (right panel).   \\ 
+To study noise with normal-shaped PSD, we generate 1000 datasets of 3000 points each.
+This will provide a number of samples that is large enough to perform statistical analysis.
+For every dataset we will reproduce its spectrum with every optimization method.
+We will show that the quality of reconstruction is very different in different parts of the spectrum: the minimum of the error will be achieved around the peaks, increasing fastly in the tails
+ \\ \\ 
+\subsubsection{Final Prediction Error (FPE)}
+The estimate of the recursive order obtained with FPE on the 1000 realization of noise, is found to lay between 30 and 70.
+Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum: a logarithmic plot of the mean of all the spectra versus the true spectrum is reported in \textbf{Figure \ref{fig:FPEmean}} with 90\% credible regions.
+It is evident that, despite not too large error bars, the tail of the normal distribution is not comprised in the interval of the estimation.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width = \linewidth]{Images/NormalPSD/FPEpsdAndResiduals.pdf}
+    \caption{Ensemble Average for FPE reconstruction with 90\% Credibility Regions (above) and percentage error (below)}
+    \label{fig:FPEmean}
+\end{figure}
+To assess the magnitude of the disagreement, we compute the relative error between the mean spectrum $\hat{S}$ and the true spectrum $S$ for every frequency bin
+\begin{equation}
+    \nonumber
+    r(f) = \frac{|\hat{S}(f) - S(f)|}{S(f)}\,.
+\end{equation}
+We found this to be a better choice than residuals since relative errors are not dependent on the magnitude of the data.
+Second Fig \ref{fig:FPEmean} (below) shows that the largest error in the reconstruction of the spectrum is in the tails, while in the bulk of the distribution we have a good approximation of the true function.
+For the entire spectrum, the mean relative error $\Bar{r}$ is 
+\begin{equation}
+    \nonumber
+    \bar{r}_{\bar S, FPE} = 0.025.
+\end{equation}
+FPE performs well in reconstructing the spectrum even with relatively short time series, with a deviation from the true spectrum around 2\%, but it is evident that it is not able to capture all the features of the spectrum, since the reconstruction in the tails is inaccurate.
+If we compute the average percentage error in three standard deviations from the mean we gain an order of magnitude estimate of the accuracy of the reconstruction, obtaining:
+\begin{equation}
+    \nonumber
+    \bar{r}_{\bar S, FPE}^{\mu \pm 3\sigma} = 0.005. 
+\end{equation}
+and we can state the the method is highly performant in this frequency window. 
+Since it is not always possible to work with more then one realization for a process, as in observational experiments, it is important to understand how this method behaves when we apply it on a single time series.
+In that case, the mean error and its standard deviation are:
+\begin{equation}
+    \nonumber
+    \bar r_{FPE} = 0.16; \quad \sigma_{r} = 0.16\,.
+\end{equation}
+The average error of a reconstruction based on a single time series is bigger than the one obtained considering the average of all the reconstructions.
+Standard deviation of the errors assume small values and 90\% credibility regions (Cr.) in \ref{fig:FPEmean} are tight.
+This means that if we randomly take two of the reconstructed spectra, we expect their differences to be statistically small.
+These facts are evidence for FPE to be a relatively stable optimizer.
+We already said that the estimate for filter's length lays in a well defined range of values.
+The quality of the reconstruction, as noted in Chapter 1, depends on the order of the reconstructed AR process.
+To investigate this idea further, we plot the value of the optimizer for each recursion and the residuals as a function of the estimate for the order of the autoregressive process (\textbf{Figure \ref{fig:FPEErrorOrder}}),
+
+\begin{figure}[H]
+    \noindent
+    \centering
+    \includegraphics[width = \linewidth]{Images/NormalPSD/FPEoptimizer.pdf}
+    \caption{Stability of FPE, that results in a cluster of values in a small region (left), is due to the typical behaviour of the optimizer as a function of the AR order (right).}
+    \label{fig:FPEErrorOrder}
+\end{figure}
+Foremost evidence for the stability of this optimizer is found in its values as a function of the AR order.
+Figure \ref{fig:FPEErrorOrder} shows that Values tend to decrease monotonically until a sharp and well defined minimum is reached, to start growing again after this area.
+This is representative of the typical behaviour of this optimizer as a function of the length of the filter, and explains why the estimates for AR orders are confined in a small region.
+This plot also shows that most of the values collapsed in the region of AR orders comprised between 30 an 50.
+There are only few values outside this interval, and get fewer as order's estimate increases.
+In this region, there is no strong evidence of some correlation between the estimate for M and its associated relative error.
+It is important to notice that the estimate associated with the higher values of M (around $M > 60$) exhibit the highest 'mean' relative error.
+\\
+We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length M is clustered in a region where there is no dependence of error from M.
+\subsubsection{Optimum Bayes Decision Rule (OBD)}
+The second optimizer in consideration is OBD.  (\textbf{Figure \ref{fig:OBD3000}}) shows a good agreement between the average over the reconstructions and true spectrum.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width = \linewidth]{Images/NormalPSD/OBDpsdAndResiduals.pdf}
+    \caption{Reconstruction of the spectrum from 1000 simulations with OBD optimizer. Plot is mean of all the results with 90\% credibility regions (above) and percentage error (below)}
+    \label{fig:OBD3000}
+\end{figure}
+The same behaviour of FPE is observed, a good quality reconstruction at the center of the distribution with quite small Credibility regions (Cr.) at a $90$\% credibility level.
+Qualitatively they behave similarly but quantitatively the disagreement of this method is found to be even larger than the one obtained with FPE, as can be noticed in the plot of relative errors in figure \ref{fig:OBD3000} (below) where at the extremes the inaccuracy is even larger than 300\% with respect to the real value of the spectrum.
+For the overall spectrum and within three standard deviation from the mean, we find that the reconstructed spectrum has a deviation of
+\begin{equation}
+    \nonumber
+    r_{\bar S, OBD} = 0.15, \qquad r^{\mu \pm 3\sigma}_{\bar S, OBD} = 0.018. 
+\end{equation}
+When applied to single realizations of the time series, we find
+\begin{equation}
+    \nonumber
+    \bar{r}_{OBD} = 0.237, \qquad \sigma_{r, OBD} = 0.0296.
+\end{equation}
+It is interesting to notice that overall relative error of the average spectrum is of the same order of magnitude of the average single reconstruction for the spectrum.
+Averaging over a 1000 realisations does not substantially improve our results.
+As a function of the filter length, the values taken by the OBD cluster in a small region, indicating the stability of the reconstructed process order $M$.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width = \linewidth]{Images/NormalPSD/OBDoptimizer.pdf}
+    \caption{Stability of OBD, that results in a cluster of values in a small region (left), is due to the typical behaviour of the optimizer as a function of the AR order (right)}
+       \label{fig:OBDErrorOrder}
+\end{figure}we see that again our results are clustered in a very small region, due to the behaviour of the optimizer that converge in a small area before growing again.
+The estimate of the length of the filter is around $p = 20 \div 30$.Note the difference compared to FPE (see \ref{fig:FPEErrorOrder}); for OBD the mean relative error depends on $M$ as
+\begin{equation}
+    \hat r \sim M ^ {-2.016 \pm 0.006}.
+\end{equation}
+In summary, OBD shows greater error wherever in the spectrum, due to the fact that it choose filters that are too short and in a region where quality drastically change when changing M.
+In this case we can say that we have a stable estimate for M, but due to this dependence, the estimate is not as stable as FPE's.
+\subsubsection{Criterion Autoregressive Transfer function (CAT)}
+This last criterion differs substantially from the two other optimizers. Figure \ref{fig:CATmean} (above) shows a good agreement between the estimate of the spectrum for every frequency
+\begin{figure}[H]
+    \centering
+    \includegraphics[width = \linewidth]{Images/CAT/CATnormal.pdf}
+    \caption{Mean spectrum for CAT optimizer, 1000 simulations for 3000 points datasets}
+    \label{fig:CATmean}
+\end{figure} and also tails are well captured by the average spectrum.
+However, the variance of the reconstructed spectrum is quite large, much larger than for FPE and OBD.
+In general, the relative error is nearly constant, $\sim 10\%$ over all frequencies, see  Fig.
+\ref{fig:CATmean} (below).
+The residuals, averaged over all frequencies, are:
+\begin{equation}
+    \nonumber
+    \bar r_{\bar S, CAT} = 0.016 \qquad \bar r^{\mu \pm \sigma}_{\bar S, CAT} = 0.017.
+\end{equation}
+the overall reconstruction is thus quite accurate, with deviations from the simulated spectrum of the order of $1\% \div 2\%$.\\
+The picture drastically changes in the case of individual time series realizations. The average residuals are, in fact
+\begin{equation}
+    \bar r_{CAT} = 0.42 \qquad \sigma_{CAT} = .44\,.
+\end{equation}
+The error is indeed $\sim 42\%$, hence much larger than in the case of FPE and OBD.
+Both large error bars and the large standard deviation are indications of the instability of this optimizer.
+This is exemplified by the variety of different spectra that can be obtained considering different time series.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width = \linewidth]{Images/CAT/CAToptimizer1.pdf}
+    \caption{CAT is not a stable optimizer since its optimizer do not converge to any specific region of the AR estimate order}
+    \label{fig:CAToptimizer}
+
+\end{figure}
+The reason for this behaviour, can be found in \textbf{Figure (\ref{fig:CAToptimizer})}.
+CAT does not converge to any specific value for the order of the autoregressive process.
+In turn, this results in estimates for the order that span a large range.
+Figure \ref{fig:CAToptimizer} also shows a strong dependence of the error on the estimated length of the filter.
+The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like the big excursion in the size of datas of a normal distribution, but this length is also responsible of an augmented variance in the estimate, introducing spurious peaks in the reproduction (Figure \ref{fig:GaussMedian})
+
+Averaging, the noise in each spectrum is cancelled.
+It is interesting that the cancelling of errors in averaging is property of gaussian noise (the process is gaussian, after all).
+This implies, in a sense, that each estimate of the spectrum is independent of any other, as suggested by the huge variance in the residuals.
+This instability is not a good property for an estimation, so that this estimator might be found to be optimal for repeatable experiments and not for single observation of some process.
+Another approach to this method will be proposed in section \ref{sec:ACAT}
+
+\begin{figure}
+    \centering
+    \includegraphics[width = \linewidth]{Images/NormalPSD/Gaussmedian.pdf}
+    \caption{Reconstructed power spectral density with the three method. The reported reconstruction is the one that shows the median percentage error (overall)}
+    \label{fig:GaussMedian}
+\end{figure}
+\subsection{Summary and Comparison}
+\begin{figure}
+    \centering
+    \includegraphics[width = \linewidth]{Images/NormalPSD/Normal Residuals.pdf}
+    \caption{Percentage error in the reconstruction for every optimizer, as a function of the frequency. Lower panel restricts the domain to three standard deviation to increase the definition}
+    \label{fig:NormalEnsembleResiduals}
+\end{figure}
+In our analysis, FPE and OBD optimizers are found to behave similarly while CAT shows completely different properties.
+Studying the average spectrum is not sufficient to choose which method behaves better.
+\\
+The bad behaviour in the reconstruction of the tails from OBD and FPE might be due to the fact that very small values require longer filters. 
+In fact when we consider the average for all the spectra we note that the shorter the filter the larger the error in the tails: OBD, which select the shortest filters, arrives to a $300 \%$ error at the extrema of the power spectral density, FPE which select intermediate order has got a $85\%$ error in same window, while CAT, that selects a wide range of long filters, has got a constant accuracy everywhere in the spectrum.
+\\
+Figure \ref{fig:NormalEnsembleResiduals} shows that CAT is the method that minimizes errors in the tails (upper plot), but if one restrict to three standard deviations from the mean, FPE is found to reach best accuracy (lower plot).
+The frequency averaged percentage error is reflecting this behaviour. Considering the whole reproduction, error's average and standard deviation are found to be: 
+\begin{align*}
+\centering
+&r_{\bar S, FPE} = 0.025; \quad \sigma_{\bar S, FPE} = 0.083;  \\
+&r_{\bar S, OBD} = 0.15;\quad \sigma_{\bar S, OBD} = 0.42 ; \\ 
+&r_{\bar S, CAT} = 0.015; \quad \sigma_{\bar S, CAT} = 0.012
+\end{align*}
+In is interesting to see that the average for FPE and OBD is strongly affected from the high value of the error in the tails of the distribution.
+If one consider the three standard deviations from the mean, the value of the error substantially decreases,
+\begin{align*}
+\centering
+&r_{\bar S, FPE} = 0.005; \quad \sigma_{\bar S, FPE} = 0.004;  \\
+&r_{\bar S, OBD} = 0.018;\quad \sigma_{\bar S, OBD} = 0.019 ; \\ 
+&r_{\bar S, CAT} = 0.015; \quad \sigma_{\bar S, CAT} = 0.011
+\end{align*}
+showing that the overall percentage error is just a first indicator for the quality of the reproduction, but is not sufficient to perform a full characterization. \\
+It is evident that tails do not affect CAT  average error and standard deviation at all, so that it's accuracy is approximately constant everywhere in the spectrum. \\ \ 
+Since both FPE and CAT shows pros and cons, at this level, there is no reason to choose between them.\\ 
+Analyzing the resolution of the single reconstruction, figure \ref{fig:optcomparison} clearly shows that FPE minimize overall relative error giving a stable result.\\ 
+CAT instead shows a larger error and do not converge to any specific value.
+Almost every reproduction obtained with CAT is worst with respect to those obtained with FPE, but their mean is more precise.
+\\
+It's instability is probably the cause for this behaviour. Since it does not show a strong convergence to some order with respect to others, very different lengths are selected.
+Longer lengths are able to capture a lot of details but are more likely to introduce noise. In other words, long filters reduce bias but increment the variance of the result. 
+Taking the average, for central limit theorem the error decreases as $\sim \frac{1}{\sqrt{N}}$. In this way we compensate noise and obtain a spectrum with small bias and small fluctuations.
+Considering the average, CAT reconstruct precisely and with good accuracy every part of the spectrum, even if in "regular" parts best accuracy is reached from FPE.
+For a single reproduction, CAT is very noisy and unstable, while FPE performs the better.
+\ref{fig:optcomparison}
+\begin{figure}
+    \centering
+    \includegraphics[width = \linewidth]{Images/NormalPSD/NormalPSDcomparison.pdf}
+    \caption{Relative error as function of the length of the filter.}
+    \label{fig:optcomparison}
+\end{figure}
+\begin{figure}
+    \hfill
+    \centering
+    \includegraphics[width = \linewidth]{Images/NormalPSD/ordersComparison.pdf}
+    \caption{Histogram of estimate orders for each method}
+    \label{fig:ordersCompairson}
+
+\end{figure}
+definitely shows that FPE gives the best results: it is a stable method that choose filters in the clustered area associated with minimum error, while both OBD and CAT result lies outside this area and show bigger errors.
+Two other interesting plots that stress this fact are reported in figure \ref{fig:ordersCompairson} and \ref{fig:residualsComparison}
+\begin{figure}
+    \centering
+    \includegraphics[width = \linewidth]{Images/NormalPSD/ResidualsComparison.pdf}
+    \caption{Histogram of residuals for each method}
+    \label{fig:residualsComparison}
+\end{figure}
+where the histograms for the errors and for order's estimate are reported.
+These graphs provide stronger evidence for our previous statements.
+As already noticed, FPE is very stable and provide the smallest 'overall relative error'.
+\\
+CAT histogram shows a very interesting behaviour.
+The histogram for the order show two different peaks,  a small one associated to 'short filters', and a second high peak at $m = M$.
+This means that CAT is most likely to not converge at all at some value for the regression order, and it is more likely selects a length that is comparable with the largest available.
+If short filters are selected, the overall percentage error will be low, respectively represented by the first peak in the distribution of the errors, while if long filters are selected, the error is maximized.
+
+So, even if in some cases CAT might be more accurate when taking the average over several spectra, FPE guarantees that minimum error will be reached.
+
+\subsection{Analysis of Ligo Spectrum}
+\begin{figure}
+    \centering
+    \includegraphics[width = \linewidth]{Images/LIGOsimulate/LigoPSDoriginal.pdf}
+    \caption{Original power spectral density from Ligo}
+    \label{fig:LigoPSDOriginal}
+\end{figure}
+In this subsection we concentrate again on the reconstruction of a specific, known, power spectrum.
+We analyze the properties of reconstruction of this second spectrum whose shape is the Advanced LIGO design sensitivity theoretical spectral curve (Figure \ref{fig:LigoPSDOriginal}). 
+For this analysis we will study properties only for arrays of 40960 points.
+We consider a sampling rate of $2048 samples / second$ that means we can reconstruct a Nyquist frequency of $1024 Hz$.
+We chose a longer dataset than before because algorithm works with a 1 dimensional interpolation of spectrum, and with shorter datasets the interpolated spectrum is not consistent with the original spectrum and does not capture the height of the peaks.
+Shorter choices would result in noise extracted from a different power spectrum than the one in figure \ref{fig:LigoPSDOriginal}.
+\\
+We will study statistical properties from samples obtained with 500 simulations and compare the obtained results for each optimizer. 
+We start considering statistical properties of mean spectrum obtained averaging over all the simulations and conclude with an analysis for the statistics of the reproduction of single spectrum. 
+\subsubsection{Properties for the average spectrum}
+Here we report mean properties of the reproductions of the spectrum. The spectrum we want to reconstruct is reported in Figure \ref{fig:ligospectrum},
+\begin{figure}
+    \centering
+
+        \includegraphics[width = \linewidth]{Images/LIGOsimulate/PSD.pdf}
+        \caption{A priori spectrum from Ligo database}
+        \label{fig:ligospectrum}
+\end{figure}
+shows the original power spectral density and the ensemble average of every method. For the mean, the relative errors are: 
+\begin{equation}
+    \bar r_{\bar S, FPE} = 0.0426\%, \qquad \bar r_{\bar S, OBD} = 0.129\%, \qquad \bar r_{\bar S, CAT} = 0.0211\%
+\end{equation}
+Again, if one only check overall accuracy for every method, CAT is outperforming.
+In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them.
+To check this behaviour we take a look at the two peaks separately.
+For the first one a log plot of the reconstructed peaks and their relative errors are shown in Figure \ref{fig:ligo1peak}
+\begin{figure}[h]
+        \includegraphics[width = \linewidth]{Images/LIGOsimulate/1stPeakComparison.pdf}
+        \caption{1st peak of the spectrum and it reconstruction with every optimizer (left) and associated percentage error (right)}
+        \label{fig:ligo1peak}
+\end{figure}
+and the results reinforce our previous statement.
+The behaviour of the error is very similar for CAT and FPE, showing similar shapes but very different sizes, while CAT is giving a different and way more precise result.
+Both FPE and OBD are not able to capture the width of the peak and also show larger errors, while CAT has a good behaviour in reconstructing width's also.
+The reconstruction for very sharp peaks considering the average over all the spectra is found to be strongly inaccurate for both FPE and OBD, while a good accuracy is reached with CAT method. \\ 
+Considering the estimate for the position of the peak, whose true value is found to be $16.85 \quad a.u.
+$, within numerical accuracy both FPE and CAT find its exact position, while OBD shift the peak left at $16.7 a.u.$, with a $0.9\%$ inaccuracy.\\
+Second peak results are analogous, plotting the peak and the errors value around it
+\begin{figure}[h]
+    \centering
+        \includegraphics[width = \linewidth]{Images/LIGOsimulate/2ndPeakComparison.pdf}
+        \caption{2nd peak of the spectrum, Relative error around 2nd peak of the spectrum}
+        \label{fig:ligo2peak}
+\end{figure}
+we find again that FPE and OBD are performing similarly, the magnitude of the error is maximum for OBD and minimum for CAT. \\ 
+In this case even CAT shows large errors in some points, but they are reasonable in a large part of it, still, it remains the best reconstruction.
+For this second peak, all the methods find maximum position to be at $438.05 \quad a.u.$, which is the true position of the maximum.
+Looking at Figure \ref{fig:ligo1peak} at \ref{fig:ligo2peak}, it appears to be clear that the least of the accuracy is not on the peak, but immediately around, so that one of the biggest deal is the reconstruction of peak's width, that only CAT is able to capture with high definition.
+\\
+Even if FPE shows larger error in the peaks than CAT, when the average over all the frequencies is considered, their percentage errors are of the same order of magnitude.
+This means that outside the peaks FPE is more accurate than CAT, i.e.
+CAT is still giving noisy results.\\
+As for the normal distributed noise, mean spectrum is not sufficient to declare what is the best optimizer to be used, since most of experiment are purely observational and mean spectrum over different realizations of the same stochastic process cannot be computed.
+So, it is important to also understand what is their behaviour for single realizations of the process.
+\subsubsection{Statistical properties of single realizations} 
+The errors for the single realizations behave very differently from the errors obtained when we considered the mean spectrum.
+In computing the mean of the overall errors for each single spectrum, the results we obtain are the following:
+\begin{equation}
+    \bar{r}_{FPE} = 0.13 \pm 0.01; \qquad \bar{r}_{OBD} = 0.17 \pm .02; \qquad 
+    \bar{r}_{CAT} = 0.4 \pm 0.1
+\end{equation}
+where the error is just the standard deviation of the samples.
+In this case we find a different result: FPE shows the best results and order's estimate is associated with least errors.
+Also, having found a smaller standard deviation of the datas, we can state that FPE and OBD are more stable optimizers then CAT.
+In fact, their properties as a function of filter's length are almost identical to those obtained in the study of normal spectrum.
+If we again plot the value of the errors as a function of the estimate of the order (Figure \ref{fig:LigoOrderError})
+\begin{figure}
+    \centering
+    \includegraphics[width = \linewidth]{Images/LIGOsimulate/orderVSerror.pdf}
+    \caption{Plot of relative error as a function of the estimate for filter's length}
+    \label{fig:LigoOrderError}
+\end{figure}
+it is immediate to see that, again, FPE results lies in the are of filter's length associated with a minimum of relative errors, OBD lies in a clustered window with a higher error and CAT is not convergent, but most of the results are outside the minimum error area.
+So, we arrive at the same conclusion we had when we considered the normal specturm.
+OBD has no general properties that make it a preferable optimizer then FPE or CAT while, what to choose beneath the last two is based on the problem we are dealing with: if we have one single realization for the process, we are mostly sure that FPE would catch the area associated with the best resolution possible, while CAT would result in a spurious result with, in general, a not so good resolution.
+But, in case we have several realization for the same process, this last property of CAT, together with its instability, result to be his luck and, taking the average errors tend to compensate noise in the result, originating a very well defined spectrum that show same resolution almost everywhere, no matter how small the values are (as for normal PSD case) or how sharp peaks can be, the result has got generally good accuracy, also if it is clear that we have to pay this fact with very large error bars.
+So a choice between these methods is a matter of convenience and interest for each specific problem.
+\\
+\begin{figure}
+    \centering
+    \includegraphics[width = \linewidth]{Images/LIGOsimulate/LigoHists.pdf}
+    \caption{Histograms showing the estimate for the autoregressive order and overall percentage error obtained with all the three optimizers.}
+    \label{fig:my_label}
+\end{figure}
+
+\section{Example} 
+=======
 \subsection{Analysis of the optimizers}
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END & BIB
+>>>>>>> dc2ea6b02d554f80ea49ef0c67cf14dcffc60bb8
 
 \appendix
-\section{MESA solution}
+\section{MESA solution} \label{sec:MESA_solution}
 \subsection{MESA solution}
@@
 or, defining $a_0 = 1$,
-\begin{equation}\label{eq:MESApsd}
+\begin{equation}\label{eq:MESApsd_appendix}
     S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^z\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}. 
@@
 \begin{equation}
-   \frac{P_N}{2 \pi \imath} \oint _{\mathbb S^1}\frac{z^{-s - 1}}{\sum_{n = 0}^N a_n z^n \sum_{n = 0}^N a^*_n z^{-n}} = \bar r_s. 
+   \frac{P_N}{2 \pi \imath} \oint _{\mathbb S^1}\frac{z^{-s - 1}}{\sum_{n = 0}^N a_n z^n \sum_{n = 0}^N a^*_n z^{-n}}dz = \bar r_s. 
 \end{equation}
@@
 This problem is often solved numerically with Burg's Algorithm \cite{Vos}.
-A python implementation of Standard Burg's Method can be found in Appendix \ref{app:Burg}
 
@@
 so that we can reformulate the variational problem requiring $S(f)$ to maximize the entropy gain.
-In real examples we have no access to all frequencies but only on those between minus and plus the Nyquist frequency, defined from the sampling interval $\Delta t$ as
+In a real example, we have no access to all frequencies but only on those between minus and plus the Nyquist frequency, defined from the sampling interval $\Delta t$ as 
 \begin{equation}
@@
 \end{equation}
-All our quantities are well defined in this range only, so that we are not going to maximize \ref{eq:EntropyGain}, but
+All the quantities of interest are well defined in this range only, so that we are not going to maximize \ref{eq:EntropyGain} over all frequencies, but only on the admissible ones.
+Our objective becomes:
 \begin{equation}\label{eq:EntropyGain2}
@@
 \end{equation}
-Defining $\bar r_n$ as the sample autocorrelation, our constraints will be
+\par
+Now the the objective to maximize is well defined, we can write down the constraints of our problem.
+Defining $\bar r_n$ as the sample autocorrelation, we require that the variational solution $S(f)$ for the PSD matches the empirical autocorrelation. This means:
 \begin{equation}\label{eq:MaxConstraint}
@@
     \centering
-        \includegraphics[scale = 0.43]{Images/Noise and PSD/NormalPSDautocorr.pdf}
+        \includegraphics[width = \linewidth]{Images/Noise and PSD/NormalPSDautocorr.pdf}
         \caption{A priori spectrum (left) and autocorrelation (right)}

### 2021-02-17 scripts/welch.py
Default nfft in Welch is now segment length

@@ -39,7 +39,6 @@ def psd(strain,
     # number of samples per segments
     N = int(segment_duration*sampling_rate)
     
-    if nfft is None: nfft = len(strain)
     # compute the PSD and its frequencies
     frequencies, psd = welch(strain,
                              fs              = sampling_rate,

### 2021-02-17 scripts/mem_welch_comparison.py
White noise and Ligo noise comparison

@@ -18,12 +18,6 @@ from scipy.interpolate import interp1d
 
 
 
-def compute_spectra(time_series, dt):
-    welch_frequency, welch_spectrum = [], []
-    MESA_frequency, MESA_spectrum = [], [] 
-    for i in range(len(time_series)): 
-        None 
-
 def relative_error(real, estimate):
     return np.abs(real - estimate) / real 
 
@@ -32,17 +26,15 @@ def relative_error(real, estimate):
 if __name__ == '__main__': 
     #Sampling variables 
     dt = 1. / 4096
-    times = np.array([1, 5, 10, 100, 1000])
+    times = np.array([1, 5, 10, 100])#, 1000])
     N = times // dt 
-    segment_duration = np.array([521, 1024, 2048, 8192, 32768]) 
+    segment_length = np.array([521, 1024, 2048, 8192])  #, 32768])
     Ny = 0.5 / dt 
     
     #Generate normal PSD and import Ligo PSD 
-    wnoise_frequency = np.arange(0, N[0] // 2 + 1) / times[0]
-    wnoise_spectrum = np.repeat(1, wnoise_frequency.size) * dt 
+    wnoise_frequency, wnoise_spectrum = np.arange(0, N[0] // 2 + 1) / times[0], np.repeat(1, N[0] // 2 + 1) * dt
     ligo_frequency, ligo_spectrum = np.loadtxt('LIGO-P1200087-v18-AdV_DESIGN_psd.dat', unpack=True)     
-    ligo_interp = interp1d(ligo_frequency, ligo_spectrum)
-    #Lists containing all simulated arrays, frequencies and psds 
+    #Generating Noise 
     print('Generating white noise series')
     Wtime, Wtime_series, Wfrequency, Wfrequency_series, Wpsd = generate_data(wnoise_frequency,
                                                                         wnoise_spectrum,
@@ -53,6 +45,36 @@ if __name__ == '__main__':
                                                                              ligo_spectrum,
                                                                              times[-1],
                                                                              1 / dt)
+    w_interp = interp1d(wnoise_frequency, wnoise_spectrum)   
+    l_interp = interp1d(ligo_frequency, ligo_spectrum, fill_value = 'extrapolate')    
+    M = MESA()
+    
+
[…]

### 2021-02-17 scripts/mem_welch_comparison.py
Added saving and plots

@@ -24,6 +24,8 @@ def relative_error(real, estimate):
 
 
 if __name__ == '__main__': 
+    import os 
+    save_dir = os.getcwd() + '/comparisons_2/'
     #Sampling variables 
     dt = 1. / 4096
     times = np.array([1, 5, 10, 100])#, 1000])
@@ -49,7 +51,8 @@ if __name__ == '__main__':
     l_interp = interp1d(ligo_frequency, ligo_spectrum, fill_value = 'extrapolate')    
     M = MESA()
     
-
+    init_plotting()
+    
     for i, n in enumerate(segment_length):
         #White Noise
         w_welchFreq, w_welchSpectrum = welch.psd(Wtime_series[:int(N[i])], 1 / dt, n * dt)
@@ -65,13 +68,16 @@ if __name__ == '__main__':
         l_welchFreq, l_welchSpectrum = welch.psd(Ltime_series[:int(N[i])], 1 / dt, n * dt )
         M.solve(Ltime_series[:int(N[i])])
         l_mesaSpectrum, l_mesaFreq = M.spectrum(dt)
-        
+
         fig2, ax2 = plt.subplots()
         ax2.loglog(l_welchFreq[: n // 2], l_welchSpectrum[: n // 2], color = 'blue')
         ax2.loglog(l_mesaFreq[: int(N[i] // 2)], l_mesaSpectrum[:int(N[i] // 2)], color = 'red')
         ax2.loglog(ligo_frequency, ligo_spectrum, '--', color = 'k')
         ax2.set_title('Ligo Noise reconstrcution with {} points'.format(N[i]))
         ax2.set_xlim(10, 2048)
+        
+        fig.savefig(save_dir + 'White Noise {} points.pdf'.format(int(N[i])))
+        fig2.savefig(save_dir +'Ligo Noise {} points.pdf'.format(int(N[i])))
         
 
     

### 2021-02-18 paper/mesa.tex
second chunk edited

@@
 without highlighting its connection with the time structure of the process itself, thus masking some important properties that will explore further in what follows.
-The latter definition in (\ref{eq:psd-f-definition}) gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating for an unknown process.
+The latter definition in (\ref{eq:psd-f-definition}) gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.
 %Under such assumption, it can be shown [cite something] that the problem of characterizing the noise properties is greatly simplified and its properties are summarized by a single \textit{autocorrelation function}, which describes how an observation at a given time correlates with another observation a later time.
@@
 %Thus, in many real application, the knowledge of the PSD is crucial for the understanding of the process and indeed algorithms for PSD computation are a standard part of every software package for computational physics.
-An ubiquitous method for such computation is due to Welch \cite{Welch} and it based on the method of periodograms.
+A ubiquitous method for such computation is due to Welch \cite{Welch} and it based on the method of periodograms.
 Basically, the PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many windowed batches and averaging the squared moduli of their Fourier transforms.
@@
 \section{Theoretical foundations}
-The Maximum Entropy principle (MAXENT) is among the most important results in Bayesian Statistics.
+The Maximum Entropy principle (MAXENT) is among the most important results in Bayesian probability theory.
 It provides a way to uniquely assign probabilities to a phenomenon in a way that best represent our state of knowledge, while at the same time it is non committal with unavailable information.
@@
 \sschmidt{This section is yet to prune}
-To introduce MAXENT as a variational problem, we will workout some simple examples to understand how information's entropy and `uncertainty' are related\footnote{
-The word `information' will not be used to indicate our knowledge on the system under study: this is what we will call `evidence'.
-Information will be intended as in communication theory: it specifies the quantity of details needed to provide a full description of the system under study.
-}.
-
-If we consider a perfectly sinusoidal signal, knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information entropy content.
-Even less information is required if we are studying a system whose outcome is certain (has probability $p = 1$).
-If the amount of information $I$ can be considered to zero, a communication is not even needed.  
+Before introducing MAXENT as a variational problem, we will workout some simple examples to contexualise how information's entropy
+and `uncertainty' are related.
+First of all,  it is worth noting that the word `information' will not be used to indicate our knowledge on the system under study: this is what we will call `evidence'.
+Information will be intended as in communication theory: it specifies the length of the message necessary to provide
+a full description of the system under study.
+
+For instance, consider a perfectly sinusoidal signal: knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: 
+it has a low information entropy content. Even less information is required if we are studying a system whose outcome is 
+certain (has probability $p = 1$). If the amount of information $I$ can be considered to zero, a communication is not even needed.  
 Shannon \cite{Shannon} proposed the quantity
@@
 to represent the quantity of information brought by an outcome $x$ with probability $p(x)$.
-It has the nice property of being an additive quantity and it is monotonically decreasing as a function of $p \in [0, 1]$.
+It is additive quantity as well as monotonically decreasing as a function of $p \in [0, 1]$.
 The more uncertain the outcome, the higher the information it brings.
@@
 \end{equation}
-this describes a situation of `maximum ignorance'; we have no evidence at all and we have no instruments to predict what the outcome of such an experiment will be.  \\ 
-This can be generalized in the case of $N$ events. Assigning same probability
-\begin{equation*}
-    P_1 = \dots = P_N = \frac{1}{N}
-\end{equation*}
-for each outcome, i.e. appealing to the indifference principle (IP), is equivalent to the requirement for our experiment to be the most unpredictable. 
-All we know is that one over $N$ events is going to happen. With this evidence, we cannot expect one outcome to occur over the others. 
-To report the results of such an experiment, we need to communicate every single outcome: it 'carries' high information. 
-This description also makes clearer the role played by evidence: it has to be used in probability assignment in order to make the result 'more certain' or 'less informative'.
-It is clear from the point of view of communication theory: the higher the evidence on a system, the lower the information I need to send to describe it.
-\par
-It is now clear that any assignment different from the 'most uncertain' is violating the hypothesis for it to being non committal to non-available evidence.
+this describes a situation of `maximum ignorance'; we have no evidence at all and we have no way of predicting what the outcome of such an experiment will be.
+The generalisation to $N$ possible outcomes is straightforward.
+\\
+%This can be generalized in the case of $N$ events. Assigning same probability
+%\begin{equation*}
+%    P_1 = \dots = P_N = \frac{1}{N}
+%\end{equation*}
+%for each outcome, i.e. appealing to the indifference principle (IP), is equivalent to the requirement for our experiment to be the most unpredictable. 
+%All we know is that one over $N$ events is going to happen. With this evidence, we cannot expect one outcome to occur over the others. 
+%To report the results of such an experiment, we need to communicate every single outcome: it 'carries' high information. 
+In this context, the role played by evidence should also be clear: it dictates the probability assignment in order to make the result 'more certain' or 'less informative'.
+Any assignment different from the 'most uncertain' is violating the principle of it to being non committal to non-available evidence.
 We can only assign a more certain probability, i.e.
@@
 the most informative.
-Assigning probabilities requires maximization of uncertainty using evidence as constraint.
-This looks like a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized.
+From the point of view of communication theory: the higher the evidence on a system, the lower the information I need to send to describe it.
+\par
+MAXENT ratifies the aforementioned principle by stating that probabilities should be assigned by maximizing uncertainty using evidence as constraint. 
+This defines a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized.
 $H$ has to be a 'generalization' of information defined in equation (\ref{eq:information}) to be something like the 'expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
-
-Some properties for $H$ can be worked out easily. 
-From previous discussion, given a fixed number $N$ of possible outcomes, we want $H$ to be maximized by the choice $p_1 = \dots = p_N$.
-Consider now the opposite situation, in which probability is assigned appealing to IP and we have a different number of possible outcomes.
-Intuitively, the most unpredictable experiment is the one with larger number of outcomes: probability of "getting it wrong" increases as $\frac{N -1}{N}$.
+%
+%Some properties for $H$ can be worked out easily. 
+%From previous discussion, given a fixed number $N$ of possible outcomes, we want $H$ to be maximized by the choice $p_1 = \dots = p_N$.
+%Consider now the opposite situation, in which probability is assigned appealing to IP and we have a different number of possible outcomes.
+%Intuitively, the most unpredictable experiment is the one with larger number of outcomes: probability of "getting it wrong" increases as $\frac{N -1}{N}$.
 So when equal probabilities are assigned, we want
-\begin{equation}
-    \nonumber
-    H\left[\frac{1}{N}, \dots, \frac{1}{N}\right]
-\end{equation}
-to increase monotonically with N.
-
-We also want our function to be continuous in its parameters: any small change in probability has to slightly affect uncertainty. \\
-Last is composition property. Consider three events with probability $p_1, p_2, p_3$ such that $p_2 + p_3 = 1 - p_1 = q$. We require 
-\begin{equation}
-    H\left(p_1, p_2, p_3\right) = H(p_1, q) + qH\left(\frac{p_2}{q}, \frac{p_3}{q}\right),
-\end{equation}
-so that uncertainty is additive.
-
-Shannon \cite{Shannon} showed that the only functional form satisfying our assumptions is given by:
+%\begin{equation}
+%    \nonumber
+%    H\left[\frac{1}{N}, \dots, \frac{1}{N}\right]
+%\end{equation}
+%to increase monotonically with N.
+%
+%We also want our function to be continuous in its parameters: any small change in probability has to slightly affect uncertainty. \\
+%Last is composition property. Consider three events with probability $p_1, p_2, p_3$ such that $p_2 + p_3 = 1 - p_1 = q$. We require 
+%\begin{equation}
+%    H\left(p_1, p_2, p_3\right) = H(p_1, q) + qH\left(\frac{p_2}{q}, \frac{p_3}{q}\right),
+%\end{equation}
+%so that uncertainty is additive.
+
+Shannon \cite{Shannon} showed that the only functional form satisfying continuity with respect to its parameters and additivity is:
 \begin{equation}\label{eq:entropy}
@@
 \end{equation}
-and $H$ is what we will call entropy.
-
-This interpretation is straigthforward: if we compare it with Eq.~(\ref{eq:information}), we see that the entropy is defined as the 'average information'.
+and $H$ is what we will call (information) entropy.
+
+%This interpretation is straightforward: if we compare it with Eq.~(\ref{eq:information}), we see that the entropy is defined as the 'average information'.
 The probability assignments made maximizing it with respect to $p$, is somehow requiring  the 'expected value' of uncertainty to be a maximum.
@@
 %\end{equation}
-
-To show how the maximization of entropy, supplemented with additional evidence, can provide a unique probability assignment, we will work out three examples.
-In each of them, we will add more evindence (i.e. constraints on the optimization problem) and we will obtain different results.
-
-First of all, we want to show that, in case in which our only constraint is exhaustiveness
-$\sum_{i = 1}^{N}p_i = 1 $,
-MAXENT reduces to indifference principle as it should be. The variational problem is defined as
-\begin{equation}\label{eq:entropyNormConst}
-    \delta\left[H - \lambda\sum_{i = 1}^N (p_i - 1)\right] = 0 
-\end{equation}
-with
-\begin{equation}
-    \delta H\left[p_1, \dots, p_N\right] = \sum_{i = 1}^N\frac{\partial H}{\partial p_i}\delta p_i = - \sum_{i = 1}^N\left(\log p_i + 1\right),
-\end{equation}
-so that equation \ref{eq:entropyNormConst} reduce to
-\begin{equation}\nonumber 
-    \sum_{i = 1}^N\left[\log p_i + \lambda + 1 \right]\delta p_i = 0\,.
-\end{equation}
-The solution of the previous equation is easily worked out requiring every term to be zero: 
-\begin{equation}\nonumber 
-    p_i = e^{-\lambda - 1}.
-\end{equation}
-Applying the normalization condition, it is immediate to show that
-\begin{equation}
-    p_i = \frac{1}{N}.
-\end{equation}
-MAXENT reduces to the IP when only the normalization constraint is used.
-
-In the second example, we will add the constraint that the distribution has a known mean value: $\int \d{x} x p(x) = \bar{x}$. For simplicity we will work in the continuous case.
-The variational problem can be rewritten as: 
-\begin{equation}
-    \int \left[\log p(x) + 1 + \lambda + \mu x \right]\delta p(x) = 0
-\end{equation}
-with constraints: 
+\wdp{I need to review and correct the following}
+To show how the maximization of entropy, supplemented with additional evidence, can provide a unique probability assignment, 
+we use a simple, pedagogical example. Consider the case in which we want assign a probability distribution $p(x)$ to a variable $x$ 
+for which we know mean $\mu$ and covariance $\Sigma$. Our `evidence' is then coded as the following constraints:
+\begin{align}
+$\sum_{i = 1}^{N}p_i &= 1 $\\
+$\sum_{i = 1}^{N}x_i p_i &= \mu $\\
+$\sum_{i = 1}^{N}(x_i - \mu)(x_j - \mu) p_i &= \Sigma$\,
+\end{align}
+where the first row states the requirement of exhaustivity.
+\wdp{i think we want the multivariate case since i) the gaussian case is also on wikipedia ii) we actually will need that  later on.}
+
+%First of all, we want to show that, in case in which our only constraint is exhaustiveness
+%$\sum_{i = 1}^{N}p_i = 1 $,
+%MAXENT reduces to indifference principle as it should be. The variational problem is defined as
+%\begin{equation}\label{eq:entropyNormConst}
+%    \delta\left[H - \lambda\sum_{i = 1}^N (p_i - 1)\right] = 0 
+%\end{equation}
+%with
+%\begin{equation}
+%    \delta H\left[p_1, \dots, p_N\right] = \sum_{i = 1}^N\frac{\partial H}{\partial p_i}\delta p_i = - \sum_{i = 1}^N\left(\log p_i + 1\right),
+%\end{equation}
+%so that equation \ref{eq:entropyNormConst} reduce to
+%\begin{equation}\nonumber 
+%    \sum_{i = 1}^N\left[\log p_i + \lambda + 1 \right]\delta p_i = 0\,.
+%\end{equation}
+%The solution of the previous equation is easily worked out requiring every term to be zero: 
+%\begin{equation}\nonumber 
+%    p_i = e^{-\lambda - 1}.
+%\end{equation}
+%Applying the normalization condition, it is immediate to show that
+%\begin{equation}
+%    p_i = \frac{1}{N}.
+%\end{equation}
+%MAXENT reduces to the IP when only the normalization constraint is used.
+%
+%In the second example, we will add the constraint that the distribution has a known mean value: $\int \d{x} x p(x) = \bar{x}$. For simplicity we will work in the continuous case.
+%The variational problem can be rewritten as: 
+%\begin{equation}
+%    \int \left[\log p(x) + 1 + \lambda + \mu x \right]\delta p(x) = 0
+%\end{equation}
+%with constraints: 
+%\begin{equation*}
+%    \int \d{x} \; p(x)  = 1 \text{  and  }
+%    \int \d{x} \; x p(x) = \bar x.
+%\end{equation*}
+%For simplicity, we consider $x$ to be a variable ranging in the interval $(-\infty,+\infty)$, but this can be easily computed for every other interval.
+%
+%The general solution for $p(x)$ is:
+%\begin{equation}
+%    p(x) = e^{-1 - \lambda - \mu x} 
+%\end{equation}
+%where $\lambda$ and $\mu$ are two Lagrange multipliers
+%The normalization conditions implies 
+%\begin{equation}\nonumber 
+%   1 =  \int \d{x} \; p(x) = e^{-1-\lambda}\int e^{-\mu x} dx, 
+%\end{equation}
+%so that: 
+%\begin{equation}\nonumber 
+%    e^{-1 -\lambda} = \mu.
+%\end{equation}
+%The fixed average condition provides a value for $\mu$ in terms of $\bar{x}$:
+%\begin{equation}
+%	\mu = \frac{1}{\bar{x}}
+%\end{equation}
+%The solution is then an exponential distribution:
+%\begin{equation}
+%    p(x) = \frac{1}{\bar{x}} \exp\left(-\frac{x}{\bar{x}}\right). \nonumber 
+%\end{equation}
+
+%Lastly, and more interestingly, we also consider a constraint on the variance. For this reason, we set:
+%\begin{align*}
+%    \int \d{x} \; p(x)  = 1 \text{  and  }
+%	\int \d{x} \; x p(x) = \bar{x} \\
+%	\text{  and  }
+%	\int \d{x} \; (x-\bar{x})^2 p(x)  = \sigma^2
+%\end{align*}
+%Performing the same computation as above, we obtain a well known result:
+%\begin{equation}
+%	p(x) = \frac{1}{\sigma\sqrt{2\pi}} e^{-\frac{1}{2}\left(\frac{x-\bar{x}}{\sigma}\right)^2}
+%\end{equation}
+
+Take a vector  $\vec x(t)$ with components in $\mathbb{R}^N$
 \begin{equation*}
-    \int \d{x} \; p(x)  = 1 \text{  and  }
-    \int \d{x} \; x p(x) = \bar x.
+    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
 \end{equation*}
-For simplicity, we consider $x$ to be a variable ranging in the interval $(-\infty,+\infty)$, but this can be easily computed for every other interval.
-
-The general solution for $p(x)$ is:
-\begin{equation}
-    p(x) = e^{-1 - \lambda - \mu x} 
-\end{equation}
-where $\lambda$ and $\mu$ are two Lagrange multipliers
-The normalization conditions implies 
-\begin{equation}\nonumber 
-   1 =  \int \d{x} \; p(x) = e^{-1-\lambda}\int e^{-\mu x} dx, 
-\end{equation}
-so that: 
-\begin{equation}\nonumber 
-    e^{-1 -\lambda} = \mu.
-\end{equation}
-The fixed average condition provides a value for $\mu$ in terms of $\bar{x}$:
-\begin{equation}
-	\mu = \frac{1}{\bar{x}}
-\end{equation}
-The solution is then an exponential distribution:
-\begin{equation}
-    p(x) = \frac{1}{\bar{x}} \exp\left(-\frac{x}{\bar{x}}\right). \nonumber 
-\end{equation}
-
-Lastly, and more interestingly, we also consider a constraint on the variance. For this reason, we set:
-\begin{align*}
-    \int \d{x} \; p(x)  = 1 \text{  and  }
-	\int \d{x} \; x p(x) = \bar{x} \\
-	\text{  and  }
-	\int \d{x} \; (x-\bar{x})^2 p(x)  = \sigma^2
-\end{align*}
-Performing the same computation as above, we obtain a well known result:
-\begin{equation}
-	p(x) = \frac{1}{\sigma\sqrt{2\pi}} e^{-\frac{1}{2}\left(\frac{x-\bar{x}}{\sigma}\right)^2}
-\end{equation}
-
-%Take a vector  $\vec x(t)$ with components in $\mathbb{R}^N$
-%\begin{equation*}
-%    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
-%\end{equation*}
-%We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals.
+We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals.
 If we put constraints in both mean $\vec \mu(t)$ and covariance matrix $C_(t_i, t_j)$, MAXENT allow us to assign for $\vec x(t)$ a multivariate normal distribution \cite{gregory_2005}:
-%\begin{align}
-%    p&\left(\vec x(t)\vert I\right) = \nonumber \\
-%    &\frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right). 
-%\end{align}
+\begin{align}
+    p&\left(\vec x(t)\vert I\right) = \nonumber \\
+    &\frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right). 
+\end{align}
+
 Thus MAXENT provide a deeper insight into the Gaussian distribution.
@@
 \subsection{Maximum Entropy Spectral Analysis} \label{sec:MESA}
-The computation of the power spectral density is an important component in stationary time series analysis and characterization.
-In principle, if the autocorrelation was known exactly (i.e. at every time $\tau$), the computation of the PSD would be just a single Fourier transform.
-However, in every realistic setting, we are dealing with a finite number of samples $N$ and it is not possible such precise computation.
+The computation of the power spectral density is an important component in time series analysis and characterization.
+In principle, if the autocorrelation was known exactly (i.e. at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD 
+would reduce to a single Fourier transform.
+However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process, hence such computation
+is impossible.
 Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases with $k$ as $\sigma \sim \frac{1}{\sqrt{N - k}}$, so that only a few values for the autocorrelation function can actually be computed reliably.
@@
 Still, study of ensemble average is important to highlight some interesting properties.
-In this way we can fully characterize the performance of each optimizer
+In this way we can fully characterize the performance of each optimizer .
 We will attentively study the role of the filter length estimate M in determining the accuracy of the reconstruction.
@@
 \end{figure}
-
-\section{Example} 
-=======
-\subsection{Analysis of the optimizers}
-
-\subsection{Comparison with Welch method}
-
-\section{Examples}
-
-	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END & BIB
->>>>>>> dc2ea6b02d554f80ea49ef0c67cf14dcffc60bb8
 

### 2021-02-18 paper/mesa.tex
change my  color for  comments
edits up to section 2

@@
 % comments command
-\newcommand{\wdp}[1]{{\textcolor{green}{\texttt{SS: #1}} }}
+\newcommand{\wdp}[1]{{\textcolor{magenta}{\texttt{WDP: #1}} }}
 \newcommand{\amartini}[1]{{\textcolor{blue}{\texttt{SS: #1}} }}
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE
-	\title{Maximum Entropy Spectral Analysis}
+	\title{Maximum Entropy Spectral Analysis: a case study}
 	\author{Alessandro \surname{Martini}}
@@
 
-The study of the properties of a noisy process is a crucial task in many fields of physics and mathematical physics [Here we should make some examples...].
+The study of the properties of  stochastic processes is a crucial task in many fields of physics and mathematical physics [Here we should make some examples...].
 A particularly interesting class of stochastic processes are the so-called \textit{wide-sense} stationary processes.
@@
 \begin{equation}
-	R(\tau) = \mathbf{E}[x_t \cdot x_{t+\tau}]
+	C(\tau) = \mathbf{E}[x_t \cdot x_{t+\tau}]
 \end{equation}
@@
 \begin{equation}
-	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau R(\tau) e^{-i 2 \pi f \tau}\,.
+	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau C(\tau) e^{-i 2 \pi f \tau}\,.
 \end{equation}
@@
 
+The maximisation of entropy, supplemented by evidence in the form of constraints to which the sought-for probability distribution must obey,
+gives rise to several of the most common probability distributions utilised commonly by data scientists.
+For instance, whenever the only constraint available is the normalisation of the probability distribution, the entropy is maximised by the uniform distribution.
+If the information available is instead the expected value -- the mean -- entropy is maximised by the  exponential distribution. Of particular importance is 
+the case in which, in addition to the mean, also the variance is known. MAXENT leads to the Gaussian distribution. 
+We retain this particularly interesting from the foundational point of view, since it provides a deeper insight into the ubiquitous Gaussian distribution.
+It is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy.
+For this reason, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process.
+In some sense, we can interpret the central limit theorem as the natural `statistical' evolution toward a configuration that maximizes entropy.
+
+For this work, we are particularly interested in the mutli-dimensional case.
+Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realisation of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of auto-covariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution\cite{gregory_2005}:
+\begin{align}
+    p&\left((x_1, \ldots, x_n)\vert I\right) = \nonumber \\
+    &\frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right)\,. 
+\end{align}
+We have already seen that for a wide-sense stationary process the mean function is independent of time, hence redefined to be equal to zero without loss of generality, and the auto-covariance function is dependent only on the time lag $\tau \equiv t_i - t_j$.
+One can thus choose a sampling rate $\Delta t$ so that $C_{ij} = C((i-j)\Delta t)$.
+The autocovariance matrix thus becomes a Toeplitz matrix.
+Toeplitz matrices are asymptotically equivalent to circulant matrices and thus diagonalised by the discrete Fourier transform base\cite{Gray}.
+%R. M. Gray, Foundations and Trends⃝R in Communications and Information Theory 2, 155 (2005), URL https://doi.org/10.1561/0100000006.
+Some simple algebra shows that the time-domain multivariate Gaussian can be transformed into the equivalent frequency domain 
+probability distribution:
+\begin{align}
+p&\left((\tilde{x}_1, \ldots, \tilde{x}_n)\vert I\right) = \nonumber \\
+    &\frac{1}{\left(2 \pi \det S\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i}(\tilde{x}_i^2S^{-1}_{i} \right)\,,
+\end{align}
+where the matrix $S$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
+Many readers will recognised the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\cite{woodward}
+%P. M. Woodward, D. W. Fry, and W. Higinbotham, Probability and information theory, with applications to radar; 2nd ed. (Pergamon Press, Oxford, 1964), URL http://cds.cern.ch/record/2031792.
+and of gravitational waves data analysis, e.g. \cite{finn:1992}.
+% L. S. Finn, Phys. Rev. D46, 5236 (1992), gr-qc/9209010
+Thanks to MAXENT, the problem of defining the probability distribution describing a wide-sense stationary process is thus 
+entirely reduced to the estimation of the PSD or, equivalently, the auto-covariance function.
+
+%
+%
+
 %This interpretation is straightforward: if we compare it with Eq.~(\ref{eq:information}), we see that the entropy is defined as the 'average information'.
@@
 %\end{equation}
-\wdp{I need to review and correct the following}
-To show how the maximization of entropy, supplemented with additional evidence, can provide a unique probability assignment, 
-we use a simple, pedagogical example. Consider the case in which we want assign a probability distribution $p(x)$ to a variable $x$ 
-for which we know mean $\mu$ and covariance $\Sigma$. Our `evidence' is then coded as the following constraints:
-\begin{align}
-$\sum_{i = 1}^{N}p_i &= 1 $\\
-$\sum_{i = 1}^{N}x_i p_i &= \mu $\\
-$\sum_{i = 1}^{N}(x_i - \mu)(x_j - \mu) p_i &= \Sigma$\,
-\end{align}
-where the first row states the requirement of exhaustivity.
-\wdp{i think we want the multivariate case since i) the gaussian case is also on wikipedia ii) we actually will need that  later on.}
+%\wdp{I need to review and correct the following}
+%To show how the maximization of entropy, supplemented with additional evidence, can provide a unique probability assignment, 
+%we use a simple, pedagogical example. Consider the case in which we want assign a probability distribution $p(x)$ to a variable $x$ 
+%for which we know mean $\mu$ and covariance $\Sigma$. Our `evidence' is then coded as the following constraints:
+%\begin{align}
+%\sum_{i = 1}^{N}p_i &= 1 \\
+%\sum_{i = 1}^{N}x_i p_i &= \mu \\
+%\sum_{i = 1}^{N}(x_i - \mu)(x_j - \mu) p_i &= \Sigma\,
+%\end{align}
+%where the first row states the requirement of exhaustivity.
+%\wdp{i think we want the multivariate case since i) the gaussian case is also on wikipedia ii) we actually will need that  later on.}
 
@@
 %\end{equation}
-
-Take a vector  $\vec x(t)$ with components in $\mathbb{R}^N$
-\begin{equation*}
-    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
-\end{equation*}
-We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals.
+%
+%Take a vector  $\vec x(t)$ with components in $\mathbb{R}^N$
+%\begin{equation*}
+%    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
+%\end{equation*}
+%We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals.
 If we put constraints in both mean $\vec \mu(t)$ and covariance matrix $C_(t_i, t_j)$, MAXENT allow us to assign for $\vec x(t)$ a multivariate normal distribution \cite{gregory_2005}:
-\begin{align}
-    p&\left(\vec x(t)\vert I\right) = \nonumber \\
-    &\frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right). 
-\end{align}
-
-Thus MAXENT provide a deeper insight into the Gaussian distribution.
-It is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy.
-For this reason, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process.
-In some sense, we can interpret the central limit theorem as the natural `statistical' evolution toward a configuration that maximizes entropy.
+%\begin{align}
+%    p&\left(\vec x(t)\vert I\right) = \nonumber \\
+%    &\frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right). 
+%\end{align}
 

### 2021-02-18 paper/mesa.tex
Partial restyling of validation

@@
 \section{Validation of the model}
+It is now clear that MESA algorithm provides a recursive formula for computing the coefficients $a_k$ in eq.~\ref{eq:MESApsd}.
+The number $N$ of such coefficients is the same as the number of time steps considered in the computation of the autocorrelation $\bar{r}_n$.
+In a perfect scenario, this would be equal to the number of points the autocorrelation is computed at (equivalent to the length of data considered).
+However, the computation of high order coefficients of the autocorrelation is unstable and for high enough $n$, the the estimation for  $\bar{n}_n$ shows a very high variance, broadly scaling as $\sim \left(\sqrt{N - n}\right)^{-1}$.
+
+It is then clear that the choice of how many points of the autocorrelation to consider is important: on one hand it is advisable to include as many knowledge of the autocorrelation as possible, leading to include all the known $\bar{r}_n$; on the other hand, including values of the autocorrelation that are not reliably estimated, can be counterproductive.
+Indeed, the order $N$ of the autocorrelation to be considered (or equivalently the order $N$ of the underlying autoregressive process) is the only input required by the user and a careful balance between this two necessities must be made when applying the algorithm.
+
+This section is devoted to an extensive study on how to make such choice.
+First of all (\ref{sec:optimizers}), we are going to define three different ``metrics" to measure how well the algorithm is performed: we will refer to them as \textit{optimizers}.
+The basic idea is to monitor, as the autoregressive order considered increase, the performance of the algorithm (in some sense) and pick, among the orders the one that yields better results.
+Second, the performance of different optimizers will be assessed by performing the spectral analysis on different time series: a simple white gaussian noise (\ref{sec:gaussian_noise_validation}) and a more complicate PSD (\ref{sec:arp_validation}).
+In a last subsection~\ref{sec:LIGO_validation} real data from LIGO-Virgo gravitational waves observatory.
+
 \subsection{Choosing an autoregressive order}\label{sec:optimizers} 
@@
 The following chapter will be devoted to the study of the statistical properties of the previous optimizers to understand which choice provides the best quality in the reproduction of some known power spectral densities.
-\section{Analysis of the optimizers}
-\section{Analysis of optimizers - Introduction}
+\subsection{Choosing an optimizer - Gaussian Noise} \ref{sec:gaussian_noise_validation}
 To apply Burg's method, it is a fundamental task to understand what is the best choice for estimating the order for the autoregressive stochastic process.
@@
 
-\section{Gaussian noise} 
 This section is devoted to the analysis of the optimizations methods, applied on pure Gaussian noise process $\vec n = (n_1, n_2, \dots, n_N),\text{ with } n_i = n(t_i)$, assuming our process to be a wide-sense stationary process, i.e.
@@
 We choose to sample both real and imaginary part from two identical normal distribution distributing variance in a proper way between real and imaginary components.
-\section{Test on Time Series}
+\subsection{Choosing an optimizer - AR(p) Time Series} \label{sec:arp_validation}
 It is now possible to test the three methods on a given time series.
@@
 We will then study in depth the Advanced LIGO design sensitivity power spectrum \cite{Ligo}
-\subsection{Normal Power Spectral Density}
+
 \begin{figure}
@@
 \end{figure}
-\subsection{Summary and Comparison}
+\subsubsection{Conclusive Remarks}
 \begin{figure}
@@
 
-\subsection{Analysis of Ligo Spectrum}
+\subsection{Analysis of Ligo Spectrum} \label{sec:LIGO_validation}
 \begin{figure}
@@
 \end{figure}
+
+\section{Comparison with Welch method}
+
+\section{Final remarks and future prospects}
 

### 2021-02-18 scripts/mem_welch_comparison.py
Added init options

@@ -25,59 +25,73 @@ def relative_error(real, estimate):
 
 if __name__ == '__main__': 
     import os 
-    save_dir = os.getcwd() + '/comparisons_2/'
+    save_dir = os.getcwd() + '/alessandro_fake_comparisons/'
+    seglen_factor = 1
+    save = False 
+    white_noise = False 
+    Ligo_noise = True 
+    
     #Sampling variables 
     dt = 1. / 4096
-    times = np.array([1, 5, 10, 100])#, 1000])
-    N = times // dt 
-    segment_length = np.array([521, 1024, 2048, 8192])  #, 32768])
     Ny = 0.5 / dt 
+    times = np.array([1, 5, 10, 100, 1000])
+    N = times // dt 
+    segment_length = np.array([521, 1024, 2048, 8192, 32768]) 
+    segment_length *= seglen_factor
+
     
     #Generate normal PSD and import Ligo PSD 
-    wnoise_frequency, wnoise_spectrum = np.arange(0, N[0] // 2 + 1) / times[0], np.repeat(1, N[0] // 2 + 1) * dt
-    ligo_frequency, ligo_spectrum = np.loadtxt('LIGO-P1200087-v18-AdV_DESIGN_psd.dat', unpack=True)     
-    #Generating Noise 
-    print('Generating white noise series')
-    Wtime, Wtime_series, Wfrequency, Wfrequency_series, Wpsd = generate_data(wnoise_frequency,
+    if white_noise: 
+        wnoise_frequency, wnoise_spectrum = np.arange(0, N[0] // 2 + 1) / times[0], np.repeat(1, N[0] // 2 + 1) * dt
+        print('Generating white noise series')
+        Wtime, Wtime_series, Wfrequency, Wfrequency_series, Wpsd = generate_data(wnoise_frequency,
                                                                         wnoise_spectrum,
                                                                         times[-1],
                                                                         1/ dt)
-    print('Generating Ligo noise series')
-    Ltime, Ltime_series, Lfrequency, Lfrequency_series, Lpsd = generate_data(ligo_frequency,
+        w_interp = interp1d(wnoise_frequency, wnoise_spectrum) 
+        
+    #ligo_frequency, ligo_spectrum = np.loadtxt('LIGO-P1200087-v18-AdV_DESIGN_psd.dat', unpack=True)     
+    if Ligo_noise: 
[…]

### 2021-02-18 paper/mesa.tex

@@
 It is now clear that MESA algorithm provides a recursive formula for computing the coefficients $a_k$ in eq.~\ref{eq:MESApsd}.
-The number $N$ of such coefficients is the same as the number of time steps considered in the computation of the autocorrelation $\bar{r}_n$.
+The number $M$ of such coefficients is the same as the number of time steps considered in the computation of the autocorrelation $\bar{r}_m$.
 In a perfect scenario, this would be equal to the number of points the autocorrelation is computed at (equivalent to the length of data considered).
-However, the computation of high order coefficients of the autocorrelation is unstable and for high enough $n$, the the estimation for  $\bar{n}_n$ shows a very high variance, broadly scaling as $\sim \left(\sqrt{N - n}\right)^{-1}$.
-
-It is then clear that the choice of how many points of the autocorrelation to consider is important: on one hand it is advisable to include as many knowledge of the autocorrelation as possible, leading to include all the known $\bar{r}_n$; on the other hand, including values of the autocorrelation that are not reliably estimated, can be counterproductive.
-Indeed, the order $N$ of the autocorrelation to be considered (or equivalently the order $N$ of the underlying autoregressive process) is the only input required by the user and a careful balance between this two necessities must be made when applying the algorithm.
+However, the computation of high order coefficients of the autocorrelation is unstable and for high enough $n$, the the estimation for  $\bar{n}_m$ shows a very high variance, broadly scaling as $\sim \left(\sqrt{M - m}\right)^{-1}$.
+
+%OR WE CAN KEEP THIS (SOMEHOW MODIFIED)
+%Before being able to apply MESA, we need to address the question of the best length $M$ of the filter that provides the optimal estimate of the power spectrum.
+If the autocorrelation matrix was perfectly know, the answer would be simply $M = N$.
+%In real cases, such a knowledge is not available and one has to estimate the autocorrelation function from the data.
+The accuracy for the estimate decreases as the time lag $k$ increases as $\sim \left(\sqrt{N - k}\right)^{-1}$.
+%The accuracy of the estimate should then increase as the length of the autocorrelation function increases, but, from a certain point on, inaccuracy in the estimate for $R$ becomes so large that the estimate for the spectrum is negatively affected by this fact.
+
+It is then clear that the choice of how many points of the autocorrelation to consider is important: on one hand it is advisable to include as many knowledge of the autocorrelation as possible, leading to include all the known $\bar{r}_m$; on the other hand, including values of the autocorrelation that are not reliably estimated, can be counterproductive.
+Indeed, the order $M$ of the autocorrelation to be considered (or equivalently the order $M$ of the underlying autoregressive process) is the only input required by the user and a careful balance between this two necessities must be made when applying the algorithm.
 
@@
 
+\sschmidt{Qui ho provato a mettere un po' di ordine alla sezione.
+Ho scritto l'introduzione sopra per spiegare i passaggi logici della validation.
+Spero che vada bene.
+La parte sotto è tutta da rivedere e probabilmente c'è molto da asciugare: sono quasi 10 pagine!
+Mi metto a farlo subito}
+
 \subsection{Choosing an autoregressive order}\label{sec:optimizers} 
-Before being able to apply MESA, we need to address the question of the best length $M$ of the filter that provides the optimal estimate of the power spectrum.
-If the autocorrelation matrix was perfectly know, the answer would be simply $M = N$.
-\\
-In real cases, such a knowledge is not available and one has to estimate the autocorrelation function from the data.
-The accuracy for the estimate decreases as the time lag $k$ increases as $\sim \left(\sqrt{N - k}\right)^{-1}$.
-\\
-The accuracy of the estimate should then increase as the length of the autocorrelation function increases, but, from a certain point on, inaccuracy in the estimate for $R$ becomes so large that the estimate for the spectrum is negatively affected by this fact.
-A compromise between the two behaviour is proposed to be \cite{doi:10.1190/1.1440902}
+\sschmidt{optimizer è la parola giusta? Optimizer è un algoritmo... Si potrebbe chiamare optimizer ``metrica"?}
+A good choice for the autoregressive order $M$ is proposed to be \cite{doi:10.1190/1.1440902}
 \begin{equation}\label{eq:MMAx}
@@
 \end{equation}
-Still, this is not a satisfactory answer since it is an upper limit for the estimate of the length $M$.
-Various approaches to optimize the order $M$ have been proposed in literature; in the following we present them.
+However, this is just an upper limit for the estimate of the length $M$ and the optimal algorithm could employ much less points.
+We then need a more sophisticated method for computing the right value for $M$.
+Various optimizers to assess the algorithm performance have been proposed in literature. We summarize them below:
+
 \begin{itemize}
@@
 \begin{equation}
-    FPE: \min_{M}E\left((x_t - \hat x_t) ^ 2\right)
-\end{equation}
-with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.  This criterion is equivalent to the choice of $M$ such that 
-\begin{equation}
-    \min_{M}P_{m} \frac{N + m + 1}{N - m - 1}
-\end{equation}
-with $P_M$ defined by the prediction error filter equation. 
+    FPE(m) = \mathbb{E}\left[ \left((x_t - \hat x_t) ^ 2\right) \right]
+\end{equation}
+with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.
+Minimizing FPE is the same as minimizing the quantity: 
+\begin{equation}
+    P_{m} \frac{N + m + 1}{N - m - 1}
+\end{equation}
+with $P_M$ defined by the prediction error filter equation \sschmidt{What is this equation?}.
 \item \textbf{Criterion Autoregressive Transfer function (CAT)}
-This second criterion is due to Parzen \cite{bhansali1986}, and requires $M$ to be chosen as the length $m$ the minimizes the quantity 
-\begin{equation}
-    \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
+This second optimizer is due to Parzen \cite{bhansali1986}, and it is defined as
+\begin{equation}
+    CAT(m) = \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
 \end{equation}
@@
 The last criterion we will consider is the OBD (Rao, 1982)\cite{doi:10.1029/WR018i004p01097}.
-With the technique of asymptotic integration, $M$ is proposed to be the length that maximizes the posterior distribution, condition met selecting $m$ that minimize
-   \begin{equation}
-        (N - m - 2) \log(P_m) + m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2.
+With the technique of asymptotic integration, th optimizer is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}:
+   \begin{equation}
+        OBD(m) = (N - m - 2) \log(P_m) + m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2.
     \end{equation}
 \end{itemize}
-Once an optimizer is chosen to select the best recursion order $M$, we then proceed as follows:  we solve the Levinson recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached.
+
+Once an optimizer is chosen, the choice of the best recursion order is straightforward: we solve the Levinson recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached.
 Then, the order $M$ is selected to be the one that minimize the chosen optimizer value.
-\\
-PARLARE QUI DELL'EARLY STOP? 
-The following chapter will be devoted to the study of the statistical properties of the previous optimizers to understand which choice provides the best quality in the reproduction of some known power spectral densities.
-\subsection{Choosing an optimizer - Gaussian Noise} \ref{sec:gaussian_noise_validation}
+
+In a real implementation of the algorithm, computing all the recursion up to $M_{max}$ can result in a significant waste of computational power: the optimal value is often $M_{opt}<< M_{max}$ and in such cases, computing all the values of $m$ until $M_{max}$ is not useful.
+In practice, we can apply an \textit{early stopping} procedure: every few iterations we look for the best order of $M_{opt}$; if this value does not change for a while, the computation is stopped and a good (local) minimum of the optimizer is found.
+
+The following sections will be devoted to the study of the statistical properties of the optimizers above: we need to understand which choice provides the best quality in the reproduction of some known power spectral densities.
+
+\subsection{Choosing an optimizer - Gaussian Noise} \label{sec:gaussian_noise_validation}
 To apply Burg's method, it is a fundamental task to understand what is the best choice for estimating the order for the autoregressive stochastic process.
@@
 \section{Comparison with Welch method}
+
+\sschmidt{In questa sezione potremmo mettere i vari confronti con Welch}
+\sschmidt{Mettiamo da qualche parte una descrizione dettagliata del codice?}
 

### 2021-02-18 scripts/mem_welch_comparison.py
Added .show() for plots

@@ -42,20 +42,18 @@ if __name__ == '__main__':
     
     #Generate normal PSD and import Ligo PSD 
     if white_noise: 
-        wnoise_frequency, wnoise_spectrum = np.arange(0, N[0] // 2 + 1) / times[0], np.repeat(1, N[0] // 2 + 1) * dt
         print('Generating white noise series')
+        wnoise_frequency, wnoise_spectrum = np.arange(0, N[0] // 2 + 1) / times[0], np.repeat(1, N[0] // 2 + 1) * dt
         Wtime, Wtime_series, Wfrequency, Wfrequency_series, Wpsd = generate_data(wnoise_frequency,
                                                                         wnoise_spectrum,
                                                                         times[-1],
                                                                         1/ dt)
         w_interp = interp1d(wnoise_frequency, wnoise_spectrum) 
         
-    #ligo_frequency, ligo_spectrum = np.loadtxt('LIGO-P1200087-v18-AdV_DESIGN_psd.dat', unpack=True)     
     if Ligo_noise: 
-        ligo_frequency, ligo_spectrum, _ = np.loadtxt('GWTC1_GW150914_PSDs.dat', unpack = True)
-    #Generating Noise 
-   
         print('Generating Ligo noise series')
+        #ligo_frequency, ligo_spectrum = np.loadtxt('LIGO-P1200087-v18-AdV_DESIGN_psd.dat', unpack=True)     
+        ligo_frequency, ligo_spectrum, _ = np.loadtxt('GWTC1_GW150914_PSDs.dat', unpack = True)
         Ltime, Ltime_series, Lfrequency, Lfrequency_series, Lpsd = generate_data(ligo_frequency,
                                                                              ligo_spectrum,
                                                                              times[-1],
@@ -92,7 +90,8 @@ if __name__ == '__main__':
             ax2.set_xlim(ligo_frequency.min(), ligo_frequency.max())
             if save: 
                 fig2.savefig(save_dir +'GW150914 {} points ({} seglen).pdf'.format(int(N[i]), seglen_factor))
-        
+                
+        plt.show() 
 
     
     

### 2021-02-18 scripts/mem_welch_comparison.py
Added option to multiply seglen

@@ -37,7 +37,7 @@ if __name__ == '__main__':
     times = np.array([1, 5, 10, 100, 1000])
     N = times // dt 
     segment_length = np.array([521, 1024, 2048, 8192, 32768]) 
-    segment_length *= seglen_factor
+    if seglen_factor != 1: segment_length *= seglen_factor
 
     
     #Generate normal PSD and import Ligo PSD 

### 2021-02-19 paper/mesa.tex

@@
 The basic idea is to monitor, as the autoregressive order considered increase, the performance of the algorithm (in some sense) and pick, among the orders the one that yields better results.
-Second, the performance of different optimizers will be assessed by performing the spectral analysis on different time series: a simple white gaussian noise (\ref{sec:gaussian_noise_validation}) and a more complicate PSD (\ref{sec:arp_validation}).
+Second, the performance of different optimizers will be assessed by performing the spectral analysis on time series with given analytical PSD, sec.~\ref{sec:arp_validation}.
 In a last subsection~\ref{sec:LIGO_validation} real data from LIGO-Virgo gravitational waves observatory.
@@
 Spero che vada bene.
-La parte sotto è tutta da rivedere e probabilmente c'è molto da asciugare: sono quasi 10 pagine!
-Mi metto a farlo subito}
+La parte sotto è ancora da rivedere e probabilmente c'è molto da asciugare: sono quasi 10 pagine!}
 
@@
 With the technique of asymptotic integration, th optimizer is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}:
-   \begin{equation}
-        OBD(m) = (N - m - 2) \log(P_m) + m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2.
-    \end{equation}
+   \begin{align}
+        OBD&(m) = (N - m - 2) \log(P_m) \nonumber\\
+        &+ m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2.
+    \end{align}
 \end{itemize}
@@
 
-\subsection{Choosing an optimizer - Gaussian Noise} \label{sec:gaussian_noise_validation}
-To apply Burg's method, it is a fundamental task to understand what is the best choice for estimating the order for the autoregressive stochastic process.
-Before applying it to real datasets, we will study main properties of the three different optimizer, introduce in section \ref{sec:optimizers} on distinct, simulated, realizations of some AR(p) processes.
-We will generate different Gauss noise time series with a given known power spectral density (PSD) to understand how each method performs in different situation.
-The analysis will concern the quality, to be defined later, of the reconstruction of the spectrum itself.
-We will consider two different kinds of spectra: i) a Gaussian spectrum with a given mean and variance; ii) an Advanced LIGO design sensitivity theoretical spectral curve.
-As a part of our investigation, we will consider both ensemble-averaged, indicated with $\bar S$ and non averaged quantities.
-\\
-In either cases, we will quantify the agreement between the reconstructed quantities and the expected ones.\\
-A short description of the principle of the algorithm used to construct the time series is also given.
-
-This section is devoted to the analysis of the optimizations methods, applied on pure Gaussian noise process $\vec n = (n_1, n_2, \dots, n_N),\text{ with } n_i = n(t_i)$, assuming our process to be a wide-sense stationary process, i.e.
-autocorrelation matrix is a Toeplitz matrix depending only on time intervals.\\
-The spectrum chosen for our first analysis is a Gaussian distribution
-\begin{equation}
-    \nonumber
-    S(f) = \frac{1}{\sqrt{2 \pi} \sigma}e^{-\frac{(f - \mu)^2}{2 \sigma ^2}}
-\end{equation}
-with mean $\mu =2.5 \quad a.u.$ and standard deviation $\sigma =  0.5 \quad  a.u.$. 
-To generate noise with a power spectrum that is known, we have to first understand what is the role played by the PSD. \\ 
-Via MAXENT principle, we can state that noise in normally distributed with 0 mean and some covariance matrix $C_{ij} = C(t_i, t_j)$, i.e.
-\begin{equation}
-    \nonumber
-    p(\vec{n}(t) \vert I) \propto \exp{\left\{-\frac{1}{2}n_i C^{-1}_{ij}n_j\right\}},
-\end{equation}
+\subsection{Choosing an optimizer - A gaussian PSD} \label{sec:arp_validation}
+We test the performance of the three optimizers on a random time series with a known power spectral density.
+We take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$, see fig.~\ref{fig:autocorr}.
+Of course, the time series will be generated by sampling from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$ and performing the inverse Fast Fourier Transform on the frequency series.
+
+We will apply Burg method with each optimizer on a number of simulated time series, obtaining a variety of estimate for PSD.
+We will study statistical properties for both single reproductions and ensemble average. In either cases, we will quantify the agreement between the reconstructed quantities and the expected ones.
+In many realistic cases, such ensemble average is not accessible (one car rarely observe multiple time series realizations for the \textit{same} process), so that single reproduction studies are probably more useful.
+Still, study of ensemble average is important to highlight some interesting properties.
+In this way we can fully characterize the performance of each optimizer .
+n what follows, we will indicate with $\bar S$ any ensemble averaged quantity.
+
+We will attentively study the role of the filter length estimate M in determining the accuracy of the reconstruction.
+We will consider every optimizer separately and compare result only at the end of this section.
+In section~\ref{sec:LIGO_validation}, we will perform the analysis again using, for the purpose of generating the time series, the Advanced LIGO design sensitivity PSD \cite{Ligo}.
+
 \begin{figure}
@@
 	\label{fig:autocorr}
-\end{figure}and, since process has zero mean, autocovariance and autocorrelation matrices are identical. \\ 
-Using the properties of Toeplitz matrices (see appendix \ref{app:Toeplitz}) and the relation between autocorrelation and power spectral density, it is immediate to verify that, in the limit of large number of time samples, the distribution of noise in frequency domain is given by:
-\begin{equation}
-    p(\tilde{n}(f_i) \vert I) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}},
-\end{equation}
-so that the role of PSD appears to be clear: it is the diagonal covariance matrix for the noise distribution in Fourier Domain: \\
-\begin{equation}
-    \nonumber
-    <\tilde n_i \tilde n_j > = S(f_i)\delta_{ij}.
-\end{equation}
-With this information, it is easy to generate any time series with power spectral density of any shape.
-Choosing a priori values for sampling interval $dt$ and total time of observation $T$, recalling Nyquist theory of sampling, we generate arrays of given length N / 2 in frequency domain from 0 to the Nyquist Frequency $f_{Ny} = \frac{1}{2 dt}$.
-We then convert noise to the time domain using an Inverse Fast Fourier Transform.\cite{oliphant2006guide}\cite{van2011numpy} \\ Basic idea for fast algorithm used to construct noise can be found in ~\cite{noiseGen}.
-We worked it out in a slightly different way, the imaginary component is not obtained sampling a random complex phase from a uniform distribution.
-We choose to sample both real and imaginary part from two identical normal distribution distributing variance in a proper way between real and imaginary components.
-\subsection{Choosing an optimizer - AR(p) Time Series} \label{sec:arp_validation}
-It is now possible to test the three methods on a given time series.
-We will concentrate on a power spectral density that is a normal distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$.
-We will apply Burg method with each optimizer on a large amount of simulated dataset, obtaining a variety of estimate for PSD.
-We will study statistical properties for both single reproductions and ensemble average.
-In many cases, such as gravitational waves, one cannot observe multiple time series realizations for a fixed process, so that single reproduction studies are probably more useful.
-Still, study of ensemble average is important to highlight some interesting properties.
-In this way we can fully characterize the performance of each optimizer .
-We will attentively study the role of the filter length estimate M in determining the accuracy of the reconstruction.
-We will consider every optimizer separately and compare result only at the end of this section.
-We will then study in depth the Advanced LIGO design sensitivity power spectrum \cite{Ligo}
+\end{figure}
 
@@
 \end{figure}
-Fig \ref{fig:noise} shows a realization of the gaussian process in the frequency (left panel) and time domain (right panel).   \\ 
-To study noise with normal-shaped PSD, we generate 1000 datasets of 3000 points each.
-This will provide a number of samples that is large enough to perform statistical analysis.
-For every dataset we will reproduce its spectrum with every optimization method.
-We will show that the quality of reconstruction is very different in different parts of the spectrum: the minimum of the error will be achieved around the peaks, increasing fastly in the tails
- \\ \\ 
+
+For our studies, we generate a dataset of 1000 time series of 3000 points each: see fig.~\ref{fig:noise} for an example of a time series in frequency and time domain.
+For every time series we will compute the spectrum with every optimization method.
+
+\sschmidt{Sono arrivato fino a qui nelle modifiche}
+
 \subsubsection{Final Prediction Error (FPE)}
@@
 
-\subsection{Analysis of Ligo Spectrum} \label{sec:LIGO_validation}
+\subsection{Choosing an optimizer - LIGO Spectrum} \label{sec:LIGO_validation}
 \begin{figure}

### 2021-02-19 paper/mesa.tex

@@
 \subsection{Maximum Entropy Principle} \label{sec:MAXENT}
-\sschmidt{This section is yet to prune}
+
 Before introducing MAXENT as a variational problem, we will workout some simple examples to contexualise how information's entropy
@@
 \subsection{Maximum Entropy Spectral Analysis} \label{sec:MESA}
-The computation of the power spectral density is an important component in time series analysis and characterization.
+
 In principle, if the autocorrelation was known exactly (i.e. at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD 
@@
 is impossible.
-Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases with $k$ as $\sigma \sim \frac{1}{\sqrt{N - k}}$, so that only a few values for the autocorrelation function can actually be computed reliably.
-This bring us the core of our problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function?
-The Maximum Entropy Principle can guide use in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method.
+Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases with $k$ as $\sigma \sim 1/\sqrt{N - k}$, so that only a few values for the autocorrelation function can actually be computed reliably.
+This bring us the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function?
+MAXENT can guide use in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method.
 The latter assumes that the unknown values of the autocorrelation are $0$.
@@
 \par
-As in the previous examples, we need to set up a variational problem where the entropy eq.~\ref{eq:entropy} is maximized with some problem-specific constraints. 
+As in the previous examples, one needs to set up a variational problem where the entropy, Eq.~(\ref{eq:entropy}), is maximized 
+subject to some problem-specific constraints. 
 In our case, they are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation (wherever an estimate of this is available).
 \par
-Before doing so, there is a technical problem to solve: our definition of entropy (eq.~\ref{eq:entropy}) depends on a probability distribution, not on the power spectrum.
-We then need to to formulate the variational problem in terms of the power spectral density $S(f)$ alone.
-A way out of this is considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA}.
-One can show that the 'entropy gain' obtained filtering white noise is:
+Before doing so, there is a technical problem to solve: the definition of entropy (Eq.~\ref{eq:entropy}) depends on a probability distribution, 
+not on the PSD. The variational problem can be formulated in terms of the power spectral density $S(f)$ alone by
+considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA}. 
+The entropy gain obtained by filtering white noise is: 
 \begin{equation}\label{eq:EntropyGain}
@@
 \end{equation}
-so that we can reformulate the variational problem requiring $S(f)$ to maximize the entropy gain \sschmidt{Non ho capito questa parte...:(}.
-In real examples we have no access to all frequencies but only on those between minus and plus the Nyquist frequency, defined from the sampling interval $\Delta t$ as
-\begin{equation}
-    Ny = \frac{1}{2 \Delta t}.
-\end{equation}
-All our quantities are well defined in this range only, so that we are not going to maximize \ref{eq:EntropyGain}, but
+hence the variational problem can be stated by requiring that $S(f)$ maximises the entropy gain \sschmidt{Non ho capito questa parte...:(}. 
+Time-limited observations do not provide access to the full frequency range, but only between minus and plus the Nyquist frequency, 
+that, in turn is given by the sampling interval $\Delta t$ as $Ny \equiv \frac{1}{2 \Delta t}$. Hence, the finite-size entropy
+gain is 
 \begin{equation}\label{eq:EntropyGain2}
-    \Delta H = \int_{- Ny}^{Ny}\log S(f) df.
-\end{equation}
-
-Now the the objective to maximize is well defined, we can write down the constraints of our problem.
-Defining $\bar r_n$ as the sample autocorrelation, we require that the variational solution $S(f)$ for the PSD matches the empirical autocorrelation. This means:
+    \Delta H = \int_{- Ny}^{Ny}\log S(f) df\,.
+\end{equation}
+
+Consider a realisation of a stochastic process $(x_1,\ldots,x_n)$ with sample autocorrelations $\bar r_i,\,i=0,\ldots, n/2$, we require that the variational solution $S(f)$ for the PSD matches the empirical autocorrelation:
 \begin{equation}\label{eq:MaxConstraint}
-\int_{-Ny}^{Ny} S(f) e^{\imath 2 \pi f n \Delta t} df = \bar r_{n}.
+\int_{-Ny}^{Ny} S(f) e^{\imath 2 \pi f n \Delta t} df = \bar r_{n}\,.
 \end{equation}
@@
 It requires no assumptions on the unavailable estimates for $\bar r_n$ and is consistent with the sample autocorrelation function by construction.
-Remarkably, it admits an analytical solution in a closed expression for $S(f)$.
+Remarkably, the variational problem admits a closed-form analytical expression for $S(f)$:
 
 \begin{equation}\label{eq:MESApsd}
-    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^z\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}, 
-\end{equation}
-with $\Delta t$ the sampling interval $a_0$ = 1 and the coefficients $P_N$ and $a_i (i \neq 0)$ to be determined.
+    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^s\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}, 
+\end{equation}
+with $\Delta t$ the sampling interval, $z=\exp{(2\pi i f)}$, $a_0$ = 1 and the coefficients $P_N$ and $a_i (i \neq 0)$ to be determined.
 The details of the derivation and the actual values of the coefficients $a_i$ can be found in appendix \ref{sec:MESA_solution}.
@@
 \subsection{Autoregressive Process Analogy} \label{sec:autoregr}
+
 The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the study
 of autoregressive processes.\cite{doi:10.1029/RG013i001p00183}.
-An autoregressive stationary process of order $p$ (AR(p)) is a time series whose values satisfy the following expression: 
+An autoregressive stationary process of order $p$ (AR($p$)) is a time series whose values satisfy the following expression: 
 \begin{equation} \label{eq:AR_p}
@@
 where $b_1, \ldots, b_p$ are real coefficients and $\nu_t$ is white noise with a given variance $\sigma^2$.
-Thus, an AR(p) process models the dependence on every point on the $p$ past observation, thus being potentially able to model complex autocorrelation between data.
+Thus, an AR($p$) process models the dependence of the value of the process at time $t$ on every past $p$ observations, 
+thus being potentially able to model complex autocorrelation between data.
 \par
 To show the analogy, we compute the PSD of such process and we show that it is formally equivalent to the PSD obtained in eq.~\ref{eq:MESApsd}.
-We start taking the z \sschmidt{What is the z transform} transform of the previous eq.~\ref{eq:AR_p}: 
+We start taking the z \sschmidt{What is the z transform} \wdp{it is a relative of the Fourier transform, z are the roots of the unit in the complex plane} transform of the previous eq.~\ref{eq:AR_p}: 
 \begin{align}
@@
 The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.
-Equation~\ref{eq:ARspectrum} and eq.~\ref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
+Eq.~(\ref{eq:ARspectrum}) and Eq.~(\ref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
 This shows that the MAXENT estimation of the PSD computes the coefficients of the autoregressive process that is able to model the data faithfully.

### 2021-02-19 paper/mesa.tex
Change in definition of z

@@
 \end{equation}
-with $\Delta t$ the sampling interval, $z=\exp{(2\pi i f)}$, $a_0$ = 1 and the coefficients $P_N$ and $a_i (i \neq 0)$ to be determined.
+with $\Delta t$ the sampling interval, $z=e^{-2\pi i f \Delta t}$, $a_0$ = 1 and the coefficients $P_N$ and $a_i (i \neq 0)$ to be determined.
 The details of the derivation and the actual values of the coefficients $a_i$ can be found in appendix \ref{sec:MESA_solution}.

### 2021-02-21 paper/mesa.tex

--- 2021-02-19 paper/mesa.tex
@@
 \end{equation}
-with $\Delta t$ the sampling interval, $z=\exp{(2\pi i f)}$, $a_0$ = 1 and the coefficients $P_N$ and $a_i (i \neq 0)$ to be determined.
+with $\Delta t$ the sampling interval, $z=\exp{(2\pi i f\Delta t)}$, $a_0$ = 1 and the coefficients $P_N$ and $a_i (i \neq 0)$ to be determined.
 The details of the derivation and the actual values of the coefficients $a_i$ can be found in appendix \ref{sec:MESA_solution}.

### 2021-02-22 paper/mesa.tex
Added some text

@@
 We take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$, see fig.~\ref{fig:autocorr}.
-Of course, the time series will be generated by sampling from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$ and performing the inverse Fast Fourier Transform on the frequency series.
+Of course, the time series will be generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$ and performing the inverse Fast Fourier Transform on the frequency series.
 
@@
 For every time series we will compute the spectrum with every optimization method.
+To assess the magnitude of the disagreement, we compute the relative error between the mean spectrum $\hat{S}$ and the true (analytical) spectrum $S$ for every frequency bin:
+\begin{equation}
+r(f) = \frac{|\hat{S}(f) - S(f)|}{S(f)}\; .
+\end{equation}
+
+We report the average values of the relative error for the different optimizers in Table~\ref{tab:relative_errors}.
 
 \sschmidt{Sono arrivato fino a qui nelle modifiche}
-
-\subsubsection{Final Prediction Error (FPE)}
-The estimate of the recursive order obtained with FPE on the 1000 realization of noise, is found to lay between 30 and 70.
-Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum: a logarithmic plot of the mean of all the spectra versus the true spectrum is reported in \textbf{Figure \ref{fig:FPEmean}} with 90\% credible regions.
-It is evident that, despite not too large error bars, the tail of the normal distribution is not comprised in the interval of the estimation.
+\begin{table}
+	\caption{Average relative error for different optimizers, computed in different ways. We consider a set of $1000$ realization of a noisy time series of $3000$ points.
+	We report the average error  $\bar{r}_{\bar{S}}$ on the whole set of simulations, the error $\bar{r}_{\bar S}^{\mu \pm 3\sigma} $ averaged only on data points between $3\sigma$ and the error $\bar{r}$ on a single, randomly drawn realization.
+	\sschmidt{I think we need to discuss on which performance error to include in the estimation}
+	}
+	\label{tab:relative_errors}
+	\def\arraystretch{1.5}
+	\begin{ruledtabular}
+		\begin{tabular}{ l c c c }
+			\multirow{2}{*}{Optimizer}& \multicolumn{3}{c}{Relative Error}\\
+			& $\bar{r}_{\bar{S}}$	& $\bar{r}_{\bar S}^{\mu \pm 3\sigma} $ & $\bar{r}$\\
+			\hline \hline
+			FPE 			& $6.9 \; (46.9\%)$	& $7 \; (1.6\%)$ & \\ 
+			\hline
+			CAT 			& $4.5 \; (30.6\%)$	& $194 \; (45.3\%)$ & \\
+			\hline
+			OBD 				& $1.7 \; (11.6\%)$	& $206 \; (48.1\%)$ & \\
+		\end{tabular}
+	\end{ruledtabular}
+\end{table}
+
 \begin{figure}[H]
@@
 \end{figure}
-To assess the magnitude of the disagreement, we compute the relative error between the mean spectrum $\hat{S}$ and the true spectrum $S$ for every frequency bin
-\begin{equation}
-    \nonumber
-    r(f) = \frac{|\hat{S}(f) - S(f)|}{S(f)}\,.
-\end{equation}
-We found this to be a better choice than residuals since relative errors are not dependent on the magnitude of the data.
-Second Fig \ref{fig:FPEmean} (below) shows that the largest error in the reconstruction of the spectrum is in the tails, while in the bulk of the distribution we have a good approximation of the true function.
-For the entire spectrum, the mean relative error $\Bar{r}$ is 
-\begin{equation}
-    \nonumber
+
+\subsubsection{Final Prediction Error (FPE)}
+The results of our investigation on te performance of FPE are reported in fig.~\ref{fig:FPEmean}: we plot the ``true" spectrum, the reconstructed spectrum as well as the $90\%$ credible region of the predicted spectrum.
+Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum; however, we note that the reconstruction is not very accurate in the low frequency region.
+
+For the entire spectrum, the mean relative error $\bar{r}$ is 
+\begin{equation}  
     \bar{r}_{\bar S, FPE} = 0.025.
@@
 
-\sschmidt{In questa sezione potremmo mettere i vari confronti con Welch}
+It is really interesting to have a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
+We perform such comparison on the spectrum of LIGO (even though similar conclusions can be drawn from any other shape of the PSD).
+
+We perform two sets of experiments.
+In the first one (see fig.~), we simulate data from an analytical PSD and we make the comparison with them: this is to endure the we have a baseline PSD to compare the data with.
+In a second experiment (see fig.~), we employ the data released to the public byt the LIGO/Virgo collaboration.
+In each experiment, we vary the length of the data used for the estimation: this is also useful to assess how the computation depends on the data available.
+We set the total observation time $T = 1, 5, 10, 100, 1000 \SI{}{s}$
+For the MESA algorithm, we employ FPE optimizer.
+For the Welch algorithm, we employ a Tukey window, an overlap fraction of $0.5$ for the segments and a length of segments $L = 512, 1024, 2048, 8192, 32768$ datapoints.
+In all cases, the sampling rate is set to $\SI{4096}{Hz}$.
+
+First of all, we note that using a longer time series results in a more stable estimation of the PSD, especially in the low frequency sector.
+This is somehow obvious: more data provide more information available for spectral estimation.
+This fact is more important at lower frequencies, where the natural time length is large.
+Turning to the comparison between MESA and Welch's method, we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
+has a large number of spurious peaks) rather than PSD provided by MESA.
+The different behavior is more manifest in the low frequency sector.
+The result does not change much by changing the hyperparameters required for Welch algorithm.
+\sschmidt{Other observations??}
+
+\begin{figure}
+	\label{fig:mem_welch_realdata}
+	\caption{Blablabla}
+\end{figure}
+
 \sschmidt{Mettiamo da qualche parte una descrizione dettagliata del codice?}

### 2021-02-22 scripts/optimisers_comparison.py
Code that compares loss functions behaviour

@@ -1,0 +1,150 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Feb 19 10:17:55 2021
+
+@author: alessandro
+"""
+import sys
+sys.path.insert(0, '..')
+from memspectrum import MESA 
+from GenerateTimeSeries import generate_data 
+import numpy as np
+import scipy.stats 
+import matplotlib.pyplot as plt
+from init_plotting import init_plotting
+
+def relative_error(real, estimate): 
+    return np.abs(real - estimate) / real 
+
+PSD = 'ligo' #normal or ligo 
+save = True 
+savedir = '../paper/Images/optimisers_comparison/' + PSD + '/'
+methods = ['FPE', 'CAT', 'OBD']
+M = MESA() 
+init_plotting() 
+colors = {'FPE': 'r', 'CAT': 'k', 'OBD': 'green'}
+
+#Initialize single spectrum dictionaries 
+spectra, optimisers, orders, errors = {}, {}, {}, {} 
+for method in methods: 
+    spectra[method], optimisers[method], orders[method], errors[method] = [], [], [], []
+
+#ensemble dictionaries 
+median, p5, p95, ensemble_error = {}, {}, {}, {}
+
+#Generating noise 
+if PSD.lower() == 'normal': 
+    number_of_simulations, dt, number_of_points = 1000, 1./10, 3000 
+    f = np.linspace(0, .5 / dt, number_of_points) 
+    spectrum = scipy.stats.norm.pdf(f, 2.5, .5)
+    
+elif PSD.lower() == 'ligo': 
+    number_of_simulations, dt, number_of_points = 500, 1./2048, 40960
+    f, spectrum = np.loadtxt('LIGO-P1200087-v18-AdV_DESIGN_psd.dat', unpack = True)
+    
+
+T = number_of_points * dt 
+
+
+for i in range(number_of_simulations): 
+    time, time_series, frequency, frequency_series, psd =\
+        generate_data(f, spectrum, T, 1 / dt)
+    for method in methods:
+        p, a_k, opt = M.solve(time_series, optimisation_method = method, early_stop = False)
+        optimisers[method].append(opt)
+        orders[method].append(M.a_k.size)
+        spectra[method].append(M.spectrum(dt)[0][: number_of_points // 2])    
+
+#Set unity measures 
+if PSD == 'ligo': 
+    y = r"$PSD \left(\frac{1}{Hz} \right)$"
+
+    
[…]

### 2021-02-24 scripts/optimisers_comparison.py
Can now load data

@@ -13,13 +13,15 @@ import numpy as np
 import scipy.stats 
 import matplotlib.pyplot as plt
 from init_plotting import init_plotting
+from scipy.interpolate import interp1d 
 
 def relative_error(real, estimate): 
     return np.abs(real - estimate) / real 
 
 PSD = 'ligo' #normal or ligo 
-save = True 
+save = False
 savedir = '../paper/Images/optimisers_comparison/' + PSD + '/'
+simulate_data = False 
 methods = ['FPE', 'CAT', 'OBD']
 M = MESA() 
 init_plotting() 
@@ -35,27 +37,38 @@ median, p5, p95, ensemble_error = {}, {}, {}, {}
 
 #Generating noise 
 if PSD.lower() == 'normal': 
-    number_of_simulations, dt, number_of_points = 1000, 1./10, 3000 
-    f = np.linspace(0, .5 / dt, number_of_points) 
-    spectrum = scipy.stats.norm.pdf(f, 2.5, .5)
-    
-elif PSD.lower() == 'ligo': 
-    number_of_simulations, dt, number_of_points = 500, 1./2048, 40960
-    f, spectrum = np.loadtxt('LIGO-P1200087-v18-AdV_DESIGN_psd.dat', unpack = True)
-    
+        number_of_simulations, dt, number_of_points = 1000, 1./10, 3000 
+        f = np.linspace(0, .5 / dt, number_of_points) 
+        spectrum = scipy.stats.norm.pdf(f, 2.5, .5)
 
+elif PSD.lower() == 'ligo':  
+        number_of_simulations, dt, number_of_points = 500, 1./2048, 40960
+        f, spectrum = np.loadtxt('LIGO-P1200087-v18-AdV_DESIGN_psd.dat', unpack = True)
+        
 T = number_of_points * dt 
 
+if simulate_data: 
+    for i in range(number_of_simulations): 
+        time, time_series, frequency, frequency_series, psd =\
+            generate_data(f, spectrum, T, 1 / dt)
+        for method in methods:
+            p, a_k, opt = M.solve(time_series, optimisation_method = method, early_stop = False)
+            optimisers[method].append(opt)
+            orders[method].append(M.a_k.size)
+            spectra[method].append(M.spectrum(dt)[0][: number_of_points // 2])    
 
-for i in range(number_of_simulations): 
-    time, time_series, frequency, frequency_series, psd =\
-        generate_data(f, spectrum, T, 1 / dt)
[…]

### 2021-02-24 memspectrum.py
Removed abs

@@ -499,7 +499,7 @@ class MESA(object):
             a.append(new_a)
             P.append(P[-1] * (1 - k * k.conj()))
             #Compute optimizer value for chosen method
-            optimization.append( np.abs(self._optimizer(P, a[-1], self.N, i + 1)) )
+            optimization.append( self._optimizer(P, a[-1], self.N, i + 1))
             
             is_nan = np.isnan(new_a).any() #checking for nans
             if np.abs(k)>1 or is_nan:

### 2021-02-24 paper/mesa.tex
Edit in validation section

@@
 \usepackage{ragged2e}
+\usepackage{cleveref}
 \usepackage{blindtext}
@@
 To show the analogy, we compute the PSD of such process and we show that it is formally equivalent to the PSD obtained in eq.~\ref{eq:MESApsd}.
-We start taking the z \sschmidt{What is the z transform} \wdp{it is a relative of the Fourier transform, z are the roots of the unit in the complex plane} transform of the previous eq.~\ref{eq:AR_p}: 
+We start taking the z \footnote{Spiegare in due righe la z transform} transform of the previous eq.~\ref{eq:AR_p}: 
 \begin{align}
@@
 This section is devoted to an extensive study on how to make such choice.
-First of all (\ref{sec:optimizers}), we are going to define three different ``metrics" to measure how well the algorithm is performed: we will refer to them as \textit{optimizers}.
-The basic idea is to monitor, as the autoregressive order considered increase, the performance of the algorithm (in some sense) and pick, among the orders the one that yields better results.
-Second, the performance of different optimizers will be assessed by performing the spectral analysis on time series with given analytical PSD, sec.~\ref{sec:arp_validation}.
-In a last subsection~\ref{sec:LIGO_validation} real data from LIGO-Virgo gravitational waves observatory.
-
-\sschmidt{Qui ho provato a mettere un po' di ordine alla sezione.
-Ho scritto l'introduzione sopra per spiegare i passaggi logici della validation.
-Spero che vada bene.
-La parte sotto è ancora da rivedere e probabilmente c'è molto da asciugare: sono quasi 10 pagine!}
+In section~\ref{sec:optimizers}, we are going to define three different \textit{loss functions} to measure how well the algorithm is able to reproduce a known PSD.
+The basic idea is to validate, as the autoregressive order considered increase, the performance of the algorithm results by measuring the loss function and pick, among the orders the one that yields better results.
+Second, the performance of different loss functions will be assessed by performing the spectral analysis on time series with given analytical PSD, sec.~\ref{sec:arp_validation}.
+In a last subsection~\ref{sec:LIGO_validation}, the same analysis is performed on real data from LIGO-Virgo gravitational waves observatory.
 
 \subsection{Choosing an autoregressive order}\label{sec:optimizers}
-\sschmidt{optimizer è la parola giusta? Optimizer è un algoritmo... Si potrebbe chiamare optimizer ``metrica"?}
+
 A good choice for the autoregressive order $M$ is proposed to be \cite{doi:10.1190/1.1440902}
@@
 We then need a more sophisticated method for computing the right value for $M$.
-Various optimizers to assess the algorithm performance have been proposed in literature. We summarize them below:
+Various loss functions to assess the algorithm performance have been proposed in literature.
+We summarize them below:
 
@@
 \item \textbf{Criterion Autoregressive Transfer function (CAT)}
-This second optimizer is due to Parzen \cite{bhansali1986}, and it is defined as
+This second loss function is due to Parzen \cite{bhansali1986}, and it is defined as
 \begin{equation}
@@
 The last criterion we will consider is the OBD (Rao, 1982)\cite{doi:10.1029/WR018i004p01097}.
-With the technique of asymptotic integration, th optimizer is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}:
+With the technique of asymptotic integration, th loss function is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}:
    \begin{align}
@@
 
-Once an optimizer is chosen, the choice of the best recursion order is straightforward: we solve the Levinson recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached.
-Then, the order $M$ is selected to be the one that minimize the chosen optimizer value.
-
-In a real implementation of the algorithm, computing all the recursion up to $M_{max}$ can result in a significant waste of computational power: the optimal value is often $M_{opt}<< M_{max}$ and in such cases, computing all the values of $m$ until $M_{max}$ is not useful.
-In practice, we can apply an \textit{early stopping} procedure: every few iterations we look for the best order of $M_{opt}$; if this value does not change for a while, the computation is stopped and a good (local) minimum of the optimizer is found.
-
-The following sections will be devoted to the study of the statistical properties of the optimizers above: we need to understand which choice provides the best quality in the reproduction of some known power spectral densities.
-
-\subsection{Choosing an optimizer - A gaussian PSD} \label{sec:arp_validation}
-We test the performance of the three optimizers on a random time series with a known power spectral density.
+Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached.
+Then, the order $M$ is selected to be the one that minimize the loss function.
+
+In a real implementation of the algorithm, computing all the recursion up to $M_{max}$ can result in a significant waste of computational power: the optimal value is often $M_{opt}<< M_{max}$ and, in such cases, computing all the values of $m$ until $M_{max}$ is not useful.
+In practice, we can apply an \textit{early stopping} procedure: every few iterations we look for the best order of $M_{opt}$; if this value does not change for a while, the computation is stopped and a good (local) minimum of the loss function is found.
+
+The following sections will be devoted to the study of the statistical properties of the loss functions introduced above: we need to understand which choice provides the best quality in the reproduction of some known power spectral densities.
+
+\subsection{Choosing a loss function - A gaussian PSD} \label{sec:arp_validation}
+We test the performance of the three loss functions on a random time series with a known power spectral density.
 We take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$, see fig.~\ref{fig:autocorr}.
@@
 
-We will apply Burg method with each optimizer on a number of simulated time series, obtaining a variety of estimate for PSD.
-We will study statistical properties for both single reproductions and ensemble average. In either cases, we will quantify the agreement between the reconstructed quantities and the expected ones.
-In many realistic cases, such ensemble average is not accessible (one car rarely observe multiple time series realizations for the \textit{same} process), so that single reproduction studies are probably more useful.
-Still, study of ensemble average is important to highlight some interesting properties.
-In this way we can fully characterize the performance of each optimizer .
-n what follows, we will indicate with $\bar S$ any ensemble averaged quantity.
-
-We will attentively study the role of the filter length estimate M in determining the accuracy of the reconstruction.
-We will consider every optimizer separately and compare result only at the end of this section.
-In section~\ref{sec:LIGO_validation}, we will perform the analysis again using, for the purpose of generating the time series, the Advanced LIGO design sensitivity PSD \cite{Ligo}.
+We will apply Burg method with each loss function on a number of simulated time series, obtaining a variety of estimate for PSD.
+The disagreement between the PSD $S_i(f)$ estimated from the i-th simulated time series and the target PSD $S(f)$ is measured with the relative error $r_i$, averaged over all frequencies:
+\begin{equation}
+r_i = \frac{1}{f_{Ny}}\int_0^{f_{Ny}} \d{f} \frac{|S_i(f) - S(f)|}{S(f)}
+\end{equation}
+where $f_{Ny}$ is the Nyquist frequency of the time series. Of course in practice, the integral will be computed by averaging over a discrete grid of frequencies.
+
+We will consider every loss function separately and compare the results only at the end of this section.
+For each loss function, we will compute the ensemble averaged PSD estimation (as well as the $90\%$ confidence level) and the frequency dependence of the error (ensemble averaged):
+\begin{equation}\label{eq:ensemble_averaged_error}
+	r(f) = \sum_i \frac{|S_i(f) - S(f)|}{S(f)} \;.
+\end{equation}
+In figure~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} we report the averaged PSD for the three loss functions, as well as the ensemble averaged error in eq.~\ref{eq:ensemble_averaged_error}.
+
+Furthermore, we display in a contour plot the correlation between $r_i$ and the autoregressive order $m_i$ chosen for each independently drawn time series.
+This is done in figure~\ref{fig:optcomparison}.
+
+For our studies, we generate a dataset of 1000 time series of 3000 points each: see fig.~\ref{fig:noise} for an example of a time series in frequency and time domain.
+In section~\ref{sec:LIGO_validation}, we will perform the analysis again using the Advanced LIGO design sensitivity PSD \cite{Ligo}.
 
@@
         \includegraphics[width = \linewidth]{Images/NormalPSD/NormalNoise.pdf}
-    \caption{Noise in both time and frequency domain. Frequency noise behaviour is representative of the chosen spectrum, having its maximum amplitude around $f = 2.5 au$}
+    \caption{Noise in both time and frequency domain.
+Frequency noise behaviour is representative of the chosen spectrum, having its maximum amplitude around $f = 2.5 au$ \sschmidt{Togliamo questa figura?}}
     \label{fig:noise}
@@
 
-For our studies, we generate a dataset of 1000 time series of 3000 points each: see fig.~\ref{fig:noise} for an example of a time series in frequency and time domain.
-For every time series we will compute the spectrum with every optimization method.
-To assess the magnitude of the disagreement, we compute the relative error between the mean spectrum $\hat{S}$ and the true (analytical) spectrum $S$ for every frequency bin:
-\begin{equation}
-r(f) = \frac{|\hat{S}(f) - S(f)|}{S(f)}\; .
-\end{equation}
-
-We report the average values of the relative error for the different optimizers in Table~\ref{tab:relative_errors}.
-
-\sschmidt{Sono arrivato fino a qui nelle modifiche}
-\begin{table}
-	\caption{Average relative error for different optimizers, computed in different ways. We consider a set of $1000$ realization of a noisy time series of $3000$ points.
-	We report the average error  $\bar{r}_{\bar{S}}$ on the whole set of simulations, the error $\bar{r}_{\bar S}^{\mu \pm 3\sigma} $ averaged only on data points between $3\sigma$ and the error $\bar{r}$ on a single, randomly drawn realization.
-	\sschmidt{I think we need to discuss on which performance error to include in the estimation}
-	}
-	\label{tab:relative_errors}
-	\def\arraystretch{1.5}
-	\begin{ruledtabular}
-		\begin{tabular}{ l c c c }
-			\multirow{2}{*}{Optimizer}& \multicolumn{3}{c}{Relative Error}\\
-			& $\bar{r}_{\bar{S}}$	& $\bar{r}_{\bar S}^{\mu \pm 3\sigma} $ & $\bar{r}$\\
-			\hline \hline
-			FPE 			& $6.9 \; (46.9\%)$	& $7 \; (1.6\%)$ & \\ 
-			\hline
-			CAT 			& $4.5 \; (30.6\%)$	& $194 \; (45.3\%)$ & \\
-			\hline
-			OBD 				& $1.7 \; (11.6\%)$	& $206 \; (48.1\%)$ & \\
-		\end{tabular}
-	\end{ruledtabular}
-\end{table}
-
-\begin{figure}[H]
+\begin{figure}[t]
 	\centering
-	\includegraphics[width = \linewidth]{Images/NormalPSD/FPEpsdAndResiduals.pdf}
+	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/FPE_spectrum_estim.pdf}
 	\caption{Ensemble Average for FPE reconstruction with 90\% Credibility Regions (above) and percentage error (below)}
@@
 \end{figure}
-
-\subsubsection{Final Prediction Error (FPE)}
-The results of our investigation on te performance of FPE are reported in fig.~\ref{fig:FPEmean}: we plot the ``true" spectrum, the reconstructed spectrum as well as the $90\%$ credible region of the predicted spectrum.
-Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum; however, we note that the reconstruction is not very accurate in the low frequency region.
-
-For the entire spectrum, the mean relative error $\bar{r}$ is 
-\begin{equation}  
-    \bar{r}_{\bar S, FPE} = 0.025.
-\end{equation}
-FPE performs well in reconstructing the spectrum even with relatively short time series, with a deviation from the true spectrum around 2\%, but it is evident that it is not able to capture all the features of the spectrum, since the reconstruction in the tails is inaccurate.
-If we compute the average percentage error in three standard deviations from the mean we gain an order of magnitude estimate of the accuracy of the reconstruction, obtaining:
-\begin{equation}
-    \nonumber
-    \bar{r}_{\bar S, FPE}^{\mu \pm 3\sigma} = 0.005. 
-\end{equation}
-and we can state the the method is highly performant in this frequency window. 
-Since it is not always possible to work with more then one realization for a process, as in observational experiments, it is important to understand how this method behaves when we apply it on a single time series.
-In that case, the mean error and its standard deviation are:
-\begin{equation}
-    \nonumber
-    \bar r_{FPE} = 0.16; \quad \sigma_{r} = 0.16\,.
-\end{equation}
-The average error of a reconstruction based on a single time series is bigger than the one obtained considering the average of all the reconstructions.
-Standard deviation of the errors assume small values and 90\% credibility regions (Cr.) in \ref{fig:FPEmean} are tight.
-This means that if we randomly take two of the reconstructed spectra, we expect their differences to be statistically small.
-These facts are evidence for FPE to be a relatively stable optimizer.
-We already said that the estimate for filter's length lays in a well defined range of values.
-The quality of the reconstruction, as noted in Chapter 1, depends on the order of the reconstructed AR process.
-To investigate this idea further, we plot the value of the optimizer for each recursion and the residuals as a function of the estimate for the order of the autoregressive process (\textbf{Figure \ref{fig:FPEErrorOrder}}),
-
-\begin{figure}[H]
-    \noindent
+\begin{figure}[t]
 	\centering
-    \includegraphics[width = \linewidth]{Images/NormalPSD/FPEoptimizer.pdf}
-    \caption{Stability of FPE, that results in a cluster of values in a small region (left), is due to the typical behaviour of the optimizer as a function of the AR order (right).}
-    \label{fig:FPEErrorOrder}
-\end{figure}
-Foremost evidence for the stability of this optimizer is found in its values as a function of the AR order.
-Figure \ref{fig:FPEErrorOrder} shows that Values tend to decrease monotonically until a sharp and well defined minimum is reached, to start growing again after this area.
-This is representative of the typical behaviour of this optimizer as a function of the length of the filter, and explains why the estimates for AR orders are confined in a small region.
-This plot also shows that most of the values collapsed in the region of AR orders comprised between 30 an 50.
-There are only few values outside this interval, and get fewer as order's estimate increases.
-In this region, there is no strong evidence of some correlation between the estimate for M and its associated relative error.
-It is important to notice that the estimate associated with the higher values of M (around $M > 60$) exhibit the highest 'mean' relative error.
-\\
-We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length M is clustered in a region where there is no dependence of error from M.
-\subsubsection{Optimum Bayes Decision Rule (OBD)}
-The second optimizer in consideration is OBD.  (\textbf{Figure \ref{fig:OBD3000}}) shows a good agreement between the average over the reconstructions and true spectrum.
-\begin{figure}[H]
+	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/OBD_spectrum_estim.pdf}
+	\caption{Reconstruction of the spectrum from 1000 simulations with OBD optimizer. Plot is mean of all the results with 90\% credibility regions (above) and percentage error (below)}
+	\label{fig:OBDmean}
+\end{figure}
+
+\begin{figure}[t]
 	\centering
-    \includegraphics[width = \linewidth]{Images/NormalPSD/OBDpsdAndResiduals.pdf}
-    \caption{Reconstruction of the spectrum from 1000 simulations with OBD optimizer. Plot is mean of all the results with 90\% credibility regions (above) and percentage error (below)}
-    \label{fig:OBD3000}
-\end{figure}
-The same behaviour of FPE is observed, a good quality reconstruction at the center of the distribution with quite small Credibility regions (Cr.) at a $90$\% credibility level.
-Qualitatively they behave similarly but quantitatively the disagreement of this method is found to be even larger than the one obtained with FPE, as can be noticed in the plot of relative errors in figure \ref{fig:OBD3000} (below) where at the extremes the inaccuracy is even larger than 300\% with respect to the real value of the spectrum.
-For the overall spectrum and within three standard deviation from the mean, we find that the reconstructed spectrum has a deviation of
-\begin{equation}
-    \nonumber
-    r_{\bar S, OBD} = 0.15, \qquad r^{\mu \pm 3\sigma}_{\bar S, OBD} = 0.018. 
-\end{equation}
-When applied to single realizations of the time series, we find
-\begin{equation}
-    \nonumber
-    \bar{r}_{OBD} = 0.237, \qquad \sigma_{r, OBD} = 0.0296.
-\end{equation}
-It is interesting to notice that overall relative error of the average spectrum is of the same order of magnitude of the average single reconstruction for the spectrum.
-Averaging over a 1000 realisations does not substantially improve our results.
-As a function of the filter length, the values taken by the OBD cluster in a small region, indicating the stability of the reconstructed process order $M$.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width = \linewidth]{Images/NormalPSD/OBDoptimizer.pdf}
-    \caption{Stability of OBD, that results in a cluster of values in a small region (left), is due to the typical behaviour of the optimizer as a function of the AR order (right)}
-       \label{fig:OBDErrorOrder}
-\end{figure}we see that again our results are clustered in a very small region, due to the behaviour of the optimizer that converge in a small area before growing again.
-The estimate of the length of the filter is around $p = 20 \div 30$.Note the difference compared to FPE (see \ref{fig:FPEErrorOrder}); for OBD the mean relative error depends on $M$ as
-\begin{equation}
-    \hat r \sim M ^ {-2.016 \pm 0.006}.
-\end{equation}
-In summary, OBD shows greater error wherever in the spectrum, due to the fact that it choose filters that are too short and in a region where quality drastically change when changing M.
-In this case we can say that we have a stable estimate for M, but due to this dependence, the estimate is not as stable as FPE's.
-\subsubsection{Criterion Autoregressive Transfer function (CAT)}
-This last criterion differs substantially from the two other optimizers. Figure \ref{fig:CATmean} (above) shows a good agreement between the estimate of the spectrum for every frequency
-\begin{figure}[H]
-    \centering
-    \includegraphics[width = \linewidth]{Images/CAT/CATnormal.pdf}
+	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/CAT_spectrum_estim.pdf}
 	\caption{Mean spectrum for CAT optimizer, 1000 simulations for 3000 points datasets}
 	\label{fig:CATmean}
-\end{figure} and also tails are well captured by the average spectrum.
-However, the variance of the reconstructed spectrum is quite large, much larger than for FPE and OBD.
-In general, the relative error is nearly constant, $\sim 10\%$ over all frequencies, see  Fig.
-\ref{fig:CATmean} (below).
-The residuals, averaged over all frequencies, are:
-\begin{equation}
-    \nonumber
-    \bar r_{\bar S, CAT} = 0.016 \qquad \bar r^{\mu \pm \sigma}_{\bar S, CAT} = 0.017.
-\end{equation}
-the overall reconstruction is thus quite accurate, with deviations from the simulated spectrum of the order of $1\% \div 2\%$.\\
-The picture drastically changes in the case of individual time series realizations. The average residuals are, in fact
-\begin{equation}
-    \bar r_{CAT} = 0.42 \qquad \sigma_{CAT} = .44\,.
-\end{equation}
-The error is indeed $\sim 42\%$, hence much larger than in the case of FPE and OBD.
-Both large error bars and the large standard deviation are indications of the instability of this optimizer.
-This is exemplified by the variety of different spectra that can be obtained considering different time series.
-
-\begin{figure}[H]
+\end{figure}
+
+\begin{figure}
 	\centering
-    \includegraphics[width = \linewidth]{Images/CAT/CAToptimizer1.pdf}
-    \caption{CAT is not a stable optimizer since its optimizer do not converge to any specific region of the AR estimate order}
-    \label{fig:CAToptimizer}
-
-\end{figure}
-The reason for this behaviour, can be found in \textbf{Figure (\ref{fig:CAToptimizer})}.
-CAT does not converge to any specific value for the order of the autoregressive process.
-In turn, this results in estimates for the order that span a large range.
-Figure \ref{fig:CAToptimizer} also shows a strong dependence of the error on the estimated length of the filter.
-The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like the big excursion in the size of datas of a normal distribution, but this length is also responsible of an augmented variance in the estimate, introducing spurious peaks in the reproduction (Figure \ref{fig:GaussMedian})
-
-Averaging, the noise in each spectrum is cancelled.
+	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/error_length_contour.pdf}
+	\caption{Relative error as function of the length of the filter.}
+	\label{fig:optcomparison}
+\end{figure}
+
+\paragraph{Final Prediction Error (FPE)}
+The results of our investigation on the performance of FPE are reported in fig.~\ref{fig:FPEmean}.
+Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum; however, we note that the reconstruction is not very accurate in the low frequency region. 
+Furthermore, the $90\%$ credible region is very small: this means that if we randomly take two of the reconstructed spectra, we expect their differences to be statistically small.
+These facts are evidence for FPE to be a reliable loss function.
+
+By looking at the red series in fig.~\ref{fig:optcomparison}, we note that the autoregressive orders obtained with FPE are clustered in a very small region.
+This is a desiderable property: FPE provides a stable estimation of the autoregressive order, which does not affect much the reconstruction error.
+We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length M is clustered in a region where there is no dependence of error from M.
+\paragraph{Optimum Bayes Decision Rule (OBD)}
+The second loss function in consideration is OBD.
+The results in fig.~\ref{fig:OBDmean} show a good agreement between the average over the reconstructions and the true spectrum.
+Qualitatively the same behaviour of FPE is observed: a good quality reconstruction at the intermediate and high frequencies with a low $90\%$ confidence level; and a worse performance at low frequencies.
+However, when looking at the error, the disagreement of this method is found to be larger than the one obtained with FPE: in the worst case, the error can gain a factor of $2$ as compared with FPE.
+
+When plotting the error as a function of the autoregressive length (green series in fig.~\ref{fig:optcomparison}), the values taken by the OBD cluster in a small region, indicating the stability of the reconstructed process order $M$.
+We note that on average, OBD tends to choose smaller values of $M$ with respect to FPE: this can probably explain the observed large errors.
+
+\paragraph{Criterion Autoregressive Transfer function (CAT)}
+The performance of CAT are shown in fig.~\ref{fig:CATmean} and it is striking that CAT differs substantially from the other two loss functions considered.
+The ensambled average PSD matches very well the ``true" PSD. This is also true even in the low frequency region, where the OBD and FPE had a poor performance.
+However, the variance of the reconstructed spectrum is quite large (much larger than for FPE and OBD), and the relative error is quite high $~10\%$ and it approximately constant over all the frequencies.
+
+The reason for this behaviour, can be found in fig.~\ref{fig:optcomparison} (black series). CAT does not converge to any specific value for the order of the autoregressive process.
+In turn, the estimates for the order span a large range of values, hence the large variance observed.
+Figure~\ref{fig:optcomparison} also shows a strong dependence of the error on the estimated length of the filter.
+The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like outliers in different realization of a noisy timeseries, but this is also responsible of an augmented variance in the estimate, by introducing spurious peaks in the reproduction.
+
+When averaging the diffrerent PSD estimation, the noise in each spectrum cancels.
 It is interesting that the cancelling of errors in averaging is property of gaussian noise (the process is gaussian, after all).
@@
 This instability is not a good property for an estimation, so that this estimator might be found to be optimal for repeatable experiments and not for single observation of some process.
-Another approach to this method will be proposed in section \ref{sec:ACAT}
-
-\begin{figure}
-    \centering
-    \includegraphics[width = \linewidth]{Images/NormalPSD/Gaussmedian.pdf}
-    \caption{Reconstructed power spectral density with the three method. The reported reconstruction is the one that shows the median percentage error (overall)}
-    \label{fig:GaussMedian}
-\end{figure}
-\subsubsection{Conclusive Remarks}
-\begin{figure}
-    \centering
-    \includegraphics[width = \linewidth]{Images/NormalPSD/Normal Residuals.pdf}
-    \caption{Percentage error in the reconstruction for every optimizer, as a function of the frequency. Lower panel restricts the domain to three standard deviation to increase the definition}
-    \label{fig:NormalEnsembleResiduals}
-\end{figure}
-In our analysis, FPE and OBD optimizers are found to behave similarly while CAT shows completely different properties.
-Studying the average spectrum is not sufficient to choose which method behaves better.
-\\
-The bad behaviour in the reconstruction of the tails from OBD and FPE might be due to the fact that very small values require longer filters. 
-In fact when we consider the average for all the spectra we note that the shorter the filter the larger the error in the tails: OBD, which select the shortest filters, arrives to a $300 \%$ error at the extrema of the power spectral density, FPE which select intermediate order has got a $85\%$ error in same window, while CAT, that selects a wide range of long filters, has got a constant accuracy everywhere in the spectrum.
-\\
-Figure \ref{fig:NormalEnsembleResiduals} shows that CAT is the method that minimizes errors in the tails (upper plot), but if one restrict to three standard deviations from the mean, FPE is found to reach best accuracy (lower plot).
-The frequency averaged percentage error is reflecting this behaviour. Considering the whole reproduction, error's average and standard deviation are found to be: 
-\begin{align*}
-\centering
-&r_{\bar S, FPE} = 0.025; \quad \sigma_{\bar S, FPE} = 0.083;  \\
-&r_{\bar S, OBD} = 0.15;\quad \sigma_{\bar S, OBD} = 0.42 ; \\ 
-&r_{\bar S, CAT} = 0.015; \quad \sigma_{\bar S, CAT} = 0.012
-\end{align*}
-In is interesting to see that the average for FPE and OBD is strongly affected from the high value of the error in the tails of the distribution.
-If one consider the three standard deviations from the mean, the value of the error substantially decreases,
-\begin{align*}
-\centering
-&r_{\bar S, FPE} = 0.005; \quad \sigma_{\bar S, FPE} = 0.004;  \\
-&r_{\bar S, OBD} = 0.018;\quad \sigma_{\bar S, OBD} = 0.019 ; \\ 
-&r_{\bar S, CAT} = 0.015; \quad \sigma_{\bar S, CAT} = 0.011
-\end{align*}
-showing that the overall percentage error is just a first indicator for the quality of the reproduction, but is not sufficient to perform a full characterization. \\
-It is evident that tails do not affect CAT  average error and standard deviation at all, so that it's accuracy is approximately constant everywhere in the spectrum. \\ \ 
-Since both FPE and CAT shows pros and cons, at this level, there is no reason to choose between them.\\ 
-Analyzing the resolution of the single reconstruction, figure \ref{fig:optcomparison} clearly shows that FPE minimize overall relative error giving a stable result.\\ 
-CAT instead shows a larger error and do not converge to any specific value.
+
+\paragraph{Final remarks on the choice of the loss function}
+
+In our analysis, FPE and OBD loss functions are found to behave similarly while CAT shows completely different properties.
+CAT provides an accurate average spectrum over all the frequencies and large amount of variability; in turn OBD and FPE provide a poorer average in the low frequency tails with a lower variability (FPE with less noise than OBD).
+
+The bad behavior in the reconstruction of the tails from OBD and FPE might be due to the fact that very small values require longer filters. 
+In fact when we consider the average for all the spectra we note that short filters yield larger the error in the tails.
+OBD, which select the shortest filters, can provide an error as large as $300 \%$ at the extrema, as compared with $85\%$ error od FPE.
+In turn, CAT is $5\%$ accurate in those regions.
+
+However, we see in figure \ref{fig:optcomparison} that a single PSD estimation with FPE is still able to provide the lowest averaged error over all frequencies, while CAT can reach error error $5$ times larger.
+This is to say that CAT is the loss function that minimizes errors in the tails, FPE is found to reach best overall accuracy.
+
+The conclusion is clear: even if in some cases CAT might be more accurate when taking the average over several estimation, FPE guarantees that the single estimation will be more accurate.
+As in any common situation we cannot perform such averaging over different realization of the same time series, we must prefer FPE over CAT (let alone OBD, which even though qualitatively similar to FPE has worse performance).
+However, if we indeed can measure the PSD by averaging over different time series, probably using CAT as a loss function is the best choice.
+\sschmidt{This is similar to Welch: averaging over different time series to get a reliable result. Should we mention it?}
+\sschmidt{C'erano ancora molti commenti: ho sintetizzato molto. Va bene? O si vuole aggiungere altro?}
+
+%It is evident that tails do not affect CAT  average error and standard deviation at all, so that it's accuracy is approximately constant everywhere in the spectrum. \\ \ 
+%Since both FPE and CAT shows pros and cons, at this level, there is no reason to choose between them.\\ 
+%Analyzing the resolution of the single reconstruction, figure \ref{fig:optcomparison} clearly shows that FPE minimize overall relative error giving a stable result.\\ 
+%CAT instead shows a larger error and do not converge to any specific value.
 Almost every reproduction obtained with CAT is worst with respect to those obtained with FPE, but their mean is more precise.
 \\
-It's instability is probably the cause for this behaviour. Since it does not show a strong convergence to some order with respect to others, very different lengths are selected.
-Longer lengths are able to capture a lot of details but are more likely to introduce noise. In other words, long filters reduce bias but increment the variance of the result. 
-Taking the average, for central limit theorem the error decreases as $\sim \frac{1}{\sqrt{N}}$. In this way we compensate noise and obtain a spectrum with small bias and small fluctuations.
-Considering the average, CAT reconstruct precisely and with good accuracy every part of the spectrum, even if in "regular" parts best accuracy is reached from FPE.
+%It's instability is probably the cause for this behaviour. Since it does not show a strong convergence to some order with respect to others, very different lengths are selected.
+%Longer lengths are able to capture a lot of details but are more likely to introduce noise. In other words, long filters reduce bias but increment the variance of the result. 
+%Taking the average, for central limit theorem the error decreases as $\sim \frac{1}{\sqrt{N}}$. In this way we compensate noise and obtain a spectrum with small bias and small fluctuations.
+%Considering the average, CAT reconstruct precisely and with good accuracy every part of the spectrum, even if in "regular" parts best accuracy is reached from FPE.
 For a single reproduction, CAT is very noisy and unstable, while FPE performs the better.
 \ref{fig:optcomparison}
-\begin{figure}
-    \centering
-    \includegraphics[width = \linewidth]{Images/NormalPSD/NormalPSDcomparison.pdf}
-    \caption{Relative error as function of the length of the filter.}
-    \label{fig:optcomparison}
-\end{figure}
-\begin{figure}
-    \hfill
-    \centering
-    \includegraphics[width = \linewidth]{Images/NormalPSD/ordersComparison.pdf}
-    \caption{Histogram of estimate orders for each method}
-    \label{fig:ordersCompairson}
-
-\end{figure}
-definitely shows that FPE gives the best results: it is a stable method that choose filters in the clustered area associated with minimum error, while both OBD and CAT result lies outside this area and show bigger errors.
+%definitely shows that FPE gives the best results: it is a stable method that choose filters in the clustered area associated with minimum error, while both OBD and CAT result lies outside this area and show bigger errors.
 Two other interesting plots that stress this fact are reported in figure \ref{fig:ordersCompairson} and \ref{fig:residualsComparison}
-\begin{figure}
-    \centering
-    \includegraphics[width = \linewidth]{Images/NormalPSD/ResidualsComparison.pdf}
-    \caption{Histogram of residuals for each method}
-    \label{fig:residualsComparison}
-\end{figure}
-where the histograms for the errors and for order's estimate are reported.
+%where the histograms for the errors and for order's estimate are reported.
 These graphs provide stronger evidence for our previous statements.
@@
 \\
-CAT histogram shows a very interesting behaviour.
+%CAT histogram shows a very interesting behaviour.
 The histogram for the order show two different peaks,  a small one associated to 'short filters', and a second high peak at $m = M$.
@@
 
-So, even if in some cases CAT might be more accurate when taking the average over several spectra, FPE guarantees that minimum error will be reached.
-
-\subsection{Choosing an optimizer - LIGO Spectrum} \label{sec:LIGO_validation}
+\subsection{Choosing a loss function - LIGO Spectrum} \label{sec:LIGO_validation}
 \begin{figure}
@@
 For the first one a log plot of the reconstructed peaks and their relative errors are shown in Figure \ref{fig:ligo1peak}
-\begin{figure}[h]
+\begin{figure}[t]
         \includegraphics[width = \linewidth]{Images/LIGOsimulate/1stPeakComparison.pdf}
@@
 Second peak results are analogous, plotting the peak and the errors value around it
-\begin{figure}[h]
+\begin{figure}[t]
     \centering
@@
     \centering
-    \includegraphics[width = \linewidth]{Images/LIGOsimulate/orderVSerror.pdf}
+    \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/error_length_contour.pdf}
     \caption{Plot of relative error as a function of the estimate for filter's length}
@@
 \\
-\begin{figure}
-    \centering
-    \includegraphics[width = \linewidth]{Images/LIGOsimulate/LigoHists.pdf}
-    \caption{Histograms showing the estimate for the autoregressive order and overall percentage error obtained with all the three optimizers.}
-    \label{fig:my_label}
-\end{figure}
 
@@
 \begin{figure}
+
 	\label{fig:mem_welch_realdata}
-	\caption{Blablabla}
-\end{figure}
-
-\sschmidt{Mettiamo da qualche parte una descrizione dettagliata del codice?}
-
+	\caption{Make here a comparison between Welch}
+
+\end{figure}
+
+\sschmidt{Mettiamo da qualche parte una descrizione dettagliata del codice? No: solo rimandiamo al codice nell'intro}
+
+\section{Forecast analysis}
 \section{Final remarks and future prospects}
@@
 \end{document}
-
-%%%%%%%%%%%%%% BELOW THE OLD PART... USEFUL TO TAKE INSPIRATION FROM
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction}
-The Maximum Entropy principle (MAXENT) is one of the most important results in Bayesian Statistics.
-It provides a way to uniquely assign probabilities to a phenomenon in a way that best represent our state of knowledge, while at the same time it is non committal with unavailable information.
-Its domain of application turned out to be wider than expected.
-In fact, thanks to Burg \cite{burg1975maximum}, this method has also been applied to perform high quality computation of power spectral densities of time series.
-After a short introduction to Jayne's MAXENT, we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form.
-Since its natural domain of application are time series, we will also provide a brief explanation of the basic concepts in stochastic processes that are needed to fully understand the method and the cases in which it can be correctly applied.
-\section{Maximum Entropy}
-One of the main problems in statistics is the assignment of a probability for an event.
-There have been several proposals from both 'objective' and 'subjective' approaches to probability and statistic.
-The former assert that concept of probability is intrinsic to the event itself and can be computed as the 'percentage of occurrences of an attribute in a given ensemble':
-speaking of probability only makes sense when an ensemble is considered. 
-For the latter approach instead, probability is a representation of the state of knowledge of the observer about the event itself: the higher the probability, the higher our belief that the event will occur and vice versa.
-It appears to be clear that for the Bayesian, information will play a key role in the a priori assignment of the probability for an event, while for the frequentist prior assignment does not make sense at all, since the whole information is contained in the experiment.
-\\
-So, assigning probability given some prior information is a pure Bayesian problem.
-
-\subsection{Indifference principle}
-The indifference principle (IP) is one of the easiest principle that can be used to assign a probability to some events. 
-Consider a set of independent, mutually exclusive hypotheses: 
-\begin{equation}\nonumber
-    \left\{H_i\right\} \quad i = 1 \dots N
-\end{equation}
-so that they are exhaustive: 
-\begin{equation}\nonumber
-    \sum_{i = 1}^{N} P(H_i \vert I) = 1.
-\end{equation}
-If there is no reason to prefer one hypothesis with respect to another, the IP states that we should assign the same probability for every event: 
-\begin{equation}
-    P(H_i\vert I) = \frac{1}{N}. 
-\end{equation}
-This can be easily generalized to the continuous case. Consider, for simplicity, a one dimensional case, and suppose we have the additional information: $a < x < b$.\\ 
-In the continuous case, the probability $p(x\vert I)$ is meant as the probability for the event to assume values in a given subset $\left[x, x +dx\right]$.
-Requiring events to be mutually exclusive is requiring their interception to be the empty set, while the exhaustiveness condition is the normalization condition:
-\begin{equation}
-    \nonumber
-    \int_a^b p(x\vert I)dx = 1.
-\end{equation}
-In this case, the IP reduces to the computation of the normalization constant: 
-\begin{equation}
-    p(x\vert I) = \frac{1}{b-a}
-\end{equation}
-and allow us to easily introduce uniform priors, the simplest prior distribution for Bayesian computation.
-The only hypothesis is for our events to be mutually exclusive and exhaustive.
-In this case, since there is no evidence for a given hypothesis to be more plausible than another, according to IP we have to assign the same probability for every event.
-\\
-If that is not the case, how can we generalize IP principle for a larger amount of information?\\ 
-Jaynes\cite{JaynesArticle}\cite{jaynes2003ptl} formulated an exhaustive answer as a variational problem and rely to the concept of information entropy: 
-the distribution that reflects knowledge being maximally non committal to unknown information is uniquely determined maximizing Shannon's entropy\cite{Shannon} using information as constraints. \\ 
-This is now known as the 'Maximum Entropy Principle' (MAXENT)
-
-\subsection{Maximum Entropy Principle}
-To introduce MAXENT as a variational problem, we will workout some simple examples to understand how information's entropy and 'uncertainty' are related. \\ 
-The word 'information' will not be used to indicate our knowledge on the system under study: this is what we will call 'evidence'.
-Information will be intended as in communication theory: it specifies the quantity of details needed to provide a full description of the system under study.
-\\
-If we consider a perfectly sinusoidal signal, knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information entropy content . \\ 
-Even less information is required if we are studying a system whose outcome is certain (has probability $p = 1$).
-If the quantity of information can be considered to be $I = 0$, a communication is not even needed.  
-Shannon \cite{Shannon} proposed the quantity
-\begin{equation}\label{eq:information}
-    I_i = \log_2 \frac{1}{p(x_i)}
-\end{equation}
-to represent the quantity of information brought by an outcome $x_i$ with probability $p_i$.
-It has the nice property of being an additive quantity and it is monotonically decreasing as a function of $p \in [0, 1]$.
-The more uncertain the outcome, the higher the information it brings.
-\\
-What happens if we consider not one outcome, but a larger set of different events?
-What is  the 'most uncertain' outcome?
-Since we are intending probability as a representation of the state of knowledge, one of the first requirement for our assignment is the agreement with common sense.
-\\
-Consider an experiment based on the occurrence of one of two different outcomes $E_1, E_2$ with given probabilities $P_1$ and $P_2$. What are the values that make our outcome the most uncertain? \\ 
-If $P_1$ and $P_2$ are largely different, like $P_1 = 0.999$ and $P_2 = 0.001$, we have enough evidence to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome.
-\\
-The most unpredictable experiment is the one in which 
-\begin{equation}\nonumber
-    P_1 = P_2 = \frac{1}{2}:
-\end{equation}
-this describes a situation of 'maximum ignorance'; we have no evidence at all and we have no instruments to predict what the outcome of such an experiment will be.  \\ 
-This can be generalized in the case of $N$ events. Assigning same probability
-\begin{equation*}
-    P_1 = \dots = P_N = \frac{1}{N}
-\end{equation*}
-for each outcome, i.e. appealing to the indifference principle, is equivalent to the requirement for our experiment to be the most unpredictable. 
-All we know is that one over $N$ events is going to happen. With this evidence, we cannot expect one outcome to occur over the others. 
-To report the results of such an experiment, we need to communicate every single outcome: it 'carries' high information. 
-This description also makes clearer the role played by evidence: it has to be used in probability assignment in order to make the result 'more certain' or 'less informative'.
-It is clear from the point of view of communication theory: the higher the evidence on a system, the lower the information I need to send to describe it.
-\\
-It is now clear that any assignment different from the 'most uncertain' is violating the hypothesis for it to being non committal to non-available evidence.
-We can only assign a more certain probability, i.e.
-less informative, if we have evidence for it.
-If we have none, the result has to be the most unpredictable, i.e.
-the most informative.
-Assigning probabilities requires maximization of uncertainty using evidence as constraint.
-This looks like a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized.
-$H$ has to be a 'generalization' of information defined in equation (\ref{eq:information}) to be something like the 'expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
-\\
-Some properties for $H$ can be worked out easily. 
-From previous discussion, given a fixed number $N$ of possible outcomes, we want $H$ to be maximized by the choice $p_1 = \dots = p_N$. \\ 
-Consider now the opposite situation, in which probability is assigned appealing to IP and we have a different number of possible outcomes: 
-\begin{itemize}
-\centering
-    \item $P_1 = P_2 = \frac{1}{2}$
-    \item $P_1 = P_2 = P_2 = \frac{1}{3}$
-    \item $P_1 = \dots = P_4 = \frac{1}{4}$
-\end{itemize}
-Intuitively, the most unpredictable experiment is the one with larger number of outcomes: probability of getting it wrong increases as $\frac{N -1}{N}$.
-So when equal probabilities are assigned, we want
-\begin{equation}
-    \nonumber
-    H\left[\frac{1}{N}, \dots, \frac{1}{N}\right]
-\end{equation}
-to increase monotonically with N. \\ 
-We also want our function to be continuous in its parameters: any small change in probability has to slightly affect uncertainty. \\
-Last is composition property. Consider three events with probability $p_1, p_2, p_3$ such that $p_2 + p_3 = 1 - p_1 = q$. We require 
-\begin{equation}
-    H\left(p_1, p_2, p_3\right) = H(p_1, q) + qH\left(\frac{p_2}{q}, \frac{p_3}{q}\right),
-\end{equation}
-so that uncertainty is additive. \\ 
-Shannon \cite{Shannon} showed that the only functional form satisfying our assumptions is given by:
-\begin{equation}\label{eq:entropy}
-    H[p_1, \dots, p_N] = - \sum_{i = 1}^N p_i\log{p_i},
-\end{equation}
-or, in the continuous case:
-\begin{equation}
-    H[p(x)] = - \int p(x)\ln p(x) dx,
-\end{equation}
-and $H$ is what we will call entropy.
-This result is very interesting, if we compare it with Eq.~(\ref{eq:information}), we see that the entropy is defined as the 'average information'.
-The probability assignments made maximizing it with respect to $p$, is somehow requiring  the 'expected value' of uncertainty to be a maximum.
-Such an assignment is in agreement with our hypothesis of being 'non committal' with unavailable evidence, since \ref{eq:entropy} is all we need to assign probabilities uniquely.
-Evidence can be added in form of constraints in the maximization.
-\\ It is easy to verify that this expression for entropy verifies the aforementioned properties.
-Assigning the same value for each probability $p_i$, it is straightforward to verify $H$ increases with N: 
-\begin{equation}
-    -\sum_{i = 1}^N \frac{1}{N}\log{\frac{1}{N}} = - \log\frac{1}{N} = \log N.
-\end{equation}
-We want to show that, in case in which our only constraint is exhaustiveness \begin{equation}\nonumber 
-    \sum_{i = 1}^{N}p_i = 1,
-\end{equation}
-MAXENT reduces to indifference principle as it should be. The variational problem is defined as
-\begin{equation}\label{eq:entropyNormConst}
-    \delta\left[H - \lambda\sum_{i = 1}^N (p_i - 1)\right] = 0 
-\end{equation}
-with
-\begin{equation}
-    \delta H\left[p_1, \dots, p_N\right] = \sum_{i = 1}^N\frac{\partial H}{\partial p_i}\delta p_i = - \sum_{i = 1}^N\left(\log p_i + 1\right),
-\end{equation}
-so that equation \ref{eq:entropyNormConst} reduce to
-\begin{equation}\nonumber 
-    \sum_{i = 1}^N\left[\log p_i + \lambda + 1 \right]\delta p_i = 0\,.
-\end{equation}
-The solution of the previous equation is easily worked out requiring every term to be zero: 
-\begin{equation}\nonumber 
-    p_i = e^{-\lambda - 1}.
-\end{equation}
-Applying the normalization condition, it is immediate to show that
-\begin{equation}
-    p_i = \frac{1}{N}.
-\end{equation}
-MAXENT reduces to the IP when only the normalization constraint is used. What happens if more than one constraint is used? \\
-For simplicity, we will work through the next examples in the continuous case. Suppose the mean of the distribution is known $\int xP(x) dx = \bar{x}$. The variational problem can be rewritten as: 
-\begin{equation}
-    \int \left[\log p(x) + 1 + \lambda + \mu x \right]\delta p(x) = 0
-\end{equation}
-and its solution in term of $p_i$ is: 
-\begin{equation}
-    p(x) = e^{-1 - \lambda - \mu x} 
-\end{equation}
-and constraints: 
-\begin{equation} \nonumber
-    \int p(x) dx  = 1 \text { and }
-    \int x p(x) dx = \bar x.
-\end{equation}
-For simplicity, we consider $x$ to be a variable ranging in the interval $(-\infty,+\infty)$, but this can be easily computed for every other interval. The normalization conditions implies 
-\begin{equation}\nonumber 
-   1 =  \int p(x \vert I) dx = e^{-1-\lambda}\int e^{-\mu x} dx, 
-\end{equation}
-so that: 
-\begin{equation}\nonumber 
-    e^{-1 -\lambda} = \mu.
-\end{equation}
-In the continuous case MAXENT assigns for $x$ an exponential distribution
-\begin{equation}
-    p(x \vert I) = \mu e^{-\mu x}. \nonumber 
-\end{equation}
-It is clear from the functional form that the Lagrange multiplier has to be $\mu = \frac{1}{\bar x}$ to satisfy the second constraint. \\ 
-The most interesting and useful result of MAXENT is obtained when a constraint on the variance is also considered. Take a vector  $\vec x(t)$ with components in $\mathbb{R}^N$
-\begin{equation*}
-    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
-\end{equation*}
-We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals.
-If we put constraints in both mean $\vec \mu(t)$ and covariance matrix $C_(t_i, t_j)$, MAXENT allow us to assign for $\vec x(t)$ a multivariate normal distribution \cite{gregory_2005}:
-\begin{equation}
-    p\left(\vec x(t)\vert I\right) = 
-    \frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right). 
-\end{equation}
-With MAXENT, the often unjustified assumptions for the data to be normally distributed finds an elegant explanation.
-It is not only the central limit distribution for finite variance processes, it is also the distribution that maximizes the entropy and, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process.
-In some sense, we can interpret the central limit theorem as the natural 'statistical' evolution toward a configuration that maximizes entropy.
-
-\section{Time Series}
-Before we introduce MAXENT applications to spectral analysis, we provide a short review on time series and some concepts that will be used in following chapters. \\ 
-Consider a stochastic process, i.e.
-any process with some non deterministic component that takes values in some space $\bar X = \left\{x_1 \right\}$.
-A time series X of the process, i.e.
-a realization, is a set of subsequent values assumed by the variable at different times, i.e.
-\begin{equation}
-    X = \left\{x_1, x_2, \dots, x_N \right\}; \qquad \text{ with } x_i = x(t_i)
-\end{equation}
-For our goal is important to understand how a single time series can be used to infer properties on the stochastic process itself.
-A full characterization for the stochastic process can only be provided if all the joint probability distributions for the process are known
-\begin{equation}\nonumber 
-    p(x_1, t_1 | I); \quad \dots \quad ;  p(x_1, t_1; \dots x_n, t_n | I);  \quad \dots  
-\end{equation}
-or, equivalently, all momenta for the distribution are provided: 
-\begin{align}
-    \nonumber &\mu(t_1) = \int  x_1p(x_1, t_1 \vert I)dx_1  \\
-    \nonumber 
-    & R(t_1, t_2) = \int  x_1x_2p(x_1, t_1;x_2, t_2\vert I) dx_1dx_2\\
-    \nonumber 
-    \vdots \\ \nonumber
-    &\mu_n(t_1, \dots, t_n) = \int  x_1\dots x_np(x_1; \dots; x_n \vert I) dx_1dx_2\dots dx_n\\ 
-    \nonumber
-    &\vdots
-\end{align}
-Since both descriptions require $N \to \infty$, they are impossible to achieve and very often an estimate of just the first two moments is given, hence only a partial characterization can be made. \\ 
-The first momentum is known as the mean for the process, while second moment is known as 'Autocorrelation', and gives information on the deterministic dependence of a process from its previous values.
-It is a central quantity in time series analysis and is related to the autocovariance function via:
-\begin{equation}\nonumber 
-    C(t_1, t_2) = R(t_1, t_2) - \mu(t_1)\mu(t_2).
-\end{equation} \\
-Study of time series analysis is often a hard task, but there are some special cases that simplify the analysis: stationary and autoregressive processes.  
-\subsection{Stationary processes}
-\subsubsection{Stationarity conditions}
-There are several definitions of stationarity for a process \cite{Mantegna2007Introduction}. A stochastic process is said to be 'strictly stationary' if the joint distribution for the process: 
-\begin{equation}
-p(x_1,t_1; \dots; x_n,t_n \vert I) = p(x_1,t_1 + \Delta t; \dots; x_n, t_n +\Delta t\vert I)
-\end{equation}
-is constant in time whatever the order $n$ and time lag $\Delta t$.
-This implies that its moments are also constant in time.
-In this case, the time series consists in independent identically distributed terms.
-Such a process is rarely encountered and more general definitions are needed for real case studies.
-\\
-On of the most important is the 'wide sense stationary process', whose constraints are: 
-\begin{equation}
-    \mu_1(t) = \mu; \quad R(t_1, t_2) = R(t_2 - t_1) = R(\tau).
-\end{equation}
-Such a process has constant mean and autocorrelation that only depends on time lag between the considered values. \\
-The mean does not bring high information about the process, and can always be reabsorbed in the process defining 
-\begin{equation}
-    \nonumber
-    y(t) = x(t) - \mu.
-\end{equation}
-y(t) is a stationary process with 0 mean but same statistical properties of $x(t)$.
-$R(\tau)$ brings information on the amount of dependence of the time series on its previous values: if some process repeats similarly after $\bar t$ seconds, autocorrelation will have a secondary maximum at $\tau = \bar t$.
-Also, if the function is periodic of period $T$, the autocorrelation will have same periodicity.
-In general, autocorrelation represents most of the information we can obtain from the study of a stationary stochastic process, and it is rather fundamental.
-Its computation is even more important in the case of a Gaussian process.
-For such a process it is the only moment required to fully characterize it.
-One of the most important results regarding the autocorrelation function, is the Wiener-Khinchin theorem.
-It ensures that, for a wide sense stationary process, the frequency domain counterpart of the autocorrelation function is the power spectral density:
-\begin{equation}
-    R(\tau) = \int  S(\omega) e^{-\imath \omega \tau} d\omega,
-\end{equation}
-Autocorrelation and power spectral density bring the same information about the process.
-It is fundamental to develop methods to accurately estimate the power spectral density in such a way that it is consistent with the autocorrelation function.
-\subsubsection{Gaussian noise}
-MAXENT is a powerful tool to study noise.
-If one considers a wide sense stationary process  $\vec n(t) = \left(n_1, n_2, \dots, n_k \right)$ with 0 mean and known autocorrelation matrix, one can assign uniquely a normal distribution:
-\begin{equation}\label{eq:noisePDF}
-    p\left(\vec n(t)\vert I\right) = \frac{1}{\left(2\pi\det R \right)^{\frac{k}{2}}}\exp\left(-\frac{1}{2}\sum_{i, j}n_i R^{-1}_{ij}n_j\right)\,.
-\end{equation}
-In the case of a wide sense stationary process, being $R_{ij} = R(t_j - t_i)$ dependent on time steps only, the autocorrelation matrix is a Toeplitz Matrix (see Appendix \ref{app:Toeplitz}) and can be diagonalized, in the limit of large number of samples, by a Discrete Fourier Transform:
-\begin{equation}
-    S = F R F^{-1}. 
-\end{equation}
-For the Wiener-Khinchin theorem, $S$ is the power spectral density for the noise itself. Defining
-\begin{equation}
-    \tilde n(\omega) = F\left[n(t) \right],
-\end{equation}
-with $F$ the Fourier transform, the probability distribution (\ref{eq:noisePDF}) for noise in frequency domain (FD) becomes  
-\begin{equation}
-    p(\tilde n(f) \vert I) = \frac{1}{\left(2 \pi \det S \right)} \exp \left(-\frac{1}{2}\sum_{i} \frac{\tilde n_i \tilde n_i}{S_i} \right).
-\end{equation}
-Power spectral density plays a central role in noise study: it is the covariance matrix for noise in frequency domain. Also, in large N domain, frequency noise is serially uncorrelated. \\
-There is a very special case of Gaussian Noise that we have to mention: white noise. White noise can be represented as Gaussian noise with autocorrelation function
-\begin{equation}
-    R_{ij} = \sigma^2 \delta_{ij},
-\end{equation}
- it is a strictly stationary process whose autocorrelation function is 0 for any time lag different from 0, and its realization is made of independent identically distributed samples.
- Going in the Fourier Domain
-\begin{equation}
-S_{ij} = \sigma^2 F_{ik}\delta_{kl}F^{-1}_{lj} = \sigma^2 \delta_{ij} = R_{ij}, 
-\end{equation}
-we show that power spectral density for white noise is constant over frequency.
-
-\subsection{Autoregressive process}\label{sec:ARprocess}
-The description of a process is simplified if we are dealing with some autoregressive process of order p, or simply an AR(p) process.
-For an AR(p) process, each value depends linearly from its previous values
-\begin{equation}
-    x_i = \sum_{i = 1}^{p}a_{i}x_{i - p} + \nu_i, 
-\end{equation}
-with $a_i$ some constant coefficients and $\nu_i$ white noise. \\
-An AR(1) process is known as a Markov process.
-Complete knowledge on AR(p) can be achieved with the computation of $p + 1$ momenta (or conditional distributions) only.
-We prove it for a Markov process.\\
-Consider the $n'th$ order joint probability distribution
-\begin{equation}\label{eq:NthOrderPDF}
-    p(x_1, t_1; x_2, t_2; \dots; x_n, t_n | I),
-\end{equation}
-from the definition of conditional probability
-\begin{equation}\nonumber 
-p(x,y \vert I) = p(x \vert y I) p(y \vert I)
-\end{equation}
-one can rewrite \ref{eq:NthOrderPDF} as 
-\begin{equation} \nonumber 
-    p(x_n, t_n; \dots; x_1, t_1 \vert I) = p(x_n, t_n \vert x_{n-1}, t_{n-1}; \dots x_1, t_1 I) p(x_{n-1}, t_{n-1} \dots x_1, t_1 \vert I). 
-\end{equation}
-Since under Markov assumption, conditional $pdf$ only depends on the previous value for the variable 
-\begin{equation} \nonumber 
-    p(x_n, t_n \vert x_{n-1}, t_{n-1}; \dots, x_{1}, t_1 I) = p(x_n, t_n \vert x_{n-1}. t_{n-1} I) 
-\end{equation}
-reiterating the decomposition of the joint probability, we find that
-\begin{equation}
-\left(\ref{eq:NthOrderPDF} \right)=  p(x_n, t_n \vert x_{n-1}, t_{n-1}, I )p(x_{n-1}, t_{n-1}\vert x_{n-2}, t_{n-2}, I) \dots p(x_1, t_1\vert  I),
-\end{equation}
-so that for every order $n$, the probability distribution can be reduced to the product of $n$ second order probability distributions. \\
-The knowledge of the distribution for the variable and of the second order conditional distribution fully characterize a Markov process.\\
-The same reasoning can be repeated for any order $p$.
-For an AR(p) process the $n$-th order joint probability distribution ($n > p$) can be rewritten as a product of the $p$-th order conditional distributions, so that they completely characterize the process.
-
-\subsection{ARMA process and Autoregressive stationary process}
-There are two fundamental results for an autoregressive process that are strictly related with Burg's Algorithm (see section \ref{sec:MESA}) for spectral analysis, and they are all about autoregressive stationary processes and their generalization, ARMA(p,q) processes.
-\subsubsection{ARMA process}
-We will only need to define ARMA(p,q) and their relation with AR(p) processes without stepping in many details. These concepts are needed to give a stronger background to our future results.  \\
-The generalization of the Autoregressive Process, the $ARMA(p,q)$ process, is defined by the following relation: 
-\begin{equation}
-    x_t = \sum_{i = 1}^p \alpha_i x_{t - i} + \sum_{i = 1}^q \beta_i \nu_{t - i} + \nu_t,
-\end{equation}
-again, $\nu$ is white noise and both $\alpha$ and $\beta$ are constants.
-It is evident that an autoregressive process is a special case of the more general $ARMA(p,q)$ process, since  $Ar(p) = ARMA(p,0)$.
-We will not enter in the details of the description of $ARMA(p,q)$ processes, but they are important for two main reasons.
-\\
-First, $ARMA$ is a parsimonious way to characterize an $AR$ process.
-It is, in fact, possible to prove that an $ARMA(p,q)$ process, with both $p, q$ finite, is equivalent to some $AR(\infty)$ process \cite{Haykin}.
-\\
-The second and even more important result about these processes is Wold's Theorem\cite{Wold}: 
-\begin{theorem}
-Every real valued stationary stochastic process $P$ can be written as an $ARMA(p,q)$ process. 
-\end{theorem}
-This theorem makes clear the central role played by the ARMA representation: it allow us to fully characterize any stochastic process if its coefficients are known.
-This is usually an hard task, but, since we know the $AR / ARMA$ equivalence, we can find an approximate description of the process in terms of an $AR$ process, all is needed is a way to estimate the autoregressive coefficients.
-\\
-\subsubsection{Autoregressive stationary process}
-The last important ingredient is the condition of stationarity for AR processes. The conditions of stationarity for a stochastic process are well known. Defining the 'lag' operator $B$ such that 
-\begin{equation}
-    \nonumber 
-    B x_{i} = x_{i - 1},
-\end{equation}
-one can rewrite the autoregressive process as 
-\begin{equation} \nonumber 
-    x_i - \sum_{k = 1}^p a_k B^k x_i = \nu_i, 
-\end{equation}
-or, defining $b_0 = 1$ and $b_k = - a_k$
-\begin{equation}
-    \left(\sum_{k = 0}^p b_k B^{k}\right) x_i = \nu_i.
-\end{equation}
-The polynomial on the left side is known as the 'characteristic polynomial' of the autoregressive process $\phi (B) = \sum_{k = 0}^p b_k B^k$. 
-The condition of stationarity for autoregressive processes is
-\begin{theorem}
-An autoregressive process of order $p$ is stationary if all the roots $\beta_i$ of its characteristic polynomials lay outside the unit circle, i.e. $\vert\beta_i\vert > 1$.
-\end{theorem}
-Consider for simplicity a Markov process that only takes real values
-\begin{equation}\nonumber 
-    x_t - a_1 x_{t- 1} = \nu_t; \quad \text{ with } a_i \in \mathbb{R}
-\end{equation}
-whose characteristic polynomial is
-\begin{equation}
-    \nonumber 
-    \phi(B) = 1 - a_1 B .
-\end{equation}
-Let $\beta$ be the root of the polynomial $\phi(\beta) = 0$. 
-If $\vert \beta \vert > 1$ ($\vert \beta \vert < 1$) $a_1$ has to be $\vert a_1 \vert < 1$ ($\vert a_1 \vert > 1$).
-It is easy to see that if condition of stationarity is not met, the process explodes and $x_t$ will quickly diverge.
-\\
-These few concepts are fundamental in stochastic processes and necessary for what follows.
-\section{Maximum Entropy Spectral Analysis}\label{sec:MESA}
-From our previous discussion on time series, the importance of the autocorrelation function should be clear.
-Since it contains the same information of a power spectral density, high resolution computation of a power spectrum from a single time series $\vec x(t)$ of length $N$ is a fundamental task in signal analysis.
-\\
-Since we are dealing with a finite number of samples $N$, it is not possible to have an accurate computation of the autocorrelation function for every time interval.
-Also, the error $\sigma$ in the estimate increases with the time interval $k$ as $\sigma \sim \frac{1}{\sqrt{N - k}}$, so that only a few values for the autocorrelation function can actually be computed.
-\\
-The problem we want to address is how to compute a PSD from the partial knowledge of the autocorrelation function. \\
-The standard approach of peridiograms \cite{Lomb} \cite{Scargle} consists in taking the Fourier Transform of the sample autocorrelation, written as \begin{equation}
-    \rho = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\}.
-\end{equation}
-The sequence $W$ is a window function that can be chosen in several different way, each method having pros and cons for the final estimate of the power spectral density. \\ 
-The choice of a window function is arbitrary and made by trial and error, until a satisfactory compromise between variance and resolution of the estimate of power spectral density  is reached.
-A high frequency resolution implies high variance and vice-versa.
-Another problem is that the result depends on this choice, and this fact may raise the question of which we should consider as the "real" power spectral density.
-\\
-Another problem with this approach is the requirement for the window to be $0$ outside the interval: we are arbitrarily assuming $\rho_j = 0$ for $j > M$, and modifying the estimate if a non-rectangular window is chosen.
-Making assumptions on unavailable data and modifying the ones we have at our disposal does not seem a completely fair approach.
-\\
-MAXENT gives us the possibility to develop a completely new way for the estimation of PSDs, without any a priori assumptions on the unavailable data.
-Our requirements (the constraints for the variational problem) are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation.
-\\
-To do so, there is a problem we have to solve: our definition of entropy (equation \ref{eq:entropy}) depends on the probability distribution, not the power spectrum.
-We need to find a way to change the definition's domain for the entropy to formulate the variational problem in terms of the power spectral density $S(f)$ alone.
-A way out of this problem is considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA}.
-One can show that the 'entropy gain' obtained filtering white noise is:
-\begin{equation}\label{eq:EntropyGain}
-    \Delta H = \int_{-\infty}^{\infty}\log S(f) df , 
-\end{equation}
-so that we can reformulate the variational problem requiring $S(f)$ to maximize the entropy gain.
-In a real example, we have no access to all frequencies but only on those between minus and plus the Nyquist frequency, defined from the sampling interval $\Delta t$ as 
-\begin{equation}
-    Ny = \frac{1}{2 \Delta t}.
-\end{equation}
-All the quantities of interest are well defined in this range only, so that we are not going to maximize \ref{eq:EntropyGain} over all frequencies, but only on the admissible ones.
-Our objective becomes:
-\begin{equation}\label{eq:EntropyGain2}
-    \Delta H = \int_{- Ny}^{Ny}\log S(f) df.
-\end{equation}
-\par
-Now the the objective to maximize is well defined, we can write down the constraints of our problem.
-Defining $\bar r_n$ as the sample autocorrelation, we require that the variational solution $S(f)$ for the PSD matches the empirical autocorrelation. This means:
-\begin{equation}\label{eq:MaxConstraint}
-    \int_{-Ny}^{Ny} S(f) e^{\imath 2 \pi f n \Delta t} df = \bar r_{n}.
-\end{equation}
-This approach on power spectral density estimate was developed by Burg \cite{burg1975maximum} and it is a way out from the problems raised in periodograms estimate.
-It requires no assumptions on the unavailable estimates for $\bar r_n$ and is consistent with the sample autocorrelation function by construction.
-The hard task is now to show that this problem has a solution.
-Remarkably, it admits an analytical solution in a closed expression for $S(f)$, that we now work out.
-\subsection{MESA solution}
-We will follow the solution proposed by Burg in his PhD thesis \cite{burg1975maximum}.
-Instead of directly solving the variational problem for the power spectral density, we will rewrite it in terms of the constraints (\ref{eq:MaxConstraint}) of the problem.
-$S(f)$ is defined as the Fourier Transform of the autocorrelation function, as:
-\begin{equation}
-    S(f) = \frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{- \imath 2 \pi n \Delta t}.
-\end{equation}
-Substituting in \ref{eq:EntropyGain2}, the variational problem reduces to the maximization of: 
-\begin{equation}
-    \int_{-Ny}^{Ny} 
-    \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{-\imath 2 \pi f n \Delta t}
-    \right).
-\end{equation}
-As we said before, not all the values for the autocorrelation are known.
-Generally, we have access to a finite number of values for the autocorrelation function, ranging in a symmetric interval $n \in [-N, N]$.
-In the previous equation, since the summation index is ranging between $-\infty$ and $+\infty$, we are correctly assuming that also the unknown values have to be used in the computation of the power spectral density.
-But, since they are not known, we have to require the maximization of entropy to be independent of these values:
-\begin{equation}\nonumber 
-    \frac{\delta H}{\delta \bar r_s} = 0 \text{ for } \vert s \vert > N. 
-\end{equation}
-Using the explicit form for $H$, this becomes
-\begin{equation}
-      \frac{\delta H}{\delta \bar r_s} = \frac{1}{2Ny}\int_{-Ny}^{Ny}S(f)^{-1}e^{-\imath 2 \pi f s \Delta t} = \lambda_s = 0 \text{ for } \vert s \vert > N. 
-\end{equation}
-We see that $S(f)$ can be expressed in terms of the $\lambda s$ via a Fourier Series
-\begin{equation}\label{eq:PSDconstraint}
-    S(f)^{-1} = \sum_{s = -N}^N \lambda_s e^{-\imath 2 \pi f s \Delta t},
-\end{equation}
-This is a non-linear closed expression for the power spectral density. The requirement for $S(f)$ to be positive definite implies that: 
-\begin{equation}
-    \nonumber 
-    \lambda_{-s} = \lambda_s^* 
-\end{equation}
-With this property, and defining $z = e^{-\imath 2 \pi f \Delta t}$, $S(f)^{-1}$ can be rewritten as a polynomial in $z$ and $z^*$
-\begin{equation}
-    \label{eq:zExp}
-    S(f)^{-1} = \lambda_0 + \sum_{s = i}^N \lambda_s z^s + \sum_{s = 1}^N \lambda^*_s z^{-s}.
-\end{equation}
-Let $z_0$ be a root for the polynomial
-\begin{equation}
-    \nonumber
-    \lambda_0 + \sum_{s = 1}^N \lambda_s z_0^s + \sum_{s = 1}^N \lambda^*_s z_0^{-s} = 0,
-\end{equation}
-it is easy to show that $(z_0^*)^{-1}$ is also a root for the equation: 
-\begin{align}
-    \nonumber 
-    &\lambda_0 + \sum_{s = 1}^{N} \lambda s \left(\frac{1}{z_0^s} \right)^* + \sum_{s = 1}^N \lambda^*_s \left(\frac{1 }{z_0^{-s}}\right)^* = \\ \nonumber 
-    &\lambda_0 + \sum_{s = 1}^N \lambda_s (z_0^{-s})^* + \sum_{s=1}^N \left(\lambda_s (z_0^s)\right)^* = \\ \nonumber 
-    &\left(\lambda_0 + \sum_{s=1}^N \lambda_s z_0^s + \sum_{s = 1}^N \lambda^*_s z_0^{-s}\right)^* = 0 
-\end{align}
-This means that for every pole $z_0$ laying outside (inside) the unit circle, there is another pole inside (outside) the unit circle.
-So, if there are no roots on the unit circle, $N$ roots lay outside and $N$ inside of it.
-\\
-These properties allow us to rewrite the Fourier expansion (\ref{eq:zExp}) as \cite{1975STIN...7714318B}
-\begin{equation}\nonumber 
-    S(f)^{-1} = \frac{1}{P_N \Delta t} \left(1 + \sum_{s}^N a_s z^s\right)\left(1 + \sum_s^N  a^*_s z^{-s}\right),
-\end{equation}
-or, defining $a_0 = 1$,
-\begin{equation}\label{eq:MESApsd}
-    S(f) = \frac{P_N \Delta_t}{\left(\sum_{s=0}^N a_s z^z\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}. 
-\end{equation}
-The MESA expression for the power spectral density admits a non-linear, all pole representation when evaluated in $z = e^{-\imath 2 \pi f \Delta t}$.
-All the roots of the first polynomial can be chosen to lay outside the unit circle, so that all the roots of the second polynomial will lay inside of it \cite{1975STIN...7714318B}.
-If one is able to compute the values of the coefficients $a_i$ the PSD is uniquely determined from the previous equation. \\ 
-To compute the $a_i$-s, we substitute in equation
-(\ref{eq:MaxConstraint}) with the above representation for $S(f)$ in terms of the $a_i$ coefficients.
-Changing  variable from $f$ to $z(f) = e^{-\imath 2 \pi f \Delta t}$, the integral (\ref{eq:MaxConstraint}) becomes a contour integral in the unit circle centered at the origin of the complex plane:
-\begin{equation}
-   \frac{P_N}{2 \pi \imath} \oint _{\mathbb S^1}\frac{z^{-s - 1}}{\sum_{n = 0}^N a_n z^n \sum_{n = 0}^N a^*_n z^{-n}} = \bar r_s. 
-\end{equation}
-Sending $s \to s - r$, multiplying by $a^*_s$ and summing over all possible values for $s$ the equation becomes 
-\begin{align} \nonumber 
-    \sum_{s = 0}^N a_s \bar r_{s - r} &= \frac{P_N}{2 \pi \imath}\oint \frac{z^{r - 1} \sum_{s = 0}^N a^*s z^{-s}}{\sum_{s = 0}^N a_s z^s \sum_{s = 0}^N a^*s z^{-s}} dz\\
-    & = \frac{P_N}{2 \pi \imath}\oint \frac{z^{r -1}}{\sum_{s = 0}^N a_s z^s}dz\label{eq:ErFilter}
-\end{align}
-Since all the zeros of the denominator lay outside the unit circle, the contour integral is 0.
-The only exception is for $r = 0$ with a residual in the origin of the axis.
-For $r = 0$ the integral can be computed via the Cauchy integral formula and (\ref{eq:ErFilter}) becomes:
-\begin{align}\label{eq:errorFilter1}
-    \sum_{s = 0}^N a_s \bar r_{r - s} &= P_N \quad \text{ if } r = 0 \\ \label{eq:errorFilter2}
-    \sum_{s = 0}^N a_s \bar r_{r - s} & = 0 \qquad \text{ if } r \neq 0.
-\end{align}
-This kind of equations are solved by the use of the Levinson recursion \cite{doi:10.1002/sapm1946251261}
-\subsection{Levinson Recursion}
-The equations (\ref{eq:errorFilter1}) and (\ref{eq:errorFilter2}) can be rewritten as a Matrix equation of the form
-\begin{equation}\label{eq:predictionErrorFilter}
-    \begin{pmatrix}
-    \bar r_0 &  \bar r_{-1} & \dots & \bar r_{-N}\\
-    \bar r_{1} & \bar r_ 0 & \dots & \bar r_{1 - N} \\ 
-    \bar r_{2} & \bar r_{1} & \dots & \bar r_{2 - N}
-    \\
-    \vdots & \vdots & \ddots & \vdots  \\
-    \bar r_{N} & \bar r_{N - 1} & \dots  &\bar r_0
-    \end{pmatrix}
-    \begin{pmatrix}
-    1 \\   a_1 \\   a_ 2 \\  \vdots \\   a_N
-    \end{pmatrix} = 
-    \begin{pmatrix}
-    P_N \\   0 \\  0 \\ \vdots  \\   0
-    \end{pmatrix}
-\end{equation}
-know as the prediction error filter equation.
-It is a matrix equation involing the Toeplitz Matrix $R_{ij}$ and the vector $\vec a$, known as the forward prediction error filter.
-The approach for solving the equations is the Levinson - Durbin recursion.
-\\
-Our goal is to obtain the $N$-th order equation (\ref{eq:predictionErrorFilter}) from the $N-1$-th order equation: 
-\begin{equation}\label{eq:Forward}
-    \begin{pmatrix}
-    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N}\\
-    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} \\ 
-    \vdots & \vdots & \ddots & \vdots \\ 
-    \bar r_{N - 1} & \bar r_{N -2} & \dots & \bar r_0
-    \end{pmatrix} 
-    \begin{pmatrix}
-    1  \\   b_1  \\  \vdots \\   b_{N - 1}
-    \end{pmatrix} =
-    \begin{pmatrix}
-    P_{N -1} \\  0 \\ \vdots \\ 0
-    \end{pmatrix} .
-\end{equation}
-This equation is not sufficient to solve the recursion.
-Since the autocorrelation matrix is self-adjoint, from equations (\ref{eq:errorFilter1}) and (\ref{eq:errorFilter2}) and noting that the first and the last row of the autocorrelation matrix are respectively the complex conjugate reverse, is easy to show that
-\begin{equation}\label{eq:Backward}
-     \begin{pmatrix}
-    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N}\\
-    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} \\ 
-    \vdots & \vdots & \ddots & \vdots \\ 
-    \bar r_{N - 1} & \bar r_{N -2} & \dots & \bar r_0
-    \end{pmatrix} 
-    \begin{pmatrix}
-    b^*_{N-1} \\ b^*_{N- 2} \\\vdots \\ 1
-    \end{pmatrix} 
-    = 
-    \begin{pmatrix}
-    0 \\ 0 \\ \vdots \\ P_{N-1}
-    \end{pmatrix} 
-\end{equation}
-holds. The vector obtained reflecting b and taking its complex conjugate is known as ``backward prediction error". \\ 
-Stepping from order $N - 1$ to order $N$, the autocorrelation matrix grows from size $(N,N)$ to size $(N + 1, N + 1)$, this implies that a new component $b_n$ for the prediction error filter is needed.
-If we take it to be $b_N = 0$, this choice does not modify the first $N$ component of the right side of equation (\ref{eq:Forward}) (or the last $N$ components of (\ref{eq:Backward})), i.e.:
-
-\begin{equation}\label{eq:OrderN}
-    \begin{pmatrix}
-    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N} & \bar r_{- N}\\
-    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} & \bar r_{1 - N}\\ 
-    \vdots & \vdots & \ddots & \vdots & \vdots \\ 
-    \bar r_{N -1} & \bar r_{N-2} & \dots & \bar r_0 & \bar r_{-1}
-    \\
-    \bar r_{N } & \bar r_{N -1} & \dots & \bar r_1 & r_0 
-    \end{pmatrix} 
-    \begin{pmatrix}
-    1  \\   b_1  \\  \vdots \\   b_{N - 1} \\ 0 
-    \end{pmatrix} =
-    \begin{pmatrix}
-    P_{N -1} \\  0 \\ \vdots \\ 0 \\ \Delta_N
-    \end{pmatrix} .
-\end{equation}
-with 
-\begin{equation}
-    \Delta_N = \sum_{n=0}^N \bar r_{N - n}b_n.
-\end{equation}
-Adding the previous expansion for both forward and backward equations, one obtains 
-\begin{equation}\nonumber 
-        \begin{pmatrix}
-    \bar r_0 & \bar r_{-1} & \dots & \bar r_{- N}\\
-    \bar r_{1} & \bar r_0 & \dots & \bar r_{1 - N}\\ 
-    \vdots & \vdots & \ddots & \vdots \\ 
-    \bar r_{N -1} & \bar r_{N-2} & \dots &  \bar r_{-1}
-    \\
-    \bar r_{N } & \bar r_{N -1} & \dots  & r_0 
-    \end{pmatrix} 
-    \begin{pmatrix}
-    1  \\   b_1  \\  \vdots \\   b_{N - 1} \\ 0 
-    \end{pmatrix} 
-    + c_N 
-    \begin{pmatrix}
-    0 \\ b^*_{N -1} \\ \vdots  \\ b_1 \\ 1 
-    \end{pmatrix}
-    =
-    \begin{pmatrix}
-    P_{N -1} \\  0 \\ \vdots \\ 0 \\ \Delta_N
-    \end{pmatrix} .
-    + c_N
-    \begin{pmatrix}
-    \Delta^*_N \\ 0 \\\vdots \\ 0 \\ P_{N-1}
-    \end{pmatrix}.
-\end{equation}
-With the reflection coefficient $c_N$ to be computed. Requiring the previous equation to be equivalent to \ref{eq:predictionErrorFilter}, one obtains
-\begin{equation}\nonumber 
-\begin{cases}
-    P_{N - 1} + c_N\Delta^*_N = P_N \\
-    c_N P_{N - 1} + \Delta_N = 0
-\end{cases}
-\end{equation}
-which can be easily solved for $c_N$ and $P_N$ as 
-\begin{equation}
-    c_N = -\frac{\Delta_N}{P_{N-1}}; \quad \text{ and } P_N = P_{N - 1}\left(1 - \vert c_N \vert ^2 \right). 
-\end{equation}
-Having computed the value for the reflection coefficient, the $N$-th order forward prediction error filter is found to be 
-\begin{equation}
-    a_i = b_i + c_N b^*_{N - i}.
-\end{equation}
-To obtain the solution for the $N$-th order, one starts $\bar r_0 = P_0$ and recursively computes the new values until the desired order is computed. 
-This problem is often solved numerically with Burg's Algorithm \cite{Vos}.
-A python implementation of Standard Burg's Method can be found in Appendix \ref{app:Burg} 
-\section{The Autoregressive process analogy}
-The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the analysis
-of autoregressive processes.\cite{doi:10.1029/RG013i001p00183}.
-Consider an autoregressive stationary process of order $p$: 
-\begin{equation*}
-    x_t - b_1 x_{t-1} - b_2 x_{t-2} \dots b_n x_{t - n} = \nu_t, 
-\end{equation*}
-taking the z transform of the previous: 
-\begin{align}
-    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t
-\end{align}
-and calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
-in the $z$ domain, the process takes the form:
-\begin{equation}
-    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
-\end{equation}
-Since we assumed stationarity for the process, $\tilde{x}(z)$ is analytic both on and inside the unit circle.
-Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
-\begin{equation}\label{eq:ARspectrum}
-    S(f) = \vert \tilde x(z)\vert ^ 2 = 
-    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}.
-\end{equation}
-The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$. \\ 
-Comparing (\ref{eq:ARspectrum}) with (\ref{eq:MESApsd}), identifying $b_1 = - a_i$ and $P_N \Delta t= \sigma ^ 2$, the expression for the power spectral density obtained with MESA is formally equivalent to the power spectral density of some stationary autoregressive process.
-\\
-Having computed the prediction error filter one can easily obtain an estimate of the $AR$ coefficients for the process just changing sign to the prediction error filter $a_i$.
-Taken a time series, MESA allows us to compute the power spectrum and to find an effective description in terms of an autoregressive process. 
-Wold's theorem ensures that this (approximate) description for the phenomenon is legitimate when the process is found to be stationary.
-So, MESA defines a bridge between spectral estimation and the study of autoregressive processes, an unexpected and fundamental result.
-\\
-Another very remarkable process for this estimate, is that the linear predictor obtained solving the error filter equation, is the best linear predictor, i.e.
-it minimizes error, and is equivalent to a least square fit for the coefficients.
-\cite{VDBos}  \\
-To prove that it minimizes error, we will follow the original proof provided by Burg \cite{burg1975maximum}. Consider a general linear predictor $-c_i$: 
-\begin{equation}
-    x_{t} - \sum_{i = 1}^p (-c_i)x_{t - i} = \sum_{i = 0}^p c_i x_{t - i} = e'_t, 
-\end{equation}
-Denoting with $E\left[\cdot\right]$ the expectation value of some variable, the mean square error is given by: 
-\begin{equation}
-E\left[\sum_{i = 0}^p c_i x_{t-i}\sum_{j = 0}^p c^*_j x^*_{t - j}\right] = \sum_{i,j} c^*_j E\left[x^*_{t-j}x_{t-i} \right] c_j = \sum_{i, j}c^*_j R(j - i) c_i > 0,
-\end{equation}
-which, being $R$ positive definite and defined by the multiplication of a complex number times its complex conjugate, is always positive definite. 
-Let $-\vec{b}$ be the prediction error filter obtained with MESA. Multiplying the left hand side of \ref{eq:predictionErrorFilter} for the complex conjugate of the forward prediction error, 
-\begin{equation}\nonumber 
-    \sum_{i,j}b^*_jR(j-i)b_i = P_N
-\end{equation}
-is obtained . \\ 
-The (positive) quantity  
-\begin{equation}\label{eq:linearPredictor}
-    \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i),
-\end{equation}
-using Einstein notation on repeated indices and recall that $c_0 = b_0 = 1$, can be rewritten as 
-\begin{align}
-    & c^*_j R(j - i) c_i - c^*_j R(j - i) b_i - b^*_j R(j - i) c_i + b^*_j R(j - i)  b_i =\\ \nonumber 
-    &c^*_j R(j - i) c_i - c^*_0 P_N - P_N c_0 + P_N =\\ \nonumber
-    & c^*_j R(j - i) c_i - P_N.
-\end{align}
-Moving $P_N$ on the left hand side, it becomes 
-\begin{equation}\label{eq:bestLinearPredictor}
-    \sum_{i, j}c^*_j R(j-i)c_i = \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i) + P_N.
-\end{equation}
-Being all the quantities positive definite, the left hand side is minimized by the choice $c_i = b_i$.
-Also, known $P_N$, the previous equation gives an estimate on the additional mean square error associated at choices that differ from the prediction error filter.
-
-\section{Choosing an autoregressive order}\label{sec:optimizers} 
-Before being able to apply MESA, we need to address the question of the best length $M$ of the filter that provides the optimal estimate of the power spectrum.
-If the autocorrelation matrix was perfectly know, the answer would be simply $M = N$.
-\\
-In real cases, such a knowledge is not available and one has to estimate the autocorrelation function from the data.
-The accuracy for the estimate decreases as the time lag $k$ increases as $\sim \left(\sqrt{N - k}\right)^{-1}$.
-\\
-The accuracy of the estimate should then increase as the length of the autocorrelation function increases, but, from a certain point on, inaccuracy in the estimate for $R$ becomes so large that the estimate for the spectrum is negatively affected by this fact.
-A compromise between the two behaviour is proposed to be \cite{doi:10.1190/1.1440902}
-\begin{equation}\label{eq:MMAx}
-M_{max} = 2N / \ln{(2N)}
-\end{equation}
-Still, this is not a satisfactory answer since it is an upper limit for the estimate of the length $M$.
-Various approaches to optimize the order $M$ have been proposed in literature; in the following we present them.
-\begin{itemize}
-\item \textbf{Final prediction Error (FPE)} 
-The first criterion is due to Akaike (1970) \cite{Akaike1970StatisticalPI}. He proposed that M has to be chosen as the length that reduces the error when the filter is used as a predictor: 
-\begin{equation}
-    FPE: \min_{M}E\left((x_t - \hat x_t) ^ 2\right)
-\end{equation}
-with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.  This criterion is equivalent to the choice of $M$ such that 
-\begin{equation}
-    \min_{M}P_{m} \frac{N + m + 1}{N - m - 1}
-\end{equation}
-with $P_M$ defined by the prediction error filter equation. 
-\item \textbf{Criterion Autoregressive Transfer function (CAT)}
-This second criterion is due to Parzen \cite{bhansali1986}, and requires $M$ to be chosen as the length $m$ the minimizes the quantity 
-\begin{equation}
-    \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
-\end{equation}
-It has the interesting property that the distance between the original filter and the estimate is minimum and asymptotically unbiased.
-
-\item \textbf{Optimum Bayes Decision rule (OBD)} 
-The last criterion we will consider is the OBD (Rao, 1982)\cite{doi:10.1029/WR018i004p01097}.
-With the technique of asymptotic integration, $M$ is proposed to be the length that maximizes the posterior distribution, condition met selecting $m$ that minimize
-   \begin{equation}
-        (N - m - 2) \log(P_m) + m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2.
-    \end{equation}
-\end{itemize}
-Once an optimizer is chosen to select the best recursion order $M$, we then proceed as follows:  we solve the Levinson recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached.
-Then, the order $M$ is selected to be the one that minimize the chosen optimizer value.
-\\
-The following chapter will be devoted to the study of the statistical properties of the previous optimizers to understand which choice provides the best quality in the reproduction of some known power spectral densities.
-\section{Conclusions}
-In this chapter, we investigated several aspects of the maximum entropy approach to spectral analysis.
-We saw how MAXENT was originally introduced and how it solves the problem of probability assignments given some information about the phenomenon under study.
-We have shown that the best assignment we can make using no more assumptions than those given by the available evidence, is the one obtained maximizing Shannon's Entropy \ref{eq:entropy}.
-\\
-The domain of application of MAXENT is even wider.
-With a proper redefinition and analogy with filters, we have been able to rewrite the variational problem in terms of the power spectral density and the constraints in terms of the sample autocorrelation.
-This allowed us to define MESA to compute power spectral densities of a large range of spectra, obtaining an analytical representation for the power spectral density, that result to be defined for every frequency comprised between plus or minus Nyquist frequency.
-\\
-The analogy with autoregressive processes can be taken even further, and MESA is found to provide an estimate of the best linear predictor for a stationary autoregressive process.
-With the support of Wold's theorem, we can state that MESA defines a bridge between the study of spectral densities and analysis of stationary processes.
-\\
-Even more remarkable is the fact that, in principle, this approach requires no assumption at all about the process, and all the information is directly extracted from the dataset. 
-But, from a practical point of view, there is one a priori choice that one has to make: the optimizer to better approximate the length of the autoregressive process.
-This problem arises because we do not have a full knowledge of the autocorrelation function, but just an estimate whose error increases with the order of the time window.
-We briefly introduced three different optimizers, FPE, CAT and OBD, and in the next chapter we will study the properties of each of them.
-
-\section{Analysis of optimizers - Introduction}
-To apply Burg's method, it is a fundamental task to understand what is the best choice for estimating the order for the autoregressive stochastic process.
-Before applying it to real datasets, we will study main properties of the three different optimizer, introduce in section \ref{sec:optimizers} on distinct, simulated, realizations of some AR(p) processes.
-We will generate different Gauss noise time series with a given known power spectral density (PSD) to understand how each method performs in different situation.
-The analysis will concern the quality, to be defined later, of the reconstruction of the spectrum itself.
-We will consider two different kinds of spectra: i) a Gaussian spectrum with a given mean and variance; ii) an Advanced LIGO design sensitivity theoretical spectral curve.
-As a part of our investigation, we will consider both ensemble-averaged, indicated with $\bar S$ and non averaged quantities.
-\\
-In either cases, we will quantify the agreement between the reconstructed quantities and the expected ones.\\
-A short description of the principle of the algorithm used to construct the time series is also given.
-
-\section{Gaussian noise} 
-This section is devoted to the analysis of the optimizations methods, applied on pure Gaussian noise process $\vec n = (n_1, n_2, \dots, n_N),\text{ with } n_i = n(t_i)$, assuming our process to be a wide-sense stationary process, i.e.
-autocorrelation matrix is a Toeplitz matrix depending only on time intervals.\\
-The spectrum chosen for our first analysis is a Gaussian distribution
-\begin{equation}
-    \nonumber
-    S(f) = \frac{1}{\sqrt{2 \pi} \sigma}e^{-\frac{(f - \mu)^2}{2 \sigma ^2}}
-\end{equation}
-with mean $\mu =2.5 \quad a.u.$ and standard deviation $\sigma =  0.5 \quad  a.u.$. 
-To generate noise with a power spectrum that is known, we have to first understand what is the role played by the PSD. \\ 
-Via MAXENT principle, we can state that noise in normally distributed with 0 mean and some covariance matrix $C_{ij} = C(t_i, t_j)$, i.e.
-\begin{equation}
-    \nonumber
-    p(\vec{n}(t) \vert I) \propto \exp{\left\{-\frac{1}{2}n_i C^{-1}_{ij}n_j\right\}},
-\end{equation}
-\begin{figure}
-    \centering
-        \includegraphics[width = \linewidth]{Images/Noise and PSD/NormalPSDautocorr.pdf}
-        \caption{A priori spectrum (left) and autocorrelation (right)}
-        \label{fig:autocorr}
-\end{figure}and, since process has zero mean, autocovariance and autocorrelation matrices are identical. \\ 
-Using the properties of Toeplitz matrices (see appendix \ref{app:Toeplitz}) and the relation between autocorrelation and power spectral density, it is immediate to verify that, in the limit of large number of time samples, the distribution of noise in frequency domain is given by:
-\begin{equation}
-    p(\tilde{n}(f_i) \vert I) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}},
-\end{equation}
-so that the role of PSD appears to be clear: it is the diagonal covariance matrix for the noise distribution in Fourier Domain: \\
-\begin{equation}
-    \nonumber
-    <\tilde n_i \tilde n_j > = S(f_i)\delta_{ij}.
-\end{equation}
-With this information, it is easy to generate any time series with power spectral density of any shape.
-Choosing a priori values for sampling interval $dt$ and total time of observation $T$, recalling Nyquist theory of sampling, we generate arrays of given length N / 2 in frequency domain from 0 to the Nyquist Frequency $f_{Ny} = \frac{1}{2 dt}$.
-We then convert noise to the time domain using an Inverse Fast Fourier Transform.\cite{oliphant2006guide}\cite{van2011numpy} \\ Basic idea for fast algorithm used to construct noise can be found in ~\cite{noiseGen}.
-We worked it out in a slightly different way, the imaginary component is not obtained sampling a random complex phase from a uniform distribution.
-We choose to sample both real and imaginary part from two identical normal distribution distributing variance in a proper way between real and imaginary components.
-\section{Test on Time Series}
-It is now possible to test the three methods on a given time series.
-We will concentrate on a power spectral density that is a normal distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$.
-We will apply Burg method with each optimizer on a large amount of simulated dataset, obtaining a variety of estimate for PSD.
-We will study statistical properties for both single reproductions and ensemble average.
-In many cases, such as gravitational waves, one cannot observe multiple time series realizations for a fixed process, so that single reproduction studies are probably more useful.
-Still, study of ensemble average is important to highlight some interesting properties.
-In this way we can fully characterize the performance of each optimizer
-We will attentively study the role of the filter length estimate M in determining the accuracy of the reconstruction.
-We will consider every optimizer separately and compare result only at the end of this section.
-We will then study in depth the Advanced LIGO design sensitivity power spectrum \cite{Ligo}
-\subsection{Normal Power Spectral Density}
-\begin{figure}
-    \centering
-        \includegraphics[scale = .43]{Images/NormalPSD/NormalNoise.pdf}
-    \caption{Noise in both time and frequency domain. Frequency noise behaviour is representative of the chosen spectrum, having its maximum amplitude around $f = 2.5 au$}
-    \label{fig:noise}
-\end{figure}
-Fig \ref{fig:noise} shows a realization of the gaussian process in the frequency (left panel) and time domain (right panel).   \\ 
-To study noise with normal-shaped PSD, we generate 1000 datasets of 3000 points each.
-This will provide a number of samples that is large enough to perform statistical analysis.
-For every dataset we will reproduce its spectrum with every optimization method.
-We will show that the quality of reconstruction is very different in different parts of the spectrum: the minimum of the error will be achieved around the peaks, increasing fastly in the tails
- \\ \\ 
-\subsubsection{Final Prediction Error (FPE)}
-The estimate of the recursive order obtained with FPE on the 1000 realization of noise, is found to lay between 30 and 70.
-Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum: a logarithmic plot of the mean of all the spectra versus the true spectrum is reported in \textbf{Figure \ref{fig:FPEmean}} with 90\% credible regions.
-It is evident that, despite not too large error bars, the tail of the normal distribution is not comprised in the interval of the estimation.
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = .4]{Images/NormalPSD/FPEpsdAndResiduals.pdf}
-    \caption{Ensemble Average for FPE reconstruction with 90\% Credibility Regions (above) and percentage error (below)}
-    \label{fig:FPEmean}
-\end{figure}
-To assess the magnitude of the disagreement, we compute the relative error between the mean spectrum $\hat{S}$ and the true spectrum $S$ for every frequency bin
-\begin{equation}
-    \nonumber
-    r(f) = \frac{|\hat{S}(f) - S(f)|}{S(f)}\,.
-\end{equation}
-We found this to be a better choice than residuals since relative errors are not dependent on the magnitude of the data.
-Second Fig \ref{fig:FPEmean} (below) shows that the largest error in the reconstruction of the spectrum is in the tails, while in the bulk of the distribution we have a good approximation of the true function.
-For the entire spectrum, the mean relative error $\Bar{r}$ is 
-\begin{equation}
-    \nonumber
-    \bar{r}_{\bar S, FPE} = 0.025.
-\end{equation}
-FPE performs well in reconstructing the spectrum even with relatively short time series, with a deviation from the true spectrum around 2\%, but it is evident that it is not able to capture all the features of the spectrum, since the reconstruction in the tails is inaccurate.
-If we compute the average percentage error in three standard deviations from the mean we gain an order of magnitude estimate of the accuracy of the reconstruction, obtaining:
-\begin{equation}
-    \nonumber
-    \bar{r}_{\bar S, FPE}^{\mu \pm 3\sigma} = 0.005. 
-\end{equation}
-and we can state the the method is highly performant in this frequency window. 
-Since it is not always possible to work with more then one realization for a process, as in observational experiments, it is important to understand how this method behaves when we apply it on a single time series.
-In that case, the mean error and its standard deviation are:
-\begin{equation}
-    \nonumber
-    \bar r_{FPE} = 0.16; \quad \sigma_{r} = 0.16\,.
-\end{equation}
-The average error of a reconstruction based on a single time series is bigger than the one obtained considering the average of all the reconstructions.
-Standard deviation of the errors assume small values and 90\% credibility regions (Cr.) in \ref{fig:FPEmean} are tight.
-This means that if we randomly take two of the reconstructed spectra, we expect their differences to be statistically small.
-These facts are evidence for FPE to be a relatively stable optimizer.
-We already said that the estimate for filter's length lays in a well defined range of values.
-The quality of the reconstruction, as noted in Chapter 1, depends on the order of the reconstructed AR process.
-To investigate this idea further, we plot the value of the optimizer for each recursion and the residuals as a function of the estimate for the order of the autoregressive process (\textbf{Figure \ref{fig:FPEErrorOrder}}),
-
-\begin{figure}[H]
-    \noindent
-    \centering
-    \includegraphics[scale = .3]{Images/NormalPSD/FPEoptimizer.pdf}
-    \caption{Stability of FPE, that results in a cluster of values in a small region (left), is due to the typical behaviour of the optimizer as a function of the AR order (right).}
-    \label{fig:FPEErrorOrder}
-\end{figure}
-Foremost evidence for the stability of this optimizer is found in its values as a function of the AR order.
-Figure \ref{fig:FPEErrorOrder} shows that Values tend to decrease monotonically until a sharp and well defined minimum is reached, to start growing again after this area.
-This is representative of the typical behaviour of this optimizer as a function of the length of the filter, and explains why the estimates for AR orders are confined in a small region.
-This plot also shows that most of the values collapsed in the region of AR orders comprised between 30 an 50.
-There are only few values outside this interval, and get fewer as order's estimate increases.
-In this region, there is no strong evidence of some correlation between the estimate for M and its associated relative error.
-It is important to notice that the estimate associated with the higher values of M (around $M > 60$) exhibit the highest 'mean' relative error.
-\\
-We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length M is clustered in a region where there is no dependence of error from M.
-\subsubsection{Optimum Bayes Decision Rule (OBD)}
-The second optimizer in consideration is OBD.  (\textbf{Figure \ref{fig:OBD3000}}) shows a good agreement between the average over the reconstructions and true spectrum.
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = .45]{Images/NormalPSD/OBDpsdAndResiduals.pdf}
-    \caption{Reconstruction of the spectrum from 1000 simulations with OBD optimizer. Plot is mean of all the results with 90\% credibility regions (above) and percentage error (below)}
-    \label{fig:OBD3000}
-\end{figure}
-The same behaviour of FPE is observed, a good quality reconstruction at the center of the distribution with quite small Credibility regions (Cr.) at a $90$\% credibility level.
-Qualitatively they behave similarly but quantitatively the disagreement of this method is found to be even larger than the one obtained with FPE, as can be noticed in the plot of relative errors in figure \ref{fig:OBD3000} (below) where at the extremes the inaccuracy is even larger than 300\% with respect to the real value of the spectrum.
-For the overall spectrum and within three standard deviation from the mean, we find that the reconstructed spectrum has a deviation of
-\begin{equation}
-    \nonumber
-    r_{\bar S, OBD} = 0.15, \qquad r^{\mu \pm 3\sigma}_{\bar S, OBD} = 0.018. 
-\end{equation}
-When applied to single realizations of the time series, we find
-\begin{equation}
-    \nonumber
-    \bar{r}_{OBD} = 0.237, \qquad \sigma_{r, OBD} = 0.0296.
-\end{equation}
-It is interesting to notice that overall relative error of the average spectrum is of the same order of magnitude of the average single reconstruction for the spectrum.
-Averaging over a 1000 realisations does not substantially improve our results.
-As a function of the filter length, the values taken by the OBD cluster in a small region, indicating the stability of the reconstructed process order $M$.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = .3]{Images/NormalPSD/OBDoptimizer.pdf}
-    \caption{Stability of OBD, that results in a cluster of values in a small region (left), is due to the typical behaviour of the optimizer as a function of the AR order (right)}
-       \label{fig:OBDErrorOrder}
-\end{figure}we see that again our results are clustered in a very small region, due to the behaviour of the optimizer that converge in a small area before growing again.
-The estimate of the length of the filter is around $p = 20 \div 30$.Note the difference compared to FPE (see \ref{fig:FPEErrorOrder}); for OBD the mean relative error depends on $M$ as
-\begin{equation}
-    \hat r \sim M ^ {-2.016 \pm 0.006}.
-\end{equation}
-In summary, OBD shows greater error wherever in the spectrum, due to the fact that it choose filters that are too short and in a region where quality drastically change when changing M.
-In this case we can say that we have a stable estimate for M, but due to this dependence, the estimate is not as stable as FPE's.
-\subsubsection{Criterion Autoregressive Transfer function (CAT)}
-This last criterion differs substantially from the two other optimizers. Figure \ref{fig:CATmean} (above) shows a good agreement between the estimate of the spectrum for every frequency
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = .43]{Images/CAT/CATnormal.pdf}
-    \caption{Mean spectrum for CAT optimizer, 1000 simulations for 3000 points datasets}
-    \label{fig:CATmean}
-\end{figure} and also tails are well captured by the average spectrum.
-However, the variance of the reconstructed spectrum is quite large, much larger than for FPE and OBD.
-In general, the relative error is nearly constant, $\sim 10\%$ over all frequencies, see  Fig.
-\ref{fig:CATmean} (below).
-The residuals, averaged over all frequencies, are:
-\begin{equation}
-    \nonumber
-    \bar r_{\bar S, CAT} = 0.016 \qquad \bar r^{\mu \pm \sigma}_{\bar S, CAT} = 0.017.
-\end{equation}
-the overall reconstruction is thus quite accurate, with deviations from the simulated spectrum of the order of $1\% \div 2\%$.\\
-The picture drastically changes in the case of individual time series realizations. The average residuals are, in fact
-\begin{equation}
-    \bar r_{CAT} = 0.42 \qquad \sigma_{CAT} = .44\,.
-\end{equation}
-The error is indeed $\sim 42\%$, hence much larger than in the case of FPE and OBD.
-Both large error bars and the large standard deviation are indications of the instability of this optimizer.
-This is exemplified by the variety of different spectra that can be obtained considering different time series.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = .43]{Images/CAT/CAToptimizer1.pdf}
-    \caption{CAT is not a stable optimizer since its optimizer do not converge to any specific region of the AR estimate order}
-    \label{fig:CAToptimizer}
-
-\end{figure}
-The reason for this behaviour, can be found in \textbf{Figure (\ref{fig:CAToptimizer})}.
-CAT does not converge to any specific value for the order of the autoregressive process.
-In turn, this results in estimates for the order that span a large range.
-Figure \ref{fig:CAToptimizer} also shows a strong dependence of the error on the estimated length of the filter.
-The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like the big excursion in the size of datas of a normal distribution, but this length is also responsible of an augmented variance in the estimate, introducing spurious peaks in the reproduction (Figure \ref{fig:GaussMedian})
-
-Averaging, the noise in each spectrum is cancelled.
-It is interesting that the cancelling of errors in averaging is property of gaussian noise (the process is gaussian, after all).
-This implies, in a sense, that each estimate of the spectrum is independent of any other, as suggested by the huge variance in the residuals.
-This instability is not a good property for an estimation, so that this estimator might be found to be optimal for repeatable experiments and not for single observation of some process.
-Another approach to this method will be proposed in section \ref{sec:ACAT}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale = .45]{Images/NormalPSD/Gaussmedian.pdf}
-    \caption{Reconstructed power spectral density with the three method. The reported reconstruction is the one that shows the median percentage error (overall)}
-    \label{fig:GaussMedian}
-\end{figure}
-\subsection{Summary and Comparison}
-\begin{figure}
-    \centering
-    \includegraphics[scale = .3]{Images/NormalPSD/Normal Residuals.pdf}
-    \caption{Percentage error in the reconstruction for every optimizer, as a function of the frequency. Lower panel restricts the domain to three standard deviation to increase the definition}
-    \label{fig:NormalEnsembleResiduals}
-\end{figure}
-In our analysis, FPE and OBD optimizers are found to behave similarly while CAT shows completely different properties.
-Studying the average spectrum is not sufficient to choose which method behaves better.
-\\
-The bad behaviour in the reconstruction of the tails from OBD and FPE might be due to the fact that very small values require longer filters. 
-In fact when we consider the average for all the spectra we note that the shorter the filter the larger the error in the tails: OBD, which select the shortest filters, arrives to a $300 \%$ error at the extrema of the power spectral density, FPE which select intermediate order has got a $85\%$ error in same window, while CAT, that selects a wide range of long filters, has got a constant accuracy everywhere in the spectrum.
-\\
-Figure \ref{fig:NormalEnsembleResiduals} shows that CAT is the method that minimizes errors in the tails (upper plot), but if one restrict to three standard deviations from the mean, FPE is found to reach best accuracy (lower plot).
-The frequency averaged percentage error is reflecting this behaviour. Considering the whole reproduction, error's average and standard deviation are found to be: 
-\begin{align*}
-\centering
-&r_{\bar S, FPE} = 0.025; \quad \sigma_{\bar S, FPE} = 0.083;  \\
-&r_{\bar S, OBD} = 0.15;\quad \sigma_{\bar S, OBD} = 0.42 ; \\ 
-&r_{\bar S, CAT} = 0.015; \quad \sigma_{\bar S, CAT} = 0.012
-\end{align*}
-In is interesting to see that the average for FPE and OBD is strongly affected from the high value of the error in the tails of the distribution.
-If one consider the three standard deviations from the mean, the value of the error substantially decreases,
-\begin{align*}
-\centering
-&r_{\bar S, FPE} = 0.005; \quad \sigma_{\bar S, FPE} = 0.004;  \\
-&r_{\bar S, OBD} = 0.018;\quad \sigma_{\bar S, OBD} = 0.019 ; \\ 
-&r_{\bar S, CAT} = 0.015; \quad \sigma_{\bar S, CAT} = 0.011
-\end{align*}
-showing that the overall percentage error is just a first indicator for the quality of the reproduction, but is not sufficient to perform a full characterization. \\
-It is evident that tails do not affect CAT  average error and standard deviation at all, so that it's accuracy is approximately constant everywhere in the spectrum. \\ \ 
-Since both FPE and CAT shows pros and cons, at this level, there is no reason to choose between them.\\ 
-Analyzing the resolution of the single reconstruction, figure \ref{fig:optcomparison} clearly shows that FPE minimize overall relative error giving a stable result.\\ 
-CAT instead shows a larger error and do not converge to any specific value.
-Almost every reproduction obtained with CAT is worst with respect to those obtained with FPE, but their mean is more precise.
-\\
-It's instability is probably the cause for this behaviour. Since it does not show a strong convergence to some order with respect to others, very different lengths are selected.
-Longer lengths are able to capture a lot of details but are more likely to introduce noise. In other words, long filters reduce bias but increment the variance of the result. 
-Taking the average, for central limit theorem the error decreases as $\sim \frac{1}{\sqrt{N}}$. In this way we compensate noise and obtain a spectrum with small bias and small fluctuations.
-Considering the average, CAT reconstruct precisely and with good accuracy every part of the spectrum, even if in "regular" parts best accuracy is reached from FPE.
-For a single reproduction, CAT is very noisy and unstable, while FPE performs the better.
-\ref{fig:optcomparison}
-\begin{figure}
-    \centering
-    \includegraphics[scale = .3]{Images/NormalPSD/NormalPSDcomparison.pdf}
-    \caption{Relative error as function of the length of the filter.}
-    \label{fig:optcomparison}
-\end{figure}
-\begin{figure}
-    \hfill
-    \centering
-    \includegraphics[scale = .45]{Images/NormalPSD/ordersComparison.pdf}
-    \caption{Histogram of estimate orders for each method}
-    \label{fig:ordersCompairson}
-
-\end{figure}
-definitely shows that FPE gives the best results: it is a stable method that choose filters in the clustered area associated with minimum error, while both OBD and CAT result lies outside this area and show bigger errors.
-Two other interesting plots that stress this fact are reported in figure \ref{fig:ordersCompairson} and \ref{fig:residualsComparison}
-\begin{figure}
-    \centering
-    \includegraphics[scale = .46]{Images/NormalPSD/ResidualsComparison.pdf}
-    \caption{Histogram of residuals for each method}
-    \label{fig:residualsComparison}
-\end{figure}
-where the histograms for the errors and for order's estimate are reported.
-These graphs provide stronger evidence for our previous statements.
-As already noticed, FPE is very stable and provide the smallest 'overall relative error'.
-\\
-CAT histogram shows a very interesting behaviour.
-The histogram for the order show two different peaks,  a small one associated to 'short filters', and a second high peak at $m = M$.
-This means that CAT is most likely to not converge at all at some value for the regression order, and it is more likely selects a length that is comparable with the largest available.
-If short filters are selected, the overall percentage error will be low, respectively represented by the first peak in the distribution of the errors, while if long filters are selected, the error is maximized.
-
-So, even if in some cases CAT might be more accurate when taking the average over several spectra, FPE guarantees that minimum error will be reached.
-
-\subsection{Analysis of Ligo Spectrum}
-\begin{figure}
-    \centering
-    \includegraphics[scale = .45]{Images/LIGOsimulate/LigoPSDoriginal.pdf}
-    \caption{Original power spectral density from Ligo}
-    \label{fig:LigoPSDOriginal}
-\end{figure}
-In this subsection we concentrate again on the reconstruction of a specific, known, power spectrum.
-We analyze the properties of reconstruction of this second spectrum whose shape is the Advanced LIGO design sensitivity theoretical spectral curve (Figure \ref{fig:LigoPSDOriginal}). 
-For this analysis we will study properties only for arrays of 40960 points.
-We consider a sampling rate of $2048 samples / second$ that means we can reconstruct a Nyquist frequency of $1024 Hz$.
-We chose a longer dataset than before because algorithm works with a 1 dimensional interpolation of spectrum, and with shorter datasets the interpolated spectrum is not consistent with the original spectrum and does not capture the height of the peaks.
-Shorter choices would result in noise extracted from a different power spectrum than the one in figure \ref{fig:LigoPSDOriginal}.
-\\
-We will study statistical properties from samples obtained with 500 simulations and compare the obtained results for each optimizer. 
-We start considering statistical properties of mean spectrum obtained averaging over all the simulations and conclude with an analysis for the statistics of the reproduction of single spectrum. 
-\subsubsection{Properties for the average spectrum}
-Here we report mean properties of the reproductions of the spectrum. The spectrum we want to reconstruct is reported in Figure \ref{fig:ligospectrum},
-\begin{figure}
-    \centering
-
-        \includegraphics[scale = 0.45]{Images/LIGOsimulate/PSD.pdf}
-        \caption{A priori spectrum from Ligo database}
-        \label{fig:ligospectrum}
-\end{figure}
-shows the original power spectral density and the ensemble average of every method. For the mean, the relative errors are: 
-\begin{equation}
-    \bar r_{\bar S, FPE} = 0.0426\%, \qquad \bar r_{\bar S, OBD} = 0.129\%, \qquad \bar r_{\bar S, CAT} = 0.0211\%
-\end{equation}
-Again, if one only check overall accuracy for every method, CAT is outperforming.
-In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them.
-To check this behaviour we take a look at the two peaks separately.
-For the first one a log plot of the reconstructed peaks and their relative errors are shown in Figure \ref{fig:ligo1peak}
-\begin{figure}[h]
-        \includegraphics[scale = 0.45]{Images/LIGOsimulate/1stPeakComparison.pdf}
-        \caption{1st peak of the spectrum and it reconstruction with every optimizer (left) and associated percentage error (right)}
-        \label{fig:ligo1peak}
-\end{figure}
-and the results reinforce our previous statement.
-The behaviour of the error is very similar for CAT and FPE, showing similar shapes but very different sizes, while CAT is giving a different and way more precise result.
-Both FPE and OBD are not able to capture the width of the peak and also show larger errors, while CAT has a good behaviour in reconstructing width's also.
-The reconstruction for very sharp peaks considering the average over all the spectra is found to be strongly inaccurate for both FPE and OBD, while a good accuracy is reached with CAT method. \\ 
-Considering the estimate for the position of the peak, whose true value is found to be $16.85 \quad a.u.
-$, within numerical accuracy both FPE and CAT find its exact position, while OBD shift the peak left at $16.7 a.u.$, with a $0.9\%$ inaccuracy.\\
-Second peak results are analogous, plotting the peak and the errors value around it
-\begin{figure}[h]
-    \centering
-        \includegraphics[scale = 0.45]{Images/LIGOsimulate/2ndPeakComparison.pdf}
-        \caption{2nd peak of the spectrum, Relative error around 2nd peak of the spectrum}
-        \label{fig:ligo2peak}
-\end{figure}
-we find again that FPE and OBD are performing similarly, the magnitude of the error is maximum for OBD and minimum for CAT. \\ 
-In this case even CAT shows large errors in some points, but they are reasonable in a large part of it, still, it remains the best reconstruction.
-For this second peak, all the methods find maximum position to be at $438.05 \quad a.u.$, which is the true position of the maximum.
-Looking at Figure \ref{fig:ligo1peak} at \ref{fig:ligo2peak}, it appears to be clear that the least of the accuracy is not on the peak, but immediately around, so that one of the biggest deal is the reconstruction of peak's width, that only CAT is able to capture with high definition.
-\\
-Even if FPE shows larger error in the peaks than CAT, when the average over all the frequencies is considered, their percentage errors are of the same order of magnitude.
-This means that outside the peaks FPE is more accurate than CAT, i.e.
-CAT is still giving noisy results.\\
-As for the normal distributed noise, mean spectrum is not sufficient to declare what is the best optimizer to be used, since most of experiment are purely observational and mean spectrum over different realizations of the same stochastic process cannot be computed.
-So, it is important to also understand what is their behaviour for single realizations of the process.
-\subsubsection{Statistical properties of single realizations} 
-The errors for the single realizations behave very differently from the errors obtained when we considered the mean spectrum.
-In computing the mean of the overall errors for each single spectrum, the results we obtain are the following:
-\begin{equation}
-    \bar{r}_{FPE} = 0.13 \pm 0.01; \qquad \bar{r}_{OBD} = 0.17 \pm .02; \qquad 
-    \bar{r}_{CAT} = 0.4 \pm 0.1
-\end{equation}
-where the error is just the standard deviation of the samples.
-In this case we find a different result: FPE shows the best results and order's estimate is associated with least errors.
-Also, having found a smaller standard deviation of the datas, we can state that FPE and OBD are more stable optimizers then CAT.
-In fact, their properties as a function of filter's length are almost identical to those obtained in the study of normal spectrum.
-If we again plot the value of the errors as a function of the estimate of the order (Figure \ref{fig:LigoOrderError})
-\begin{figure}
-    \centering
-    \includegraphics[scale = .43]{Images/LIGOsimulate/orderVSerror.pdf}
-    \caption{Plot of relative error as a function of the estimate for filter's length}
-    \label{fig:LigoOrderError}
-\end{figure}
-it is immediate to see that, again, FPE results lies in the are of filter's length associated with a minimum of relative errors, OBD lies in a clustered window with a higher error and CAT is not convergent, but most of the results are outside the minimum error area.
-So, we arrive at the same conclusion we had when we considered the normal specturm.
-OBD has no general properties that make it a preferable optimizer then FPE or CAT while, what to choose beneath the last two is based on the problem we are dealing with: if we have one single realization for the process, we are mostly sure that FPE would catch the area associated with the best resolution possible, while CAT would result in a spurious result with, in general, a not so good resolution.
-But, in case we have several realization for the same process, this last property of CAT, together with its instability, result to be his luck and, taking the average errors tend to compensate noise in the result, originating a very well defined spectrum that show same resolution almost everywhere, no matter how small the values are (as for normal PSD case) or how sharp peaks can be, the result has got generally good accuracy, also if it is clear that we have to pay this fact with very large error bars.
-So a choice between these methods is a matter of convenience and interest for each specific problem.
-\\
-\begin{figure}
-    \centering
-    \includegraphics[scale = .43]{Images/LIGOsimulate/LigoHists.pdf}
-    \caption{Histograms showing the estimate for the autoregressive order and overall percentage error obtained with all the three optimizers.}
-    \label{fig:my_label}
-\end{figure}
-
-\section{Another approach to CAT} \label{sec:ACAT}
-We stated that, if one consider a single realization for some stochastic process the choice for the optimizer should rely on FPE because of its stability and better single estimate, while CAT, when applied on a single dataset, does not assure a good resolution for our estimate.
-\\
-It was found that CAT shows its best properties when considering the average over a big number of reconstructions, where noisy results tend to compensate each other reaching a very good quality reconstruction even for not too long arrays (as in the case of Gaussian PSD where we considered simulations of 3000 points).
-This led us to propose another approach to CAT: instead of applying Burg's method on the whole data set, we can divide it in smaller arrays of data and apply it to each one of them taking their mean at the end of the computation.
-In this way, accuracy can be made incredibly higher than the one obtained with the standard approach.
-\\
-Unluckly, it is very hard to investigate how one should divide the dataset to reach minimum error.
-If one consider a known power spectrum, the problem could be solved numerically, but unuckly in real life experiments we do not know what true spectrum is, and such an approach is non-sense.
-\\
-The non linear form for the estimator together with the finite-order recursion make it hard or even impossible to find general properties for the estimate, and so we will not investigate many of that questions that could arise.\\
-For examples, what is the best choice for the length of the sub data-set?
-If we choose too short arrays, we will mediate over a lot of different spectra, decreasing variance, but it is probable that we will not be able to capture all the features of the spectrum.
-Too long subsets, instead, would result in too few reconstructed spectra, so that taking their average would not reduce variance in a significant way.
-It seems a very hard task to answer this question uniquely, and maybe such an answer does not even exist in a general case.
-One prescription on the length of the sub arrays may concern some consideration about frequencies: if we want to accurately reconstruct low frequency behaviour, one has to make long time observations, so for those cases longer datasets are required.
-Since taking overlapping arrays does not seems to affect negatively single estimates, this could be a good prescription to maximize the number of reconstructions for a given length. \\
-Stressing to its maximum the hypothesis that such a procedure is harmless, one could try any overlapping sequence to maximize the number of reconstructions.
-To study this approach, since we want every reconstruction to be more than slightly different from each other, we chose an overlapping that is half the length for the sub arrays.
-We test this new approach on LIGO dataset on the same noise simulations as before. We chose sub segments of length L = 3000, 5000, 7000 and 10000.. \\
-In figure \ref{fig:alternativeCAT} we compare the result associated with with this new approach for the lengths written above.
-For each of the 500 reproductions, we chose the one associated with the median overall error.
-\begin{figure}
-    \centering
-    \includegraphics[scale = .45]{Images/AlternativeCAT/AlternativeCAT.pdf}
-    \caption{Alternative CAT approach to Ligo Spectrum for different lengths of the sub arrays}
-    \label{fig:alternativeCAT}
-\end{figure}
-As one can expect from CAT behaviour results obtained with larger arrays are associated with increasing noise.
-Considering the reconstruction for single peaks, it is found that accuracy is certainly dependent from the length, but it seems to give the best result for $L = 7000$, and not for $L = 10000$ as we were expecting.
-This difference can be understood if one considers the number of realizations over which we take our average: for arrays of $7000$ points we are taking the average over 12 different realizations, while for arrays of $10000$ points we are taking our average on eight different realizations only.
-This is probably the reason why this higher length is not associated with higher resolution.
-We clearly see that the reconstruction achieved using $L = 7000$ is the best in capturing the width of the peaks and their position, but none of the reconstructions is able to accurately reproduce the second peak.
-We want to compare these results with the standard approach, and with both the results obtained with FPE and with CAT.
-
-\subsection{Alternative CAT vs FPE}
-The first comparison we make between FPE and alternative CAT regards the distribution of the overall errors. 
-\begin{figure}
-    \centering
-    \includegraphics[scale = .45]{Images/AlternativeCAT/histACatVSFpe.pdf}
-    \caption{Histogram of the percentage error as a function of the length of the sub arrays. The blue panel shows the histogram obtained from FPE}
-    \label{fig:aCATvsFPEhist}
-\end{figure}
-Figure \ref{fig:aCATvsFPEhist} shows the distribution of the percentage error as a function of the length of the segments of the data.
-The support of the histogram is, for every length, shorter than the support the one expect to be obtained with CAT (y-axis of figure \ref{fig:LigoOrderError}), so that this approach is found to be more stable than standard CAT.
-\\
-Shorter arrays are found to minimize the error, as could be expected from the fact that the variance is reduced from this choice, but non of the chosen lengths reaches the overall accuracy of FPE. \\ 
-From the previous histograms, the most meaningful comparison of FPE with alternative CAT is obtained using as preferable length the reconstruction '3000'. \\ 
-The mean and standard deviations of the percentage error are, in the two cases 
-\begin{align} 
-    \bar &r_{FPE} = 0.13; \qquad 
-    \sigma_{FPE} = 0.01 \\ 
-    \bar &r_{aCAT} = 0.158; \quad 
-    \sigma_{aCAT} = 0.007
-\end{align}
-With this alternative approach, we have been able to obtain reconstructions for the spectra whose accuracy is almost comparable with the one obtained with FPE.
-But the comparison of the overall error is just a rough indicator of the accuracy in the reconstruction, for that reason we compare the median error reconstruction of both methods.
-\begin{figure}
-    \centering
-    \includegraphics[scale = .45]{Images/AlternativeCAT/FPEaCATwhole.pdf}
-    \caption{Comparison of FPE method with alternative cat with length of 3000 points (above panel) and their percentage error (below panel)}
-    \label{fig:aCATvsFPEwhole}
-\end{figure}
-Figure \ref{fig:aCATvsFPEwhole} shows both the reconstruction obtained with FPE and alternative CAT (above panel) and the percentage error for each frequency (below panel).
-It is evident that the two result are very similar.
-Alternative CAT provides a sightly better accuracy in the reproduction for the first peak.
-But, for the rest of the reconstruction FPE has to be preferred.
-Regarding the second peak, FPE accuracy is largely better.
-Alternative CAT flattens the peak and its maximum is not localized, while the reconstruction of FPE is representative of true position of the peak.
-Also, in the rest of the spectrum the result of FPE still shows less noise.
-\\
-From thie brief comparison, we can conclude that alternative CAT approach shouldn't be used with "short" arrays (short with respect to the total length N of the time series), since FPE provides a better estimate.
-\\
-Since for short arrays the best quality is still achieved by the use of FPE, a second interesting comparison is the one between FPE with the alternative CAT estimate on longer arrays.
-From figure \ref{fig:alternativeCAT}, it is evident that the best compromise between noise and accuracy in the reproduction is obtained by the arrays of 7000 points.
-It is clear that, outside peaks, the smaller error is achieved by FPE, so that te interesting part will be the expected gain in the definition of the peaks.
-\begin{figure}
-    \centering
-    \includegraphics[scale = .43]{Images/AlternativeCAT/FPEaCAT7000.pdf}
-    \caption{Comparison of the alternative CAT with 7000 points subsegments of data with FPE}
-    \label{fig:aCAT7000FPE}
-\end{figure}
-Figure \ref{fig:aCAT7000FPE} shows the reconstruction obtained with FPE (red line) and the one obtained with the alternative CAT approach, for sub-segments of 7000 data and a 50\% overlapping.
-This result is more interesting than the previous.
-As expected, the noise in the reproduction is larger, as can also be noticed looking at the error in the below panel, but the introduced noise is a reasonable quantity.
-Concerning the peaks, the first peaks reconstruction has gained a large accuracy in the reproduction, for both its width and position.
-The second peak is reconstructed in a comparable way for both the methods.
-FPE captures in a better way the position of the peak, while alternative CAT gains in the reproduction of the width and general shape.
-\\
-Concluding, this new approach might be a valid alternative to FPE if 'not-too-short' sub segments of data are chosen.
-In this way, introducing a reasonable amount of noise, we are able to capture some details with more accuracy.
-But if we deal with shorter datasets, like 3000 points, without any a priori knowledge on the shape of the spectrum, is this alternative method convenient?
-In chapter 4 we will see that this method is the best to characterize brown noise spectrum even with dataset of $1000$ points only, but such an investigation would require a long analysis and would deviate too much from our main interests.
-Since FPE has minimum requirement of external input we will prefer this standard approach, even if this alternative approach seems to have its merits. 
-\subsection{Alternative VS Standard CAT}
-We now compare the performance of this alternative approach with the standard approach with CAT optimizer. 
-\begin{figure}
-    \centering
-    \includegraphics[scale = .45]{Images/AlternativeCAT/AlternativeVSstandard.pdf}
-    \caption{Spectral reconstruction obtained with standard CAT approach and alternative approach}
-    \label{fig:alternativevsstandard}
-\end{figure}
-Figure \ref{fig:alternativevsstandard} show that the alternative approach provides a reconstruction that presents a largely decreased quantity of noise.
-This decreased level of noise has as counterpart a decreased accuracy in the reconstruction of the width of both the peaks.
-But, even with this loss, the estimate is still very representative of the obtained spectrum, and might be a good price to pay to reduce the variance of the estimate.
-\begin{figure}
-    \centering
-    \includegraphics[scale = .45]{Images/AlternativeCAT/CATvsALTerrors.pdf}
-    \caption{Percentage error of spectral estimate with CAT (red) and alternative CAT (green) }
-    \label{fig:CATALTerrors}
-\end{figure}
-Figure \ref{fig:CATALTerrors} shows that loss in the definition due to the averaging over the subset is concentrated in a very small number of points of the estimate, and  negligible with respect to the gain obtained in the whole spectrum.
-\\
-In this case, the reduction of noise is important because we selected the spectrum associated with the median overall error, highly influenced from noise.
-If CAT is 'lucky enough' to select filters whose length is comparable with FPE, probably such an achievement wouldn't be reached, but we expect the alternative estimate won't show any big loss.
-So, since this alternative approach gives better result for long filters and comparable results for shorter filter, being also more stable, is probably to be preferred to the standard approach with CAT as optimizer.
-Clearly, such an approach has a great disadvantage: we lose all the information regarding the prediction error filter.
-So, if we want to characterize the process, FPE is with no doubt the best choice.
-While, if we are only interested in reconstructing the power spectral density, alternative CAT's approach might be worth an attempt.
-\section{Comparison of MESA with Welch Method}
-We now compare the performance of MESA algorithm with estimate obtained with Welch method \cite{Welch1967}. Welch introduced his method in 1967, and it is based of Fast Fourier Transform. 
-Estimate is obtained averaging over modified peridiograms computed on subset of the original dataset.
-Here, as in the alternative approach to CAT, one has to choose an opportune length for the arrays and their overlapping.
-Else, a window is required to emulate period of the Fourier Transform, and with this choice one is obliged to assume data to be zero outside this window.
-The shape of the reconstruction is strongly dependent from window's choice, so, in general, Welch method is not a stable estimate of the spectra, since it strongly depends on this arbitrary choice.
-We compare the reconstruction obtained with the two methods in figure \ref{fig:WelchVSfpe}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = .3]{Images/WelchVSfpe.pdf}
-    \caption{Comparison of the reconstructed spectrum obtained with FPE (red line) and Welch's method (green line)}
-    \label{fig:WelchVSfpe}
-\end{figure}
-
-Graphically both methods are found to perform similarly.
-Since MESA is analytical, graph obtained with this method is smoother than one obtained with Welch.
-In fact, Welch spectrum is obtained performing a FFT and so its output length is related to the length of the subsets.
-Computing percentage errors, we obtain
-\begin{equation}
-    \nonumber
-    \bar r_{MESA} = 0.11; \qquad \bar r_{Welch} = 0.26
-\end{equation}
-so that better overall accuracy is reached with MESA. A plot of both the errors of FPE and Welch method is reported in figure \ref{fig:WelchError}
-
-\begin{figure}
-    \centering
-    \includegraphics[scale = .3]{Images/WelchErrors.pdf}
-    \caption{Comparison of percentage error of reconstruction with MESA and Welch method}
-    \label{fig:WelchError}
-\end{figure}It is evident that greater error for Welch is due to two main reasons.
-First reason is the reconstruction of the second peak.
-While for the first peak accuracy of the two methods is comparable, regarding the second peak Welch method shows a huge disagreement on the peak, with a value of percentage error near 4000\%.
-Clearly, being one in more than 1000 points, its own contribute to overall error is not sufficient to justify the different accuracy of the two methods.
-To explain this we have to look at noise, in fact, it is evident that in the other parts of the spectrum Welch method shows an error that is almost everywhere bigger than the one obtained with MESA.
-Taking care of noise justify the big difference of the accuracy between the two methods.
-We can conclude that, even if Welch method is a fine reconstruction for the spectrum, MESA has to be preferred for several reasons: firstly, accuracy is mostly everywhere better than the one obtained with Welch method and secondly MESA is more stable, does not require to work with arbitrary choices that significantly change results and even more important MESA does not require a modification of data, being maximally non committal to not available data, while Welch require them to be zero.
-So we can conclude that MESA is a better approach for both the quality of the reconstruction and scientific reasons.
-\section{Conclusions}
-In this chapter we studied the properties for the reconstruction of power spectral density using maximum entropy principle.
-We tried to understand what is the best criterion to infer the recursion order and that provides the better estimate of the power spectral density.
-We studied all the optimizers applying them to a normal shaped spectrum Ligo's.
-FPE and OBD optimizers are found to be very similar optimizers, for both quality of the reconstruction and stability of the method itself, but OBD shows everywhere a larger error.
-Figures {\ref{fig:optcomparison}} and \ref{fig:LigoOrderError} clearly explain the reason for this lower quality: percentage error is dependent on the estimate for filter's length, and the range where filter's  chosen from OBD lays is associated with higher error with respect to those reached by FPE, where errors seems to be independent from the choice for the autoregressive order and reach their minimum values.
-So, this allowed us to exclude OBD from the race for the best optimizer.
-\\
-The last comparison was between CAT and FPE.
-If one only considers the regions of estimated lengths, CAT estimates lay in a region associated with larger errors and it is also found to be a very unstable optimizer, since it does not show any convergence property and it is more likely to choose an order comparable with maximum length available (figure \ref{fig:ordersCompairson}).
-This behaviour led us to choose FPE as the best method to infer what is the best recursion order that allow us to reproduce the spectrum.
-But an analysis of the spectrum obtained as an average over all the simulations showed important and very interesting features of CAT: it is able to reproduce the spectrum with the same accuracy everywhere, also where FPE is found to fail.
-This is probably due to the high variability of the results reproduced from CAT.
-\\
-This led us to propose a different approach to MESA when CAT is used.
-We apply it to different subsets (of fixed length and overlapping) of the original dataset, taking the average over all the reconstructions at the end.
-This approach is found to have a better reproduction than the standard approach (figure (\ref{fig:alternativevsstandard})), providing an estimate that reduces noise with a reasonable loss in terms of details.
-Unlucky this require to introduce some arbitrary parameters for the estimate, such as the length of the smaller data extracted from the original dataset and their overlapping.
-A lot of interesting questions arises in the application of this method but, since their investigation is not our purpose, further investigation is postponed.
-At the end we compared the performance of MESA with Welch method and found that MESA has to be preferred for several reasons, firstly because it result in a better quality reconstruction for the spectrum, secondly because it requires a smaller number of 'arbitrary' assumptions, making it a more stable estimator.
-Thirdly, it requires no assumption on not available data, while Welch method requires the choice of a window that forces the data outside the window to be zero.
-So, MESA is a better performing estimate with less commitment with data and has to be preferred to Welch method.
-
-\chapter{Solar Spots Analysis}
-\section{Introduction}
-Having studied in detail MESA and understood what optimizer provides a better reproduction of spectral densities, in this chapter we provide a first and simple application of the method on the number of Sun spots as a function of time.
-MESA is a natural resolution to the
-problems raised and studied by Yule and Walker, since the autoregressive
-component naturally arises from the data and does not require any major work.   \\
-Such a study is interesting for two reasons: it is an application on a simple dataset where most of the variables can be easily taken under
-control. Secondly, the study of Sun spots allow us to apply MESA for both
-spectral analysis and time series forecasting. 
-At first, we estimate the PSD of the time series on the whole dataset.
-Secondly, we will estimate the filter on a reduced dataset and use the result to forecast the time series.
-We will then compare the estimate with the unused, observed, data.
-\section{Analysis of Sun Spots}
-First attempts to give a description of the periodical behaviour of solar
-activity is due to Yule and Walker \cite{doi:10.1098/rsta.1927.0007}\cite{1931RSPSA.131..518W}, who first proposed that the time series of solar spots does not have to be regarded as some
-perfect sinusoidal signal with a mere superposition of fluctuations, but is
-better described if a deterministic autoregressive component is considered
-in the description. A description with some deterministic dependence
-on the previous values of the series, allowed to reconstruct with higher precision the
-time dependence of the amplitude of the process. 
-In their description the choice for autoregressive order
-has to be made with different trials. It seems to be clear that, at least
-from this point of view, after a method for selecting the autoregressive 
-order has been chosen, MESA provides a natural solution to this study. 
-\begin{figure}
-    \centering
-    \includegraphics[scale = .45]{Images/SunSpots/SunSpotsNumbers.pdf}
-    \caption{Monthly number of solar spots.}
-    \label{fig:sunSpotsPlot}
-\end{figure}
-Fig \ref{fig:sunSpotsPlot} shows the monthly time series of Zurich Sunspots Number in a time period from $1749$ until $1983$. The periodicity in the time series is eye catching. \\
-We are going to estimate the time series spectrum as a means of capturing periodicity in our dataset, MESA is applied with the FPE chosen as optimizer.  
-\begin{figure}
-    \centering
-    \includegraphics[scale = .45]{Images/SunSpots/SunSpotsPSD&ACR.pdf}
-    \caption{Top panel: power spectral density for Zurich Monthly Number of Sun spots. Bottom panel: reconstructed autocorrelation function.}
-    \label{fig:sunSpotsPSD&ACR}
-\end{figure}
-Figure \ref{fig:sunSpotsPSD&ACR} (bottom panel) shows the the autocorrelation function.
-It is a slowly decaying function of time whose peaks are located at multiples of 121 months (around 10 years), reproducing the periodicity of the sample.
-A slowly decaying function is a sign of the process to be long range.
-Figure \ref{fig:sunSpotsPSD&ACR} (top panel) shows the frequency estimate of the Power Spectral Density obtained with MESA.
-The 0 frequency component is different from zero, indicating that the mean of the process is different from zero.
-There is evidence of some well defined periodicity at some frequency $f_0$, superimposed to some $f^{-\alpha}$ noise, both $f_0$ and $\alpha$ to be determined.
-MESA is also able to capture secondary peaks, at frequencies that are multiple of the fundamental frequency.
-\\
-To study it, we consider the power spectral density as a superimposition of colored noise and a Cauchy resonance.
-A fit of PSD background, see the top panel in Figure \ref{fig:SunSpotsFit}, shows that the decreasing component of the power spectral density can be actually treated as colored noise of the form:
-\begin{equation}
-    f_{noise} = (1.1 \pm .2) f^{-1.74 \pm 0.03}
-\end{equation}
-\begin{figure}
-    \centering
-    \includegraphics[scale = .45]{Images/SunSpots/SunSpotsFit.pdf}
-    \caption{Power spectral density against fitting of background with colored noise (top) and with a Cauchy distribution (bottom).}
-    \label{fig:SunSpotsFit}
-\end{figure}
-
-In this way, the non periodic component can be eliminated from the spectrum and the peak can be fit.
-Even if not perfectly symmetrical, we chose a Cauchy distribution to fit the peak (Figure \ref{fig:SunSpotsFit}, below), parameterized as
-\begin{equation}
-    C(f) = \frac{\hat{h} }{1 +\left( \frac{f - \hat f_0}{\hat \gamma}\right)^2}.
-\end{equation}
-The fit provided for the parameters the following values
-\begin{equation}
-    \nonumber 
-    \hat h = (3.35 \pm 0.02)\cdot 10^5; \quad 
-    \hat f_0 = (8.104 \pm 0.003) \cdot 10^{-3} \mbox{months}^{-1}; \quad 
-    \hat \gamma = (5.49 \pm 0.05) \cdot 10^{-4} \mbox{months}^{-1}
-\end{equation}
-and the periodicity in Sun spots is found to be 
-\begin{equation*}
-    10.28 \pm 0.05 y_{ears}.
-\end{equation*}
-It is interesting to understand the meaning of the parameters by taking an inverse Fourier Transform, the autocorrelation correspondent to a pure Cauchy shaped PSD is:
-\begin{equation}
-    K \gamma e^{-\gamma \vert t \vert } e^{-\imath f_0 t}.
-\end{equation}
-$f_0$ is the natural frequency of the oscillating component for the power spectrum, while $\gamma$ is the inverse of a lifetime $\gamma = \frac{1}{\tau}$ and can be taken as a typical time scale for the autocorrelation of the periodic component, but this is not true for the whole dataset since its power spectral density is, at first approximation, the sum of the two components.
-Since they do not provide additional information, secondary peaks (Figure \ref{fig:sunSpotsPSD&ACR}) have been neglected.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale = .47]{Images/SunSpots/SunSpotsTotalFit.pdf}
-    \caption{Fits to the spectrum considered with two components, one $f$ noise component and one Cauchy component for the resonance.}
-    \label{fig:PSDendFit}
-\end{figure}
-Figure \ref{fig:PSDendFit} shows the fit obtained adding the $f$ noise component with a Cauchy resonance.
-The result is accurate enough to describe the main component of the power spectral density: having neglected secondary peaks does not affect the computation of the quantity of interest.
-
-\section{Time Series Forcasting}
-With MESA we have been able to compute a smooth and high quality power spectral density for the time series of Sun spots, estimating the period of solar activity to be $  10.28 \pm 0.05 y_{ears}$.
-This is just a first result of the application of MESA.
-On the other hand, the computation of the prediction error filter allows us to predict the future values for our observations as well as their credibility regions.
-Since we want to compare the results of the forecast with real data, we apply MESA again on our dataset, excluding the 200 last data points.
-We will then compare the excised data with the predictions of MESA.
-Having computed the prediction error filter from the data, eliminating the first value (which is one) and changing the sign to the remaining, we have an estimate of the autoregressive coefficients.
-In this way, The process is estimated to be an $AR(96)$ process.
-\begin{figure}
-    \centering
-    \includegraphics[scale = .47]{Images/SunSpots/SunSpotsARcoefficients.pdf}
-    \caption{Estimate of the autoregressive coefficients for Sun Spots time series.}
-    \label{fig:SunSpotsFilter}
-\end{figure}
-Figure \ref{fig:SunSpotsFilter} shows estimate for the autoregressive coefficient of the time series. 
-The behaviour of the coefficient as a function of the length is characteristic, the dependence is different from zero and positive for only few values, and becomes a noisy fluctuation around $0$ of both positive and negative components after few observations.
-The standard deviation $\sqrt{P_N}$ of the superimposed white noise is estimated to be $\sqrt{P_N} = 14.77$.
-Having computed the filter and variance for the $AR$ predictor, new values can be inferred by
-\begin{equation}
-    x_t = \sum_{i = 1}^{96}a_i x_{t-i} + \nu_t \text{ with } \nu_t \sim \mathbf{N}(0, \sqrt{P_N}).
-\end{equation}
-In every step of the prediction, the deterministic component has to be summed to the white noise component.
-Since the process is linearly dependent on its previous values, this random component will propagate in the time series.
-For this reason, in forecasting it makes no sense to consider a realization only: even if a single prediction is an admissible realization for the process, taken alone it has no statistical meaning.
-\\
-For this reason, to have a fair comparison between simulated and observed realization, we will simulate 3000 series of 200 points each. \\ 
-Since random error propagates in the time series, we expect our prediction to be a consistent estimate of the observed time series for a number of points shorter than filter's length.
-In this interval, in fact, both simulation and real data are still dependent on the observed data.
-For predictions of length comparable with AR order or longer, our prediction is based on simulated data only and can be largely different from the observed series.
-\begin{figure}
-    \centering
-    \includegraphics[scale = .47]{Images/SunSpots/SunSpotsForecast.pdf}
-    \caption{Forecasting on Sun spots time series using MESA linear predictor. Observed time series is compared with ensemble mean and 95\% Cr.}
-    \label{fig:SunSpotsForecast}
-\end{figure}
-Figure \ref{fig:SunSpotsForecast} shows the numerical prediction of MESA (red dashed line), computed as a mean over the ensemble, and the future values assumed by the series (black solid line) for 200 months, credibility regions at 95\% are also reported.
-\\
-The vertical line at an abscissa of $96$ provides a visual separation between a time interval shorter and longer than the autoregressive order for the process.
-Is evident that this line separates two different regimes for the estimate.
-At earlier times, our inference is almost everywhere consistent with the observed data, and they are everywhere comprised in the $95\%$ regions.
-\\
-At later times, the quality of inference is degraded, because the inference in this region is no more dependent on observed data.
-Even if the credible regions become larger there are some values laying outside them.
-Secondly, the observed data are no more consistent with the estimate obtained from the mean.
-\\
-The only information we can achieve in this area is about the increasing / decreasing trend of the series.
-This happens because this trend is determined by the deterministic component of the process.
-\\
-So, we conclude that a proper inference with MESA can be obtained for a time period shorter than the length of the filter, in that case a period of $70\sim 80 months$, i.e.
-around $6$ years, which is a remarkable result.
-\\
-It is important to stress a fact: to obtain all the results, we assumed no knowledge on the process itself.
-To study power spectral density, gain quantitative information and make high quality forecast, we did not assume any physical model.
-All we obtained was directly contained in the data themselves and MESA extracted as much information as possible.
-This shows how the PSD - Autorgressive duality is powerful.
-With the support of Wold's theorem, given a stationary process, MESA is a tool that allow to find an approximate description of its ARMA(p,q) dual without making any assumption on the process.
-
-\section{Conclusions}
-In this section we applied MESA to a first simple dataset: the number of Sun spots.
-Even if simple, this dataset makes clear some important features of the method.
-\\ Firstly, we have been able to estimate accurately the power spectral density of the process.
-We have seen that, at least at the leading order, it can be treated as the sum of a colored noise component and a periodic component, described by a Cauchy resonance.
-The f noise component has an exponential decay with exponent $\alpha = -1.74 \pm 0.03 $ .
-The periodic component is associated to a frequency
-\begin{equation}
-    \nonumber 
-    \hat f_0 = \left(8.104 \pm 0.003 \right) \cdot 10^{-3}. months^{-1}
-\end{equation}
-Secondly, we worked out with the autoregressive duality computing its AR coefficient from the prediction error filter.
-The AR process is found to be well approximated by an $AR(96)$ process.
-Having computed the coefficients, we showed how MESA can be used to  forecast over a time period shorter than the AR order $p$.
-This is because in that range, our estimate depends on both observed and simulated data.
-With no knowledge at all on the physics behind the process, we have been able to perform accurate inference and predictions with the only application of MESA.
-This is a very remarkable result and shows how MAXENT provide a powerful tool to make inference without superimposing any model to the dataset.
-
-\newpage 
-\chapter{Sinusoidal Signals}
-\section{Introduction}
-The estimation of parameters describing theoretical models is one of the key point in statistical Analysis.
-Considering some dataset and model, dependent on a set of parameters $\vec \theta$, the objective is to find the values for $\vec \theta$ that better describe the observation.
-\\
-Such an estimate can only be performed if noise is correctly characterized, otherwise the inference might be imprecise.
-Since data come out as a superposition of signal and noise, the characterization of noise is often made 'off-source', i.e.
-studied in a time window of proper length that precedes the stretch of data containing the signal.
-Such an analysis assumes the noise to be stationary for a period of time and is seldom justified.
-In this chapter, we explore a novel idea that allows to perform an "on-source" characterization of noise by the use of MESA.
-After a short explanation of the basic idea, we will apply this new method to a toy problem of sinusoidal signal superimposed with colored noise.
-\section{Bayesian inference}
-\subsection{Bayes Theorem}
-Bayes theorem is the fundamental tool for Bayesian inference. \\ 
-Let $\vec \theta \in \mathbb{R}^N$ be the set of parameters we want to infer.
-The objective in Bayesian Analysis is the computation of the posterior distribution for the parameters given some data $D$, prior information $I$ and model $M$.
-\\
-Starting from Bayes theorem, the goal is to determine the distribution $P(\vec \theta \vert D M I)$ such that 
-\begin{equation}\label{eq:Bayes}
-    P(\vec \theta \vert D M I) = \frac{p(D\vert \vec \theta M I)}{p(D \vert M I )} p(\vec \theta \vert M I) 
-\end{equation}
-holds. \\ 
-The posterior on the left hand side indicates how the parameters are distributed according to the model, prior information and the observation of some data $D$. \\ 
-On the right hand side $p(\vec \theta \vert M I)$ is the prior distribution for the parameters given only model and prior information, while $p(D \vert \vec \theta M I)$ is the Likelihood.
-The quantity in the denominator $p(D\vert M I)$ is a purely Bayesian quantity called evidence.
-For sake of simplicity, since we are not going to use it explicitly, being computed as
-\begin{equation} 
-    p(D \vert M I) = \int d\vec \theta p(D \vert \vec \theta M I) p(\vec \theta \vert M I),
-\end{equation}
-we can consider it as a normalization factor. \\ It is clear that, to perform Bayesian inference,
-we need to correctly determine the previous quantities. \\ 
-Concerning the prior distribution, MAXENT provides a way to assign them, but there are several more methods and a huge literature about it \cite{jaynes2003ptl}. \\ 
-\subsubsection{Assigning Likelihood}
-Our main interest is the determination of the likelihood from the information and a Model $M$. \\
-The Model $M$ defines functional relations between physical quantities, dependent on the set of parameters $\vec \theta$.
-Given some value $x$ for a variable, the model $M$ allows us to determine the value of a causally related variable $z$, via
-\begin{equation}
-    \nonumber 
-    \vec z = M(\vec x \vert \vec \theta).
-\end{equation}
-The previous relations are our physical laws that we want to test and the determination of the parameters is a fundamental task to assert the validity of some Model or theory.
-In general, we do not observes pairs of values $(z_i, x_i)$, but we have to deal with some error that can mask the deterministic relation between the two.
-Out data will be "couples" of points of the form
-\begin{equation}
-    D = (y_i, x_i),
-\end{equation}
-where $y_i$ is determined from the model $M$ and some error $e_i$, via 
-\begin{equation}
-    y_i = z_i + e_i .
-\end{equation}
-In general, each of the previous quantities, the observation $y_i$, the prediction from the model $z_i$ and the error $y_i$ will have their own probability distributions.
-The Likelihood can be obtained marginalizing the joint probability distribution over $z_i$ and $e_i$, assuming them to be independent:
-\begin{align}\nonumber 
-    p_y(y_i \vert M \vec \theta I) &= \int de_i dz_i p(y_i, z_i, e_i \vert M \vec \theta I) = \int de_i dz_i p_y(y_i \vert z_i e_i M \vec \theta I) p_z(z_i \vert M \vec \theta I) p_e(e_i \vert M \vec \theta I)
-\end{align}
-Since $y_i$ is uniquely determined from the model prediction $z_i$ and the error $e_i$, its conditional distribution is 
-\begin{equation}\nonumber 
-    p(y_i \vert z_i e_i M \vec \theta I) = \delta(y_i - z_i - e_i),
-\end{equation}
-so that 
-\begin{align}\nonumber 
-    p_y(y_i \vert M \vec \theta I) &= \int de_i dz_i \delta(y_i - z_i - e_i) p_z(z_i \vert M \vec \theta I) p_e(e_i \vert M \vec \theta I )\\ \nonumber 
-    &= \int dz_i p_z(z_i \vert M \vec \theta I) p_e((y_i - z_i) \vert M \vec \theta I) 
-\end{align}
-that is the convolution between the distribution of the model and the distribution for the error.
-For a deterministic model, such that $z_i = m(x_i \vec \theta)$ is uniquely determined, the previous can be furhter simplified
-\begin{align}
-    p(y_i \vert M \vec \theta I) &= \int dz_i \delta\left(z_i - m(x_i \vert \vec \theta)\right) p_e\left((y_i - z_i) \vert M \vec \theta I\right) \\ \nonumber 
-    & = p_e\left((y_i - m(x_i\vert \vec \theta)) \vert M \vec \theta I\right)
-\end{align}
-and for such a model the likelihood is uniquely determined from the distribution of the errors.
-For independent data $y_i$ and errors $e_i$, given a set of observations, the Likelihood can be rewritten as
-\begin{equation}
-    p(\vec y \vert M \vec \theta I) = \prod_{i}p_e\left(y_i - m(x_i \vert \vec \theta) \vert M \vec \theta I \right).
-\end{equation}
-Given some deterministic model $M$, the Likelihood is uniquely determined by the distribution of the errors,
-so that a full characterization of the error is a necessary step to compute the posterior for the distribution. \\ 
-\newpage
-\section{Power Spectral density Computation}
-From the previous discussion, the role of power spectral density in parameter estimation appears to be clear: Parameter estimation can only be done if an estimate for the variance of the noise is given.
-\\
-We briefly review how this is done for signals and propose a new 'on-source' method based on MAXENT. 
-\subsection{Pure noise signals} 
-Computation of power spectral density for pure noise signals of the form
-\begin{equation}
-    y(t) = n(t)
-\end{equation}
-is the easiest task, since our observation is coincident with noise only and we don't have to extract it from the data.
-Also if simple, this is the very start to study even more complex data that are a superposition of signal and noise.
-The most traditional approach is based on Bayes theorem.
-We can consider noise to be normally distributed with 0 mean and with some variance to be determined. To infer on the posterior distribution, a prior distribution for $\sigma$ have to be assigned
-\begin{equation}
-    \nonumber
-    p(\sigma \vert D I) = \frac{p(D \vert \sigma D I)}{p(D \vert I)}p(\sigma \vert I).
-\end{equation}
-Firstly, we need to rewrite the previous in Frequency domain with discrete time sampling.
-Under these conditions, the so-called Whittle Likelihood \cite{Rover} provides a discretized version of a normal Likelihood in frequency Domain:
-\begin{equation}
-    \label{eq:Whittle}
-    -\frac{1}{2}\log L(\tilde n(f_i) \vert \vec S) = - \sum_{j = - N /2}^{N / 2}\left(\log(S(f_j) +  \frac{\frac{\Delta_t}{N}\vert \tilde \tilde n \vert ^2}{S(f_j)} \right), 
-\end{equation}
-being $\tilde n$ the Fourier Transform of noise.
-In this domain, a prior on Power Spectral Density has to be assigned also.
-In this way, the posterior on the power spectral density can be uniquely determined as
-\begin{equation}
-    p(S(f) \vert D I ) = p(\tilde n \vert S I) p(S(f) \vert I) 
-\end{equation}
-This approach allows us to determine a posterior for the power spectral density and make an estimate with error bars. 
-This easy example can become very complex in real cases, since the number of parameters can be arbitrarily large and a priori assignment for each of them is needed. \\
-But, neglecting technicalities, such an approach is still possible. \\ 
-It is interesting to see how MESA enters in this case. Having a realization of pure noise, the determination of power spectral density simply reduces to the application of MESA to the time series.
-The disadvantage is that, with respect to Bayesian inference, we will only have an estimate for the power spectral density and no other quantities.
-The advantage is that in complicated cases, this estimate is model independent, and no assumption on noise behaviour are needed.
-Such a simple examples makes clear what are the main differences between MESA and model based approaches, both having their pro and cons depending on the case of study.
-\subsection{Off Source Estimate}
-Even if pure noise signals are not in general our main goal, they provide a useful way to perform parameter estimation in case of data that are a superposition of both signal and noise:
-\begin{equation}
-\nonumber 
-d(t) = z(t) + e(t).  
-\end{equation}
-In cases like this, the main objective is to compute the Posterior for the parameters $\vec \theta$ so that an estimate on their value $\vec \theta_0$ based on the observed Data can be performed. \\ 
-Equation (\ref{eq:Bayes}) only formally solve this problem: to apply it, assigning likelihood is a necessary step and so posterior can be computed only after an estimate for power spectral density and noise characterization is needed.
-\\
-One of the most common techniques for solving this problem is the "off source noise characterization" \cite{offsource}.
-With off source studies, noise characterization is not made in the same time window where signal also lives. If signal duration is $T$, noise is studied in the $T$ seconds before signal detection. \\ 
-In this way, our problem can be divided in two subsequent steps, the first one being the characterization of a pure noise signal and the second one being the estimate of parameters via Bayes theorem, that can be made with proper algorithms like MCMC.
-\\
-Off source characterization is an efficient and fast way to characterize noise and is very proficient when noise is stationary. 
-But in the case of non stationary noise such an approach shouldn't be applied since it could bring to misleading and wrong results.
-
-\subsection{On-source Estimate}
-Our objective is to propose an 'On-source' computation to characterize noise in the same time window where signal lives, to perform a better estimate of parameters in case of non stationary noise.
-To avoid making assumptions on power spectral density shape, we will use MESA for its estimate and computation.
-This will allow us to reduce at a minimum the number of assumptions needed to perform the estimate.
-\\
-We will explain our idea referring to Markov Chain Monte Carlo (MCMC) algorithm. 
-\subsubsection{The Idea in a nutshell}
-The conceptual simplicity of this class of algorithms is ideal to explain the idea of on source estimate without introducing blinding technicalities.  \\ 
-Given some observations $D(t)$ described with some model $M(\vec \theta)$, MCMC algorithm allow to sample parameters space to find the maximum of posterior distribution for the parameters $\vec \theta$.
-\\
-Let N be the number of recursion for the algorithm, and suppose Likelihood and prior are assigned. The structure of MCMC for posterior sampling is the following:
-
-\begin{enumerate}
-    \item Set $N = N_0$; 
-    \item Set $i = 0$; 
-    \item Randomly samples $\vec \theta_i$ values from prior; 
-    \item Compute value for Posterior on samples $\vec \theta_i$
-    \begin{equation*}
-        P_i = P(\vec \theta = \vec \theta_i \vert D M) = p(D \vert \vec \theta_i M) * p(\vec \theta_i \vert M) 
-    \end{equation*}
-    \item Update parameters $\theta_i$
-    \begin{equation}\nonumber 
-        \vec \theta_{k} = \vec \theta_i + \vec \epsilon_i \qquad \text{with: } \vec \epsilon_i \sim D(0, \vec \sigma) 
-    \end{equation}
-    \item Compute $P_k$
-    \item If: 
-        \begin{enumerate}
-            \item $P_k \geq P_i$: Set $\vec \theta_{i+1} = \vec \theta_{k}$
-            \item $P_k < P_i$, Set: \\ 
-            $\vec \theta_{i + 1}= \theta_k$ with probability $\frac{P_k}{P_i}$ or \\ $\vec \theta_{i+1} = \theta_i$ with probability $1 - \frac{P_k}{P_i}$
-        \end{enumerate}
-    \item if $i = N_0$: end
-    \item increment $i \rightarrow i + 1$ 
-    \item back to step 5. 
-\end{enumerate}
-With such an algorithm one can produce an arbitrary number of samples representative of the peak for the posterior distribution. \\ 
-If the power spectral density is given, it has to be used to compute the posterior value. What happens if we don't know it? \\ 
-The simplest approach is considering it as a parameter and infer it simply sampling a larger posterior distribution of the kind: 
-\begin{equation}\nonumber 
-    P(S_1, \dots, S_{N}, \theta_1, \dots, \theta_N \vert D M I). 
-\end{equation}
-This would require the introduction of a prior distribution for every component of the PSD.
-Conceptually it is a practicable approach, but for large data set of $N$ points one should infer $N / 2$ values for the power spectral density, and complexity would increase accordingly.
-\\
-Our objective is to propose a different approach based on MAXENT. At first, we note that at step $(6)$ the computation of the Posterior implies that we are computing the quantity: 
-\begin{equation}\nonumber 
-    d(t) - z_i(t),  
-\end{equation}
-with $z_i$ the model computed on the current values $\vec \theta_i$ for the parameters. 
-Collaterally, an estimate of the noise $n_i(t) = d(t) - z_i(t)$ is provided as a function of the current value for the parameters. \\ 
-Having a sample of possible realization for noise, the basic idea is to compute $z_i$ power spectral density with MESA and use this estimate in the evaluation of the posterior.  \\ 
-The modification needed in the algorithm to insert the computation of PSD is very simple, and can be resumed as follows: 
-\begin{enumerate}
-    \item Set $N = N_0$; 
-    \item Set $i = 0$; 
-    \item Randomly samples $\vec \theta_i$ values from prior; 
-    \item Compute noise as $n_i = d - z_i$
-    \item Compute power spectral density $S_i$ applying MESA on $n_i$
-    \item Compute value for Posterior on samples $\vec \theta_i$ and $S_i$
-    \begin{equation*}
-        P_i = P(\vec \theta = \vec \theta_i \vert D S_i M) = p(D \vert \vec \theta_i S_i M) * p(\vec \theta_i \vert M) 
-    \end{equation*}
-    \item Update parameters $\theta_i$
-    \begin{equation}\nonumber 
-        \vec \theta_{k} = \vec \theta_i + \vec \epsilon_i \qquad \text{with: } \vec \epsilon_i \sim D(0, \vec \sigma) 
-    \end{equation}
-    \item Compute noise as $n_k = d - z_k$
-    \item Compute power spectral density $S_k$ applying MESA on $n_k$
-    \item Compute $P_k$
-    \item If: 
-        \begin{enumerate}
-            \item $P_k \geq P_i$: Set $\vec \theta_{i+1} = \vec \theta_{k}$
-            \item $P_k < P_i$, Set: \\ 
-            $\vec \theta_{i + 1}= \theta_k$ with probability $\frac{P_k}{P_i}$ or \\ $\vec \theta_{i+1} = \theta_i$ with probability $1 - \frac{P_k}{P_i}$
-        \end{enumerate}
-    \item if $i = N_0$: end
-    \item increment $i \rightarrow i + 1$ 
-    \item back to step 5. 
-\end{enumerate}
-This method has many great advantages. \\
-Firstly, it doesn't require any assumption on prior for power spectral density and so minimizes the number of assumptions needed. \\ 
-Secondly, we end up having a large number of samples for power spectral density also, and its posterior can be reconstructed. \\ 
-The most important implication is that we can marginalize the posterior for the parameters with respect to the power spectral density, obtaining a conservative - model independent estimate of their distribution.
-\\
-One objection could be raised at this point. 
-We saw that MESA spectrum is analytical when stationary noise is considered, since it has no poles on the unitary circles where PSD is actually evaluated.
-One could state that we are trying to abandon the too restrictive hypothesis for noise to be stationary, studying it with some method that requires is stationarity.
-\\
-To answer this objection some specifications are needed, in fact, this statement is correct.
-The assumption of stationarity is a simplifying assumption that allow us to study noise with very powerful tools.
-Our objective is not to completely relax this assumption, but to make it less restrictive.
-With off-source analysis, we are assuming noise to be "approximately stationary"  for a time period that is (at least) twice the duration of the signal.
-\\
-On source analysis requires signal to be approximately stationary in the same time window where signal lives, that is a less stringent hypothesis.
-This also clarify the domain of application of this on-source method.
-If the typical time scale of significant changes of noise is larger than the duration of the signal, on-source parameters estimate with MESA might provide an efficient solution to the problem.
-If typical time scale are shorter or comparable with signal duration, this method cannot ensure to provide a correct estimate.
-
-\section{Sinusoidal Signal} 
-As proof-of-principle we perform on-source study with the above method on two different time series: a sinusoidal signal with the superimposition of brown noise and a sinusoidal signal with a drift in the frequency, superimposed to noise generated from Ligo Power Spectral Density.
-For both its stability and fast computation, FPE is the ideal optimizer to perform this kind of analysis.
-Off-source estimate of the data will also be provided using Welch Method and Alternative CAT.
-Since noise is stationary, we expect the on-source method to provide an equivalent estimate on parameters.
-\subsection{Brown Noise}
-As first step, we generate the sinusoidal signal 
-\begin{equation}
-    a_1 \sin(2\pi f t) + a_2 \cos(2\pi f t)
-\end{equation}
-with values for amplitude and frequency parameters chosen to be
-\begin{equation}\nonumber 
-    a_1 = 5, a_2 = 7, f = 7.3. 
-\end{equation}
-The signal is superimposed with Brownian noise, whose PSD is 
-\begin{equation}
-    S(f) = \frac{3}{f^2 + 0.01}
-\end{equation}
-whit 0.01 inserted to avoid the divergence at f= 0.
-Figure (\ref{fig:BrownData}) represents the dataset we will analyze.
-First half of the data are a pure noise signal, while the second half a the superimposition of noise and sinusoidal signal with a signal to noise ratio (SNR) $\sim 37\%$
-
-\begin{figure}
-    \centering
-    \includegraphics[scale = 0.40]{Images/BrownData/BrownData.pdf}
-    \caption{Dataset under analysis.
-First half of the data are pure noise signal d(t) = n(t), needed to perform off-source characterization of noise, while the second half is the superposition of noise and sinusoidal signal d(t) = n(t) + z(t), that will be used to perform parameter estimation with both off-source/on-source methods}
-    \label{fig:BrownData}
-\end{figure}
-This dataset is composed of 2000 points.
-The whole allow us to perform off-source analysis computing the power spectral density off noise in the first 1000 and parameter estimation with the second 1000.
-This second half will be the only one needed to perform on-source  analysis.
-The results of parameters estimation and PSD reconstruction will be then compared.
-\subsubsection{Off-source analysis}
-As first step to perform off-source analysis, the reconstruction of the power spectral density is needed. Such an estimate is obtained with both Welch Method and Alternative CAT method. 
-\begin{figure}
-    \centering
-    \includegraphics[scale = .40]{Images/BrownData/OffSource_PSDcomparison.pdf}
-    \caption{Comparison of true power spectral density with estimate obtained with Welch method (green) and Alternative CAT (red).
-Alternative CAT estimate is made on sub arrays of 100 points with a 50\% superimposition}
-    \label{fig:brownPSD_offSource}
-\end{figure}
-Figure \ref{fig:brownPSD_offSource} compares the estimate of the power spectral density obtained with Welch Method (green) and Alternative Cat (Red).
-Both of the estimates are representative for the power spectral density, but aCAT minimizes the variance of the result.
-The parameter estimation is obtained with the MCMC sampling of a Nested Sampling
-Algorithm \cite{Skilling..735..395S} \cite{john_veitch_2020_4109277}. 
-Figures (\ref{fig:BrownCATposterior})shows the posterior for the parameters computed on alternative CAT (above) and Welch  (below) power spectral densities 
-\begin{figure}
-    \centering
-    \includegraphics[scale = .47]{Images/BrownData/BrownOffPosteriors.pdf}
-    \caption{Posterior distributions for the parameters obtained using alternative CAT (above) and Welch (below).}
-    \label{fig:BrownCATposterior}
-\end{figure}
-The estimate for the parameters with Welch's
-is 
-\begin{equation}\nonumber 
-    \hat a_{1; Welch} = 4.86 \pm 0.04 ; \quad \hat a_{2;  Welch} = 7.10 \pm 0.05 \quad \hat f_{Welch} = 7.30000 \pm 4 \cdot 10^{-5}
-\end{equation}
-The frequency estimate is unbiased and with a small error.
-The estimate for the second amplitude $a2$ lays in one standard deviation from the mean, while for the 1st amplitude parameter the real value is found to lay in 3.5 standard deviation.
-The results obtained with alternative CAT are more accurate
-\begin{equation}
-    \hat a_{1; aCAT} = 4.94 \pm 0.07 ;\quad 
-    \hat a_{2; aCAT} = 6.96 \pm 0.06;\quad 
-    \hat f_{aCAT} =  7.30000 \pm 0.00005
-\end{equation}
-and a correct estimate for the parameters, since their true values are always comprised in one standard deviation from the mean.
-In similar cases where noise is stationary, the off-source approach can give results that are consistent with the real values for a signal.
-\subsubsection{On-Source Analysis}
-On source estimate is performed directly on the dataset, so in the second half part of the data in figure \ref{fig:BrownData}, and the posterior sampling is obtained via the same Nested sampling algorithm used before.
-\begin{figure}
-    \centering
-    \includegraphics[scale = .47]{Images/BrownData/BrownOnPosteriors.pdf}
-    \caption{Posterior distribution for the parameters obtained as on-source estimate}
-    \label{fig:brownOnPosterior}
-\end{figure}
-Figure (\ref{fig:brownOnPosterior}) shows the posterior distribution for the parameters obtained with the on source estimate.
-With respect to the posteriors obtained with the off-source estimate, the domain is larger and so is the error in the estimate.
-the larger support for the posterior in the on source case is due to the fact that the power spectral density is not 'a priori' known, but evaluated step by step with the parameters.
-The uncertainty on the posterior affect the parameter estimation, resulting in larger posterior distributions.
-The estimate for the parameters obtained with on-source analysis is
-\begin{equation}\label{brownONestimate}
-    \hat a_1 = 5.0 \pm 0.2 ; 
-    \quad \hat a_2 = 7.0 \pm 0.4;
-    \quad f = 7.3000 \pm 0.0001 
-\end{equation}
-Remarkably, the estimate results to be unbiased for every parameter, concerning the standard deviation it result to be one order of magnitude larger than that obtained with the off-source estimate.\newline
-\begin{figure}
-    \centering
-    \includegraphics[scale = .43]{Images/BrownData/On_Source_PSDposterior.pdf}
-    \caption{On-source reconstructed posterior with 90\% credibility regions. FPE estimate on real error is also reported. }
-    \label{fig:brownOnSourcePSD}
-\end{figure}Figure (\ref{fig:brownOnSourcePSD}) show the on-source estimate for the power spectral density obtain with $90\%$ Credbility regions.
-FPE estimate on the real Error is also reported.
-The estimate is computed taking the median over all the reconstruction obtained in the Monte Carlo.
-Median is chosen over the mean to be more representative for its greater stability.
-\\
-The true power spectral density is not consistent with the median of the ensemble, neither comprised in the $90\%$ Credibility.
-But it also has to be noticed that the median on-source estimate is almost coincident with the FPE estimate for the power spectral density, that is computed on the real error.
-This fact allow us to state that the off-set in the estimate is due to MESA and not to on-source estimate.
-The fact that the estimate of PSD on real error is coincident with our estimate is rather a good result, and the estimate of noise stress this fact.
-\begin{figure}
-    \centering
-    \includegraphics[scale = .43]{Images/BrownData/OnSource_ReconstructedNoise.pdf}
-    \caption{'On-source' estimate of noise}
-    \label{fig:brownNoise}
-\end{figure}
-Figure (\ref{fig:brownNoise}) compares the real error (black) with its estimate (red crosses).
-This estimate is computed taking the median of the errors over all the possible models reconstructed with MCMC algorithm.
-The agreement is impressive and gives even more evidence that the disagreement between real PSD and the reconstruction is due to MESA only.
-This is easily understood comparing MESA analytical solution with the functional form we used to generate noise.
-Neglecting constants, we are requiring
-\begin{equation}\nonumber 
-    \nonumber 
-    \biggr \vert\sum_{i = 0}^p a_i e^{- \imath 2 \pi f i \Delta t}\biggr \vert ^2 \sim f^2. 
-\end{equation}
-The expansion on the left hand side could be consistent with $f$ if long orders are selected.
-In general $f^{-2}$ noise is brownian motion spectrum, that is an $AR(1)$ non-stationary process.
-Under these assumptions we are working outside the limits of MESA, and so we are not ensured that we will obtain a consistent estimate.
-In fact, a check on the autoregressive coefficient shows that the filter obtained with FPE is of order 3, too short to be an accurate approximation.
-It interesting to note that the estimate obtained with alternative CAT method in the previous section, shown in figure \ref{fig:brownPSD_offSource}, is a way more accurate estimate for the power spectral density.
-Reminding that CAT chooses long filters, it is more able to capture the power low behaviour of the power spectral density, and the variance of the estimate is reduced taking the average over different realizations.
-All these facts underline that the non consistent estimate for the power spectral density is not to be considered as a failure of our proposal for on-source estimates.
-\newpage 
-\section{Ligo Noise}
-The same analysis will be repeated on a sinusoidal signal with a linear drift $f'$ in the frequencies 
-\begin{equation}
-    z(t) = a_1 \sin\left(2 \pi (f + f' t) t\right) + a_2 \cos\left(2\pi(f + f' t)t \right), 
-\end{equation}
-whose parameters are valued 
-\begin{equation}\nonumber 
-    a_1 = 10^{-21}; \quad a_2 = 5\cdot 10^{-22}; \quad f = 128, \quad f' = 1.4 
-\end{equation}
-the sampling interval and observation time are respectively 
-\begin{equation*}
-    dt = \frac{1}{2048}; \quad T = 8,
-\end{equation*}
-both parameters and sampling time expressed in arbitrary units. 
-The signal is superimposed with some noise randomly generated from the Ligo Power Spectral Density studied in chapter 2. 
-\begin{figure}
-    \centering
-    \includegraphics[scale = .43]{Images/LigoOnSource/ObservedData.pdf}
-    \caption{Simulated dataset used to perform the analysis}
-    \label{fig:LigoData}
-\end{figure}
-Figure \ref{fig:LigoData} shows the dataset used for both 'off-source' and 'on-source' analysis.
-As a main difference from the previous case, the amplitude for noise is much greater than the amplitude of the signal.
-\\
-The first 4 seconds of observation are pure noise, while the second are a superimposition of the deterministic signal $z(t)$ and noise $n(t)$.
-Again, the first 4 seconds will be used to study the power spectral density with the off-source estimate, while the last 4 seconds to perform parameter estimation.
-For these last for second SNR is reduced to $\sim 10^{-3}$.
-\subsection{Off-Source Analysis}
-The off source analysis is performed characterizing noise with the alternative CAT approach and Welch approach.
-The reconstruction of the power spectral densities with the above methods is reported in figure \ref{fig:offPSDLigo}.
-\begin{figure}
-    \centering
-    \includegraphics[scale = .43]{Images/LigoOnSource/offPSDcomparison.pdf}
-    \caption{Comparison of the estimate of the power spectral density obtained with Alternative CAT (red line) and Welch method (green line)}
-    \label{fig:offPSDLigo}
-\end{figure}
-Both peaks are recognized but not perfectly characterized, especially the first peak's height is underestimated.
-Still, alternative CAT's estimate shows improved variance with respect to Welch's method.
-A comparison of the posterior for the parameters obtained with both methods is shown in figure (\ref{fig:posteriorLigoComparison}).
-\begin{figure}
-    \centering
-    \includegraphics[scale = .46]{Images/LigoOnSource/PosteriorsComparison.pdf}
-    \caption{Posteriors for the parameters of the model for both Alternative CAT (above) and Welch (Below)}
-    \label{fig:posteriorLigoComparison}
-\end{figure}
-Alternative CAT results are peaked around the true results, while the posteriors obtained with Welch's method are slightly biased.
-The estimate for the parameters obtained with alternative CAT are found to be
-\begin{align*}
-\centering
-    \hat a_1 = (1.005 \pm 0.005) \cdot 10^{-21}; \quad 
-    &\hat a_2 = (4.99 \pm 0.09) \cdot 10^{-22}; \\
-    \hat f = 128.000 \pm 0.002; \quad 
-    &\hat f' = 1.3998 \pm 0.0004, 
-\end{align*}
-so that the true values are always found to lay in one standard deviation from the mean. As can be seen from the posteriors, the estimate for the parameters obtained with Welch's method 
-\begin{align*}
-\centering
-    \hat a_1 = (0.999 \pm 0.003) \cdot 10^{-21}; \quad 
-    &\hat a_2 = (5.29 \pm 0.05) \cdot 10^{-22}; \\
-    \hat f = 127.998 \pm 0.001; \quad 
-    &\hat f' = 1.4004 \pm 0.0002
-\end{align*}
-gives less accurate results.
-The only parameters that lays in one standard deviation from the mean is the first amplitude parameter $a_1$.
-Both drift and frequency are correctly estimated in 2 standard deviations, that is a reasonable result, while for the second amplitude parameter (a2) the estimate is outside 3 standard deviations.
-The big variance of Welch's method in power spectral density estimate is
-naturally affecting the estimate on the parameters, but still providing reasonable results, but the averaged MESA based results of CAT are way more accurate.
-
-\subsection{On-Source Analysis}
-We now provide the result of the characterization of noise and the parameter estimate using the proposed on-source estimate.
-This case is of particular interest because signal noise ratio (SNR) is around $\sim 10^{-3}$, so that we are trying to characterize the dominant component of noise and separate it from a relatively weak signal.
-Also in the study of brownian noise, even if SNR incredibly decreased, the result in the estimate is very accurate. 
-\begin{figure}
-    \centering
-    \includegraphics[scale = .43]{Images/LigoOnSource/PArametersPosteriors.pdf}
-    \caption{Posterior distribution for the parameters obtained with the on-source estimate}
-    \label{fig:LigoOnPosteriors}
-\end{figure}
-Figure \ref{fig:LigoOnPosteriors} shows the posterior distribution computed for each parameter using the proposal method.
-Again the uncertainty on the estimate of the power spectral density propagates in the estimate for the parameters increasing the domain of the posterior and resulting in larger error bars.
-The estimate for each parameter is found to be
-\begin{align}\label{LigoEstimate} 
-    \hat a_1 = (1.00 \pm 0.01) \cdot 10^{-21}; \quad 
-    &\hat f = 128.003 \pm, 0.004 \\ \nonumber 
-    \hat a_2 = (4.7 \pm 0.2) \cdot 10^{-22} ; \quad 
-    &\hat f' = 1.3993 \pm 0.0009
-\end{align}
-all the estimate but the second amplitude are very accurate and the true value lays inside the error defined by one standard deviation.
-For the second amplitude parameter the result is comprised in two standard deviation.
-\\
-So, even for a weak signal we have been able to accurately resolve the deterministic signal described by the sinusoidal model. \\ 
-But the added value of this new approach is the characterization of noise and the computation of its power spectral density. 
-\begin{figure}
-    \centering
-    \includegraphics[scale = .43]{Images/LigoOnSource/PSDestimate.pdf}
-    \caption{Estimate for the power spectral density obtained with the on-source analysis.}
-    \label{fig:LOnSourcePSD}
-\end{figure}
-Figure \ref{fig:LOnSourcePSD} (above) shows the estimate for the power spectral density obtained with the on-source estimate, computed with FPE optimizer.
-The result of the estimate is accurate and representative of the real power spectral density.
-The credibility regions for the estimate are tight.
-This is probably due to the stability of FPE.
-We expect that for slightly small changes in the parameters the estimate for the power spectral density won't be largely affected.
-Since frequency and drift estimate are narrow, this is reflected on the posterior obtained for the power spectral density.
-This result is clearly related to the estimate and capability of the algorithm to discriminate between noise and signal.
-\begin{figure}
-    \centering
-    \includegraphics[scale = .43] {Images/LigoOnSource/ErrorEstimate.pdf}
-    \caption{Comparison of the real noise (black dashed line) with the estimate obtained with the on-source analysis (red solid line)}
-    \label{fig:LErrorEstim}
-\end{figure}
-Figure \ref{fig:LErrorEstim} shows the real error, superimposed to the data, and its estimate obtained with the on-source algorithm.
-This is computed as the ensemble median over all the reconstructed time series for the noise.
-The accuracy and precision in the reconstruction are incredibly high.
-The percentage error in the reconstruction of noise is found to be $r = 0.78 \%$, with a standard deviation $\sigma_r = 62.4\%$.
-The credibility regions are very tight, their average width is found to be $(3 \pm 1) \cdot 10^{-23}$, so that they are very narrow with respect to the values of noise time series, even though 100\% of the real values are found to lay between them.
-This explains why the accuracy in the reproduction is in one standard deviation around
-$0.2 \% \div 2 \%$ and also the very tight credibility regions in the estimate for the power spectral density. \\ 
-\begin{figure}
-    \centering
-    \includegraphics[scale = .43]{Images/LigoOnSource/NoiseEstimateBisect.pdf}
-    \caption{Noise estimate VS real noise plot. The red dashed line represents the bisector of the axis}
-    \label{fig:Bisector}
-\end{figure}
-To study the accuracy in the reproduction, the curve obtained in a "real error VS error estimate" is fitted with a straight line to see if the result is consistent with the bisector of the axis.
-Figure \ref{fig:Bisector} shows both the plot of the estimated error VS the real error and the bisector of the axis.
-The fit with a line 
-\begin{equation}\nonumber 
-    y = m x + q,
-\end{equation}
-gives as estimate for slope and intercept the following values 
-\begin{equation}
-    \hat m = 1.00000 \pm 0.00001; \quad \hat q = (0.02 \pm 1) \cdot 10^{-25},
-\end{equation}
-their functional relation is perfectly coincident with the bisector of the axis with an incredibly high precision, so that the small differences in the reproduction are probably statistically meaningless.
-With the on-source study parameters can be estimate with great accuracy, just renouncing to some precision in the result.
-Even for small values of SNR, this technique allow to distinguish between noise and signal in the data, reconstructing both in a very accurate way, just with the specification of the model for the signal.
-
-\newpage 
-\section{Conclusions}
-In this chapter we briefly discussed the parameter estimate and the role played by the computation of power spectral density for noise to achieve this goal.
-Since likelihood distribution is uniquely determined from distribution of noise, a method to characterize this distribution is necessary.
-The standard off-source analysis requires to assume that noise is stationary for a time length that is at least twice the time of the observation of the signal.
-That assumption can lead to inaccurate parameter estimate when dealing with real noise, so that less restrictive one is needed.
-\\ For this reason, we proposed a new method to perform parameter estimation, introducing MAXENT computation of power spectral density in a MCMC algorithm.
-This means that only the strain of data where the signal live is needed.
-This new approach has been applied to two different 'toy models'.
-The first consists in a sinusoid with constant frequency, superimposed to some colored noise, with a SNR around $36 \%$.
-\\
-In the second example we complicated the model by the introduction of a linear drift in the frequencies and superimposed with noise generated to be consistent with the Ligo power spectral density.
-In this second case, the signal noise ratio is reduced to $\sim 10^{-3}$.
-Comparing with the standard 'off-source' analysis, this new approach enlarges the error in the estimate of the parameters but lead to very accurate results: all the true values for the parameters have been recovered in no more than two standard deviations from the mean (eq: \ref{brownONestimate} and \ref{LigoEstimate}).
-The real improvement regards the characterization of noise.
-Just using the strain of data where the signal actually lives, this approach has the capability to discriminate with huge accuracy which is the deterministic component of the data and which is simply noise, also for very low value of SN ratio.
-The estimate of noise (Figure \ref{fig:brownNoise} and \ref{fig:LErrorEstim}) is extremely accurate and precise: even if the credibility regions are very sharp, they always comprise the real values.
-Even the reconstruction of the power spectral density is extremely accurate, and most of the differences between the real power spectral density are attributable to MESA.
-\\
-Even if the aforementioned results are based on simple toy-models, the obtained results provide enough evidence that this method could refined to try an implementation on different, real, contexts.
-
-\newpage 
-\chapter*{Conclusions and future developments}
-In this thesis we introduced the maximum entropy principle (MAXENT) with some of its most relevant applications to the study of stationary time series.
-In chapter one MAXENT has been introduced and reformulated, according to the original work of Burg, to be applied in the estimate of power spectral densities.
-This method is known as 'Maximum Entropy Spectral Analysis' (MESA).
-\\
-Burg's approach admits a closed expression for the estimator of the power spectral density  in terms of some coefficients known as 'forward prediction error coefficients' (equation \ref{eq:MESApsd}). 
-These $p$ coefficients provide a link between the spectral estimate and the characterization of the observed stochastic process.
-In fact, simply changing their sign, these new coefficient are a 'linear predictor' for the time series (equation \ref{eq:ARspectrum}).
-This means that they allow us to express the observed time series in terms of some autoregressive process of order $p$ (AR(p)).
-\\
-Even more remarkably, the coefficients obtained with MESA are found to minimize the error in the prediction (equation \ref{eq:bestLinearPredictor}), i.e.
-their computation is equivalent to a least square fitting for the time series with an $AR(p)$ model.
-It is evident that importance of MESA goes far beyond his original domain of application: with no assumption on the physical model behind the observed data,  it accurately reconstructs the spectrum and characterizes the process..\\
-Burg's algorithm is recursive and, for dataset whose autocorrelation function is not perfectly know, requires some criterion to estimate the recursive order.
-Different criteria to address this problem can be found in literature, and in chapter two we characterized three of them: Final Prediction Error (FPE), Criterion on Auoregressive Transfert function (CAT) and Optimum Bayes Decision rule (OBD).
-The study considered the accuracy in the reproduction of two different spectra: a normal distribution and the a simplified spectrum of noise for the LIGO experiment (figure \ref{fig:LigoPSDOriginal}).
-\\
-The error in the reproduction resulted to be strongly dependent on the estimate of the recursive order $p$ (Figures \ref{fig:optcomparison} and \ref{fig:LigoOrderError}), and in both cases the estimate for $p$ associated with the smallest error is obtained with FPE.
-OBD lays in a region of shorter filter and larger error .
-The reconstructions obtained with these two methods are similar, but OBD provides estimates with a larger bias.
-CAT, instead, shows a completely different behaviour.
-It selects larger orders than FPE,  but does not converge to any specific order.
-These long arrays are associated to a decreased bias in the estimate but with higher variance: it introduces a large number of spurious peaks in the reproduction.
-\\
-For its peculiar properties, CAT's estimates might be improved averaging on various estimates obtained from subsets of the original dataset.
-The result will depend on both the length of the subsets and their overlapping.
-Short arrays (and so shorter filters) provide more biased results but less variance, while the opposite holds for longer arrays.
-For 'short' arrays the estimate is similar to the results obtained with FPE, but with lower accuracy.
-For longer arrays, it provides a similar result to the standard CAT but with reduced noise, so that this alternative approach makes sense if relatively ``long" arrays are chosen.
-In chapter 4 this method is found to be way more accurate than $FPE$ (Figures (\ref{fig:brownPSD_offSource}) and (\ref{fig:brownOnSourcePSD})) in the reconstruction of power spectral density for Brown Noise.
-\\
-In every application of this alternative method, the overlapping between the segments of data was fixed to $50\%$, so that a future perspective could be the characterization of the method for different lengths and overlapping.
-\\
-This modified approach has several cons.
-MESA is a powerful approach also because it is free of prior assumptions.
-The introduction of this alternative approach require some a priori choices on length and overlapping of sub segments of data that are hard to justify, we can only appeal to some 'expected behaviour' of the estimate with no general arguments supporting this choice.
-Also, it increases the complexity of the algorithm.
-But the worst loss regards the link with autoregressive processes: taking the average, every information on the autoregressive coefficients is completely lost.
-\\
-For both its accuracy and its stability, FPE has been chosen as the best optimizer to improve the estimate on the recursive order.\\ 
-There are two other aspects that would deserve some studies. Firstly, for both spectra we investigated the estimate for fixed length of the time series.
-Since $AR(p)$ processes are, for $p$ finite, just an approximation for the more general $ARMA(p,q)$ process, we expect the estimate for $p$ to be dependent on the length $N$ of the input time series.
-It would be interesting to see how and if the length of the estimate of recursive order is dependent on the original length of the time series $N$.
-\\
-Secondly, since our interest was in characterizing noise, we studied the three optimizer focusing on power spectral density only.
-To achieve a full characterization of the criteria, one should generate an $AR(p)$ process with known coefficients and compare them with the estimates for length and values obtained with the three methods.\\
-Finally, having characterized the three optimizer, in chapter 3 we applied MESA on a very simple time series: sun spots number.
-We obtained a description of the time series in terms of a AR(96) process.
-Fitting the spectrum (figure \ref{fig:sunSpotsPSD&ACR}) with some colored noise for the background and with a Cauchy resonance for the peak, we found a periodicity in the series of $10.28 \pm 0.05 years$.
-Knowing the forward prediction error, we have also been able to accurately predict the future values of the time series for $\sim 80$ months (figure \ref{fig:SunSpotsForecast}).
-\\
-The last chapter developed a new proposal to perform 'on-source' bayesian parameter estimation, inserting MESA computation in a Markov Chain Monte Carlo algorithm (MCMC).
-With this approach, a model for the power spectral density is constructed "in real time" with the sampling for the parameters.
-In this way, with a relatively small number of assumptions, posterior samples are obtained for both the parameters of the model and the power spectral density.
-This proposal is convenient because, given some observed data $d(t) = z(t) + n(t)$, it allow us to separate the deterministic signal component $z(t)$ from the noise $n(t)$ directly 'on' the data.
-This approach allow us to relax the restrictive hypothesis on the stationarity for noise, that are usually required when noise is characterized 'off-source'.
-\\
-This approach has been tested on a sinusoidal signal superimposed to brown noise and noise generated with the same LIGO Spectrum studied in the second chapter.
-We obtained very promising results.
-Concerning the estimate of parameters, we have great accuracy in the estimate (equations \ref{brownONestimate} and \ref{LigoEstimate}), with the loss of one order of magnitude in precision.
-This lost is probably related to the uncertainty in the estimate of the power spectral density.
-The most remarkable result is about the noise reconstruction from the data: the algorithm has been able to almost perfectly distinguish and separate noise and signal from the data (figures \ref{fig:brownNoise} and \ref{fig:LErrorEstim}).
-A linear least square fit of the estimated noise as a function of the real noise is found to be perfectly coincident with the bisector of the axis, up to a precision of $10^{-5}$, so that the differences between the real values for noise and its estimate are negligible from a statistical point of view.
-\\
-From the beginning, MESA have been applied to reproduce the simplified power spectral density of LIGO noise.
-The reason behind this choice concerns our future perspectives: the very next step will be the application of the new on-source method to public data of LIGO-Virgo collaboration, that will be a test bench for our analysis: the noise in the detector is not perfectly stationary as the one we generated in our simulation.
-But, since this analysis is also performed 'off-source', considering the results obtained with our toy models, it seems very reasonable to believe that the on-source method can be applied on gravitational waves data with optimum results.
-We are confident that we can be able to characterize almost exactly the noise component in the apparatus at the time the signal was caught.
-\\
-The complete absence of assumptions on noise, makes this method a worth trying approach to several process.
-Special candidates are observations whose noise component cannot be completely understood and modeled.
-A simple example, reminiscent of our sinusoidal model, is the transit of exoplanets around a star.
-The partial eclipse due to the transit of the exoplanet in front of the star, will result in an observed luminosity that is periodic in time.
-This oscillatory, deterministic behaviour, will be superimposed to noise from various sources, such as stellar activity, the presence of the interstellar medium and many others.
-If one wants to isolate the phenomenon itself, not caring to understand what is every component of noise, MESA on-source estimate might be a good approach.
-Another, king size, example is LISA.
-For such an apparatus, it is impossible to provide a satisfactory model for noise: every signal coming from the galaxy,  such as stochastic background, binaries etc..
-is a source for noise, and it is impossible to characterize them all.
-Also, noise is highly non stationary so that off-source analysis can lead to very misleading results.
-Having modeled what the "signal" is, one might attempt to use our on-source estimate to discriminate between signal and noise, giving up on characterizing every source of noise and focusing on reconstructing the overall noise only.
-Clearly, the aforementioned examples require a lot of technical work to do, and MESA on-source analysis is just "the core" of way more complex analysis framework.
-For the exoplanet complications may arise from non-uniform sampling interval and from the fact that data can be recorded for 6 months only.
-For LISA a problem might be the understanding of what has to be considered a signal and what is noise.
-MESA on source estimate seems to be a promising approach to separate noise and signal in a variety of cases, but is just the core for most  applications: every implementation requires some ad hoc analysis and characterizations.

### 2021-02-24 scripts/optimisers_comparison.py
Added plots and script

@@ -18,14 +18,17 @@ from scipy.interpolate import interp1d
 def relative_error(real, estimate): 
     return np.abs(real - estimate) / real 
 
-PSD = 'ligo' #normal or ligo 
-save = False
+PSD = 'normal' #normal or ligo 
+save = True
 savedir = '../paper/Images/optimisers_comparison/' + PSD + '/'
 simulate_data = False 
 methods = ['FPE', 'CAT', 'OBD']
 M = MESA() 
-init_plotting() 
+init_plotting()
+plt.close() 
 colors = {'FPE': 'r', 'CAT': 'k', 'OBD': 'green'}
+
+print("Dealing with ",PSD)
 
 #Initialize single spectrum dictionaries 
 spectra, optimisers, orders, errors = {}, {}, {}, {} 
@@ -57,19 +60,32 @@ if simulate_data:
             orders[method].append(M.a_k.size)
             spectra[method].append(M.spectrum(dt)[0][: number_of_points // 2])    
 
+        #saving data
+    s, o, orde = [], [], []
+    for i, method in enumerate(methods): #initializing data
+        s.append(spectra[method]); o.append(optimisers[method]); orde.append(orders[method])
+    np.save('plot_data/normal_spectra.npy', np.array(s))
+    np.save('plot_data/normal_optimisers.npy', np.array(o))
+    np.save('plot_data/normal_orders.npy', np.array(orde))
+    
 elif not simulate_data:
     frequency = np.linspace(0, .5 / dt, number_of_points // 2 + 1)
     psd_int = interp1d(f, spectrum, fill_value='extrapolate')
     psd = psd_int(frequency)
     
     if PSD.lower() == 'ligo': 
-        s = np.load('ligo_spectra.npy')
-        o = np.load('ligo_optimisers.npy')
-        orde = np.load('ligo_orders.npy')
-        
-        for i, method in enumerate(methods):
-            spectra[method] = s[i]; optimisers[method] = o[i]; orders[method] = orde[i]
-#Set unity measures 
+        s = np.load('plot_data/ligo_spectra.npy')
+        o = np.load('plot_data/ligo_optimisers.npy')
+        orde = np.load('plot_data/ligo_orders.npy')
+    if PSD.lower() == 'normal': 
+        s = np.load('plot_data/normal_spectra.npy')
+        o = np.load('plot_data/normal_optimisers.npy')
[…]

### 2021-02-25 paper/mesa.tex
Validation part completed

@@
 The disagreement between the PSD $S_i(f)$ estimated from the i-th simulated time series and the target PSD $S(f)$ is measured with the relative error $r_i$, averaged over all frequencies:
-\begin{equation}
+\begin{equation}\label{eq:freq_erorr}
 r_i = \frac{1}{f_{Ny}}\int_0^{f_{Ny}} \d{f} \frac{|S_i(f) - S(f)|}{S(f)}
@@
 \end{figure}
-In this subsection we concentrate again on the reconstruction of a specific, known, power spectrum.
-We analyze the properties of reconstruction of this second spectrum whose shape is the Advanced LIGO design sensitivity theoretical spectral curve (Figure \ref{fig:LigoPSDOriginal}). 
-For this analysis we will study properties only for arrays of 40960 points.
-We consider a sampling rate of $2048 samples / second$ that means we can reconstruct a Nyquist frequency of $1024 Hz$.
-We chose a longer dataset than before because algorithm works with a 1 dimensional interpolation of spectrum, and with shorter datasets the interpolated spectrum is not consistent with the original spectrum and does not capture the height of the peaks.
-Shorter choices would result in noise extracted from a different power spectrum than the one in figure \ref{fig:LigoPSDOriginal}.
-\\
-We will study statistical properties from samples obtained with 500 simulations and compare the obtained results for each optimizer. 
-We start considering statistical properties of mean spectrum obtained averaging over all the simulations and conclude with an analysis for the statistics of the reproduction of single spectrum. 
-\subsubsection{Properties for the average spectrum}
-Here we report mean properties of the reproductions of the spectrum. The spectrum we want to reconstruct is reported in Figure \ref{fig:ligospectrum},
+In this subsection we concentrate again on the reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve see fig.~\ref{fig:LigoPSDOriginal}.
+For this analysis we will study properties on $500$ time series arrays of $40960$ points each. We set a sampling rate of $\SI{2048}{Hz}$ (this is $\SI{20}{s} of data$).
+We chose a longer dataset to capture the complicated structure of the PSD. A shorter time batch would not have allowed us to reconstruct faithfully all the peaks of the empirical PSD.
+
 \begin{figure}
     \centering
-
      \includegraphics[width = \linewidth]{Images/LIGOsimulate/PSD.pdf}
@@
 \end{figure}
-shows the original power spectral density and the ensemble average of every method. For the mean, the relative errors are: 
-\begin{equation}
-    \bar r_{\bar S, FPE} = 0.0426\%, \qquad \bar r_{\bar S, OBD} = 0.129\%, \qquad \bar r_{\bar S, CAT} = 0.0211\%
-\end{equation}
-Again, if one only check overall accuracy for every method, CAT is outperforming.
-In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them.
-To check this behaviour we take a look at the two peaks separately.
-For the first one a log plot of the reconstructed peaks and their relative errors are shown in Figure \ref{fig:ligo1peak}
+In figure \ref{fig:ligospectrum}, we report the original spectrum and the ensemble averaged spectrum obtained with the three methods (no confidence level).
+In figure \ref{fig:ligo1peak} and \ref{fig:ligo2peak}, we plot a details od fig.~\ref{fig:ligospectrum} around the two peaks of the spectrum
+In figure~\ref{fig:LigoOrderError}, we report, as above, a scatter plot with the frequency averaged error (eq.~\ref{eq:freq_erorr}) and the length of the autoregressive model obtained with each loss function.
+
+The conclusions are very similar to those above. When averaged on multiple realization, CAT loss function is able to capture the spectrum very precisely.
+In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them (figs.\ref{fig:ligo1peak} and \ref{fig:ligo2peak}).
+However, the computation on a single noise realization yields larger error, as manifest in fig.~\ref{fig:LigoOrderError}.
+This supports the conclusion that CAT is not a stable optimizer: different computation of the same spectrum can output very different results.
+On the other hand, FPE seems more robust and provides a smaller variance (and frequency averaged error), despite having a worse performance on the average.
+We find again that FPE and OBD have similar performance: indeed, OBD shows qualitatively the same behaviour of FPE, although with slightly worse performance in the error.
+
+We can then draw some general conclusions:
+\begin{itemize}
+	\item There are no reasons to prefer OBD over CAT or FPE.
+	\item If we have one single realization for the process, we recommend the use of FPE, that would get the best resolution possible.
+In this situation, CAT would provide spurious and unreliable results, with large error.
+	\item But, in case we have several realization for the same process, this last property of CAT, together with its instability, result to be his luck: averaging the spectrum over different time series tend to compensate noise in the results, originating a very well precise spectral estimation.
+The same is not true for FPE.
+\end{itemize}
+
+The choice of the loss function between CAT and FPE seems problem-dependent.
+
 \begin{figure}[t]
@@
 \end{figure}
-and the results reinforce our previous statement.
-The behaviour of the error is very similar for CAT and FPE, showing similar shapes but very different sizes, while CAT is giving a different and way more precise result.
-Both FPE and OBD are not able to capture the width of the peak and also show larger errors, while CAT has a good behaviour in reconstructing width's also.
-The reconstruction for very sharp peaks considering the average over all the spectra is found to be strongly inaccurate for both FPE and OBD, while a good accuracy is reached with CAT method. \\ 
-Considering the estimate for the position of the peak, whose true value is found to be $16.85 \quad a.u.
-$, within numerical accuracy both FPE and CAT find its exact position, while OBD shift the peak left at $16.7 a.u.$, with a $0.9\%$ inaccuracy.\\
-Second peak results are analogous, plotting the peak and the errors value around it
 \begin{figure}[t]
@@
 \end{figure}
-we find again that FPE and OBD are performing similarly, the magnitude of the error is maximum for OBD and minimum for CAT. \\ 
-In this case even CAT shows large errors in some points, but they are reasonable in a large part of it, still, it remains the best reconstruction.
-For this second peak, all the methods find maximum position to be at $438.05 \quad a.u.$, which is the true position of the maximum.
-Looking at Figure \ref{fig:ligo1peak} at \ref{fig:ligo2peak}, it appears to be clear that the least of the accuracy is not on the peak, but immediately around, so that one of the biggest deal is the reconstruction of peak's width, that only CAT is able to capture with high definition.
-\\
-Even if FPE shows larger error in the peaks than CAT, when the average over all the frequencies is considered, their percentage errors are of the same order of magnitude.
-This means that outside the peaks FPE is more accurate than CAT, i.e.
-CAT is still giving noisy results.\\
-As for the normal distributed noise, mean spectrum is not sufficient to declare what is the best optimizer to be used, since most of experiment are purely observational and mean spectrum over different realizations of the same stochastic process cannot be computed.
-So, it is important to also understand what is their behaviour for single realizations of the process.
-\subsubsection{Statistical properties of single realizations} 
-The errors for the single realizations behave very differently from the errors obtained when we considered the mean spectrum.
-In computing the mean of the overall errors for each single spectrum, the results we obtain are the following:
-\begin{equation}
-    \bar{r}_{FPE} = 0.13 \pm 0.01; \qquad \bar{r}_{OBD} = 0.17 \pm .02; \qquad 
-    \bar{r}_{CAT} = 0.4 \pm 0.1
-\end{equation}
-where the error is just the standard deviation of the samples.
-In this case we find a different result: FPE shows the best results and order's estimate is associated with least errors.
-Also, having found a smaller standard deviation of the datas, we can state that FPE and OBD are more stable optimizers then CAT.
-In fact, their properties as a function of filter's length are almost identical to those obtained in the study of normal spectrum.
-If we again plot the value of the errors as a function of the estimate of the order (Figure \ref{fig:LigoOrderError})
 \begin{figure}
@@
 \end{figure}
-it is immediate to see that, again, FPE results lies in the are of filter's length associated with a minimum of relative errors, OBD lies in a clustered window with a higher error and CAT is not convergent, but most of the results are outside the minimum error area.
-So, we arrive at the same conclusion we had when we considered the normal specturm.
-OBD has no general properties that make it a preferable optimizer then FPE or CAT while, what to choose beneath the last two is based on the problem we are dealing with: if we have one single realization for the process, we are mostly sure that FPE would catch the area associated with the best resolution possible, while CAT would result in a spurious result with, in general, a not so good resolution.
-But, in case we have several realization for the same process, this last property of CAT, together with its instability, result to be his luck and, taking the average errors tend to compensate noise in the result, originating a very well defined spectrum that show same resolution almost everywhere, no matter how small the values are (as for normal PSD case) or how sharp peaks can be, the result has got generally good accuracy, also if it is clear that we have to pay this fact with very large error bars.
-So a choice between these methods is a matter of convenience and interest for each specific problem.
-\\
 

### 2021-02-25 paper/mesa.tex
Change in plots and captions

@@
 	\includegraphics[width = \linewidth]{Images/Noise and PSD/NormalPSDautocorr.pdf}
-	\caption{A priori spectrum (left) and autocorrelation (right)}
+	\caption{A priori spectrum (left) and autocorrelation (right) \sschmidt{This should be done with the same style as the others}}
 	\label{fig:autocorr}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/FPE_spectrum_estim.pdf}
-	\caption{Ensemble Average for FPE reconstruction with 90\% Credibility Regions (above) and percentage error (below)}
+	\caption{Average spectrum for FPE loss function with 90\% Credibility Regions (top) and percentage error (bottom).
+The average is computed with $1000$ realization of a $3000$ points long time series.}
 	\label{fig:FPEmean}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/OBD_spectrum_estim.pdf}
-	\caption{Reconstruction of the spectrum from 1000 simulations with OBD optimizer. Plot is mean of all the results with 90\% credibility regions (above) and percentage error (below)}
+	\caption{Average spectrum for OBD loss function with 90\% Credibility Regions (top) and percentage error (bottom).
+The average is computed with $1000$ realization of a $3000$ points long time series.}
 	\label{fig:OBDmean}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/CAT_spectrum_estim.pdf}
-	\caption{Mean spectrum for CAT optimizer, 1000 simulations for 3000 points datasets}
+	\caption{Average spectrum for CAT loss function with 90\% Credibility Regions (top) and percentage error (bottom).
+The average is computed with $1000$ realization of a $3000$ points long time series.}
 	\label{fig:CATmean}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/error_length_contour.pdf}
-	\caption{Relative error as function of the length of the filter.}
+	\caption{For each of the $1000$ independent realization of the time series, we plot the relative error $r$ (as in eq.~\ref{eq:freq_erorr}) against the length $m$ of autoregressive process.
+The time series are randomly drawn with a gaussian PSD in fig.`\ref{fig:autocorr}.
+Different colors refers to different choices for the loss function.
+Histograms for the distribution od the individual quantities are also represented.}
 	\label{fig:optcomparison}
@@
     \includegraphics[width = \linewidth]{Images/LIGOsimulate/LigoPSDoriginal.pdf}
-    \caption{Original power spectral density from Ligo}
+    \caption{Original power spectral density from Ligo \sschmidt{Remove this?? It's already in the other figure}}
     \label{fig:LigoPSDOriginal}
@@
     \centering
-     \includegraphics[width = \linewidth]{Images/LIGOsimulate/PSD.pdf}
-      \caption{A priori spectrum from Ligo database}
+     \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/compare_estimates.pdf}
+      \caption{Average spectrum for the three different loss function as compared with the ``true" PSD. The average is computed with $1000$ realization of a $3000$ points long time series.}
        \label{fig:ligospectrum}
@@
         \includegraphics[width = \linewidth]{Images/LIGOsimulate/1stPeakComparison.pdf}
-        \caption{1st peak of the spectrum and it reconstruction with every optimizer (left) and associated percentage error (right)}
+        \caption{1st peak of the spectrum and it reconstruction with every optimizer (left) and associated percentage error (right) \sschmidt{This should be done with the same style as the others}}
         \label{fig:ligo1peak}
@@
         \includegraphics[width = \linewidth]{Images/LIGOsimulate/2ndPeakComparison.pdf}
-        \caption{2nd peak of the spectrum, Relative error around 2nd peak of the spectrum}
+        \caption{2nd peak of the spectrum, Relative error around 2nd peak of the spectrum \sschmidt{This should be done with the same style as the others}}
         \label{fig:ligo2peak}
@@
     \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/error_length_contour.pdf}
-    \caption{Plot of relative error as a function of the estimate for filter's length}
+    \caption{For each of the $500$ independent realization of the time series, we plot the relative error $r$ (as in eq.~\ref{eq:freq_erorr}) against the length $m$ of autoregressive process.
+The time series are randomly drawn with a the analytical LIGO PSD in fig.`\ref{fig:ligospectrum}.
+Different colors refers to different choices for the loss function.
+Histograms for the distribution od the individual quantities are also represented.}
     \label{fig:LigoOrderError}
@@
 It is really interesting to have a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
-We perform such comparison on the spectrum of LIGO (even though similar conclusions can be drawn from any other shape of the PSD).
+We perform such comparison on the spectrum of LIGO (even though similar conclusions can be drawn from any other PSD).
 
 We perform two sets of experiments.
-In the first one (see fig.~), we simulate data from an analytical PSD and we make the comparison with them: this is to endure the we have a baseline PSD to compare the data with.
+In the first one (see fig.~), we simulate data from an analytical PSD and we make the comparison with them: this is to ensure the we have a baseline PSD to compare the data with.
 In a second experiment (see fig.~), we employ the data released to the public byt the LIGO/Virgo collaboration.
@@
 This fact is more important at lower frequencies, where the natural time length is large.
+
 Turning to the comparison between MESA and Welch's method, we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
@@
 \begin{figure}
-
 	\label{fig:mem_welch_realdata}
-	\caption{Make here a comparison between Welch}
-
+	\caption{Make here a comparison between Welch - True data}
+	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_False.pdf}
+\end{figure}
+\begin{figure}
+	\label{fig:mem_welch_fakedata}
+	\caption{Make here a comparison between Welch - Fake data}
+	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_True.pdf}
 \end{figure}

### 2021-02-25 scripts/compare_LVC_data.py

@@ -4,7 +4,8 @@ from memspectrum import MESA
 from welch import *
 import numpy as np
 import matplotlib.pyplot as plt
-from init_plotting import init_plotting
+#from init_plotting import init_plotting
+from style_sheet import *
 import pandas as pd
 from scipy.interpolate import interp1d
 
@@ -63,6 +64,10 @@ if compute:
 if plot:
 	if use_fake_data:
 		true_PSD = np.loadtxt(true_PSD_file)
+		
+	fig_overall = init_plotting()
+	fig_overall.clf()
+	fig_overall.set_size_inches(3.4,3.4*2)
 	for i, T in enumerate(T_list):
 		PSDs = np.loadtxt("plot_data/plot_{}_{}.txt".format(T, use_fake_data))
 		N = PSDs.shape[0]
@@ -84,7 +89,28 @@ if plot:
 		filename = save_folder+"comparison_LVC_data_T{}_fake_{}.pdf".format(T, use_fake_data)
 		plt.savefig(filename)
 		print("Save file @ {}".format(filename))
-		plt.show()
+		
+			#overall plot
+		ax = fig_overall.add_subplot(5,1,i+1)#, sharex = True)
+#		set_ax_style(ax)
+		ax.set_title(r"$T = {}s$".format(T))
+		ax.loglog(freqs, PSD_Welch, c = 'b', zorder = 0, label = "Welch")
+		ax.loglog(freqs, PSD_MESA, c = 'r', zorder = 1, label = "MESA")
+		if use_fake_data:
+			ax.loglog(true_PSD[:,0], true_PSD[:,1], '--', c = 'k', zorder = 2, label = "True")
+		#ax.set_ylabel(r"$PSD \left(\frac{1}{Hz} \right)$") #No ylabel
+		ax.set_xlim(20,1024)
+		if i == len(T_list)-1:
+			ax.set_xlabel(r"$f(Hz)$")
+		if i == 0:
+			ax.legend(loc = 'upper center', ncol = 2+int(use_fake_data))#, fontsize='x-small')
+	fig_overall.tight_layout()
+	filename = save_folder+"comparison_LVC_data_overall_fake_{}.pdf".format(use_fake_data)
+	plt.savefig(filename)
+	print("Save file @ {}".format(filename))
+	fig_overall.savefig(filename)
+	
+	plt.show()
 
 		
 

### 2021-02-25 scripts/init_plotting.py

@@ -2,9 +2,9 @@ import matplotlib.pyplot as plt
 
 def init_plotting():
     """
-    Initialises matplotlib *rcParams*
+    Initialises matplotlib *rcParams* and opens a figure
     
-    :return: no return
+    :return: a figure
     """
     
     # plotting options
@@ -35,3 +35,5 @@ def init_plotting():
     plt.gca().xaxis.set_ticks_position('bottom')
     plt.gca().yaxis.set_ticks_position('left')
     return fig
+    
+

### 2021-02-25 scripts/optimisers_comparison.py

@@ -12,13 +12,13 @@ from GenerateTimeSeries import generate_data
 import numpy as np
 import scipy.stats 
 import matplotlib.pyplot as plt
-from init_plotting import init_plotting
+from style_sheet import init_plotting
 from scipy.interpolate import interp1d 
 
 def relative_error(real, estimate): 
     return np.abs(real - estimate) / real 
 
-PSD = 'normal' #normal or ligo 
+PSD = 'ligo' #normal or ligo 
 save = True
 savedir = '../paper/Images/optimisers_comparison/' + PSD + '/'
 simulate_data = False 
@@ -51,7 +51,8 @@ elif PSD.lower() == 'ligo':
 T = number_of_points * dt 
 
 if simulate_data: 
-    for i in range(number_of_simulations): 
+    for i in range(number_of_simulations):
+        if i % 5 ==0: print("Simulating time series {}/{}".format(i, number_of_simulations))
         time, time_series, frequency, frequency_series, psd =\
             generate_data(f, spectrum, T, 1 / dt)
         for method in methods:
@@ -143,7 +144,7 @@ for i, method in enumerate(methods):
     if PSD == 'normal': ax3[0].set_xlim(ax3[0].get_xlim()[0], .5 / dt)
     elif PSD == 'ligo': ax3[0].set_xlim(f.min(), .5 / dt)
     ax3[0].fill_between(frequency[:-1], p5[method], p95[method], color = 'blue', alpha = .5)
-    ax3[1].loglog(frequency[:-1], ensemble_error[method], '.', color = 'k')
+    ax3[1].loglog(frequency[:-1], ensemble_error[method], '.', ms = 1,  color = 'k')
     ax3[1].set_xlabel(r'$f(Hz)$'); ax3[0].set_ylabel(y); ax3[1].set_ylabel('Percentage Error')
     fig3.tight_layout()
 
@@ -168,6 +169,7 @@ ord_fig.legend(loc = 'lower right')
 ax4.loglog(frequency, psd, '--', color = 'k') 
 ax4.set_xlim(f.min(), .5 / dt)
 fig4.legend(loc = 'upper right')
+fig4.tight_layout()
 if save: 
     err_fig.savefig(savedir + 'errors_hist.pdf', bbox_inches = 'tight')
     ord_fig.savefig(savedir + 'error_VS_order_comparison.pdf', bbox_inches = 'tight')

### 2021-02-25 scripts/style_sheet.py

@@ -1,0 +1,44 @@
+import matplotlib.pyplot as plt 
+import matplotlib.font_manager
+
+def set_ax_style( ax = None):
+    if ax is None:
+    	ax = plt.gca()
+    ax.spines['right'].set_color('none')
+    ax.spines['top'].set_color('none')
+    ax.xaxis.set_ticks_position('bottom')
+    ax.yaxis.set_ticks_position('left')
+    return
+
+def init_plotting(ratio=1., scale = 1.):
+    plt.rcParams['figure.figsize'] = (scale*3.405, scale*3.405*ratio)
+    plt.rcParams['font.size'] = 7
+    plt.rcParams['font.family'] = 'DejaVu Sans'
+    #plt.rcParams['font.sans-serif'] = ['Bitstream Vera Sans']
+    plt.rcParams['axes.labelsize'] = plt.rcParams['font.size']
+    plt.rcParams['axes.titlesize'] = plt.rcParams['font.size']
+    plt.rcParams['legend.fontsize'] = plt.rcParams['font.size']
+    plt.rcParams['xtick.labelsize'] = 6
+    plt.rcParams['ytick.labelsize'] = 6
+    plt.rcParams['xtick.major.size'] = 3
+    plt.rcParams['xtick.minor.size'] = 3
+    plt.rcParams['xtick.major.width'] = 1
+    plt.rcParams['xtick.minor.width'] = 1
+    plt.rcParams['ytick.major.size'] = 3
+    plt.rcParams['ytick.minor.size'] = 3
+    plt.rcParams['ytick.major.width'] = 1
+    plt.rcParams['ytick.minor.width'] = 1
+    plt.rcParams['legend.frameon'] = False
+    plt.rcParams['legend.loc'] = 'center left'
+    plt.rcParams['axes.linewidth'] = 1
+    plt.rcParams['contour.negative_linestyle'] = 'solid'
+    plt.rcParams['axes.spines.right'] = False
+    plt.rcParams['axes.spines.top'] = False
+    plt.rcParams['xtick.minor.top'] = False
+    plt.rcParams['ytick.minor.left'] = True
+    plt.rcParams['xtick.major.top'] = False
+    plt.rcParams['ytick.major.left'] = True
+    fig = plt.figure()
+    return fig
+    
+

### 2021-03-01 climate_data/analyse_climate_data.py
Added analysis temp series

@@ -1,0 +1,49 @@
+#####
+#Loading climate data
+#Request ID: 72b941a9-4c27-4d17-8061-3525cfcbd646
+
+#Variable:2m
+#temperatureYear:2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020
+#Month:January, February, March, April, May, June, July, August, September, October, November, December
+#Day:01, 02, 03, 04, 05, 06, 07, 08, 09, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31
+#Time:00:00, 01:00, 02:00, 03:00, 04:00, 05:00, 06:00, 07:00, 08:00, 09:00, 10:00, 11:00, 12:00, 13:00, 14:00, 15:00, 16:00, 17:00, 18:00, 19:00, 20:00, 21:00, 22:00, 23:00
+#Sub-region extraction:North 45.5°, West 9.1°, South 45.4°, East 9.2°
+#Format:GRIB
+
+######
+
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+sys.path.insert(0,'..')
+import memspectrum
+
+data = np.loadtxt('data_@1h_len95689.dat')
+
+srate = 1./3600 # 1/(1 hours)
+T = len(data)/srate
+
+#computing spectrum
+M = memspectrum.MESA()
+M.solve(data)
+f = np.linspace(1./(T), 0.5*srate,1000)
+spec = M.spectrum(1./srate, f)
+
+plt.figure()
+plt.plot(np.linspace(0, len(data)/24., len(data)), data)
+plt.xlabel("Days")
+plt.ylabel("Temperature (K)")
+plt.savefig("temp_timeseries.pdf")
+
+plt.figure()
+unit_shift = 3600*24 #frequency is made 1/day
+plt.loglog(f * unit_shift , spec, label = "spectrum", c = 'k')
+plt.axvline(1./(3600.)*unit_shift, label = "hour", c = 'r', ls = '--')
+plt.axvline(1./(3600.*24.)*unit_shift, label = "day", c = 'b', ls = '--')
+plt.axvline(1./(3600.*24.*365.)*unit_shift, label = "year", c = 'g', ls = '--')
+plt.xlabel("frequency (1/day)")
+plt.ylabel("PSD")
+plt.legend()
+plt.savefig("spectrum.pdf")
+
+plt.show()

### 2021-03-01 climate_data/download_data.py

@@ -1,0 +1,47 @@
+import cdsapi
+
+c = cdsapi.Client()
+
+c.retrieve(
+    'reanalysis-era5-land',
+    {
+        'format': 'grib',
+        'area': [
+            45.5, 9.1, 45.4,
+            9.2,
+        ],
+        'time': [
+            '00:00', '01:00', '02:00',
+            '03:00', '04:00', '05:00',
+            '06:00', '07:00', '08:00',
+            '09:00', '10:00', '11:00',
+            '12:00', '13:00', '14:00',
+            '15:00', '16:00', '17:00',
+            '18:00', '19:00', '20:00',
+            '21:00', '22:00', '23:00',
+        ],
+        'day': [
+            '01', '02', '03',
+            '04', '05', '06',
+            '07', '08', '09',
+            '10', '11', '12',
+            '13', '14', '15',
+            '16', '17', '18',
+            '19', '20', '21',
+            '22', '23', '24',
+            '25', '26', '27',
+            '28', '29', '30',
+            '31',
+        ],
+        'month': [
+            '01', '02', '03',
+            '04', '05', '06',
+            '07', '08', '09',
+            '10', '11', '12',
+        ],
+        'year': [
+            '2019', '2020',
+        ],
+        'variable': '2m_temperature',
+    },
+    'download.grib')

### 2021-03-01 climate_data/test_grib.py

@@ -1,0 +1,80 @@
+# https://github.com/ecmwf/cfgrib
+# TO download data: https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-land?tab=overview
+
+import xarray as xr
+import numpy as np
+import matplotlib.pyplot as plt
+
+ds = xr.open_dataset('adaptor.mars.internal-1614589133.4434555-10786-9-72b941a9-4c27-4d17-8061-3525cfcbd646.grib', engine='cfgrib', backend_kwargs={'read_keys': ['experimentVersionNumber']})
+print(ds)
+print(ds.t2m.attrs)
+#help(xr.open_dataset)
+#help(ds)
+
+print("UFFAAAAAA")
+
+print(ds.data_vars['t2m'])
+
+print("UFFAAAAAAAAAAAAAAAAAAAAA\n")
+
+print(ds.data_vars['t2m'].time) #actually keeps the day
+print(ds.data_vars['t2m'].step)
+print(ds.data_vars['t2m'].latitude)
+print(ds.data_vars['t2m'].longitude)
+print(ds.data_vars['t2m'].shape)
+
+data = ds.data_vars['t2m']
+
+	#doing a time series
+b = np.array(ds.data_vars['t2m'][:,:,0,0])
+print(b.shape)
+print(b[:,:])
+
+b = b.flatten()
+print(b)
+b = b[~np.isnan(b)]
+print(b)
+t = np.linspace(0, len(b), len(b)) #hours
+
+np.savetxt('data_@1h_len{}.dat'.format(len(b)),b)
+quit()
+
+plt.plot(t, b)
+plt.show()
+
+quit()
+a = []
+for i in range(2):
+	print(float(ds.data_vars['t2m'].time[i]))
+	for k in range(data.shape[1]):
+		t = float(ds.data_vars['t2m'][i,k,50,150])
+		print(float(ds.data_vars['t2m'].step[k])/1e9, t)
+		a.append(t)
+
+plt.plot(a)
+plt.show()
+quit()
+	#doing a latitude map
+	#For some reasons there are a lot of nans around: probably this is required because not every resolution has the same angles??
+
+lat_list = []
+k = 2
+for i in range(10,int(data.shape[2]/2)):
+	long_list = []
+	for j in range(10,int(data.shape[3]/2)):
+		if not np.isnan(float(ds.data_vars['t2m'][1,k,i,j])):
+			#print(k,i,j)
[…]

### 2021-03-01 climate_data/analyse_climate_data.py
Forecast added

@@ -12,6 +12,8 @@
 
 ######
 
+#PROBLEM: the code does not catch the trend over the year timescale!!
+
 import numpy as np
 import matplotlib.pyplot as plt
 import sys
@@ -29,12 +31,20 @@ M.solve(data)
 f = np.linspace(1./(T), 0.5*srate,1000)
 spec = M.spectrum(1./srate, f)
 
+#forecasting
+N_forecast = 24*1000 #10 day of forecasting
+predictions = M.forecast(data[:-N_forecast], N_forecast, 100)
+
+
+	#plot time series
 plt.figure()
-plt.plot(np.linspace(0, len(data)/24., len(data)), data)
+t = np.linspace(0, len(data)/24., len(data))
+plt.plot(t, data)
 plt.xlabel("Days")
 plt.ylabel("Temperature (K)")
 plt.savefig("temp_timeseries.pdf")
 
+	#plot spectrum
 plt.figure()
 unit_shift = 3600*24 #frequency is made 1/day
 plt.loglog(f * unit_shift , spec, label = "spectrum", c = 'k')
@@ -46,4 +56,18 @@ plt.ylabel("PSD")
 plt.legend()
 plt.savefig("spectrum.pdf")
 
+	#plot forecasting
+plt.figure()
+plt.title("{} days of forecasting".format(int(N_forecast/24)))
+l, m, h = np.percentile(predictions,[5,50, 95],axis=0)
+plt.plot(t[-(M.get_p()+N_forecast):],data[-(M.get_p()+N_forecast):],linewidth=1.5,color='r',zorder=2, label = 'data')
+plt.axvline(t[-N_forecast])
+plt.plot(t[-N_forecast:],m,'--',linewidth=0.5, color='k', zorder = 0)
+plt.fill_between(t[-N_forecast:],l,h,facecolor='turquoise',alpha=0.5, zorder = 1)
+plt.savefig("forecast.pdf")
+
 plt.show()
+
+
+
+

### 2021-03-02 paper/mesa.tex
another big chunk of edits

@@
 considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA}. 
-The entropy gain obtained by filtering white noise is: 
+The entropy gain obtained by filtering white noise with a sampling rate $\Delta t$ and Nyquist frequency $Ny \equiv \frac{1}{2 \Delta t}$ is:
+
 \begin{equation}\label{eq:EntropyGain}
-    \Delta H = \int_{-\infty}^{\infty}\log S(f) df , 
-\end{equation}
-hence the variational problem can be stated by requiring that $S(f)$ maximises the entropy gain \sschmidt{Non ho capito questa parte...:(}. 
-Time-limited observations do not provide access to the full frequency range, but only between minus and plus the Nyquist frequency, 
-that, in turn is given by the sampling interval $\Delta t$ as $Ny \equiv \frac{1}{2 \Delta t}$. Hence, the finite-size entropy
-gain is 
-\begin{equation}\label{eq:EntropyGain2}
     \Delta H = \int_{- Ny}^{Ny}\log S(f) df\,.
@@
 \end{equation}
-This approach on power spectral density estimate was developed by Burg \cite{burg1975maximum} and it is a way out from the problems raised in periodograms estimate.
+This approach on power spectral density estimate was developed by Burg~\cite{burg1975maximum} and it is a way out from the problems raised in periodograms estimates.
 It requires no assumptions on the unavailable estimates for $\bar r_n$ and is consistent with the sample autocorrelation function by construction.
@@
 To show the analogy, we compute the PSD of such process and we show that it is formally equivalent to the PSD obtained in eq.~\ref{eq:MESApsd}.
-We start taking the z \footnote{Spiegare in due righe la z transform} transform of the previous eq.~\ref{eq:AR_p}: 
+We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~(\ref{eq:AR_p}):
 \begin{align}
@@
 \end{equation}
-Since we assumed a stationary process, $\tilde{x}(z)$ is analytic both on and inside the unit circle.
+Since we assumed a wide-sense stationary process, $\tilde{x}(z)$ is analytic both on and inside the unit circle.
 Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
@@
     S(f) = \vert \tilde x(z)\vert ^ 2 = 
-    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}.
+    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}\,.
 \end{equation}
@@
 Eq.~(\ref{eq:ARspectrum}) and Eq.~(\ref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
-This shows that the MAXENT estimation of the PSD computes the coefficients of the autoregressive process that is able to model the data faithfully.
-
-\sschmidt{I don't understand the next of this section...}
-Having computed the prediction error filter one can easily obtain an estimate of the $AR$ coefficients for the process just changing sign to the prediction error filter $a_i$.
-Taken a time series, MESA allows us to compute the power spectrum and to find an effective description in terms of an autoregressive process. 
-Wold's theorem ensures that this (approximate) description for the phenomenon is legitimate when the process is found to be stationary.
+This shows that the MAXENT estimation of the PSD also computes the coefficients of the autoregressive 
+process that model the observed process.
+
+%\sschmidt{I don't understand the next of this section...}
+%Having computed the prediction error filter one can easily obtain an estimate of the $AR$ coefficients for the process just changing sign to the prediction error filter $a_i$.
+%Taken a time series, MESA allows us to compute the power spectrum and to find an effective description in terms of an autoregressive process. 
+%Wold's theorem ensures that this (approximate) description for the phenomenon is legitimate when the process is found to be stationary.
 So, MESA defines a bridge between spectral estimation and the study of autoregressive processes, an unexpected and fundamental result.
 \\
-Another very remarkable process for this estimate, is that the linear predictor obtained solving the error filter equation, is the best linear predictor, i.e.
+%Another very remarkable process for this estimate, is that the linear predictor obtained solving the error filter equation, is the best linear predictor, i.e.
 it minimizes error, and is equivalent to a least square fit for the coefficients.
 \cite{VDBos}  \\
-To prove that it minimizes error, we will follow the original proof provided by Burg \cite{burg1975maximum}. Consider a general linear predictor $-c_i$: 
-\begin{equation}
-    x_{t} - \sum_{i = 1}^p (-c_i)x_{t - i} = \sum_{i = 0}^p c_i x_{t - i} = e'_t, 
-\end{equation}
-Denoting with $E\left[\cdot\right]$ the expectation value of some variable, the mean square error is given by: 
-\begin{equation}
-E\left[\sum_{i = 0}^p c_i x_{t-i}\sum_{j = 0}^p c^*_j x^*_{t - j}\right] = \sum_{i,j} c^*_j E\left[x^*_{t-j}x_{t-i} \right] c_j = \sum_{i, j}c^*_j R(j - i) c_i > 0,
-\end{equation}
-which, being $R$ positive definite and defined by the multiplication of a complex number times its complex conjugate, is always positive definite. 
-Let $-\vec{b}$ be the prediction error filter obtained with MESA. Multiplying the left hand side of \ref{eq:predictionErrorFilter} for the complex conjugate of the forward prediction error, 
-\begin{equation}\nonumber 
-    \sum_{i,j}b^*_jR(j-i)b_i = P_N
-\end{equation}
-is obtained . \\ 
-The (positive) quantity  
-\begin{equation}\label{eq:linearPredictor}
-    \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i),
-\end{equation}
-using Einstein notation on repeated indices and recall that $c_0 = b_0 = 1$, can be rewritten as 
-\begin{align}
-    & c^*_j R(j - i) c_i - c^*_j R(j - i) b_i - b^*_j R(j - i) c_i + b^*_j R(j - i)  b_i =\\ \nonumber 
-    &c^*_j R(j - i) c_i - c^*_0 P_N - P_N c_0 + P_N =\\ \nonumber
-    & c^*_j R(j - i) c_i - P_N.
-\end{align}
-Moving $P_N$ on the left hand side, it becomes 
-\begin{equation}\label{eq:bestLinearPredictor}
-    \sum_{i, j}c^*_j R(j-i)c_i = \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i) + P_N.
-\end{equation}
-Being all the quantities positive definite, the left hand side is minimized by the choice $c_i = b_i$.
+%To prove that it minimizes error, we will follow the original proof provided by Burg \cite{burg1975maximum}. Consider a general linear predictor $-c_i$: 
+%\begin{equation}
+%    x_{t} - \sum_{i = 1}^p (-c_i)x_{t - i} = \sum_{i = 0}^p c_i x_{t - i} = e'_t, 
+%\end{equation}
+%Denoting with $E\left[\cdot\right]$ the expectation value of some variable, the mean square error is given by: 
+%\begin{equation}
+%E\left[\sum_{i = 0}^p c_i x_{t-i}\sum_{j = 0}^p c^*_j x^*_{t - j}\right] = \sum_{i,j} c^*_j E\left[x^*_{t-j}x_{t-i} \right] c_j = \sum_{i, j}c^*_j R(j - i) c_i > 0,
+%\end{equation}
+%which, being $R$ positive definite and defined by the multiplication of a complex number times its complex conjugate, is always positive definite. 
+%Let $-\vec{b}$ be the prediction error filter obtained with MESA. Multiplying the left hand side of \ref{eq:predictionErrorFilter} for the complex conjugate of the forward prediction error, 
+%\begin{equation}\nonumber 
+%    \sum_{i,j}b^*_jR(j-i)b_i = P_N
+%\end{equation}
+%is obtained . \\ 
+%The (positive) quantity  
+%\begin{equation}\label{eq:linearPredictor}
+%    \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i),
+%\end{equation}
+%using Einstein notation on repeated indices and recall that $c_0 = b_0 = 1$, can be rewritten as 
+%\begin{align}
+%    & c^*_j R(j - i) c_i - c^*_j R(j - i) b_i - b^*_j R(j - i) c_i + b^*_j R(j - i)  b_i =\\ \nonumber 
+%    &c^*_j R(j - i) c_i - c^*_0 P_N - P_N c_0 + P_N =\\ \nonumber
+%    & c^*_j R(j - i) c_i - P_N.
+%\end{align}
+%Moving $P_N$ on the left hand side, it becomes 
+%\begin{equation}\label{eq:bestLinearPredictor}
+%    \sum_{i, j}c^*_j R(j-i)c_i = \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i) + P_N.
+%\end{equation}
+%Being all the quantities positive definite, the left hand side is minimized by the choice $c_i = b_i$.
 Also, known $P_N$, the previous equation gives an estimate on the additional mean square error associated at choices that differ from the prediction error filter.
@@
 \subsection{Forecasting} \label{sec:forecasting}
-The link between MESA and autoregressive process allow us to forecast from a given time series, thus providing plausible future observation for the data.
-Indeed, an autoregressive process of order $p$ models the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation:
+The link between MESA and autoregressive processes of particular interest.
+Given the solution to the Levinson recursion to determine the $a_k$, see Appendix A, we have automatically the coefficients of the AR with equivalent PSD, hence we are able to exploit Eq.~(\ref{eq:AR_p}) to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
+Indeed, an AR($p$) process models the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation:
 \begin{align}\label{eq:p_forecast}
@@
 \end{align}
-The interpretation of eq.~\ref{eq:p_forecast} is straightforward: $x_t$ follows a Guassian distribution with a fixed variance an a mean value $m_t = \sum_{i = 1}^p b_i x_{t-i}$ computed from past observations.
-Equation~\ref{eq:p_forecast} provides then a well defined probability framework for predicting future observations: this is a very useful feature of the method, which does not have an equivalent in any other spectral analysis computation.
+The interpretation of Eq.~(\ref{eq:p_forecast}) is straightforward: $x_t$ follows a Gaussian distribution with a fixed variance an a mean value $m_t = \sum_{i = 1}^p b_i x_{t-i}$ computed from past observations.
+Eq.~(\ref{eq:p_forecast}) provides then a well defined probability framework for predicting future observations: this is a very useful feature of MESA, that does not have an equivalent in any other spectral analysis computation.
 
 \section{Validation of the model}
-It is now clear that MESA algorithm provides a recursive formula for computing the coefficients $a_k$ in eq.~\ref{eq:MESApsd}.
+It is now clear that MESA provides a recursive formula for computing the coefficients $a_k$ in Eq.~(\ref{eq:MESApsd}).
 The number $M$ of such coefficients is the same as the number of time steps considered in the computation of the autocorrelation $\bar{r}_m$.
@@
 
-It is then clear that the choice of how many points of the autocorrelation to consider is important: on one hand it is advisable to include as many knowledge of the autocorrelation as possible, leading to include all the known $\bar{r}_m$; on the other hand, including values of the autocorrelation that are not reliably estimated, can be counterproductive.
-Indeed, the order $M$ of the autocorrelation to be considered (or equivalently the order $M$ of the underlying autoregressive process) is the only input required by the user and a careful balance between this two necessities must be made when applying the algorithm.
-
-This section is devoted to an extensive study on how to make such choice.
-In section~\ref{sec:optimizers}, we are going to define three different \textit{loss functions} to measure how well the algorithm is able to reproduce a known PSD.
-The basic idea is to validate, as the autoregressive order considered increase, the performance of the algorithm results by measuring the loss function and pick, among the orders the one that yields better results.
-Second, the performance of different loss functions will be assessed by performing the spectral analysis on time series with given analytical PSD, sec.~\ref{sec:arp_validation}.
-In a last subsection~\ref{sec:LIGO_validation}, the same analysis is performed on real data from LIGO-Virgo gravitational waves observatory.
-
-\subsection{Choosing an autoregressive order}\label{sec:optimizers}
-
-A good choice for the autoregressive order $M$ is proposed to be \cite{doi:10.1190/1.1440902}
+It is then clear that the choice of how many samples of the discrete autocorrelation to consider is important: 
+on the one hand it is advisable to include as much knowledge of the autocorrelation as possible, leading to include all the known $\bar{r}_m$; on the other hand, including values of the autocorrelation that are not reliably estimated, can be counterproductive.
+Indeed, the order $M$ of the autocorrelation to be considered (or equivalently the order $M$ of the underlying autoregressive process) is the only tuning parameter of MESA and a careful balance between this two necessities must be made when applying the algorithm.
+
+The reminder of this section is devoted to an extensive study on how to make such choice.
+In Section~\ref{sec:optimizers}, we are going to define three different \textit{loss functions} to measure how well the 
+algorithm is able to reproduce a known PSD.
+The basic idea is to validate, as the autoregressive order considered increase, the performance of the algorithm results 
+by measuring the loss function and pick, among the orders the one that yields better results.
+Second, the performance of different loss functions will be assessed by performing the spectral analysis on time series with given analytical PSD, Sec.~\ref{sec:arp_validation}.
+In a last subsection~\ref{sec:LIGO_validation}, the same analysis is performed on real data from the LIGO-Virgo gravitational waves observatories.
+
+\subsection{Choice of the autoregressive order}\label{sec:optimizers}
+
+Guided from numerical experiments, an indication on the upper bound to the autoregressive order $M_{max}$ is~\cite{doi:10.1190/1.1440902}:
 \begin{equation}\label{eq:MMAx}
-M_{max} = 2N / \ln{(2N)}
-\end{equation}
-However, this is just an upper limit for the estimate of the length $M$ and the optimal algorithm could employ much less points.
-We then need a more sophisticated method for computing the right value for $M$.
+M_{max} = 2N / \ln{(2N)}\,,
+\end{equation}
+where $N$ is the number of points in the time-series.
+However, this is just a plausible upper limit on the order of the AR process $m$ and the optimal algorithm could employ fewer points.
+We then need a more sophisticated method for computing the right value for $m$.
 Various loss functions to assess the algorithm performance have been proposed in literature.
-We summarize them below:
+We summarise them below:
 
 \begin{itemize}
-\item \textbf{Final prediction Error (FPE)} 
-The first criterion is due to Akaike (1970) \cite{Akaike1970StatisticalPI}. He proposed that M has to be chosen as the length that reduces the error when the filter is used as a predictor: 
+\item \textbf{Final prediction Error } 
+The first criterion is due to Akaike~\cite{Akaike1970StatisticalPI}. He proposed that $m$ should be chosen as the 
+length that minimises the error when the filter is used as a predictor, the \emph{final prediction} error (FPE): 
 \begin{equation}
@@
 with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.
-Minimizing FPE is the same as minimizing the quantity: 
-\begin{equation}
-    P_{m} \frac{N + m + 1}{N - m - 1}
-\end{equation}
-with $P_M$ defined by the prediction error filter equation \sschmidt{What is this equation?}.
+Minimizing FPE is equivalent to minimizing the quantity: 
+\begin{equation}
+    \mathcal{L}_{\rm FPE}(m) = P_{m} \frac{N + m + 1}{N - m - 1}
+\end{equation}
+with $P_m$ defined by the prediction error filter equation \sschmidt{What is this equation?}.
 \item \textbf{Criterion Autoregressive Transfer function (CAT)}
-This second loss function is due to Parzen \cite{bhansali1986}, and it is defined as
-\begin{equation}
-    CAT(m) = \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
-\end{equation}
-It has the interesting property that the distance between the original filter and the estimate is minimum and asymptotically unbiased.
-
-\item \textbf{Optimum Bayes Decision rule (OBD)} 
-The last criterion we will consider is the OBD (Rao, 1982)\cite{doi:10.1029/WR018i004p01097}.
-With the technique of asymptotic integration, th loss function is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}:
+This second loss function is due to Parzen~\cite{bhansali1986}, and it is defined as
+\begin{equation}
+    \mathcal{L}_{\rm CAT}(m) = \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
+\end{equation}
+It has the interesting property that the distance between the original filter and the estimate is minimum and asymptotically unbiased. \wdp{please expand}
+
+\item \textbf{Optimum Bayes Decision rule} 
+The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}.
+With the technique of asymptotic integration, the loss function is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}:
    \begin{align}
-        OBD&(m) = (N - m - 2) \log(P_m) \nonumber\\
-        &+ m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2.
+        \mathcal{L}_{\rm OBD}&(m) = (N - m - 2) \log(P_m) \nonumber\\
+        &+ m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2\,.
     \end{align}
@@
 
-Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached.
-Then, the order $M$ is selected to be the one that minimize the loss function.
-
-In a real implementation of the algorithm, computing all the recursion up to $M_{max}$ can result in a significant waste of computational power: the optimal value is often $M_{opt}<< M_{max}$ and, in such cases, computing all the values of $m$ until $M_{max}$ is not useful.
-In practice, we can apply an \textit{early stopping} procedure: every few iterations we look for the best order of $M_{opt}$; if this value does not change for a while, the computation is stopped and a good (local) minimum of the loss function is found.
+Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson\wdp{add ref} recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached.
+Then, the order $m$ is selected to be the one that minimizes the loss function.
+
+In a real implementation of the algorithm, computing all the recursion up to $M_{max}$ can result in a significant waste of computational power: the optimal value is often $m_{opt} << M_{max}$ and, in such cases, computing all the values of $m$ until $M_{max}$ is not useful.
+In practice, we can apply an \textit{early stopping} procedure: every few iterations we look for the best order of $m_{opt}$; if this value does not change for a while, we assume that a good (local) minimum of the loss function is found and the computation is stopped.
 
@@
 \subsection{Choosing a loss function - A gaussian PSD} \label{sec:arp_validation}
-We test the performance of the three loss functions on a random time series with a known power spectral density.
-We take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$, see fig.~\ref{fig:autocorr}.
-Of course, the time series will be generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$ and performing the inverse Fast Fourier Transform on the frequency series.
-
-We will apply Burg method with each loss function on a number of simulated time series, obtaining a variety of estimate for PSD.
-The disagreement between the PSD $S_i(f)$ estimated from the i-th simulated time series and the target PSD $S(f)$ is measured with the relative error $r_i$, averaged over all frequencies:
+We test the performance of the three loss functions on a random time series generated with a known power spectral density.
+We take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$, see Fig.~\ref{fig:autocorr}.
+The time series will be generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$, following the procedure outlined in ~\cite{}, and performing the inverse Fast Fourier Transform on the frequency series.
+
+We will apply the Burg method choosing in turn each of the loss functions on an ensemble of simulated time series, 
+thus obtaining an ensemble of PSD estimates, that will aloows the statistical characterisation of the performance of each loss function.
+The disagreement between the PSD $S_i(f)$ estimated from the $i$-th simulated time series and the target PSD $S(f)$ is measured via 
+the relative error $r_i$, averaged over all frequencies:
 \begin{equation}\label{eq:freq_erorr}
-r_i = \frac{1}{f_{Ny}}\int_0^{f_{Ny}} \d{f} \frac{|S_i(f) - S(f)|}{S(f)}
-\end{equation}
-where $f_{Ny}$ is the Nyquist frequency of the time series. Of course in practice, the integral will be computed by averaging over a discrete grid of frequencies.
-
-We will consider every loss function separately and compare the results only at the end of this section.
-For each loss function, we will compute the ensemble averaged PSD estimation (as well as the $90\%$ confidence level) and the frequency dependence of the error (ensemble averaged):
+r_i = \frac{1}{Ny}\sum_{f_j=0}^{f_j = Ny} \frac{|S_i(f_j) - S(f_j)|}{S(f_j)}
+\end{equation}
+where $Ny$ is the Nyquist frequency of the time series.
+
+For each loss function, we compute the ensemble-averaged PSD (as well as the $90\%$ confidence level) and the ensemble-averaged frequency dependence of the error:
 \begin{equation}\label{eq:ensemble_averaged_error}
@@
 \end{equation}
-In figure~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} we report the averaged PSD for the three loss functions, as well as the ensemble averaged error in eq.~\ref{eq:ensemble_averaged_error}.
-
-Furthermore, we display in a contour plot the correlation between $r_i$ and the autoregressive order $m_i$ chosen for each independently drawn time series.
-This is done in figure~\ref{fig:optcomparison}.
-
-For our studies, we generate a dataset of 1000 time series of 3000 points each: see fig.~\ref{fig:noise} for an example of a time series in frequency and time domain.
-In section~\ref{sec:LIGO_validation}, we will perform the analysis again using the Advanced LIGO design sensitivity PSD \cite{Ligo}.
+Figs.~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} show the averaged PSD for the three loss functions, 
+as well as the ensemble-averaged error from Eq.~(\ref{eq:ensemble_averaged_error}).
+Furthermore, Fig.~\ref{fig:optcomparison} displays the correlation between $r_i$ and the autoregressive order $m_i$ chosen for each independently drawn time series.
+
+For this investigation, we generate a dataset of 1000 time series of 3000 points each, see Fig.~\ref{fig:noise} for an example of a time series in frequency and time domain.
+%In section~\ref{sec:LIGO_validation}, we will perform the analysis again using the Advanced LIGO design sensitivity PSD \cite{Ligo}.
 
@@
 	\includegraphics[width = \linewidth]{Images/Noise and PSD/NormalPSDautocorr.pdf}
-	\caption{A priori spectrum (left) and autocorrelation (right) \sschmidt{This should be done with the same style as the others}}
+	\caption{Synthetic spectrum (left) and relative autocorrelation function (right) used to simulate the data to assess the performance of the three loss functions considered in Sec.~\ref{sec:arp_validation}\sschmidt{This should be done with the same style as the others}.}
 	\label{fig:autocorr}
@@
     \caption{Noise in both time and frequency domain.
-Frequency noise behaviour is representative of the chosen spectrum, having its maximum amplitude around $f = 2.5 au$ \sschmidt{Togliamo questa figura?}}
+Frequency noise behaviour is representative of the chosen spectrum, having its maximum amplitude around $f = 2.5 au$ \sschmidt{Togliamo questa figura?}\wdp{direi di si}}
     \label{fig:noise}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/FPE_spectrum_estim.pdf}
-	\caption{Average spectrum for FPE loss function with 90\% Credibility Regions (top) and percentage error (bottom).
+	\caption{Average spectrum for $\mathcal{L}_{\rm FPE}(m)$ with 90\% confidence regions (top) and percentage error (bottom).
 The average is computed with $1000$ realization of a $3000$ points long time series.}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/OBD_spectrum_estim.pdf}
-	\caption{Average spectrum for OBD loss function with 90\% Credibility Regions (top) and percentage error (bottom).
+	\caption{Average spectrum for $\mathcal{L}_{\rm OBD}(m)$ loss function with 90\% confidence regions (top) and percentage error (bottom).
 The average is computed with $1000$ realization of a $3000$ points long time series.}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/CAT_spectrum_estim.pdf}
-	\caption{Average spectrum for CAT loss function with 90\% Credibility Regions (top) and percentage error (bottom).
+	\caption{Average spectrum for $\mathcal{L}_{\rm CAT}(m)$ loss function with 90\% confidence regions (top) and percentage error (bottom).
 The average is computed with $1000$ realization of a $3000$ points long time series.}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/error_length_contour.pdf}
-	\caption{For each of the $1000$ independent realization of the time series, we plot the relative error $r$ (as in eq.~\ref{eq:freq_erorr}) against the length $m$ of autoregressive process.
-The time series are randomly drawn with a gaussian PSD in fig.`\ref{fig:autocorr}.
-Different colors refers to different choices for the loss function.
-Histograms for the distribution od the individual quantities are also represented.}
+	\caption{Relative error $r$ (as in Eq.~(\ref{eq:freq_error})) vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with gaussian PSD given in Fig.~\ref{fig:autocorr}.
+Different colours refers to different choices for the loss function: in red $\mathcal{L}_{\rm FPE}(m)$, in green $\mathcal{L}_{\rm OBD}(m)$ and in black $\mathcal{L}_{\rm CAT}(m)$.
+The top and right histograms show the marginal distributions.}
 	\label{fig:optcomparison}

### 2021-03-02 paper/mesa.tex
Fixed appendix

@@
 \newcommand{\wdp}[1]{{\textcolor{magenta}{\texttt{WDP: #1}} }}
-\newcommand{\amartini}[1]{{\textcolor{blue}{\texttt{SS: #1}} }}
+\newcommand{\amartini}[1]{{\textcolor{blue}{\texttt{AM: #1}} }}
 \newcommand{\sschmidt}[1]{{\textcolor{red}{\texttt{SS: #1}} }}
@@
 \subsection{MESA solution}
-We will follow the solution proposed by Burg in his PhD thesis \cite{burg1975maximum}.
-Instead of directly solving the variational problem for the power spectral density, we will rewrite it in terms of the constraints (\ref{eq:MaxConstraint}) of the problem.
-$S(f)$ is defined as the Fourier Transform of the autocorrelation function, as:
-\begin{equation}
-    S(f) = \frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{- \imath 2 \pi n \Delta t}.
-\end{equation}
-Substituting in \ref{eq:EntropyGain2}, the variational problem reduces to the maximization of: 
+The proposed solution follows the original demonstration due to Burg \cite{burg1975maximum}.
+The constraints (\ref{eq:MaxConstraint}) can be directly implemented by substituting $S(f)$ as the Fourier transform of the autocorrelation function
+\begin{equation}
+    S(f) = \frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{- \imath 2 \pi n \Delta t}
+\end{equation}
+in equation \ref{eq:EntropyGain}. In this way the variational problem is reduced to the maximization of: 
 \begin{equation}
     \int_{-Ny}^{Ny}  
-    \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{-\imath 2 \pi f n \Delta t} 
+    \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r(n\Delta t) e^{-\imath 2 \pi f n \Delta t} 
     \right) df.
 \end{equation}
-As we said before, not all the values for the autocorrelation are known.
-Generally, we have access to a finite number of values for the autocorrelation function, ranging in a symmetric interval $n \in [-N, N]$.
-In the previous equation, since the summation index is ranging between $-\infty$ and $+\infty$, we are correctly assuming that also the unknown values have to be used in the computation of the power spectral density.
-But, since they are not known, we have to require the maximization of entropy to be independent of these values:
+Since the autocorrelation function can only be computed for a finite time interval $n \in [-N, N]$, the requirement for our estimate to be independent from the unkown values of the autocorrelation function can be implemented as:
 \begin{equation}\nonumber 
@@
 \end{equation}
-Using the explicit form for $H$, this becomes
-\begin{equation}
-      \frac{\delta H}{\delta \bar r_s} = \frac{1}{2Ny}\int_{-Ny}^{Ny} S(f)^{-1}e^{-\imath 2 \pi f s \Delta t } df = \lambda_s = 0 \text{ for } \vert s \vert > N. 
-\end{equation}
-We see that $S(f)$ can be expressed in terms of the $\lambda s$ via a Fourier Series
+By explicitly computing the derivative: 
+\begin{equation}
+      \frac{\delta H}{\delta \bar r_s} = \frac{1}{2Ny}\int_{-Ny}^{Ny} S(f)^{-1}e^{-\imath 2 \pi f s \Delta t } df = \lambda_s
+\end{equation}
+is easily seen that $S(f)$ can be expressed via a Fourier Series
 \begin{equation}\label{eq:PSDconstraint}
-    S(f)^{-1} = \sum_{s = -N}^N \lambda_s e^{-\imath 2 \pi f s \Delta t},
-\end{equation}
-This is a non-linear closed expression for the power spectral density. The requirement for $S(f)$ to be positive definite implies that: 
-\begin{equation}
-    \nonumber 
-    \lambda_{-s} = \lambda_s^* 
-\end{equation}
-With this property, and defining $z = e^{-\imath 2 \pi f \Delta t}$, $S(f)^{-1}$ can be rewritten as a polynomial in $z$ and $z^*$
+    S(f)^{-1} = \sum_{s = -\infty}^\infty \lambda_s e^{-\imath 2 \pi f s \Delta t}
+\end{equation}
+with the properties $\lambda_s = 0$ for $\vert s \vert > N$ and $\lambda_s = \lambda_{-s}^*$.  
+Defining $z = e^{-\imath 2 \pi f \Delta t}$ the previous Fourier expansion becomes a Laurent Polynomial in $z$: 
 \begin{equation}
@@
 \end{equation}
-Let $z_0$ be a root for the polynomial
-\begin{equation}
-    \nonumber
-    \lambda_0 + \sum_{s = 1}^N \lambda_s z_0^s + \sum_{s = 1}^N \lambda^*_s z_0^{-s} = 0,
-\end{equation}
-it is easy to show that $(z_0^*)^{-1}$ is also a root for the equation: 
-\begin{align}
-    \nonumber 
-    &\lambda_0 + \sum_{s = 1}^{N} \lambda s \left(\frac{1}{z_0^s} \right)^* + \sum_{s = 1}^N \lambda^*_s \left(\frac{1 }{z_0^{-s}}\right)^* = \\ \nonumber 
-    &\lambda_0 + \sum_{s = 1}^N \lambda_s (z_0^{-s})^* + \sum_{s=1}^N \left(\lambda_s (z_0^s)\right)^* = \\ \nonumber 
-    &\left(\lambda_0 + \sum_{s=1}^N \lambda_s z_0^s + \sum_{s = 1}^N \lambda^*_s z_0^{-s}\right)^* = 0 
-\end{align}
-This means that for every pole $z_0$ laying outside (inside) the unit circle, there is another pole inside (outside) the unit circle.
-So, if there are no roots on the unit circle, $N$ roots lay outside and $N$ inside of it.
+It is easy to show that if $z_0$ is a root for the polynomial $(z_0^*)^{-1}$ is also a root: for every root laying outside the unit circle there will be another root inside of it and vice-versa. \\ 
+These properties allow us to rewrite the Fourier expansion (\ref{eq:zExp}) as \cite{1975STIN...7714318B}:
+\begin{equation}\label{eq:MESApsd_appendix}
+    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^z\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}
+\end{equation}
+with $a_0 = 1$.
+The vector obtained as $(1, a_1, \dots, a_N)$ is known as the prediction error filter.
+The power spectral density $S(f)$ is uniquely determined if both the prediction error filter and $P_N$ coefficients are computed.
 \\
-These properties allow us to rewrite the Fourier expansion (\ref{eq:zExp}) as \cite{1975STIN...7714318B}
-\begin{equation}\nonumber 
-    S(f)^{-1} = \frac{1}{P_N \Delta t} \left(1 + \sum_{s}^N a_s z^s\right)\left(1 + \sum_s^N  a^*_s z^{-s}\right),
-\end{equation}
-or, defining $a_0 = 1$,
-\begin{equation}\label{eq:MESApsd_appendix}
-    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^z\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}. 
-\end{equation}
-The MESA expression for the power spectral density admits a non-linear, all pole representation when evaluated in $z = e^{-\imath 2 \pi f \Delta t}$.
+%The MESA expression for the power spectral density admits a non-linear, all pole representation when evaluated in $z = e^{-\imath 2 \pi f \Delta t}$.
 All the roots of the first polynomial can be chosen to lay outside the unit circle, so that all the roots of the second polynomial will lay inside of it \cite{1975STIN...7714318B}.
-If one is able to compute the values of the coefficients $a_i$ the PSD is uniquely determined from the previous equation. \\ 
-To compute the $a_i$-s, we substitute in equation
-(\ref{eq:MaxConstraint}) with the above representation for $S(f)$ in terms of the $a_i$ coefficients.
-Changing  variable from $f$ to $z(f) = e^{-\imath 2 \pi f \Delta t}$, the integral (\ref{eq:MaxConstraint}) becomes a contour integral in the unit circle centered at the origin of the complex plane:
+This is equivalent to (wide-sense) stationarity of the observed process.
+%If one is able to compute the values of the coefficients $a_i$ the PSD is uniquely determined from the previous equation. \\ 
+To compute the $a_i$-s it is convenient to rewerite the equation 
+(\ref{eq:MaxConstraint}) in terms of the Laurent Polynomial and then integrating over $z$. In this way the equation becomes:
 \begin{equation}
@@
 \end{equation}
-Sending $s \to s - r$, multiplying by $a^*_s$ and summing over all possible values for $s$ the equation becomes 
-\begin{align} \nonumber 
-    \sum_{s = 0}^N a_s \bar r_{s - r} &= \frac{P_N}{2 \pi \imath}\oint \frac{z^{r - 1} \sum_{s = 0}^N a^*s z^{-s}}{\sum_{s = 0}^N a_s z^s \sum_{s = 0}^N a^*s z^{-s}} dz\\
-    & = \frac{P_N}{2 \pi \imath}\oint \frac{z^{r -1}}{\sum_{s = 0}^N a_s z^s}dz\label{eq:ErFilter}
-\end{align}
-Since all the zeros of the denominator lay outside the unit circle, the contour integral is 0.
-The only exception is for $r = 0$ with a residual in the origin of the axis.
-For $r = 0$ the integral can be computed via the Cauchy integral formula and (\ref{eq:ErFilter}) becomes:
+Substituting $s \to s - r$, multiplying by $a^*_s$ and summing over $s$ the previous equation becomes 
+\begin{equation}
+    \sum_{s = 0}^N a_s \bar r_{s - r} = \frac{P_N}{2 \pi \imath}\oint \frac{z^{r -1}}{\sum_{s = 0}^N a_s z^s}dz.\label{eq:ErFilter}
+\end{equation}
+From now on only wide-sense stationary processes will be considered: for such a process all poles lay outside the unit circle so that the previous integral can be easily computed obtaining the following results:
 \begin{align}\label{eq:errorFilter1}
@@
 \end{align}
-This kind of equations are solved by the use of the Levinson recursion \cite{doi:10.1002/sapm1946251261}
+The $N$ equations obtained for $r = 0$ can be solved via the Lenvoinson-Durbin recursion \cite{doi:10.1002/sapm1946251261} and are sufficient to uniquely determine prediction error filter $\vec a$ and $P_N$.
+Defining the quantities
+\begin{align}
+\Delta_N &= \sum_{n = 0}^{N}R(N - n + 1)a_n \\ 
+c_N &= - \frac{\Delta_N}{P_N},
+\end{align} the Levinson recursion is: 
+\begin{equation} \label{eq:Levinson1}
+P_N = P_{N -1}\left(1 - \vert c_{N - 1} \vert ^2\right)
+\end{equation}
+and 
+\begin{equation} \label{eq:Levinson2}
+\begin{pmatrix}
+1 \\ a_1 \\ \vdots \\ a_{N - 1} \\ a_N
+\end{pmatrix}
+= 
+\begin{pmatrix}
+1 \\ b_1 \\ \vdots \\ b_{N -1} \\ 0
+\end{pmatrix}
++ c_{N-1}
+\begin{pmatrix}
+0 \\ b_{N -1}^* \\ \vdots \\ b^*_1 \\ 1
+\end{pmatrix}. 
+\end{equation}
+with $b$ the forward prediction error coefficients at order $N-1$. 
+The recursion can be solved for every order if the 0'th order elements are computed. Reminding that $a_0 = 1$ whatever the order the other coefficients are easily found to be: 
+\begin{equation}
+P_0 = R(0); \quad \Delta_0 = R(1); \quad c_0 = -\frac{R(1)}{R(0)}. 
+\end{equation}
+These expressions allow us to compute $\vec a$ and $P_N$ to whatever order by simply iterating \ref{eq:Levinson1} and \ref{eq:Levinson2}.
+Substituing them in equation \ref{eq:MESApsd_appendix} the problem of the estimation for the power spectral density via maximum entropy principle is solved.
+\amartini{Io lascerei questo senza tutta la descrizione esplicita della Levinson recursion, che comunque lascio sotto se vogliamo tenerla}
+
 \subsection{Levinson Recursion}

### 2021-03-02 paper/mesa.tex
some notes

--- 2021-03-02 paper/mesa.tex
@@
 with $P_m$ defined by the prediction error filter equation \sschmidt{What is this equation?}.
+\wdp{comment on N that goes to infinity, just minimising the variance of the process}
 \item \textbf{Criterion Autoregressive Transfer function (CAT)}
@@
 The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}.
-With the technique of asymptotic integration, the loss function is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}:
+With the technique of asymptotic integration, the loss function is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}, \wdp{please check this equation}:
    \begin{align}

### 2021-03-03 paper/mesa.tex
Some figs removed

--- 2021-03-02 paper/mesa.tex
@@
 the relative error $r_i$, averaged over all frequencies:
-\begin{equation}\label{eq:freq_erorr}
+\begin{equation}\label{eq:freq_error}
 r_i = \frac{1}{Ny}\sum_{f_j=0}^{f_j = Ny} \frac{|S_i(f_j) - S(f_j)|}{S(f_j)}
@@
 %In section~\ref{sec:LIGO_validation}, we will perform the analysis again using the Advanced LIGO design sensitivity PSD \cite{Ligo}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width = \linewidth]{Images/Noise and PSD/NormalPSDautocorr.pdf}
-	\caption{Synthetic spectrum (left) and relative autocorrelation function (right) used to simulate the data to assess the performance of the three loss functions considered in Sec.~\ref{sec:arp_validation}\sschmidt{This should be done with the same style as the others}.}
-	\label{fig:autocorr}
-\end{figure}
-
-\begin{figure}
-    \centering
-        \includegraphics[width = \linewidth]{Images/NormalPSD/NormalNoise.pdf}
-    \caption{Noise in both time and frequency domain.
-Frequency noise behaviour is representative of the chosen spectrum, having its maximum amplitude around $f = 2.5 au$ \sschmidt{Togliamo questa figura?}\wdp{direi di si}}
-    \label{fig:noise}
-\end{figure}
 
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/error_length_contour.pdf}
-	\caption{Relative error $r$ (as in Eq.~(\ref{eq:freq_error})) vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with gaussian PSD given in Fig.~\ref{fig:autocorr}.
+	\caption{Relative error $r$ (as in eq.~(\ref{eq:freq_error})) vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with gaussian PSD given in Fig.~\ref{fig:autocorr}.
 Different colours refers to different choices for the loss function: in red $\mathcal{L}_{\rm FPE}(m)$, in green $\mathcal{L}_{\rm OBD}(m)$ and in black $\mathcal{L}_{\rm CAT}(m)$.
@@
 \subsection{Choosing a loss function - LIGO Spectrum} \label{sec:LIGO_validation}
-\begin{figure}
-    \centering
-    \includegraphics[width = \linewidth]{Images/LIGOsimulate/LigoPSDoriginal.pdf}
-    \caption{Original power spectral density from Ligo \sschmidt{Remove this?? It's already in the other figure}}
-    \label{fig:LigoPSDOriginal}
-\end{figure}
-In this subsection we concentrate again on the reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve see fig.~\ref{fig:LigoPSDOriginal}.
+In this subsection we concentrate again on the reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve see fig.~\ref{fig:ligospectrum}.
 For this analysis we will study properties on $500$ time series arrays of $40960$ points each. We set a sampling rate of $\SI{2048}{Hz}$ (this is $\SI{20}{s} of data$).
@@
      \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/compare_estimates.pdf}
-      \caption{Average spectrum for the three different loss function as compared with the ``true" PSD. The average is computed with $1000$ realization of a $3000$ points long time series.}
+      \caption{Average spectrum for the three different loss function as compared with the ``true" PSD. The average is computed with $500$ realization of a $40960$ points long time series.}
        \label{fig:ligospectrum}
@@
 In figure \ref{fig:ligo1peak} and \ref{fig:ligo2peak}, we plot a details od fig.~\ref{fig:ligospectrum} around the two peaks of the spectrum
-In figure~\ref{fig:LigoOrderError}, we report, as above, a scatter plot with the frequency averaged error (eq.~\ref{eq:freq_erorr}) and the length of the autoregressive model obtained with each loss function.
+In figure~\ref{fig:LigoOrderError}, we report, as above, a scatter plot with the frequency averaged error (eq.~\ref{eq:freq_error}) and the length of the autoregressive model obtained with each loss function.
 
@@
     \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/error_length_contour.pdf}
-    \caption{For each of the $500$ independent realization of the time series, we plot the relative error $r$ (as in eq.~\ref{eq:freq_erorr}) against the length $m$ of autoregressive process.
+    \caption{For each of the $500$ independent realization of the time series, we plot the relative error $r$ (as in eq.~\ref{eq:freq_error}) against the length $m$ of autoregressive process.
 The time series are randomly drawn with a the analytical LIGO PSD in fig.`\ref{fig:ligospectrum}.

### 2021-03-03 memspectrum.py
Added fixed option

@@ -45,13 +45,13 @@ class optimizer:
     def __init__(self, method):
         """
         Implements various method to choose the best recursive order for Burg's Algorithm
-        Avilable methods are "FPE", "OBD", "CAT", "AIC". The most representative order is 
+        Avilable methods are "FPE", "OBD", "CAT", "AIC", "Fixed". The most representative order is 
         chosen to be the one that minimized the related function. 
         Parameters
         ----------
         method : 'str'
             Selects the method to be used to estimate the best order between "FPE",
-            "OBD", "CAT", "AIC"
+            "OBD", "CAT", "AIC", "Fixed"
 
 
         """
@@ -367,7 +367,7 @@ class MESA(object):
         """
         Computes the power spectral density of the attribute data for the class
         using standard Burg method recursive and a Faster (but less stable) version. 
-        Default is Fast. 
+        Default is Fast.
 
         Parameters
         ----------
@@ -381,7 +381,8 @@ class MESA(object):
         optimisation_method: 'str'     
             Method used to select the best recursive order. The order is chosen
             minimizing the corresponding method. 
-            Available methods are "FPE", "OBD", "CAT", "AIC".
+            Available methods are "FPE", "OBD", "CAT", "AIC", "Fixed".
+            If optimisation_method is "Fixed", the autoregressive order is always set to m, without looking for a minimum.
             Deafult is "FPE".   
         
         method: 'str'                  
@@ -396,6 +397,7 @@ class MESA(object):
             Default is True. Breaks the iteration if there is no new global 
             maximum after 100 iterations. 
             Recommended for every optimisation method but CAT.
+            Has no effect with "Fixed" optimizer.
 
         Returns
         -------
@@ -423,9 +425,12 @@ class MESA(object):
         self.regularisation = regularisation
         self.early_stop = early_stop
         if m is None:
-            self.mmax = int(2*self.N/np.log(2.*self.N))
+            self.mmax = int(2*self.N/np.log(2.*self.N)) +1
         else:
-            self.mmax = m
+            self.mmax = m + 1
+        
+        if optimisation_method == 'Fixed':
+            self.early_stop = False
            
         if method.lower() == "fast":
             self._method = self._FastBurg

### 2021-03-03 scripts/optimisers_comparison.py
Ligo comparison on peaks added

@@ -14,12 +14,13 @@ import scipy.stats
 import matplotlib.pyplot as plt
 from style_sheet import init_plotting
 from scipy.interpolate import interp1d 
+import matplotlib.ticker as ticker
 
 def relative_error(real, estimate): 
     return np.abs(real - estimate) / real 
 
 PSD = 'ligo' #normal or ligo 
-save = True
+save = False
 savedir = '../paper/Images/optimisers_comparison/' + PSD + '/'
 simulate_data = False 
 methods = ['FPE', 'CAT', 'OBD']
@@ -77,6 +78,7 @@ elif not simulate_data:
     if PSD.lower() == 'ligo': 
         s = np.load('plot_data/ligo_spectra.npy')
         o = np.load('plot_data/ligo_optimisers.npy')
+    
         orde = np.load('plot_data/ligo_orders.npy')
     if PSD.lower() == 'normal': 
         s = np.load('plot_data/normal_spectra.npy')
@@ -98,9 +100,11 @@ ord_fig, ord_ax = plt.subplots()            #Error VS order for each method
 err_fig, err_ax = plt.subplots(1, 3)        #Error hist for each method 
 aro_fig, aro_ax = plt.subplots(1, 3)        #AR order estimate hist for each method
 fig4, ax4 = plt.subplots()                  #Plot with every spectral estimate    
-
+if PSD == 'ligo': fig5, ax5 = plt.subplots(1, 2)      
+            #Plot with zoom over two peaks for Ligo Spectrum 
 ord_ax.set_xlabel('Filter Length estimate'); ord_ax.set_ylabel('Frequency averaged error')
 ax4.set_ylabel(y); ax4.set_xlabel(r"$f(Hz)$")
+ax5[0].set_ylabel(y); ax5[0].set_xlabel(r"$f(Hz)$"); ax5[1].set_xlabel(r"$f(Hz)$")
 err_ax[0].set_ylabel('Percentage')
 aro_ax[0].set_ylabel('Percentage')
 
@@ -155,7 +159,9 @@ for i, method in enumerate(methods):
     aro_ax[i].hist(orders[method], bins = 70, density = True, color = 'k', histtype = 'step')
 
     ax4.loglog(frequency[:-1], median[method], color = colors[method], label = method)
-    
+    ax5[0].loglog(frequency[:-1], median[method], color = colors[method])
[…]

### 2021-03-03 paper/mesa.tex
Appendix completed

--- 2021-03-02 paper/mesa.tex
@@
 with $P_m$ defined by the prediction error filter equation \sschmidt{What is this equation?}.
-\wdp{comment on N that goes to infinity, just minimising the variance of the process}
 \item \textbf{Criterion Autoregressive Transfer function (CAT)}
@@
 The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}.
-With the technique of asymptotic integration, the loss function is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}, \wdp{please check this equation}:
+With the technique of asymptotic integration, the loss function is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}:
    \begin{align}
@@
 \subsection{MESA solution}
-The proposed solution follows the original demonstration due to Burg \cite{burg1975maximum}.
-The constraints (\ref{eq:MaxConstraint}) can be directly implemented by substituting $S(f)$ as the Fourier transform of the autocorrelation function
+To derive the expression for the MAXENT spectral estimator the approach proposed by Burg \cite{burg1975maximum} will be followed.
+The constraints \ref{eq:MaxConstraint} are not implemented via the standard Lagrange Multipliers approach while the entropy gain \ref{eq:EntropyGain} will be re-written by explicitly writing $S(f)$ as the Fourier Transform of the sample autocorrelation function:
 \begin{equation}
-    S(f) = \frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{- \imath 2 \pi n \Delta t}
+    S(f) = \frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{- \imath 2 \pi n \Delta t},
 \end{equation}
-in equation \ref{eq:EntropyGain}. In this way the variational problem is reduced to the maximization of: 
+and becomes
 \begin{equation}
-    \int_{-Ny}^{Ny}  
+    \Delta H = \int_{-Ny}^{Ny}  
     \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r(n\Delta t) e^{-\imath 2 \pi f n \Delta t} 
@@
 \end{equation}
-Since the autocorrelation function can only be computed for a finite time interval $n \in [-N, N]$, the requirement for our estimate to be independent from the unkown values of the autocorrelation function can be implemented as:
+Deriving with respect to the autocorrelation function $r_s$
+\begin{equation}
+      \frac{\delta H}{\delta \bar r_s} = \frac{1}{2Ny}\int_{-Ny}^{Ny} S(f)^{-1}e^{-\imath 2 \pi f s \Delta t } df = \lambda_s. 
+\end{equation} 
+its easily shown that $S(f)^{-1}$ can be written as a Fourier Expansion in terms of some $\lambda$s coefficients.
+THe determination of the values for the $\lambda$s uniquely solve the problem of power-spectral density estimation.
+Some properties for the coefficients can be worked out easily.
+First, since $S(f)$ is positive definite the $\lambda$s coefficient show the property
+\begin{equation}
+\nonumber 
+\lambda_s = \lambda_{-s}^*. 
+\end{equation}
+The second property is obtained considering that the autocorrelation function can only be computed for a finite time interval $n \in [-N, N]$, the requirement for our estimate to be independent from the unkown values of the autocorrelation function can be implemented as:
 \begin{equation}\nonumber 
-    \frac{\delta H}{\delta \bar r_s} = 0 \text{ for } \vert s \vert > N.
+    \frac{\delta H}{\delta \bar r_s} = 0 \text{ for } \vert s \vert > N,
 \end{equation}
-By explicitly computing the derivative: 
+that means 
 \begin{equation}
-      \frac{\delta H}{\delta \bar r_s} = \frac{1}{2Ny}\int_{-Ny}^{Ny} S(f)^{-1}e^{-\imath 2 \pi f s \Delta t } df = \lambda_s
+\nonumber 
+\lambda_s = 0 \text{ for } \vert s \vert > N. 
 \end{equation}
-is easily seen that $S(f)$ can be expressed via a Fourier Series
+From the previous equations is easily seen that $S(f)$ can be expressed via a Fourier Series 
 \begin{equation}\label{eq:PSDconstraint}
-    S(f)^{-1} = \sum_{s = -\infty}^\infty \lambda_s e^{-\imath 2 \pi f s \Delta t}
+    S(f)^{-1} = \sum_{s = -N}^N \lambda_s e^{-\imath 2 \pi f s \Delta t}.
 \end{equation}
-with the properties $\lambda_s = 0$ for $\vert s \vert > N$ and $\lambda_s = \lambda_{-s}^*$.  
 Defining $z = e^{-\imath 2 \pi f \Delta t}$ the previous Fourier expansion becomes a Laurent Polynomial in $z$: 
@@
     \label{eq:zExp}
-    S(f)^{-1} = \lambda_0 + \sum_{s = i}^N \lambda_s z^s + \sum_{s = 1}^N \lambda^*_s z^{-s}.
+    S(f)^{-1} = \lambda_0 + \sum_{s = 1}^N \lambda_s z^s + \sum_{s = 1}^N \lambda^*_s z^{-s}.
 \end{equation}
-It is easy to show that if $z_0$ is a root for the polynomial $(z_0^*)^{-1}$ is also a root: for every root laying outside the unit circle there will be another root inside of it and vice-versa. \\ 
+It is easy to show that if $z_0$ is a root for the polynomial $(z_0^*)^{-1}$ is also a root: for every root laying outside the unit circle there will be another root inside of it and vice-versa.
 These properties allow us to rewrite the Fourier expansion (\ref{eq:zExp}) as \cite{1975STIN...7714318B}:
@@
 \\
-%The MESA expression for the power spectral density admits a non-linear, all pole representation when evaluated in $z = e^{-\imath 2 \pi f \Delta t}$.
-All the roots of the first polynomial can be chosen to lay outside the unit circle, so that all the roots of the second polynomial will lay inside of it \cite{1975STIN...7714318B}.
-This is equivalent to (wide-sense) stationarity of the observed process.
-%If one is able to compute the values of the coefficients $a_i$ the PSD is uniquely determined from the previous equation. \\ 
 To compute the $a_i$-s it is convenient to rewerite the equation 
@@
 \end{equation}
-From now on only wide-sense stationary processes will be considered: for such a process all poles lay outside the unit circle so that the previous integral can be easily computed obtaining the following results:
+From now on only wide-sense stationary processes will be considered: for such a process all poles lay outside the unit circle so that the previous integral can be easily computed obtaining the following, well known, equations:
 \begin{align}\label{eq:errorFilter1}
@@
 \end{align}
-The $N$ equations obtained for $r = 0$ can be solved via the Lenvoinson-Durbin recursion \cite{doi:10.1002/sapm1946251261} and are sufficient to uniquely determine prediction error filter $\vec a$ and $P_N$.
-Defining the quantities
+The solution of the previous equations fully determine the functional form of the power spectral density estimator \ref{eq:MESApsd_appendix}, and the method for solving such equations is the Levinson-Durbin recursion \cite{doi:10.1002/sapm1946251261}.
+The recursion can be easily schemitized as follows:
+After defining the quantities  
 \begin{align}
@@
 with $b$ the forward prediction error coefficients at order $N-1$. 
-The recursion can be solved for every order if the 0'th order elements are computed. Reminding that $a_0 = 1$ whatever the order the other coefficients are easily found to be: 
+The 0'th order element can be easily initialized reminding that $a_0 = 1$ whatever the order the other coefficients. $P_0$ can be determined from \ref{eq:errorFilter1} and its value is: 
 \begin{equation}
-P_0 = R(0); \quad \Delta_0 = R(1); \quad c_0 = -\frac{R(1)}{R(0)}. 
+P_0 = R(0),
+\end{equation}
+ while $\Delta_0$ and $c_0$ are uniquely determined from their definitions and are
+\begin{equation}
+\Delta_0 = R(1); \quad c_0 = -\frac{R(1)}{R(0)}. 
 \end{equation}
@@
 Substituing them in equation \ref{eq:MESApsd_appendix} the problem of the estimation for the power spectral density via maximum entropy principle is solved.
-\amartini{Io lascerei questo senza tutta la descrizione esplicita della Levinson recursion, che comunque lascio sotto se vogliamo tenerla}
-
-\subsection{Levinson Recursion}
-The equations (\ref{eq:errorFilter1}) and (\ref{eq:errorFilter2}) can be rewritten as a Matrix equation of the form
-\begin{equation}\label{eq:predictionErrorFilter}
-    \begin{pmatrix}
-    \bar r_0 &  \bar r_{-1} & \dots & \bar r_{-N}\\
-    \bar r_{1} & \bar r_ 0 & \dots & \bar r_{1 - N} \\ 
-    \bar r_{2} & \bar r_{1} & \dots & \bar r_{2 - N}
-    \\
-    \vdots & \vdots & \ddots & \vdots  \\
-    \bar r_{N} & \bar r_{N - 1} & \dots  &\bar r_0
-    \end{pmatrix}
-    \begin{pmatrix}
-    1 \\   a_1 \\   a_ 2 \\  \vdots \\   a_N
-    \end{pmatrix} = 
-    \begin{pmatrix}
-    P_N \\   0 \\  0 \\ \vdots  \\   0
-    \end{pmatrix}
-\end{equation}
-know as the prediction error filter equation.
-It is a matrix equation involing the Toeplitz Matrix $R_{ij}$ and the vector $\vec a$, known as the forward prediction error filter.
-The approach for solving the equations is the Levinson - Durbin recursion.
-\\
-Our goal is to obtain the $N$-th order equation (\ref{eq:predictionErrorFilter}) from the $N-1$-th order equation: 
-\begin{equation}\label{eq:Forward}
-    \begin{pmatrix}
-    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N}\\
-    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} \\ 
-    \vdots & \vdots & \ddots & \vdots \\ 
-    \bar r_{N - 1} & \bar r_{N -2} & \dots & \bar r_0
-    \end{pmatrix} 
-    \begin{pmatrix}
-    1  \\   b_1  \\  \vdots \\   b_{N - 1}
-    \end{pmatrix} =
-    \begin{pmatrix}
-    P_{N -1} \\  0 \\ \vdots \\ 0
-    \end{pmatrix} .
-\end{equation}
-This equation is not sufficient to solve the recursion.
-Since the autocorrelation matrix is self-adjoint, from equations (\ref{eq:errorFilter1}) and (\ref{eq:errorFilter2}) and noting that the first and the last row of the autocorrelation matrix are respectively the complex conjugate reverse, is easy to show that
-\begin{equation}\label{eq:Backward}
-     \begin{pmatrix}
-    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N}\\
-    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} \\ 
-    \vdots & \vdots & \ddots & \vdots \\ 
-    \bar r_{N - 1} & \bar r_{N -2} & \dots & \bar r_0
-    \end{pmatrix} 
-    \begin{pmatrix}
-    b^*_{N-1} \\ b^*_{N- 2} \\\vdots \\ 1
-    \end{pmatrix} 
-    = 
-    \begin{pmatrix}
-    0 \\ 0 \\ \vdots \\ P_{N-1}
-    \end{pmatrix} 
-\end{equation}
-holds. The vector obtained reflecting b and taking its complex conjugate is known as ``backward prediction error". \\ 
-Stepping from order $N - 1$ to order $N$, the autocorrelation matrix grows from size $(N,N)$ to size $(N + 1, N + 1)$, this implies that a new component $b_n$ for the prediction error filter is needed.
-If we take it to be $b_N = 0$, this choice does not modify the first $N$ component of the right side of equation (\ref{eq:Forward}) (or the last $N$ components of (\ref{eq:Backward})), i.e.:
-
-\begin{equation}\label{eq:OrderN}
-    \begin{pmatrix}
-    \bar r_0 & \bar r_{-1} & \dots &\bar r_{1 - N} & \bar r_{- N}\\
-    \bar r_{1} & \bar r_0 & \dots & \bar r_{2 - N} & \bar r_{1 - N}\\ 
-    \vdots & \vdots & \ddots & \vdots & \vdots \\ 
-    \bar r_{N -1} & \bar r_{N-2} & \dots & \bar r_0 & \bar r_{-1}
-    \\
-    \bar r_{N } & \bar r_{N -1} & \dots & \bar r_1 & r_0 
-    \end{pmatrix} 
-    \begin{pmatrix}
-    1  \\   b_1  \\  \vdots \\   b_{N - 1} \\ 0 
-    \end{pmatrix} =
-    \begin{pmatrix}
-    P_{N -1} \\  0 \\ \vdots \\ 0 \\ \Delta_N
-    \end{pmatrix} .
-\end{equation}
-with 
-\begin{equation}
-    \Delta_N = \sum_{n=0}^N \bar r_{N - n}b_n.
-\end{equation}
-Adding the previous expansion for both forward and backward equations, one obtains 
-\begin{equation}\nonumber 
-        \begin{pmatrix}
-    \bar r_0 & \bar r_{-1} & \dots & \bar r_{- N}\\
-    \bar r_{1} & \bar r_0 & \dots & \bar r_{1 - N}\\ 
-    \vdots & \vdots & \ddots & \vdots \\ 
-    \bar r_{N -1} & \bar r_{N-2} & \dots &  \bar r_{-1}
-    \\
-    \bar r_{N } & \bar r_{N -1} & \dots  & r_0 
-    \end{pmatrix} 
-    \begin{pmatrix}
-    1  \\   b_1  \\  \vdots \\   b_{N - 1} \\ 0 
-    \end{pmatrix} 
-    + c_N 
-    \begin{pmatrix}
-    0 \\ b^*_{N -1} \\ \vdots  \\ b_1 \\ 1 
-    \end{pmatrix}
-    =
-    \begin{pmatrix}
-    P_{N -1} \\  0 \\ \vdots \\ 0 \\ \Delta_N
-    \end{pmatrix} .
-    + c_N
-    \begin{pmatrix}
-    \Delta^*_N \\ 0 \\\vdots \\ 0 \\ P_{N-1}
-    \end{pmatrix}.
-\end{equation}
-With the reflection coefficient $c_N$ to be computed. Requiring the previous equation to be equivalent to \ref{eq:predictionErrorFilter}, one obtains
-\begin{equation}\nonumber 
-\begin{cases}
-    P_{N - 1} + c_N\Delta^*_N = P_N \\
-    c_N P_{N - 1} + \Delta_N = 0
-\end{cases}
-\end{equation}
-which can be easily solved for $c_N$ and $P_N$ as 
-\begin{equation}
-    c_N = -\frac{\Delta_N}{P_{N-1}}; \quad \text{ and } P_N = P_{N - 1}\left(1 - \vert c_N \vert ^2 \right). 
-\end{equation}
-Having computed the value for the reflection coefficient, the $N$-th order forward prediction error filter is found to be 
-\begin{equation}
-    a_i = b_i + c_N b^*_{N - i}.
-\end{equation}
-To obtain the solution for the $N$-th order, one starts $\bar r_0 = P_0$ and recursively computes the new values until the desired order is computed. 
-This problem is often solved numerically with Burg's Algorithm \cite{Vos}.
+While Burg's method for spectral analysis is solved via Levinson recursion, another faster recursion \cite{Vos} is implemented in memspectrum package.
 

### 2021-03-03 paper/mesa.tex
Changed peaks figure

@@
 In figure \ref{fig:ligospectrum}, we report the original spectrum and the ensemble averaged spectrum obtained with the three methods (no confidence level).
-In figure \ref{fig:ligo1peak} and \ref{fig:ligo2peak}, we plot a details od fig.~\ref{fig:ligospectrum} around the two peaks of the spectrum
+In figure \ref{fig:ligoPeaks} we plot a details od fig.~\ref{fig:ligospectrum} around the two peaks of the spectrum
 In figure~\ref{fig:LigoOrderError}, we report, as above, a scatter plot with the frequency averaged error (eq.~\ref{eq:freq_error}) and the length of the autoregressive model obtained with each loss function.
@@
 The conclusions are very similar to those above. When averaged on multiple realization, CAT loss function is able to capture the spectrum very precisely.
-In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them (figs.\ref{fig:ligo1peak} and \ref{fig:ligo2peak}).
+In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them (figs.\ref{fig:ligoPeaks}).
 However, the computation on a single noise realization yields larger error, as manifest in fig.~\ref{fig:LigoOrderError}.
@@
 \begin{figure}[t]
-        \includegraphics[width = \linewidth]{Images/LIGOsimulate/1stPeakComparison.pdf}
-        \caption{1st peak of the spectrum and it reconstruction with every optimizer (left) and associated percentage error (right) \sschmidt{This should be done with the same style as the others}}
-        \label{fig:ligo1peak}
-\end{figure}
-\begin{figure}[t]
-    \centering
-        \includegraphics[width = \linewidth]{Images/LIGOsimulate/2ndPeakComparison.pdf}
-        \caption{2nd peak of the spectrum, Relative error around 2nd peak of the spectrum \sschmidt{This should be done with the same style as the others}}
-        \label{fig:ligo2peak}
+        \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/compare_estimates_peaks.pdf}
+        \caption{Details of peaks of the spectrum and their reconstruction with every optimizer}
+        \label{fig:ligoPeaks}
 \end{figure}

### 2021-03-03 memspectrum.py
Change in init to have a nice load file

@@ -176,23 +176,23 @@ class MESA(object):
     """
     Class the implement the reproduces the Maximum Entropy Spectrum of a given 
     time-series. 
-    
-    init: data: `np.ndarray` shape (N,)
-    solve(): 
     """
-    def __init__(self, *args, **kwargs):
+    def __init__(self,filename = None, *args, **kwargs):
         """ 
         Class that implements Burg method to estimate power spectral densitiy of
         time series. 
         
         Parameters
         ----------
-        data: 'np.ndarray'       
-            Time series with power spectral density to be computed 
+        filename: 'str'       
+            Name of file from which the model is loaded.
+            If None, the model is not initialized
         """
         self.P = None
         self.a_k = None #If a_k and P are None, the model is not already fitted
         self.optimization = None
+        if isinstance(filename,str):
+        	self.load(filename)
 
     def save(self,filename):
         """

### 2021-03-03 scripts/autoregressive_plots.py
Added autoregressive_plots

@@ -1,0 +1,90 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+import sys
+sys.path.insert(0,'..')
+import memspectrum
+
+
+
+N_process = 10 #10 simulated autoregressive processes
+N_data = 1000 #number of timesteps for each autoregressive process
+p_max = 100
+
+save_dir = 'arp_data/'
+
+srate = 1.
+
+load = True
+
+	#initializing the models
+true_arp_list = []
+rec_MESA_list = []
+data_list = []
+loss_functions = ['FPE', 'CAT', 'OBD']
+
+for i in range(N_process):
+
+	if load:
+		true_arp_list.append(memspectrum.MESA(save_dir+'arp_model_{}'.format(i)))
+		data_list.append(np.loadtxt(save_dir+'data_{}'.format(i)))
+		rec_MESA_list.append( (memspectrum.MESA(save_dir+'MESA_model_{}_{}'.format(loss_functions[0],i)), 
+			memspectrum.MESA(save_dir+'MESA_model_{}_{}'.format(loss_functions[1],i)),
+			memspectrum.MESA(save_dir+'MESA_model_{}_{}'.format(loss_functions[2],i)) )
+			)
+	else:
+		true_arp_list.append(memspectrum.MESA())
+			#getting p
+		p = np.random.choice(range(2,p_max))
+		
+		a_k = np.random.normal(0, 0.2, size = (p,))
+		a_k = np.multiply(a_k,np.exp(-0.1*np.arange(0,len(a_k),1)))
+		P = np.random.exponential(0.05)
+
+		print('Process {} has p = {} and P, a_k = {}, {}'.format(i,p,P,a_k))
+		
+			#initializing the model
+		true_arp_list[-1].N = N_data
+		true_arp_list[-1].P = P
+		true_arp_list[-1].a_k = a_k
+		
+		true_arp_list[-1].save(save_dir+'arp_model_{}'.format(i))
+		
+			#forecasting
+		data_0 = np.random.rand(p) #initial data
+		burn_steps = 100000
+		
+		data = true_arp_list[-1].forecast(data_0, burn_steps+N_data)[0,:] #(burn_steps+N_data,)
+		
+		data_list.append(data[-N_data:])
+		#print("Time series: ", data_list[-1])
+		
+		#plt.plot(data_list[-1])
+		#plt.show()
[…]

### 2021-03-03 memspectrum.py
Added countor plots

@@ -425,9 +425,9 @@ class MESA(object):
         self.regularisation = regularisation
         self.early_stop = early_stop
         if m is None:
-            self.mmax = int(2*self.N/np.log(2.*self.N)) +1
+            self.mmax = int(2*self.N/np.log(2.*self.N))
         else:
-            self.mmax = m + 1
+            self.mmax = m
         
         if optimisation_method == 'Fixed':
             self.early_stop = False
@@ -679,7 +679,7 @@ class MESA(object):
         p : 'int'
             Order of the autoregressive process that define the PSD
         """
-        return self.a_k.size - 1
+        return self.a_k.size - 1 #why -1???
     
     def forecast(self, data, length, number_of_simulations = 1, P = None, include_data = False, verbose = False):
         """

### 2021-03-03 scripts/autoregressive_plots.py

@@ -1,5 +1,6 @@
 import numpy as np
 import matplotlib.pyplot as plt
+from style_sheet import init_plotting
 
 import sys
 sys.path.insert(0,'..')
@@ -7,15 +8,16 @@ import memspectrum
 
 
 
-N_process = 10 #10 simulated autoregressive processes
-N_data = 1000 #number of timesteps for each autoregressive process
-p_max = 100
+
+N_process = 50 #10 simulated autoregressive processes
+N_data = 2000 #number of timesteps for each autoregressive process
+p_max = 30
 
 save_dir = 'arp_data/'
-
-srate = 1.
+plot_dir = '../paper/Images/arp_errors/'
 
 load = True
+plot = True
 
 	#initializing the models
 true_arp_list = []
@@ -37,7 +39,7 @@ for i in range(N_process):
 			#getting p
 		p = np.random.choice(range(2,p_max))
 		
-		a_k = np.random.normal(0, 0.2, size = (p,))
+		a_k = np.random.normal(0, 0.1, size = (p,))
 		a_k = np.multiply(a_k,np.exp(-0.1*np.arange(0,len(a_k),1)))
 		P = np.random.exponential(0.05)
 
@@ -75,8 +77,46 @@ for i in range(N_process):
 
 
 	#doing plots
+if plot:
+	N_p_plot = 1 #N of series to be shown in the err vs p plot
+	colors = {'FPE': 'r', 'CAT': 'k', 'OBD': 'green'}
+	fig_p = init_plotting()
+	ax_p = fig_p.gca()
+	
+	fig_scatter = init_plotting()
+	ax_scatter = fig_scatter.gca()
+	
+		#computing erorrs
+	diff_p = np.zeros((N_process,3))
+	diff_ak = np.zeros((N_process,N_data,3))
+	for i in range(N_process):
+		for j, l in enumerate(loss_functions):
+			diff_p[i,j] = - true_arp_list[i].get_p() + rec_MESA_list[i][j].get_p()
+			#print(j, true_arp_list[i].get_p() , rec_MESA_list[i][j].get_p(),diff_p[i,j])
+			ak_true = np.concatenate([true_arp_list[i].a_k, np.zeros((N_data-len(true_arp_list[i].a_k),))]) #(N_data,)
[…]

### 2021-03-03 scripts/autoregressive_plots.py
Added countor plots -bis

@@ -106,6 +106,7 @@ if plot:
 	ax_scatter.legend(loc = 'upper right')
 	ax_p.legend()
 	ax_p.set_xlim([0,200])
+	ax_p.set_ylim([-0.15,0.15])
 	ax_p.set_xlabel("p")
 	ax_p.set_ylabel("Difference in a_k")
 	ax_scatter.set_xlabel("p - p_true")

### 2021-03-05 paper/mesa.tex
another big chunk of edits

@@
 p&\left((\tilde{x}_1, \ldots, \tilde{x}_n)\vert I\right) = \nonumber \\
-    &\frac{1}{\left(2 \pi \det S\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i}(\tilde{x}_i^2S^{-1}_{i} \right)\,,
+    &\frac{1}{\left(2 \pi \det S\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i}\tilde{x}_i^2S^{-1}_{i} \right)\,,
 \end{align}
 where the matrix $S$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
-Many readers will recognised the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\cite{woodward}
+Many readers will recognise the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\cite{woodward}
 %P. M. Woodward, D. W. Fry, and W. Higinbotham, Probability and information theory, with applications to radar; 2nd ed. (Pergamon Press, Oxford, 1964), URL http://cds.cern.ch/record/2031792.
@@
 In a real implementation of the algorithm, computing all the recursion up to $M_{max}$ can result in a significant waste of computational power: the optimal value is often $m_{opt} << M_{max}$ and, in such cases, computing all the values of $m$ until $M_{max}$ is not useful.
-In practice, we can apply an \textit{early stopping} procedure: every few iterations we look for the best order of $m_{opt}$; if this value does not change for a while, we assume that a good (local) minimum of the loss function is found and the computation is stopped.
+In practice, we can apply an \textit{early stop} procedure: every few iterations we look for the best order of $m_{opt}$; if this value does not change for a while, we assume that a good (local) minimum of the loss function is found and the computation is stopped.
 
@@
 
-\subsection{Choosing a loss function - A gaussian PSD} \label{sec:arp_validation}
+\subsection{Choice of the loss function: Gaussian PSD} \label{sec:arp_validation}
 We test the performance of the three loss functions on a random time series generated with a known power spectral density.
-We take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$, see Fig.~\ref{fig:autocorr}.
-The time series will be generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$, following the procedure outlined in ~\cite{}, and performing the inverse Fast Fourier Transform on the frequency series.
+In this first experiment, we take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$, see Fig.~\ref{fig:autocorr}.
+The time series are generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$, following the procedure outlined in~\cite{}, and performing an inverse Fast Fourier Transform on the frequency series.
+For this investigation, we generate a dataset of 1000 time series of 3000 points each, see Fig.~\ref{fig:noise} for an example of a time series in frequency and time domain.
 
-We will apply the Burg method choosing in turn each of the loss functions on an ensemble of simulated time series, 
-thus obtaining an ensemble of PSD estimates, that will aloows the statistical characterisation of the performance of each loss function.
+We then apply the Burg method choosing in turn each of the loss functions on an ensemble of simulated time series, 
+thus obtaining an ensemble of PSD estimates, that will allows the statistical characterisation of the performance of each loss function.
 The disagreement between the PSD $S_i(f)$ estimated from the $i$-th simulated time series and the target PSD $S(f)$ is measured via 
@@
 as well as the ensemble-averaged error from Eq.~(\ref{eq:ensemble_averaged_error}).
-Furthermore, Fig.~\ref{fig:optcomparison} displays the correlation between $r_i$ and the autoregressive order $m_i$ chosen for each independently drawn time series.
+Furthermore, Fig.~\ref{fig:optcomparison} displays the correlation between $r_i$ and the autoregressive 
+order $m_i$ chosen for each independently drawn time series.
 
-For this investigation, we generate a dataset of 1000 time series of 3000 points each, see Fig.~\ref{fig:noise} for an example of a time series in frequency and time domain.
 %In section~\ref{sec:LIGO_validation}, we will perform the analysis again using the Advanced LIGO design sensitivity PSD \cite{Ligo}.
@@
 \paragraph{Final Prediction Error (FPE)}
-The results of our investigation on the performance of FPE are reported in fig.~\ref{fig:FPEmean}.
-Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum; however, we note that the reconstruction is not very accurate in the low frequency region. 
+The results of our investigation on the performance of FPE are shown in Fig.~\ref{fig:FPEmean}.
+Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum; 
+however, we note that the reconstruction is not very accurate in the low frequency region. 
 Furthermore, the $90\%$ credible region is very small: this means that if we randomly take two of the reconstructed spectra, we expect their differences to be statistically small.
@@
 
-By looking at the red series in fig.~\ref{fig:optcomparison}, we note that the autoregressive orders obtained with FPE are clustered in a very small region.
-This is a desiderable property: FPE provides a stable estimation of the autoregressive order, which does not affect much the reconstruction error.
-We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length M is clustered in a region where there is no dependence of error from M.
+By looking at therRed series in Fig.~\ref{fig:optcomparison}, we note that the autoregressive orders obtained with FPE are clustered in a very small region.
+This is a desirable property: FPE, in fact, provides a stable estimation of the autoregressive order, which does not affect much the reconstruction error.
+We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length $m$ is clustered in a region where there is no dependence of error from $m$.
 \paragraph{Optimum Bayes Decision Rule (OBD)}
-The second loss function in consideration is OBD.
-The results in fig.~\ref{fig:OBDmean} show a good agreement between the average over the reconstructions and the true spectrum.
+The second loss function in consideration is OBD. As for the FPE loss function, the results in Fig.~\ref{fig:OBDmean} show a 
+good agreement between the average over the reconstructions and the true spectrum.
 Qualitatively the same behaviour of FPE is observed: a good quality reconstruction at the intermediate and high frequencies with a low $90\%$ confidence level; and a worse performance at low frequencies.
-However, when looking at the error, the disagreement of this method is found to be larger than the one obtained with FPE: in the worst case, the error can gain a factor of $2$ as compared with FPE.
+However, when looking at the error, the disagreement of this method is found to be larger than the one obtained with FPE: in the worst case, the error can be as large as a factor of $2$ when compared with FPE.
 
-When plotting the error as a function of the autoregressive length (green series in fig.~\ref{fig:optcomparison}), the values taken by the OBD cluster in a small region, indicating the stability of the reconstructed process order $M$.
-We note that on average, OBD tends to choose smaller values of $M$ with respect to FPE: this can probably explain the observed large errors.
+The error as a function of the autoregressive length (green series in Fig.~\ref{fig:optcomparison}) cluster in a small region, indicating the stability of the reconstructed process order $m$.
+We note that on average, OBD tends to choose smaller values of $m$ with respect to FPE.
 
 \paragraph{Criterion Autoregressive Transfer function (CAT)}
-The performance of CAT are shown in fig.~\ref{fig:CATmean} and it is striking that CAT differs substantially from the other two loss functions considered.
-The ensambled average PSD matches very well the ``true" PSD. This is also true even in the low frequency region, where the OBD and FPE had a poor performance.
-However, the variance of the reconstructed spectrum is quite large (much larger than for FPE and OBD), and the relative error is quite high $~10\%$ and it approximately constant over all the frequencies.
+The performance of CAT is shown in Fig.~\ref{fig:CATmean}. At a first glance CAT differs substantially from the other two loss functions considered.
+The ensemble-average PSD matches very well the underlying ``true" PSD. This is also true even in the low frequency region, 
+where both OBD and FPE had a poor performance.
+However, the variance of the reconstructed spectrum is quite large (much larger than for FPE and OBD), and the relative error is quite high, $\sim 10\%$ and it is approximately constant over all the frequencies.
 
-The reason for this behaviour, can be found in fig.~\ref{fig:optcomparison} (black series). CAT does not converge to any specific value for the order of the autoregressive process.
+The reason for this behaviour becomes clear from Fig.~\ref{fig:optcomparison} (black series). CAT does not converge 
+to any specific value for the order of the autoregressive process.
 In turn, the estimates for the order span a large range of values, hence the large variance observed.
-Figure~\ref{fig:optcomparison} also shows a strong dependence of the error on the estimated length of the filter.
-The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like outliers in different realization of a noisy timeseries, but this is also responsible of an augmented variance in the estimate, by introducing spurious peaks in the reproduction.
+Fig.~\ref{fig:optcomparison} also shows a strong dependence of the error on the estimated length of the filter.
+The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like outliers in different realization of the time-series, but this is also responsible of an increased variance in the estimate, by introducing spurious peaks in the reproduction.
 
-When averaging the diffrerent PSD estimation, the noise in each spectrum cancels.
-It is interesting that the cancelling of errors in averaging is property of gaussian noise (the process is gaussian, after all).
+When averaging the different PSD estimates, the noise in each spectrum cancels, as expected for Gaussian noise. 
 This implies, in a sense, that each estimate of the spectrum is independent of any other, as suggested by the huge variance in the residuals.
-This instability is not a good property for an estimation, so that this estimator might be found to be optimal for repeatable experiments and not for single observation of some process.
+This lack of stability is not a good property for the estimation estimation of the PSD from a single realisation of the time-series, however, thanks to the averaging out of the errors, this estimator seems optimal in the case of repeatable experiments and ensemble-averages.
 
@@
 
-In our analysis, FPE and OBD loss functions are found to behave similarly while CAT shows completely different properties.
-CAT provides an accurate average spectrum over all the frequencies and large amount of variability; in turn OBD and FPE provide a poorer average in the low frequency tails with a lower variability (FPE with less noise than OBD).
+In our analysis, the FPE and OBD loss functions are found to behave similarly while CAT shows fairly different properties.
+CAT provides an accurate average spectrum over all the frequencies at the price of a large variance; in turn OBD and FPE 
+provide a poorer average in the low frequency tails, however they also display a  smaller variance, with FPE showing the lowest.
 
-The bad behavior in the reconstruction of the tails from OBD and FPE might be due to the fact that very small values require longer filters. 
-In fact when we consider the average for all the spectra we note that short filters yield larger the error in the tails.
-OBD, which select the shortest filters, can provide an error as large as $300 \%$ at the extrema, as compared with $85\%$ error od FPE.
+The poor low-frequency reconstruction from OBD and FPE might be due to the fact that very small values require longer filters. 
+In fact, when we consider the average for all the spectra we note that short filters yield larger the error in the tails.
+OBD, which select the shortest filters, can provide an error as large as $300 \%$ at the extrema, as compared with $85\%$ error of FPE.
 In turn, CAT is $5\%$ accurate in those regions.
 
-However, we see in figure \ref{fig:optcomparison} that a single PSD estimation with FPE is still able to provide the lowest averaged error over all frequencies, while CAT can reach error error $5$ times larger.
-This is to say that CAT is the loss function that minimizes errors in the tails, FPE is found to reach best overall accuracy.
+However, we also note, see Fig.~\ref{fig:optcomparison}, that the psD inferred from a single time-series realisation with FPE is still able 
+to provide the lowest averaged error over all frequencies, while CAT can reach errors $5$ times larger.
+Hence, while CAT is the loss function that minimizes errors in the low-frequency end of the spectrum, 
+FPE obtains the best overall accuracy.
 
@@
 As in any common situation we cannot perform such averaging over different realization of the same time series, we must prefer FPE over CAT (let alone OBD, which even though qualitatively similar to FPE has worse performance).
-However, if we indeed can measure the PSD by averaging over different time series, probably using CAT as a loss function is the best choice.
-\sschmidt{This is similar to Welch: averaging over different time series to get a reliable result. Should we mention it?}
-\sschmidt{C'erano ancora molti commenti: ho sintetizzato molto. Va bene? O si vuole aggiungere altro?}
+However, if we indeed can measure the PSD by averaging over different time series, using CAT as a loss function is the best choice. In this sense
+we retain CAT to provide the best, and most similar in spirit, alternative to the commonly employed WelchPsd estimation method whenever 
+ensemble-averages are needed and justified.
 
@@
 
-\subsection{Choosing a loss function - LIGO Spectrum} \label{sec:LIGO_validation}
-In this subsection we concentrate again on the reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve see fig.~\ref{fig:ligospectrum}.
-For this analysis we will study properties on $500$ time series arrays of $40960$ points each. We set a sampling rate of $\SI{2048}{Hz}$ (this is $\SI{20}{s} of data$).
-We chose a longer dataset to capture the complicated structure of the PSD. A shorter time batch would not have allowed us to reconstruct faithfully all the peaks of the empirical PSD.
+\subsection{Choice of the loss function: LIGO Spectrum} \label{sec:LIGO_validation}
+We continue our characterisation of the various loss functions considered in this work, by investigating the 
+reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve\wdp{add ref}.
+For this analysis, we generate $500$ time series of $40960$ points each, sampled with a sampling rate of $\SI{2048}{Hz}$, hence we fix the duration
+of the time-series to $\SI{20}{s}$. The chosen length is convenient to capture fairly accurately the low-frequency features of the LIGO PSD.
 
@@
 \end{figure}
-In figure \ref{fig:ligospectrum}, we report the original spectrum and the ensemble averaged spectrum obtained with the three methods (no confidence level).
-In figure \ref{fig:ligoPeaks} we plot a details od fig.~\ref{fig:ligospectrum} around the two peaks of the spectrum
-In figure~\ref{fig:LigoOrderError}, we report, as above, a scatter plot with the frequency averaged error (eq.~\ref{eq:freq_error}) and the length of the autoregressive model obtained with each loss function.
+Our findings are summarised in Figs.~\ref{fig:ligospectrum} and ~\ref{fig:LigoOrderError}. The former shows the original spectrum (dashed line)
+and the ensemble-averaged reconstructed PSD adopting the FPE (green line), OBD(red line) and CAT(black line).
+In all cases, the spectrum is well reconstructed, but with a fairly distinct behaviour at low frequency, where CAT -- as in the Gaussian case -- better captures and resolves the distinct
+spectral feature at $\sim 17$ Hz, see also the left panel in Fig.~\ref{fig:ligoPeaks}, as well as the high frequency line at $438$Hz, right panel in Fig.~\ref{fig:ligoPeaks}. 
+Fig.~\ref{fig:LigoOrderError} shows the distribution of recovered AR orders $m$ against the relative frequency-averaged error. The behaviour
+of the three loss functions is very similar to what found in the Gaussian PSD case: ODB infers the smallest orders and gives average errors around $20$\%, FPE consistently estimates orders of a few hundreds and shows the smallest errors $\sim 15$ \%while CAT does not show any preference towards any AR order and displays wildly varying errors.
+Yet again, when the PSD is averaged over multiple realization of the time, CAT is able to capture the spectrum very precisely.
+In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them (figs.\ref{fig:ligoPeaks}).
+Hence, also the study of simulated Advanced LIGO data seems to indicate that whenever and wherever ensemble-averaged PSD are necessary, CAT is the optimal choice of loss function.
+However, on a single time-series realisation, FPE is the more robust choice.
 
-The conclusions are very similar to those above. When averaged on multiple realization, CAT loss function is able to capture the spectrum very precisely.
-In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them (figs.\ref{fig:ligoPeaks}).
-However, the computation on a single noise realization yields larger error, as manifest in fig.~\ref{fig:LigoOrderError}.
-This supports the conclusion that CAT is not a stable optimizer: different computation of the same spectrum can output very different results.
-On the other hand, FPE seems more robust and provides a smaller variance (and frequency averaged error), despite having a worse performance on the average.
-We find again that FPE and OBD have similar performance: indeed, OBD shows qualitatively the same behaviour of FPE, although with slightly worse performance in the error.
-
-We can then draw some general conclusions:
+Let us summarise some key general conclusions:
 \begin{itemize}
@@
 In this situation, CAT would provide spurious and unreliable results, with large error.
-	\item But, in case we have several realization for the same process, this last property of CAT, together with its instability, result to be his luck: averaging the spectrum over different time series tend to compensate noise in the results, originating a very well precise spectral estimation.
-The same is not true for FPE.
-\end{itemize}
+	\item In the case of several realizations of the same process, CAT ensemble-average properties provide very precise spectral estimation.\end{itemize}
 
-The choice of the loss function between CAT and FPE seems problem-dependent.
-
+Therefore, the choice of loss function, at least in between CAT and FPE, depends on the problem one  is attempting to solve.
 \begin{figure}[t]

### 2021-03-05 paper/mesa.tex
added opt plots
Removed old img + some changes

@@
 
+\subsection{How well the autoregressive order is recovered?}
+
+\begin{figure}
+	\centering
+	\includegraphics[width = \linewidth]{Images/arp_errors/scatter_deltap_ptrue}
+	\caption{Reconstructed value for the autoregressive order plotted against the true value of the autoregressive order.
+	The reconstructed autoregressive orders are computed from a time series randomly drawn with an $AR(p)$ model, with the three different loss functions under investigation.
+	}
+	\label{fig:p_vs_ptrue}
+\end{figure}
+
+We now address the issue of how well the autoregressive order (i.e. the number of $a_k$ coefficients employed) is estimated with each loss function.
+In doing so, we generate $100$ autoregressive processes $AR(p)$ with a random value of $p$, drawn such that $log(p)~\mathcal{U}[log(2), log(5000)]$.
+Each coefficient $a_k$ is assigned as follows: $a_k = \alpha_k e^{-0.05\cdot k}$, where $\alpha ~ \mathbb{N}(0,0.1)$ and the exponential ensures that the coefficients falls with distance (as in any realistic finite memory process).
+We report the result of this investigation in fig.~\ref{fig:p_vs_ptrue}
+
+Observations below!!!!
+
 \section{Comparison with Welch method}

### 2021-03-05 memspectrum.py

@@ -425,9 +425,9 @@ class MESA(object):
         self.regularisation = regularisation
         self.early_stop = early_stop
         if m is None:
-            self.mmax = int(2*self.N/np.log(2.*self.N))
+            self.mmax = int(2*self.N/np.log(2.*self.N)) + 1
         else:
-            self.mmax = m
+            self.mmax = m +1
         
         if optimisation_method == 'Fixed':
             self.early_stop = False

### 2021-03-05 scripts/autoregressive_plots.py

@@ -7,15 +7,15 @@ sys.path.insert(0,'..')
 import memspectrum
 
 
-N_process = 50 #10 simulated autoregressive processes
-N_data = 2000 #number of timesteps for each autoregressive process
-p_max = 30
+N_process = 100 #10 simulated autoregressive processes
+N_data = 50000 #number of timesteps for each autoregressive process
+p_max = 5000
 
 save_dir = 'arp_data/'
 plot_dir = '../paper/Images/arp_errors/'
 
-load = True
-plot = True
+load = False
+plot = False
 
 	#initializing the models
 true_arp_list = []
@@ -27,7 +27,7 @@ for i in range(N_process):
 
 	if load:
 		true_arp_list.append(memspectrum.MESA(save_dir+'arp_model_{}'.format(i)))
-		data_list.append(np.loadtxt(save_dir+'data_{}'.format(i)))
+		#data_list.append(np.loadtxt(save_dir+'data_{}'.format(i)))
 		rec_MESA_list.append( (memspectrum.MESA(save_dir+'MESA_model_{}_{}'.format(loss_functions[0],i)), 
 			memspectrum.MESA(save_dir+'MESA_model_{}_{}'.format(loss_functions[1],i)),
 			memspectrum.MESA(save_dir+'MESA_model_{}_{}'.format(loss_functions[2],i)) )
@@ -35,13 +35,15 @@ for i in range(N_process):
 	else:
 		true_arp_list.append(memspectrum.MESA())
 			#getting p
-		p = np.random.choice(range(2,p_max))
+		p = np.random.choice(np.linspace(np.log10(2),np.log10(p_max), 10000))
+		p = int(10**(p))
 		
 		a_k = np.random.normal(0, 0.1, size = (p,))
-		a_k = np.multiply(a_k,np.exp(-0.1*np.arange(0,len(a_k),1)))
+		a_k = np.multiply(a_k,np.exp(-0.05*np.arange(0,len(a_k),1)))
+		a_k[0] = 1 #you forgot this part!!! Now you need to start again :(
 		P = np.random.exponential(0.05)
 
-		print('Process {} has p = {} and P, a_k = {}, {}'.format(i,p,P,a_k))
+		print('Process {} has p = {} and P, a_k = {}, {}'.format(i,p,P,None))
 		
[…]

### 2021-03-05 paper/mesa.tex
Updated loss functions descriptions

--- 2021-03-03 paper/mesa.tex
@@
 \end{equation}
-with $P_m$ defined by the prediction error filter equation \sschmidt{What is this equation?}.
-\wdp{comment on N that goes to infinity, just minimising the variance of the process}
+with $P_m$ defined by the prediction error filter equation \ref{eq:errorFilter1} \sschmidt{What is this equation?}.
 \item \textbf{Criterion Autoregressive Transfer function (CAT)}
-This second loss function is due to Parzen~\cite{bhansali1986}, and it is defined as
+This second loss function has been written by Parzen and studied into details by Banshali ~\cite{bhansali1986}. 
+It is based on the assumption that the observed process is an infinite order autoregressive process 
 \begin{equation} 
-    \mathcal{L}_{\rm CAT}(m) = \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
+x_t = \sum_{i = 1}^{\infty}a_{i}x_{t-i} + \nu_t
 \end{equation} 
-It has the interesting property that the distance between the original filter and the estimate is minimum and asymptotically unbiased. \wdp{please expand}
+and tries to select the order $m$ as the best finite-order approximation for the observed process.
+Being N the number of samples it has the property that $m \to \infty$ as $N \to \infty$.
+This is an interesting result since any real-valued
+stochastic process can be written as a ARMA(p,q) process (Wold theorem) i.e. an $AR(\infty)$ process thus this is a physically significant property. 
+The loss function has the functional form: 
+\begin{equation}
+    \mathcal{L}_{\rm CAT}(m) = \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m},
+\end{equation}
+and the so chosen order is found to be asymptotically unbiased for $P_N$.
 
 \item \textbf{Optimum Bayes Decision rule} 
-The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}.
-With the technique of asymptotic integration, the loss function is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}, \wdp{please check this equation}:
+The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
+time series for the process that has to be described as an autoregressive process of order $m$ (AR(m)) with m to be determined.
+The OBD is obtained chosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$.
+Each
+hypothesis is uniquely determined from both the length and the values of the filter under study.
+Chosing a Likelihood of the form \ref{eq:p_forecast} and uniform priors for both the coefficients and the hypothesis Rao has shown that chosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
+maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of:
    \begin{align}
@@
 \end{equation}
-with $a_0 = 1$.
+with $a_0 = 1$ and $\Delta t$ the uniform sampling interval for the time series.
 The vector obtained as $(1, a_1, \dots, a_N)$ is known as the prediction error filter.

### 2021-03-10 paper/mesa.tex
Removed merge strings

--- 2021-03-05 paper/mesa.tex
@@
 \end{equation}
-with $P_m$ defined by the prediction error filter equation \sschmidt{What is this equation?}.
-\wdp{comment on N that goes to infinity, just minimising the variance of the process}
+with $P_m$ defined by the prediction error filter equation \ref{eq:errorFilter1}. \\ 
+If the $N \to \infty$ limit is considered,remembering $m_{max} = 2N / \log(2N)$, we see that N dominates both the numerator and the denominator of the fraction: in large $N$ limit using Akaike's loss function is equivalent to the minimization of the variance $P_m$ of the process.
+
 \item \textbf{Criterion Autoregressive Transfer function (CAT)}
-This second loss function is due to Parzen~\cite{bhansali1986}, and it is defined as
+This second loss function has been written by Parzen and studied into details by Banshali ~\cite{bhansali1986}. 
+It is based on the assumption that the observed process is an infinite order autoregressive process 
 \begin{equation} 
-    \mathcal{L}_{\rm CAT}(m) = \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
+x_t = \sum_{i = 1}^{\infty}a_{i}x_{t-i} + \nu_t
 \end{equation} 
-It has the interesting property that the distance between the original filter and the estimate is minimum and asymptotically unbiased. \wdp{please expand}
+and tries to select the order $m$ as the best finite-order approximation for the observed process.
+Being N the number of samples it has the property that $m \to \infty$ as $N \to \infty$.
+Since any real-valued stochastic process can be written as a ARMA(p,q) process (Wold theorem) i.e.
+an $AR(\infty)$ process, this is a physically significant property.
+The loss function has the functional form: 
+\begin{equation}
+    \mathcal{L}_{\rm CAT}(m) = \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m},
+\end{equation}
+and the so chosen order is found to be asymptotically unbiased for $P_N$.
 
 \item \textbf{Optimum Bayes Decision rule} 
-The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}.
-With the technique of asymptotic integration, the loss function is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}, \wdp{please check this equation}:
+The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
+time series for the process that has to be described as an autoregressive process of order $m$ (AR(m)) with m to be determined.
+The OBD is obtained chosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$.
+Each
+hypothesis is uniquely determined from both the length and the values of the filter under study.
+Chosing a Likelihood of the form \ref{eq:p_forecast} and uniform priors for both the coefficients and the hypothesis Rao has shown that chosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
+maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of \wdp{please check this equation}:
+
    \begin{align}
@@
 
-Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson\wdp{add ref} recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached.
+Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson \cite{doi:10.1002/sapm1946251261} recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached.
 Then, the order $m$ is selected to be the one that minimizes the loss function.
@@
 its easily shown that $S(f)^{-1}$ can be written as a Fourier Expansion in terms of some $\lambda$s coefficients.
-THe determination of the values for the $\lambda$s uniquely solve the problem of power-spectral density estimation.
+The determination of the values for the $\lambda$s uniquely solve the problem of power-spectral density estimation.
 Some properties for the coefficients can be worked out easily.

### 2021-03-10 memspectrum.py
Added autocovariance

@@ -190,6 +190,8 @@ class MESA(object):
         """
         self.P = None
         self.a_k = None #If a_k and P are None, the model is not already fitted
+        self.N = None
+        self.mu = None
         self.optimization = None
         if isinstance(filename,str):
         	self.load(filename)
@@ -198,7 +200,7 @@ class MESA(object):
         """
         Save the class to file (if the spectral density analysis is already performed).
         The output file can be used to load the class with method load()
-        File is a 1D array with the format: [P, N, a_k]. The header holds the shapes for each array
+        File is a 1D array with the format: [P, N, mu, a_k]. The header holds the shapes for each array
         
         Parameters
         ----------
@@ -208,8 +210,8 @@ class MESA(object):
         if self.P is None or self.a_k is None:
             raise RuntimeError("PSD analysis is not performed yet: unable to save the model. You should call solve() before saving to file") 
         
-        to_save = np.concatenate([[self.P], [self.N], self.a_k])
-        header = "(1,1,{})".format(len(self.a_k))
+        to_save = np.concatenate([[self.P], [self.N],[self.mu], self.a_k])
+        header = "(1,1,1,{})".format(len(self.a_k))
         np.savetxt(filename, to_save, header = header)
         return
         
@@ -236,11 +238,11 @@ class MESA(object):
         shapes = eval(first_line.translate({ord('#'): None, ord(' '): None}))
             #checking for the header
         if not isinstance(shapes, tuple):
-            if len(shapes) != 3 or not np.all([isinstance(s, int) for s in shapes]):
+            if len(shapes) != 4 or not np.all([isinstance(s, int) for s in shapes]):
                 raise ValueError("Wrong format for the header: unable to load the model")
 
             #assigning values
-        self.P, self.N, self.a_k = np.split(data, np.cumsum(shapes)[:-1])
+        self.P, self.N, self.mu, self.a_k = np.split(data, np.cumsum(shapes)[:-1])
         self.N = int(self.N.real)
         
         return
@@ -319,8 +321,6 @@ class MESA(object):
             (positive) frequencies to evaluate the spectrum at (shape (N,))
             must be equally spaced
 
-        
-
         Returns: 
         ----------
         if no frequency array is given 
@@ -356,7 +356,31 @@ class MESA(object):
             raise ValueError("Type of frequencies not understood: expected to be None or np.ndarray but given {} insted".format(type(frequencies)))
 
         return
+    
[…]

### 2021-03-10 climate_data/test_autocov.py
Test autocov

@@ -1,0 +1,98 @@
+"""
+Two misteries:
+	1) Why the empirical autocovariance and the mesa autocovariance have different scales?
+	2) Which is the relation between autocovariance and autocorrelation?
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+sys.path.insert(0,'..')
+import memspectrum
+import scipy.signal
+
+sys.path.insert(0,'../scripts')
+import welch
+
+def empirical_autocorrelation(data):
+	corr = scipy.signal.correlate(data, data, mode="full")
+		#computing average value
+	lags = scipy.signal.correlation_lags(data.size, data.size, mode="full")
+	N_divide = len(data)-np.abs(lags) #number of data summed at each lag
+	return np.divide(corr,N_divide)
+
+
+data = np.loadtxt('data_@1h_len95689.dat')
+#data = data - np.mean(data)# + 3.242
+
+srate = 1/3600.
+
+M = memspectrum.MESA()
+M.solve(data, method ='fast', optimisation_method = 'FPE', early_stop = True)
+M.save("test_model_FPE")
+M.load("test_model_FPE")
+
+K = -2.342932 #1- np.mean(data)*np.sum(M.a_k)
+M.a_k = M.a_k/K
+spec_scaled, f = M.spectrum(1./srate)
+
+M.solve(data, method ='fast', optimisation_method = 'FPE', early_stop = True)
+spec, f = M.spectrum(1./srate)
+S_0 = spec[np.where(f==0)]
+print(S_0)
+
+	#spectrum of zero mean data
+M_zero = memspectrum.MESA()
+M_zero.solve(data- np.mean(data), method ='fast', optimisation_method = 'FPE', early_stop = True)
+spec_zero, f = M_zero.spectrum(1./srate)
+S_0_zero = spec_zero[np.where(f==0)]
+print(S_0_zero)
+
+print(np.allclose(spec_zero, spec), np.allclose(spec, spec_scaled*1/K**2))
+
+plt.figure()
+plt.loglog(f,spec_scaled*1/K**2, label = "symmetry")
+plt.loglog(f,spec_zero, label = 'zero mean')
+plt.loglog(f,spec, label = 'normal data')
+plt.legend()
+#plt.show()
+
+
+autocov_empirical = empirical_autocorrelation(data-np.mean(data))
[…]

### 2021-03-10 memspectrum.py
Bug fix in MESA.compute_autocovariance

@@ -339,6 +339,8 @@ class MESA(object):
             ValueError if frequencies greater then Nyquist frequencies are given 
             
         """
+        if self.a_k is None:
+            raise RuntimeError("Model is not initialized: unable to compute spectrum. Call MESA.solve() or load a model from file to initialize the model")
         f_ny = .5 / dt 
         spec, f_spec = self._spectrum(dt, self.N)
         
@@ -357,7 +359,7 @@ class MESA(object):
 
         return
     
-    def compute_autocovariance(dt, normalize = False):
+    def compute_autocovariance(self, dt, normalize = False):
         """
         Compute the autocovariance of the data based on the autoregressive coefficients.
         The autocovariance is defined as: C(tau) = E_t[(x_t -mu)(x_t+tau -mu)]
@@ -375,8 +377,12 @@ class MESA(object):
         autocov: 'np.ndarray'                   
             Autocovariance of the model
         """
-        spec, f = M.spectrum(dt)
-        autocov = np.fft.irfft(spec*np.sqrt(self.mu)) #or there is a +1 in there...
+        spec, f = self.spectrum(dt)
+        spec = spec[:int(self.N/2)]
+        f = f[:int(self.N/2)]
+        autocov = np.fft.irfft(spec) #or there is a +1 in there...
+        #autocov -= np.square(self.mu)
+        #print(self.mu)
         if normalize:
             autocov /= autocov[0]
         return autocov       

### 2021-03-22 scripts/autoregressive_plots.py
Dirichlet in autoregressive plots

@@ -11,11 +11,11 @@ N_process = 100 #10 simulated autoregressive processes
 N_data = 50000 #number of timesteps for each autoregressive process
 p_max = 5000
 
-save_dir = 'arp_data/'
+save_dir = 'arp_data_dirichlet/'
 plot_dir = '../paper/Images/arp_errors/'
 
-load = True
-plot = True
+load = False
+plot = False
 
 	#initializing the models
 true_arp_list = []
@@ -37,9 +37,11 @@ for i in range(N_process):
 			#getting p
 		p = np.random.choice(np.linspace(np.log10(2),np.log10(p_max), 10000))
 		p = int(10**(p))
+		#p=int(p_max)
 		
-		a_k = np.random.normal(0, 0.1, size = (p,))
-		a_k = np.multiply(a_k,np.exp(-0.05*np.arange(0,len(a_k),1)))
+		a_k = np.random.dirichlet([1. for i in range(p)])
+		a_k = a_k * (2*np.random.binomial(1,0.5,a_k.shape)-1) #flipping the sign at random
+		
 		a_k[0] = 1 #you forgot this part!!! Now you need to start again :(
 		P = np.random.exponential(0.05)
 
@@ -48,23 +50,21 @@ for i in range(N_process):
 			#initializing the model
 		true_arp_list[-1].N = N_data
 		true_arp_list[-1].P = P
+		true_arp_list[-1].mu = 0.
 		true_arp_list[-1].a_k = a_k
 		
 		true_arp_list[-1].save(save_dir+'arp_model_{}'.format(i))
 		
 			#forecasting
 		data_0 = np.random.rand(p) #initial data
-		burn_steps = 1000*p
+		burn_steps = 100*p
 		
 		data = true_arp_list[-1].forecast(data_0, burn_steps+N_data)[0,:] #(burn_steps+N_data,)
 		
 		data_list.append(data[-N_data:])
 		#print("Time series: ", data_list[-1])
 		
-		#plt.plot(data_list[-1])
-		#plt.show()
 		np.savetxt(save_dir+'data_{}'.format(i), data_list[-1])
-		
 		
 			#reconstructing with MESA
 		temp_MESA =  (memspectrum.MESA(), memspectrum.MESA(), memspectrum.MESA()) 
@@ -125,6 +125,9 @@ if plot:
 	
[…]

### 2021-03-22 paper/mesa.tex
small edits

@@
 These are stochastic processes that display an invariance of their statistical properties, such as their two-point autocovariance function, with respect to time translation.
-If $x(t)$ is a wide-sense stationary process, it is completely determined by the knowledge of the auto-covariance function
+If $x(t)$ is a wide-sense stationary process, it is completely determined by the knowledge of the autocorrelation function
 \begin{equation}
@@
 \end{equation}
-The sequence $W$ is a window function that can be chosen in several different way, each choice presenting advantages and disadvantages for the final estimate of the PSD. \\ 
+The sequence $W$ is a window function that can be chosen in several different way, each choice presenting advantages and disadvantages for the final estimate of the PSD.
+\sschmidt{Here we say different things about periodigrams.
+We should be more clear...}
+
 The choice of a window function is arbitrary and typically is made by trial and error, until a satisfactory compromise between variance and resolution of the estimate of PSD is reached.
@@
 This provide a natural and simple machinery to forecast a time series, thus predicting past observations based on previous ones.
+
 In this paper we discuss the details of the Maximum entropy principle, its application to Power Spectral Density estimation with Burg algorithm and the link between Burg's algorithm and autoregressive process.
-Furthermore, we present and describe a new code, `memspectrum`, that provides a robust python implementation of the algorithm.
+Furthermore, we present and describe a new code, \texttt{memspectrum}, that provides a robust python implementation of the algorithm.
 We provide a thorough assessment of the performance of our code and we validate our results performing a number of tests.
@@
 After a short introduction to Jayne's MAXENT (sec.~\ref{sec:MAXENT}), we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form (sec.~\ref{sec:MESA}).
-Next, we will discuss an interesting link between Burg's method and autoregressive processes (sec.~\ref{sec:autoregr}) and in sec.~\ref{sec:forecasting} we will use such link to introduce a method for forecasting a time series.
+Next, we will discuss an interesting link between Burg's method and autoregressive processes (sec.~\ref{sec:autoregr}) and in sec.~\ref{sec:forecasting} we will use such link for straightforwardly forecasting a time series.
 
@@
 %To report the results of such an experiment, we need to communicate every single outcome: it 'carries' high information. 
-In this context, the role played by evidence should also be clear: it dictates the probability assignment in order to make the result 'more certain' or 'less informative'.
-Any assignment different from the 'most uncertain' is violating the principle of it to being non committal to non-available evidence.
+In this context, the role played by evidence should also be clear: it dictates the probability assignment in order to make the result `more certain' or `less informative'.
+Any assignment different from the `most uncertain' is violating the principle of it to being non committal to non-available evidence.
 We can only assign a more certain probability, i.e.
@@
 This defines a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized.
-$H$ has to be a 'generalization' of information defined in equation (\ref{eq:information}) to be something like the 'expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
+$H$ has to be a `generalization' of information defined in equation (\ref{eq:information}) to be something like the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
 %
@@
 For instance, whenever the only constraint available is the normalisation of the probability distribution, the entropy is maximised by the uniform distribution.
-If the information available is instead the expected value -- the mean -- entropy is maximised by the  exponential distribution. Of particular importance is 
-the case in which, in addition to the mean, also the variance is known. MAXENT leads to the Gaussian distribution. 
+If the information available is instead the expected value -- the mean -- entropy is maximised by the  exponential distribution.
+Of particular importance is the case in which, in addition to the mean, also the variance is known. MAXENT leads to the Gaussian distribution. 
 We retain this particularly interesting from the foundational point of view, since it provides a deeper insight into the ubiquitous Gaussian distribution.
@@
 \end{align}
+
 We have already seen that for a wide-sense stationary process the mean function is independent of time, hence redefined to be equal to zero without loss of generality, and the auto-covariance function is dependent only on the time lag $\tau \equiv t_i - t_j$.
 One can thus choose a sampling rate $\Delta t$ so that $C_{ij} = C((i-j)\Delta t)$.
-The autocovariance matrix thus becomes a Toeplitz matrix.
+The autocovariance matrix thus becomes a Toeplitz matrix\footnote{
+We remind the reader that a Toeplix matrix is a matrix in the form:
+\begin{pmatrix}
+	a_0 & a_1 & a_2 & \ldots & \ldots& \ldots &a_n\\
+	a_{-1} & a_0 & a_1 & \ldots & \ldots& \ldots &a_{n-1}\\
+	a_{-2} & a_{-1} & a_0 & \ldots & \ldots & \ldots  &a_{n-2}\\
+	\vdots & \vdots & \vdots & \vdots & \vdots & \vdots &\vdots\\
+	a_{-n +1} & \ldots & \ldots & \ldots& a_{-1} & a_0    &a_{1}\\
+	a_{-n} & \ldots & \ldots & \ldots& a_{-2} & a_{-1} & a_0
+\end{pmatrix}
+
+}.
 Toeplitz matrices are asymptotically equivalent to circulant matrices and thus diagonalised by the discrete Fourier transform base\cite{Gray}.
@@
 \end{align}
-where the matrix $S$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
+where the matrix $S = S_i \delta_{ij}$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
 Many readers will recognise the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\cite{woodward}
@@
 is impossible.
-Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases with $k$ as $\sigma \sim 1/\sqrt{N - k}$, so that only a few values for the autocorrelation function can actually be computed reliably.
+Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases as $\sigma \sim 1/\sqrt{N - k}$\footnote{
+This is clearly understood: when computing the autocorrelation at order $k$, only $N-k$ examples of the product $x_t x_{t+k}$ are available and the variance of the average value goes as the inverse of the square root of the points considered.
+}, so that only a few values for the autocorrelation function can actually be computed reliably.
 This bring us the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function?
@@
 thus being potentially able to model complex autocorrelation between data.
-\par
-To show the analogy, we compute the PSD of such process and we show that it is formally equivalent to the PSD obtained in eq.~\ref{eq:MESApsd}.
+
+It turns out that the maximums entropy principle provides a representation of the time series as an $AR(p)$ process.
+The Burg's algorithm computes the autoregressive coefficients that are suitable to the available data.
+Thanks to Wold's theorem \cite{something}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general.
+
+To show the analogy, we compute the PSD $S_{AR(p)}$ of an $AR(p)$ process and we show that it is formally equivalent to the PSD obtained in eq.~\ref{eq:MESApsd}.
+This will also provide a direct expression for the autoregressive coefficients $b_i$ and for the noise variance $\sigma^2$.
 We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~(\ref{eq:AR_p}):
@@
 \begin{equation}\label{eq:ARspectrum}
-    S(f) = \vert \tilde x(z)\vert ^ 2 = 
+    S_{AR(p)}(f) = \vert \tilde x(z)\vert ^ 2 = 
     \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}\,.
@@
 \subsection{Forecasting} \label{sec:forecasting}
-The link between MESA and autoregressive processes of particular interest.
+The link between MESA and autoregressive processes is of particular interest.
 Given the solution to the Levinson recursion to determine the $a_k$, see Appendix A, we have automatically the coefficients of the AR with equivalent PSD, hence we are able to exploit Eq.~(\ref{eq:AR_p}) to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
@@
 We now address the issue of how well the autoregressive order (i.e. the number of $a_k$ coefficients employed) is estimated with each loss function.
-In doing so, we generate $100$ autoregressive processes $AR(p)$ with a random value of $p$, drawn such that $log(p)~\mathcal{U}[log(2), log(5000)]$.
-Each coefficient $a_k$ is assigned as follows: $a_k = \alpha_k e^{-0.05\cdot k}$, where $\alpha ~ \mathbb{N}(0,0.1)$ and the exponential ensures that the coefficients falls with distance (as in any realistic finite memory process).
+In doing so, we generate $100$ autoregressive processes $AR(p)$ with a random value of $p$, drawn such that $log(p)~\mathcal{U}_{[log(2), log(5000)]}$.
+Each coefficient $a_k$ is assigned according to a Dirichlet distribution $\mathrm{Dir}([1,\hdots, 1])$. The sign of $a_k$ is assigned randomly according to a binomial distribution.
 We report the result of this investigation in fig.~\ref{fig:p_vs_ptrue}

### 2021-03-22 paper/mesa.tex
Edits to text, temperature plots and autoregressive plots

@@
 We now address the issue of how well the autoregressive order (i.e. the number of $a_k$ coefficients employed) is estimated with each loss function.
-In doing so, we generate $100$ autoregressive processes $AR(p)$ with a random value of $p$, drawn such that $log(p)~\mathcal{U}_{[log(2), log(5000)]}$.
+In doing so, we generate $100$ autoregressive processes $AR(p)$ with a random value of $p$, drawn such that $log(p) \sim\mathcal{U}_{[log(2), log(5000)]}$.
 Each coefficient $a_k$ is assigned according to a Dirichlet distribution $\mathrm{Dir}([1,\hdots, 1])$. The sign of $a_k$ is assigned randomly according to a binomial distribution.
-We report the result of this investigation in fig.~\ref{fig:p_vs_ptrue}
+We report the result of this investigation in fig.~\ref{fig:p_vs_ptrue}.
 
-Observations below!!!!
+It can be seen that OBD and FPE loss functions show very similar behaviour.
+They are {\it very} realiable in capturing the correct autoregressive order up to a certain threshold ($p\sim 100$ for OBD and $p\sim 200$ for FPE).
+For any autoregressive process of order higher than the threshold, the optimization underestimates badly the actual value of $p$.
+For this reason, FPE and OBD seems realiable only for simple autoregressive processes.
+For complicated processes, it seems they underfit the problem (i.e.
+they output a model simpler that those required).
 
-\section{Comparison with Welch method}
+On the other hand, CAT loss function shows a different behaviour.
+The selected autoregressive order $p$ is always close to the maximum possible value, regardless the actual value $p_{true}$ of the underlying process.
+This produces always a model that is more complex than those obtained with FPE and OBD and this can explain the origin of the high variance in the estimation observed in the experiments described in the previous section.
+Of course, CAT does not perform a good job in reconstructing an autoregressive process.
+However, for high order autoregressive process the error introduced by CAT is more tolerable than the error introduced by FPE and OBD.
+This is a good reason to prefer CAT optimizer in such situations (see sec.~\ref{sec:temp_timeseries} for another example of this).
+
+\sschmidt{Shall we add more?}
+
+\section{Comparison with Welch method}\label{sec:Welch_comparison}
 
@@
 
-\section{Forecast analysis}
-\section{Final remarks and future prospects}
+\section{Application: Temperature Time Series} \label{sec:temp_timeseries}
+We apply our code to the study of the temperature time series.
+This will allow to test the code performance on a different time series, of great importance in physics.
+We will also assess the quality of the forecasting predictions: this task has a great impact also on practical applications.
+We downloaded the historical reanalysis data from the Copernicus project (or whathever it is) \cite{}.
+Data refers to the temperature measure at coordinates $\textrm{N}\;\ang{45;5;} \;\; \textrm{E}\;\ang{9;1;}}$, corresponding about to the city of Milan, Italy.
+The temperature is measured once per hour for almost 31 years from 31st December 1989 to 30th November 2020.
 
+In fig.~\ref{fig:temp_spectrum}, we compute the spectrum over all the data. We employ both FPE and CAT loss functions.
+First of all, we note that at every frequency the estimation given by FPE is much less noisy the one from CAT: this issue was already studied in the previous sections. 
+We clearly see that there is a peak on the PSD at the frequency $f_D = \SI{1}{day^{-1}}$ corresponding to a day. This is somehow trivial and highly expected.
+Higher order harmonics (corresponding to integer multiples of $f_D$) are also visible up to the Nyquist frequency ($f_D = \SI{0.5}{hour^{-1}}$): they are required to explain the high variability of the daily temperature cycle throughout the year.
+
+Looking at the low frequencies, it is striking to note that FPE does not capture the peak $f_{yr} = \SI{1}{yr^{-1}}$ located at the frequency of a year.
+On the other hand, the peak is captured well by the CAT loss function.
+This can be understood by looking at fig.~\ref{fig:p_vs_ptrue}.
+It is shown that CAT systematically select very long filters (i.e.
+large $p_{CAT}$ values for the autoregressive orders), whereas FPE tends to select smaller values for the order $p_{FPE}$ of the autoregressive process, often understimating badly the actual value.
+In this example, we have $p_{CAT} \sim  36500 \sim  \SI{4}{yr}$, whereas $p_{FPE} \sim  1300 \sim  \SI{2}{months}$.
+Clearly, the autoregressive process selected by FPE cannot produce a meaningful prediction on the timescale of a few months and for this reason it is unable to capture the peak at $f_{yr}$.
+In other words, the model is too simple to capture \textit{both} the behaviour at high and low frequencies (separated by approximately 3 orders of magnitude).
+On the other hand, the model chosen by CAT is much more complex and is able to model also the high frequency behaviour.
+The ``actual" length of the autoregressive process should be closer to the choice of CAT.
+
+\sschmidt{
+Some remarks:
+\begin{itemize}
+	\item What happens if we downsample the data? FPE should be able to capture the yearly trend...
+	\item Remove yearly trend from the data?
+	\item Other plots?
+\end{itemize}
+}
+
+\sschmidt{Other comments below...}
+
+\begin{figure}
+	\caption{Spectrum for the historical temperature time series.
+The two lines refers to different loss functions (CAT and FPE); the inset shows the harmonics of the fundamental frequency of a day.
+Two vertical lines are drawn in correspondence to the frequency $f_D$ of a day and $f_{yr}$ of a year.}
+	\label{fig:temp_spectrum}
+	\includegraphics{Images/climate_plots/temp_spectrum.pdf}
+\end{figure}
+
+\section{Final remarks and future prospects} \label{sec:conclusion}
+\blindtext
+\blindtext
+\blindtext
+
+\pagebreak
 \appendix

### 2021-03-22 climate_data/analyse_climate_data.py

@@ -27,7 +27,8 @@ import welch
 
 #data = np.loadtxt('data_@1h_len95689.dat')[(195-30)*24:(195+30)*24] #summer
 #data = np.loadtxt('data_@1h_len95689.dat')[(2567-30)*24:(2567+30)*24] #winter
-data = np.loadtxt('data_@1h_len95689.dat')
+#data = np.loadtxt('data_@1h_len95689.dat')
+data = np.loadtxt('data_@1h_len271009_1989-2020.dat')
 
 srate = 1./3600 # 1/(1 hours)
 T = len(data)/srate
@@ -35,20 +36,30 @@ T = len(data)/srate
 	#togliere ciclo annuale
 	#togliere trend delle medie annuali
 
+mu = np.mean(data)
 f_0 = 1/(365)
 t = np.linspace(0, len(data)/24., len(data))
 A = (max(data)-min(data))/2.*0.5
 phi_0 = 2.8
-y = A*np.cos(2*np.pi*f_0*t +phi_0) +np.mean(data)
+if True:
+	def f(t, A, f_0,phi_0):
+		return A*np.cos(2*np.pi*f_0*t +phi_0) +mu
+	print("Initial guess for amplitude, freq and phase for annual trend: ",A, f_0,phi_0)
+	popt, pcov = scipy.optimize.curve_fit(f, t, data, (A, f_0,phi_0))
+	print("Final guess for amplitude, freq and phase for annual trend: ",*popt)
+	A, f_0,phi_0 = popt
+	y = f(t, A, f_0,phi_0)
+else:
+	y = f(t, A, f_0,phi_0)
 #data = data -y
 
 #computing spectrum
 M = memspectrum.MESA()
 if False:
-	M.solve(data-np.mean(data), method ='fast', optimisation_method = 'CAT', early_stop = False)
-	M.save("model_CAT_zeromean")
+	M.solve(data, method ='fast', optimisation_method = 'CAT', early_stop = False)
+	M.save("model_CAT_long")
 else:
-	M.load("model_CAT_zeromean")
+	M.load("model_CAT_long")
 N = len(data)
 spec, f = M.spectrum(1./srate)
 spec = spec[:int(N/2)]
@@ -60,11 +71,12 @@ if False:
 	f = f[:int(len(f)/2)-1]
 
 #forecasting
-N_forecast = 24*5 #1000 day of forecasting
[…]

### 2021-03-22 climate_data/test_autocov.py

@@ -18,8 +18,8 @@ def empirical_autocorrelation(data):
 		#computing average value
 	lags = scipy.signal.correlation_lags(data.size, data.size, mode="full")
 	N_divide = len(data)-np.abs(lags) #number of data summed at each lag
-	return np.divide(corr,N_divide)
-
+	return np.divide(corr,N_divide) #*np.mean(np.square(data)
+	
 
 data = np.loadtxt('data_@1h_len95689.dat')
 #data = data - np.mean(data)# + 3.242
@@ -31,8 +31,9 @@ M.solve(data, method ='fast', optimisation_method = 'FPE', early_stop = True)
 M.save("test_model_FPE")
 M.load("test_model_FPE")
 
-K = -2.342932 #1- np.mean(data)*np.sum(M.a_k)
+K = 2000.342932 #1- np.mean(data)*np.sum(M.a_k)
 M.a_k = M.a_k/K
+M.P = M.P/K**2
 spec_scaled, f = M.spectrum(1./srate)
 
 M.solve(data, method ='fast', optimisation_method = 'FPE', early_stop = True)
@@ -45,53 +46,76 @@ M_zero = memspectrum.MESA()
 M_zero.solve(data- np.mean(data), method ='fast', optimisation_method = 'FPE', early_stop = True)
 spec_zero, f = M_zero.spectrum(1./srate)
 S_0_zero = spec_zero[np.where(f==0)]
-print(S_0_zero)
+print(np.sum(M_zero.a_k), np.sum(M.a_k))
 
-print(np.allclose(spec_zero, spec), np.allclose(spec, spec_scaled*1/K**2))
+print(np.allclose(spec_zero, spec), np.allclose(spec, spec_scaled))
+
+print(M.a_k, M_zero.a_k)
+print(M.P, M_zero.P)
+print(np.sum(M.a_k), np.sum(M_zero.a_k))
+print(np.allclose(M.a_k[:10],M_zero.a_k[:10]))
+plt.plot(M_zero.a_k)
+plt.plot(M.a_k)
+#plt.show()
+
+#quit()
 
 plt.figure()
-plt.loglog(f,spec_scaled*1/K**2, label = "symmetry")
+plt.loglog(f,spec_scaled*1/K**0, label = "symmetry")
 plt.loglog(f,spec_zero, label = 'zero mean')
 plt.loglog(f,spec, label = 'normal data')
 plt.legend()
-#plt.show()
+plt.show()
 
 
[…]

### 2021-03-22 climate_data/test_grib.py

@@ -5,7 +5,11 @@ import xarray as xr
 import numpy as np
 import matplotlib.pyplot as plt
 
-ds = xr.open_dataset('adaptor.mars.internal-1614589133.4434555-10786-9-72b941a9-4c27-4d17-8061-3525cfcbd646.grib', engine='cfgrib', backend_kwargs={'read_keys': ['experimentVersionNumber']})
+ds = xr.open_dataset('adaptor.mars.internal-1614589133.4434555-10786-9-72b941a9-4c27-4d17-8061-3525cfcbd646.grib', engine='cfgrib', backend_kwargs={'read_keys': ['experimentVersionNumber']}) #20091231-20201130
+#ds = xr.open_dataset('adaptor.mars.internal-1614619196.5842516-30992-8-350d6fd6-6cf6-4e7e-bed2-b92cca5f47ec.grib', engine='cfgrib', backend_kwargs={'read_keys': ['experimentVersionNumber']}) #19991231-20091231
+#ds = xr.open_dataset('adaptor.mars.internal-1614619254.7736595-26399-6-0b5cc738-4574-4489-9cfe-8b807abaf151.grib', engine='cfgrib', backend_kwargs={'read_keys': ['experimentVersionNumber']}) #19891231-19991231
+
+
 print(ds)
 print(ds.t2m.attrs)
 #help(xr.open_dataset)
@@ -36,7 +40,7 @@ b = b[~np.isnan(b)]
 print(b)
 t = np.linspace(0, len(b), len(b)) #hours
 
-np.savetxt('data_@1h_len{}.dat'.format(len(b)),b)
+np.savetxt('data_@1h_len{}_.dat'.format(len(b)),b)
 quit()
 
 plt.plot(t, b)
@@ -76,5 +80,67 @@ data_np = np.array(lat_list)
 print(data_np.shape)
 plt.imshow(data_np)
 plt.show()
+
+
+"""
+array(['1999-12-31T00:00:00.000000000', '2000-01-01T00:00:00.000000000',
+       '2000-01-02T00:00:00.000000000', ..., '2009-12-29T00:00:00.000000000',
+       '2009-12-30T00:00:00.000000000', '2009-12-31T00:00:00.000000000'],
+      dtype='datetime64[ns]')
+Coordinates:
+    number   int64 ...
+  * time     (time) datetime64[ns] 1999-12-31 2000-01-01 ... 2009-12-31
+    surface  int64 ...
+Attributes:
[…]

### 2021-03-22 memspectrum.py

@@ -210,7 +210,7 @@ class MESA(object):
         if self.P is None or self.a_k is None:
             raise RuntimeError("PSD analysis is not performed yet: unable to save the model. You should call solve() before saving to file") 
         
-        to_save = np.concatenate([[self.P], [self.N],[self.mu], self.a_k])
+        to_save = np.concatenate([[self.P], [float(self.N)],[self.mu], self.a_k])
         header = "(1,1,1,{})".format(len(self.a_k))
         np.savetxt(filename, to_save, header = header)
         return
@@ -393,7 +393,8 @@ class MESA(object):
               optimisation_method = "FPE",
               method              = "Fast",
               regularisation      = 1.0,
-              early_stop          = True):
+              early_stop          = True,
+              verbose = False):
         """
         Computes the power spectral density of the attribute data for the class
         using standard Burg method recursive and a Faster (but less stable) version. 
@@ -429,6 +430,9 @@ class MESA(object):
             maximum after 100 iterations. 
             Recommended for every optimisation method but CAT.
             Has no effect with "Fixed" optimizer.
+        
+        verbose: 'boolean'
+            Whether to print the status of the mesa solution
 
         Returns
         -------
@@ -456,6 +460,7 @@ class MESA(object):
         self.mu = np.mean(data)
         self.regularisation = regularisation
         self.early_stop = early_stop
+        self.verbose = verbose
         if m is None:
             self.mmax = int(2*self.N/np.log(2.*self.N))
         else:
@@ -524,6 +529,7 @@ class MESA(object):
         old_idx = 0
         #Loop for the FastBurg Algorithm
         for i in range(self.mmax):
+            if self.verbose: sys.stderr.write('\r\tIteration {0} of {1}'.format(i + 1, self.mmax))
             #Update prediction error filter and reflection error coefficient
             k, new_a = self._updateCoefficients(a[-1], g)
             #Update variables. Check paper for indices at j-th loop.
@@ -555,6 +561,8 @@ class MESA(object):
                     break
         if not self.early_stop:
             idx = np.nanargmin(optimization)
+
+        if self.verbose: sys.stderr.write('\n')
 
         return P[idx], a[idx], np.array(optimization)
    
@@ -647,6 +655,7 @@ class MESA(object):
         old_idx = 0
         #Burg's recursion
         for i in range(self.mmax):
[…]

### 2021-03-22 scripts/autoregressive_plots.py

@@ -14,8 +14,8 @@ p_max = 5000
 save_dir = 'arp_data_dirichlet/'
 plot_dir = '../paper/Images/arp_errors/'
 
-load = False
-plot = False
+load = True
+plot = True
 
 	#initializing the models
 true_arp_list = []

### 2021-03-22 scripts/plot_climate_data.py

@@ -1,0 +1,66 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from style_sheet import init_plotting
+
+import sys
+sys.path.insert(0,'..')
+import memspectrum
+
+plot_dir = '../paper/Images/climate_plots/'
+
+data = np.loadtxt('../climate_data/data_@1h_len271009_1989-2020.dat')
+
+srate = 1./3600 #in Hz 1/(1 hours)
+M_CAT = memspectrum.MESA()
+M_FPE = memspectrum.MESA()
+M_CAT.load("../climate_data/model_CAT_long")
+M_FPE.load("../climate_data/model_FPE_long")
+
+N = len(data)
+spec_FPE, f_FPE = M_FPE.spectrum(1./srate)
+spec_FPE = spec_FPE[:int(N/2)]
+f_FPE = f_FPE[:int(N/2)]
+
+spec_CAT, f_CAT = M_CAT.spectrum(1./srate)
+spec_CAT = spec_CAT[:int(N/2)]
+f_CAT = f_CAT[:int(N/2)]
+
+print(len(M_CAT.a_k),len(M_FPE.a_k))
+
+	#plot spectrum
+fig = init_plotting()
+ax = fig.gca()
+
+unit_shift = 3600*24 #frequency is made 1/day
+ax.loglog(f_CAT * unit_shift , spec_CAT, label = "CAT", c = 'k', zorder = 1)
+ax.loglog(f_FPE * unit_shift , spec_FPE, label = "FPE", c = 'r', zorder = 2)
+
+ax.axvline(1./(3600.*24.*365.)*unit_shift,lw = .5, c = 'b', ls = '--', zorder = 0)
+ax.axvline(1./(3600.*24.)*unit_shift,lw = .5, c = 'b', ls = '--', zorder = 0)
+ax.set_xlabel("frequency (1/day)")
+ax.set_ylabel("PSD (1/Hz)")
+
+ax.legend(loc = 'upper left')
+
+
+	#inset
+axins =  ax.inset_axes([0.25, 0.65, 0.75, 0.35])
+ids_CAT = np.where(f_CAT>.75/unit_shift)
+ids_FPE = np.where(f_FPE>.75/unit_shift)
+axins.loglog(f_CAT[ids_CAT] * unit_shift , spec_CAT[ids_CAT], label = "CAT", c = 'k')
[…]

### 2021-03-23 paper/mesa.tex
Added forecasting in temperatures

@@
 
-\sschmidt{Other comments below...}
+We now assess the accuracy of the forecasting of new observations of the time series.
+Based on actual data, we try to predict future values as described in sec.~\ref{sec:forecasting}.
+We produce $N =100$ indepedent predictions and we compute the median as well as the $90\%$ confidence interval.
+This is compared with the actual measured temparature values.
+The predictions spans a two years range of time.
+We report our results in fig.~\ref{fig:temp_spectrum}.
+
+We note the observed difference is always well included in the $90\%$ confidence interval: the forecasting predictions seem reliable.
+On the other hand, the confidence interval is pretty large (almost as large as $\SI{15}{K}$), thus making easy for the actual data to fit the predictions.
+We believe that the origin of such large uncertainty is due to the complexity of the time series considered: any realistic model, even more complex than our autoregressive process, might not produce drastically lower uncertainties (in such a long term).
+
+\sschmidt{Data should be divided in train and test set for the purpose of forecasting!!}
+
+\sschmidt{Other comments below??}
 
@@
 	\includegraphics{Images/climate_plots/temp_spectrum.pdf}
+\end{figure}
+
+\begin{figure}
+	\caption{Difference between the historical temperature $T$ and the forecast temperature $T_{forecast}$ for two years of data.
+The red shadow denotes the $90\%$ confidence interval of the predictions.
+The model is trained with CAT loss function.
+The discrepancy between predicted and actual values is always everywhere included in the $90\%$ confidence interval.}
+	\label{fig:temp_forecast}
+	\includegraphics{Images/climate_plots/forecast_accuracy.pdf}
 \end{figure}

### 2021-03-23 scripts/plot_climate_data.py

@@ -10,6 +10,7 @@ plot_dir = '../paper/Images/climate_plots/'
 
 data = np.loadtxt('../climate_data/data_@1h_len271009_1989-2020.dat')
 
+	#spectrum computation
 srate = 1./3600 #in Hz 1/(1 hours)
 M_CAT = memspectrum.MESA()
 M_FPE = memspectrum.MESA()
@@ -27,12 +28,39 @@ f_CAT = f_CAT[:int(N/2)]
 
 print(len(M_CAT.a_k),len(M_FPE.a_k))
 
+	#Forecasting
+N_forecast = 24*2*365
+
+if False:
+	forecast = M_CAT.forecast(data[:len(M_CAT.a_k)],N_forecast, 100, verbose = True)
+	np.savetxt('forecast_climate_data.dat', forecast)
+else:
+	forecast = np.loadtxt('forecast_climate_data.dat')
+l,m,h = np.percentile(forecast,[5,50,95],axis = 0)
+true = data[len(M_CAT.a_k):len(M_CAT.a_k)+N_forecast]
+
+	
+	#plot forecasting
+fig = init_plotting()
+ax = fig.gca()
+
+t_grid = np.linspace(0,N_forecast/24.,N_forecast)
+ax.plot(t_grid, (true-m), c = 'b', zorder = 1)
+ax.fill_between(t_grid, true-h, true-l, color = 'r',alpha = 0.3, zorder = 0)
+#ax.plot(t_grid, (true-m)/(h-l), c = 'b', zorder = 1)
+#ax.plot(np.linspace(0,N_forecast,N_forecast), true-h, c = 'r')
+#ax.plot(np.linspace(0,N_forecast,N_forecast), true- l, c = 'r')
+ax.set_xlabel("Time (days)")
+ax.set_ylabel(r"$T - T_{forecast} (K)$")
+plt.tight_layout()
+plt.savefig(plot_dir+"forecast_accuracy.pdf")
+
 	#plot spectrum
 fig = init_plotting()
 ax = fig.gca()
 
 unit_shift = 3600*24 #frequency is made 1/day
-ax.loglog(f_CAT * unit_shift , spec_CAT, label = "CAT", c = 'k', zorder = 1)
+ax.loglog(f_CAT * unit_shift , spec_CAT, label = "CAT", c = 'b', zorder = 1)
[…]

### 2021-03-23 paper/mesa.tex
Edits in theory and appendix

@@
 
+\sschmidt{This section might be simplified...}
+
 Before introducing MAXENT as a variational problem, we will workout some simple examples to contexualise how information's entropy
@@
 it has a low information entropy content. Even less information is required if we are studying a system whose outcome is 
-certain (has probability $p = 1$). If the amount of information $I$ can be considered to zero, a communication is not even needed.  
+certain (has probability $p = 1$), as in this case, a communication is not even needed.  
 Shannon \cite{Shannon} proposed the quantity
@@
 The more uncertain the outcome, the higher the information it brings.
-\par
-What happens if we consider a larger set of different events?
-What is  the `most uncertain' outcome?
-Since we are intending probability as a representation of the state of knowledge, one of the first requirement for our assignment is the agreement with common sense.
-Consider an experiment based on the occurrence of one of two different outcomes $E_1, E_2$ with given probabilities $P_1$ and $P_2$. What are the values that make our outcome the most uncertain?
-If $P_1$ and $P_2$ are largely different, like $P_1 = 0.999$ and $P_2 = 0.001$, we have enough evidence to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome.
-The most unpredictable experiment is the one in which 
+
+We now seek to generalize the measure for the information in the case where different outcomes are possible.
+To do so, let us consider an experiment with two different random outcomes $E_1, E_2$, with given probabilities $P_1$ and $P_2$, and we ask ourselfs which are the probability assignments that makes the outcome more uncertain (i.e.
+maximize the information).
+If $P_1$ and $P_2$ are largely different, for istance $P_1 = 0.999$ and $P_2 = 0.001$, we clearly have enough evidence to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome.
+The information content will be very low.
+On the other hand, most unpredictable experiment happens when 
 \begin{equation}\nonumber
@@
 this describes a situation of `maximum ignorance'; we have no evidence at all and we have no way of predicting what the outcome of such an experiment will be.
-The generalisation to $N$ possible outcomes is straightforward.
-\\
-%This can be generalized in the case of $N$ events. Assigning same probability
-%\begin{equation*}
-%    P_1 = \dots = P_N = \frac{1}{N}
-%\end{equation*}
-%for each outcome, i.e. appealing to the indifference principle (IP), is equivalent to the requirement for our experiment to be the most unpredictable. 
-%All we know is that one over $N$ events is going to happen. With this evidence, we cannot expect one outcome to occur over the others. 
-%To report the results of such an experiment, we need to communicate every single outcome: it 'carries' high information. 
-In this context, the role played by evidence should also be clear: it dictates the probability assignment in order to make the result `more certain' or `less informative'.
+The information content of this system must be high.
+
+Of course, when assigning probabilities for a set of events, all the evidence available must be taken into account.
+On the other hand, the probability assignment must not introduce additional evidence in the system and thus the need to be ``the most uncertain possible given the knoledge of the system".
+
+%In this context, the role played by evidence should be clear: it dictates the probability assignment in order to make the result `more certain' or `less informative'.
 Any assignment different from the `most uncertain' is violating the principle of it to being non committal to non-available evidence.
@@
 \par
-MAXENT ratifies the aforementioned principle by stating that probabilities should be assigned by maximizing uncertainty using evidence as constraint. 
+MAXENT puts the aforementioned principle into a more precise mathematical context by stating that probabilities should be assigned by maximizing uncertainty using evidence as constraint. 
 This defines a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized.
@@
 If the information available is instead the expected value -- the mean -- entropy is maximised by the  exponential distribution.
+
 Of particular importance is the case in which, in addition to the mean, also the variance is known. MAXENT leads to the Gaussian distribution. 
@@
 not on the PSD. The variational problem can be formulated in terms of the power spectral density $S(f)$ alone by
-considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA}. 
+considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA} \sschmidt{Which form has this filter?}. 
 The entropy gain obtained by filtering white noise with a sampling rate $\Delta t$ and Nyquist frequency $Ny \equiv \frac{1}{2 \Delta t}$ is:
-
 \begin{equation}\label{eq:EntropyGain}
@@
 
-Consider a realisation of a stochastic process $(x_1,\ldots,x_n)$ with sample autocorrelations $\bar r_i,\,i=0,\ldots, n/2$, we require that the variational solution $S(f)$ for the PSD matches the empirical autocorrelation:
+Before maximising the entropy gain, we need to include the evidence available as a form of mathematical constraints for the assignment of $S(f)$.
+This is equivalent in imposing that the variational solution $S(f)$ for the PSD matches the empirical autocorrelation.
+Let us define a realisation of a stochastic process $(x_1,\ldots,x_N)$ with sample autocorrelations $\bar r_k,\,k=0,\ldots, N/2$, then the PSD must satisfy the following equation:
 \begin{equation}\label{eq:MaxConstraint}
-\int_{-Ny}^{Ny} S(f) e^{\imath 2 \pi f n \Delta t} df = \bar r_{n}\,.
-\end{equation}
-This approach on power spectral density estimate was developed by Burg~\cite{burg1975maximum} and it is a way out from the problems raised in periodograms estimates.
-It requires no assumptions on the unavailable estimates for $\bar r_n$ and is consistent with the sample autocorrelation function by construction.
-Remarkably, the variational problem admits a closed-form analytical expression for $S(f)$:
-
+\int_{-Ny}^{Ny} S(f) e^{\imath 2 \pi f k \Delta t} df = \bar r_{k}\,.
+\end{equation}
+
+Thus, by maximizing eq.~\ref{eq:EntropyGain} with constraints in eq.~\ref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
+This approach on PSD computation provide a result consistent with the epirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
+Most importantly, the variational problem admits a closed-form analytical expression for $S(f)$.
+The expression was first found by Burg~\cite{burg1975maximum}:
 \begin{equation}\label{eq:MESApsd}
@@
 \end{equation}
-with $\Delta t$ the sampling interval, $z=\exp{(2\pi i f\Delta t)}$, $a_0$ = 1 and the coefficients $P_N$ and $a_i (i \neq 0)$ to be determined.
-The details of the derivation and the actual values of the coefficients $a_i$ can be found in appendix \ref{sec:MESA_solution}.
+where $\Delta t$ is the sampling interval of the time series, $z=\exp{(2\pi i f\Delta t)}$, $a_0$ = 1.
+The vector obtained as $(1, a_1, \dots, a_N)$ is also known as the \textit{prediction error filter}.
+The coefficients $P_N$ and $a_s (s > 0)$ are to be determined by an iterative process (called Burg's algorithm).
+The details of the derivation and the actual form for the coefficients $a_s$ can be found in appendix \ref{sec:MESA_solution}.
 
@@
 \subsection{MESA solution}
-To derive the expression for the MAXENT spectral estimator the approach proposed by Burg \cite{burg1975maximum} will be followed.
-The constraints \ref{eq:MaxConstraint} are not implemented via the standard Lagrange Multipliers approach while the entropy gain \ref{eq:EntropyGain} will be re-written by explicitly writing $S(f)$ as the Fourier Transform of the sample autocorrelation function:
+We derive the expression for the MAXENT spectral estimator following the approach proposed by Burg \cite{burg1975maximum}.
+Unlike the standard approach, we do not enforce the constraints in eq.~\ref{eq:MaxConstraint} with the standard Lagrange Multipliers approach.
+We write instead the PSD $S(f)$ as the Fourier Transform of the sample autocorrelation function: 
 \begin{equation}
@@
 \end{equation}
-and becomes
+and the entropy gain expression eq.~\ref{eq:EntropyGain} becomes:
 \begin{equation}
     \Delta H = \int_{-Ny}^{Ny}  
-    \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r(n\Delta t) e^{-\imath 2 \pi f n \Delta t} 
+    \log\left(\frac{1}{2 Ny}\sum_{n = -\infty}^{\infty} \bar r_n e^{-\imath 2 \pi f n \Delta t} 
     \right) df.
 \end{equation}
-Deriving with respect to the autocorrelation function $r_s$
-\begin{equation}
-      \frac{\delta H}{\delta \bar r_s} = \frac{1}{2Ny}\int_{-Ny}^{Ny} S(f)^{-1}e^{-\imath 2 \pi f s \Delta t } df = \lambda_s. 
-\end{equation} 
-its easily shown that $S(f)^{-1}$ can be written as a Fourier Expansion in terms of some $\lambda$s coefficients.
-The determination of the values for the $\lambda$s uniquely solve the problem of power-spectral density estimation.
-Some properties for the coefficients can be worked out easily.
-First, since $S(f)$ is positive definite the $\lambda$s coefficient show the property
+Note that this expression already takes into account the contraints in eq.~\ref{eq:MaxConstraint}.
+
+We now introduce a set of coefficients $\lambda_s$, defined as the derivative of $\Delta H$ with respect to the autocorrelation function $r_s$.
+Excplitely they are:
+\begin{equation} \label{eq:lamdas}
+      \lambda_s \coloneqq \frac{\delta H}{\delta \bar r_s} = \frac{1}{2Ny}\int_{-Ny}^{Ny} S(f)^{-1}e^{-\imath 2 \pi f s \Delta t } df
+\end{equation} 
+and we will show that $S(f)^{-1}$ can be written as a Fourier Expansion in terms of such coefficients.
+Thenm the determination of the values for the $\lambda_s$ uniquely solves the problem of power-spectral density estimation.
+
+Some properties for the coefficients can be worked out easily. First, since $S(f)$ is positive definite the $\lambda_s$ show the property 
 \begin{equation}
@@
 \end{equation}
-The second property is obtained considering that the autocorrelation function can only be computed for a finite time interval $n \in [-N, N]$, the requirement for our estimate to be independent from the unkown values of the autocorrelation function can be implemented as:
+The second property is obtained considering that the autocorrelation function $r_n$ can only be computed for a finite time interval $n \in [-N, N]$ and that the PSD estimation must not depend on the unavailable values $r_n$: this is part of the constraint in eq.~\ref{eq:MaxConstraint}
+This requirement can be implemented as:  
 \begin{equation}\nonumber 
@@
 \end{equation}
-From the previous equations is easily seen that $S(f)$ can be expressed via a Fourier Series 
+
+From eq.~\ref{eq:lamdas} and from the properties above, is easily seen \sschmidt{How exactly?} that $S(f)$ can be expressed via a Fourier Series 
 \begin{equation}\label{eq:PSDconstraint}
@@
 with $a_0 = 1$ and $\Delta t$ the uniform sampling interval for the time series.
-The vector obtained as $(1, a_1, \dots, a_N)$ is known as the prediction error filter.
+The vector obtained as $(1, a_1, \dots, a_N)$ is the prediction error filter.
 The power spectral density $S(f)$ is uniquely determined if both the prediction error filter and $P_N$ coefficients are computed.
-\\
-To compute the $a_i$-s it is convenient to rewerite the equation 
-(\ref{eq:MaxConstraint}) in terms of the Laurent Polynomial and then integrating over $z$. In this way the equation becomes:
-\begin{equation}
+
+To compute the $a_s$ is convenient to plug into eq.~\ref{eq:MaxConstraint} the Laurent Polynomial exansion for $S(f)$ eq.~\ref{eq:MESApsd_appendix} and then integrating over $z$.
+In this way the equation becomes:
+\begin{equation}
+    \label{eq:contourintegral}
    \frac{P_N}{2 \pi \imath} \oint _{\mathbb S^1}\frac{z^{-s - 1}}{\sum_{n = 0}^N a_n z^n \sum_{n = 0}^N a^*_n z^{-n}}dz = \bar r_s. 
 \end{equation}
-Substituting $s \to s - r$, multiplying by $a^*_s$ and summing over $s$ the previous equation becomes 
-\begin{equation}
-    \sum_{s = 0}^N a_s \bar r_{s - r} = \frac{P_N}{2 \pi \imath}\oint \frac{z^{r -1}}{\sum_{s = 0}^N a_s z^s}dz.\label{eq:ErFilter}
-\end{equation}
-From now on only wide-sense stationary processes will be considered: for such a process all poles lay outside the unit circle so that the previous integral can be easily computed obtaining the following, well known, equations:
-\begin{align}\label{eq:errorFilter1}
+Substituting $s \to s - r$, multiplying by $a^*_s$ and summing over $s$, the previous equation becomes 
+\begin{equation}
+    \label{eq:errorFilter}
+    \sum_{s = 0}^N a_s \bar r_{s - r} = \frac{P_N}{2 \pi \imath}\oint \frac{z^{r -1}}{\sum_{s = 0}^N a_s z^s}dz
+\end{equation}
+For a wide-sense stationary processes \sschmidt{Justify this assumption!}, all the poles lay outside the unit circle so that the previous integral can be easily computed obtaining the following, well known, equations:
+\begin{align}
+    \label{eq:errorFilter1}
     \sum_{s = 0}^N a_s \bar r_{r - s} &= P_N \quad \text{ if } r = 0 \\ \label{eq:errorFilter2}
@@
 \end{align}
-The solution of the previous equations fully determine the functional form of the power spectral density estimator \ref{eq:MESApsd_appendix}, and the method for solving such equations is the Levinson-Durbin recursion \cite{doi:10.1002/sapm1946251261}.
-The recursion can be easily schemitized as follows:
-After defining the quantities  
+
+\section{Levinson recursion} \label{sec:MESA_solution}
+The solution of the eqs.~(\ref{eq:errorFilter1}-\ref{eq:errorFilter2}) fully determines the functional form of the power spectral density estimator \ref{eq:MESApsd_appendix}.
+The method for solving the equations is called the Levinson-Durbin recursion \cite{doi:10.1002/sapm1946251261} and it is described in the following.
+For each order $N$ of the iteration we define the quantities:
 \begin{align}
-\Delta_N &= \sum_{n = 0}^{N}R(N - n + 1)a_n \\ 
+\Delta_N &= \sum_{n = 0}^{N} a_n \bar{r}_{N - n + 1} \\ 
 c_N &= - \frac{\Delta_N}{P_N},
-\end{align} the Levinson recursion is: 
+\end{align}
+
+The Levinson recursion computes the $N$th order quantities given the $N-1$th order quantities: 
 \begin{equation} \label{eq:Levinson1}
@@
 \end{equation}
-with $b$ the forward prediction error coefficients at order $N-1$. 
-The 0'th order element can be easily initialized reminding that $a_0 = 1$ whatever the order the other coefficients. $P_0$ can be determined from \ref{eq:errorFilter1} and its value is: 
+where $b$ is the forward prediction error coefficients at order $N-1$. 
+The 0-th order element can be easily initialized reminding that $a_0 = 1$ (always) and that $P_0$ can be determined from \ref{eq:errorFilter1}.
+Its values turns out to be: 
 \begin{equation}
@@
 \end{equation}
- while $\Delta_0$ and $c_0$ are uniquely determined from their definitions and are
+$\Delta_0$ and $c_0$ are uniquely determined from their definitions and they are:
 \begin{equation}
@@
 \end{equation}
-These expressions allow us to compute $\vec a$ and $P_N$ to whatever order by simply iterating \ref{eq:Levinson1} and \ref{eq:Levinson2}.
+
+These expressions allow us to compute $\vec a$ and $P_N$ to any order by simply iterating \ref{eq:Levinson1} and \ref{eq:Levinson2}.
 Substituing them in equation \ref{eq:MESApsd_appendix} the problem of the estimation for the power spectral density via maximum entropy principle is solved.
-While Burg's method for spectral analysis is solved via Levinson recursion, another faster recursion \cite{Vos} is implemented in memspectrum package.
+Burg's method for spectral analysis is solved via Levinson is implmenteed in the released \texttt{memspectrum} package.
+Another faster recursion method is available in \cite{Vos} and it is also available in \texttt{memspectrum}.
 

### 2021-03-23 paper/mesa.tex
Eqrefs

@@
 without highlighting its connection with the time structure of the process itself, thus masking some important properties that will explore further in what follows.
-The latter definition in (\ref{eq:psd-f-definition}) gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.
+The latter definition in~\eqref{eq:psd-f-definition} gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.
 %Under such assumption, it can be shown [cite something] that the problem of characterizing the noise properties is greatly simplified and its properties are summarized by a single \textit{autocorrelation function}, which describes how an observation at a given time correlates with another observation a later time.
@@
 This defines a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized.
-$H$ has to be a `generalization' of information defined in equation (\ref{eq:information}) to be something like the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
+$H$ has to be a `generalization' of information defined in equation \eqref{eq:information} to be something like the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
 %
@@
 
-%This interpretation is straightforward: if we compare it with Eq.~(\ref{eq:information}), we see that the entropy is defined as the 'average information'.
+%This interpretation is straightforward: if we compare it with Eq.~(\eqref{eq:information}), we see that the entropy is defined as the 'average information'.
 The probability assignments made maximizing it with respect to $p$, is somehow requiring  the 'expected value' of uncertainty to be a maximum.
-Such an assignment is in agreement with our hypothesis of being `non committal' with unavailable evidence, since \ref{eq:entropy} is all we need to assign probabilities uniquely.
+Such an assignment is in agreement with our hypothesis of being `non committal' with unavailable evidence, since \eqref{eq:entropy} is all we need to assign probabilities uniquely.
 Evidence can be added in form of constraints in the maximization.
@@
 %\end{equation}
-%so that equation \ref{eq:entropyNormConst} reduce to
+%so that equation \eqref{eq:entropyNormConst} reduce to
 %\begin{equation}\nonumber 
@@
 \par
-As in the previous examples, one needs to set up a variational problem where the entropy, Eq.~(\ref{eq:entropy}), is maximized 
+As in the previous examples, one needs to set up a variational problem where the entropy, Eq.~\eqref{eq:entropy}, is maximized 
 subject to some problem-specific constraints. 
@@
 
-Thus, by maximizing eq.~\ref{eq:EntropyGain} with constraints in eq.~\ref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
+Thus, by maximizing eq.~\eqref{eq:EntropyGain} with constraints in eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
 This approach on PSD computation provide a result consistent with the epirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
@@
 
-To show the analogy, we compute the PSD $S_{AR(p)}$ of an $AR(p)$ process and we show that it is formally equivalent to the PSD obtained in eq.~\ref{eq:MESApsd}.
+To show the analogy, we compute the PSD $S_{AR(p)}$ of an $AR(p)$ process and we show that it is formally equivalent to the PSD obtained in eq.~\eqref{eq:MESApsd}.
 This will also provide a direct expression for the autoregressive coefficients $b_i$ and for the noise variance $\sigma^2$.
-We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~(\ref{eq:AR_p}):
+We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~(\eqref{eq:AR_p}):
 \begin{align}
@@
 The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.
-Eq.~(\ref{eq:ARspectrum}) and Eq.~(\ref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
+Eq.~(\eqref{eq:ARspectrum}) and Eq.~(\eqref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
 This shows that the MAXENT estimation of the PSD also computes the coefficients of the autoregressive 
@@
 The link between MESA and autoregressive processes is of particular interest.
-Given the solution to the Levinson recursion to determine the $a_k$, see Appendix A, we have automatically the coefficients of the AR with equivalent PSD, hence we are able to exploit Eq.~(\ref{eq:AR_p}) to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
+Given the solution to the Levinson recursion to determine the $a_k$, see Appendix A, we have automatically the coefficients of the AR with equivalent PSD, hence we are able to exploit Eq.~\ref{eq:AR_p} to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
 Indeed, an AR($p$) process models the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation:
@@
 hypothesis is uniquely determined from both the length and the values of the filter under study.
-Chosing a Likelihood of the form \ref{eq:p_forecast} and uniform priors for both the coefficients and the hypothesis Rao has shown that chosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
+Chosing a Likelihood of the form \eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypothesis Rao has shown that chosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
 maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of \wdp{please check this equation}:
@@
 
-Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson \cite{doi:10.1002/sapm1946251261} recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached.
+Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson \cite{doi:10.1002/sapm1946251261} recursion until $M_{max}$, as given in Eq.~\eqref{eq:MMAx}, iterations are reached.
 Then, the order $m$ is selected to be the one that minimizes the loss function.
@@
 Figs.~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} show the averaged PSD for the three loss functions, 
-as well as the ensemble-averaged error from Eq.~(\ref{eq:ensemble_averaged_error}).
+as well as the ensemble-averaged error from Eq.~\eqref{eq:ensemble_averaged_error}.
 Furthermore, Fig.~\ref{fig:optcomparison} displays the correlation between $r_i$ and the autoregressive 
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/error_length_contour.pdf}
-	\caption{Relative error $r$ (as in eq.~(\ref{eq:freq_error})) vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with gaussian PSD given in Fig.~\ref{fig:autocorr}.
+	\caption{Relative error $r$ (as in eq.~\eqref{eq:freq_error}) vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with gaussian PSD given in Fig.~\ref{fig:autocorr}.
 Different colours refers to different choices for the loss function: in red $\mathcal{L}_{\rm FPE}(m)$, in green $\mathcal{L}_{\rm OBD}(m)$ and in black $\mathcal{L}_{\rm CAT}(m)$.
@@
     \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/error_length_contour.pdf}
-    \caption{For each of the $500$ independent realization of the time series, we plot the relative error $r$ (as in eq.~\ref{eq:freq_error}) against the length $m$ of autoregressive process.
+    \caption{For each of the $500$ independent realization of the time series, we plot the relative error $r$ (as in eq.~\eqref{eq:freq_error}) against the length $m$ of autoregressive process.
 The time series are randomly drawn with a the analytical LIGO PSD in fig.`\ref{fig:ligospectrum}.
@@
 \appendix
-\section{MESA solution} \label{sec:MESA_solution}
-\subsection{MESA solution}
+\section{Computing the PSD} 
+\subsection{MESA solution}\label{sec:MESA_solution}
 We derive the expression for the MAXENT spectral estimator following the approach proposed by Burg \cite{burg1975maximum}.
-Unlike the standard approach, we do not enforce the constraints in eq.~\ref{eq:MaxConstraint} with the standard Lagrange Multipliers approach.
+Unlike the standard approach, we do not enforce the constraints in eq.~\eqref{eq:MaxConstraint} with the standard Lagrange Multipliers approach.
 We write instead the PSD $S(f)$ as the Fourier Transform of the sample autocorrelation function: 
@@
 \end{equation}
-and the entropy gain expression eq.~\ref{eq:EntropyGain} becomes:
+and the entropy gain expression eq.~\eqref{eq:EntropyGain} becomes:
 \begin{equation}
@@
 \end{equation}
-Note that this expression already takes into account the contraints in eq.~\ref{eq:MaxConstraint}.
+Note that this expression already takes into account the contraints in eq.~\eqref{eq:MaxConstraint}.
 
@@
 \end{equation}
-The second property is obtained considering that the autocorrelation function $r_n$ can only be computed for a finite time interval $n \in [-N, N]$ and that the PSD estimation must not depend on the unavailable values $r_n$: this is part of the constraint in eq.~\ref{eq:MaxConstraint}
+The second property is obtained considering that the autocorrelation function $r_n$ can only be computed for a finite time interval $n \in [-N, N]$ and that the PSD estimation must not depend on the unavailable values $r_n$: this is part of the constraint in eq.~\eqref{eq:MaxConstraint}
 This requirement can be implemented as:  
@@
 
-From eq.~\ref{eq:lamdas} and from the properties above, is easily seen \sschmidt{How exactly?} that $S(f)$ can be expressed via a Fourier Series 
+From eq.~\eqref{eq:lamdas} and from the properties above, is easily seen \sschmidt{How exactly?} that $S(f)$ can be expressed via a Fourier Series 
 \begin{equation}\label{eq:PSDconstraint}
@@
 It is easy to show that if $z_0$ is a root for the polynomial $(z_0^*)^{-1}$ is also a root: for every root laying outside the unit circle there will be another root inside of it and vice-versa.
-These properties allow us to rewrite the Fourier expansion (\ref{eq:zExp}) as \cite{1975STIN...7714318B}:
+These properties allow us to rewrite the Fourier expansion \eqref{eq:zExp} as \cite{1975STIN...7714318B}:
 \begin{equation}\label{eq:MESApsd_appendix}
@@
 
-To compute the $a_s$ is convenient to plug into eq.~\ref{eq:MaxConstraint} the Laurent Polynomial exansion for $S(f)$ eq.~\ref{eq:MESApsd_appendix} and then integrating over $z$.
+To compute the $a_s$ is convenient to plug into eq.~\eqref{eq:MaxConstraint} the Laurent Polynomial exansion for $S(f)$ eq.~\eqref{eq:MESApsd_appendix} and then integrating over $z$.
 In this way the equation becomes:
@@
 
-\section{Levinson recursion} \label{sec:MESA_solution}
-The solution of the eqs.~(\ref{eq:errorFilter1}-\ref{eq:errorFilter2}) fully determines the functional form of the power spectral density estimator \ref{eq:MESApsd_appendix}.
+\subsection{Levinson recursion} \label{sec:LevinsonRecursion}
+The solution of the eqs.~(\ref{eq:errorFilter1}-\ref{eq:errorFilter2}) fully determines the functional form of the power spectral density estimator \eqref{eq:MESApsd_appendix}.
 The method for solving the equations is called the Levinson-Durbin recursion \cite{doi:10.1002/sapm1946251261} and it is described in the following.
@@
 where $b$ is the forward prediction error coefficients at order $N-1$. 
-The 0-th order element can be easily initialized reminding that $a_0 = 1$ (always) and that $P_0$ can be determined from \ref{eq:errorFilter1}.
+The 0-th order element can be easily initialized reminding that $a_0 = 1$ (always) and that $P_0$ can be determined from \eqref{eq:errorFilter1}.
 Its values turns out to be: 
@@
 
-These expressions allow us to compute $\vec a$ and $P_N$ to any order by simply iterating \ref{eq:Levinson1} and \ref{eq:Levinson2}.
-Substituing them in equation \ref{eq:MESApsd_appendix} the problem of the estimation for the power spectral density via maximum entropy principle is solved.
+These expressions allow us to compute $\vec a$ and $P_N$ to any order by simply iterating \eqref{eq:Levinson1} and \eqref{eq:Levinson2}.
+Substituing them in equation \eqref{eq:MESApsd_appendix} the problem of the estimation for the power spectral density via maximum entropy principle is solved.
 Burg's method for spectral analysis is solved via Levinson is implmenteed in the released \texttt{memspectrum} package.

### 2021-03-24 climate_data/test_autocov.py
Test autocov

@@ -17,7 +17,7 @@ def empirical_autocorrelation(data):
 	corr = scipy.signal.correlate(data, data, mode="full")
 		#computing average value
 	lags = scipy.signal.correlation_lags(data.size, data.size, mode="full")
-	N_divide = len(data)-np.abs(lags) #number of data summed at each lag
+	N_divide = len(data)#-np.abs(lags) #number of data summed at each lag
 	return np.divide(corr,N_divide) #*np.mean(np.square(data)
 	
 
@@ -50,22 +50,12 @@ print(np.sum(M_zero.a_k), np.sum(M.a_k))
 
 print(np.allclose(spec_zero, spec), np.allclose(spec, spec_scaled))
 
-print(M.a_k, M_zero.a_k)
-print(M.P, M_zero.P)
-print(np.sum(M.a_k), np.sum(M_zero.a_k))
-print(np.allclose(M.a_k[:10],M_zero.a_k[:10]))
-plt.plot(M_zero.a_k)
-plt.plot(M.a_k)
-#plt.show()
-
-#quit()
-
 plt.figure()
 plt.loglog(f,spec_scaled*1/K**0, label = "symmetry")
 plt.loglog(f,spec_zero, label = 'zero mean')
 plt.loglog(f,spec, label = 'normal data')
 plt.legend()
-plt.show()
+#plt.show()
 
 
 autocov_empirical = empirical_autocorrelation(data-np.mean(data))
@@ -87,7 +77,7 @@ else:
 mu_avg_sq[len(data)] = -np.square(np.mean(data))
 #print("Mean: ", -np.mean(data[k:]+data[:-k]) * np.mean(data) +np.mean(data)**2, -np.mean(data)**2)
 
-autocorr_empirical += mu_avg_sq
+#autocorr_empirical += mu_avg_sq
 
 lags = scipy.signal.correlation_lags(data.size, data.size, mode="full")
 
@@ -109,13 +99,14 @@ plt.plot(lags[N_start:N_start+delta_T], autocov_mesa[:delta_T])
 
 plt.figure()
 plt.plot(lags[N_start:N_start+delta_T], autocov_mesa_zero[:delta_T]/autocov_mesa_zero[0])
-#plt.plot(lags[N_start:N_start+delta_T], autocorr_empirical[N_start:N_start+delta_T]-mu_avg_sq[N_start:N_start+delta_T])
[…]

### 2021-03-24 paper/mesa.tex
Change in tex

@@
 %Thus, in many real application, the knowledge of the PSD is crucial for the understanding of the process and indeed algorithms for PSD computation are a standard part of every software package for computational physics.
+\sschmidt{Spiega bene i due metodi di windowing!}
 A ubiquitous method for such computation is due to Welch \cite{Welch} and it based on the method of periodograms.
@@
 This will also provide a direct expression for the autoregressive coefficients $b_i$ and for the noise variance $\sigma^2$.
-We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~(\eqref{eq:AR_p}):
+We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~\eqref{eq:AR_p}:
 \begin{align}
@@
 \appendix
-\section{Computing the PSD} 
+\section{Details of PSD computation} 
 \subsection{MESA solution}\label{sec:MESA_solution}

### 2021-03-24 paper/mesa.tex
Added plots

@@
 
-\section{Application: Temperature Time Series} \label{sec:temp_timeseries}
+\section{Applications} \label{sec:applications}
+\subsection{Temperature Time Series} \label{sec:temp_timeseries}
 We apply our code to the study of the temperature time series.
@@
 On the other hand, the peak is captured well by the CAT loss function.
-This can be understood by looking at fig.~\ref{fig:p_vs_ptrue}.
+In analogy as what observed for the peak at $f_D$, in the CAT spectru we observe a higher order harmonic at a frequency $2 \cdot f_{yr}$.
+As in the previous case, this is required to better model the temperature variation in the year.
+
+The failure of FPE in catching the low frequency trend can be understood by looking at fig.~\ref{fig:p_vs_ptrue}.
 It is shown that CAT systematically select very long filters (i.e.
@@
 	\includegraphics{Images/climate_plots/forecast_accuracy.pdf}
+\end{figure}
+
+\subsection{Seismic waveforms} \label{sec:seismic_timeseries}
+\blindtext
+\begin{figure}
+	\caption{Do a caption}
+	\label{fig:seismic_spectrum}
+	\includegraphics{Images/seismic_plots/seismic_spectrum.pdf}
+\end{figure}
+
+\begin{figure}
+	\caption{Do a caption}
+	\label{fig:seismic_forecast}
+	\includegraphics{Images/seismic_plots/forecast_accuracy.pdf}
+\end{figure}
+
+\subsection{Forecasting the LIGO strain} \label{sec:LIGO_forecasting}
+\blindtext
+\begin{figure}
+	\caption{Do a caption}
+	\label{fig:LIGO_forecast}
+	\includegraphics{Images/forecast_LIGO/forecast_accuracy.pdf}
 \end{figure}

### 2021-03-24 scripts/plot_LIGO_forecast.py

@@ -1,0 +1,85 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from style_sheet import init_plotting
+
+import sys
+sys.path.insert(0,'..')
+import memspectrum
+
+import h5py
+
+plot_dir = '../paper/Images/forecast_LIGO/'
+
+#Time line: GPS:1164603392	UTC: 2016-12-01T04:56:15
+
+data = h5py.File('L-L1_GWOSC_O2_4KHZ_R1-1164603392-4096.hdf5', 'r')
+data = np.array(data['strain']['Strain'])
+
+	#spectrum computation
+srate = 4096 #in Hz
+T_train = 1000
+M = memspectrum.MESA()
+if False:
+	M.solve(data[:int(T_train*srate)], optimisation_method = 'CAT', method = 'Fast' , verbose = True, m = 100000, early_stop = True)
+	M.save('LIGO_model_CAT')
+else:
+	M.load('LIGO_model_CAT')
+
+	#Forecasting
+N_forecast = srate*10 #100 s of forecasting
+
+if False:
+	forecast = M.forecast(data[int(T_train*srate)-len(M.a_k):int(T_train*srate)],N_forecast, 100, verbose = True)
+	np.savetxt('forecast_LIGO_data.dat', forecast)
+else:
+	forecast = np.loadtxt('forecast_LIGO_data.dat')
+l,m,h = np.percentile(forecast,[5,50,95],axis = 0)
+sigma = np.std(forecast,axis = 0)
+true = data[int(T_train*srate):int(T_train*srate)+N_forecast]
+#quit()
+
+	#plot forecasting
+fig = init_plotting()
+plt.close()
+fig, (ax,ax2) = plt.subplots(nrows=2, sharex=True)
+
+t_grid = np.linspace(0,N_forecast/srate,N_forecast)
+ax.plot(t_grid, -(true-m), c = 'b', zorder = 1)
+ax.fill_between(t_grid, -true+h, -true+l, color = 'r',alpha = 0.3, zorder = 0)
+#ax.plot(t_grid, (true-m)/(h-l), c = 'b', zorder = 1)
+#ax.plot(np.linspace(0,N_forecast,N_forecast), true-h, c = 'r')
+#ax.plot(np.linspace(0,N_forecast,N_forecast), true- l, c = 'r')
[…]

### 2021-03-24 scripts/plot_climate_data.py

@@ -45,13 +45,13 @@ fig = init_plotting()
 ax = fig.gca()
 
 t_grid = np.linspace(0,N_forecast/24.,N_forecast)
-ax.plot(t_grid, (true-m), c = 'b', zorder = 1)
-ax.fill_between(t_grid, true-h, true-l, color = 'r',alpha = 0.3, zorder = 0)
+ax.plot(t_grid, -(true-m), c = 'b', zorder = 1)
+ax.fill_between(t_grid, -true+h, -true+l, color = 'r',alpha = 0.3, zorder = 0)
 #ax.plot(t_grid, (true-m)/(h-l), c = 'b', zorder = 1)
 #ax.plot(np.linspace(0,N_forecast,N_forecast), true-h, c = 'r')
 #ax.plot(np.linspace(0,N_forecast,N_forecast), true- l, c = 'r')
 ax.set_xlabel("Time (days)")
-ax.set_ylabel(r"$T - T_{forecast} (K)$")
+ax.set_ylabel(r"$T_{forecast} -T (K)$")
 plt.tight_layout()
 plt.savefig(plot_dir+"forecast_accuracy.pdf")
 
@@ -64,6 +64,7 @@ ax.loglog(f_CAT * unit_shift , spec_CAT, label = "CAT", c = 'b', zorder = 1)
 ax.loglog(f_FPE * unit_shift , spec_FPE, label = "FPE", c = 'r', zorder = 2)
 
 ax.axvline(1./(3600.*24.*365.)*unit_shift,lw = .5, c = 'b', ls = '--', zorder = 0)
+#ax.axvline(1./(3600.*24.*365.*0.5)*unit_shift,lw = .5, c = 'b', ls = '--', zorder = 0)
 ax.axvline(1./(3600.*24.)*unit_shift,lw = .5, c = 'b', ls = '--', zorder = 0)
 ax.set_xlabel("frequency (1/day)")
 ax.set_ylabel("PSD (1/Hz)")

### 2021-03-24 scripts/plot_seismic_data.py

@@ -1,0 +1,114 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from style_sheet import init_plotting
+
+import sys
+sys.path.insert(0,'..')
+import memspectrum
+
+from obspy import read
+
+plot_dir = '../paper/Images/seismic_plots/'
+
+st = read('knmi-fdsnws.mseed')
+
+print(st)
+print(st[1].stats)
+print(st[1].data)
+
+data = np.array(st[1].data)[:1000000]
+data = data/1e4 #pay attention to this scaling
+print("Len data (days): ", len(data)/40./(3600.*24))
+
+del st
+
+###################
+
+	#spectrum computation
+srate = 40. #in Hz 1/(1 hours)
+M_CAT = memspectrum.MESA()
+M_FPE = memspectrum.MESA()
+
+T_train = 10000
+
+if False:
+	M_CAT.solve(data[:int(T_train*srate)], optimisation_method = 'CAT', method = 'Standard' , verbose = True, m = 10000, early_stop = True)	
+	M_CAT.save('seismic_model_CAT')
+	M_FPE.solve(data[:int(T_train*srate)], optimisation_method = 'FPE', method = 'Standard' , verbose = True, m = 10000, early_stop = True)	
+	M_FPE.save('seismic_model_FPE')
+else:
+	M_CAT.load("seismic_model_CAT")
+	M_FPE.load("seismic_model_FPE")
+
+N = int(T_train*srate)
+spec_FPE, f_FPE = M_FPE.spectrum(1./srate)
+spec_FPE = spec_FPE[:int(N/2)]*1e8
+f_FPE = f_FPE[:int(N/2)]
+
+spec_CAT, f_CAT = M_CAT.spectrum(1./srate)
+spec_CAT = spec_CAT[:int(N/2)]*1e8
+f_CAT = f_CAT[:int(N/2)]
+
+print(len(M_CAT.a_k),len(M_FPE.a_k))
+
+	#Forecasting
+N_forecast = int(40.*60*10)
+
+if False:
+	forecast = M_CAT.forecast(data[int(T_train*srate) - len(M_CAT.a_k):int(T_train*srate)], N_forecast, 100, verbose = True)
+	np.savetxt('forecast_seismic_data.dat', forecast)
+else:
[…]

### 2021-03-25 paper/mesa.tex
Change in intro

@@
 or equivalently by the knowledge of their \emph{power spectral density} (PSD) $S(f)$. Thanks to the Wiener-Khinchin theorem the two are related by a Fourier transform: 
-\begin{equation}
+\begin{equation}\label{eq:psd-autocorrelation}
 	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau C(\tau) e^{-i 2 \pi f \tau}\,.
@@
 %Thus, in many real application, the knowledge of the PSD is crucial for the understanding of the process and indeed algorithms for PSD computation are a standard part of every software package for computational physics.
-\sschmidt{Spiega bene i due metodi di windowing!}
-A ubiquitous method for such computation is due to Welch \cite{Welch} and it based on the method of periodograms.
+
+An ubiquitous method for such computation is due to Welch \cite{Welch} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
 Basically, the PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many windowed batches and averaging the squared moduli of their Fourier transforms.
-The standard approach of periodograms \cite{Lomb,Scargle} consists in taking the Fourier Transform of the sample autocorrelation, written as \begin{equation}
-    \rho = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\}.
-\end{equation}
+This approach is equivalent \cite{Lomb,Scargle} to taking the Fourier Transform of the sample autocorrelation $\rho$, written as
+\begin{equation}
+    \rho_{Windowed} = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\},
+\end{equation}
+where $M$ is the maximum time lag at which the autocorrelation is computed.
 The sequence $W$ is a window function that can be chosen in several different way, each choice presenting advantages and disadvantages for the final estimate of the PSD.
-\sschmidt{Here we say different things about periodigrams.
-We should be more clear...}
 
@@
 A high frequency resolution implies high variance and vice-versa.
-Moreover, the final result depends on the choice of window function, hence begging the question of what the ``real" PSD is.
-\\
-Another problem with this approach is the requirement for the window to be $0$ outside the interval: we are arbitrarily assuming $\rho_j = 0$ for $j > M$, and modifying the estimate if a non-rectangular window is chosen.
-Making assumptions on unobserved data and modifying the ones we have at our disposal introduces information about the process that is not really there.
-Moreover, this method requires a number of arbitrary choices to be made, e.g.
-the number of time slices, the kind of Window function, the overlap between consecutive slices, etc...
+Besides the window function, Welch method requires a number of arbitrary choices to be made, as the number of time slices and the overlap between consecutive slices.
 All the knobs must be tuned by hand and their choice can dramatically affect the PSD estimation.
-\par
+The final PSD estimation depends on all such choices, hence begging the question of what the ``real" PSD is.
+
+Another problem with this approach is the requirement for the window to be $0$ outside the interval in which the autocorrelation is computed.
+We are arbitrarily assuming $\rho_j = 0$ for $j > M$ and modifying the estimate (i.e. the data) if a non-rectangular window is chosen.
+Making assumptions on unobserved data and modifying the ones we have at our disposal introduces a piece of ``spurious" information about the process that we are not allowed to use.
+
 An appealing alternative approach, based on the Maximum Entropy principle \cite{Jaynes}, has been put forward by Burg \cite{burg1975maximum}.
-Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the arbitrary parameters, as it provides an iterative closed form expression for the spectrum of a stochastic time series.
+Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the arbitrary parameters, as it provides an iterative closed form expression for the spectrum of a stochastic stationary time series.
 Furthermore, it embeds the PSD estimation problem into an elegant theoretical framework and makes minimal assumptions on the nature of the data.
@@
 In this paper we discuss the details of the Maximum entropy principle, its application to Power Spectral Density estimation with Burg algorithm and the link between Burg's algorithm and autoregressive process.
-Furthermore, we present and describe a new code, \texttt{memspectrum}, that provides a robust python implementation of the algorithm.
+Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many problem posed by the ubiquitous Welch algorithm (or other similar methods).
+To facilitate this goal, we present and describe a new code, \texttt{memspectrum}, that provides a robust python implementation of the algorithm.
+The code is ready to use and available at the PyPI repository\footnote{
+Available at link: \url{https://pypi.org/project/memspectrum/}.
+}
 We provide a thorough assessment of the performance of our code and we validate our results performing a number of tests.
-We compare our results with those of spectral analysis carried out with the standard Welch method.
-\par
+We also compare our results with those of spectral analysis carried out with the standard Welch method.
+In order to apply our model on a realistic setting, we analyse some time series of broad interest in the scientific community.
+
 Our paper is organized as follows: blablabla...

### 2021-03-25 paper/mesa.tex
LIGO forecast section

@@
 \subsection{Forecasting the LIGO strain} \label{sec:LIGO_forecasting}
-\blindtext
+As a last example of and application of the maximum entropy method on real data, we come back to the strain produced by the LIGO-Virgo interferometers and we forecast the future observations.
+The spectral analysis is performed with CAT and FPE loss functions on $\SI{1000}{s}$ of data \cite{Abbott_2021} sampled at $\SI{4096}{Hz}$ observed at Livingston observatory starting from GPS time $1164603392$.
+We then forecast the following $\SI{100}{s}$ of observations with the model optimized with CAT. The results are shown in fig.~\ref{fig:LIGO_forecast}.
+
+By looking at the figure, we note that the forecasted prediction is always in the $90\%$ confidence interval.
+The confidence interval, however, is very broad and the prediction is not very accurate (being the same order of magnitude of the strain).
+By looking at the standard deviation $\sigma$ of the predictions (bottom panel in fig.~\ref{fig:LIGO_forecast}), we note that it increases very quickly in the first $\SI{0.5}{s}$ and then its trend stabilizes to a constant value.
+The order of the autoregressive process selected is $p_{CAT} = 57766 \sim \SI{14}{s}$, which is much larger than the region in which $\sigma$ is small $\sim \SI{0.2}{s}$.
+This means that the series is very difficult to predict: the knowledge of past observations is not very helpful in predicting future observations.
+Similar performances in the sigma are observed also with the FPE model: in this case FPE selects a large prediction filter $p_{FPE} = 29924$ , similar to those selected by CAT.
+
+Despite poor predictions on long timescales, the autoregressive process obtained with MESA still can be useful on short timescales.
+Indeed, a precise prediction of the strain time series can be beneficial in anomaly detection and removal of glitches.
+The predictions can form an expected baseline for the strain; any anomaly (i.e.
+a glitch or even a transient of physical origin) can show up as a large departure from expected trend.
+Moreover, if a glitch is detected, its shape can be estimated (as well as the confidence level) by subracting the expected signal with the actual signal.
+This can work: a typical glitch \cite{Zevin_2017} can last as long as $\SI{0.2}{s}$, which is close to the time scale in which the forecasting is realiable.
+
+Of course, more sophisticated forecasting methods are available in the Machine Learning literature (see e.g.
+\cite{LSTM_article, oord2016wavenet}) and they can be trained for precision forecasting.
+However, a Maximum-Entropy trained autoregressive process can be a simple baseline for comparing such more advanced methods and might suffice for many purposes.
+This opens a promising path in GW data analysis and other fields of physics might also take advantage from this.
+
 \begin{figure}
-	\caption{Do a caption}
+	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
+The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
+The confidence interval is computed on  $100$ independent random realization of the predictions.
+	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
+	}
 	\label{fig:LIGO_forecast}
@@
 \blindtext
+
+	\begin{acknowledgments}
+This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center (https://www.gw-openscience.org/ ), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration.
+LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS), and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector.
+Additional support for Advanced LIGO was provided by the Australian Research Council.
+Virgo is funded, through the European Gravitational Observatory (EGO), by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef, with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain.
+	\end{acknowledgments}
 

### 2021-03-25 scripts/plot_LIGO_forecast.py

@@ -19,24 +19,34 @@ data = np.array(data['strain']['Strain'])
 srate = 4096 #in Hz
 T_train = 1000
 M = memspectrum.MESA()
+M_FPE = memspectrum.MESA()
 if False:
 	M.solve(data[:int(T_train*srate)], optimisation_method = 'CAT', method = 'Fast' , verbose = True, m = 100000, early_stop = True)
 	M.save('LIGO_model_CAT')
+	M_FPE.solve(data[:int(T_train*srate)], optimisation_method = 'FPE', method = 'Fast' , verbose = True, m = 100000, early_stop = True)
+	M_FPE.save('LIGO_model_FPE')
 else:
 	M.load('LIGO_model_CAT')
+	M_FPE.load('LIGO_model_FPE')
 
 	#Forecasting
 N_forecast = srate*10 #100 s of forecasting
 
 if False:
 	forecast = M.forecast(data[int(T_train*srate)-len(M.a_k):int(T_train*srate)],N_forecast, 100, verbose = True)
-	np.savetxt('forecast_LIGO_data.dat', forecast)
+	np.savetxt('forecast_LIGO_data_CAT.dat', forecast)
+	forecast_FPE = M_FPE.forecast(data[int(T_train*srate)-len(M.a_k):int(T_train*srate)],N_forecast, 100, verbose = True)
+	np.savetxt('forecast_LIGO_data_FPE.dat', forecast_FPE)
 else:
-	forecast = np.loadtxt('forecast_LIGO_data.dat')
+	forecast = np.loadtxt('forecast_LIGO_data_CAT.dat')
+	forecast_FPE = np.loadtxt('forecast_LIGO_data_FPE.dat')
 l,m,h = np.percentile(forecast,[5,50,95],axis = 0)
 sigma = np.std(forecast,axis = 0)
+sigma_FPE = np.std(forecast_FPE,axis = 0)
 true = data[int(T_train*srate):int(T_train*srate)+N_forecast]
 #quit()
+
+print("Model length: ",len(M.a_k), len(M_FPE.a_k))
 
 	#plot forecasting
 fig = init_plotting()
@@ -51,14 +61,15 @@ ax.fill_between(t_grid, -true+h, -true+l, color = 'r',alpha = 0.3, zorder = 0)
[…]

### 2021-03-25 scripts/plot_seismic_data.py

@@ -8,6 +8,8 @@ import memspectrum
 
 from obspy import read
 
+#http://rdsa.knmi.nl/fdsnws/dataselect/1/builder
+
 plot_dir = '../paper/Images/seismic_plots/'
 
 st = read('knmi-fdsnws.mseed')

### 2021-03-25 paper/mesa.tex
Added and fixed refs

@@
 
-An ubiquitous method for such computation is due to Welch \cite{Welch} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
+An ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
 Basically, the PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many windowed batches and averaging the squared moduli of their Fourier transforms.
@@
 
-An appealing alternative approach, based on the Maximum Entropy principle \cite{Jaynes}, has been put forward by Burg \cite{burg1975maximum}.
+An appealing alternative approach, based on the Maximum Entropy principle \cite{JaynesArticle,jaynes2003ptl}, has been put forward by Burg \cite{burg1975maximum}.
 Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the arbitrary parameters, as it provides an iterative closed form expression for the spectrum of a stochastic stationary time series.
@@
 where the matrix $S = S_i \delta_{ij}$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
-Many readers will recognise the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\cite{woodward}
+Many readers will recognise the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\cite{prob_information_theory}
 %P. M. Woodward, D. W. Fry, and W. Higinbotham, Probability and information theory, with applications to radar; 2nd ed. (Pergamon Press, Oxford, 1964), URL http://cds.cern.ch/record/2031792.
-and of gravitational waves data analysis, e.g. \cite{finn:1992}.
-% L. S. Finn, Phys. Rev. D46, 5236 (1992), gr-qc/9209010
+and of gravitational waves data analysis, e.g. \cite{Finn_1992, FINDCHIRP}.
 Thanks to MAXENT, the problem of defining the probability distribution describing a wide-sense stationary process is thus 
@@
 The Burg's algorithm computes the autoregressive coefficients that are suitable to the available data.
-Thanks to Wold's theorem \cite{something}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general.
+Thanks to Wold's theorem \cite{Wold_theorem}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general.
 
@@
 We test the performance of the three loss functions on a random time series generated with a known power spectral density.
-In this first experiment, we take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$, see Fig.~\ref{fig:autocorr}.
-The time series are generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$, following the procedure outlined in~\cite{}, and performing an inverse Fast Fourier Transform on the frequency series.
-For this investigation, we generate a dataset of 1000 time series of 3000 points each, see Fig.~\ref{fig:noise} for an example of a time series in frequency and time domain.
+In this first experiment, we take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$.
+The time series are generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$, following the procedure outlined in sec.~\ref{sec:forecasting}, and performing an inverse Fast Fourier Transform on the frequency series.
+For this investigation, we generate a dataset of 1000 time series of 3000 points each.
 
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/error_length_contour.pdf}
-	\caption{Relative error $r$ (as in eq.~\eqref{eq:freq_error}) vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with gaussian PSD given in Fig.~\ref{fig:autocorr}.
+	\caption{Relative error $r$ (as in eq.~\eqref{eq:freq_error}) vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with a gaussian PSD.
 Different colours refers to different choices for the loss function: in red $\mathcal{L}_{\rm FPE}(m)$, in green $\mathcal{L}_{\rm OBD}(m)$ and in black $\mathcal{L}_{\rm CAT}(m)$.
@@
 We perform two sets of experiments.
-In the first one (see fig.~), we simulate data from an analytical PSD and we make the comparison with them: this is to ensure the we have a baseline PSD to compare the data with.
-In a second experiment (see fig.~), we employ the data released to the public byt the LIGO/Virgo collaboration.
+In the first one (see fig.~\ref{fig:mem_welch_fakedata}), we simulate data from an analytical PSD and we make the comparison with them: this is to ensure the we have a baseline PSD to compare the data with.
+In a second experiment (see fig.~\ref{fig:mem_welch_realdata}), we employ the data released to the public by the LIGO/Virgo collaboration \cite{Abbott_2021}.
 In each experiment, we vary the length of the data used for the estimation: this is also useful to assess how the computation depends on the data available.
@@
 \begin{figure}
+	\label{fig:mem_welch_fakedata}
+	\caption{Make here a comparison between Welch - Fake data}
+	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_True.pdf}
+\end{figure}
+\begin{figure}
 	\label{fig:mem_welch_realdata}
@@
 	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_False.pdf}
-\end{figure}
-\begin{figure}
-	\label{fig:mem_welch_fakedata}
-	\caption{Make here a comparison between Welch - Fake data}
-	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_True.pdf}
 \end{figure}
@@
 We will also assess the quality of the forecasting predictions: this task has a great impact also on practical applications.
-We downloaded the historical reanalysis data from the Copernicus project (or whathever it is) \cite{}.
+We used the historical reanalysis data from the ``ERA5-Land hourly data from 1981 to present" dataset, downloaded from the Climate Data Store \cite{ERA5_dataset}
+
 Data refers to the temperature measure at coordinates $\textrm{N}\;\ang{45;5;} \;\; \textrm{E}\;\ang{9;1;}}$, corresponding about to the city of Milan, Italy.
@@
 We clearly see that there is a peak on the PSD at the frequency $f_D = \SI{1}{day^{-1}}$ corresponding to a day. This is somehow trivial and highly expected.
-Higher order harmonics (corresponding to integer multiples of $f_D$) are also visible up to the Nyquist frequency ($f_D = \SI{0.5}{hour^{-1}}$): they are required to explain the high variability of the daily temperature cycle throughout the year.
+Higher order harmonics (corresponding to integer multiples of $f_D$) are also visible up to the Nyquist frequency ($f_D = \SI{0.5}{hour^{-1}}$): they corresponds to signals that preserve the 1 day period while, at the same time, capturing the complex variability of the daily temperature cycle throughout the year.
 
@@
 \blindtext
+
+\cite{seismic_data}
+
+\blindtext
+
 \begin{figure}
@@
 	\begin{acknowledgments}
-This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center (https://www.gw-openscience.org/ ), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration.
+\sschimidt{Is it fine to ackonowledge people helping with temperature data and seismic data?}
+%S.S. wants to thank Maurizio Maugeri and Costanza Rossi for useful discussions.
+
+This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center (\url{https://www.gw-openscience.org/}), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration.
 LIGO Laboratory and Advanced LIGO are funded by the United States National Science Foundation (NSF) as well as the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS), and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector.

### 2021-03-25 scripts/plot_climate_data.py

@@ -69,7 +69,7 @@ ax.axvline(1./(3600.*24.)*unit_shift,lw = .5, c = 'b', ls = '--', zorder = 0)
 ax.set_xlabel("frequency (1/day)")
 ax.set_ylabel("PSD (1/Hz)")
 
-ax.legend(loc = 'upper left')
+ax.legend(loc = 'upper left')questo
 
 
 	#inset

### 2021-03-25 paper/mesa.tex
Tiny change

@@
 We perform two sets of experiments.
-In the first one (see fig.~\ref{fig:mem_welch_fakedata}), we simulate data from an analytical PSD and we make the comparison with them: this is to ensure the we have a baseline PSD to compare the data with.
+In the first one (see fig.~\ref{fig:mem_welch_fakedata}), we simulate data from the realesed LIGO intererferometer PSD and we make the comparison with them: this is to ensure the we have a baseline PSD to compare the data with.
 In a second experiment (see fig.~\ref{fig:mem_welch_realdata}), we employ the data released to the public by the LIGO/Virgo collaboration \cite{Abbott_2021}.
@@
 \begin{figure}
+	\caption{Make here a comparison between Welch - Fake data}
 	\label{fig:mem_welch_fakedata}
-	\caption{Make here a comparison between Welch - Fake data}
 	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_True.pdf}
 \end{figure}
+
 \begin{figure}
+	\caption{Make here a comparison between Welch - True data}
 	\label{fig:mem_welch_realdata}
-	\caption{Make here a comparison between Welch - True data}
 	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_False.pdf}
 \end{figure}
-
-\sschmidt{Mettiamo da qualche parte una descrizione dettagliata del codice? No: solo rimandiamo al codice nell'intro}
 
@@
 
+\sschimidt{Is it fine to ackonowledge people helping with temperature data and seismic data?}
 	\begin{acknowledgments}
-\sschimidt{Is it fine to ackonowledge people helping with temperature data and seismic data?}
 %S.S. wants to thank Maurizio Maugeri and Costanza Rossi for useful discussions.

### 2021-03-25 paper/mesa.tex
Edit in caption welch comparison

@@
 \begin{figure}
-	\caption{Make here a comparison between Welch - Fake data}
+	\caption{Comparison between PSD estimation with Welch method (blue) and Maximum Entropy (red) for \textit{synthetic} LIGO data.
+The estimation is based data with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.
+In dashed black is reported the true PSD (from which data are generated).}
 	\label{fig:mem_welch_fakedata}
@@
 \begin{figure}
-	\caption{Make here a comparison between Welch - True data}
+	\caption{Comparison between PSD estimation with Welch method (blue) and Maximum Entropy (red) for \textit{real} LIGO data.
+The estimation is based data with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
 	\label{fig:mem_welch_realdata}
@@
 On the other hand, the peak is captured well by the CAT loss function.
-In analogy as what observed for the peak at $f_D$, in the CAT spectru we observe a higher order harmonic at a frequency $2 \cdot f_{yr}$.
+In analogy as what observed for the peak at $f_D$, in the CAT spectrum we observe a higher order harmonic at a frequency $2 \cdot f_{yr}$.
 As in the previous case, this is required to better model the temperature variation in the year.

### 2021-03-25 paper/mesa.tex
Edits in temperature&LIGO forecasting

@@
 We note the observed difference is always well included in the $90\%$ confidence interval: the forecasting predictions seem reliable.
-On the other hand, the confidence interval is pretty large (almost as large as $\SI{15}{K}$), thus making easy for the actual data to fit the predictions.
-We believe that the origin of such large uncertainty is due to the complexity of the time series considered: any realistic model, even more complex than our autoregressive process, might not produce drastically lower uncertainties (in such a long term).
+On the other hand, the confidence interval is pretty large (almost as large as $\SI{15}{K}$), thus making ``easy" for the actual data to fit the predictions.
+Indeed, the prediction model, while suitable for spectral estimation, is nothing more than a linear regression (plus noise term).
+Such a simple model hardly catches the complex trend in the variability of the temperature daily trend during a year.
+For more precise predictions, probably one should consider nonlinear regression model, tapping into the wealth of non linear predictors offered by the field of Deep Lerning.
 
@@
 
-Of course, more sophisticated forecasting methods are available in the Machine Learning literature (see e.g.
+As noted above, for pushing further the prediction performance from our simple linear predictor, more sophisticated forecasting methods are available in the Machine Learning literature (see e.g.
 \cite{LSTM_article, oord2016wavenet}) and they can be trained for precision forecasting.

### 2021-03-25 paper/mesa.tex
Draft for the conclusions

@@
 The result does not change much by changing the hyperparameters required for Welch algorithm.
-\sschmidt{Other observations??}
+\sschmidt{Elaborate a bit more on this... Right now it is a very poor comment (blame is on me!)}
 
@@
 \section{Final remarks and future prospects} \label{sec:conclusion}
-\blindtext
-\blindtext
-\blindtext
-
-\sschimidt{Is it fine to ackonowledge people helping with temperature data and seismic data?}
+
+We presented a case study of the application of Maximum Entropy principle in the realm of spectral estimation: it provides a closed form estimation of the power spectral density, bridging the spectral analysis with the study of autoregressive processes.
+The method has the towfold advantage of avoiding any unjustified assumption on the observed time series and of providing a simple context for forecasting the time series.
+We provided a ready to use python code, called `memspectrum' and we performed numerical studies of its performance.
+We investigated the use of different loss functions for the purpose of correctly estimating the autoregressive order for the estimation and we found that different loss functions are required for different situations.
+Finally, we performed a comparison between PSDs computed with MESA and standard Welch's algorithm and we applied our method to several time series of interest in physics.
+
+As outlined in sec.~\ref{sec:Welch_comparison}, the MESA provides PSD estimates with less noise than the standard Welch algorithm, thus making it a good alternative for it.
+The use of MESA is particularly useful for short time series samples, where Welch method is outperformed in accuracy and conficence.
+It has to be said that our method can be quite computationally expensive, as compared to standard Welch's algorithm. However, for short time series, the computational cost is still managable.
+
+Besides providing an alternative to standard methods of spectral estimate, MESA opens up several oppurtunities in the context of time series analysis.
+
+First of all, as short time series are required to produce meaningful PSD estimations, it can be suitable in the context of PE...
+\textbf{##############PE part here#############}
+
+Furthermore, MESA provides a simple, but robust and quite accurate, predictor for the time series. This fact is remarkable and can be used in time series analysis for several purposes.
+As already discussed in sec~\ref{sec:LIGO_forecasting}, an anomaly detection pipeline can be easily built using the forecasts of MESA: the predictions can form a baseline to compare the actual observations with.
+Whenever the observed data are outside the expectations, an anomaly detection can be claimed.
+Of course such predictions can be done with a more accurate (perhaps nonlinear model); however MESA has the advantage of being simple and fast to train, while providing good (short term) predictions.
+
+\sschmidt{Other stuff with forecasting?}
+
+We believe that Maximum Entropy Spectral Analysis, despite being an ``old" theory, has still the potential to become the standard in future time series analysis, due to its theorical elegance and to its reliable performance in a vast range of application.
+
 	\begin{acknowledgments}
-%S.S. wants to thank Maurizio Maugeri and Costanza Rossi for useful discussions.
+%We want to thank Maurizio Maugeri and Costanza Rossi for useful discussions.
+\sschmidt{Is it fine to ackonowledge people helping with temperature data and seismic data?}
 

### 2021-03-25 paper/mesa.tex
Modified theoretical foundations

@@
 \subsection{Maximum Entropy Principle} \label{sec:MAXENT}
-
-\sschmidt{This section might be simplified...}
-
-Before introducing MAXENT as a variational problem, we will workout some simple examples to contexualise how information's entropy
-and `uncertainty' are related.
-First of all,  it is worth noting that the word `information' will not be used to indicate our knowledge on the system under study: this is what we will call `evidence'.
-Information will be intended as in communication theory: it specifies the length of the message necessary to provide
-a full description of the system under study.
-
+Before introducing MAXENT principle and its formulation, we will workout some simple examples to specify what the core quantities of the problem are and the 
+role they play.
+These are the 'evidence' and the 'information'.
+Let us start with the 'information' for one or more events: it is a measure of the degree of uncertainty on the outcomes of some experiment and specifies the length of the message necessary to provide a full description of the system under study.
 For instance, consider a perfectly sinusoidal signal: knowledge of amplitude, frequency and phase are sufficient to fully reproduce it:
@@
 to represent the quantity of information brought by an outcome $x$ with probability $p(x)$.
-It is additive quantity as well as monotonically decreasing as a function of $p \in [0, 1]$.
-The more uncertain the outcome, the higher the information it brings.
-
+It is additive quantity as well as monotonically decreasing as a function of $p \in [0, 1]$: the more uncertain the outcome, the higher the information it brings.
 We now seek to generalize the measure for the information in the case where different outcomes are possible.
@@
 \end{equation}
-this describes a situation of `maximum ignorance'; we have no evidence at all and we have no way of predicting what the outcome of such an experiment will be.
-The information content of this system must be high.
-
-Of course, when assigning probabilities for a set of events, all the evidence available must be taken into account.
-On the other hand, the probability assignment must not introduce additional evidence in the system and thus the need to be ``the most uncertain possible given the knoledge of the system".
-
-%In this context, the role played by evidence should be clear: it dictates the probability assignment in order to make the result `more certain' or `less informative'.
-Any assignment different from the `most uncertain' is violating the principle of it to being non committal to non-available evidence.
-We can only assign a more certain probability, i.e.
-less informative, if we have evidence for it.
-If we have none, the result has to be the most unpredictable, i.e.
-the most informative.
-From the point of view of communication theory: the higher the evidence on a system, the lower the information I need to send to describe it.
+this describes a situation of `maximum ignorance':  the information content of this system must be high. If an experiments with $N$ outcomes
+is  considered, the most uncertain experiment is the one in which 
+\begin{equation}\label{eq:indifference}
+P_1 = \dots = P_N = \frac{1}{N}.
+\end{equation}
+In the above cases our knowledge on the system was the total number of different outcomes.
+Let us now reverse the perspective: what is the most unbiased probability assignment we can make if all we know is that one over $N$ outcomes will occur?
+Since we have no reason to prefere one outcome over the others, the least biased assignment is the one equivalent to (\ref{eq:indifference}) and is found to be coincident with the most uncertain experiment.
+This result represents the core idea of MAXENT principle: the least biased probability assignment can be obtained maximizing the "expected" information's entropy.
+\\
+In the previous example we had no reason to prefere one outcome over the others but that's not always the case.
+This hypothesis can be relaxed introducing the concept of 'evidence', that represents the knowledge we posses on the system under study.
+With proper evidence some outcomes can be preferred to others.
+Probability assignment must obey to all the evidence we possess.
 \par
@@
 This defines a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized.
-$H$ has to be a `generalization' of information defined in equation \eqref{eq:information} to be something like the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
-%
-%Some properties for $H$ can be worked out easily. 
-%From previous discussion, given a fixed number $N$ of possible outcomes, we want $H$ to be maximized by the choice $p_1 = \dots = p_N$.
-%Consider now the opposite situation, in which probability is assigned appealing to IP and we have a different number of possible outcomes.
-%Intuitively, the most unpredictable experiment is the one with larger number of outcomes: probability of "getting it wrong" increases as $\frac{N -1}{N}$.
-So when equal probabilities are assigned, we want
-%\begin{equation}
-%    \nonumber
-%    H\left[\frac{1}{N}, \dots, \frac{1}{N}\right]
-%\end{equation}
-%to increase monotonically with N.
-%
-%We also want our function to be continuous in its parameters: any small change in probability has to slightly affect uncertainty. \\
-%Last is composition property. Consider three events with probability $p_1, p_2, p_3$ such that $p_2 + p_3 = 1 - p_1 = q$. We require 
-%\begin{equation}
-%    H\left(p_1, p_2, p_3\right) = H(p_1, q) + qH\left(\frac{p_2}{q}, \frac{p_3}{q}\right),
-%\end{equation}
-%so that uncertainty is additive.
+$H$ has to be a `generalization' of information defined in equation \ref{eq:information} to be something like the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
 
@@
 and $H$ is what we will call (information) entropy.
+
+The maximisation of entropy, supplemented by evidence in the form of constraints to which the sought-for probability distribution must obey,
+gives rise to several of the most common probability distributions utilised commonly by data scientists.
+For instance, whenever the only constraint available is the normalisation of the probability distribution, the entropy is maximised by the uniform distribution.
+If the information available is instead the expected value -- the mean -- entropy is maximised by the  exponential distribution
 

### 2021-03-25 scripts/compare_LVC_data.py
New plots for Welch comparison

@@ -9,7 +9,8 @@ from style_sheet import *
 import pandas as pd
 from scipy.interpolate import interp1d
 
-plot = True
+plot = False
+plot_same = True
 compute = False
 generate_fake_data = False
 use_fake_data = False
@@ -59,7 +60,59 @@ if compute:
 		PSD_MESA = M.spectrum(1./srate, freqs)
 		
 		np.savetxt("plot_data/plot_{}_{}.txt".format(T, use_fake_data), np.column_stack([freqs, PSD_MESA, PSD_Welch]))
+
+if plot_same:
+	if use_fake_data:
+		true_PSD = np.loadtxt(true_PSD_file)
+
+	fig = init_plotting()
+	fig.clf()
+	fig.set_size_inches(3.4,3.4*2)
+	ax = fig.gca()
+
+	ax.set_xlabel(r"$f(Hz)$")
+	ax.set_ylabel(r"$PSD (a.u.)$") #No ylabel
 	
+	yticks = []
+	
+	for i, T in enumerate(T_list):
+		PSDs = np.loadtxt("plot_data/plot_{}_{}.txt".format(T, use_fake_data))
+		N = PSDs.shape[0]
+		freqs, PSD_MESA, PSD_Welch = PSDs[:int(N/2),0], PSDs[:int(N/2),1], PSDs[:int(N/2),2]
+
+		offset = .1
+		if use_fake_data: step = 4.2
+		else: step = 4.
+		
+			#scaling PSDs
+		PSD_MESA =PSD_MESA/ 1e-41 *10**(step*i)*offset
+		PSD_Welch = PSD_Welch/ 1e-41 * 10**(step*i)
+
+		ax.loglog(freqs, PSD_Welch , c = 'g', zorder = 0, label = "Welch")
+		ax.loglog(freqs, PSD_MESA , c = 'r', zorder = 1, label = "MESA")
+		
+		yticks.append(10**(step*i)/1000.)
+		print(yticks)
+		
+		if i == 0:
+			ax.legend(loc = 'upper center', ncol = 2+int(use_fake_data))
+		
+		if use_fake_data:
+			ax.loglog(true_PSD[:,0], true_PSD[:,1]/1e-41*10**(step*i), '--', c = 'k', zorder = 2)
+			ax.loglog(true_PSD[:,0], true_PSD[:,1]/1e-41*10**(step*i)*offset, '--', c = 'k', zorder = 2, label = "True")
+		
+	ax.set_xlim(20,1024)
[…]

### 2021-03-26 paper/mesa.tex
Tiny change in paper
Added PE to conclusions

@@
 
-The maximisation of entropy, supplemented by evidence in the form of constraints to which the sought-for probability distribution must obey,
-gives rise to several of the most common probability distributions utilised commonly by data scientists.
-For instance, whenever the only constraint available is the normalisation of the probability distribution, the entropy is maximised by the uniform distribution.
-If the information available is instead the expected value -- the mean -- entropy is maximised by the  exponential distribution.
-
 Of particular importance is the case in which, in addition to the mean, also the variance is known. MAXENT leads to the Gaussian distribution. 
@@
 probability distribution:
-\begin{align}
+\begin{align}\label{eq:Whittle}
 p&\left((\tilde{x}_1, \ldots, \tilde{x}_n)\vert I\right) = \nonumber \\
@@
 
-First of all, as short time series are required to produce meaningful PSD estimations, it can be suitable in the context of PE...
-\textbf{##############PE part here#############}
+First of all, as short time series are required to produce meaningful PSD estimations, it can be suitable in the context of parameter estimation.
+To compute the posterior distribution for the parameters the knowledge of prior distributions and Likelihood for noise is required.
+A full description for noise can be obtained in Fourier Domain by the choice of the Whittle Likelihood \ref{eq:Whittle}, that is completely determined by the knowledge of the PSD for noise.
+While traditional methods might be instable with respect to the hypothesis (such as Welch) or require a Large number of prior assumptions, MESA represents a stable and parsimonious way to characterize noise and compute posterior distributions.
 

### 2021-03-26 paper/mesa.tex
Updated MESA-Welch Comparison

@@
 This fact is more important at lower frequencies, where the natural time length is large.
-
-Turning to the comparison between MESA and Welch's method, we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
-has a large number of spurious peaks) rather than PSD provided by MESA.
-The different behavior is more manifest in the low frequency sector.
+We can also note that MESA converges to the real spectrum faster than Welch method, providing a better estimate expecially for short time series.
+This behaviour is more evident if the low frequencies are considered.
+An accurate profile reconstruction can be obtained with MESA using a 5 seconds-strain only, while Welch method requires at least 10 seconds of data to obtain a comparable profile.
+Another important element is noise: we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
+has a large number of spurious peaks) rather than PSD provided by MESA with FPE loss function.
+If longer datasets are considered, even if MESA is found to be a little more accurate and less noisy than Welch, all of the above differences become negligible.
+As we already said Welch method is very instable with respect to the windowing choices we made.
+We tried to find the best compromize between noise and accuracy for the reconstruction, but different choices can be made.
+We want to stress that this fact doesn't invalidate our discussion: a more accurate reconstruction can only be obtained introducing a greater amount of noise: a greater accuracy requires longer sub-segments of data, reducing the number of spectra over which we average.
+Reducing noise require the opposite choices, and we can conclude that MESA represents a better compromise between noise and accuracy than Welch's method.
 The result does not change much by changing the hyperparameters required for Welch algorithm.
-\sschmidt{Elaborate a bit more on this... Right now it is a very poor comment (blame is on me!)}
 

### 2021-03-26 paper/mesa.tex
Review intro & Welch comparison

@@
 \subsection{Maximum Entropy Principle} \label{sec:MAXENT}
-Before introducing MAXENT principle and its formulation, we will workout some simple examples to specify what the core quantities of the problem are and the 
-role they play.
-These are the 'evidence' and the 'information'.
-Let us start with the 'information' for one or more events: it is a measure of the degree of uncertainty on the outcomes of some experiment and specifies the length of the message necessary to provide a full description of the system under study.
-For instance, consider a perfectly sinusoidal signal: knowledge of amplitude, frequency and phase are sufficient to fully reproduce it:
-it has a low information entropy content. Even less information is required if we are studying a system whose outcome is 
-certain (has probability $p = 1$), as in this case, a communication is not even needed.  
+
+Before introducing MAXENT principle, we will define through some simple examples the two core concepts of the problem and the role they play: the 'evidence' and the 'information'.
+Let us start with the 'information' (or entropy): it is a measure of the degree of uncertainty on the outcomes of some experiment and specifies the length of the message necessary to provide a full description of the system under study.
+For instance, consider a perfectly sinusoidal signal: knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information content.
+Even less information is required if we are studying a system whose outcome is certain (has probability $p = 1$), as in this case, a communication is not even needed.
 Shannon \cite{Shannon} proposed the quantity
@@
 \end{equation}
-to represent the quantity of information brought by an outcome $x$ with probability $p(x)$.
+to measure the quantity of information brought by an outcome $x$ with probability $p(x)$.
 It is additive quantity as well as monotonically decreasing as a function of $p \in [0, 1]$: the more uncertain the outcome, the higher the information it brings.
-We now seek to generalize the measure for the information in the case where different outcomes are possible.
-To do so, let us consider an experiment with two different random outcomes $E_1, E_2$, with given probabilities $P_1$ and $P_2$, and we ask ourselfs which are the probability assignments that makes the outcome more uncertain (i.e.
-maximize the information).
-If $P_1$ and $P_2$ are largely different, for istance $P_1 = 0.999$ and $P_2 = 0.001$, we clearly have enough evidence to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome.
+
+We can generalize the definition of information in the case where two different outcomes $E_1, E_2$, with given probabilities $P_1$ and $P_2$, are possible.
+To gain some intuition on the problem, we ask ourselfs which are the probability assignments that makes the outcome more uncertain (i.e. maximize the information).
+If $P_1$ and $P_2$ are largely different, for istance $P_1 = 0.999$ and $P_2 = 0.001$, we are allowed to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome.
 The information content will be very low.
@@
 \end{equation}
-this describes a situation of `maximum ignorance':  the information content of this system must be high. If an experiments with $N$ outcomes
-is  considered, the most uncertain experiment is the one in which 
-\begin{equation}\label{eq:indifference}
-P_1 = \dots = P_N = \frac{1}{N}.
-\end{equation}
-In the above cases our knowledge on the system was the total number of different outcomes.
-Let us now reverse the perspective: what is the most unbiased probability assignment we can make if all we know is that one over $N$ outcomes will occur?
+this describes a situation of `maximum ignorance': the information content of this system must be high.
+Any generalization of eq.~\eqref{eq:information}, must then have its maximum when $ P_1 = P_2$.
+For $N$ events, system with the highest possible information content is when:
+\begin{equation}\nonumber
+    P_1 = \hdots = P_N = \frac{1}{N}:
+\end{equation}
+
+Shannon \cite{Shannon} showed that the only functional form satisfying continuity with respect to its parameters, additivity and that has a maximum for equal probability events is:
+\begin{equation}\label{eq:entropy}
+    H[p_1, \dots, p_N] = - \sum_{i = 1}^N p_i\log{p_i},
+\end{equation}
+The equation can be interpreted as the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
+In the continuous case:
+\begin{equation}
+    H[p(x)] = - \int p(x)\ln p(x) dx,
+\end{equation}
+We call the function $H$ (information) entropy.
+
+We now turn the core of our problem: how do we make a probability assignment for a set of events, that keeps into account our knoledge of the system and, at the same time, it is non committal with unavailable knowledge?
+The knowledge at our disposal about the system under study is what we call `evidence' and of course, any probability assignment must agree with it.
+In the above cases our knowledge on the system was only the total number $N$ of different outcomes -- this is a minimal requirement. Of course, more complex evidence constraints can be applied.
+
+It is very common that the constraints provided by the evidence are not enough for setting the probabilities for each event: in this case, it is reasonable to assume that the probability assignment should make the experiment as unpredictable as possible.
+In other words, the amount of `information' introduced by the the probability assignment should be as high as possible.
+
+%Let us now reverse the perspective: what is the most unbiased probability assignment we can make if all we know is that one over $N$ outcomes will occur?
 Since we have no reason to prefere one outcome over the others, the least biased assignment is the one equivalent to (\ref{eq:indifference}) and is found to be coincident with the most uncertain experiment.
@@
 \\
-In the previous example we had no reason to prefere one outcome over the others but that's not always the case.
+%In the previous example we had no reason to prefere one outcome over the others but that's not always the case.
 This hypothesis can be relaxed introducing the concept of 'evidence', that represents the knowledge we posses on the system under study.
@@
 Probability assignment must obey to all the evidence we possess.
-\par
-MAXENT puts the aforementioned principle into a more precise mathematical context by stating that probabilities should be assigned by maximizing uncertainty using evidence as constraint. 
+
+MAXENT puts the aforementioned principle into a more precise mathematical context by stating that probabilities should be assigned by maximizing uncertainty (information entropy) using evidence as constraint.
 This defines a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized.
-$H$ has to be a `generalization' of information defined in equation \ref{eq:information} to be something like the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
-
-Shannon \cite{Shannon} showed that the only functional form satisfying continuity with respect to its parameters and additivity is:
-\begin{equation}\label{eq:entropy}
-    H[p_1, \dots, p_N] = - \sum_{i = 1}^N p_i\log{p_i},
-\end{equation}
-or, in the continuous case:
-\begin{equation}
-    H[p(x)] = - \int p(x)\ln p(x) dx,
-\end{equation}
-and $H$ is what we will call (information) entropy.
 
 The maximisation of entropy, supplemented by evidence in the form of constraints to which the sought-for probability distribution must obey,
-gives rise to several of the most common probability distributions utilised commonly by data scientists.
-For instance, whenever the only constraint available is the normalisation of the probability distribution, the entropy is maximised by the uniform distribution.
-If the information available is instead the expected value -- the mean -- entropy is maximised by the  exponential distribution
+gives rise to several of the most common probability distributions commonly employed in statistics.
+For instance, whenever the only constraint available is the normalisation of the probability distribution (i.e.
+no evidence is available), the entropy is maximised by the uniform distribution.
+If we have evidence to constraint the expected value, the information entropy is maximised by the exponential distribution
 
@@
 p&\left((\tilde{x}_1, \ldots, \tilde{x}_n)\vert I\right) = \nonumber \\
-    &\frac{1}{\left(2 \pi \det S\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i}\tilde{x}_i^2S^{-1}_{i} \right)\,,
+    &\frac{1}{\left(2 \pi \det S\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{ij}\tilde{x}_i S^{-1}_{ij} {x}_j \right)\,,
 \end{align}
-where the matrix $S = S_i \delta_{ij}$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
+where the matrix $S_{ij} = S_i \delta_{ij}$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
 Many readers will recognise the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\cite{prob_information_theory}
@@
 An accurate profile reconstruction can be obtained with MESA using a 5 seconds-strain only, while Welch method requires at least 10 seconds of data to obtain a comparable profile.
-Another important element is noise: we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
+This is true in the low frequency sector, where the estimate requires longer time windows.
+Another important element is the estimation noise: we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
 has a large number of spurious peaks) rather than PSD provided by MESA with FPE loss function.
 If longer datasets are considered, even if MESA is found to be a little more accurate and less noisy than Welch, all of the above differences become negligible.
-As we already said Welch method is very instable with respect to the windowing choices we made.
-We tried to find the best compromize between noise and accuracy for the reconstruction, but different choices can be made.
-We want to stress that this fact doesn't invalidate our discussion: a more accurate reconstruction can only be obtained introducing a greater amount of noise: a greater accuracy requires longer sub-segments of data, reducing the number of spectra over which we average.
+
+As already discussed Welch, method is very instable with respect to the windowing choices we made.
+We tried to find the best compromize between noise and accuracy for the reconstruction, but different choices can be made, possibly providing more accurate results.
+We want to stress that this fact doesn't invalidate our discussion: one of the most appealing advantages of MESA is the absence of fine tuning.
+A more accurate reconstruction with Welch can only be obtained introducing a greater amount of noise: a greater accuracy requires longer sub-segments of data, reducing the number of spectra over which we average.
 Reducing noise require the opposite choices, and we can conclude that MESA represents a better compromise between noise and accuracy than Welch's method.
@@
 	\label{fig:mem_welch_fakedata}
-	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_True.pdf}
+	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_fake_True.pdf}
 \end{figure}
@@
 	\label{fig:mem_welch_realdata}
-	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_False.pdf}
+	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_fake_False.pdf}
 \end{figure}

### 2021-03-26 paper/mesa.tex
Various changes

@@
 Burg method of Maximum Entropy Spectral Analysis (MESA) provides a powerful tool to perform spectral estimation of a time-series.
-The method is a generalization of Jaynes maximum entropy principle and provides the estimate for the spectrum in terms of the coefficients of some autoregressive process AR(p) or order $p$.
+The method is a generalization of Jaynes maximum entropy principle and provides the estimate for the spectrum in terms of the coefficients of some autoregressive process AR(p) of order $p$.
 A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
@@
 		\email{s.schmidt@uu.nl}
-        \affiliation{add me}      
+        \affiliation{Institute for Gravitational and Subatomic Physics (GRASP), Utrecht University, Princetonplein 1, 3584 CC, Utrecht, The Netherlands}      
 	\author{Walter \surname{Del Pozzo}}
@@
 
-The study of the properties of  stochastic processes is a crucial task in many fields of physics and mathematical physics [Here we should make some examples...].
+The study of the properties of  stochastic processes is a crucial task in many fields of physics and mathematical physics \sschmidt{Here we should make some examples... Or at least cite something}.
 A particularly interesting class of stochastic processes are the so-called \textit{wide-sense} stationary processes.
@@
 Basically, the PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many windowed batches and averaging the squared moduli of their Fourier transforms.
-This approach is equivalent \cite{Lomb,Scargle} to taking the Fourier Transform of the sample autocorrelation $\rho$, written as
-\begin{equation}
-    \rho_{Windowed} = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\},
-\end{equation}
-where $M$ is the maximum time lag at which the autocorrelation is computed.
+This approach is equivalent \cite{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
+\begin{equation}
+    \rho_{W} = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\},
+\end{equation}
+where $\rho$ is the empirical autocorrelation and $M$ is the maximum time lag at which the autocorrelation is computed.
 The sequence $W$ is a window function that can be chosen in several different way, each choice presenting advantages and disadvantages for the final estimate of the PSD.
@@
 In the continuous case:
-\begin{equation}
+\begin{equation} \label{eq:entropy_continuos}
     H[p(x)] = - \int p(x)\ln p(x) dx,
@@
 
-%
-%
-
-%This interpretation is straightforward: if we compare it with Eq.~(\eqref{eq:information}), we see that the entropy is defined as the 'average information'.
-The probability assignments made maximizing it with respect to $p$, is somehow requiring  the 'expected value' of uncertainty to be a maximum.
-Such an assignment is in agreement with our hypothesis of being `non committal' with unavailable evidence, since \eqref{eq:entropy} is all we need to assign probabilities uniquely.
-Evidence can be added in form of constraints in the maximization.
-
-%It is easy to verify that this expression for entropy verifies the aforementioned properties. 
-%Assigning the same value for each probability $p_i$, it is straightforward to verify $H$ increases with N: 
-%\begin{equation}
-%    -\sum_{i = 1}^N \frac{1}{N}\log{\frac{1}{N}} = - \log\frac{1}{N} = \log N.
-%\end{equation}
-%\wdp{I need to review and correct the following}
-%To show how the maximization of entropy, supplemented with additional evidence, can provide a unique probability assignment, 
-%we use a simple, pedagogical example. Consider the case in which we want assign a probability distribution $p(x)$ to a variable $x$ 
-%for which we know mean $\mu$ and covariance $\Sigma$. Our `evidence' is then coded as the following constraints:
-%\begin{align}
-%\sum_{i = 1}^{N}p_i &= 1 \\
-%\sum_{i = 1}^{N}x_i p_i &= \mu \\
-%\sum_{i = 1}^{N}(x_i - \mu)(x_j - \mu) p_i &= \Sigma\,
-%\end{align}
-%where the first row states the requirement of exhaustivity.
-%\wdp{i think we want the multivariate case since i) the gaussian case is also on wikipedia ii) we actually will need that  later on.}
-
-%First of all, we want to show that, in case in which our only constraint is exhaustiveness
-%$\sum_{i = 1}^{N}p_i = 1 $,
-%MAXENT reduces to indifference principle as it should be. The variational problem is defined as
-%\begin{equation}\label{eq:entropyNormConst}
-%    \delta\left[H - \lambda\sum_{i = 1}^N (p_i - 1)\right] = 0 
-%\end{equation}
-%with
-%\begin{equation}
-%    \delta H\left[p_1, \dots, p_N\right] = \sum_{i = 1}^N\frac{\partial H}{\partial p_i}\delta p_i = - \sum_{i = 1}^N\left(\log p_i + 1\right),
-%\end{equation}
-%so that equation \eqref{eq:entropyNormConst} reduce to
-%\begin{equation}\nonumber 
-%    \sum_{i = 1}^N\left[\log p_i + \lambda + 1 \right]\delta p_i = 0\,.
-%\end{equation}
-%The solution of the previous equation is easily worked out requiring every term to be zero: 
-%\begin{equation}\nonumber 
-%    p_i = e^{-\lambda - 1}.
-%\end{equation}
-%Applying the normalization condition, it is immediate to show that
-%\begin{equation}
-%    p_i = \frac{1}{N}.
-%\end{equation}
-%MAXENT reduces to the IP when only the normalization constraint is used.
-%
-%In the second example, we will add the constraint that the distribution has a known mean value: $\int \d{x} x p(x) = \bar{x}$. For simplicity we will work in the continuous case.
-%The variational problem can be rewritten as: 
-%\begin{equation}
-%    \int \left[\log p(x) + 1 + \lambda + \mu x \right]\delta p(x) = 0
-%\end{equation}
-%with constraints: 
-%\begin{equation*}
-%    \int \d{x} \; p(x)  = 1 \text{  and  }
-%    \int \d{x} \; x p(x) = \bar x.
-%\end{equation*}
-%For simplicity, we consider $x$ to be a variable ranging in the interval $(-\infty,+\infty)$, but this can be easily computed for every other interval.
-%
-%The general solution for $p(x)$ is:
-%\begin{equation}
-%    p(x) = e^{-1 - \lambda - \mu x} 
-%\end{equation}
-%where $\lambda$ and $\mu$ are two Lagrange multipliers
-%The normalization conditions implies 
-%\begin{equation}\nonumber 
-%   1 =  \int \d{x} \; p(x) = e^{-1-\lambda}\int e^{-\mu x} dx, 
-%\end{equation}
-%so that: 
-%\begin{equation}\nonumber 
-%    e^{-1 -\lambda} = \mu.
-%\end{equation}
-%The fixed average condition provides a value for $\mu$ in terms of $\bar{x}$:
-%\begin{equation}
-%	\mu = \frac{1}{\bar{x}}
-%\end{equation}
-%The solution is then an exponential distribution:
-%\begin{equation}
-%    p(x) = \frac{1}{\bar{x}} \exp\left(-\frac{x}{\bar{x}}\right). \nonumber 
-%\end{equation}
-
-%Lastly, and more interestingly, we also consider a constraint on the variance. For this reason, we set:
-%\begin{align*}
-%    \int \d{x} \; p(x)  = 1 \text{  and  }
-%	\int \d{x} \; x p(x) = \bar{x} \\
-%	\text{  and  }
-%	\int \d{x} \; (x-\bar{x})^2 p(x)  = \sigma^2
-%\end{align*}
-%Performing the same computation as above, we obtain a well known result:
-%\begin{equation}
-%	p(x) = \frac{1}{\sigma\sqrt{2\pi}} e^{-\frac{1}{2}\left(\frac{x-\bar{x}}{\sigma}\right)^2}
-%\end{equation}
-%
-%Take a vector  $\vec x(t)$ with components in $\mathbb{R}^N$
-%\begin{equation*}
-%    \vec x(t) = \left(x_1, \dots, x_N \right), \quad \text{ with } x_i = x(t_i),
-%\end{equation*}
-%We assume $\vec x$ to be time dependent and real valued because we will work with time dependent single channel signals.
-If we put constraints in both mean $\vec \mu(t)$ and covariance matrix $C_(t_i, t_j)$, MAXENT allow us to assign for $\vec x(t)$ a multivariate normal distribution \cite{gregory_2005}:
-%\begin{align}
-%    p&\left(\vec x(t)\vert I\right) = \nonumber \\
-%    &\frac{1}{\left(2 \pi \det C\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{i,j}(x_i-\mu_i) (x_j-\mu_j)C^{-1}_{ij} \right). 
-%\end{align}
-
 \subsection{Maximum Entropy Spectral Analysis} \label{sec:MESA}
@@
 Clearly, this assumption is unjustified and MAXENT is a good way to drop this assumption.}.
-\par
+
 As in the previous examples, one needs to set up a variational problem where the entropy, Eq.~\eqref{eq:entropy}, is maximized 
@@
 In our case, they are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation (wherever an estimate of this is available).
-\par
+
 Before doing so, there is a technical problem to solve: the definition of entropy (Eq.~\ref{eq:entropy}) depends on a probability distribution, 
-not on the PSD. The variational problem can be formulated in terms of the power spectral density $S(f)$ alone by
-considering our signal as the result of filtering a white noise process with a filter whose power response is exactly $S(f)$ \cite{AblesMESA} \sschmidt{Which form has this filter?}. 
-The entropy gain obtained by filtering white noise with a sampling rate $\Delta t$ and Nyquist frequency $Ny \equiv \frac{1}{2 \Delta t}$ is:
+not on the PSD.
+It can be proved  \cite{AblesMESA, Bartlett} that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of filtering a white noise process with a filter with transfer function $S(f)$\footnote{
+A filter with transfer function $T(f)$ takes in input a time series $x_t$ and outputs a times series $y_t$ such that:
+$$T(f) = \frac{\tilde{y}(f)}{\tilde{x}(f)}$$
+where $\tilde{x}(f)$ denotes the Fourier transform of $x_t$ (and similarly for $y_t$)
+}.
+The difference in entropy between the input and the output time series (i.e. the entropy gain) obtained by such filter applied on white noise is:
 \begin{equation}\label{eq:EntropyGain}
@@
 \end{equation}
+where $\Delta t$ is sampling rate and $Ny \equiv \frac{1}{2 \Delta t}$  is the Nyquist frequency.
+Thus maximising eq.~\eqref{eq:EntropyGain} is equivalent to maximising eq.~\eqref{eq:entropy_continuos}.
 
@@
 The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.
-Eq.~(\eqref{eq:ARspectrum}) and Eq.~(\eqref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
+
+Eq.~\eqref{eq:ARspectrum} and Eq.~\eqref{eq:MESApsd} are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
 This shows that the MAXENT estimation of the PSD also computes the coefficients of the autoregressive 
 process that model the observed process.
-
-%\sschmidt{I don't understand the next of this section...}
-%Having computed the prediction error filter one can easily obtain an estimate of the $AR$ coefficients for the process just changing sign to the prediction error filter $a_i$.
-%Taken a time series, MESA allows us to compute the power spectrum and to find an effective description in terms of an autoregressive process. 
-%Wold's theorem ensures that this (approximate) description for the phenomenon is legitimate when the process is found to be stationary.
-So, MESA defines a bridge between spectral estimation and the study of autoregressive processes, an unexpected and fundamental result.
-\\
-%Another very remarkable process for this estimate, is that the linear predictor obtained solving the error filter equation, is the best linear predictor, i.e.
-it minimizes error, and is equivalent to a least square fit for the coefficients.
-\cite{VDBos}  \\
-%To prove that it minimizes error, we will follow the original proof provided by Burg \cite{burg1975maximum}. Consider a general linear predictor $-c_i$: 
-%\begin{equation}
-%    x_{t} - \sum_{i = 1}^p (-c_i)x_{t - i} = \sum_{i = 0}^p c_i x_{t - i} = e'_t, 
-%\end{equation}
-%Denoting with $E\left[\cdot\right]$ the expectation value of some variable, the mean square error is given by: 
-%\begin{equation}
-%E\left[\sum_{i = 0}^p c_i x_{t-i}\sum_{j = 0}^p c^*_j x^*_{t - j}\right] = \sum_{i,j} c^*_j E\left[x^*_{t-j}x_{t-i} \right] c_j = \sum_{i, j}c^*_j R(j - i) c_i > 0,
-%\end{equation}
-%which, being $R$ positive definite and defined by the multiplication of a complex number times its complex conjugate, is always positive definite. 
-%Let $-\vec{b}$ be the prediction error filter obtained with MESA. Multiplying the left hand side of \ref{eq:predictionErrorFilter} for the complex conjugate of the forward prediction error, 
-%\begin{equation}\nonumber 
-%    \sum_{i,j}b^*_jR(j-i)b_i = P_N
-%\end{equation}
-%is obtained . \\ 
-%The (positive) quantity  
-%\begin{equation}\label{eq:linearPredictor}
-%    \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i),
-%\end{equation}
-%using Einstein notation on repeated indices and recall that $c_0 = b_0 = 1$, can be rewritten as 
-%\begin{align}
-%    & c^*_j R(j - i) c_i - c^*_j R(j - i) b_i - b^*_j R(j - i) c_i + b^*_j R(j - i)  b_i =\\ \nonumber 
-%    &c^*_j R(j - i) c_i - c^*_0 P_N - P_N c_0 + P_N =\\ \nonumber
-%    & c^*_j R(j - i) c_i - P_N.
-%\end{align}
-%Moving $P_N$ on the left hand side, it becomes 
-%\begin{equation}\label{eq:bestLinearPredictor}
-%    \sum_{i, j}c^*_j R(j-i)c_i = \sum_{j = 0}^p\sum_{i = 0}^p(c^*_j - b^*_j)R(j-i)(c_i - b_i) + P_N.
-%\end{equation}
-%Being all the quantities positive definite, the left hand side is minimized by the choice $c_i = b_i$.
-Also, known $P_N$, the previous equation gives an estimate on the additional mean square error associated at choices that differ from the prediction error filter.
 
@@
 Chosing a Likelihood of the form \eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypothesis Rao has shown that chosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
-maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of \wdp{please check this equation}:
+maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of:
 
@@
 We continue our characterisation of the various loss functions considered in this work, by investigating the 
-reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve\wdp{add ref}.
+reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve\cite{LIGO_PSDs}.
 For this analysis, we generate $500$ time series of $40960$ points each, sampled with a sampling rate of $\SI{2048}{Hz}$, hence we fix the duration
@@
 
-\sschmidt{Shall we add more?}
-
 \section{Comparison with Welch method}\label{sec:Welch_comparison}
-
+\sschmidt{Make a good plot here...}
 It is really interesting to have a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
@@
 	\label{fig:mem_welch_fakedata}
-	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_fake_True.pdf}
+	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_True.pdf}
 \end{figure}
@@
 	\label{fig:mem_welch_realdata}
-	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_fake_False.pdf}
+	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_False.pdf}
 \end{figure}
@@
 
-\sschmidt{
-Some remarks:
-\begin{itemize}
-	\item What happens if we downsample the data? FPE should be able to capture the yearly trend...
-	\item Remove yearly trend from the data?
-	\item Other plots?
-\end{itemize}
-}
-
 We now assess the accuracy of the forecasting of new observations of the time series.
@@
 \sschmidt{Data should be divided in train and test set for the purpose of forecasting!!}
-
-\sschmidt{Other comments below??}
 
@@
 
-\cite{seismic_data}
+\cite{seismic_data}\\
+Are data non stationary??\\
+Cosa sono i picchi?\\
+Che cos'è la time series che guardiamo?
 
@@
 \section{Final remarks and future prospects} \label{sec:conclusion}
-
+\sschmidt{Check this section}
 We presented a case study of the application of Maximum Entropy principle in the realm of spectral estimation: it provides a closed form estimation of the power spectral density, bridging the spectral analysis with the study of autoregressive processes.
@@
 
-First of all, as short time series are required to produce meaningful PSD estimations, it can be suitable in the context of parameter estimation.
-To compute the posterior distribution for the parameters the knowledge of prior distributions and Likelihood for noise is required.
-A full description for noise can be obtained in Fourier Domain by the choice of the Whittle Likelihood \ref{eq:Whittle}, that is completely determined by the knowledge of the PSD for noise.
+A first application is based on the observation that meaningful PSD estimations can be extracted from short time series (unlike the standard approach).
+This can make the approach particularly conveniente in the context of Bayesian parameter estimation, whenever the likelihood of the model for a time series takes the form
+\begin{equation}
+	\mathcal{L}(d_t | \theta) \propto exp\left[ -\frac{1}{2} \frac{(\tilde{d}(f)- \tilde{x}(f;\theta))^2}{S(f)} \right]
+\end{equation}
+where $\tilde{\hphantom{a}}$ denotes Fourier transform and $x_t(\theta)$ is a prediction for the time series dependent on some parameter $\theta$.
+This likelihood corresponds to looking for a signal $x_t$ into noisy data.
+Usually the PSD is estimated offline on a large batch of data, that might also contain the signal to analyse and in general it might not reflect the structure of noise in the time slice to analyse.
+MESA can offer a way out fromm this problem: at each evaluation od the likelihood, a new spectral analysis is performed by computing the PSD on $d_t-x_t$.
+In this way, the PSD would depend also on $\theta$ and would effectively model the residual of noise subtracted with the signal.
+%To compute the posterior distribution for the parameters the knowledge of prior distributions and Likelihood for noise is required.
+A full description for noise can be obtained in Fourier Domain by the choice of the Whittle Likelihood \eqref{eq:Whittle}, that is completely determined by the knowledge of the PSD for noise.
 While traditional methods might be instable with respect to the hypothesis (such as Welch) or require a Large number of prior assumptions, MESA represents a stable and parsimonious way to characterize noise and compute posterior distributions.
 
-Furthermore, MESA provides a simple, but robust and quite accurate, predictor for the time series. This fact is remarkable and can be used in time series analysis for several purposes.
+Furthermore, MESA provides a simple, but robust and quite accurate in the short term, predictor for the time series.
+This fact is remarkable and can be used in time series analysis for several purposes.
 As already discussed in sec~\ref{sec:LIGO_forecasting}, an anomaly detection pipeline can be easily built using the forecasts of MESA: the predictions can form a baseline to compare the actual observations with.
@@
 
-From eq.~\eqref{eq:lamdas} and from the properties above, is easily seen \sschmidt{How exactly?} that $S(f)$ can be expressed via a Fourier Series 
+From eq.~\eqref{eq:lamdas} and from the properties above, is easily seen from the properties of the Fourier transform that $S(f)$ can be expressed via a Fourier Series 
 \begin{equation}\label{eq:PSDconstraint}

### 2021-03-26 scripts/compare_LVC_data.py

@@ -9,11 +9,11 @@ from style_sheet import *
 import pandas as pd
 from scipy.interpolate import interp1d
 
-plot = False
+plot = True
 plot_same = True
 compute = False
 generate_fake_data = False
-use_fake_data = False
+use_fake_data = True
 
 	#folder to save the plot at
 save_folder = '../paper/Images/comparison_LVC_data/'
@@ -46,7 +46,7 @@ if compute:
 
 	for i, T in enumerate(T_list):
 		print("Batch length: {}s".format(T))
-		data_T = data[:int(srate*T)]
+		data_T = data[-int(srate*T)-4096*5:-4096*5]
 		
 		M = MESA()
 		M.solve(data_T, early_stop = True, method = 'Standard')

### 2021-03-26 scripts/plot_seismic_data.py

@@ -34,9 +34,9 @@ M_FPE = memspectrum.MESA()
 T_train = 10000
 
 if False:
-	M_CAT.solve(data[:int(T_train*srate)], optimisation_method = 'CAT', method = 'Standard' , verbose = True, m = 10000, early_stop = True)	
+	M_CAT.solve(data[-int(T_train*srate):], optimisation_method = 'CAT', method = 'Standard' , verbose = True, m = 10000, early_stop = True)	
 	M_CAT.save('seismic_model_CAT')
-	M_FPE.solve(data[:int(T_train*srate)], optimisation_method = 'FPE', method = 'Standard' , verbose = True, m = 10000, early_stop = True)	
+	M_FPE.solve(data[-int(T_train*srate):], optimisation_method = 'FPE', method = 'Standard' , verbose = True, m = 10000, early_stop = True)	
 	M_FPE.save('seismic_model_FPE')
 else:
 	M_CAT.load("seismic_model_CAT")

### 2021-03-31 memspectrum.py
Change in memspectrum and others

@@ -305,7 +305,7 @@ class MESA(object):
 
 
 
-    def spectrum(self, dt, frequencies = None): 
+    def spectrum(self, dt, frequencies = None, onesided = False): 
         """
         Computes the power spectral density of the model. Default returns power 
         spectral density and frequency array automatically computed by sampling theory. 
@@ -319,14 +319,18 @@ class MESA(object):
             
         frequencies: 'np.ndarray'        
             (positive) frequencies to evaluate the spectrum at (shape (N,))
-            must be equally spaced
+            If None, a equally spaced frequency grid is used (and returned)
+           
+        onesided: 'bool'
+            Whether the one sided PSD (only positive frequencies) shall be returned.
+            It has effect only if a frequency array is not given
 
         Returns: 
         ----------
         if no frequency array is given 
             spectrum: 'np.ndarray'           
                 Two sided power spectral density (shape = (N,))
-            frequencies: 'np.ndaarray'      
+            frequencies: 'np.ndarray'      
                 Frequencies at which power spectral density is evaluated (shape = (N,))
             
         if frequency array is given:
@@ -345,7 +349,10 @@ class MESA(object):
         spec, f_spec = self._spectrum(dt, self.N)
         
         if frequencies is None:
-            return spec, f_spec
+            if onesided:
+                return spec[:self.N//2], f_spec[:self.N//2]
+            else:
+                return spec, f_spec
         
         elif isinstance(frequencies, np.ndarray):
             if np.max(frequencies) > f_ny *1.01: 

### 2021-03-31 scripts/plot_climate_data.py

@@ -18,13 +18,8 @@ M_CAT.load("../climate_data/model_CAT_long")
 M_FPE.load("../climate_data/model_FPE_long")
 
 N = len(data)
-spec_FPE, f_FPE = M_FPE.spectrum(1./srate)
-spec_FPE = spec_FPE[:int(N/2)]
-f_FPE = f_FPE[:int(N/2)]
-
-spec_CAT, f_CAT = M_CAT.spectrum(1./srate)
-spec_CAT = spec_CAT[:int(N/2)]
-f_CAT = f_CAT[:int(N/2)]
+spec_FPE, f_FPE = M_FPE.spectrum(1./srate, onesided = True)
+spec_CAT, f_CAT = M_CAT.spectrum(1./srate, onesided = True)
 
 print(len(M_CAT.a_k),len(M_FPE.a_k))
 
@@ -69,7 +64,7 @@ ax.axvline(1./(3600.*24.)*unit_shift,lw = .5, c = 'b', ls = '--', zorder = 0)
 ax.set_xlabel("frequency (1/day)")
 ax.set_ylabel("PSD (1/Hz)")
 
-ax.legend(loc = 'upper left')questo
+ax.legend(loc = 'upper left')
 
 
 	#inset

### 2021-03-31 scripts/plot_seismic_data.py

@@ -10,6 +10,8 @@ from obspy import read
 
 #http://rdsa.knmi.nl/fdsnws/dataselect/1/builder
 
+#[y] = v
+
 plot_dir = '../paper/Images/seismic_plots/'
 
 st = read('knmi-fdsnws.mseed')
@@ -64,6 +66,10 @@ else:
 l,m,h = np.percentile(forecast[:,:N_forecast],[5,50,95],axis = 0)
 sigma = np.std(forecast[:,:N_forecast], axis = 0)
 true = data[int(T_train*srate):int(T_train*srate)+N_forecast]
+
+	#plot data
+fig = init_plotting()
+plt.plot(np.linspace(0, len(data)/srate/60, len(data)),data*1e4)
 
 	
 	#plot forecasting

### 2021-03-31 paper/mesa.tex
Removed seismic stuff
Cited rationale of MAXENT

@@
 
-An appealing alternative approach, based on the Maximum Entropy principle \cite{JaynesArticle,jaynes2003ptl}, has been put forward by Burg \cite{burg1975maximum}.
+An appealing alternative approach, based on the Maximum Entropy principle \cite{JaynesArticle,jaynes2003ptl, Jaynes_MAXENT}, has been put forward by Burg \cite{burg1975maximum}.
 Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the arbitrary parameters, as it provides an iterative closed form expression for the spectrum of a stochastic stationary time series.
@@
 
-\subsection{Seismic waveforms} \label{sec:seismic_timeseries}
-\blindtext
-
-\cite{seismic_data}\\
-Are data non stationary??\\
-Cosa sono i picchi?\\
-Che cos'è la time series che guardiamo?
-
-\blindtext
-
-\begin{figure}
-	\caption{Do a caption}
-	\label{fig:seismic_spectrum}
-	\includegraphics{Images/seismic_plots/seismic_spectrum.pdf}
-\end{figure}
-
-\begin{figure}
-	\caption{Do a caption}
-	\label{fig:seismic_forecast}
-	\includegraphics{Images/seismic_plots/forecast_accuracy.pdf}
-\end{figure}
-
 \subsection{Forecasting the LIGO strain} \label{sec:LIGO_forecasting}

### 2021-04-01 memspectrum.py
Normalized one-sided PSD

@@ -350,12 +350,12 @@ class MESA(object):
         
         if frequencies is None:
             if onesided:
-                return spec[:self.N//2], f_spec[:self.N//2]
+                return spec[:self.N//2] * np.sqrt(2), f_spec[:self.N//2]
             else:
                 return spec, f_spec
         
         elif isinstance(frequencies, np.ndarray):
-            if np.max(frequencies) > f_ny *1.01: 
+            if np.max(frequencies) > f_ny * 1.01: 
                 warnings.warn("Some of the required frequencies are higher than the Nyquist frequency ({} Hz): a zero PSD is returned for f>Nyquist".format(f_ny), UserWarning)
                 
             f_interp = np.interp(frequencies, f_spec[:int(self.N/2+0.5)], spec.real[:int(self.N/2+0.5)], left = 0., right = 0.)

### 2021-04-06 paper/mesa.tex
a pass at the abstract

@@
 \begin{abstract}
-Burg method of Maximum Entropy Spectral Analysis (MESA) provides a powerful tool to perform spectral estimation of a time-series.
-The method is a generalization of Jaynes maximum entropy principle and provides the estimate for the spectrum in terms of the coefficients of some autoregressive process AR(p) of order $p$.
+The Maximum Entropy Spectral Analysis (MESA) method, developed by Burg, provides a powerful tool to perform spectral estimation of a time-series.
+The method relies on a generalization of Jaynes' maximum entropy principle and provides the means of inferring the spectrum of a stochastic process in terms of the coefficients of some autoregressive process AR(p) of order $p$.
 A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
-We provide a ready to use implementation of the algorithm in the form of a python package `memspectrum`.
-We characterize our code performing a power spectral density analysis on some syntethic generated data (with known power spectral density) and we compare different criteria for stopping the recursion.
-Furthermore, we compare the performance of our code with the traditional Welch algorithm, using both syntethic data and real gravitational waves data from the LIGO-Virgo collaboration.
-We find that, when compared to the Welch's method, the Burg's method provides a power spectral density (PSD) estimation with a systematically lower variance.
-This is particularly manifest in the case of a little number of datapoints, making Burg method most suitable to work in this regime.
+We provide a ready-to-use implementation of the algorithm in the form of a python package `memspectrum`.
+We characterize our implementation by performing a power spectral density analysis on syntethic data (with known power spectral density) and we compare different criteria for stopping the recursion.
+Furthermore, we compare the performance of our code with the traditional Welch algorithm, using both syntethic data and real gravitational wave data from the LIGO-Virgo collaboration.
+We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
+systematically lower variance. This is particularly manifest in the case of a little number of data points, making Burg method most suitable to work in this regime.
 %Several methods are characterized on synthetic noise having a known power spectral density (PSD) via the computation of the percentage error and then compared.

### 2021-04-09 paper/mesa.tex
tiny fix in mesa.tex

@@
 	\begin{acknowledgments}
-%We want to thank Maurizio Maugeri and Costanza Rossi for useful discussions.
-\sschmidt{Is it fine to ackonowledge people helping with temperature data and seismic data?}
-
 This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center (\url{https://www.gw-openscience.org/}), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration.

### 2021-04-14 paper/mesa.tex
intro

--- 2021-04-06 paper/mesa.tex
@@
 The Maximum Entropy Spectral Analysis (MESA) method, developed by Burg, provides a powerful tool to perform spectral estimation of a time-series.
-The method relies on a generalization of Jaynes' maximum entropy principle and provides the means of inferring the spectrum of a stochastic process in terms of the coefficients of some autoregressive process AR(p) of order $p$.
+The method relies on a Jaynes' maximum entropy principle and provides the means of inferring the spectrum of a stochastic process in terms of the coefficients of some autoregressive process AR($p$) of order $p$.
 A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
@@
 We characterize our implementation by performing a power spectral density analysis on syntethic data (with known power spectral density) and we compare different criteria for stopping the recursion.
-Furthermore, we compare the performance of our code with the traditional Welch algorithm, using both syntethic data and real gravitational wave data from the LIGO-Virgo collaboration.
+Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using both syntethic data and real gravitational wave data from the LIGO-Virgo collaboration.
 We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
-systematically lower variance. This is particularly manifest in the case of a little number of data points, making Burg method most suitable to work in this regime.
+systematically lower variance. This is particularly manifest in the case of a little number of data points, making Burg's method most suitable to work in this regime.
 %Several methods are characterized on synthetic noise having a known power spectral density (PSD) via the computation of the percentage error and then compared.
@@
 
-The study of the properties of stochastic processes is a crucial task in many fields of physics and mathematical physics \sschmidt{Here we should make some examples... Or at least cite something}.
-A particularly interesting class of stochastic processes are the so-called \textit{wide-sense} stationary processes.
+The study of the properties of stochastic processes is a crucial task in many fields of physics, astronomy, quantitative biology, astronomy as well as engineering.
+Of particular interest for several areas of physics is the study of stochastic time series.
+Among those, a special place is occupied by the so-called \textit{wide-sense} stationary processes.
 These are stochastic processes that display an invariance of their statistical properties, such as their two-point autocovariance function, with respect to time translation.
@@
 \end{equation}
-or equivalently by the knowledge of their \emph{power spectral density} (PSD) $S(f)$. Thanks to the Wiener-Khinchin theorem the two are related by a Fourier transform: 
+or, equivalently, by the knowledge of their \emph{power spectral density} (PSD) $S(f)$. Thanks to the Wiener-Khinchin theorem the two are related by a Fourier transform: 
 \begin{equation}\label{eq:psd-autocorrelation}
@@
 \end{equation}
-In some literature, the PSD is introduced as  
+In some literature, especially in the context of gravitational waves physics, e.g. ~\cite{Finn_1992}, the PSD is introduced as  
 \begin{equation}\label{eq:psd-f-definition}
@@
 The latter definition in~\eqref{eq:psd-f-definition} gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.
-%Under such assumption, it can be shown [cite something] that the problem of characterizing the noise properties is greatly simplified and its properties are summarized by a single \textit{autocorrelation function}, which describes how an observation at a given time correlates with another observation a later time.
-More formally, the properties of a zero mean time series $x_t$ are described by the function:
-%\begin{equation}
-%	R(\tau) = \mathbf{E}[x_t \cdot x_{t+\tau}]
-%\end{equation}
-%where the average is computed on different times $t$.
-%Thus, a stationary noise time series is completely characterized by the autocorellation function.
-%\par
-%More often, the study is performed in frequency domain: indeed, for stationary noise the components in frequency $\tilde{x}(f)$ of a stationary time series are conditionally independent from each other.
-It can be shown that:
-%\begin{equation}
-%	\mathbf{E}[\tilde{x}(f) \cdot \tilde{x}(f^\prime)] = S(f) \delta(f-f^\prime)
-%\end{equation}
-%This equation defines the function \textit{power spectral density} (PSD) $S(f)$.
-%The PSD has a straightforward interpretation: it measures how much signal ``power" is located in each frequency bin.
-Furthermore, it is not difficult to show that it is the Fourier transform of the autocorrelation function:
-%\begin{equation}
-%	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau R(\tau) e^{-i 2 \pi f \tau}
-%\end{equation}
-%A power spectral density fully describes a stationary noisy time series and this shows how important is in the study of noise.
-%\par
-%Thus, in many real application, the knowledge of the PSD is crucial for the understanding of the process and indeed algorithms for PSD computation are a standard part of every software package for computational physics.
 
-An ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
-Basically, the PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many windowed batches and averaging the squared moduli of their Fourier transforms.
-This approach is equivalent \cite{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
+A ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
+Basically, the PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
+This approach is equivalent~\cite{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
 \begin{equation}
@@
 A high frequency resolution implies high variance and vice-versa.
-Besides the window function, Welch method requires a number of arbitrary choices to be made, as the number of time slices and the overlap between consecutive slices.
+Besides the window function, Welch' method requires a number of arbitrary choices to be made, since the number of time slices and the overlap between consecutive slices.
 All the knobs must be tuned by hand and their choice can dramatically affect the PSD estimation.
-The final PSD estimation depends on all such choices, hence begging the question of what the ``real" PSD is.
+The final PSD estimate depends on all such choices, hence begging the question of what the ``real" PSD is.
 
@@
 We are arbitrarily assuming $\rho_j = 0$ for $j > M$ and modifying the estimate (i.e. the data) if a non-rectangular window is chosen.
-Making assumptions on unobserved data and modifying the ones we have at our disposal introduces a piece of ``spurious" information about the process that we are not allowed to use.
+Making assumptions on unobserved data and modifying the ones we have at our disposal introduces ``spurious" information about the process that we in general do not really have.
 
 An appealing alternative approach, based on the Maximum Entropy principle \cite{JaynesArticle,jaynes2003ptl, Jaynes_MAXENT}, has been put forward by Burg \cite{burg1975maximum}.
-Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the arbitrary parameters, as it provides an iterative closed form expression for the spectrum of a stochastic stationary time series.
+Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the algorithm parameters, since it provides an iterative closed form expression for the spectrum of a stochastic stationary time series.
 Furthermore, it embeds the PSD estimation problem into an elegant theoretical framework and makes minimal assumptions on the nature of the data.
 Lastly and most importantly, it provides a robust link between spectral density estimation and the field of autoregressive processes.
-This provide a natural and simple machinery to forecast a time series, thus predicting past observations based on previous ones.
+This provide a natural and simple machinery to forecast a time series, thus predicting future observations based on previous ones.
 
-In this paper we discuss the details of the Maximum entropy principle, its application to Power Spectral Density estimation with Burg algorithm and the link between Burg's algorithm and autoregressive process.
-Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many problem posed by the ubiquitous Welch algorithm (or other similar methods).
-To facilitate this goal, we present and describe a new code, \texttt{memspectrum}, that provides a robust python implementation of the algorithm.
-The code is ready to use and available at the PyPI repository\footnote{
+In this paper, we discuss the details of the Maximum entropy principle, its application to the problem of PSD estimation with Burg algorithm and the link between Burg's algorithm and autoregressive process.
+Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many problem posed by Welch's algorithm (or other similar methods).
+To facilitate this goal, we present and describe a new code, \texttt{memspectrum}, that provides a robust and easy-to-use python implementation of the algorithm\footnote{
 Available at link: \url{https://pypi.org/project/memspectrum/}.
-}
-We provide a thorough assessment of the performance of our code and we validate our results performing a number of tests.
+}.
+We provide a thorough assessment of the performance of our code and we validate our results performing a number of tests on simulated and real data.
 We also compare our results with those of spectral analysis carried out with the standard Welch method.
@@
 
-Our paper is organized as follows: blablabla...
+Our paper is organized as follows: we being by briefly reviewing the theoretical foundations of the maximum entropy principle in Sec.~\ref{sec:foundations}.
+Sec.~\ref{sec:validation} presents the validation of Burg's method as well as of our implementation on simulated data.
+In Sec.~\ref\label{sec:Welch_comparison} we compare the results from \texttt{memspectrum} with the Welch method; Sec.~\ref{sec:applications} presents a few applications to time series and, finally, we conclude with a discussion in Sec.~\ref{sec:conclusion}.
 
-\section{Theoretical foundations}
+\section{Theoretical foundations}\label{sec:foundations}
 The Maximum Entropy principle (MAXENT) is among the most important results in Bayesian probability theory.
@@
 
-\section{Validation of the model}
+\section{Validation of the model}\label{sec:validation}
 It is now clear that MESA provides a recursive formula for computing the coefficients $a_k$ in Eq.~(\ref{eq:MESApsd}).
@@
 	\begin{acknowledgments}
-%We want to thank Maurizio Maugeri and Costanza Rossi for useful discussions.
-\sschmidt{Is it fine to ackonowledge people helping with temperature data and seismic data?}
-
+We are grateful to S.~Shore and PLEASE ADD WHOEVER WE INTERACTED WITH IN THIS PAPER for useful discussions.
 This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center (\url{https://www.gw-openscience.org/}), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration.

### 2021-04-15 paper/mesa.tex
First bunch of edits

@@
 systematically lower variance. This is particularly manifest in the case of a little number of data points, making Burg's method most suitable to work in this regime.
-%Several methods are characterized on synthetic noise having a known power spectral density (PSD) via the computation of the percentage error and then compared.
 
@@
 
-The study of the properties of stochastic processes is a crucial task in many fields of physics, astronomy, quantitative biology, astronomy as well as engineering.
+The study of the properties of stochastic processes is a crucial task in many fields of physics, astronomy, quantitative biology, astronomy as well as engineering and finance.
 Of particular interest for several areas of physics is the study of stochastic time series.
@@
 A high frequency resolution implies high variance and vice-versa.
-Besides the window function, Welch' method requires a number of arbitrary choices to be made, since the number of time slices and the overlap between consecutive slices.
-All the knobs must be tuned by hand and their choice can dramatically affect the PSD estimation.
-The final PSD estimate depends on all such choices, hence begging the question of what the ``real" PSD is.
+Besides the window function, Welch's method requires a number of arbitrary choices to be made, such as the number of time slices and the overlap between consecutive slices.
+All these knobs must be tuned by hand and their choice can dramatically affect the PSD estimation, hence begging the question of what the ``real" PSD is.
 
@@
 In this paper, we discuss the details of the Maximum entropy principle, its application to the problem of PSD estimation with Burg algorithm and the link between Burg's algorithm and autoregressive process.
-Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many problem posed by Welch's algorithm (or other similar methods).
+Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many problems posed by the Welch's algorithm (or other similar methods).
 To facilitate this goal, we present and describe a new code, \texttt{memspectrum}, that provides a robust and easy-to-use python implementation of the algorithm\footnote{
@@
 We provide a thorough assessment of the performance of our code and we validate our results performing a number of tests on simulated and real data.
-We also compare our results with those of spectral analysis carried out with the standard Welch method.
+We also compare our results with those of spectral analysis carried out with the standard Welch's method.
 In order to apply our model on a realistic setting, we analyse some time series of broad interest in the scientific community.
@@
 Sec.~\ref{sec:validation} presents the validation of Burg's method as well as of our implementation on simulated data.
-In Sec.~\ref\label{sec:Welch_comparison} we compare the results from \texttt{memspectrum} with the Welch method; Sec.~\ref{sec:applications} presents a few applications to time series and, finally, we conclude with a discussion in Sec.~\ref{sec:conclusion}.
+In Sec.~\ref{sec:Welch_comparison} we compare the results from \texttt{memspectrum} with the Welch method; Sec.~\ref{sec:applications} presents a few applications to time series and, finally, we conclude with a discussion in Sec.~\ref{sec:conclusion}.
 
@@
 
-It is very common that the constraints provided by the evidence are not enough for setting the probabilities for each event: in this case, it is reasonable to assume that the probability assignment should make the experiment as unpredictable as possible.
+It is very common that the constraints provided by the evidence are not enough for setting the probabilities for each event: in this case, it is reasonable to assume that the probability assignment should make the experiment as unpredictable as possible\footnote{
+In ref.~\cite{Jaynes_MAXENT} this statement is made more precise and justified more thoroughly, with arguments based on combinatorial analysis.
+}.
 In other words, the amount of `information' introduced by the the probability assignment should be as high as possible.
 
-%Let us now reverse the perspective: what is the most unbiased probability assignment we can make if all we know is that one over $N$ outcomes will occur?
-Since we have no reason to prefere one outcome over the others, the least biased assignment is the one equivalent to (\ref{eq:indifference}) and is found to be coincident with the most uncertain experiment.
-This result represents the core idea of MAXENT principle: the least biased probability assignment can be obtained maximizing the "expected" information's entropy.
-\\
-%In the previous example we had no reason to prefere one outcome over the others but that's not always the case.
-This hypothesis can be relaxed introducing the concept of 'evidence', that represents the knowledge we posses on the system under study.
-With proper evidence some outcomes can be preferred to others.
-Probability assignment must obey to all the evidence we possess.
-
 MAXENT puts the aforementioned principle into a more precise mathematical context by stating that probabilities should be assigned by maximizing uncertainty (information entropy) using evidence as constraint.
-This defines a variational problem, where a numerical measure of uncertainty $H\left[p_1, \dots, p_N\right]$ has to maximized.
+This defines a variational problem, where the information entropy $H\left[p_1, \dots, p_N\right]$, defined in eq.~\eqref{eq:entropy}, has to maximized.
 
@@
 We retain this particularly interesting from the foundational point of view, since it provides a deeper insight into the ubiquitous Gaussian distribution.
-It is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy.
+Indeed, the Gaussian distribution is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy.
 For this reason, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process.
@@
 For this work, we are particularly interested in the mutli-dimensional case.
-Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realisation of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of auto-covariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution\cite{gregory_2005}:
+Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realisation of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of autocovariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution\cite{gregory_2005}:
 \begin{align}
@@
 
-We have already seen that for a wide-sense stationary process the mean function is independent of time, hence redefined to be equal to zero without loss of generality, and the auto-covariance function is dependent only on the time lag $\tau \equiv t_i - t_j$.
+For a wide-sense stationary process the mean function is independent of time, hence it can be redefined to be equal to zero without loss of generality, and the auto-covariance function is dependent only on the time lag $\tau \equiv t_i - t_j$.
 One can thus choose a sampling rate $\Delta t$ so that $C_{ij} = C((i-j)\Delta t)$.
@@
 p&\left((\tilde{x}_1, \ldots, \tilde{x}_n)\vert I\right) = \nonumber \\
-    &\frac{1}{\left(2 \pi \det S\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{ij}\tilde{x}_i S^{-1}_{ij} {x}_j \right)\,,
+    &\frac{1}{\left(2 \pi \det S\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{ij}\tilde{x}_i S^{-1}_{ij} \tilde{x}_j \right)\,,
 \end{align}
@@
 Thanks to MAXENT, the problem of defining the probability distribution describing a wide-sense stationary process is thus 
-entirely reduced to the estimation of the PSD or, equivalently, the auto-covariance function.
+entirely reduced to the estimation of the PSD or, equivalently, the autocovariance function.
 
@@
 
-As in the previous examples, one needs to set up a variational problem where the entropy, Eq.~\eqref{eq:entropy}, is maximized 
+As in the previous examples, one needs to set up a variational problem where the entropy eq.~\eqref{eq:entropy} is maximized 
 subject to some problem-specific constraints. 
@@
 
-Before doing so, there is a technical problem to solve: the definition of entropy (Eq.~\ref{eq:entropy}) depends on a probability distribution, 
-not on the PSD.
-It can be proved  \cite{AblesMESA, Bartlett} that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of filtering a white noise process with a filter with transfer function $S(f)$\footnote{
+Before doing so, there is a technical problem to solve: the definition of entropy depends on a probability distribution, not on the PSD.
+It can be proved \cite{AblesMESA, Bartlett} that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of filtering a white noise process with a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
 A filter with transfer function $T(f)$ takes in input a time series $x_t$ and outputs a times series $y_t$ such that:
@@
 
-Eq.~\eqref{eq:ARspectrum} and Eq.~\eqref{eq:MESApsd} are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
+Eq.~\eqref{eq:ARspectrum} and eq.~\eqref{eq:MESApsd} are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
 This shows that the MAXENT estimation of the PSD also computes the coefficients of the autoregressive 
@@
 The link between MESA and autoregressive processes is of particular interest.
-Given the solution to the Levinson recursion to determine the $a_k$, see Appendix A, we have automatically the coefficients of the AR with equivalent PSD, hence we are able to exploit Eq.~\ref{eq:AR_p} to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
-Indeed, an AR($p$) process models the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation:
+Given the solution to the Levinson recursion to determine the $a_k$, see Appendix~\ref{sec:appendix}, we authomatically know the coefficients of the equivalent AR process, hence we are able to exploit eq.~\ref{eq:AR_p} to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
+Indeed, for an AR($p$) process the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation has the form:
 \begin{align}\label{eq:p_forecast}
@@
 In a perfect scenario, this would be equal to the number of points the autocorrelation is computed at (equivalent to the length of data considered).
-However, the computation of high order coefficients of the autocorrelation is unstable and for high enough $n$, the the estimation for  $\bar{n}_m$ shows a very high variance, broadly scaling as $\sim \left(\sqrt{M - m}\right)^{-1}$.
+However, the computation of high order coefficients of the autocorrelation is unstable and for high enough $m$, as the estimation for  $\bar{r}_m$ shows a very high variance, broadly scaling as $\sim \left(\sqrt{M - m}\right)^{-1}$.
 
@@
 on the one hand it is advisable to include as much knowledge of the autocorrelation as possible, leading to include all the known $\bar{r}_m$; on the other hand, including values of the autocorrelation that are not reliably estimated, can be counterproductive.
-Indeed, the order $M$ of the autocorrelation to be considered (or equivalently the order $M$ of the underlying autoregressive process) is the only tuning parameter of MESA and a careful balance between this two necessities must be made when applying the algorithm.
+The order $M$ of the autocorrelation to be considered (or equivalently the order $M$ of the underlying autoregressive process) is the only tuning parameter of MESA and a careful balance between this two necessities must be made when applying the algorithm.
 
@@
 \end{equation}
-where $N$ is the number of points in the time-series.
+where $N$ is the number of observed points in the time-series.
 However, this is just a plausible upper limit on the order of the AR process $m$ and the optimal algorithm could employ fewer points.
@@
 \end{equation}
-with $P_m$ defined by the prediction error filter equation \ref{eq:errorFilter1}. \\ 
-If the $N \to \infty$ limit is considered,remembering $m_{max} = 2N / \log(2N)$, we see that N dominates both the numerator and the denominator of the fraction: in large $N$ limit using Akaike's loss function is equivalent to the minimization of the variance $P_m$ of the process.
+with $P_m$ being the estimated noise variance at order $m$, see eq.~\eqref{eq:errorFilter1}.
+
+In the $N \to \infty$ limit, remembering $m_{max} \sim 2N / \log(2N)$, Akaike's loss function is equivalent to the minimization of the variance $P_m$ of the white noise of the underlying $AR(p)$ model.
 
@@
 The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
-time series for the process that has to be described as an autoregressive process of order $m$ (AR(m)) with m to be determined.
+time series for the process that has to be described as an autoregressive process of order $m$ ($AR(m)$), with $m$ to be determined.
 The OBD is obtained chosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$.
@@
 In this first experiment, we take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$.
-The time series are generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$, following the procedure outlined in sec.~\ref{sec:forecasting}, and performing an inverse Fast Fourier Transform on the frequency series.
-For this investigation, we generate a dataset of 1000 time series of 3000 points each.
+The time series are generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$ and performing an inverse Fast Fourier Transform on the frequency series.
+For this investigation, we generate a dataset of $N=1000$ time series of $3000$ points each.
 
@@
 The disagreement between the PSD $S_i(f)$ estimated from the $i$-th simulated time series and the target PSD $S(f)$ is measured via 
-the relative error $r_i$, averaged over all frequencies:
+the \textit{frequency-averaged relative error} $r_i$:
 \begin{equation}\label{eq:freq_error}
-r_i = \frac{1}{Ny}\sum_{f_j=0}^{f_j = Ny} \frac{|S_i(f_j) - S(f_j)|}{S(f_j)}
+	r_i = \frac{1}{N_f}\sum_{f_j=\{0, \hdots , Ny\}} \frac{|S_i(f_j) - S(f_j)|}{S(f_j)}
 \end{equation}
-where $Ny$ is the Nyquist frequency of the time series.
+where $Ny$ is the Nyquist frequency of the time series and $N_f$ is the number of the discrete frequencies the PSD is evaluated at.
 
-For each loss function, we compute the ensemble-averaged PSD (as well as the $90\%$ confidence level) and the ensemble-averaged frequency dependence of the error:
+For each loss function, we compute the ensemble-averaged PSD (as well as the $90\%$ confidence level) and the \textit{ensemble-averaged relative error}:
 \begin{equation}\label{eq:ensemble_averaged_error}
-	r(f) = \sum_i \frac{|S_i(f) - S(f)|}{S(f)} \;.
+	r(f) = \frac{1}{N} \sum_i \frac{|S_i(f) - S(f)|}{S(f)} \;.
 \end{equation}
-Figs.~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} show the averaged PSD for the three loss functions, 
-as well as the ensemble-averaged error from Eq.~\eqref{eq:ensemble_averaged_error}.
-Furthermore, Fig.~\ref{fig:optcomparison} displays the correlation between $r_i$ and the autoregressive 
+For each loss function, figs.~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} show the averaged PSD, 
+as well as the ensemble-averaged relative error from Eq.~\eqref{eq:ensemble_averaged_error}.
+Furthermore, Fig.~\ref{fig:optcomparison} displays the relation between $r_i$ and the autoregressive 
 order $m_i$ chosen for each independently drawn time series.
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/FPE_spectrum_estim.pdf}
-	\caption{Average spectrum for $\mathcal{L}_{\rm FPE}(m)$ with 90\% confidence regions (top) and percentage error (bottom).
+	\caption{Average spectrum for $\mathcal{L}_{\rm FPE}(m)$ with 90\% confidence regions (top) and ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error} (bottom).
 The average is computed with $1000$ realization of a $3000$ points long time series.}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/OBD_spectrum_estim.pdf}
-	\caption{Average spectrum for $\mathcal{L}_{\rm OBD}(m)$ loss function with 90\% confidence regions (top) and percentage error (bottom).
+	\caption{Average spectrum for $\mathcal{L}_{\rm OBD}(m)$ loss function with 90\% confidence regions (top) and ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error} (bottom).
 The average is computed with $1000$ realization of a $3000$ points long time series.}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/CAT_spectrum_estim.pdf}
-	\caption{Average spectrum for $\mathcal{L}_{\rm CAT}(m)$ loss function with 90\% confidence regions (top) and percentage error (bottom).
+	\caption{Average spectrum for $\mathcal{L}_{\rm CAT}(m)$ loss function with 90\% confidence regions (top) and ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error} (bottom).
 The average is computed with $1000$ realization of a $3000$ points long time series.}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/error_length_contour.pdf}
-	\caption{Relative error $r$ (as in eq.~\eqref{eq:freq_error}) vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with a gaussian PSD.
+	\caption{Frequency-averaged relative error $r$ eq.~\eqref{eq:freq_error} vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with a gaussian PSD.
 Different colours refers to different choices for the loss function: in red $\mathcal{L}_{\rm FPE}(m)$, in green $\mathcal{L}_{\rm OBD}(m)$ and in black $\mathcal{L}_{\rm CAT}(m)$.
@@
 
-By looking at therRed series in Fig.~\ref{fig:optcomparison}, we note that the autoregressive orders obtained with FPE are clustered in a very small region.
+By looking at fig.~\ref{fig:optcomparison} (red series), we note that the autoregressive orders obtained with FPE are clustered in a very small region.
 This is a desirable property: FPE, in fact, provides a stable estimation of the autoregressive order, which does not affect much the reconstruction error.
@@
 \paragraph{Optimum Bayes Decision Rule (OBD)}
-The second loss function in consideration is OBD. As for the FPE loss function, the results in Fig.~\ref{fig:OBDmean} show a 
-good agreement between the average over the reconstructions and the true spectrum.
+The second loss function in consideration is OBD and the results are reported in fig.~\ref{fig:OBDmean}.
+As for the FPE case, they show a good agreement between the average over the reconstructions and the true spectrum.
 Qualitatively the same behaviour of FPE is observed: a good quality reconstruction at the intermediate and high frequencies with a low $90\%$ confidence level; and a worse performance at low frequencies.
@@
 \paragraph{Criterion Autoregressive Transfer function (CAT)}
-The performance of CAT is shown in Fig.~\ref{fig:CATmean}. At a first glance CAT differs substantially from the other two loss functions considered.
+The performance of CAT is shown in fig.~\ref{fig:CATmean}. At a first glance CAT differs substantially from the other two loss functions considered.
 The ensemble-average PSD matches very well the underlying ``true" PSD. This is also true even in the low frequency region, 
-where both OBD and FPE had a poor performance.
+where both OBD and FPE showed poor performance.
 However, the variance of the reconstructed spectrum is quite large (much larger than for FPE and OBD), and the relative error is quite high, $\sim 10\%$ and it is approximately constant over all the frequencies.
 
-The reason for this behaviour becomes clear from Fig.~\ref{fig:optcomparison} (black series). CAT does not converge 
+The reason for this behaviour becomes clear from fig.~\ref{fig:optcomparison} (black series). CAT does not converge 
 to any specific value for the order of the autoregressive process.
@@
 
-The poor low-frequency reconstruction from OBD and FPE might be due to the fact that very small values require longer filters. 
-In fact, when we consider the average for all the spectra we note that short filters yield larger the error in the tails.
+The poor low-frequency reconstruction from OBD and FPE might be due to the fact that the first tends to select shorter filters, whereas long filters are required to model low frequency correlation.
+This seems to be confirmed by looking at fig.~\ref{fig:optcomparison}.
 OBD, which select the shortest filters, can provide an error as large as $300 \%$ at the extrema, as compared with $85\%$ error of FPE.
-In turn, CAT is $5\%$ accurate in those regions.
+In turn, CAT is $5\%$ accurate in these regions.
 
-However, we also note, see Fig.~\ref{fig:optcomparison}, that the psD inferred from a single time-series realisation with FPE is still able 
-to provide the lowest averaged error over all frequencies, while CAT can reach errors $5$ times larger.
-Hence, while CAT is the loss function that minimizes errors in the low-frequency end of the spectrum, 
-FPE obtains the best overall accuracy.
+However, we also note that the when inferring PSDs from a single time-series realisation, FPE provide the lowest averaged error over all frequencies, while CAT can reach errors $5$ times larger (see again fig.~\ref{fig:optcomparison}).
+Hence, while CAT is the loss function that minimizes errors in the low-frequency end of the spectrum, FPE obtains the best overall accuracy.
 
-The conclusion is clear: even if in some cases CAT might be more accurate when taking the average over several estimation, FPE guarantees that the single estimation will be more accurate.
+The conclusion is clear: even if in some cases CAT is more accurate when taking the average over several estimation, FPE guarantees that the single estimation is more faithful.
 As in any common situation we cannot perform such averaging over different realization of the same time series, we must prefer FPE over CAT (let alone OBD, which even though qualitatively similar to FPE has worse performance).
-However, if we indeed can measure the PSD by averaging over different time series, using CAT as a loss function is the best choice. In this sense
-we retain CAT to provide the best, and most similar in spirit, alternative to the commonly employed WelchPsd estimation method whenever 
-ensemble-averages are needed and justified.
-
-%It is evident that tails do not affect CAT  average error and standard deviation at all, so that it's accuracy is approximately constant everywhere in the spectrum. \\ \ 
-%Since both FPE and CAT shows pros and cons, at this level, there is no reason to choose between them.\\ 
-%Analyzing the resolution of the single reconstruction, figure \ref{fig:optcomparison} clearly shows that FPE minimize overall relative error giving a stable result.\\ 
-%CAT instead shows a larger error and do not converge to any specific value.
-Almost every reproduction obtained with CAT is worst with respect to those obtained with FPE, but their mean is more precise.
-\\
-%It's instability is probably the cause for this behaviour. Since it does not show a strong convergence to some order with respect to others, very different lengths are selected.
-%Longer lengths are able to capture a lot of details but are more likely to introduce noise. In other words, long filters reduce bias but increment the variance of the result. 
-%Taking the average, for central limit theorem the error decreases as $\sim \frac{1}{\sqrt{N}}$. In this way we compensate noise and obtain a spectrum with small bias and small fluctuations.
-%Considering the average, CAT reconstruct precisely and with good accuracy every part of the spectrum, even if in "regular" parts best accuracy is reached from FPE.
-For a single reproduction, CAT is very noisy and unstable, while FPE performs the better.
-\ref{fig:optcomparison}
-%definitely shows that FPE gives the best results: it is a stable method that choose filters in the clustered area associated with minimum error, while both OBD and CAT result lies outside this area and show bigger errors.
-Two other interesting plots that stress this fact are reported in figure \ref{fig:ordersCompairson} and \ref{fig:residualsComparison}
-%where the histograms for the errors and for order's estimate are reported.
-These graphs provide stronger evidence for our previous statements.
-As already noticed, FPE is very stable and provide the smallest 'overall relative error'.
-\\
-%CAT histogram shows a very interesting behaviour.
-The histogram for the order show two different peaks,  a small one associated to 'short filters', and a second high peak at $m = M$.
-This means that CAT is most likely to not converge at all at some value for the regression order, and it is more likely selects a length that is comparable with the largest available.
-If short filters are selected, the overall percentage error will be low, respectively represented by the first peak in the distribution of the errors, while if long filters are selected, the error is maximized.
+However, if we indeed can measure the PSD by averaging over different time series, using CAT as a loss function is the best choice.
+In this sense we retain CAT to provide the best, and most similar in spirit, alternative to the commonly employed Welch estimation method whenever ensemble-averages are needed and justified.
 
@@
 reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve\cite{LIGO_PSDs}.
-For this analysis, we generate $500$ time series of $40960$ points each, sampled with a sampling rate of $\SI{2048}{Hz}$, hence we fix the duration
+For this analysis, we generate $N = 500$ time series of $40960$ points each, sampled with a sampling rate of $\SI{2048}{Hz}$, hence we fix the duration
 of the time-series to $\SI{20}{s}$. The chosen length is convenient to capture fairly accurately the low-frequency features of the LIGO PSD.
+We report our findings in figs.~\ref{fig:ligospectrum}~and~\ref{fig:LigoOrderError}.
 
@@
 \end{figure}
-Our findings are summarised in Figs.~\ref{fig:ligospectrum} and ~\ref{fig:LigoOrderError}. The former shows the original spectrum (dashed line)
-and the ensemble-averaged reconstructed PSD adopting the FPE (green line), OBD(red line) and CAT(black line).
-In all cases, the spectrum is well reconstructed, but with a fairly distinct behaviour at low frequency, where CAT -- as in the Gaussian case -- better captures and resolves the distinct
-spectral feature at $\sim 17$ Hz, see also the left panel in Fig.~\ref{fig:ligoPeaks}, as well as the high frequency line at $438$Hz, right panel in Fig.~\ref{fig:ligoPeaks}. 
+
+Figure~\ref{fig:ligospectrum} shows the original spectrum (dashed line) and the ensemble-averaged reconstructed PSD adopting the FPE (green line), OBD(red line) and CAT(black line).
+In all cases, the spectrum is well reconstructed, but with a fairly distinct behaviour at low frequency, where CAT -- as in the Gaussian case -- better captures and resolves the distinct spectral feature at $\sim \SI{17}{Hz}$.
+In fig.~\ref{fig:ligoPeaks}, we report the reconstructed spectra around the two peaks at $\sim \SI{17}{Hz}$ and $\sim \SI{438}{Hz}$ (left and right panel respectively).
+
 Fig.~\ref{fig:LigoOrderError} shows the distribution of recovered AR orders $m$ against the relative frequency-averaged error. The behaviour
-of the three loss functions is very similar to what found in the Gaussian PSD case: ODB infers the smallest orders and gives average errors around $20$\%, FPE consistently estimates orders of a few hundreds and shows the smallest errors $\sim 15$ \%while CAT does not show any preference towards any AR order and displays wildly varying errors.
+of the three loss functions is very similar to what found in the Gaussian PSD case (compare it with fig.~\ref{fig:optcomparison}): ODB infers the smallest orders and gives average errors around $20$\%, FPE consistently estimates orders of a few hundreds and shows the smallest errors $\sim 15$ \% while CAT does not show any preference towards any AR order and displays wildly varying errors.
 Yet again, when the PSD is averaged over multiple realization of the time, CAT is able to capture the spectrum very precisely.
-In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them (figs.\ref{fig:ligoPeaks}).
+In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them.
 Hence, also the study of simulated Advanced LIGO data seems to indicate that whenever and wherever ensemble-averaged PSD are necessary, CAT is the optimal choice of loss function.
@@
 Therefore, the choice of loss function, at least in between CAT and FPE, depends on the problem one is attempting to solve.
-\begin{figure}[t]
-        \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/compare_estimates_peaks.pdf}
-        \caption{Details of peaks of the spectrum and their reconstruction with every optimizer}
-        \label{fig:ligoPeaks}
-\end{figure}
 \begin{figure}
@@
     \label{fig:LigoOrderError}
+\end{figure}
+
+\begin{figure}[t]
+        \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/compare_estimates_peaks.pdf}
+        \caption{Details of peaks of the spectrum and their reconstruction with every optimizer}
+        \label{fig:ligoPeaks}
 \end{figure}
@@
 In all cases, the sampling rate is set to $\SI{4096}{Hz}$.
+For the Welch algorithm, we use the standard implementation provided by the python library \texttt{scipy} \cite{numpy,scipy}.
 
@@
 \appendix
-\section{Details of PSD computation} 
+\section{Details of PSD computation} \label{sec:appendix}
 \subsection{MESA solution}\label{sec:MESA_solution}

### 2021-04-15 paper/mesa.tex
Second bunch of edits

@@
 A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
-We provide a ready-to-use implementation of the algorithm in the form of a python package `memspectrum`.
+We provide a ready-to-use implementation of the algorithm in the form of a python package \texttt{memspectrum}.
 We characterize our implementation by performing a power spectral density analysis on syntethic data (with known power spectral density) and we compare different criteria for stopping the recursion.
@@
 We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
-systematically lower variance. This is particularly manifest in the case of a little number of data points, making Burg's method most suitable to work in this regime.
+systematically lower variance and bias. This is particularly manifest in the case of a little number of data points, making Burg's method most suitable to work in this regime.
 
@@
 This is somehow obvious: more data provide more information available for spectral estimation.
-This fact is more important at lower frequencies, where the natural time length is large.
-We can also note that MESA converges to the real spectrum faster than Welch method, providing a better estimate expecially for short time series.
-This behaviour is more evident if the low frequencies are considered.
+This fact is more important at lower frequencies, where the natural time scale is large.
+Indeed, the higher frequencies are alway better resolved than the lower ones: there is alway more information in the high frequency part of a signal.
+
+We also note that MESA converges to the real spectrum faster than Welch method, providing a better estimate even in the case of short time series.
+Although observed at every frequency, this behaviour is more evident in the low frequency region.
 An accurate profile reconstruction can be obtained with MESA using a 5 seconds-strain only, while Welch method requires at least 10 seconds of data to obtain a comparable profile.
-This is true in the low frequency sector, where the estimate requires longer time windows.
-Another important element is the estimation noise: we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
+
+Another important element is the noise noise of the spectral estimation: we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
 has a large number of spurious peaks) rather than PSD provided by MESA with FPE loss function.
@@
 
-As already discussed Welch, method is very instable with respect to the windowing choices we made.
+Finally, as already discussed Welch, method is very instable with respect to the windowing choices we made.
 We tried to find the best compromize between noise and accuracy for the reconstruction, but different choices can be made, possibly providing more accurate results.
 We want to stress that this fact doesn't invalidate our discussion: one of the most appealing advantages of MESA is the absence of fine tuning.
-A more accurate reconstruction with Welch can only be obtained introducing a greater amount of noise: a greater accuracy requires longer sub-segments of data, reducing the number of spectra over which we average.
+%A more accurate reconstruction with Welch can only be obtained introducing a greater amount of noise: a greater accuracy requires longer sub-segments of data, reducing the number of spectra over which we average.
 Reducing noise require the opposite choices, and we can conclude that MESA represents a better compromise between noise and accuracy than Welch's method.
@@
 We apply our code to the study of the temperature time series.
-This will allow to test the code performance on a different time series, of great importance in physics.
+This will allow to test our model performance on an example of great importance in physics.
 We will also assess the quality of the forecasting predictions: this task has a great impact also on practical applications.
-We used the historical reanalysis data from the ``ERA5-Land hourly data from 1981 to present" dataset, downloaded from the Climate Data Store \cite{ERA5_dataset}
-
+We used the historical reanalysis data from the ``ERA5-Land hourly data from 1981 to present" dataset, downloaded from the Climate Data Store \cite{ERA5_dataset}.
 Data refers to the temperature measure at coordinates $\textrm{N}\;\ang{45;5;} \;\; \textrm{E}\;\ang{9;1;}}$, corresponding about to the city of Milan, Italy.
@@
 In this example, we have $p_{CAT} \sim  36500 \sim  \SI{4}{yr}$, whereas $p_{FPE} \sim  1300 \sim  \SI{2}{months}$.
-Clearly, the autoregressive process selected by FPE cannot produce a meaningful prediction on the timescale of a few months and for this reason it is unable to capture the peak at $f_{yr}$.
-In other words, the model is too simple to capture \textit{both} the behaviour at high and low frequencies (separated by approximately 3 orders of magnitude).
+Clearly, the autoregressive process selected by FPE cannot produce a meaningful prediction on the timescale longer than a few weeks and for this reason it is unable to capture the peak at $f_{yr}$.
+In other words, the model is too simple to model \textit{both} the behaviour at high and low frequencies (separated by approximately 3 orders of magnitude).
 On the other hand, the model chosen by CAT is much more complex and is able to model also the high frequency behaviour.
@@
 We now assess the accuracy of the forecasting of new observations of the time series.
-Based on actual data, we try to predict future values as described in sec.~\ref{sec:forecasting}.
+Based on actual data, we try to predict future values as described in sec.~\ref{sec:forecasting} (of course the MESA is performed with CAT loss function).
 We produce $N =100$ indepedent predictions and we compute the median as well as the $90\%$ confidence interval.
@@
 For more precise predictions, probably one should consider nonlinear regression model, tapping into the wealth of non linear predictors offered by the field of Deep Lerning.
-
-\sschmidt{Data should be divided in train and test set for the purpose of forecasting!!}
 
@@
 \subsection{Forecasting the LIGO strain} \label{sec:LIGO_forecasting}
-As a last example of and application of the maximum entropy method on real data, we come back to the strain produced by the LIGO-Virgo interferometers and we forecast the future observations.
+As a last example of an application of the maximum entropy method on real data, we come back to the strain produced by the LIGO-Virgo interferometers and we forecast the future observations.
 The spectral analysis is performed with CAT and FPE loss functions on $\SI{1000}{s}$ of data \cite{Abbott_2021} sampled at $\SI{4096}{Hz}$ observed at Livingston observatory starting from GPS time $1164603392$.
@@
 This means that the series is very difficult to predict: the knowledge of past observations is not very helpful in predicting future observations.
-Similar performances in the sigma are observed also with the FPE model: in this case FPE selects a large prediction filter $p_{FPE} = 29924$ , similar to those selected by CAT.
+Similar performances in $\sigma$ are observed also with the FPE model: in this case FPE selects a large prediction filter $p_{FPE} = 29924$ , similar to those selected by CAT.
 
@@
 The method has the towfold advantage of avoiding any unjustified assumption on the observed time series and of providing a simple context for forecasting the time series.
-We provided a ready to use python code, called `memspectrum' and we performed numerical studies of its performance.
+We provided a ready to use python code, called \texttt{memspectrum} and we performed numerical studies of its performance.
 We investigated the use of different loss functions for the purpose of correctly estimating the autoregressive order for the estimation and we found that different loss functions are required for different situations.
@@
 \end{equation}
-where $\tilde{\hphantom{a}}$ denotes Fourier transform and $x_t(\theta)$ is a prediction for the time series dependent on some parameter $\theta$.
-This likelihood corresponds to looking for a signal $x_t$ into noisy data.
+where $\tilde{\hphantom{a}}$ denotes Fourier transform and $x_t(\theta)$ is a prediction for a signal hidden in the observed time series $d(t)$, dependent on some parameter $\theta$.
+This likelihood corresponds to looking for a signal $x_t$ into some (possibly very) noisy data.
 Usually the PSD is estimated offline on a large batch of data, that might also contain the signal to analyse and in general it might not reflect the structure of noise in the time slice to analyse.
-MESA can offer a way out fromm this problem: at each evaluation od the likelihood, a new spectral analysis is performed by computing the PSD on $d_t-x_t$.
+MESA can offer a way out from this problem: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on $d_t-x_t$.
 In this way, the PSD would depend also on $\theta$ and would effectively model the residual of noise subtracted with the signal.
@@
 	\begin{acknowledgments}
-We are grateful to S.~Shore and PLEASE ADD WHOEVER WE INTERACTED WITH IN THIS PAPER for useful discussions.
+We are grateful to S.~Shore, M.~Maugeri and C.~Rossi for useful discussions.\\
 This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center (\url{https://www.gw-openscience.org/}), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration.
@@
 \end{equation}
-and the entropy gain expression eq.~\eqref{eq:EntropyGain} becomes:
+and, plugging it in the entropy gain expression eq.~\eqref{eq:EntropyGain}, we obtain:
 \begin{equation}
@@
 and we will show that $S(f)^{-1}$ can be written as a Fourier Expansion in terms of such coefficients.
-Thenm the determination of the values for the $\lambda_s$ uniquely solves the problem of power-spectral density estimation.
+Then, the determination of the values for the $\lambda_s$ uniquely solves the problem of power-spectral density estimation.
 
-Some properties for the coefficients can be worked out easily. First, since $S(f)$ is positive definite the $\lambda_s$ show the property 
+Some properties for the coefficients can be worked out easily. First, since $S(f)$ is real, the $\lambda_s$ show the property 
 \begin{equation}
@@
 
-To compute the $a_s$ is convenient to plug into eq.~\eqref{eq:MaxConstraint} the Laurent Polynomial exansion for $S(f)$ eq.~\eqref{eq:MESApsd_appendix} and then integrating over $z$.
+To compute the $a_s$ is convenient to plug into eq.~\eqref{eq:MaxConstraint} the Laurent Polynomial exansion for $S(f)$ eq.~\eqref{eq:MESApsd_appendix} and then integrating over $z$ (taking values on ${\mathbb S^1}$).
 In this way the equation becomes:
@@
 \end{equation}
-where $b$ is the forward prediction error coefficients at order $N-1$. 
+where $b$ holds the value of the $a_s$ coeffiecients at order $N-1$. 
 The 0-th order element can be easily initialized reminding that $a_0 = 1$ (always) and that $P_0$ can be determined from \eqref{eq:errorFilter1}.
@@
 
+\newpage
+
 	\bibliography{Bibliography.bib}

### 2021-04-15 examples/sound_MESA.py
Changed name to optimizer class

@@ -74,22 +74,7 @@ plt.yscale('log')
 plt.xlabel("frequency (Hz)")
 plt.legend()
 
-#comparison between two different values of N
-PSD_N, f_PSD_N = M._spectrum(dt, len(data_MESA))
-PSD_ak, f_PSD_ak = M._spectrum(dt, len(M.a_k))
-
-
-plt.figure()
-plt.title("Comparison between two different choices of N")
-plt.loglog(f_PSD_N[:int(len(data_MESA)/2)+1],PSD_N.real[:int(len(data_MESA)/2)+1], c='r')
-plt.loglog(f_PSD_ak[:int(len(M.a_k)/2)+1], PSD_ak.real[:int(len(M.a_k)/2)+1], c = 'b')
-plt.axvline(1./(dt*len(data_MESA)))
-plt.axvline(1./(dt*len(M.a_k)))
-plt.axvline(1./(dt*2.))
-
 plt.show()
-
-
 
 
 

### 2021-04-15 examples/sunspots.py

@@ -13,7 +13,7 @@ if __name__ == "__main__":
     datafile = "data/zuerich-monthly-sunspot-numbers-.tsv"
     _, data = np.genfromtxt(datafile, unpack = True)
     T = len(data)
-    print([d for d in data])
+    #print([d for d in data])
     dt = 1.
     srate = 1./dt
     N = data.shape[0]
@@ -21,7 +21,7 @@ if __name__ == "__main__":
     t = np.arange(0,T,step=dt)
     M = MESA()
     P, ak, _ = M.solve(data, method = "Fast", optimisation_method = "FPE", m = int(2*N/(2*np.log(N))))
-    print("p = {0}".format(len(ak)))
+    print("Length of the selected filter: p = {0}".format(len(ak)))
     PSD, _     = M.spectrum(dt)[:N//2]
 
     fig = plt.figure(1)

### 2021-04-15 memspectrum.py

@@ -40,7 +40,7 @@ except:
 
 #################
 
-class optimizer:
+class loss_function:
 
     def __init__(self, method):
         """
@@ -88,7 +88,7 @@ class optimizer:
         Returns
         -------
         'np.float'
-            The value of FPE optimizer.
+            The value of FPE loss function.
 
         """
         return P[-1] * (N + m + 1) / (N - m - 1)
@@ -110,7 +110,7 @@ class optimizer:
         Returns
         -------
         'np.float'
-            The value of AIC optimizer.
+            The value of AIC loss function.
 
         """
         return np.log(P[-1]) + (2 * m) / N
@@ -132,7 +132,7 @@ class optimizer:
         Returns
         -------
         'np.float'
-            The value of CAT optimizer.
+            The value of CAT loss function.
 
         """
         if m == 0:
@@ -162,7 +162,7 @@ class optimizer:
         Returns
         -------
         'np.float'
-            The value of OBD optimizer.
+            The value of OBD loss function.
 
         """
         P_m = P[-1]
@@ -451,7 +451,7 @@ class MESA(object):
             Default is True. Breaks the iteration if there is no new global 
             maximum after 100 iterations. 
             Recommended for every optimisation method but CAT.
-            Has no effect with "Fixed" optimizer.
+            Has no effect with "Fixed" loss function.
         
         verbose: 'boolean'
             Whether to print the status of the mesa solution
@@ -499,7 +499,7 @@ class MESA(object):
             print("Method {0} unknown! Valid choices are 'Fast' and 'Standard'".format(method))
             exit(-1)
         
-        self._optimizer = optimizer(optimisation_method)
+        self._loss_function = loss_function(optimisation_method)
         self.P, self.a_k, self.optimization = self._method()
         del self.data
         return self.P, self.a_k, self.optimization
@@ -547,8 +547,8 @@ class MESA(object):
             #Append values to respective lists
             a.append(new_a)
             P.append(P[-1] * (1 - k * k.conj()))
-            #Compute optimizer value for chosen method
-            optimization.append( self._optimizer(P, a[-1], self.N, i + 1))
+            #Compute loss function value for chosen method
+            optimization.append( self._loss_function(P, a[-1], self.N, i + 1))
             
             is_nan = np.isnan(new_a).any() #checking for nans
             if np.abs(k)>1 or is_nan:
@@ -673,7 +673,7 @@ class MESA(object):
[…]

### 2021-04-15 paper/mesa.tex
Welch-MESA comparison changed

@@
 \section{Comparison with Welch method}\label{sec:Welch_comparison}
-\sschmidt{Make a good plot here...}
 It is really interesting to have a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
@@
 
-We perform two sets of experiments.
-In the first one (see fig.~\ref{fig:mem_welch_fakedata}), we simulate data from the realesed LIGO intererferometer PSD and we make the comparison with them: this is to ensure the we have a baseline PSD to compare the data with.
-In a second experiment (see fig.~\ref{fig:mem_welch_realdata}), we employ the data released to the public by the LIGO/Virgo collaboration \cite{Abbott_2021}.
-In each experiment, we vary the length of the data used for the estimation: this is also useful to assess how the computation depends on the data available.
+In order to ensure the we have a baseline PSD to compare the data with, we simulate data from the realesed LIGO intererferometer PSD and we employ both Welch's method and MESA to estimate the PSD.
+We vary the length of the data used for the estimation: this is also useful to assess how the computation depends on the data available.
 We set the total observation time $T = 1, 5, 10, 100, 1000 \SI{}{s}$
 For the MESA algorithm, we employ FPE optimizer.
-For the Welch algorithm, we employ a Tukey window, an overlap fraction of $0.5$ for the segments and a length of segments $L = 512, 1024, 2048, 8192, 32768$ datapoints.
+For the Welch algorithm, we employ a Tukey window, an overlap fraction of $0.5$ for the segments and a length of segments $L = 512, 1024, 2048, 8192, 32768$ datapoints, depending on the observation time.
 In all cases, the sampling rate is set to $\SI{4096}{Hz}$.
 For the Welch algorithm, we use the standard implementation provided by the python library \texttt{scipy} \cite{numpy,scipy}.
+The estimations with MESA and Welch's algorithm are reported in fig.~\ref{fig:MESA_LIGO_data} and fig.~\ref{fig:welch_LIGO_data} respectively.
 
@@
 
-We also note that MESA converges to the real spectrum faster than Welch method, providing a better estimate even in the case of short time series.
+We also note that MESA converges to the ``real" spectrum much faster than Welch's method, providing a better estimate even in the case of short time series.
 Although observed at every frequency, this behaviour is more evident in the low frequency region.
 An accurate profile reconstruction can be obtained with MESA using a 5 seconds-strain only, while Welch method requires at least 10 seconds of data to obtain a comparable profile.
+Furthermore, MESA method is able to model all the details of the peak at around $\sim \SI{40}{Hz}$ (even with $T=\SI{100}{s}$), while the Welch's algorithm fails to do so even with an observation time of $T=\SI{1000}{s}$.
 
-Another important element is the noise noise of the spectral estimation: we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
-has a large number of spurious peaks) rather than PSD provided by MESA with FPE loss function.
-If longer datasets are considered, even if MESA is found to be a little more accurate and less noisy than Welch, all of the above differences become negligible.
+Another important element is the noise of the spectral estimation: we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
+has a large number of spurious peaks) rather than PSD provided by MESA with FPE loss function (the picture might change if instead CAT loss function is employed).
+This is especially true at high frequencies and for long observation times $T$.
 
@@
 
-\begin{figure}
-	\caption{Comparison between PSD estimation with Welch method (blue) and Maximum Entropy (red) for \textit{synthetic} LIGO data.
-The estimation is based data with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.
-In dashed black is reported the true PSD (from which data are generated).}
-	\label{fig:mem_welch_fakedata}
-	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_True.pdf}
+\begin{figure}[!t]
+	\caption{Comparison between analytic (dashed line) and estimated (red line) spectrum.
+The estimation is perfomed with Maximum Entropy method on \textit{synthetic} data.
+The estimation is performed with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
+	\label{fig:MESA_LIGO_data}
+	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_MESA.pdf}
 \end{figure}
 
-\begin{figure}
-	\caption{Comparison between PSD estimation with Welch method (blue) and Maximum Entropy (red) for \textit{real} LIGO data.
-The estimation is based data with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
-	\label{fig:mem_welch_realdata}
-	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_overall_fake_False.pdf}
+\begin{figure}[!t]
+	\caption{Comparison between analytic (dashed line) and estimated (green line) spectrum.
+The estimation is perfomed with Welch's method on \textit{synthetic} data.
+The estimation is performed with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
+	\label{fig:welch_LIGO_data}
+	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_Welch.pdf}
 \end{figure}
@@
 As a last example of an application of the maximum entropy method on real data, we come back to the strain produced by the LIGO-Virgo interferometers and we forecast the future observations.
+We employ the data released to the public by the LIGO/Virgo collaboration \cite{Abbott_2021}.,
 The spectral analysis is performed with CAT and FPE loss functions on $\SI{1000}{s}$ of data \cite{Abbott_2021} sampled at $\SI{4096}{Hz}$ observed at Livingston observatory starting from GPS time $1164603392$.

### 2021-04-16 paper/mesa.tex
minor edits

--- 2021-04-15 paper/mesa.tex
@@
 As outlined in sec.~\ref{sec:Welch_comparison}, the MESA provides PSD estimates with less noise than the standard Welch algorithm, thus making it a good alternative for it.
-The use of MESA is particularly useful for short time series samples, where Welch method is outperformed in accuracy and conficence.
+The use of MESA is particularly useful for short time series samples, where Welch method is outperformed in accuracy and confidence.
 It has to be said that our method can be quite computationally expensive, as compared to standard Welch's algorithm. However, for short time series, the computational cost is still managable.
@@
 
-We believe that Maximum Entropy Spectral Analysis, despite being an ``old" theory, has still the potential to become the standard in future time series analysis, due to its theorical elegance and to its reliable performance in a vast range of application.
+We believe that Maximum Entropy Spectral Analysis, despite being an ``old" theory, has still the potential to become the standard in future time series analysis, due to its theoretical elegance and to its reliable performance in a vast range of application.
 

### 2021-04-16 paper/mesa.tex
conclusions and more references

@@
 \section{Final remarks and future prospects} \label{sec:conclusion}
-\sschmidt{Check this section}
-We presented a case study of the application of Maximum Entropy principle in the realm of spectral estimation: it provides a closed form estimation of the power spectral density, bridging the spectral analysis with the study of autoregressive processes.
-The method has the towfold advantage of avoiding any unjustified assumption on the observed time series and of providing a simple context for forecasting the time series.
-We provided a ready to use python code, called \texttt{memspectrum} and we performed numerical studies of its performance.
-We investigated the use of different loss functions for the purpose of correctly estimating the autoregressive order for the estimation and we found that different loss functions are required for different situations.
-Finally, we performed a comparison between PSDs computed with MESA and standard Welch's algorithm and we applied our method to several time series of interest in physics.
-
-As outlined in sec.~\ref{sec:Welch_comparison}, the MESA provides PSD estimates with less noise than the standard Welch algorithm, thus making it a good alternative for it.
+In this work, we presented a case study of the application of Maximum Entropy principle in the realm of spectral estimation.
+Albeit the methodology hereby presented is grounded on solid theoretical foundations and its merits are widely recognised, Maximum Entropy methods have yet to be adopted routinely in the study of problems related to time series.
+The superior nature of maximum entropy methods and in particular of Burg's method is exemplified by the closed form estimate of the power spectral density and by the theoretical bridge it lies between spectral analysis and autoregressive processes.
+Moreover, the method presents, in our view, two main advantages when compared with more traditional ones; first there is no need to choose an arbitrary window function to correct the data and second it provides as straightforward way to compute predictions given past observations.
+Accompanying this work, we provide a publicly available \texttt{Python} implementation, called \texttt{memspectrum}, that we used to perform numerical studies of its behaviour and to  demonstrate its reliability.
+Since the order of the autoregressive process is not determined by the theory, we opted for an in-depth investigation of several proposals in the literature and found that different loss functions are required for different situations, with the FPE loss function being the most indicated to deal with gravitational wave data.
+Along these lines, we directly compared the PSDs computed with MESA with the canonical Welch's algorithm.
+As outlined in sec.~\ref{sec:Welch_comparison}, MESA provides PSD estimates with smaller variance than Welch algorithm.
 The use of MESA is particularly useful for short time series samples, where Welch method is outperformed in accuracy and confidence.
-It has to be said that our method can be quite computationally expensive, as compared to standard Welch's algorithm. However, for short time series, the computational cost is still managable.
-
-Besides providing an alternative to standard methods of spectral estimate, MESA opens up several oppurtunities in the context of time series analysis.
-
-A first application is based on the observation that meaningful PSD estimations can be extracted from short time series (unlike the standard approach).
-This can make the approach particularly conveniente in the context of Bayesian parameter estimation, whenever the likelihood of the model for a time series takes the form
+This observation suggests a promising avenue to pursue in future developments: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalising over the detector noise PSD.
+Whenever the likelihood of the model for a time series takes the form
 \begin{equation}
@@
 \end{equation}
-where $\tilde{\hphantom{a}}$ denotes Fourier transform and $x_t(\theta)$ is a prediction for a signal hidden in the observed time series $d(t)$, dependent on some parameter $\theta$.
-This likelihood corresponds to looking for a signal $x_t$ into some (possibly very) noisy data.
-Usually the PSD is estimated offline on a large batch of data, that might also contain the signal to analyse and in general it might not reflect the structure of noise in the time slice to analyse.
+where $\tilde{\hphantom{a}}$ denotes the Fourier transform and the signal model $x(t;\theta)$, dependent on some parameter $\theta$, is a prediction for a deterministic signal buried in the observed time series $d(t)$.
+The typical approach is to estimate the PSD offline on a large batch of data and off-source (see~\cite{lalinference}), that might not reflect the structure of noise in the time slice to analyse.
 MESA can offer a way out from this problem: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on $d_t-x_t$.
-In this way, the PSD would depend also on $\theta$ and would effectively model the residual of noise subtracted with the signal.
-%To compute the posterior distribution for the parameters the knowledge of prior distributions and Likelihood for noise is required.
-A full description for noise can be obtained in Fourier Domain by the choice of the Whittle Likelihood \eqref{eq:Whittle}, that is completely determined by the knowledge of the PSD for noise.
-While traditional methods might be instable with respect to the hypothesis (such as Welch) or require a Large number of prior assumptions, MESA represents a stable and parsimonious way to characterize noise and compute posterior distributions.
+In this way, the PSD would depend also on $\theta$ and would effectively model the residuals. Preliminary studies suggest that this is possible~\cite{martini_thesis}.
 
@@
 This fact is remarkable and can be used in time series analysis for several purposes.
-As already discussed in sec~\ref{sec:LIGO_forecasting}, an anomaly detection pipeline can be easily built using the forecasts of MESA: the predictions can form a baseline to compare the actual observations with.
+As discussed in Sec.~\ref{sec:LIGO_forecasting}, an anomaly detection pipeline can be easily built using the forecasts of MESA: the predictions can form a baseline to compare the actual observations with.
 Whenever the observed data are outside the expectations, an anomaly detection can be claimed.
 Of course such predictions can be done with a more accurate (perhaps nonlinear model); however MESA has the advantage of being simple and fast to train, while providing good (short term) predictions.
+At the same time, several instruments present gaps in their data stream, for instance LISA is expected to show such gaps~(e.g.
+\cite{lisa_gaps} and references therein), MESA forecasting capabilities could be used to fill those gaps with predicted data from past observations.
 
-\sschmidt{Other stuff with forecasting?}
-
-We believe that Maximum Entropy Spectral Analysis, despite being an ``old" theory, has still the potential to become the standard in future time series analysis, due to its theoretical elegance and to its reliable performance in a vast range of application.
+In conclusion, we reiterate that MESA is a theoretically sound, computationally feasible and reliable way of studying the properties of stochastic processes and we hope that the investigations presented in this work will further stimulate developments and applications of this method.
 

### 2021-05-03 paper/mesa.tex
k fix
rewordings

@@
 A ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
-Basically, the PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
+The PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
 This approach is equivalent~\cite{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
@@
 We are arbitrarily assuming $\rho_j = 0$ for $j > M$ and modifying the estimate (i.e. the data) if a non-rectangular window is chosen.
-Making assumptions on unobserved data and modifying the ones we have at our disposal introduces ``spurious" information about the process that we in general do not really have.
+Making assumptions on unobserved data and modifying the ones we have at our disposal introduces ``spurious" information about the process that we, in general, do not really have.
 
@@
 
-In this paper, we discuss the details of the Maximum entropy principle, its application to the problem of PSD estimation with Burg algorithm and the link between Burg's algorithm and autoregressive process.
+In this paper, we discuss the details of the Maximum entropy principle, its application to the problem of PSD estimation with Burg's algorithm and the link between Burg's algorithm and autoregressive process.
 Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many problems posed by the Welch's algorithm (or other similar methods).
-To facilitate this goal, we present and describe a new code, \texttt{memspectrum}, that provides a robust and easy-to-use python implementation of the algorithm\footnote{
+To facilitate this goal, we present and describe a new implementation, \texttt{memspectrum}, that provides a robust and easy-to-use python implementation of the algorithm\footnote{
 Available at link: \url{https://pypi.org/project/memspectrum/}.
@@
 \section{Theoretical foundations}\label{sec:foundations}
-The Maximum Entropy principle (MAXENT) is among the most important results in Bayesian probability theory.
-It provides a way to uniquely assign probabilities to a phenomenon in a way that best represent our state of knowledge, while at the same time it is non committal with unavailable information.
+The Maximum Entropy principle (MAXENT) is among the most important results in probability theory.
+It provides a way to uniquely assign probabilities to a phenomenon in a way that best represent our state of knowledge, while being non committal with unavailable information.
 Its domain of application turned out to be wider than expected.
@@
 We can generalize the definition of information in the case where two different outcomes $E_1, E_2$, with given probabilities $P_1$ and $P_2$, are possible.
-To gain some intuition on the problem, we ask ourselfs which are the probability assignments that makes the outcome more uncertain (i.e. maximize the information).
-If $P_1$ and $P_2$ are largely different, for istance $P_1 = 0.999$ and $P_2 = 0.001$, we are allowed to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome.
+To gain some intuition on the problem, we ask ourselves which are the probability assignments that makes the outcome more uncertain (i.e. maximize the information).
+If $P_1$ and $P_2$ are largely different, for instance $P_1 = 0.999$ and $P_2 = 0.001$, we are allowed to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome.
 The information content will be very low.
@@
 \end{equation}
-We call the function $H$ (information) entropy.
+We call the functional $H$ (information) entropy\footnote{In defining the information entropy as in Eq.~(\ref{eq:entropy_continuos}) we are implicitly assuming a uniform measure over the parameter space}.
 
-We now turn the core of our problem: how do we make a probability assignment for a set of events, that keeps into account our knoledge of the system and, at the same time, it is non committal with unavailable knowledge?
-The knowledge at our disposal about the system under study is what we call `evidence' and of course, any probability assignment must agree with it.
-In the above cases our knowledge on the system was only the total number $N$ of different outcomes -- this is a minimal requirement. Of course, more complex evidence constraints can be applied.
+We now turn the core of our problem: how do we make a probability assignment for a set of events, that keeps into account our knowledge of the system and, at the same time, it is non committal towards unavailable knowledge?
+The knowledge at our disposal about the system under study is what we call `evidence' and any probability assignment must agree with it.
+In the above cases, our knowledge on the system is only the total number $N$ of different outcomes -- this is a minimal requirement. Of course, more complex evidence constraints can be applied.
 
@@
 MAXENT puts the aforementioned principle into a more precise mathematical context by stating that probabilities should be assigned by maximizing uncertainty (information entropy) using evidence as constraint.
-This defines a variational problem, where the information entropy $H\left[p_1, \dots, p_N\right]$, defined in eq.~\eqref{eq:entropy}, has to maximized.
+This defines a variational problem, where the information entropy functional $H\left[p_1, \dots, p_N\right]$, defined in eq.~\eqref{eq:entropy}, has to maximized.
 
@@
 
-For this work, we are particularly interested in the mutli-dimensional case.
-Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realisation of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of autocovariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution\cite{gregory_2005}:
+For this work, we are particularly interested in the multi-dimensional case.
+Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realisation of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of autocovariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution~\cite{gregory_2005}:
 \begin{align}
@@
 }.
-Toeplitz matrices are asymptotically equivalent to circulant matrices and thus diagonalised by the discrete Fourier transform base\cite{Gray}.
+Toeplitz matrices are asymptotically equivalent to circulant matrices and thus diagonalised by the discrete Fourier transform base~\cite{Gray}.
 %R. M. Gray, Foundations and Trends⃝R in Communications and Information Theory 2, 155 (2005), URL https://doi.org/10.1561/0100000006.
@@
 p&\left((\tilde{x}_1, \ldots, \tilde{x}_n)\vert I\right) = \nonumber \\
-    &\frac{1}{\left(2 \pi \det S\right)^{k / 2}}\exp\left(-\frac{1}{2}\sum_{ij}\tilde{x}_i S^{-1}_{ij} \tilde{x}_j \right)\,,
+    &\frac{1}{\left(2 \pi \det S\right)^{n / 2}}\exp\left(-\frac{1}{2}\sum_{ij}\tilde{x}_i S^{-1}_{ij} \tilde{x}_j \right)\,,
 \end{align}
@@
 This bring us the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function?
-MAXENT can guide use in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method.
+MAXENT can guide us in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method.
 The latter assumes that the unknown values of the autocorrelation are $0$.
@@
 Before doing so, there is a technical problem to solve: the definition of entropy depends on a probability distribution, not on the PSD.
-It can be proved \cite{AblesMESA, Bartlett} that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of filtering a white noise process with a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
+It can be shown, e.g.~\cite{AblesMESA, Bartlett}, that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of the filtering a white noise process using a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
 A filter with transfer function $T(f)$ takes in input a time series $x_t$ and outputs a times series $y_t$ such that:
@@
 where $\Delta t$ is sampling rate and $Ny \equiv \frac{1}{2 \Delta t}$  is the Nyquist frequency.
-Thus maximising eq.~\eqref{eq:EntropyGain} is equivalent to maximising eq.~\eqref{eq:entropy_continuos}.
+Thus maximising Eq.~\eqref{eq:EntropyGain} is equivalent to maximising Eq.~\eqref{eq:entropy_continuos}.
 
@@
 
-Thus, by maximizing eq.~\eqref{eq:EntropyGain} with constraints in eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
+Thus, by maximizing Eq.~\eqref{eq:EntropyGain} with constraints in eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
 This approach on PSD computation provide a result consistent with the epirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
@@
 The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the study
-of autoregressive processes.\cite{doi:10.1029/RG013i001p00183}.
+of autoregressive processes (AR)~\cite{doi:10.1029/RG013i001p00183}.
 An autoregressive stationary process of order $p$ (AR($p$)) is a time series whose values satisfy the following expression: 
@@
 Thus, an AR($p$) process models the dependence of the value of the process at time $t$ on every past $p$ observations, 
-thus being potentially able to model complex autocorrelation between data.
+thus being potentially able to model complex autocorrelation structures within observations.
 
-It turns out that the maximums entropy principle provides a representation of the time series as an $AR(p)$ process.
-The Burg's algorithm computes the autoregressive coefficients that are suitable to the available data.
-Thanks to Wold's theorem \cite{Wold_theorem}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general.
+Thanks to Wold's theorem~\cite{Wold_theorem}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general; it turns out that the maximums entropy principle provides a representation of the time series as an $AR(p)$ process ad Burg's algorithm computes the autoregressive coefficients that are suitable to the available data.
 
-To show the analogy, we compute the PSD $S_{AR(p)}$ of an $AR(p)$ process and we show that it is formally equivalent to the PSD obtained in eq.~\eqref{eq:MESApsd}.
+To show the analogy, we compute the PSD $S_{AR(p)}$ of an $AR(p)$ process and we show that it is formally equivalent to the PSD obtained in Eq.~\eqref{eq:MESApsd}.
 This will also provide a direct expression for the autoregressive coefficients $b_i$ and for the noise variance $\sigma^2$.
-We start taking the $z$ transform~\footnote{We remind the reader that the $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time time-series and returning a frequency series.} of Eq.~\eqref{eq:AR_p}:
+We start taking the $z$ transform~\footnote{The $z$ transform is the discrete-time equivalent of the Laplace transform, thus taking a discrete time-series and returning a complex frequency series.} of Eq.~\eqref{eq:AR_p}:
 \begin{align}
@@
 
-Eq.~\eqref{eq:ARspectrum} and eq.~\eqref{eq:MESApsd} are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
-This shows that the MAXENT estimation of the PSD also computes the coefficients of the autoregressive 
-process that model the observed process.
+Eqs.~\eqref{eq:ARspectrum} and ~\eqref{eq:MESApsd} are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
+This shows that the MAXENT estimation of the PSD also computes the coefficients of the AR 
+process that models the observed process.
 
 \subsection{Forecasting} \label{sec:forecasting}
-The link between MESA and autoregressive processes is of particular interest.
-Given the solution to the Levinson recursion to determine the $a_k$, see Appendix~\ref{sec:appendix}, we authomatically know the coefficients of the equivalent AR process, hence we are able to exploit eq.~\ref{eq:AR_p} to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
+The link between MESA and AR processes is of particular interest.
+Given the solution to Burg's recursion to determine the $a_k$, see Appendix~\ref{sec:appendix}, we automatically obtain the coefficients of the equivalent AR process, hence we are able to exploit Eq.~\ref{eq:AR_p} to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
 Indeed, for an AR($p$) process the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation has the form:
@@
 	p&(x_t|x_{t-1}, \ldots , x_{t-p}) \nonumber\\
-	&= \frac{1}{\sigma\sqrt{2\pi}} \exp\left[-\frac{1}{2} \left(\frac{x_t - \sum_{i = 1}^p b_i x_{t-i}}{\sigma}\right)^2\right].
+	&= \frac{1}{\sigma\sqrt{2\pi}} \exp\left[-\frac{1}{2} \left(\frac{x_t - \sum_{i = 1}^p b_i x_{t-i}}{\sigma}\right)^2\right]\,.
 \end{align}
 The interpretation of Eq.~(\ref{eq:p_forecast}) is straightforward: $x_t$ follows a Gaussian distribution with a fixed variance an a mean value $m_t = \sum_{i = 1}^p b_i x_{t-i}$ computed from past observations.
-Eq.~(\ref{eq:p_forecast}) provides then a well defined probability framework for predicting future observations: this is a very useful feature of MESA, that does not have an equivalent in any other spectral analysis computation.
+Eq.~(\ref{eq:p_forecast}) provides then a well defined probability framework for predicting future observations: this is a very useful feature of MESA, that does not have an equivalent in any other spectral analysis computations.
 
@@
 The number $M$ of such coefficients is the same as the number of time steps considered in the computation of the autocorrelation $\bar{r}_m$.
-In a perfect scenario, this would be equal to the number of points the autocorrelation is computed at (equivalent to the length of data considered).
+In an ideal scenario, this would be equal to the number of points the autocorrelation is computed at (equivalent to the length of data considered).
 However, the computation of high order coefficients of the autocorrelation is unstable and for high enough $m$, as the estimation for  $\bar{r}_m$ shows a very high variance, broadly scaling as $\sim \left(\sqrt{M - m}\right)^{-1}$.
 
-%OR WE CAN KEEP THIS (SOMEHOW MODIFIED)
-%Before being able to apply MESA, we need to address the question of the best length $M$ of the filter that provides the optimal estimate of the power spectrum.
-If the autocorrelation matrix was perfectly know, the answer would be simply $M = N$.
-%In real cases, such a knowledge is not available and one has to estimate the autocorrelation function from the data.
-The accuracy for the estimate decreases as the time lag $k$ increases as $\sim \left(\sqrt{N - k}\right)^{-1}$.
-%The accuracy of the estimate should then increase as the length of the autocorrelation function increases, but, from a certain point on, inaccuracy in the estimate for $R$ becomes so large that the estimate for the spectrum is negatively affected by this fact.
-
-It is then clear that the choice of how many samples of the discrete autocorrelation to consider is important: 
+It is then clear that the choice of the number of samples of the discrete autocorrelation to consider is important: 
 on the one hand it is advisable to include as much knowledge of the autocorrelation as possible, leading to include all the known $\bar{r}_m$; on the other hand, including values of the autocorrelation that are not reliably estimated, can be counterproductive.
-The order $M$ of the autocorrelation to be considered (or equivalently the order $M$ of the underlying autoregressive process) is the only tuning parameter of MESA and a careful balance between this two necessities must be made when applying the algorithm.
+The order $M$ of the autocorrelation to be considered (or, equivalently, the order $M$ of the underlying autoregressive process) is the only tuning parameter of MESA and a careful balance between this two necessities must be made when applying the algorithm.
 
@@
 \begin{itemize}
-\item \textbf{Final prediction Error } 
+\item \textbf{Final prediction Error} 
 The first criterion is due to Akaike~\cite{Akaike1970StatisticalPI}. He proposed that $m$ should be chosen as the 
@@
 \end{equation}
-with $P_m$ being the estimated noise variance at order $m$, see eq.~\eqref{eq:errorFilter1}.
-
-In the $N \to \infty$ limit, remembering $m_{max} \sim 2N / \log(2N)$, Akaike's loss function is equivalent to the minimization of the variance $P_m$ of the white noise of the underlying $AR(p)$ model.
+with $P_m$ being the estimated noise variance at order $m$, see Eq.~\eqref{eq:errorFilter1}. In the $N \to \infty$ limit, 
+remembering $m_{max} \sim 2N / \log(2N)$, Akaike's loss function is equivalent to the minimization of the variance $P_m$ of the white noise of the underlying $AR(p)$ model.
 
 \item \textbf{Criterion Autoregressive Transfer function (CAT)}
-This second loss function has been written by Parzen and studied into details by Banshali ~\cite{bhansali1986}. 
+This second loss function has been proposed by Parzen and studied in detail by Banshali~\cite{bhansali1986}. 
 It is based on the assumption that the observed process is an infinite order autoregressive process 
@@
 \end{equation}
-and the so chosen order is found to be asymptotically unbiased for $P_N$.
+and the so-chosen order is found to be asymptotically unbiased for $P_N$.
 
@@
 time series for the process that has to be described as an autoregressive process of order $m$ ($AR(m)$), with $m$ to be determined.
-The OBD is obtained chosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$.
+The OBD is obtained choosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$.
 Each
 hypothesis is uniquely determined from both the length and the values of the filter under study.
-Chosing a Likelihood of the form \eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypothesis Rao has shown that chosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
+Choosing a Likelihood of the form \eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypotheses, Rao has shown that choosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
 maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of:
@@
 
-Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson \cite{doi:10.1002/sapm1946251261} recursion until $M_{max}$, as given in Eq.~\eqref{eq:MMAx}, iterations are reached.
-Then, the order $m$ is selected to be the one that minimizes the loss function.
+Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson recursion~\cite{doi:10.1002/sapm1946251261} until $M_{max}$, as given in Eq.~\eqref{eq:MMAx}, iterations are reached.
+Then, the order $m$ is selected to be the one that minimizes the specified loss function.
 
@@
 
-We then apply the Burg method choosing in turn each of the loss functions on an ensemble of simulated time series, 
-thus obtaining an ensemble of PSD estimates, that will allows the statistical characterisation of the performance of each loss function.
+We then apply Burg's method choosing in turn each of the loss functions on the ensemble of simulated time series, 
+thus obtaining an ensemble of PSD estimates. Through these ensembles, we characterise statistically the performance of each loss function.
 The disagreement between the PSD $S_i(f)$ estimated from the $i$-th simulated time series and the target PSD $S(f)$ is measured via 
@@
 
-For each loss function, we compute the ensemble-averaged PSD (as well as the $90\%$ confidence level) and the \textit{ensemble-averaged relative error}:
+For each loss function, we compute the ensemble-averaged PSD (as well as the $90\%$ confidence level) and 
+the \textit{ensemble-averaged relative error}:
 \begin{equation}\label{eq:ensemble_averaged_error}
@@
 \end{equation}
-For each loss function, figs.~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} show the averaged PSD, 
+For each loss function, Figs.~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} show the averaged PSD, 
 as well as the ensemble-averaged relative error from Eq.~\eqref{eq:ensemble_averaged_error}.
@@
 order $m_i$ chosen for each independently drawn time series.
-
-%In section~\ref{sec:LIGO_validation}, we will perform the analysis again using the Advanced LIGO design sensitivity PSD \cite{Ligo}.
 
@@
 
-By looking at fig.~\ref{fig:optcomparison} (red series), we note that the autoregressive orders obtained with FPE are clustered in a very small region.
-This is a desirable property: FPE, in fact, provides a stable estimation of the autoregressive order, which does not affect much the reconstruction error.
-We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length $m$ is clustered in a region where there is no dependence of error from $m$.
+By looking at Fig.~\ref{fig:optcomparison} (red series), we note that the AR orders obtained with 
+FPE are clustered in a very small region. This is also a desirable property: FPE, in fact, provides a stable 
+estimation of the ar order, which does not affect much the reconstruction error.
+We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length $m$ is clustered in a region where there is no dependence of error on $m$.
 \paragraph{Optimum Bayes Decision Rule (OBD)}
-The second loss function in consideration is OBD and the results are reported in fig.~\ref{fig:OBDmean}.
+The second loss function we cosider is OBD and the results are summerised in Fig.~\ref{fig:OBDmean}.
 As for the FPE case, they show a good agreement between the average over the reconstructions and the true spectrum.
-Qualitatively the same behaviour of FPE is observed: a good quality reconstruction at the intermediate and high frequencies with a low $90\%$ confidence level; and a worse performance at low frequencies.
-However, when looking at the error, the disagreement of this method is found to be larger than the one obtained with FPE: in the worst case, the error can be as large as a factor of $2$ when compared with FPE.
+Qualitatively the same behaviour of FPE is observed: a good quality reconstruction at the intermediate and high frequencies with a narrow $90\%$ confidence level as well as a degrading performance at low frequencies.
+However, when looking at the error, the disagreement of this method is found to be larger compared to FPE: 
+in the worst case, the error can be as large as a factor of $2$ compared to FPE.
 
-The error as a function of the autoregressive length (green series in Fig.~\ref{fig:optcomparison}) cluster in a small region, indicating the stability of the reconstructed process order $m$.
+The error as a function of the AR length (green series in Fig.~\ref{fig:optcomparison}) clusters in a small region, indicating the stability of the reconstructed process order $m$.
 We note that on average, OBD tends to choose smaller values of $m$ with respect to FPE.
@@
 \paragraph{Criterion Autoregressive Transfer function (CAT)}
-The performance of CAT is shown in fig.~\ref{fig:CATmean}. At a first glance CAT differs substantially from the other two loss functions considered.
+The performance of CAT is shown in Fig.~\ref{fig:CATmean}. At a first glance, CAT differs substantially from the other two loss functions considered.
 The ensemble-average PSD matches very well the underlying ``true" PSD. This is also true even in the low frequency region, 
@@
 
-The reason for this behaviour becomes clear from fig.~\ref{fig:optcomparison} (black series). CAT does not converge 
-to any specific value for the order of the autoregressive process.
-In turn, the estimates for the order span a large range of values, hence the large variance observed.
+The reason for this behaviour becomes clear from Fig.~\ref{fig:optcomparison} (black series): CAT does not converge 
+to any specific value for the order of the AR process. The estimated $m$ spans a large range of values, hence the large variance observed.
 Fig.~\ref{fig:optcomparison} also shows a strong dependence of the error on the estimated length of the filter.
-The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like outliers in different realization of the time-series, but this is also responsible of an increased variance in the estimate, by introducing spurious peaks in the reproduction.
+The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like outliers in different realizations of the time-series, but this is also responsible of an increased variance in the estimate, by introducing spurious peaks in the reproduction.
 
-When averaging the different PSD estimates, the noise in each spectrum cancels, as expected for Gaussian noise. 
+When averaging the different PSD estimates, the noise in each spectrum cancels, as expected from the Gaussian nature of the AR process. 
 This implies, in a sense, that each estimate of the spectrum is independent of any other, as suggested by the huge variance in the residuals.
-This lack of stability is not a good property for the estimation estimation of the PSD from a single realisation of the time-series, however, thanks to the averaging out of the errors, this estimator seems optimal in the case of repeatable experiments and ensemble-averages.
+This lack of stability is not a good property for the estimation of the PSD from a single realisation of the time-series, however, thanks to the averaging out of errors, this estimator seems optimal in the case of repeatable experiments and ensemble-averages.
 
@@
 
-However, we also note that the when inferring PSDs from a single time-series realisation, FPE provide the lowest averaged error over all frequencies, while CAT can reach errors $5$ times larger (see again fig.~\ref{fig:optcomparison}).
+However, we also note that the when inferring PSDs from a single time-series realisation, FPE provide the lowest averaged error over all frequencies, while CAT can reach errors $5$ times larger (see again Fig.~\ref{fig:optcomparison}).
 Hence, while CAT is the loss function that minimizes errors in the low-frequency end of the spectrum, FPE obtains the best overall accuracy.
 
-The conclusion is clear: even if in some cases CAT is more accurate when taking the average over several estimation, FPE guarantees that the single estimation is more faithful.
+The conclusion is clear: even if in some cases CAT is more accurate when taking the average over several realisations of the underlying process, 
+FPE guarantees that the single estimation is more faithful.
 As in any common situation we cannot perform such averaging over different realization of the same time series, we must prefer FPE over CAT (let alone OBD, which even though qualitatively similar to FPE has worse performance).
@@
 of the time-series to $\SI{20}{s}$. The chosen length is convenient to capture fairly accurately the low-frequency features of the LIGO PSD.
-We report our findings in figs.~\ref{fig:ligospectrum}~and~\ref{fig:LigoOrderError}.
+We report our findings in Figs.~\ref{fig:ligospectrum}~and~\ref{fig:LigoOrderError}.
 
@@
      \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/compare_estimates.pdf}
-      \caption{Average spectrum for the three different loss function as compared with the ``true" PSD. The average is computed with $500$ realization of a $40960$ points long time series.}
+      \caption{Average spectrum for the three different loss function as compared with the ``true" PSD. The average is computed over $500$ realization of a $40960$ points long time series.}
        \label{fig:ligospectrum}
@@
 
-Figure~\ref{fig:ligospectrum} shows the original spectrum (dashed line) and the ensemble-averaged reconstructed PSD adopting the FPE (green line), OBD(red line) and CAT(black line).
+Fig.~\ref{fig:ligospectrum} shows the simulated spectrum (dashed line) and the ensemble-averaged reconstructed PSD adopting the FPE (green line), OBD(red line) and CAT(black line).
 In all cases, the spectrum is well reconstructed, but with a fairly distinct behaviour at low frequency, where CAT -- as in the Gaussian case -- better captures and resolves the distinct spectral feature at $\sim \SI{17}{Hz}$.
-In fig.~\ref{fig:ligoPeaks}, we report the reconstructed spectra around the two peaks at $\sim \SI{17}{Hz}$ and $\sim \SI{438}{Hz}$ (left and right panel respectively).
+In Fig.~\ref{fig:ligoPeaks}, we report the reconstructed spectra around the two peaks at $\sim \SI{17}{Hz}$ and $\sim \SI{438}{Hz}$ (left and right panel respectively).
 
@@
 Yet again, when the PSD is averaged over multiple realization of the time, CAT is able to capture the spectrum very precisely.
-In fact, even in presence of very sharp peaks, CAT reconstruction seems to be almost perfectly coincident with each of them.
-Hence, also the study of simulated Advanced LIGO data seems to indicate that whenever and wherever ensemble-averaged PSD are necessary, CAT is the optimal choice of loss function.
+In fact, even in presence of very sharp spectral features, CAT reconstruction seems to be almost perfectly coincident with each of them.
+Hence, also the study of simulated LIGO data seems to indicate that whenever and wherever ensemble-averaged PSD are necessary, CAT is the optimal choice of loss function.
 However, on a single time-series realisation, FPE is the more robust choice.
@@
 \begin{itemize}
-	\item There are no reasons to prefer OBD over CAT or FPE.
-	\item If we have one single realization for the process, we recommend the use of FPE, that would get the best resolution possible.
-In this situation, CAT would provide spurious and unreliable results, with large error.
-	\item In the case of several realizations of the same process, CAT ensemble-average properties provide very precise spectral estimation.\end{itemize}
+	\item There are no reasons to prefer OBD over CAT or FPE;
+	\item if we have one single realization for the process, we recommend the use of FPE, that would get the best resolution possible.
+In this situation, CAT would provide spurious and unreliable results, with large error;
+	\item in the case of several realizations of the same process, CAT ensemble-average properties provide very a precise spectral estimation.\end{itemize}
 
@@
 
-\subsection{How well the autoregressive order is recovered?}
+\subsection{How well is the AR order recovered?}
 
@@
 
-We now address the issue of how well the autoregressive order (i.e. the number of $a_k$ coefficients employed) is estimated with each loss function.
+We now address the issue of how well the AR order (i.e. the number of $a_k$ coefficients employed) is estimated by each loss function.
 In doing so, we generate $100$ autoregressive processes $AR(p)$ with a random value of $p$, drawn such that $log(p) \sim\mathcal{U}_{[log(2), log(5000)]}$.
-Each coefficient $a_k$ is assigned according to a Dirichlet distribution $\mathrm{Dir}([1,\hdots, 1])$. The sign of $a_k$ is assigned randomly according to a binomial distribution.
-We report the result of this investigation in fig.~\ref{fig:p_vs_ptrue}.
+Each coefficient $a_k$ is assigned according to a Dirichlet distribution $\mathrm{Dir}([1,\hdots, 1])$.
+The sign of $a_k$ is assigned randomly according to a binomial distribution.
+We report the result of this investigation in Fig.~\ref{fig:p_vs_ptrue}.
 
 It can be seen that OBD and FPE loss functions show very similar behaviour.
-They are {\it very} realiable in capturing the correct autoregressive order up to a certain threshold ($p\sim 100$ for OBD and $p\sim 200$ for FPE).
-For any autoregressive process of order higher than the threshold, the optimization underestimates badly the actual value of $p$.
-For this reason, FPE and OBD seems realiable only for simple autoregressive processes.
-For complicated processes, it seems they underfit the problem (i.e.
+They are {\it very} realiable in capturing the correct AR order up to a certain threshold ($p\sim 100$ for OBD and $p\sim 200$ for FPE).
+For any AR process of order higher than the threshold, the optimization underestimates badly the actual value of $p$.
+For this reason, FPE and OBD seem reliable only for relatively simple AR processes.
+For more complicated processes, they seem to ``underfit'' the problem (i.e.
 they output a model simpler that those required).
 
-On the other hand, CAT loss function shows a different behaviour.
-The selected autoregressive order $p$ is always close to the maximum possible value, regardless the actual value $p_{true}$ of the underlying process.
-This produces always a model that is more complex than those obtained with FPE and OBD and this can explain the origin of the high variance in the estimation observed in the experiments described in the previous section.
-Of course, CAT does not perform a good job in reconstructing an autoregressive process.
-However, for high order autoregressive process the error introduced by CAT is more tolerable than the error introduced by FPE and OBD.
-This is a good reason to prefer CAT optimizer in such situations (see sec.~\ref{sec:temp_timeseries} for another example of this).
+On the other hand, CAT shows a different behaviour.
+The selected AR order $p$ is always close to the maximum possible value, regardless the actual value $p_{true}$ of the underlying process.
+This produces always a model that is more complex than those obtained with FPE and OBD and this can explain the origin of the high variance in the estimation observed in the experiments described in the previous sections.
+Of course, CAT does not perform a good job in reconstructing the AR process.
+However, for high order AR processes, the error introduced by CAT is more tolerable than the error introduced by FPE and OBD.
+This is a good reason to prefer CAT in these situations (see also Sec.~\ref{sec:temp_timeseries} for another example of this effect).
 
 \section{Comparison with Welch method}\label{sec:Welch_comparison}
-It is really interesting to have a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
-We perform such comparison on the spectrum of LIGO (even though similar conclusions can be drawn from any other PSD).
+In this section we perform a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
+For definiteness, we focus on LIGO[which one?], although similar conclusions can be drawn from any other PSDs.
 
-In order to ensure the we have a baseline PSD to compare the data with, we simulate data from the realesed LIGO intererferometer PSD and we employ both Welch's method and MESA to estimate the PSD.
+In order to ensure the we have a baseline PSD to compare the data with, we simulate data from the released LIGO intererferometer PSD and we employ both Welch's method and MESA to estimate the PSD.
 We vary the length of the data used for the estimation: this is also useful to assess how the computation depends on the data available.
 We set the total observation time $T = 1, 5, 10, 100, 1000 \SI{}{s}$
-For the MESA algorithm, we employ FPE optimizer.
-For the Welch algorithm, we employ a Tukey window, an overlap fraction of $0.5$ for the segments and a length of segments $L = 512, 1024, 2048, 8192, 32768$ datapoints, depending on the observation time.
+For the MESA algorithm, we choose the FPE loss function. For the Welch algorithm, we employ a Tukey window with a parameter $\alpha$ 
+equal to 0.4, an overlap fraction of $1/2$ for the segments and a length of 
+segments $L = 512, 1024, 2048, 8192, 32768$ points, depending on the observation time.
 In all cases, the sampling rate is set to $\SI{4096}{Hz}$.
-For the Welch algorithm, we use the standard implementation provided by the python library \texttt{scipy} \cite{numpy,scipy}.
-The estimations with MESA and Welch's algorithm are reported in fig.~\ref{fig:MESA_LIGO_data} and fig.~\ref{fig:welch_LIGO_data} respectively.
+For the Welch algorithm, we use the standard implementation provided by the python library \texttt{scipy}~\cite{numpy,scipy}.
+The results from both methods are summarised in  Figs.~\ref{fig:MESA_LIGO_data} and~\ref{fig:welch_LIGO_data} respectively.
 
-First of all, we note that using a longer time series results in a more stable estimation of the PSD, especially in the low frequency sector.
-This is somehow obvious: more data provide more information available for spectral estimation.
-This fact is more important at lower frequencies, where the natural time scale is large.
-Indeed, the higher frequencies are alway better resolved than the lower ones: there is alway more information in the high frequency part of a signal.
+First of all, we note that using a longer time series results in a better estimation of the PSD, especially at low frequencies. 
+This is somehow obvious: longer data streams probe lower frequencies thanks to Nyquist's theorem as well as providing better estimates for the FFT, in the Welch case, and the sample autocorrelation, for MESA.
 
-We also note that MESA converges to the ``real" spectrum much faster than Welch's method, providing a better estimate even in the case of short time series.
+We also note that MESA converges to the underlying spectrum much faster than Welch's method, providing a better estimate even in the case of short time series.
 Although observed at every frequency, this behaviour is more evident in the low frequency region.
 An accurate profile reconstruction can be obtained with MESA using a 5 seconds-strain only, while Welch method requires at least 10 seconds of data to obtain a comparable profile.
-Furthermore, MESA method is able to model all the details of the peak at around $\sim \SI{40}{Hz}$ (even with $T=\SI{100}{s}$), while the Welch's algorithm fails to do so even with an observation time of $T=\SI{1000}{s}$.
+Furthermore, MESA is able to model all the details of the peak at around $\sim \SI{40}{Hz}$ (even with $T=\SI{100}{s}$), while the Welch's algorithm fails to do so even with an observation time of $T=\SI{1000}{s}$.
 
 Another important element is the noise of the spectral estimation: we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
-has a large number of spurious peaks) rather than PSD provided by MESA with FPE loss function (the picture might change if instead CAT loss function is employed).
+has a large number of spurious peaks) compared to the PSD measured with MESA and FPE loss function.
 This is especially true at high frequencies and for long observation times $T$.
 
-Finally, as already discussed Welch, method is very instable with respect to the windowing choices we made.
-We tried to find the best compromize between noise and accuracy for the reconstruction, but different choices can be made, possibly providing more accurate results.
-We want to stress that this fact doesn't invalidate our discussion: one of the most appealing advantages of MESA is the absence of fine tuning.
-%A more accurate reconstruction with Welch can only be obtained introducing a greater amount of noise: a greater accuracy requires longer sub-segments of data, reducing the number of spectra over which we average.
-Reducing noise require the opposite choices, and we can conclude that MESA represents a better compromise between noise and accuracy than Welch's method.
-The result does not change much by changing the hyperparameters required for Welch algorithm.
+Finally, as already discussed Welch's method is very dependent on the choice of window function. 
+Our choice, the Tukey window with aforementioned parameters, is the best compromise with found between noise 
+and accuracy for the reconstruction, but different choices can be made, possibly providing more accurate results than the ones reported here. 
+However, we want to stress that this fact does not invalidate our discussion but it reinforces it: one of the most appealing advantages of MESA is the minimal amount of fine tuning required.
 
@@
 \section{Applications} \label{sec:applications}
-\subsection{Temperature Time Series} \label{sec:temp_timeseries}
+\subsection{Temperature Time Series} \label{sec:temp_timeseries}[review]
 We apply our code to the study of the temperature time series.
@@
 \subsection{Forecasting the LIGO strain} \label{sec:LIGO_forecasting}
-As a last example of an application of the maximum entropy method on real data, we come back to the strain produced by the LIGO-Virgo interferometers and we forecast the future observations.
+As a last example of an application of MESA on real data, we go back to the strain produced by the LIGO-Virgo interferometers and we forecast the future observations.
 We employ the data released to the public by the LIGO/Virgo collaboration \cite{Abbott_2021}.,

### 2021-05-05 paper/mesa.tex
a final pass through the paper

@@
 Finally, as already discussed Welch's method is very dependent on the choice of window function. 
-Our choice, the Tukey window with aforementioned parameters, is the best compromise with found between noise 
+Our choice, a Tukey window with aforementioned parameters, is the best compromise with found between noise 
 and accuracy for the reconstruction, but different choices can be made, possibly providing more accurate results than the ones reported here. 
@@
 \section{Applications} \label{sec:applications}
-\subsection{Temperature Time Series} \label{sec:temp_timeseries}[review]
-We apply our code to the study of the temperature time series.
-This will allow to test our model performance on an example of great importance in physics.
-We will also assess the quality of the forecasting predictions: this task has a great impact also on practical applications.
-We used the historical reanalysis data from the ``ERA5-Land hourly data from 1981 to present" dataset, downloaded from the Climate Data Store \cite{ERA5_dataset}.
-Data refers to the temperature measure at coordinates $\textrm{N}\;\ang{45;5;} \;\; \textrm{E}\;\ang{9;1;}}$, corresponding about to the city of Milan, Italy.
-The temperature is measured once per hour for almost 31 years from 31st December 1989 to 30th November 2020.
+\subsection{Temperature Time Series} \label{sec:temp_timeseries}
+As a further example of the breadth of applicability of MESA, we applied our implementation to atmospheric temperature time series. 
+The reason for this choice is twofold: i) atmospheric temperature time series present a variety of overlapping periodicities most of which are 
+known; ii) it provides a stress test for the time series forecast analysis.  As dataset, we  used the historical reanalysis data 
+from the ``ERA5-Land hourly data from 1981 to present" dataset, downloaded from the Climate Data Store~\cite{ERA5_dataset}.
+The data consist in temperatures taken at coordinates $\textrm{N}\;\ang{45;5;} \;\; \textrm{E}\;\ang{9;1;}}$, corresponding about to the city of Milan, Italy.
+The temperatures are given on an hourly cadence for almost 31 years from 31st December 1989 to 30th November 2020.
 
-In fig.~\ref{fig:temp_spectrum}, we compute the spectrum over all the data. We employ both FPE and CAT loss functions.
-First of all, we note that at every frequency the estimation given by FPE is much less noisy the one from CAT: this issue was already studied in the previous sections. 
-We clearly see that there is a peak on the PSD at the frequency $f_D = \SI{1}{day^{-1}}$ corresponding to a day. This is somehow trivial and highly expected.
+Fig.~\ref{fig:temp_spectrum}, shows the MESA spectrum inferred from the data. As a comparison, we adopt both the FPE and the CAT loss functions.
+In agreement with what we found in previous sections, the FPE PSD is more regular compared to the one from CAT. Both spectra show a
+peak at the frequency $f_D = \SI{1}{day^{-1}}$ corresponding to one day, corresponding to the day-night cycle. This is expected.
 Higher order harmonics (corresponding to integer multiples of $f_D$) are also visible up to the Nyquist frequency ($f_D = \SI{0.5}{hour^{-1}}$): they corresponds to signals that preserve the 1 day period while, at the same time, capturing the complex variability of the daily temperature cycle throughout the year.
 
-Looking at the low frequencies, it is striking to note that FPE does not capture the peak $f_{yr} = \SI{1}{yr^{-1}}$ located at the frequency of a year.
+Looking at low frequencies, FPE does not capture the yearly variability at $f_{yr} = \SI{1}{yr^{-1}}$.
 On the other hand, the peak is captured well by the CAT loss function.
-In analogy as what observed for the peak at $f_D$, in the CAT spectrum we observe a higher order harmonic at a frequency $2 \cdot f_{yr}$.
+In analogy as what observed for the peak at $f_D$, in the CAT spectrum we observe a higher order harmonics at a frequency $2 \cdot f_{yr}$.
 As in the previous case, this is required to better model the temperature variation in the year.
 
-The failure of FPE in catching the low frequency trend can be understood by looking at fig.~\ref{fig:p_vs_ptrue}.
+The failure of FPE in capturign the low frequency trend can be understood by looking at Fig.~\ref{fig:p_vs_ptrue}.
 It is shown that CAT systematically select very long filters (i.e.
-large $p_{CAT}$ values for the autoregressive orders), whereas FPE tends to select smaller values for the order $p_{FPE}$ of the autoregressive process, often understimating badly the actual value.
+large $p_{CAT}$ values for the autoregressive orders), whereas FPE tends to select smaller values for the order $p_{FPE}$ of the autoregressive process, often underestimating the actual value.
 In this example, we have $p_{CAT} \sim  36500 \sim  \SI{4}{yr}$, whereas $p_{FPE} \sim  1300 \sim  \SI{2}{months}$.
-Clearly, the autoregressive process selected by FPE cannot produce a meaningful prediction on the timescale longer than a few weeks and for this reason it is unable to capture the peak at $f_{yr}$.
+The autoregressive process selected by FPE, hence cannot produce a meaningful prediction on the timescale longer than a few weeks and for this reason it is unable to capture the low frequency peak at $f_{yr}$.
 In other words, the model is too simple to model \textit{both} the behaviour at high and low frequencies (separated by approximately 3 orders of magnitude).
@@
 The predictions spans a two years range of time.
-We report our results in fig.~\ref{fig:temp_spectrum}.
+We report our results in Fig.~\ref{fig:temp_spectrum}.
 
@@
 \subsection{Forecasting the LIGO strain} \label{sec:LIGO_forecasting}
-As a last example of an application of MESA on real data, we go back to the strain produced by the LIGO-Virgo interferometers and we forecast the future observations.
-We employ the data released to the public by the LIGO/Virgo collaboration \cite{Abbott_2021}.,
-The spectral analysis is performed with CAT and FPE loss functions on $\SI{1000}{s}$ of data \cite{Abbott_2021} sampled at $\SI{4096}{Hz}$ observed at Livingston observatory starting from GPS time $1164603392$.
+As a last example of an application of MESA, we return to the study of the strain produced by the LIGO-Virgo interferometers 
+and we forecast the future observations. We focus on the public data released by the LIGO/Virgo collaboration~\cite{Abbott_2021}.
+We reconstruct the PSD both assuming the CAT and FPE loss functions on $\SI{1000}{s}$ of data \cite{Abbott_2021} from the Livingston observatory
+starting from GPS time $1164603392$. The data are sampled at $\SI{4096}{Hz}$.
+
 We then forecast the following $\SI{100}{s}$ of observations with the model optimized with CAT. The results are shown in fig.~\ref{fig:LIGO_forecast}.
+The prediction is always in the $90\%$ confidence interval. The confidence interval, however, is very broad and the prediction is not very accurate (being the same order of magnitude of the strain).
+By looking at the standard deviation $\sigma$ of the predictions (bottom panel in Fig.~\ref{fig:LIGO_forecast}), we note that it increases very quickly in the first $\SI{0.5}{s}$ and then it converges to a constant value.
+The order of the autoregressive process selected is $p_{CAT} = 57766 \sim \SI{14}{s}$, which is much larger than the region in which $\sigma$ is small $\sim \SI{0.2}{s}$.
+The implication is that the series is very difficult to predict: the knowledge of past observations is not very helpful to predict future observations.
+Although FPE selects a prediction filter order $p_{FPE} = 29924$ smaller than CAT,  the forecasted time series behaves very similarly.
 
-By looking at the figure, we note that the forecasted prediction is always in the $90\%$ confidence interval.
-The confidence interval, however, is very broad and the prediction is not very accurate (being the same order of magnitude of the strain).
-By looking at the standard deviation $\sigma$ of the predictions (bottom panel in fig.~\ref{fig:LIGO_forecast}), we note that it increases very quickly in the first $\SI{0.5}{s}$ and then its trend stabilizes to a constant value.
-The order of the autoregressive process selected is $p_{CAT} = 57766 \sim \SI{14}{s}$, which is much larger than the region in which $\sigma$ is small $\sim \SI{0.2}{s}$.
-This means that the series is very difficult to predict: the knowledge of past observations is not very helpful in predicting future observations.
-Similar performances in $\sigma$ are observed also with the FPE model: in this case FPE selects a large prediction filter $p_{FPE} = 29924$ , similar to those selected by CAT.
-
-Despite poor predictions on long timescales, the autoregressive process obtained with MESA still can be useful on short timescales.
-Indeed, a precise prediction of the strain time series can be beneficial in anomaly detection and removal of glitches.
+Despite poor predictions on long timescales, the AR process obtained with MESA still can be useful on short timescales.
+Indeed, a precise prediction of the strain time series can be beneficial in the detection of anomalies in the data and, eventually, their removal.
 The predictions can form an expected baseline for the strain; any anomaly (i.e.
@@
 Moreover, if a glitch is detected, its shape can be estimated (as well as the confidence level) by subracting the expected signal with the actual signal.
-This can work: a typical glitch \cite{Zevin_2017} can last as long as $\SI{0.2}{s}$, which is close to the time scale in which the forecasting is realiable.
+This can work: a typical glitch~\cite{Zevin_2017} can last as long as $\SI{0.2}{s}$, close to the time scale over which the forecasting is reliable.
 
@@
 \cite{LSTM_article, oord2016wavenet}) and they can be trained for precision forecasting.
-However, a Maximum-Entropy trained autoregressive process can be a simple baseline for comparing such more advanced methods and might suffice for many purposes.
+However, a Maximum-Entropy trained AR process can be a simple baseline for comparing such more advanced methods and might suffice for many purposes.
 This opens a promising path in GW data analysis and other fields of physics might also take advantage from this.
@@
 \section{Final remarks and future prospects} \label{sec:conclusion}
-In this work, we presented a case study of the application of Maximum Entropy principle in the realm of spectral estimation.
+We presented a case study of the application of Maximum Entropy principle to the realm of spectral estimation.
 Albeit the methodology hereby presented is grounded on solid theoretical foundations and its merits are widely recognised, Maximum Entropy methods have yet to be adopted routinely in the study of problems related to time series.
-The superior nature of maximum entropy methods and in particular of Burg's method is exemplified by the closed form estimate of the power spectral density and by the theoretical bridge it lies between spectral analysis and autoregressive processes.
-Moreover, the method presents, in our view, two main advantages when compared with more traditional ones; first there is no need to choose an arbitrary window function to correct the data and second it provides as straightforward way to compute predictions given past observations.
-Accompanying this work, we provide a publicly available \texttt{Python} implementation, called \texttt{memspectrum}, that we used to perform numerical studies of its behaviour and to  demonstrate its reliability.
-Since the order of the autoregressive process is not determined by the theory, we opted for an in-depth investigation of several proposals in the literature and found that different loss functions are required for different situations, with the FPE loss function being the most indicated to deal with gravitational wave data.
+The superior nature of maximum entropy methods, and in particular of Burg's method, is exemplified by the closed form estimate of the power spectral density and by the theoretical bridge between spectral analysis and AR processes.
+Moreover, the method presents, in our view, two main advantages when compared with more traditional ones; first there is no need to choose an arbitrary window function to correct the data and, second it provides as straightforward way to compute predictions given past observations.
+Accompanying this work, we provide a publicly available \texttt{Python} implementation, called \texttt{memspectrum}, that we used to perform the numerical studies presented in this work..
+
+Since the order of the AR process is not yet determined by the theory, we opted for an in-depth investigation of several proposals in the literature and found that different loss functions are required for different situations, with the FPE loss function being the most indicated to deal with gravitational wave data.
 Along these lines, we directly compared the PSDs computed with MESA with the canonical Welch's algorithm.
-As outlined in sec.~\ref{sec:Welch_comparison}, MESA provides PSD estimates with smaller variance than Welch algorithm.
-The use of MESA is particularly useful for short time series samples, where Welch method is outperformed in accuracy and confidence.
+As outlined in Sec.~\ref{sec:Welch_comparison}, MESA provides PSD estimates with smaller variance than Welch algorithm.
+The use of MESA is particularly useful for short time series samples, where Welch's method is outperformed in accuracy and confidence.
 This observation suggests a promising avenue to pursue in future developments: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalising over the detector noise PSD.
-Whenever the likelihood of the model for a time series takes the form
+Whenever the log-likelihood of the model for a time series takes the form
 \begin{equation}
-	\mathcal{L}(d_t | \theta) \propto exp\left[ -\frac{1}{2} \frac{(\tilde{d}(f)- \tilde{x}(f;\theta))^2}{S(f)} \right]
+	\log\mathcal{L}(d_t | \theta) \propto  -\frac{1}{2} \frac{(\tilde{d}(f)- \tilde{x}(f;\theta))^2 - \frac{1}{2}\log\left[\int  df\,S(f)} \right]}
 \end{equation}
@@
 
-Furthermore, MESA provides a simple, but robust and quite accurate in the short term, predictor for the time series.
+Furthermore, MESA provides a simple, but robust and quite accurate, albeit for short times, predictor for the time series.
 This fact is remarkable and can be used in time series analysis for several purposes.
-As discussed in Sec.~\ref{sec:LIGO_forecasting}, an anomaly detection pipeline can be easily built using the forecasts of MESA: the predictions can form a baseline to compare the actual observations with.
+As discussed in Sec.~\ref{sec:LIGO_forecasting}, an anomaly detection pipeline could be built using the forecasts of MESA: the predictions can form a baseline to compare the actual observations with.
 Whenever the observed data are outside the expectations, an anomaly detection can be claimed.
-Of course such predictions can be done with a more accurate (perhaps nonlinear model); however MESA has the advantage of being simple and fast to train, while providing good (short term) predictions.
+Of course such predictions can be done with a more accurate (perhaps nonlinear model); however MESA has the advantage of being simple and fast to construct, while providing good predictions.
 At the same time, several instruments present gaps in their data stream, for instance LISA is expected to show such gaps~(e.g.

### 2021-05-05 memspectrum.py
Changes after LL discussion

@@ -68,7 +68,7 @@ class loss_function:
         elif self.method =='AIC':
             return self._AIC(args[0], args[2], args[3])
         elif self.method =='LL':
-            return self._LL(args[4])
+            return self._LL(args[3], args[4])
         elif self.method == 'Fixed':
             return self._Fixed(args[3])
         else:
@@ -132,12 +132,14 @@ class loss_function:
         """
         return np.log(P[-1]) + (2 * m) / N
     
-    def _LL(self, spectrum):
+    def _LL(self, m, spectrum):
         """
         Implements data log-likelihood as a loss function
         Log-likelihood is composed of two terms A = 0.5*|x_f|**2/S(f) and B = 0.5*N*log(S(f))
-        A wants to be large (whyyyy?) to agree with data; |B| wants to be small to keep the model simple
-        The loss function to minimize is the balance of the two: L =  -A+|B|
+        A wants to be large to agree with data; |B| wants to be small to keep the model simple
+        The loss function to minimize is the balance of the two: L =  -A - B
+        
+        #WARNING: Untested: the performance of this loss function are not studied
         
         Parameters
         ----------
@@ -155,11 +157,13 @@ class loss_function:
         if self.data_autocorr is None:
             raise ValueError("Autocorrelation of data shall be set, before computing the LL loss function")
         assert spectrum.shape == self.data_autocorr.shape, "Log-likelihood calculation: shape of the spectrum and of data do not match"
+        T = len(spectrum)*4096
         autocorr = 0.5* np.sum(np.divide(self.data_autocorr, spectrum))
         det_S = 0.5*len(spectrum)*np.sum(np.log(spectrum))
-        LL =  - autocorr + np.abs(det_S)
-#        print(autocorr, det_S, LL) #DEBUG
-        return LL #this is a loss function, we want to minimize it (maybe)
+        LL =  - autocorr/T - det_S
+        #LL -= m*np.log(2)+np.log(m)
+        #print(autocorr, det_S, m*np.log(2)+np.log(m), LL) #DEBUG
+        return -LL #this is a loss function, we want to minimize it (maybe)
         
     
     def _CAT(self, P, N, m):
@@ -542,6 +546,9 @@ class MESA(object):
         
         if optimisation_method == 'Fixed':
             self.early_stop = False
+        
+        if optimisation_method == 'LL':
+        	warnings.warn('The performance of the \'LL\' loss function are not studied. There is no guarantee of stable and meaningful results')
            
         if method.lower() == "fast":
             self._method = self._FastBurg

### 2021-05-05 scripts/test_opt.py

@@ -6,32 +6,33 @@ import numpy as np
 import scipy.stats 
 import matplotlib.pyplot as plt
 
-data = np.loadtxt('../../GWAnomalyDetection/maxent/data/H-H1_GWOSC_4KHZ_R1-1126259447-32.txt.gz')
-#data = np.loadtxt('../climate_data/data_@1h_len95689.dat')
+data = np.loadtxt('../../GWAnomalyDetection/maxent/data/H-H1_GWOSC_4KHZ_R1-1126259447-32.txt.gz')[:int(5*4096)]
+data = np.loadtxt('../climate_data/data_@1h_len95689.dat')[:24*4*365]
 
-plot_type = 'LIGO' #'climate' #'LIGO'
+plot_type = 'climate' #'climate' #'LIGO'
 srate = 1. #4096.
 
 m_FPE = MESA()
 m_LL = MESA()
 
 print(plot_type)
+#P, a_k_LL, opt_LL = m_LL.solve(data, optimisation_method = 'LL', early_stop = False, m = 10000, verbose = True)
 
-if False:
+if True:
 	P, a_k_FPE, opt_FPE = m_FPE.solve(data, optimisation_method = 'FPE', verbose = False)
-	m_FPE.save('LL_validation/FPE_model_{}'.format(plot_type))
-	np.savetxt('LL_validation/FPE_opt_{}'.format(plot_type), opt_FPE)
+	m_FPE.save('LL_validation_call/FPE_model_{}_minLL'.format(plot_type))
+	np.savetxt('LL_validation_call/FPE_opt_{}_minLL'.format(plot_type), opt_FPE)
 
-	P, a_k_LL, opt_LL = m_LL.solve(data, optimisation_method = 'LL', early_stop = False, m = 10000, verbose = True)
-	m_LL.save('LL_validation/LL_model_{}'.format(plot_type))
-	np.savetxt('LL_validation/LL_opt_{}'.format(plot_type), opt_LL)
+	P, a_k_LL, opt_LL = m_LL.solve(data, optimisation_method = 'LL', early_stop = False, m = 10000, verbose = False)
+	m_LL.save('LL_validation_call/LL_model_{}_minLL'.format(plot_type))
+	np.savetxt('LL_validation_call/LL_opt_{}_minLL'.format(plot_type), opt_LL)
 else:
[…]

### 2021-05-05 scripts/white_noise_LL.py
Added white noise LL

@@ -1,0 +1,33 @@
+import sys
+sys.path.insert(0, '..')
+from memspectrum import MESA, loss_function
+from GenerateTimeSeries import generate_data 
+import numpy as np
+import scipy.stats 
+import matplotlib.pyplot as plt
+
+data = np.random.normal(0,1,10000)
+srate = 1.
+
+m_FPE = MESA()
+m_LL = MESA()
+
+P, a_k_FPE, opt_FPE = m_FPE.solve(data, optimisation_method = 'FPE', early_stop = False, verbose = True)
+
+P, a_k_LL, opt_LL = m_LL.solve(data, optimisation_method = 'LL', early_stop = False, m = None, verbose = True)
+
+
+plt.figure()
+plt.loglog(*m_LL.spectrum( 1./srate, onesided = True), label = 'LL', zorder = 10)
+plt.loglog(*m_FPE.spectrum( 1./srate, onesided = True), label = 'FPE',zorder = 0)
+plt.legend()
+
+fig, ax = plt.subplots(2,1, sharex = True)
+
+ax[0].plot(range(len(opt_LL)), opt_LL/np.abs(opt_LL[0]), label = 'LL')
+ax[0].axvline(len(a_k_LL), c = 'r')
+ax[1].plot(range(len(opt_FPE)), opt_FPE/np.abs(opt_FPE[0]), label = 'FPE')
+ax[1].axvline(len(a_k_FPE), c = 'r')
+ax[1].set_yscale('log')
+plt.legend()
+plt.show()

### 2021-05-06 scripts/optimisers_comparison.py
Changes to plots and memspectrum.py

@@ -231,7 +231,8 @@ for method in methods:
 ax.legend(loc = 'upper left')
 ax.set_xlabel("Filter Length estimate")
 ax.set_ylabel('Frequency averaged error')
-fig.savefig(savedir + 'error_length_contour.pdf', bbox_inches = 'tight')
+fig.savefig(savedir + 'error_length_contour.pdf', bbox_inches = 'tight', transparent = True)
+print("saved file @ ", savedir + 'error_length_contour.pdf')
 
         
 plt.show()

### 2021-05-07 paper/mesa.tex
Edits in paper

@@
 We characterize our implementation by performing a power spectral density analysis on syntethic data (with known power spectral density) and we compare different criteria for stopping the recursion.
-Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using both syntethic data and real gravitational wave data from the LIGO-Virgo collaboration.
+Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using syntethic data generated from the released spectrum by the LIGO-Virgo collaboration.
 We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
@@
 
-The study of the properties of stochastic processes is a crucial task in many fields of physics, astronomy, quantitative biology, astronomy as well as engineering and finance.
-Of particular interest for several areas of physics is the study of stochastic time series.
+The study of the properties of stochastic processes is a crucial task in many fields of physics, astronomy, quantitative biology, as well as engineering and finance.
 Among those, a special place is occupied by the so-called \textit{wide-sense} stationary processes.
@@
 
-A ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
+An ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
 The PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
@@
 Besides the window function, Welch's method requires a number of arbitrary choices to be made, such as the number of time slices and the overlap between consecutive slices.
-All these knobs must be tuned by hand and their choice can dramatically affect the PSD estimation, hence begging the question of what the ``real" PSD is.
+All these knobs must be tuned by hand and their choice can dramatically affect the PSD estimation, hence begging the question of what the ``best" PSD estimate is.
 
@@
 Lastly and most importantly, it provides a robust link between spectral density estimation and the field of autoregressive processes.
-This provide a natural and simple machinery to forecast a time series, thus predicting future observations based on previous ones.
+This provides a natural and simple machinery to forecast a time series, thus predicting future observations based on previous ones.
 
@@
 Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many problems posed by the Welch's algorithm (or other similar methods).
-To facilitate this goal, we present and describe a new implementation, \texttt{memspectrum}, that provides a robust and easy-to-use python implementation of the algorithm\footnote{
-Available at link: \url{https://pypi.org/project/memspectrum/}.
+To facilitate this goal, we present and describe a new code, \texttt{memspectrum}, that provides a robust and easy-to-use python implementation of the algorithm\footnote{
+It is available at link: \url{https://pypi.org/project/memspectrum/}.
 }.
@@
 Sec.~\ref{sec:validation} presents the validation of Burg's method as well as of our implementation on simulated data.
-In Sec.~\ref{sec:Welch_comparison} we compare the results from \texttt{memspectrum} with the Welch method; Sec.~\ref{sec:applications} presents a few applications to time series and, finally, we conclude with a discussion in Sec.~\ref{sec:conclusion}.
+In Sec.~\ref{sec:Welch_comparison} we compare the results from \texttt{memspectrum} with the Welch method; Sec.~\ref{sec:applications} presents a few applications to real time series and, finally, we conclude with a discussion in Sec.~\ref{sec:conclusion}.
 
@@
 
-Before introducing MAXENT principle, we will define through some simple examples the two core concepts of the problem and the role they play: the 'evidence' and the 'information'.
-Let us start with the 'information' (or entropy): it is a measure of the degree of uncertainty on the outcomes of some experiment and specifies the length of the message necessary to provide a full description of the system under study.
+Before introducing MAXENT principle, we will define through some simple examples the two core concepts of the problem and the role they play: the `evidence' and the `information'.
+Let us start with the `information' (or entropy): it is a measure of the degree of uncertainty on the outcomes of some experiment and specifies the length of the message necessary to provide a full description of the system under study.
 For instance, consider a perfectly sinusoidal signal: knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information content.
@@
 \end{equation}
-this describes a situation of `maximum ignorance': the information content of this system must be high.
+this describes a situation of `maximum ignorance' and the information content of such system must be high.
 Any generalization of eq.~\eqref{eq:information}, must then have its maximum when $ P_1 = P_2$.
@@
 
-Of particular importance is the case in which, in addition to the mean, also the variance is known. MAXENT leads to the Gaussian distribution. 
+Of particular importance is the case in which, in addition to the mean, also the variance is known: MAXENT leads to the Gaussian distribution. 
 We retain this particularly interesting from the foundational point of view, since it provides a deeper insight into the ubiquitous Gaussian distribution.
-Indeed, the Gaussian distribution is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy.
+Indeed, it is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy.
 For this reason, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process.
@@
 This is clearly understood: when computing the autocorrelation at order $k$, only $N-k$ examples of the product $x_t x_{t+k}$ are available and the variance of the average value goes as the inverse of the square root of the points considered.
-}, so that only a few values for the autocorrelation function can actually be computed reliably.
+}, so that only few values for the autocorrelation function can actually be computed reliably.
 This bring us the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function?
@@
 
-As in the previous examples, one needs to set up a variational problem where the entropy eq.~\eqref{eq:entropy} is maximized 
+As in the previous examples, one needs to set up a variational problem where the entropy eq.~\eqref{eq:entropy_continuos} is maximized 
 subject to some problem-specific constraints. 
@@
 where $\Delta t$ is sampling rate and $Ny \equiv \frac{1}{2 \Delta t}$  is the Nyquist frequency.
-Thus maximising Eq.~\eqref{eq:EntropyGain} is equivalent to maximising Eq.~\eqref{eq:entropy_continuos}.
+Thus maximising Eq.~\eqref{eq:EntropyGain} is equivalent to maximise Eq.~\eqref{eq:entropy_continuos}.
 
@@
 Thus, by maximizing Eq.~\eqref{eq:EntropyGain} with constraints in eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
-This approach on PSD computation provide a result consistent with the epirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
+This approach on PSD computation provides a result consistent with the epirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
+
 Most importantly, the variational problem admits a closed-form analytical expression for $S(f)$.
@@
 The coefficients $P_N$ and $a_s (s > 0)$ are to be determined by an iterative process (called Burg's algorithm).
-The details of the derivation and the actual form for the coefficients $a_s$ can be found in appendix \ref{sec:MESA_solution}.
+The details of the derivation and the actual form for the coefficients $a_s$ can be found in appendix \ref{sec:appendix}.
 
@@
 of autoregressive processes (AR)~\cite{doi:10.1029/RG013i001p00183}.
-An autoregressive stationary process of order $p$ (AR($p$)) is a time series whose values satisfy the following expression: 
+An autoregressive stationary process of order $p$, AR($p$), is a time series whose values satisfy the following expression: 
 \begin{equation} \label{eq:AR_p}
@@
 Eqs.~\eqref{eq:ARspectrum} and ~\eqref{eq:MESApsd} are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
-This shows that the MAXENT estimation of the PSD also computes the coefficients of the AR 
-process that models the observed process.
+This shows that the MAXENT estimation of the PSD models the observed times series as an AR process and provides a {\it fit} for the autoregressive coefficients.
+Furthermore, as a consequence of Wold's theorem, there is the theoretical guarantee that every stationary time series can be modelled faithfully by the MAXENT.
 
@@
 The link between MESA and AR processes is of particular interest.
-Given the solution to Burg's recursion to determine the $a_k$, see Appendix~\ref{sec:appendix}, we automatically obtain the coefficients of the equivalent AR process, hence we are able to exploit Eq.~\ref{eq:AR_p} to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
+Given the solution to Burg's recursion to determine the $a_k$, we automatically obtain the coefficients of the equivalent AR process, hence we are able to exploit Eq.~\ref{eq:AR_p} to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
 Indeed, for an AR($p$) process the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation has the form:
@@
 The interpretation of Eq.~(\ref{eq:p_forecast}) is straightforward: $x_t$ follows a Gaussian distribution with a fixed variance an a mean value $m_t = \sum_{i = 1}^p b_i x_{t-i}$ computed from past observations.
-Eq.~(\ref{eq:p_forecast}) provides then a well defined probability framework for predicting future observations: this is a very useful feature of MESA, that does not have an equivalent in any other spectral analysis computations.
+Eq.~(\ref{eq:p_forecast}) provides then a well defined probability framework for predicting future observations: this is a very useful feature of MESA, that does not have an equivalent in any other spectral analysis computation methods.
 
 \section{Validation of the model}\label{sec:validation}
-It is now clear that MESA provides a recursive formula for computing the coefficients $a_k$ in Eq.~(\ref{eq:MESApsd}).
-The number $M$ of such coefficients is the same as the number of time steps considered in the computation of the autocorrelation $\bar{r}_m$.
+It is now clear that MESA provides a recursive formula for computing the coefficients $a_k$ in Eq.~(\ref{eq:MESApsd}). The number $M$ of such coefficients is
+equivalent to the maximum order of the autocorrelation $\bar{r}_m$ considered.
 In an ideal scenario, this would be equal to the number of points the autocorrelation is computed at (equivalent to the length of data considered).
@@
 by measuring the loss function and pick, among the orders the one that yields better results.
-Second, the performance of different loss functions will be assessed by performing the spectral analysis on time series with given analytical PSD, Sec.~\ref{sec:arp_validation}.
-In a last subsection~\ref{sec:LIGO_validation}, the same analysis is performed on real data from the LIGO-Virgo gravitational waves observatories.
+Second, the performance of different loss functions will be assessed by performing the spectral analysis on time series with an analytical gaussian PSD, sec.~\ref{sec:arp_validation}.
+In a last subsection~\ref{sec:LIGO_validation}, the same analysis is performed on syntetic data generated from the analytical spectrum released by the LIGO-Virgo collaboration.
 
@@
 \item \textbf{Final prediction Error} 
-The first criterion is due to Akaike~\cite{Akaike1970StatisticalPI}. He proposed that $m$ should be chosen as the 
+The first criterion is due to Akaike~\cite{Akaike1970StatisticalPI}. It was proposed that $m$ should be chosen as the 
 length that minimises the error when the filter is used as a predictor, the \emph{final prediction} error (FPE): 
@@
 The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
-time series for the process that has to be described as an autoregressive process of order $m$ ($AR(m)$), with $m$ to be determined.
+time series for the process that has to be described as an $AR(m)$ process, with $m$ to be determined.
 The OBD is obtained choosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$.
@@
 hypothesis is uniquely determined from both the length and the values of the filter under study.
-Choosing a Likelihood of the form \eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypotheses, Rao has shown that choosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
+Choosing a Likelihood of the form eq.~\eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypotheses, Rao has shown that choosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
 maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of:
@@
 
-Fig.~\ref{fig:ligospectrum} shows the simulated spectrum (dashed line) and the ensemble-averaged reconstructed PSD adopting the FPE (green line), OBD(red line) and CAT(black line).
+Fig.~\ref{fig:ligospectrum} shows the simulated spectrum (dashed line) and the ensemble-averaged reconstructed PSD adopting the FPE (green line), OBD (red line) and CAT (black line).
 In all cases, the spectrum is well reconstructed, but with a fairly distinct behaviour at low frequency, where CAT -- as in the Gaussian case -- better captures and resolves the distinct spectral feature at $\sim \SI{17}{Hz}$.
@@
 \begin{itemize}
-	\item There are no reasons to prefer OBD over CAT or FPE;
+	\item there are no reasons to prefer OBD over CAT or FPE;
 	\item if we have one single realization for the process, we recommend the use of FPE, that would get the best resolution possible.
@@
 In this section we perform a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
-For definiteness, we focus on LIGO[which one?], although similar conclusions can be drawn from any other PSDs.
+Although similar studies can be drawn from any other PSD, for definiteness, we focus on the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog \cite{GWTC1}.
 
-In order to ensure the we have a baseline PSD to compare the data with, we simulate data from the released LIGO intererferometer PSD and we employ both Welch's method and MESA to estimate the PSD.
+We simulate data\footnote{This is to ensure that we have a baseline PSD to compare the data with} from the PSD around the event GW150914 and we employ both Welch's method and MESA to estimate the spectrum.
 We vary the length of the data used for the estimation: this is also useful to assess how the computation depends on the data available.
@@
 and accuracy for the reconstruction, but different choices can be made, possibly providing more accurate results than the ones reported here. 
-However, we want to stress that this fact does not invalidate our discussion but it reinforces it: one of the most appealing advantages of MESA is the minimal amount of fine tuning required.
+However, we want to stress that this fact does not invalidate our discussion but reinforces it: one of the most appealing advantages of MESA is the minimal amount of fine tuning required.
 
@@
 In agreement with what we found in previous sections, the FPE PSD is more regular compared to the one from CAT. Both spectra show a
-peak at the frequency $f_D = \SI{1}{day^{-1}}$ corresponding to one day, corresponding to the day-night cycle. This is expected.
+peak at the frequency $f_D = \SI{1}{day^{-1}}$ corresponding to one day, corresponding to the day-night cycle: this is expected.
 Higher order harmonics (corresponding to integer multiples of $f_D$) are also visible up to the Nyquist frequency ($f_D = \SI{0.5}{hour^{-1}}$): they corresponds to signals that preserve the 1 day period while, at the same time, capturing the complex variability of the daily temperature cycle throughout the year.
@@
 
-The failure of FPE in capturign the low frequency trend can be understood by looking at Fig.~\ref{fig:p_vs_ptrue}.
+The failure of FPE in capturing the low frequency trend can be understood by looking at Fig.~\ref{fig:p_vs_ptrue}.
 It is shown that CAT systematically select very long filters (i.e.
@@
 In other words, the model is too simple to model \textit{both} the behaviour at high and low frequencies (separated by approximately 3 orders of magnitude).
-On the other hand, the model chosen by CAT is much more complex and is able to model also the high frequency behaviour.
+On the other hand, the model chosen by CAT is much more complex and is able to model also the high frequency behaviour, at the expense of making a more noisy estimation.
 The ``actual" length of the autoregressive process should be closer to the choice of CAT.
@@
 The prediction is always in the $90\%$ confidence interval. The confidence interval, however, is very broad and the prediction is not very accurate (being the same order of magnitude of the strain).
-By looking at the standard deviation $\sigma$ of the predictions (bottom panel in Fig.~\ref{fig:LIGO_forecast}), we note that it increases very quickly in the first $\SI{0.5}{s}$ and then it converges to a constant value.
+By looking at the standard deviation $\sigma$ of the predictions (bottom panel in Fig.~\ref{fig:LIGO_forecast}), we note that it increases very quickly in the first $\SI{0.5}{s}$.
 The order of the autoregressive process selected is $p_{CAT} = 57766 \sim \SI{14}{s}$, which is much larger than the region in which $\sigma$ is small $\sim \SI{0.2}{s}$.
 The implication is that the series is very difficult to predict: the knowledge of past observations is not very helpful to predict future observations.
-Although FPE selects a prediction filter order $p_{FPE} = 29924$ smaller than CAT,  the forecasted time series behaves very similarly.
+Although FPE selects an autoregressive order $p_{FPE} = 29924 \sim \SI{7}{s}$ smaller than CAT,  the forecasted time series behaves very similarly.
 
@@
 Along these lines, we directly compared the PSDs computed with MESA with the canonical Welch's algorithm.
-As outlined in Sec.~\ref{sec:Welch_comparison}, MESA provides PSD estimates with smaller variance than Welch algorithm.
-The use of MESA is particularly useful for short time series samples, where Welch's method is outperformed in accuracy and confidence.
-This observation suggests a promising avenue to pursue in future developments: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalising over the detector noise PSD.
+As outlined in Sec.~\ref{sec:Welch_comparison}, MESA provides PSD estimates with smaller variance and better accuracy than Welch algorithm.
+The use of MESA is particularly useful for short time series samples, where Welch's method is outperformed in both precision and confidence.
+
+This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalising over the detector noise PSD.
 Whenever the log-likelihood of the model for a time series takes the form
 \begin{equation}
-	\log\mathcal{L}(d_t | \theta) \propto  -\frac{1}{2} \frac{(\tilde{d}(f)- \tilde{x}(f;\theta))^2 - \frac{1}{2}\log\left[\int  df\,S(f)} \right]}
+	\log\mathcal{L}(d_t | \theta) \propto  -\frac{1}{2} \frac{(\tilde{d}(f)- \tilde{x}(f;\theta))^2}{S(f)} - \frac{1}{2}\log\left[\int  df\,S(f)\right]
 \end{equation}
@@
 Whenever the observed data are outside the expectations, an anomaly detection can be claimed.
-Of course such predictions can be done with a more accurate (perhaps nonlinear model); however MESA has the advantage of being simple and fast to construct, while providing good predictions.
+Of course such predictions can be done with a more accurate (perhaps nonlinear model); however MESA has the advantage of being simple and fast to construct, while providing decent predictions.
 At the same time, several instruments present gaps in their data stream, for instance LISA is expected to show such gaps~(e.g.
@@
 \end{equation}
-For a wide-sense stationary processes \sschmidt{Justify this assumption!}, all the poles lay outside the unit circle so that the previous integral can be easily computed obtaining the following, well known, equations:
+For a wide-sense stationary processes, all the poles lay outside the unit circle so that the previous integral can be easily computed obtaining the following, well known, equations: 
 \begin{align}

### 2021-05-07 scripts/plot_LIGO_forecast.py

@@ -54,15 +54,15 @@ plt.close()
 fig, (ax,ax2) = plt.subplots(nrows=2, sharex=True)
 
 t_grid = np.linspace(0,N_forecast/srate,N_forecast)
-ax.plot(t_grid, -(true-m), c = 'b', zorder = 1)
+ax.plot(t_grid, -(true-m), c = 'b', zorder = 1, lw = .4)
 ax.fill_between(t_grid, -true+h, -true+l, color = 'r',alpha = 0.3, zorder = 0)
 #ax.plot(t_grid, (true-m)/(h-l), c = 'b', zorder = 1)
 #ax.plot(np.linspace(0,N_forecast,N_forecast), true-h, c = 'r')
 #ax.plot(np.linspace(0,N_forecast,N_forecast), true- l, c = 'r')
 ax.set_ylabel(r"$h_{forecast} -h$")
 
-ax2.plot(t_grid, sigma, c = 'k', zorder = 1, label = 'CAT', lw = 1)
-ax2.plot(t_grid, sigma_FPE, c = 'r', zorder = 0, label = 'FPE', lw = 1)
+ax2.plot(t_grid, sigma, c = 'k', zorder = 1, label = 'CAT', lw = .7)
+ax2.plot(t_grid, sigma_FPE, c = 'r', zorder = 0, label = 'FPE', lw = .7)
 plt.legend(loc= 'upper left')
 
 ax2.set_xlabel("Time (s)")
@@ -70,7 +70,7 @@ ax2.set_ylabel(r"$\sigma$")
 
 axins =  ax2.inset_axes([0.6, 0.15, 0.5, 0.35])
 N_inset = int(0.5*srate)
-axins.plot(t_grid[:N_inset], -(true-m)[:N_inset], c = 'b', zorder = 1)
+axins.plot(t_grid[:N_inset], -(true-m)[:N_inset], c = 'b', zorder = 1, lw = 0.5)
 axins.fill_between(t_grid[:N_inset], (-true+h)[:N_inset], (-true+l)[:N_inset], color = 'r',alpha = 0.3, zorder = 0)
 axins.set_ylabel(r"$h_{forecast} -h$")
 

### 2021-05-07 scripts/plot_climate_data.py

@@ -40,13 +40,13 @@ fig = init_plotting()
 ax = fig.gca()
 
 t_grid = np.linspace(0,N_forecast/24.,N_forecast)
-ax.plot(t_grid, -(true-m), c = 'b', zorder = 1)
+ax.plot(t_grid, -(true-m), c = 'b', zorder = 1, lw = .5)
 ax.fill_between(t_grid, -true+h, -true+l, color = 'r',alpha = 0.3, zorder = 0)
 #ax.plot(t_grid, (true-m)/(h-l), c = 'b', zorder = 1)
 #ax.plot(np.linspace(0,N_forecast,N_forecast), true-h, c = 'r')
 #ax.plot(np.linspace(0,N_forecast,N_forecast), true- l, c = 'r')
 ax.set_xlabel("Time (days)")
-ax.set_ylabel(r"$T_{forecast} -T (K)$")
+ax.set_ylabel(r"$[T_{forecast} -T]\; (K)$")
 plt.tight_layout()
 plt.savefig(plot_dir+"forecast_accuracy.pdf")
 
@@ -55,7 +55,7 @@ fig = init_plotting()
 ax = fig.gca()
 
 unit_shift = 3600*24 #frequency is made 1/day
-ax.loglog(f_CAT * unit_shift , spec_CAT, label = "CAT", c = 'b', zorder = 1)
+ax.loglog(f_CAT * unit_shift , spec_CAT, label = "CAT", c = 'k', zorder = 1)
 ax.loglog(f_FPE * unit_shift , spec_FPE, label = "FPE", c = 'r', zorder = 2)
 
 ax.axvline(1./(3600.*24.*365.)*unit_shift,lw = .5, c = 'b', ls = '--', zorder = 0)

### 2021-05-07 paper/mesa.tex
Fixed typos

@@
 We provide a ready-to-use implementation of the algorithm in the form of a python package \texttt{memspectrum}.
-We characterize our implementation by performing a power spectral density analysis on syntethic data (with known power spectral density) and we compare different criteria for stopping the recursion.
-Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using syntethic data generated from the released spectrum by the LIGO-Virgo collaboration.
+We characterize our implementation by performing a power spectral density analysis on synthetic data (with known power spectral density) and we compare different criteria for stopping the recursion.
+Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using synthetic data generated from the released spectrum by the LIGO-Virgo collaboration.
 We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
@@
 An ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
-The PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
+The PSD is obtained by slicing the observed realization $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
 This approach is equivalent~\cite{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
@@
 gives rise to several of the most common probability distributions commonly employed in statistics.
-For instance, whenever the only constraint available is the normalisation of the probability distribution (i.e.
+For instance, whenever the only constraint available is the normalization of the probability distribution (i.e.
 no evidence is available), the entropy is maximised by the uniform distribution.
@@
 For this work, we are particularly interested in the multi-dimensional case.
-Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realisation of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of autocovariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution~\cite{gregory_2005}:
+Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realization of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of autocovariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution~\cite{gregory_2005}:
 \begin{align}
@@
 }.
-Toeplitz matrices are asymptotically equivalent to circulant matrices and thus diagonalised by the discrete Fourier transform base~\cite{Gray}.
+Toeplitz matrices are asymptotically equivalent to circulant matrices and thus diagonalized by the discrete Fourier transform base~\cite{Gray}.
 %R. M. Gray, Foundations and Trends⃝R in Communications and Information Theory 2, 155 (2005), URL https://doi.org/10.1561/0100000006.
@@
 where the matrix $S_{ij} = S_i \delta_{ij}$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
-Many readers will recognise the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\cite{prob_information_theory}
+Many readers will recognize the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\cite{prob_information_theory}
 %P. M. Woodward, D. W. Fry, and W. Higinbotham, Probability and information theory, with applications to radar; 2nd ed. (Pergamon Press, Oxford, 1964), URL http://cds.cern.ch/record/2031792.
@@
 where $\Delta t$ is sampling rate and $Ny \equiv \frac{1}{2 \Delta t}$  is the Nyquist frequency.
-Thus maximising Eq.~\eqref{eq:EntropyGain} is equivalent to maximise Eq.~\eqref{eq:entropy_continuos}.
+Thus maximising Eq.~\eqref{eq:EntropyGain} is equivalent to maximize Eq.~\eqref{eq:entropy_continuos}.
 
-Before maximising the entropy gain, we need to include the evidence available as a form of mathematical constraints for the assignment of $S(f)$.
+Before maximizing the entropy gain, we need to include the evidence available as a form of mathematical constraints for the assignment of $S(f)$.
 This is equivalent in imposing that the variational solution $S(f)$ for the PSD matches the empirical autocorrelation.
-Let us define a realisation of a stochastic process $(x_1,\ldots,x_N)$ with sample autocorrelations $\bar r_k,\,k=0,\ldots, N/2$, then the PSD must satisfy the following equation:
+Let us define a realization of a stochastic process $(x_1,\ldots,x_N)$ with sample autocorrelations $\bar r_k,\,k=0,\ldots, N/2$, then the PSD must satisfy the following equation:
 \begin{equation}\label{eq:MaxConstraint}
@@
 Thus, by maximizing Eq.~\eqref{eq:EntropyGain} with constraints in eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
-This approach on PSD computation provides a result consistent with the epirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
+This approach on PSD computation provides a result consistent with the empirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
 
@@
 by measuring the loss function and pick, among the orders the one that yields better results.
-Second, the performance of different loss functions will be assessed by performing the spectral analysis on time series with an analytical gaussian PSD, sec.~\ref{sec:arp_validation}.
-In a last subsection~\ref{sec:LIGO_validation}, the same analysis is performed on syntetic data generated from the analytical spectrum released by the LIGO-Virgo collaboration.
+Second, the performance of different loss functions will be assessed by performing the spectral analysis on time series with an analytical Gaussian PSD, sec.~\ref{sec:arp_validation}.
+In a last subsection~\ref{sec:LIGO_validation}, the same analysis is performed on synthetic data generated from the analytical spectrum released by the LIGO-Virgo collaboration.
 
@@
 The first criterion is due to Akaike~\cite{Akaike1970StatisticalPI}. It was proposed that $m$ should be chosen as the 
-length that minimises the error when the filter is used as a predictor, the \emph{final prediction} error (FPE): 
+length that minimizes the error when the filter is used as a predictor, the \emph{final prediction} error (FPE): 
 \begin{equation}
@@
 We then apply Burg's method choosing in turn each of the loss functions on the ensemble of simulated time series, 
-thus obtaining an ensemble of PSD estimates. Through these ensembles, we characterise statistically the performance of each loss function.
+thus obtaining an ensemble of PSD estimates. Through these ensembles, we characterize statistically the performance of each loss function.
 The disagreement between the PSD $S_i(f)$ estimated from the $i$-th simulated time series and the target PSD $S(f)$ is measured via 
@@
 \paragraph{Optimum Bayes Decision Rule (OBD)}
-The second loss function we cosider is OBD and the results are summerised in Fig.~\ref{fig:OBDmean}.
+The second loss function we consider is OBD and the results are summarized in Fig.~\ref{fig:OBDmean}.
 As for the FPE case, they show a good agreement between the average over the reconstructions and the true spectrum.
@@
 This implies, in a sense, that each estimate of the spectrum is independent of any other, as suggested by the huge variance in the residuals.
-This lack of stability is not a good property for the estimation of the PSD from a single realisation of the time-series, however, thanks to the averaging out of errors, this estimator seems optimal in the case of repeatable experiments and ensemble-averages.
+This lack of stability is not a good property for the estimation of the PSD from a single realization of the time-series, however, thanks to the averaging out of errors, this estimator seems optimal in the case of repeatable experiments and ensemble-averages.
 
@@
 
-However, we also note that the when inferring PSDs from a single time-series realisation, FPE provide the lowest averaged error over all frequencies, while CAT can reach errors $5$ times larger (see again Fig.~\ref{fig:optcomparison}).
+However, we also note that the when inferring PSDs from a single time-series realization, FPE provide the lowest averaged error over all frequencies, while CAT can reach errors $5$ times larger (see again Fig.~\ref{fig:optcomparison}).
 Hence, while CAT is the loss function that minimizes errors in the low-frequency end of the spectrum, FPE obtains the best overall accuracy.
 
-The conclusion is clear: even if in some cases CAT is more accurate when taking the average over several realisations of the underlying process, 
+The conclusion is clear: even if in some cases CAT is more accurate when taking the average over several realizations of the underlying process, 
 FPE guarantees that the single estimation is more faithful.
@@
 \subsection{Choice of the loss function: LIGO Spectrum} \label{sec:LIGO_validation}
-We continue our characterisation of the various loss functions considered in this work, by investigating the 
+We continue our characterization of the various loss functions considered in this work, by investigating the 
 reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve\cite{LIGO_PSDs}.
@@
 Hence, also the study of simulated LIGO data seems to indicate that whenever and wherever ensemble-averaged PSD are necessary, CAT is the optimal choice of loss function.
-However, on a single time-series realisation, FPE is the more robust choice.
+However, on a single time-series realization, FPE is the more robust choice.
 
-Let us summarise some key general conclusions:
+Let us summarize some key general conclusions:
 \begin{itemize}
@@
 It can be seen that OBD and FPE loss functions show very similar behaviour.
-They are {\it very} realiable in capturing the correct AR order up to a certain threshold ($p\sim 100$ for OBD and $p\sim 200$ for FPE).
+They are {\it very} reliable in capturing the correct AR order up to a certain threshold ($p\sim 100$ for OBD and $p\sim 200$ for FPE).
 For any AR process of order higher than the threshold, the optimization underestimates badly the actual value of $p$.
@@
 For the Welch algorithm, we use the standard implementation provided by the python library \texttt{scipy}~\cite{numpy,scipy}.
-The results from both methods are summarised in  Figs.~\ref{fig:MESA_LIGO_data} and~\ref{fig:welch_LIGO_data} respectively.
+The results from both methods are summarized in  Figs.~\ref{fig:MESA_LIGO_data} and~\ref{fig:welch_LIGO_data} respectively.
 
@@
 	\caption{Comparison between analytic (dashed line) and estimated (red line) spectrum.
-The estimation is perfomed with Maximum Entropy method on \textit{synthetic} data.
+The estimation is performed with Maximum Entropy method on \textit{synthetic} data.
 The estimation is performed with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
@@
 	\caption{Comparison between analytic (dashed line) and estimated (green line) spectrum.
-The estimation is perfomed with Welch's method on \textit{synthetic} data.
+The estimation is performed with Welch's method on \textit{synthetic} data.
 The estimation is performed with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
@@
 Based on actual data, we try to predict future values as described in sec.~\ref{sec:forecasting} (of course the MESA is performed with CAT loss function).
-We produce $N =100$ indepedent predictions and we compute the median as well as the $90\%$ confidence interval.
-This is compared with the actual measured temparature values.
+We produce $N =100$ independent predictions and we compute the median as well as the $90\%$ confidence interval.
+This is compared with the actual measured temperature values.
 The predictions spans a two years range of time.
@@
 Such a simple model hardly catches the complex trend in the variability of the temperature daily trend during a year.
-For more precise predictions, probably one should consider nonlinear regression model, tapping into the wealth of non linear predictors offered by the field of Deep Lerning.
+For more precise predictions, probably one should consider nonlinear regression model, tapping into the wealth of non linear predictors offered by the field of Deep Learning.
 
@@
 a glitch or even a transient of physical origin) can show up as a large departure from expected trend.
-Moreover, if a glitch is detected, its shape can be estimated (as well as the confidence level) by subracting the expected signal with the actual signal.
+Moreover, if a glitch is detected, its shape can be estimated (as well as the confidence level) by subtracting the expected signal with the actual signal.
 This can work: a typical glitch~\cite{Zevin_2017} can last as long as $\SI{0.2}{s}$, close to the time scale over which the forecasting is reliable.
@@
 We presented a case study of the application of Maximum Entropy principle to the realm of spectral estimation.
-Albeit the methodology hereby presented is grounded on solid theoretical foundations and its merits are widely recognised, Maximum Entropy methods have yet to be adopted routinely in the study of problems related to time series.
+Albeit the methodology hereby presented is grounded on solid theoretical foundations and its merits are widely recognized, Maximum Entropy methods have yet to be adopted routinely in the study of problems related to time series.
 The superior nature of maximum entropy methods, and in particular of Burg's method, is exemplified by the closed form estimate of the power spectral density and by the theoretical bridge between spectral analysis and AR processes.
@@
 
-This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalising over the detector noise PSD.
+This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
 Whenever the log-likelihood of the model for a time series takes the form
@@
 \end{equation}
-Note that this expression already takes into account the contraints in eq.~\eqref{eq:MaxConstraint}.
+Note that this expression already takes into account the constraints in eq.~\eqref{eq:MaxConstraint}.
 
 We now introduce a set of coefficients $\lambda_s$, defined as the derivative of $\Delta H$ with respect to the autocorrelation function $r_s$.
-Excplitely they are:
+Explicitly they are:
 \begin{equation} \label{eq:lamdas}
@@
 \end{equation}
-where $b$ holds the value of the $a_s$ coeffiecients at order $N-1$. 
+where $b$ holds the value of the $a_s$ coefficients at order $N-1$. 
 The 0-th order element can be easily initialized reminding that $a_0 = 1$ (always) and that $P_0$ can be determined from \eqref{eq:errorFilter1}.
@@
 These expressions allow us to compute $\vec a$ and $P_N$ to any order by simply iterating \eqref{eq:Levinson1} and \eqref{eq:Levinson2}.
-Substituing them in equation \eqref{eq:MESApsd_appendix} the problem of the estimation for the power spectral density via maximum entropy principle is solved.
-Burg's method for spectral analysis is solved via Levinson is implmenteed in the released \texttt{memspectrum} package.
+Substituting them in equation \eqref{eq:MESApsd_appendix} the problem of the estimation for the power spectral density via maximum entropy principle is solved.
+Burg's method for spectral analysis is solved via Levinson is implemented in the released \texttt{memspectrum} package.
 Another faster recursion method is available in \cite{Vos} and it is also available in \texttt{memspectrum}.

### 2021-05-10 paper/mesa.tex
Fixed mem-welch comparison

@@
 	\maketitle
-	\tableofcontents
+	%\tableofcontents %comment this to hide the table of content
 
@@
 
-\begin{figure}[!t]
+\begin{figure*}
+
+\begin{minipage}{0.99\columnwidth}
 	\caption{Comparison between analytic (dashed line) and estimated (red line) spectrum.
-The estimation is performed with Maximum Entropy method on \textit{synthetic} data.
-The estimation is performed with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
+The estimation is performed with Maximum Entropy method on \textit{synthetic} data, with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
 	\label{fig:MESA_LIGO_data}
 	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_MESA.pdf}
-\end{figure}
-
-\begin{figure}[!t]
+\end{minipage}\hfill
+\begin{minipage}{0.99\columnwidth}
 	\caption{Comparison between analytic (dashed line) and estimated (green line) spectrum.
-The estimation is performed with Welch's method on \textit{synthetic} data.
-The estimation is performed with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
+The estimation is performed with Welch's method on \textit{synthetic} data with an increasing observation time $T = 1, 5, 10, 100, \SI{1000}{s}$.}
 	\label{fig:welch_LIGO_data}
 	\includegraphics{Images/comparison_LVC_data/comparison_LVC_data_Welch.pdf}
-\end{figure}
+\end{minipage}
+
+\end{figure*}
 

### 2021-05-26 paper/referee_response.md
First adds after referee report

# Referee response

We thank the referee for their precious work and for the comment they made.
We read carefully throught the proposed changes and we amended the text accordingly. The referee can find below thes answers to their questions/concerns

## PnP comments

- Please add citations to the LIGO and Virgo instrument papers the first time these detectors are mentioned (1411.4547, 1408.3978).
		> Done: page 9

- Please fix the broken citation at the beginning of section IV to "the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog”
		> Done, added also the link to DCC entry where the PSD are taken from

- Please link to the correct citation at the start of Section VB, as [27] links to the climate data. "We focus on the public data released by the LIGO/Virgo collaboration [27]. We reconstruct the PSD both assuming the CAT and FPE loss functions on 1000s of data [27] from the Livingston observatory starting from GPS time 1164603392. The data are sampled at 4096 Hz.”
		> Done

- "Moreover, if a glitch is detected, its shape can be estimated (as well as the confidence level) by subtracting the expected signal with the actual signal.” Here the Bayeswave papers should be cited (1410.3835, 1410.3852) as well as other key LIGO papers describing glitches, their properties, and subtraction (1508.07316, 1602.03844, 1808.03619), and I believe [28] should point to the Zevin et. al paper in [29].
		> Done: Bayeswave (as well as a study of its accuracy) is cited in sec.IV, page 10, together with 1907.06540 that assess its accuracy; papers on glitches are cited on sec. V.B, page 14

- "for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA,” Please again add the citations for the LIGO and Virgo detector papers, and also add a citation to the KAGRA detector papers (1306.6747, 1111.7185).
		> Done: pag 14

- "hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD”: Please add citations to other papers from LVC members discussing marginalization over PSD detector noise (0804.3853, 1109.0442, 1909.01934, 1409.7215, 1307.8195, 2004.05149, 2006.05292, 2101.01200, 2004.07515, 1506.00185 and the two Bayeswave papers mentioned above).
		> The relevant part has been changed so as to acknowledge previous works. Probably here a _misunderstanding_ on what marginalization means?
			Everything is at page 15: "several approaches have been developed to try to study the validity of the assumption or even to drop it [46–52]"

- Please add a citation to https://arxiv.org/abs/gr-qc/0011041, which also demonstrates the applicability of autoregressive processes to gravitational-wave power spectral density estimation.
		> Done: on sec. V.B, page 13

- I’m not sure if this qualifies as a PnP comment, but a significant omission is the lack of discussion of the Bayeswave method of PSD estimation, which is the official method for calculating PSDs for parameter estimation results by the LVC. In addition to citing the papers described above, I think it would be appropriate to add a discussion, either in the Introduction or the Final Remarks comparing MESA to this “on-source” PSD estimation and citing the relvant papers disucssing the Bayeswave PSD method in comparison to the Welch method, some repeated from the above requests (1907.06540, 2006.05292).
		> This is a very good suggestion... Are we gonna include BayesLine among our comparison? It looks like a good idea. If we do so, probably we want to say something about it in the beginning
			The paper 1907.06540 is already cited together with BayesWave

## Personal comments

We fixed all the typos that the referee pointed out. We are thankful for that.

### Comparison with Welch method:
- Maybe it’s because of the broken citation, but I don’t understand which PSD is being used in the simulation. Is it the same “analytical PSD computed for the LIGO Hanford interferometer” used in the previous section, or is it a PSD calculated using the real LIGO data from around GW150914? If so, how was that PSD calculated?
		> The two PSDs are different. In the validation part, we used the design sensitivity of aLIGO (we quoted the relevant paper as well as the DCC link for the file). In the Welch comparison part, we employed the PSD released together with GW150914 (again we quoted the GWTC-1 paper and the link).

### Applications:

- It’s not really clear to me when you perform the forecasting for the temperature data why CAT is the obvious choice for the loss function, since in the previous section you described how CAT is unable to recover the true order of the autoregressive process. Does that mean that it would be worse at forecasting based on the AR(p)? There seems to be a contradiction where CAT does a better job for the temperature data but worse for the simulation presented in section II. It would be helpful to elaborate on this.
		> We added a claryfing footnote on this (note 10, page 13). We hope it is clear.

- "Indeed, a precise prediction of the strain time series can be beneficial in the detection of anomalies in the data and, eventually, their removal.”: How does this compare to the method presented in https://arxiv.org/abs/1908.05644, where the data containing a glitch are “inpainted” with values chosen such that the inverse-PSD-filtered data is 0 at these times?
		> Work in progress: we should understand what's in the paper...

### Final Remarks and Future Prospects

- In addition to adding the citations to other LVC works presenting methods and results for marginalizing over the PSD uncertainty (see PnP comments above), a more detailed discussion of how MESA compares to these other methods (particularly the error in the PSD measurement) would be appropriate here.
		> Working on it: shall we quote some numbers about the accuracy? Or we just reitarete the qualitative findings we had?

### 2021-05-26 paper/mesa.tex
First adds after referee report

@@
 \end{equation}
-without highlighting its connection with the time structure of the process itself, thus masking some important properties that will explore further in what follows.
+without highlighting its connection with the time structure of the process itself, thus masking some important properties that will be explored further in what follows.
 The latter definition in~\eqref{eq:psd-f-definition} gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.
 
-An ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
+An ubiquitous method for such computation is due to Welch \cite{Welch1967} and it is based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
 The PSD is obtained by slicing the observed realization $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
@@
 where $\rho$ is the empirical autocorrelation and $M$ is the maximum time lag at which the autocorrelation is computed.
-The sequence $W$ is a window function that can be chosen in several different way, each choice presenting advantages and disadvantages for the final estimate of the PSD.
+The sequence $W$ is a window function that can be chosen in several different ways, each choice presenting advantages and disadvantages for the final estimate of the PSD.
 
@@
 
-Another problem with this approach is the requirement for the window to be $0$ outside the interval in which the autocorrelation is computed.
+Another drawback of this approach is the requirement for the window to be $0$ outside the interval in which the autocorrelation is computed.
 We are arbitrarily assuming $\rho_j = 0$ for $j > M$ and modifying the estimate (i.e. the data) if a non-rectangular window is chosen.
@@
 In this paper, we discuss the details of the Maximum entropy principle, its application to the problem of PSD estimation with Burg's algorithm and the link between Burg's algorithm and autoregressive process.
-Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many problems posed by the Welch's algorithm (or other similar methods).
+Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many undesired aspects of the Welch's algorithm (or other similar methods).
 To facilitate this goal, we present and describe a new code, \texttt{memspectrum}, that provides a robust and easy-to-use python implementation of the algorithm\footnote{
@@
 
-Our paper is organized as follows: we being by briefly reviewing the theoretical foundations of the maximum entropy principle in Sec.~\ref{sec:foundations}.
+Our paper is organized as follows: we begin by briefly reviewing the theoretical foundations of the maximum entropy principle in Sec.~\ref{sec:foundations}.
 Sec.~\ref{sec:validation} presents the validation of Burg's method as well as of our implementation on simulated data.
@@
 
-After a short introduction to Jayne's MAXENT (sec.~\ref{sec:MAXENT}), we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form (sec.~\ref{sec:MESA}).
+After a short introduction to Jaynes' MAXENT (sec.~\ref{sec:MAXENT}), we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form (sec.~\ref{sec:MESA}).
 Next, we will discuss an interesting link between Burg's method and autoregressive processes (sec.~\ref{sec:autoregr}) and in sec.~\ref{sec:forecasting} we will use such link for straightforwardly forecasting a time series.
@@
 We can generalize the definition of information in the case where two different outcomes $E_1, E_2$, with given probabilities $P_1$ and $P_2$, are possible.
-To gain some intuition on the problem, we ask ourselves which are the probability assignments that makes the outcome more uncertain (i.e. maximize the information).
+To gain some intuition on the problem, we ask ourselves which are the probability assignments that make the outcome more uncertain (i.e. maximize the information).
 If $P_1$ and $P_2$ are largely different, for instance $P_1 = 0.999$ and $P_2 = 0.001$, we are allowed to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome.
@@
 Any generalization of eq.~\eqref{eq:information}, must then have its maximum when $ P_1 = P_2$.
-For $N$ events, system with the highest possible information content is when:
+For $N$ events, the system with the highest possible information content is when:
 \begin{equation}\nonumber
@@
 
-We now turn the core of our problem: how do we make a probability assignment for a set of events, that keeps into account our knowledge of the system and, at the same time, it is non committal towards unavailable knowledge?
+We now turn to the core of our problem: how do we make a probability assignment for a set of events, that keeps into account our knowledge of the system and, at the same time, it is non committal towards unavailable knowledge?
 The knowledge at our disposal about the system under study is what we call `evidence' and any probability assignment must agree with it.
@@
 }.
-In other words, the amount of `information' introduced by the the probability assignment should be as high as possible.
+In other words, the amount of `information' introduced by the probability assignment should be as high as possible.
 
 MAXENT puts the aforementioned principle into a more precise mathematical context by stating that probabilities should be assigned by maximizing uncertainty (information entropy) using evidence as constraint.
-This defines a variational problem, where the information entropy functional $H\left[p_1, \dots, p_N\right]$, defined in eq.~\eqref{eq:entropy}, has to maximized.
+This defines a variational problem, where the information entropy functional $H\left[p_1, \dots, p_N\right]$, defined in eq.~\eqref{eq:entropy}, has to be maximized.
 
@@
 Of particular importance is the case in which, in addition to the mean, also the variance is known: MAXENT leads to the Gaussian distribution. 
-We retain this particularly interesting from the foundational point of view, since it provides a deeper insight into the ubiquitous Gaussian distribution.
+This derivation is particularly interesting from the foundational point of view, since it provides a deeper insight into the ubiquitous Gaussian distribution.
 Indeed, it is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy.
@@
 The autocovariance matrix thus becomes a Toeplitz matrix\footnote{
-We remind the reader that a Toeplix matrix is a matrix in the form:
+We remind the reader that a Toeplitz matrix is a matrix in the form:
 \begin{pmatrix}
@@
 \begin{align}\label{eq:Whittle}
-p&\left((\tilde{x}_1, \ldots, \tilde{x}_n)\vert I\right) = \nonumber \\
+p&\left((\tilde{x}_1, \ldots, \tilde{x}_{n/2})\vert I\right) = \nonumber \\
     &\frac{1}{\left(2 \pi \det S\right)^{n / 2}}\exp\left(-\frac{1}{2}\sum_{ij}\tilde{x}_i S^{-1}_{ij} \tilde{x}_j \right)\,,
@@
 
-Before doing so, there is a technical problem to solve: the definition of entropy depends on a probability distribution, not on the PSD.
+Before doing so, there is a technicality to solve: the definition of entropy depends on a probability distribution, not on the PSD.
 It can be shown, e.g.~\cite{AblesMESA, Bartlett}, that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of the filtering a white noise process using a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
@@
 where $\Delta t$ is sampling rate and $Ny \equiv \frac{1}{2 \Delta t}$  is the Nyquist frequency.
-Thus maximising Eq.~\eqref{eq:EntropyGain} is equivalent to maximize Eq.~\eqref{eq:entropy_continuos}.
+Thus maximising eq.~\eqref{eq:EntropyGain} is equivalent to maximizing eq.~\eqref{eq:entropy_continuos}.
 
@@
 \begin{equation}\label{eq:MESApsd}
-    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^s\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}, 
+    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=1}^N a_s z^s\right)\left(\sum_{s = 1}^N a^*_s z^{-s}\right)}, 
 \end{equation}
-where $\Delta t$ is the sampling interval of the time series, $z=\exp{(2\pi i f\Delta t)}$, $a_0$ = 1.
-The vector obtained as $(1, a_1, \dots, a_N)$ is also known as the \textit{prediction error filter}.
-The coefficients $P_N$ and $a_s (s > 0)$ are to be determined by an iterative process (called Burg's algorithm).
+where $\Delta t$ is the sampling interval of the time series, $z=\exp{(2\pi i f\Delta t)}$, $a_1$ = 1.
+The vector obtained as $(1, a_2, \dots, a_N)$ is also known as the \textit{prediction error filter}.
+The coefficients $a_s (s > 0)$, together with an overall multiplicative scale factor $P_N$, are to be determined by an iterative process (called Burg's algorithm).
+The number $N$ of such coefficients is a choice that shall be made by the user and indeed it is the only hyperparameter that needs to be tuned.
 The details of the derivation and the actual form for the coefficients $a_s$ can be found in appendix \ref{sec:appendix}.
@@
 
-Thanks to Wold's theorem~\cite{Wold_theorem}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general; it turns out that the maximums entropy principle provides a representation of the time series as an $AR(p)$ process ad Burg's algorithm computes the autoregressive coefficients that are suitable to the available data.
+Thanks to Wold's theorem~\cite{Wold_theorem}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general; it turns out that the maximum entropy principle provides a representation of the time series as an $AR(p)$ process and Burg's algorithm computes the autoregressive coefficients that are suitable to the available data.
 
@@
 \begin{equation}
-    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
+    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{n = 1}^p b_n z^n \right)}
 \end{equation}
@@
     S_{AR(p)}(f) = \vert \tilde x(z)\vert ^ 2 = 
-    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}\,.
+    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_n e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}\,.
 \end{equation}
@@
 \end{align}
-The interpretation of Eq.~(\ref{eq:p_forecast}) is straightforward: $x_t$ follows a Gaussian distribution with a fixed variance an a mean value $m_t = \sum_{i = 1}^p b_i x_{t-i}$ computed from past observations.
+The interpretation of Eq.~(\ref{eq:p_forecast}) is straightforward: $x_t$ follows a Gaussian distribution with a fixed variance and a mean value $m_t = \sum_{i = 1}^p b_i x_{t-i}$ computed from past observations.
 Eq.~(\ref{eq:p_forecast}) provides then a well defined probability framework for predicting future observations: this is a very useful feature of MESA, that does not have an equivalent in any other spectral analysis computation methods.
@@
 on the one hand it is advisable to include as much knowledge of the autocorrelation as possible, leading to include all the known $\bar{r}_m$; on the other hand, including values of the autocorrelation that are not reliably estimated, can be counterproductive.
-The order $M$ of the autocorrelation to be considered (or, equivalently, the order $M$ of the underlying autoregressive process) is the only tuning parameter of MESA and a careful balance between this two necessities must be made when applying the algorithm.
+The order $M$ of the autocorrelation to be considered (or, equivalently, the order $M$ of the underlying autoregressive process) is the only tuning parameter of MESA and a careful balance between these two necessities must be made when applying the algorithm.
 
-The reminder of this section is devoted to an extensive study on how to make such choice.
+The remainder of this section is devoted to an extensive study on how to make such choice.
 In Section~\ref{sec:optimizers}, we are going to define three different \textit{loss functions} to measure how well the 
 algorithm is able to reproduce a known PSD.
-The basic idea is to validate, as the autoregressive order considered increase, the performance of the algorithm results 
+The basic idea is to validate, as the autoregressive order considered increases, the performance of the algorithm results 
 by measuring the loss function and pick, among the orders the one that yields better results.
@@
 We test the performance of the three loss functions on a random time series generated with a known power spectral density.
-In this first experiment, we take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$.
+In this first experiment, we take the PSD to be a Gaussian distribution with mean $\mu = 2.5$ and standard deviation $\sigma = 0.5$, where the units are arbitrary.
 The time series are generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$ and performing an inverse Fast Fourier Transform on the frequency series.
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/FPE_spectrum_estim.pdf}
-	\caption{Average spectrum for $\mathcal{L}_{\rm FPE}(m)$ with 90\% confidence regions (top) and ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error} (bottom).
+	\caption{In the top panel, we show the average spectrum for $\mathcal{L}_{\rm FPE}(m)$ with 90\% confidence regions (purple shaded region), together with the median estimation (red line) and the target PSD (black dashed line).
+In the bottom panel, we display the ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error}.
 	The average is computed with $1000$ realization of a $3000$ points long time series.}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/OBD_spectrum_estim.pdf}
-	\caption{Average spectrum for $\mathcal{L}_{\rm OBD}(m)$ loss function with 90\% confidence regions (top) and ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error} (bottom).
+	\caption{In the top panel, we show the average spectrum for $\mathcal{L}_{\rm OBD}(m)$ with 90\% confidence regions (purple shaded region), together with the median estimation (red line) and the target PSD (black dashed line).
+In the bottom panel, we display the ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error}.
 The average is computed with $1000$ realization of a $3000$ points long time series.}
@@
 	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/CAT_spectrum_estim.pdf}
-	\caption{Average spectrum for $\mathcal{L}_{\rm CAT}(m)$ loss function with 90\% confidence regions (top) and ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error} (bottom).
+	\caption{In the top panel, we show the average spectrum for $\mathcal{L}_{\rm CAT}(m)$ with 90\% confidence regions (purple shaded region), together with the median estimation (red line) and the target PSD (black dashed line).
+In the bottom panel, we display the ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error}.
 The average is computed with $1000$ realization of a $3000$ points long time series.}
@@
 FPE are clustered in a very small region. This is also a desirable property: FPE, in fact, provides a stable 
-estimation of the ar order, which does not affect much the reconstruction error.
+estimation of the AR order, which does not affect much the reconstruction error.
 We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length $m$ is clustered in a region where there is no dependence of error on $m$.
@@
 Fig.~\ref{fig:optcomparison} also shows a strong dependence of the error on the estimated length of the filter.
-The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like outliers in different realizations of the time-series, but this is also responsible of an increased variance in the estimate, by introducing spurious peaks in the reproduction.
+The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like outliers in different realizations of the time-series, but this is also responsible for an increased variance in the estimate, by introducing spurious peaks in the reproduction.
 
@@
 FPE guarantees that the single estimation is more faithful.
-As in any common situation we cannot perform such averaging over different realization of the same time series, we must prefer FPE over CAT (let alone OBD, which even though qualitatively similar to FPE has worse performance).
+As in any common situation we cannot perform such averaging over different realizations of the same time series, we must prefer FPE over CAT (let alone OBD, which even though qualitatively similar to FPE has worse performance).
 However, if we indeed can measure the PSD by averaging over different time series, using CAT as a loss function is the best choice.
@@
 We continue our characterization of the various loss functions considered in this work, by investigating the 
-reconstruction of a specific, known, power spectrum: that is the Advanced LIGO design sensitivity theoretical spectral curve\cite{LIGO_PSDs}.
+reconstruction of a specific, known, power spectrum: that is the Advanced LIGO \cite{LIGO_2015, Virgo_2014, KAGRA_2012, KAGRA_2013} design sensitivity theoretical spectral curve \cite{LVK_future_2013, LIGO_PSDs}.
 For this analysis, we generate $N = 500$ time series of $40960$ points each, sampled with a sampling rate of $\SI{2048}{Hz}$, hence we fix the duration
@@
      \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/compare_estimates.pdf}
-      \caption{Average spectrum for the three different loss function as compared with the ``true" PSD. The average is computed over $500$ realization of a $40960$ points long time series.}
+      \caption{Average spectrum for the three different loss functions as compared with the ``true" PSD. The average is computed over $500$ realization of a $40960$ points long time series.}
        \label{fig:ligospectrum}
@@
 In fact, even in presence of very sharp spectral features, CAT reconstruction seems to be almost perfectly coincident with each of them.
-Hence, also the study of simulated LIGO data seems to indicate that whenever and wherever ensemble-averaged PSD are necessary, CAT is the optimal choice of loss function.
+Hence, also the study of simulated LIGO data seems to indicate that whenever and wherever ensemble-averaged PSDs are necessary, CAT is the optimal choice of loss function.
 However, on a single time-series realization, FPE is the more robust choice.
@@
 \section{Comparison with Welch method}\label{sec:Welch_comparison}
+\sschmidt{Do we want to make a comparison with BayesLine? I think so... \\ Do we want to introduce this method in the introduction? Makes sense}
 In this section we perform a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
-Although similar studies can be drawn from any other PSD, for definiteness, we focus on the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog \cite{GWTC1}.
+Although similar studies can be drawn from any other PSD, for definiteness, we focus on the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog~\cite{GWTC1, PSD_release}, and computed with BayesLine package \cite{Cornish_2015, Littenberg_2015, Chatziioannou_2019}.
 
-We simulate data\footnote{This is to ensure that we have a baseline PSD to compare the data with} from the PSD around the event GW150914 and we employ both Welch's method and MESA to estimate the spectrum.
+We simulate data\footnote{This is to ensure that we have a baseline PSD to compare the data with} from the PSD used for the analysis of the event GW150914 and we employ both Welch's method and MESA to estimate the spectrum.
 We vary the length of the data used for the estimation: this is also useful to assess how the computation depends on the data available.
@@
 Finally, as already discussed Welch's method is very dependent on the choice of window function. 
-Our choice, a Tukey window with aforementioned parameters, is the best compromise with found between noise 
+A Tukey window with aforementioned parameters is what we found to be the best compromise between noise 
 and accuracy for the reconstruction, but different choices can be made, possibly providing more accurate results than the ones reported here. 
@@
 peak at the frequency $f_D = \SI{1}{day^{-1}}$ corresponding to one day, corresponding to the day-night cycle: this is expected.
-Higher order harmonics (corresponding to integer multiples of $f_D$) are also visible up to the Nyquist frequency ($f_D = \SI{0.5}{hour^{-1}}$): they corresponds to signals that preserve the 1 day period while, at the same time, capturing the complex variability of the daily temperature cycle throughout the year.
+Higher order harmonics (corresponding to integer multiples of $f_D$) are also visible up to the Nyquist frequency ($f_D = \SI{0.5}{hour^{-1}}$): they correspond to signals that preserve the 1 day period while, at the same time, capturing the complex variability of the daily temperature cycle throughout the year.
 
@@
 On the other hand, the peak is captured well by the CAT loss function.
-In analogy as what observed for the peak at $f_D$, in the CAT spectrum we observe a higher order harmonics at a frequency $2 \cdot f_{yr}$.
+In analogy as what observed for the peak at $f_D$, in the CAT spectrum we observe a higher order harmonic at a frequency $2 \cdot f_{yr}$.
 As in the previous case, this is required to better model the temperature variation in the year.
@@
 This is compared with the actual measured temperature values.
-The predictions spans a two years range of time.
+The predictions span a two years range of time.
 We report our results in Fig.~\ref{fig:temp_spectrum}.
@@
 As a last example of an application of MESA, we return to the study of the strain produced by the LIGO-Virgo interferometers 
-and we forecast the future observations. We focus on the public data released by the LIGO/Virgo collaboration~\cite{Abbott_2021}.
-We reconstruct the PSD both assuming the CAT and FPE loss functions on $\SI{1000}{s}$ of data \cite{Abbott_2021} from the Livingston observatory
+and we forecast the future observations.
+A seminal work on the application of autoregressive models to Virgo data is presented in \cite{Cuoco_2001}, where an $AR(p)$ model is trained on the data for the purpose of estimating the PSD and to create a whithener filter.
+\sschmidt{Is this the right place to insert the citation??}
+
+We focus on the public data released by the LIGO/Virgo collaboration~\cite{Abbott_2021}.
+We reconstruct the PSD both assuming the CAT and FPE loss functions on $\SI{1000}{s}$ of data from the Livingston observatory
 starting from GPS time $1164603392$. The data are sampled at $\SI{4096}{Hz}$.
-
-We then forecast the following $\SI{100}{s}$ of observations with the model optimized with CAT. The results are shown in fig.~\ref{fig:LIGO_forecast}.
+We then forecast the following $\SI{100}{s}$ of observations with the model optimized with CAT\footnote{
+The reason for this choice is that CAT, despite showing higher variance and a worse PSD estimation accuracy, estimates sistematically high autoregressive orders.
+We believe that this feature provides a more realiable forecast, as more points are used for predictions.
+In practice, we found that CAT and FPE behave very similarly.
+}. The results are shown in fig.~\ref{fig:LIGO_forecast}.
 The prediction is always in the $90\%$ confidence interval. The confidence interval, however, is very broad and the prediction is not very accurate (being the same order of magnitude of the strain).
@@
 
+\sschmidt{How does this compare to the method presented in https://arxiv.org/abs/1908.05644? Seems like an interesting paper that tells you how to produce data after glitch removal.}
 Despite poor predictions on long timescales, the AR process obtained with MESA still can be useful on short timescales.
 Indeed, a precise prediction of the strain time series can be beneficial in the detection of anomalies in the data and, eventually, their removal.
-The predictions can form an expected baseline for the strain; any anomaly (i.e.
-a glitch or even a transient of physical origin) can show up as a large departure from expected trend.
+Indeed loud anomalies in the data, called {\it glitches}, pose a major challenge to the ability of detecting signals and intensive work has been done to mitigate the disruption in sensitivity they cause \cite{Nuttall_2015, Noise_char_2016, Zevin_2017} and to develop effective subractions techniques \cite{Pankow_2018}.
+The predictions can form an expected baseline for the strain; any anomaly (i.e. a glitch or even a transient of physical origin) can show up as a large departure from expected trend.
 Moreover, if a glitch is detected, its shape can be estimated (as well as the confidence level) by subtracting the expected signal with the actual signal.
-This can work: a typical glitch~\cite{Zevin_2017} can last as long as $\SI{0.2}{s}$, close to the time scale over which the forecasting is reliable.
+This can work: a typical glitch can last as long as $\SI{0.2}{s}$, close to the time scale over which the forecasting is reliable.
 
@@
 The use of MESA is particularly useful for short time series samples, where Welch's method is outperformed in both precision and confidence.
+\sschmidt{Add here something on accuracy on the PSD estimation, as the referee requires}
 
-This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
-Whenever the log-likelihood of the model for a time series takes the form
+This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA \cite{LIGO_2015, Virgo_2014, KAGRA_2012, KAGRA_2013}, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
+\sschmidt{There might be a misunderstanding: I guess the referee is talking about taking into account the PSD errors, while we are talking about dropping the stationariety assumption...
+The misleading word is `marginalize', I think}
+This is possible whenever the log-likelihood of the model for a time series takes the form
 \begin{equation}
@@
 where $\tilde{\hphantom{a}}$ denotes the Fourier transform and the signal model $x(t;\theta)$, dependent on some parameter $\theta$, is a prediction for a deterministic signal buried in the observed time series $d(t)$.
-The typical approach is to estimate the PSD offline on a large batch of data and off-source (see~\cite{lalinference}), that might not reflect the structure of noise in the time slice to analyse.
-MESA can offer a way out from this problem: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on $d_t-x_t$.
+In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source \cite{lalinference}: this scheme assumes stationary data and it might not reflect the structure of noise in the time slice to analyse.
+Several approaches have been developed to try to study the validity of the assumption or even to drop it \cite{R_ver_2010, R_ver_2011, Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave, Edwards_2020, Edwards_2015} \sschmidt{All these citations seems a bit too much for me...}.
+Besides them, MESA can offer an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on $d_t-x_t$.
 In this way, the PSD would depend also on $\theta$ and would effectively model the residuals. Preliminary studies suggest that this is possible~\cite{martini_thesis}.
@@
 Whenever the observed data are outside the expectations, an anomaly detection can be claimed.
-Of course such predictions can be done with a more accurate (perhaps nonlinear model); however MESA has the advantage of being simple and fast to construct, while providing decent predictions.
+Of course such predictions can be done with a more accurate (perhaps nonlinear) model; however MESA has the advantage of being simple and fast to construct, while providing decent predictions.
 At the same time, several instruments present gaps in their data stream, for instance LISA is expected to show such gaps~(e.g.

### 2021-05-26 paper/referee_response.md
Edit referee response

@@
 - Please add citations to the LIGO and Virgo instrument papers the first time these detectors are mentioned (1411.4547, 1408.3978).
-		> Done: page 9
+	+ Done: page 9
 
 - Please fix the broken citation at the beginning of section IV to "the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog”
-		> Done, added also the link to DCC entry where the PSD are taken from
+	+ Done, added also the link to DCC entry where the PSD are taken from
 
@@
 The data are sampled at 4096 Hz.”
-		> Done
+	+ Done
 
@@
 al paper in [29].
-		> Done: Bayeswave (as well as a study of its accuracy) is cited in sec.IV, page 10, together with 1907.06540 that assess its accuracy; papers on glitches are cited on sec. V.B, page 14
+	+ Done: Bayeswave (as well as a study of its accuracy) is cited in sec.IV, page 10, together with 1907.06540 that assess its accuracy; papers on glitches are cited on sec. V.B, page 14
 
 - "for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA,” Please again add the citations for the LIGO and Virgo detector papers, and also add a citation to the KAGRA detector papers (1306.6747, 1111.7185).
-		> Done: pag 14
+	+ Done: pag 14
 
 - "hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD”: Please add citations to other papers from LVC members discussing marginalization over PSD detector noise (0804.3853, 1109.0442, 1909.01934, 1409.7215, 1307.8195, 2004.05149, 2006.05292, 2101.01200, 2004.07515, 1506.00185 and the two Bayeswave papers mentioned above).
-		> The relevant part has been changed so as to acknowledge previous works. Probably here a _misunderstanding_ on what marginalization means?
+	+ The relevant part has been changed so as to acknowledge previous works. Probably here a _misunderstanding_ on what marginalization means?
 		Everything is at page 15: "several approaches have been developed to try to study the validity of the assumption or even to drop it [46–52]"
@@
 - Please add a citation to https://arxiv.org/abs/gr-qc/0011041, which also demonstrates the applicability of autoregressive processes to gravitational-wave power spectral density estimation.
-		> Done: on sec. V.B, page 13
+	+ Done: on sec. V.B, page 13
 
@@
 In addition to citing the papers described above, I think it would be appropriate to add a discussion, either in the Introduction or the Final Remarks comparing MESA to this “on-source” PSD estimation and citing the relvant papers disucssing the Bayeswave PSD method in comparison to the Welch method, some repeated from the above requests (1907.06540, 2006.05292).
-		> This is a very good suggestion... Are we gonna include BayesLine among our comparison? It looks like a good idea. If we do so, probably we want to say something about it in the beginning
-			The paper 1907.06540 is already cited together with BayesWave
+	+ This is a very good suggestion...
+Are we gonna include BayesLine among our comparison?
+It looks like a good idea.
+If we do so, probably we want to say something about it in the beginning.The paper 1907.06540 is already cited together with BayesWave
 
@@
 If so, how was that PSD calculated?
-		> The two PSDs are different.
+	+ The two PSDs are different.
 In the validation part, we used the design sensitivity of aLIGO (we quoted the relevant paper as well as the DCC link for the file).
@@
 It would be helpful to elaborate on this.
-		> We added a claryfing footnote on this (note 10, page 13). We hope it is clear.
+	+ We added a claryfing footnote on this (note 10, page 13). We hope it is clear.
 
 - "Indeed, a precise prediction of the strain time series can be beneficial in the detection of anomalies in the data and, eventually, their removal.”: How does this compare to the method presented in https://arxiv.org/abs/1908.05644, where the data containing a glitch are “inpainted” with values chosen such that the inverse-PSD-filtered data is 0 at these times?
-		> Work in progress: we should understand what's in the paper...
+	+ Work in progress: we should understand what's in the paper...
 
@@
 - In addition to adding the citations to other LVC works presenting methods and results for marginalizing over the PSD uncertainty (see PnP comments above), a more detailed discussion of how MESA compares to these other methods (particularly the error in the PSD measurement) would be appropriate here.
-		> Working on it: shall we quote some numbers about the accuracy? Or we just reitarete the qualitative findings we had?
+	+ Working on it: shall we quote some numbers about the accuracy? Or we just reitarete the qualitative findings we had?

### 2021-05-26 paper/referee_response.md
Edit referee response

@@
 In addition to citing the papers described above, I think it would be appropriate to add a discussion, either in the Introduction or the Final Remarks comparing MESA to this “on-source” PSD estimation and citing the relvant papers disucssing the Bayeswave PSD method in comparison to the Welch method, some repeated from the above requests (1907.06540, 2006.05292).
-	+ This is a very good suggestion...
+	+ Work in progress: This is a very good suggestion...
 Are we gonna include BayesLine among our comparison?
@@
 We fixed all the typos that the referee pointed out. We are thankful for that.
+
+### Validations of the model:
+- I wasn't able to get a good grasp of the practical applications of the estimation of the autoregressive process order while reading this section.
+Is gravitational-wave data expected to be stationary enough to reliably model as an autoregressive process?
+If not, what are some other examples?
+It would be very helpful to elaborate on this, perhaps when the connection to AR is introduced.
+	+ Work in progress: I guess we could just point out that it is just a way to validate the accuracy of the method and to test which loss function is better. Does it makes sense?
 
@@
 - In addition to adding the citations to other LVC works presenting methods and results for marginalizing over the PSD uncertainty (see PnP comments above), a more detailed discussion of how MESA compares to these other methods (particularly the error in the PSD measurement) would be appropriate here.
-	+ Working on it: shall we quote some numbers about the accuracy? Or we just reitarete the qualitative findings we had?
+	+ Work in progress: shall we quote some numbers about the accuracy? Or we just reitarete the qualitative findings we had?

### 2021-05-27 paper/mesa.tex
reformatted to A&A style

@@
-\documentclass[twocolumn,showpacs,preprintnumbers,nofootinbib,prd,
-superscriptaddress,10pt]{revtex4-1}
-
+\documentclass{aa}
+\bibpunct{(}{)}{;}{a}{}{,}
+\usepackage[varg]{txfonts}
 \usepackage{amsmath,amssymb}
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT
-\begin{abstract}
+
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE
+	\title{Maximum Entropy Spectral Analysis: a case study}
+	\author{Alessandro \surname{Martini}\inst{\ref{inst1}}\thanks{martini.alessandr@gmail.com} \and 
+	Stefano \surname{Schmidt}\inst{\ref{inst2}}\thanks{s.schmidt@uu.nl} \and Walter \surname{Del Pozzo}\inst{\ref{inst1}}\thanks{walter.delpozzo@unipi.it}}
+
+        \institute{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy\label{inst1} \and Institute for Gravitational and Subatomic Physics (GRASP),Utrecht University, Princetonplein 1, 3584 CC, Utrecht, The Netherlands\label{inst2}}
+
+  \abstract{
 The Maximum Entropy Spectral Analysis (MESA) method, developed by Burg, provides a powerful tool to perform spectral estimation of a time-series.
@@
 We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
-systematically lower variance and bias. This is particularly manifest in the case of a little number of data points, making Burg's method most suitable to work in this regime.
+systematically lower variance and bias. This is particularly manifest in the case of a little number of data points, making Burg's method most suitable to work in this regime.}
 
-\end{abstract}
-
-	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE
-	\title{Maximum Entropy Spectral Analysis: a case study}
-	\author{Alessandro \surname{Martini}}
-		\email{martini.alessandr@gmail.com}
-        \affiliation{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}        
-	\author{Stefano \surname{Schmidt}}
-		\email{s.schmidt@uu.nl}
-        \affiliation{Institute for Gravitational and Subatomic Physics (GRASP), Utrecht University, Princetonplein 1, 3584 CC, Utrecht, The Netherlands}      
-	\author{Walter \surname{Del Pozzo}}
-		\email{walter.delpozzo@unipi.it}
-        \affiliation{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy}
+\titlerunning{MESA: a case study}
+\authorrunning{Martini, Schmidt, Del Pozzo}
 
@@
 
-An ubiquitous method for such computation is due to Welch \cite{Welch1967} and it is based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
+An ubiquitous method for such computation is due to \citet{Welch1967} and it is based on Eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
 The PSD is obtained by slicing the observed realization $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
-This approach is equivalent~\cite{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
+This approach is equivalent~\citep{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
 \begin{equation}
@@
 
-An appealing alternative approach, based on the Maximum Entropy principle \cite{JaynesArticle,jaynes2003ptl, Jaynes_MAXENT}, has been put forward by Burg \cite{burg1975maximum}.
+An appealing alternative approach, based on the Maximum Entropy principle \citep{JaynesArticle,jaynes2003ptl, Jaynes_MAXENT}, has been put forward by~\citet{burg1975maximum}.
 Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the algorithm parameters, since it provides an iterative closed form expression for the spectrum of a stochastic stationary time series.
@@
 Its domain of application turned out to be wider than expected.
-In fact, thanks to Burg \cite{burg1975maximum}, this method has also been applied to perform high quality computation of power spectral densities of time series.
+In fact, thanks to~\citet{burg1975maximum}, this method has also been applied to perform high quality computation of power spectral densities of time series.
 
@@
 Even less information is required if we are studying a system whose outcome is certain (has probability $p = 1$), as in this case, a communication is not even needed.
-Shannon \cite{Shannon} proposed the quantity
+\citet{Shannon} proposed the quantity
 \begin{equation}\label{eq:information}
@@
 
-Shannon \cite{Shannon} showed that the only functional form satisfying continuity with respect to its parameters, additivity and that has a maximum for equal probability events is:
+\citet{Shannon} showed that the only functional form satisfying continuity with respect to its parameters, additivity and that has a maximum for equal probability events is:
 \begin{equation}\label{eq:entropy}
@@
 It is very common that the constraints provided by the evidence are not enough for setting the probabilities for each event: in this case, it is reasonable to assume that the probability assignment should make the experiment as unpredictable as possible\footnote{
-In ref.~\cite{Jaynes_MAXENT} this statement is made more precise and justified more thoroughly, with arguments based on combinatorial analysis.
+In~\citet{Jaynes_MAXENT} this statement is made more precise and justified more thoroughly, with arguments based on combinatorial analysis.
 }.
@@
 For this work, we are particularly interested in the multi-dimensional case.
-Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realization of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of autocovariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution~\cite{gregory_2005}:
+Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realization of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of autocovariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution~\citep{gregory_2005}:
 \begin{align}
@@
 }.
-Toeplitz matrices are asymptotically equivalent to circulant matrices and thus diagonalized by the discrete Fourier transform base~\cite{Gray}.
+Toeplitz matrices are asymptotically equivalent to circulant matrices and thus diagonalized by the discrete Fourier transform base~\citep{Gray}.
 %R. M. Gray, Foundations and Trends⃝R in Communications and Information Theory 2, 155 (2005), URL https://doi.org/10.1561/0100000006.
@@
 where the matrix $S_{ij} = S_i \delta_{ij}$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
-Many readers will recognize the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\cite{prob_information_theory}
-%P. M. Woodward, D. W. Fry, and W. Higinbotham, Probability and information theory, with applications to radar; 2nd ed. (Pergamon Press, Oxford, 1964), URL http://cds.cern.ch/record/2031792.
-and of gravitational waves data analysis, e.g. \cite{Finn_1992, FINDCHIRP}.
+Many readers will recognize the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\citep{prob_information_theory}
+and of gravitational waves data analysis, \citep[e.g.]{Finn_1992, FINDCHIRP}.
 Thanks to MAXENT, the problem of defining the probability distribution describing a wide-sense stationary process is thus 
@@
 
-As in the previous examples, one needs to set up a variational problem where the entropy eq.~\eqref{eq:entropy_continuos} is maximized 
+As in the previous examples, one needs to set up a variational problem where the entropy, Eq.~\eqref{eq:entropy_continuos}, is maximized 
 subject to some problem-specific constraints. 
@@
 Before doing so, there is a technicality to solve: the definition of entropy depends on a probability distribution, not on the PSD.
-It can be shown, e.g.~\cite{AblesMESA, Bartlett}, that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of the filtering a white noise process using a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
+It can be shown, ~\citep[e.g.]{AblesMESA, Bartlett}, that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of the filtering a white noise process using a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
 A filter with transfer function $T(f)$ takes in input a time series $x_t$ and outputs a times series $y_t$ such that:
@@
 where $\Delta t$ is sampling rate and $Ny \equiv \frac{1}{2 \Delta t}$  is the Nyquist frequency.
-Thus maximising eq.~\eqref{eq:EntropyGain} is equivalent to maximizing eq.~\eqref{eq:entropy_continuos}.
+Thus maximising Eq.~\eqref{eq:EntropyGain} is equivalent to maximizing eq.~\eqref{eq:entropy_continuos}.
 
@@
 
-Thus, by maximizing Eq.~\eqref{eq:EntropyGain} with constraints in eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
+Thus, by maximizing Eq.~\eqref{eq:EntropyGain} with constraints in Eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
 This approach on PSD computation provides a result consistent with the empirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
@@
 Most importantly, the variational problem admits a closed-form analytical expression for $S(f)$.
-The expression was first found by Burg~\cite{burg1975maximum}:
+The expression was first found by ~\citet{burg1975maximum}:
 \begin{equation}\label{eq:MESApsd}
@@
 The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the study
-of autoregressive processes (AR)~\cite{doi:10.1029/RG013i001p00183}.
+of autoregressive processes (AR)~\citep{doi:10.1029/RG013i001p00183}.
 An autoregressive stationary process of order $p$, AR($p$), is a time series whose values satisfy the following expression: 
@@
 
-Thanks to Wold's theorem~\cite{Wold_theorem}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general; it turns out that the maximum entropy principle provides a representation of the time series as an $AR(p)$ process and Burg's algorithm computes the autoregressive coefficients that are suitable to the available data.
+Thanks to Wold's theorem~\citep{Wold_theorem}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general; it turns out that the maximum entropy principle provides a representation of the time series as an $AR(p)$ process and Burg's algorithm computes the autoregressive coefficients that are suitable to the available data.
 
@@
 
-Guided from numerical experiments, an indication on the upper bound to the autoregressive order $M_{max}$ is~\cite{doi:10.1190/1.1440902}:
+Guided from numerical experiments, an indication on the upper bound to the autoregressive order $M_{max}$ is~\citep{doi:10.1190/1.1440902}:
 \begin{equation}\label{eq:MMAx}
@@
 \item \textbf{Final prediction Error} 
-The first criterion is due to Akaike~\cite{Akaike1970StatisticalPI}. It was proposed that $m$ should be chosen as the 
+The first criterion is due to ~\citet{Akaike1970StatisticalPI}. It was proposed that $m$ should be chosen as the 
 length that minimizes the error when the filter is used as a predictor, the \emph{final prediction} error (FPE): 
@@
 \item \textbf{Criterion Autoregressive Transfer function (CAT)}
-This second loss function has been proposed by Parzen and studied in detail by Banshali~\cite{bhansali1986}. 
+This second loss function has been proposed by Parzen and studied in detail by~\citet{bhansali1986}. 
 It is based on the assumption that the observed process is an infinite order autoregressive process 
@@
 \item \textbf{Optimum Bayes Decision rule} 
-The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
+The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\citep{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
 time series for the process that has to be described as an $AR(m)$ process, with $m$ to be determined.
@@
 hypothesis is uniquely determined from both the length and the values of the filter under study.
-Choosing a Likelihood of the form eq.~\eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypotheses, Rao has shown that choosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
+Choosing a Likelihood of the form Eq.~\eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypotheses, Rao has shown that choosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
 maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of:
@@
 
-Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson recursion~\cite{doi:10.1002/sapm1946251261} until $M_{max}$, as given in Eq.~\eqref{eq:MMAx}, iterations are reached.
+Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson recursion~\citep{doi:10.1002/sapm1946251261} until $M_{max}$, as given in Eq.~\eqref{eq:MMAx}, iterations are reached.
 Then, the order $m$ is selected to be the one that minimizes the specified loss function.
@@
 We continue our characterization of the various loss functions considered in this work, by investigating the 
-reconstruction of a specific, known, power spectrum: that is the Advanced LIGO \cite{LIGO_2015, Virgo_2014, KAGRA_2012, KAGRA_2013} design sensitivity theoretical spectral curve \cite{LVK_future_2013, LIGO_PSDs}.
+reconstruction of a specific, known, power spectrum: that is the Advanced LIGO \citep{LIGO_2015, Virgo_2014, KAGRA_2012, KAGRA_2013} design sensitivity theoretical spectral curve \citep{LVK_future_2013, LIGO_PSDs}.
 For this analysis, we generate $N = 500$ time series of $40960$ points each, sampled with a sampling rate of $\SI{2048}{Hz}$, hence we fix the duration
@@
 In this section we perform a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
-Although similar studies can be drawn from any other PSD, for definiteness, we focus on the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog~\cite{GWTC1, PSD_release}, and computed with BayesLine package \cite{Cornish_2015, Littenberg_2015, Chatziioannou_2019}.
+Although similar studies can be drawn from any other PSD, for definiteness, we focus on the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog~\citep{GWTC1, PSD_release}, and computed with BayesLine package \citep{Cornish_2015, Littenberg_2015, Chatziioannou_2019}.
 
@@
 In all cases, the sampling rate is set to $\SI{4096}{Hz}$.
-For the Welch algorithm, we use the standard implementation provided by the python library \texttt{scipy}~\cite{numpy,scipy}.
+For the Welch algorithm, we use the standard implementation provided by the python library \texttt{scipy}~\citep{numpy,scipy}.
 The results from both methods are summarized in  Figs.~\ref{fig:MESA_LIGO_data} and~\ref{fig:welch_LIGO_data} respectively.
@@
 known; ii) it provides a stress test for the time series forecast analysis.  As dataset, we  used the historical reanalysis data 
-from the ``ERA5-Land hourly data from 1981 to present" dataset, downloaded from the Climate Data Store~\cite{ERA5_dataset}.
+from the ``ERA5-Land hourly data from 1981 to present" dataset, downloaded from the Climate Data Store~\citep{ERA5_dataset}.
 The data consist in temperatures taken at coordinates $\textrm{N}\;\ang{45;5;} \;\; \textrm{E}\;\ang{9;1;}}$, corresponding about to the city of Milan, Italy.
@@
 and we forecast the future observations.
-A seminal work on the application of autoregressive models to Virgo data is presented in \cite{Cuoco_2001}, where an $AR(p)$ model is trained on the data for the purpose of estimating the PSD and to create a whithener filter.
-\sschmidt{Is this the right place to insert the citation??}
+A seminal work on the application of autoregressive models to Virgo data is presented in \citet{Cuoco_2001}, where an $AR(p)$ model is trained on the data for the purpose of estimating the PSD and to create a whitener filter.
 
-We focus on the public data released by the LIGO/Virgo collaboration~\cite{Abbott_2021}.
+We focus on the public data released by the LIGO/Virgo collaboration~\citep{Abbott_2021}.
 We reconstruct the PSD both assuming the CAT and FPE loss functions on $\SI{1000}{s}$ of data from the Livingston observatory
@@
 The reason for this choice is that CAT, despite showing higher variance and a worse PSD estimation accuracy, estimates sistematically high autoregressive orders.
-We believe that this feature provides a more realiable forecast, as more points are used for predictions.
-In practice, we found that CAT and FPE behave very similarly.
-}. The results are shown in fig.~\ref{fig:LIGO_forecast}.
+We believe that this feature provides a more reliable forecast, as more points are used for predictions.
+In practice, we found that CAT and FPE behave very similarly.}.
+The results are shown in fig.~\ref{fig:LIGO_forecast}.
 The prediction is always in the $90\%$ confidence interval. The confidence interval, however, is very broad and the prediction is not very accurate (being the same order of magnitude of the strain).
@@
 Indeed, a precise prediction of the strain time series can be beneficial in the detection of anomalies in the data and, eventually, their removal.
-Indeed loud anomalies in the data, called {\it glitches}, pose a major challenge to the ability of detecting signals and intensive work has been done to mitigate the disruption in sensitivity they cause \cite{Nuttall_2015, Noise_char_2016, Zevin_2017} and to develop effective subractions techniques \cite{Pankow_2018}.
+Indeed loud anomalies in the data, called {\it glitches}, pose a major challenge to the ability of detecting signals and intensive work has been done to mitigate the disruption in sensitivity they cause \citep{Nuttall_2015, Noise_char_2016, Zevin_2017} and to develop effective subraction techniques \citep{Pankow_2018}.
 The predictions can form an expected baseline for the strain; any anomaly (i.e. a glitch or even a transient of physical origin) can show up as a large departure from expected trend.
@@
 
-As noted above, for pushing further the prediction performance from our simple linear predictor, more sophisticated forecasting methods are available in the Machine Learning literature (see e.g.
-\cite{LSTM_article, oord2016wavenet}) and they can be trained for precision forecasting.
+As noted above, for pushing further the prediction performance from our simple linear predictor, more sophisticated forecasting methods are available in the Machine Learning literature \citep[see]{LSTM_article, oord2016wavenet} and they can be trained for precision forecasting.
 However, a Maximum-Entropy trained AR process can be a simple baseline for comparing such more advanced methods and might suffice for many purposes.
@@
 We presented a case study of the application of Maximum Entropy principle to the realm of spectral estimation.
-Albeit the methodology hereby presented is grounded on solid theoretical foundations and its merits are widely recognized, Maximum Entropy methods have yet to be adopted routinely in the study of problems related to time series.
+Albeit the methodology hereby presented is grounded on solid theoretical foundations and its merits are widely recognised, Maximum Entropy methods have yet to be adopted routinely in the study of problems related to time series.
 The superior nature of maximum entropy methods, and in particular of Burg's method, is exemplified by the closed form estimate of the power spectral density and by the theoretical bridge between spectral analysis and AR processes.
@@
 
-This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA \cite{LIGO_2015, Virgo_2014, KAGRA_2012, KAGRA_2013}, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
+This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA \citep{LIGO_2015, Virgo_2014, KAGRA_2012, KAGRA_2013}, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalising over the detector noise PSD.
 \sschmidt{There might be a misunderstanding: I guess the referee is talking about taking into account the PSD errors, while we are talking about dropping the stationariety assumption...
@@
 In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source \cite{lalinference}: this scheme assumes stationary data and it might not reflect the structure of noise in the time slice to analyse.
-Several approaches have been developed to try to study the validity of the assumption or even to drop it \cite{R_ver_2010, R_ver_2011, Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave, Edwards_2020, Edwards_2015} \sschmidt{All these citations seems a bit too much for me...}.
+Several approaches have been developed to try to study the validity of the assumption or even to drop it \citep{R_ver_2010, R_ver_2011, Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave, Edwards_2020, Edwards_2015} \sschmidt{All these citations seems a bit too much for me...}.
 Besides them, MESA can offer an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on $d_t-x_t$.
-In this way, the PSD would depend also on $\theta$ and would effectively model the residuals. Preliminary studies suggest that this is possible~\cite{martini_thesis}.
-
+In this way, the PSD would depend also on $\theta$ and would effectively model the residuals.
+Preliminary studies suggest that this is possible~\citep{martini_thesis}.
 Furthermore, MESA provides a simple, but robust and quite accurate, albeit for short times, predictor for the time series.
@@
 At the same time, several instruments present gaps in their data stream, for instance LISA is expected to show such gaps~(e.g.
-\cite{lisa_gaps} and references therein), MESA forecasting capabilities could be used to fill those gaps with predicted data from past observations.
-
+\citet{lisa_gaps} and references therein), MESA forecasting capabilities could be used to fill those gaps with predicted data from past observations.
 In conclusion, we reiterate that MESA is a theoretically sound, computationally feasible and reliable way of studying the properties of stochastic processes and we hope that the investigations presented in this work will further stimulate developments and applications of this method.
 
-\begin{acknowledgments}
-We are grateful to S.~Shore, M.~Maugeri and C.~Rossi for useful discussions.\\
+\paragraph{Acknowledgments}
+We are grateful to S.~Biscoveanu, D.~Laghi, M.~Maugeri, C.~Rossi and S.~Shore for useful comments and discussions.\\
 This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center (\url{https://www.gw-openscience.org/}), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration.
@@
 Virgo is funded, through the European Gravitational Observatory (EGO), by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef, with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain.
-\end{acknowledgments}
 
-\pagebreak
-\appendix
+	\bibliographystyle{aa}
+	\bibliography{Bibliography.bib}
+
+\begin{appendix}
 \section{Details of PSD computation} \label{sec:appendix}
 \subsection{MESA solution}\label{sec:MESA_solution}
-We derive the expression for the MAXENT spectral estimator following the approach proposed by Burg \cite{burg1975maximum}.
-Unlike the standard approach, we do not enforce the constraints in eq.~\eqref{eq:MaxConstraint} with the standard Lagrange Multipliers approach.
+We derive the expression for the MAXENT spectral estimator following the approach proposed by \citet{burg1975maximum}.
+Unlike the standard approach, we do not enforce the constraints in Eq.~\eqref{eq:MaxConstraint} with the standard Lagrange Multipliers approach.
 We write instead the PSD $S(f)$ as the Fourier Transform of the sample autocorrelation function: 
@@
 
-From eq.~\eqref{eq:lamdas} and from the properties above, is easily seen from the properties of the Fourier transform that $S(f)$ can be expressed via a Fourier Series 
+From Eq.~\eqref{eq:lamdas} and from the properties above, is easily seen from the properties of the Fourier transform that $S(f)$ can be expressed via a Fourier Series 
 \begin{equation}\label{eq:PSDconstraint}
@@
 It is easy to show that if $z_0$ is a root for the polynomial $(z_0^*)^{-1}$ is also a root: for every root laying outside the unit circle there will be another root inside of it and vice-versa.
-These properties allow us to rewrite the Fourier expansion \eqref{eq:zExp} as \cite{1975STIN...7714318B}:
+These properties allow us to rewrite the Fourier expansion \eqref{eq:zExp} as \citep{1975STIN...7714318B}:
 \begin{equation}\label{eq:MESApsd_appendix}
@@
 
-To compute the $a_s$ is convenient to plug into eq.~\eqref{eq:MaxConstraint} the Laurent Polynomial exansion for $S(f)$ eq.~\eqref{eq:MESApsd_appendix} and then integrating over $z$ (taking values on ${\mathbb S^1}$).
+To compute the $a_s$ is convenient to plug into Eq.~\eqref{eq:MaxConstraint} the Laurent Polynomial exansion for $S(f)$ eq.~\eqref{eq:MESApsd_appendix} and then integrating over $z$ (taking values on ${\mathbb S^1}$).
 In this way the equation becomes:
@@
 \subsection{Levinson recursion} \label{sec:LevinsonRecursion}
-The solution of the eqs.~(\ref{eq:errorFilter1}-\ref{eq:errorFilter2}) fully determines the functional form of the power spectral density estimator \eqref{eq:MESApsd_appendix}.
-The method for solving the equations is called the Levinson-Durbin recursion \cite{doi:10.1002/sapm1946251261} and it is described in the following.
+The solution of the Eqs.~(\ref{eq:errorFilter1}-\ref{eq:errorFilter2}) fully determines the functional form of the power spectral density estimator \eqref{eq:MESApsd_appendix}.
+The method for solving the equations is called the Levinson-Durbin recursion \citep{doi:10.1002/sapm1946251261} and it is described in the following.
 For each order $N$ of the iteration we define the quantities:
@@
 Burg's method for spectral analysis is solved via Levinson is implemented in the released \texttt{memspectrum} package.
-Another faster recursion method is available in \cite{Vos} and it is also available in \texttt{memspectrum}.
-
-\newpage
-
-	\bibliography{Bibliography.bib}
-	\bibliographystyle{ieeetr}
+Another faster recursion method is available in \citet{Vos} and it is also available in \texttt{memspectrum}.
+\end{appendix}
 

### 2021-05-27 paper/aa.tex
added the style files

%                                                                 aa.dem
% AA vers. 8.2, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                       (c) EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[rnote]{aa} % for the research notes
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
% according to the author-year natbib style

%
\documentclass{aa}

%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
\usepackage{color}
\usepackage{calc}
\usepackage{amsmath,amssymb,graphicx}
\usepackage{tensor}
\usepackage{bm}
\usepackage{times}
\usepackage[varg]{txfonts}
\usepackage[colorlinks, linkcolor=blue, pdfborder={0 0 0}, breaklinks=true]{hyperref}
\usepackage{float}
\usepackage{dcolumn}
\usepackage[nolist,nohyperlinks]{acronym}
\usepackage{xspace}
\usepackage[abs]{overpic}
\usepackage{pict2e}
\usepackage{enumitem}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{gensymb}
\usepackage[capitalize]{cleveref}
\Crefname{figure}{Fig.}{Figs.}% {<type>}{<singular>}{<plural>}
\usepackage[normalem]{ulem}
\usepackage{mathrsfs}
\interfootnotelinepenalty=10000
\usepackage{array}
\usepackage{multirow}
\usepackage{subfigure}
\usepackage{tabularx}
\usepackage{etoolbox}
\usepackage{lineno}
\linenumbers

% WATERMARK settings
%\usepackage{draftwatermark}
%\SetWatermarkText{Draft-V2}
%\SetWatermarkScale{1.3}
%\SetWatermarkColor[gray]{0.85}
%\usepackage{draftwatermark}
%\SetWatermarkText{Draft}
%\SetWatermarkScale{5}

\usepackage{scrextend}
\usepackage{textgreek}
%\renewcommand{\mymacro}[1]{{\ensuremath{#1}}}

%\renewcommand{\ThisEvent}{GW190521\xspace}
\newcommand{\Msun}{\ensuremath{\mathrm{M}_\odot}}
\newcommand{\likelihood}{\ensuremath{\mathcal{L}}}
\newcommand{\note}[1]{\textit{\color{blue}Note: #1}}
\newcommand{\rough}[1]{\textit{\color{brown} #1}}
\newcommand{\todo}[1]{\textbf{\color{red} Todo: #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[options]{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%
\begin{document}

   \title{Hydrodynamics of giant planet formation}

   \subtitle{I. Overviewing the $\kappa$-mechanism}

   \author{G. Wuchterl
          \inst{1}
          \and
          C. Ptolemy\inst{2}\fnmsep\thanks{Just to show the usage
          of the elements in the author field}
          }

   \institute{Institute for Astronomy (IfA), University of Vienna,
              T\"urkenschanzstrasse 17, A-1180 Vienna\\
              \email{wuchterl@amok.ast.univie.ac.at}
         \and
             University of Alexandria, Department of Geography, ...\\
             \email{c.ptolemy@hipparch.uheaven.space}
             \thanks{The university of heaven temporarily does not
                     accept e-mails}
             }

   \date{Received September 15, 1996; accepted March 16, 1997}

% \abstract{}{}{}{}{} 
% 5 {} token are mandatory

  \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
{Intermediate-mass black holes (IMBH), spanning the approximate mass range 
$100$--$10^5$\,\Msun, lie between black holes (BH) formed by stellar collapse
%lying between the known population of black holes (BH) formed via stellar collapse whose 
%mergers have been observed by LIGO-Virgo, 
and the super-massive BH lying at the centre of galaxies.  Mergers of IMBH binaries are 
the most intense gravitational-wave sources accessible by the Advanced detector network. 
Searches of the first two Advanced LIGO-Virgo observing runs did not yield any 
significant IMBH binary signals.  With the increased detector sensitivity of the third
observing run (O3), GW190521, a signal consistent with a binary merger of mass $\sim 150$\,\Msun\, 
provided the direct evidence of IMBH formation.  Here 
%After the first clear IMBH detection implied by the remnant of the merger event GW190521, 
we report on a dedicated search of O3 data
%from the third Advanced LIGO-Virgo observing run 
for further IMBH binary mergers, combining both modelled (template matched filter) and 
unmodelled search methods. 
We detail the most significant new candidate events: %obtained in addition to GW190521; 
\rough{these events are, however, not strongly inconsistent with instrumental noise origin
and thus do not indicate detection of further IMBH mergers.} 
%\footnote{(this statement we need to discuss with detchar)}.  
We quantify the sensitivity of the individual search methods and of the combined search
using a suite of IMBH binary signals simulated by numerical solution of general relativity, 
including the effects of spins misaligned with the binary orbital axis, and give upper 
limits on the corresponding astrophysical merger rates.  Our most stringent limit is for 
[MASSES/SPINS] at [NUMBER]\,Gpc$^{-3}$yr$^{-1}$ (90\% confidence), a factor of [NUMBER] 
more constraining than previous LIGO-Virgo limits.  We also update the estimated rate of
mergers similar to GW190521 to \todo{NUMBER}.}
  % aims heading (mandatory)

  % conclusions heading (optional), leave it empty if necessary 
   {}

   \keywords{giant planet formation --
                $\kappa$-mechanism --
                stability of gas spheres
               }

   \maketitle
%
%________________________________________________________________
%\section{Introduction}

 \input{introduction.tex}
%__________________________________________________________________
\input{data.tex}

%BOB, GV, KC, TD, DM
%\input{searches.tex}

%KC, BOB, TD, WP, PG, DM
%\input{results.tex}

%BOB, KC, TD
\input{rate.tex}

%AP
\input{conclusion.tex}

%\acknowledgments
%\input{LVCack.tex}

%\appendix
%\input{appendix.tex}
%\input{s200114f.tex}

%\section{Conclusions}

\begin{acknowledgements}

\end{acknowledgements}

%-------------------------------------------------------------------

\end{document}

### 2021-05-27 paper/mesa.tex
Clarifications in conclusion

@@
 In this section we perform a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
-Although similar studies can be drawn from any other PSD, for definiteness, we focus on the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog~\citep{GWTC1, PSD_release}, and computed with BayesLine package \citep{Cornish_2015, Littenberg_2015, Chatziioannou_2019}.
+Although similar studies can be drawn from any other PSD, for definiteness, we focus on the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog~\citep{GWTC1, PSD_release}, and computed with BayesLine package \citep{Cornish_2015, Littenberg_2015, Cornish_2020, Chatziioannou_2019}.
 
@@
 
-This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA \citep{LIGO_2015, Virgo_2014, KAGRA_2012, KAGRA_2013}, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalising over the detector noise PSD.
-\sschmidt{There might be a misunderstanding: I guess the referee is talking about taking into account the PSD errors, while we are talking about dropping the stationariety assumption...
+This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA \cite{LIGO_2015, Virgo_2014, KAGRA_2012, KAGRA_2013}, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for online PSD estimation during a Bayesian parameter estimation exercise.
+%a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
+%\sschmidt{There might be a misunderstanding: I guess the referee is talking about taking into account the PSD errors, while we are talking about dropping the stationariety assumption...
 The misleading word is `marginalize', I think}
@@
 where $\tilde{\hphantom{a}}$ denotes the Fourier transform and the signal model $x(t;\theta)$, dependent on some parameter $\theta$, is a prediction for a deterministic signal buried in the observed time series $d(t)$.
-In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source \cite{lalinference}: this scheme assumes stationary data and it might not reflect the structure of noise in the time slice to analyse.
-Several approaches have been developed to try to study the validity of the assumption or even to drop it \citep{R_ver_2010, R_ver_2011, Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave, Edwards_2020, Edwards_2015} \sschmidt{All these citations seems a bit too much for me...}.
-Besides them, MESA can offer an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on $d_t-x_t$.
+In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source \cite{lalinference}: this scheme assumes {\it stationary data} on long timescale and it might not reflect the structure of the noise in the time slice to analyse.
+Some alternatives exist for dropping this assumption and modifying the likelihood accordingly \cite{R_ver_2010, R_ver_2011, Edwards_2020, Edwards_2015}.
+MESA can add to those an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on the residual $d_t-x_t$.
 In this way, the PSD would depend also on $\theta$ and would effectively model the residuals.
-Preliminary studies suggest that this is possible~\citep{martini_thesis}.
+For this method, we would only need to assume the stationariety of the residual on the batch to analyse\footnote{
+Unlike several works in GW data analysis \cite{Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave}, this approach does not address the issue of including the PSD uncertainties in the posterior.
+Instead, it focuses on making the stationariety assumption as minimal as possible.
+}, getting, as a bonus, a posterior distribution for the PSD of the analyzed data.
+Preliminary studies suggest that this is possible~\cite{martini_thesis}.
+
+%Several approaches have been developed to try to study the validity of the assumption or even to drop it \cite{R_ver_2010, R_ver_2011, Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave, Edwards_2020, Edwards_2015} \sschmidt{All these citations seems a bit too much for me...}.
+%Besides them, MESA can offer an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on $d_t-x_t$.
+%In this way, the PSD would depend also on $\theta$ and would effectively model the residuals. Preliminary studies suggest that this is possible~\cite{martini_thesis}.
+
 Furthermore, MESA provides a simple, but robust and quite accurate, albeit for short times, predictor for the time series.

### 2021-06-07 paper/referee_response.md
Some edits to the paper

@@
 al paper in [29].
-	+ Done: Bayeswave (as well as a study of its accuracy) is cited in sec.IV, page 10, together with 1907.06540 that assess its accuracy; papers on glitches are cited on sec. V.B, page 14
+	+ Done: Bayeswave (as well as a study of its accuracy) is cited in sec.IV, page 10, together with 1907.06540 that assess its accuracy; papers on glitches are cited on sec. 5.2
 
@@
 - "hence MESA could be employed for a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD”: Please add citations to other papers from LVC members discussing marginalization over PSD detector noise (0804.3853, 1109.0442, 1909.01934, 1409.7215, 1307.8195, 2004.05149, 2006.05292, 2101.01200, 2004.07515, 1506.00185 and the two Bayeswave papers mentioned above).
-	+ The relevant part has been changed so as to acknowledge previous works. Probably here a _misunderstanding_ on what marginalization means?
-		Everything is at page 15: "several approaches have been developed to try to study the validity of the assumption or even to drop it [46–52]"
+	+ Perhaps here there was a misundenstarding, as we did not make clear our idea.
+We do not aim to include in the PE the uncertainties of the PSD (as some works indicated by the referee do).
+Instead, our idea is to estimate the PSD at _every_ LL evaluation during a PE run.
+We tried to clarify this in the conclusion in the text and in a dedicated footnote.
 
 - Please add a citation to https://arxiv.org/abs/gr-qc/0011041, which also demonstrates the applicability of autoregressive processes to gravitational-wave power spectral density estimation.
-	+ Done: on sec. V.B, page 13
+	+ Done: on sec. 5.2
 
@@
 In addition to citing the papers described above, I think it would be appropriate to add a discussion, either in the Introduction or the Final Remarks comparing MESA to this “on-source” PSD estimation and citing the relvant papers disucssing the Bayeswave PSD method in comparison to the Welch method, some repeated from the above requests (1907.06540, 2006.05292).
-	+ Work in progress: This is a very good suggestion...
-Are we gonna include BayesLine among our comparison?
-It looks like a good idea.
-If we do so, probably we want to say something about it in the beginning.The paper 1907.06540 is already cited together with BayesWave
+	+ We found that the comparison with Bayesline goes Beyond the scope of our work.
+Bayesline is a parametric model (unlike the non parametric Welch and MESA) and it is taylored to GW data analysis.
+Our discussion here aimed to be inspired from but broader than GW data analysis.
+Thus, a comparison with BayesLine would narrow too much the applicability and the interest of our work.
+	We mentioned this point in a newly added paragraph in the introduction (where we cited the two papers suggested).
 
@@
 It would be very helpful to elaborate on this, perhaps when the connection to AR is introduced.
-	+ Work in progress: I guess we could just point out that it is just a way to validate the accuracy of the method and to test which loss function is better. Does it makes sense?
+	+ We added at the beginning of the relevant section (3.4) a note on this, trying to explain the importance of such study.
 
@@
 It would be helpful to elaborate on this.
-	+ We added a claryfing footnote on this (note 10, page 13). We hope it is clear.
+	+ We added a claryfing footnote on this (note 10, page 10). We hope it is clearer now.
 
 - "Indeed, a precise prediction of the strain time series can be beneficial in the detection of anomalies in the data and, eventually, their removal.”: How does this compare to the method presented in https://arxiv.org/abs/1908.05644, where the data containing a glitch are “inpainted” with values chosen such that the inverse-PSD-filtered data is 0 at these times?
-	+ Work in progress: we should understand what's in the paper...
+	+ To answer this matter is not easy and this should be addressed in future works. Thanks for pointing out the reference: we added it
 

### 2021-06-07 paper/mesa.tex
Some edits to the paper

@@
 
-An appealing alternative approach, based on the Maximum Entropy principle \citep{JaynesArticle,jaynes2003ptl, Jaynes_MAXENT}, has been put forward by~\citet{burg1975maximum}.
+\sschmidt{This could be a footnote?}\textbf{A different approach, which provides a smooth PSD estimation, is to set up a parametric model for the PSD and to fit its parameters to the data with a Markov chain Monte Carlo \cite{Cornish_2015, Littenberg_2015}.
+Despite being effective, such method is very problem dependent, as it makes heavy assumption on the shape of the PSD.
+Moreover, it is computationally expensive and it does not come with an handy implementation available to the public.
+For all those reasons, we do not consider such method in our work.}
+
+An appealing alternative, based on the Maximum Entropy principle \citep{JaynesArticle,jaynes2003ptl, Jaynes_MAXENT}, has been put forward by~\citet{burg1975maximum}.
 Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the algorithm parameters, since it provides an iterative closed form expression for the spectrum of a stochastic stationary time series.
@@
 We now address the issue of how well the AR order (i.e. the number of $a_k$ coefficients employed) is estimated by each loss function.
-In doing so, we generate $100$ autoregressive processes $AR(p)$ with a random value of $p$, drawn such that $log(p) \sim\mathcal{U}_{[log(2), log(5000)]}$.
+This is useful to further characterize the properties of the different loss functions.
+Furthermore, as MESA is supposed to model data as an $AR(p)$ process, it is interesting to quantify its precision on some true autoregressive (and stationary) time series.
+
+We generate $100$ autoregressive processes $AR(p)$ with a random value of $p$, drawn such that $log(p) \sim\mathcal{U}_{[log(2), log(5000)]}$.
 Each coefficient $a_k$ is assigned according to a Dirichlet distribution $\mathrm{Dir}([1,\hdots, 1])$.
@@
 
-\sschmidt{How does this compare to the method presented in https://arxiv.org/abs/1908.05644? Seems like an interesting paper that tells you how to produce data after glitch removal.}
 Despite poor predictions on long timescales, the AR process obtained with MESA still can be useful on short timescales.
 Indeed, a precise prediction of the strain time series can be beneficial in the detection of anomalies in the data and, eventually, their removal.
-Indeed loud anomalies in the data, called {\it glitches}, pose a major challenge to the ability of detecting signals and intensive work has been done to mitigate the disruption in sensitivity they cause \citep{Nuttall_2015, Noise_char_2016, Zevin_2017} and to develop effective subraction techniques \citep{Pankow_2018}.
+Indeed loud anomalies in the data, called {\it glitches}, pose a major challenge to the ability of detecting signals and intensive work has been done to mitigate the disruption in sensitivity they cause \citep{Nuttall_2015, Noise_char_2016, Zevin_2017} and to develop effective subraction techniques \citep{Pankow_2018, zackay2019detecting}.
 The predictions can form an expected baseline for the strain; any anomaly (i.e. a glitch or even a transient of physical origin) can show up as a large departure from expected trend.
 Moreover, if a glitch is detected, its shape can be estimated (as well as the confidence level) by subtracting the expected signal with the actual signal.
-This can work: a typical glitch can last as long as $\SI{0.2}{s}$, close to the time scale over which the forecasting is reliable.
+This can work: a typical glitch can last as long as $\SI{0.2}{s}$, close to the time scale over which the forecasting is reliable. Future works will explore this exicting opportunity.
 
-As noted above, for pushing further the prediction performance from our simple linear predictor, more sophisticated forecasting methods are available in the Machine Learning literature \citep[see]{LSTM_article, oord2016wavenet} and they can be trained for precision forecasting.
+As noted above, for pushing further the prediction performance from our simple linear predictor, more sophisticated forecasting methods are available in the Machine Learning literature (see e.g.
+\cite{LSTM_article, oord2016wavenet}) and they can be trained for precision forecasting.
 However, a Maximum-Entropy trained AR process can be a simple baseline for comparing such more advanced methods and might suffice for many purposes.
@@
 
-This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA \cite{LIGO_2015, Virgo_2014, KAGRA_2012, KAGRA_2013}, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD, hence MESA could be employed for online PSD estimation during a Bayesian parameter estimation exercise.
+This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD.
+Hence MESA could be employed for {\it online} PSD estimation during a Bayesian parameter estimation exercise.
 %a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
@@
 where $\tilde{\hphantom{a}}$ denotes the Fourier transform and the signal model $x(t;\theta)$, dependent on some parameter $\theta$, is a prediction for a deterministic signal buried in the observed time series $d(t)$.
-In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source \cite{lalinference}: this scheme assumes {\it stationary data} on long timescale and it might not reflect the structure of the noise in the time slice to analyse.
+In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source \cite{lalinference}: this scheme assumes {\it stationary data} on a long timescale (longer than the data under study) and it might not reflect the structure of the noise in the analyzed time slice.
 Some alternatives exist for dropping this assumption and modifying the likelihood accordingly \cite{R_ver_2010, R_ver_2011, Edwards_2020, Edwards_2015}.
@@
 Unlike several works in GW data analysis \cite{Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave}, this approach does not address the issue of including the PSD uncertainties in the posterior.
-Instead, it focuses on making the stationariety assumption as minimal as possible.
-}, getting, as a bonus, a posterior distribution for the PSD of the analyzed data.
+Instead, it focuses on making minimal assumption about the nature of the data under study.
+}, making a lighter assumption on the nature of the data. Furthermore, this method would get, as a bonus, a posterior distribution for the PSD of the analyzed data.
 Preliminary studies suggest that this is possible~\cite{martini_thesis}.

### 2021-06-07 paper/referee_response.md
Some edits to the paper-bis

@@
 - In addition to adding the citations to other LVC works presenting methods and results for marginalizing over the PSD uncertainty (see PnP comments above), a more detailed discussion of how MESA compares to these other methods (particularly the error in the PSD measurement) would be appropriate here.
-	+ Work in progress: shall we quote some numbers about the accuracy? Or we just reitarete the qualitative findings we had?
+	+ We believe we discussed in details the matter in section 4, which we refer the reader to.
+At the beginning of the section, we addeed two lines explaining why the comparison shall be qualitative (being the results very problem dependent, we believe that every measure of error hardly generalises to other problems).

### 2021-06-07 paper/mesa.tex
Some edits to the paper-bis

@@
 \section{Comparison with Welch method}\label{sec:Welch_comparison}
-\sschmidt{Do we want to make a comparison with BayesLine? I think so... \\ Do we want to introduce this method in the introduction? Makes sense}
-In this section we perform a qualitative comparison between the performance of the MESA and of the standard Welch algorithm.
-Although similar studies can be drawn from any other PSD, for definiteness, we focus on the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog~\citep{GWTC1, PSD_release}, and computed with BayesLine package \citep{Cornish_2015, Littenberg_2015, Cornish_2020, Chatziioannou_2019}.
+We perform a {\it qualitative} comparison between the performance of the MESA and of the standard Welch algorithm.
+In this, we cannot avoid to be only qualitative.
+Indeed, as the results of the comparison are problem dependent, it is very hard to quantify this in a single metric.
+Although similar studies can be drawn from any other PSD, in this section we focus on a single PSD and we try to generalize some observations that we make.
+We decide to use the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog~\citep{GWTC1, PSD_release}, and computed with BayesLine package \citep{Cornish_2015, Littenberg_2015, Cornish_2020, Chatziioannou_2019}.
 

### 2021-06-08 paper/mesa.tex
Some edits to the paper- ter
minor rephrasing and a few fixes to the citations

@@
 
-\sschmidt{This could be a footnote?}\textbf{A different approach, which provides a smooth PSD estimation, is to set up a parametric model for the PSD and to fit its parameters to the data with a Markov chain Monte Carlo \cite{Cornish_2015, Littenberg_2015}.
-Despite being effective, such method is very problem dependent, as it makes heavy assumption on the shape of the PSD.
-Moreover, it is computationally expensive and it does not come with an handy implementation available to the public.
-For all those reasons, we do not consider such method in our work.}
+A alternative approach providing a smooth PSD estimation, is to adopt  a parametric model for the PSD and to fit its parameters to the data with a Markov Chain Monte Carlo \citep[e.g.]{Cornish_2015, Littenberg_2015}.
+Despite being effective, this method is problem dependent, since it needs to make definite assumptions on the shape of the PSD.
+Moreover, it can be computationally expensive and it does not come with an handy implementation available to the public.
+For all the above reasons, we did not consider such method in our work.
 
@@
 The use of MESA is particularly useful for short time series samples, where Welch's method is outperformed in both precision and confidence.
-\sschmidt{Add here something on accuracy on the PSD estimation, as the referee requires}
 
@@
 where $\tilde{\hphantom{a}}$ denotes the Fourier transform and the signal model $x(t;\theta)$, dependent on some parameter $\theta$, is a prediction for a deterministic signal buried in the observed time series $d(t)$.
-In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source \cite{lalinference}: this scheme assumes {\it stationary data} on a long timescale (longer than the data under study) and it might not reflect the structure of the noise in the analyzed time slice.
-Some alternatives exist for dropping this assumption and modifying the likelihood accordingly \cite{R_ver_2010, R_ver_2011, Edwards_2020, Edwards_2015}.
+In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source \citep{lalinference}: this scheme assumes {\it stationary data} on a long timescale (longer than the data under study) and it might not reflect the structure of the noise in the analyzed time slice.
+Some alternatives exist for dropping this assumption and modifying the likelihood accordingly \citep{R_ver_2010, R_ver_2011, Edwards_2020, Edwards_2015}.
 MESA can add to those an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on the residual $d_t-x_t$.
@@
 For this method, we would only need to assume the stationariety of the residual on the batch to analyse\footnote{
-Unlike several works in GW data analysis \cite{Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave}, this approach does not address the issue of including the PSD uncertainties in the posterior.
+Unlike several works in GW data analysis \citep{Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave}, this approach does not address the issue of including the PSD uncertainties in the posterior.
 Instead, it focuses on making minimal assumption about the nature of the data under study.
 }, making a lighter assumption on the nature of the data. Furthermore, this method would get, as a bonus, a posterior distribution for the PSD of the analyzed data.
-Preliminary studies suggest that this is possible~\cite{martini_thesis}.
+Preliminary studies suggest that this is possible~\citep{martini_thesis}.
 

### 2021-06-09 paper/mesa.tex
Response

@@
 \end{equation}
-In some literature, especially in the context of gravitational waves physics, e.g. ~\cite{Finn_1992}, the PSD is introduced as  
+In some literature, especially in the context of gravitational waves physics, e.g. ~\citep{Finn_1992}, the PSD is introduced as  
 \begin{equation}\label{eq:psd-f-definition}
@@
 \begin{equation}\label{eq:MESApsd}
-    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=1}^N a_s z^s\right)\left(\sum_{s = 1}^N a^*_s z^{-s}\right)}, 
+    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^{N} a_s z^s\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}, 
 \end{equation}
-where $\Delta t$ is the sampling interval of the time series, $z=\exp{(2\pi i f\Delta t)}$, $a_1$ = 1.
-The vector obtained as $(1, a_2, \dots, a_N)$ is also known as the \textit{prediction error filter}.
+where $\Delta t$ is the sampling interval of the time series, $z=\exp{(2\pi i f\Delta t)}$, $a_0$ = 1.
+The vector obtained as $(1, a_1, \dots, a_N)$ is also known as the \textit{prediction error filter}.
 The coefficients $a_s (s > 0)$, together with an overall multiplicative scale factor $P_N$, are to be determined by an iterative process (called Burg's algorithm).
@@
 As noted above, for pushing further the prediction performance from our simple linear predictor, more sophisticated forecasting methods are available in the Machine Learning literature (see e.g.
-\cite{LSTM_article, oord2016wavenet}) and they can be trained for precision forecasting.
+\citep{LSTM_article, oord2016wavenet, lea2016tcn}) and they can be trained for precision forecasting.
 However, a Maximum-Entropy trained AR process can be a simple baseline for comparing such more advanced methods and might suffice for many purposes.
@@
 \begin{equation}
-	\log\mathcal{L}(d_t | \theta) \propto  -\frac{1}{2} \frac{(\tilde{d}(f)- \tilde{x}(f;\theta))^2}{S(f)} - \frac{1}{2}\log\left[\int  df\,S(f)\right]
+	\log\mathcal{L}(d_t | \theta) \propto  -\frac{1}{2} \frac{|\tilde{d}(f)- \tilde{x}(f;\theta)|^2}{S(f)} - \frac{1}{2}\log\left[\int  df\,S(f)\right]
 \end{equation}
 where $\tilde{\hphantom{a}}$ denotes the Fourier transform and the signal model $x(t;\theta)$, dependent on some parameter $\theta$, is a prediction for a deterministic signal buried in the observed time series $d(t)$.
-In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source \citep{lalinference}: this scheme assumes {\it stationary data} on a long timescale (longer than the data under study) and it might not reflect the structure of the noise in the analyzed time slice.
-Some alternatives exist for dropping this assumption and modifying the likelihood accordingly \citep{R_ver_2010, R_ver_2011, Edwards_2020, Edwards_2015}.
+In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source: this scheme assumes {\it stationary data} on a long timescale (longer than the data under study) and it might not reflect the structure of the noise in the analyzed time slice.
+Some alternatives exist for dropping this assumption and modifying the likelihood accordingly \citep{R_ver_2010, R_ver_2011, Edwards_2020, Chatziioannou_2021}.
 MESA can add to those an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on the residual $d_t-x_t$.
@@
 For this method, we would only need to assume the stationariety of the residual on the batch to analyse\footnote{
-Unlike several works in GW data analysis \citep{Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave}, this approach does not address the issue of including the PSD uncertainties in the posterior.
+Unlike several works in GW data analysis \citep{Biscoveanu_2020, talbot2020gravitationalwave}, this approach does not address the issue of including the PSD uncertainties in the posterior.
 Instead, it focuses on making minimal assumption about the nature of the data under study.
@@
 Preliminary studies suggest that this is possible~\citep{martini_thesis}.
+Other studies have been done in this direction, mostly using a parametric model for the PSD \citep{Littenberg_2013, Edwards_2015, lalinference}.
 

### 2023-07-11 paper/mesa.tex
added place holder text and figures

@@
 We characterize our implementation by performing a power spectral density analysis on synthetic data (with known power spectral density) and we compare different criteria for stopping the recursion.
-Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using synthetic data generated from the released spectrum by the LIGO-Virgo collaboration.
+Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using synthetic data generated from the released spectrum by the LIGO-Virgo-Kagra collaboration.
 We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
-systematically lower variance and bias. This is particularly manifest in the case of a little number of data points, making Burg's method most suitable to work in this regime.}
+systematically lower variance and bias. This is particularly manifest in the case of a little number of data points, making Burg's method most suitable to work in this regime. 
+We conclude our investigation by utilising MESA, and its particularly easy parametrisation where the only free parameter is the order $p$ of the AR process, to marginalise over the interferometers noise PSD in conjunction with inferring the parameters of GW150914.}
 
@@
 Sec.~\ref{sec:validation} presents the validation of Burg's method as well as of our implementation on simulated data.
-In Sec.~\ref{sec:Welch_comparison} we compare the results from \texttt{memspectrum} with the Welch method; Sec.~\ref{sec:applications} presents a few applications to real time series and, finally, we conclude with a discussion in Sec.~\ref{sec:conclusion}.
+In Sec.~\ref{sec:Welch_comparison} we compare the results from \texttt{memspectrum} with the Welch method; Sec.~\ref{sec:applications} presents a few applications to real time series, including the analysis of GW150914, and, finally, we conclude with a discussion in Sec.~\ref{sec:conclusion}.
 
@@
 \section{Applications} \label{sec:applications}
-\subsection{Temperature Time Series} \label{sec:temp_timeseries}
-As a further example of the breadth of applicability of MESA, we applied our implementation to atmospheric temperature time series. 
-The reason for this choice is twofold: i) atmospheric temperature time series present a variety of overlapping periodicities most of which are 
-known; ii) it provides a stress test for the time series forecast analysis.  As dataset, we  used the historical reanalysis data 
-from the ``ERA5-Land hourly data from 1981 to present" dataset, downloaded from the Climate Data Store~\citep{ERA5_dataset}.
-The data consist in temperatures taken at coordinates $\textrm{N}\;\ang{45;5;} \;\; \textrm{E}\;\ang{9;1;}}$, corresponding about to the city of Milan, Italy.
+%\subsection{Temperature Time Series} \label{sec:temp_timeseries}
+%As a further example of the breadth of applicability of MESA, we applied our implementation to atmospheric temperature time series. 
+%The reason for this choice is twofold: i) atmospheric temperature time series present a variety of overlapping periodicities most of which are 
+%known; ii) it provides a stress test for the time series forecast analysis.  As dataset, we  used the historical reanalysis data 
+%from the ``ERA5-Land hourly data from 1981 to present" dataset, downloaded from the Climate Data Store~\citep{ERA5_dataset}.
+%The data consist in temperatures taken at coordinates $\textrm{N}\;\ang{45;5;} \;\; \textrm{E}\;\ang{9;1;}}$, corresponding about to the city of Milan, Italy.
 The temperatures are given on an hourly cadence for almost 31 years from 31st December 1989 to 30th November 2020.
-
-Fig.~\ref{fig:temp_spectrum}, shows the MESA spectrum inferred from the data. As a comparison, we adopt both the FPE and the CAT loss functions.
-In agreement with what we found in previous sections, the FPE PSD is more regular compared to the one from CAT. Both spectra show a
-peak at the frequency $f_D = \SI{1}{day^{-1}}$ corresponding to one day, corresponding to the day-night cycle: this is expected.
-Higher order harmonics (corresponding to integer multiples of $f_D$) are also visible up to the Nyquist frequency ($f_D = \SI{0.5}{hour^{-1}}$): they correspond to signals that preserve the 1 day period while, at the same time, capturing the complex variability of the daily temperature cycle throughout the year.
-
-Looking at low frequencies, FPE does not capture the yearly variability at $f_{yr} = \SI{1}{yr^{-1}}$.
+%
+%Fig.~\ref{fig:temp_spectrum}, shows the MESA spectrum inferred from the data. As a comparison, we adopt both the FPE and the CAT loss functions.
+%In agreement with what we found in previous sections, the FPE PSD is more regular compared to the one from CAT. Both spectra show a
+%peak at the frequency $f_D = \SI{1}{day^{-1}}$ corresponding to one day, corresponding to the day-night cycle: this is expected.
+%Higher order harmonics (corresponding to integer multiples of $f_D$) are also visible up to the Nyquist frequency ($f_D = \SI{0.5}{hour^{-1}}$): they correspond to signals that preserve the 1 day period while, at the same time, capturing the complex variability of the daily temperature cycle throughout the year.
+%
+%Looking at low frequencies, FPE does not capture the yearly variability at $f_{yr} = \SI{1}{yr^{-1}}$.
 On the other hand, the peak is captured well by the CAT loss function.
@@
 As in the previous case, this is required to better model the temperature variation in the year.
-
-The failure of FPE in capturing the low frequency trend can be understood by looking at Fig.~\ref{fig:p_vs_ptrue}.
+%
+%The failure of FPE in capturing the low frequency trend can be understood by looking at Fig.~\ref{fig:p_vs_ptrue}.
 It is shown that CAT systematically select very long filters (i.e.
 large $p_{CAT}$ values for the autoregressive orders), whereas FPE tends to select smaller values for the order $p_{FPE}$ of the autoregressive process, often underestimating the actual value.
-In this example, we have $p_{CAT} \sim  36500 \sim  \SI{4}{yr}$, whereas $p_{FPE} \sim  1300 \sim  \SI{2}{months}$.
-The autoregressive process selected by FPE, hence cannot produce a meaningful prediction on the timescale longer than a few weeks and for this reason it is unable to capture the low frequency peak at $f_{yr}$.
+%In this example, we have $p_{CAT} \sim  36500 \sim  \SI{4}{yr}$, whereas $p_{FPE} \sim  1300 \sim  \SI{2}{months}$.
+%The autoregressive process selected by FPE, hence cannot produce a meaningful prediction on the timescale longer than a few weeks and for this reason it is unable to capture the low frequency peak at $f_{yr}$.
 In other words, the model is too simple to model \textit{both} the behaviour at high and low frequencies (separated by approximately 3 orders of magnitude).
-On the other hand, the model chosen by CAT is much more complex and is able to model also the high frequency behaviour, at the expense of making a more noisy estimation.
-The ``actual" length of the autoregressive process should be closer to the choice of CAT.
-
-We now assess the accuracy of the forecasting of new observations of the time series. 
-Based on actual data, we try to predict future values as described in sec.~\ref{sec:forecasting} (of course the MESA is performed with CAT loss function).
+%On the other hand, the model chosen by CAT is much more complex and is able to model also the high frequency behaviour, at the expense of making a more noisy estimation.
+%The ``actual" length of the autoregressive process should be closer to the choice of CAT.
+%
+%We now assess the accuracy of the forecasting of new observations of the time series. 
+%Based on actual data, we try to predict future values as described in sec.~\ref{sec:forecasting} (of course the MESA is performed with CAT loss function).
 We produce $N =100$ independent predictions and we compute the median as well as the $90\%$ confidence interval.
@@
 We report our results in Fig.~\ref{fig:temp_spectrum}.
-
-We note the observed difference is always well included in the $90\%$ confidence interval: the forecasting predictions seem reliable.
-On the other hand, the confidence interval is pretty large (almost as large as $\SI{15}{K}$), thus making ``easy" for the actual data to fit the predictions.
-Indeed, the prediction model, while suitable for spectral estimation, is nothing more than a linear regression (plus noise term).
+%
+%We note the observed difference is always well included in the $90\%$ confidence interval: the forecasting predictions seem reliable.
+%On the other hand, the confidence interval is pretty large (almost as large as $\SI{15}{K}$), thus making ``easy" for the actual data to fit the predictions.
+%Indeed, the prediction model, while suitable for spectral estimation, is nothing more than a linear regression (plus noise term).
 Such a simple model hardly catches the complex trend in the variability of the temperature daily trend during a year.
 For more precise predictions, probably one should consider nonlinear regression model, tapping into the wealth of non linear predictors offered by the field of Deep Learning.
-
-\begin{figure}
-	\caption{Spectrum for the historical temperature time series.
+%
+%\begin{figure}
+%	\caption{Spectrum for the historical temperature time series.
 The two lines refers to different loss functions (CAT and FPE); the inset shows the harmonics of the fundamental frequency of a day.
 Two vertical lines are drawn in correspondence to the frequency $f_D$ of a day and $f_{yr}$ of a year.}
-	\label{fig:temp_spectrum}
-	\includegraphics{Images/climate_plots/temp_spectrum.pdf}
+%	\label{fig:temp_spectrum}
+%	\includegraphics{Images/climate_plots/temp_spectrum.pdf}
+%\end{figure}
+%
+%\begin{figure}
+%	\caption{Difference between the historical temperature $T$ and the forecast temperature $T_{forecast}$ for two years of data.
+The red shadow denotes the $90\%$ confidence interval of the predictions.
+The model is trained with CAT loss function.
+The discrepancy between predicted and actual values is always everywhere included in the $90\%$ confidence interval.}
+%	\label{fig:temp_forecast}
+%	\includegraphics{Images/climate_plots/forecast_accuracy.pdf}
+%\end{figure}
+
+\subsection{Marginalising over the noise PSD in GW parameter estimation: the case for GW150914}
+As an application of MESA, and its implementation in \texttt{memspectrum}, we analyse GW150914\cite{} using a Bayesian framework that allows for the marginalisation of the order  $p$ of the AR($p$) process representing the noise data stream.
+The inference algorithm proceeds as usual, see \cite{}[veitch and vecchio, ashton et al, etc etc], with a modification to the likelihood calculation.
+The likelihood function we employ is given in Eq.~\eqref{eq:Whittle} and its calculation proceeds as follows: for all proposed Monte Carlo moves $\theta\equiv\theta_{GW}\cup p$\footnote{We indicate the set of all GW parameters (component masses, spins, luminosity distance, etc.) with $\theta_{GW}$.}, for each detector included in the analysis, we compute the detector response to the template $h(t;\theta_{GW})$ in the time-domain, subtract it from the data $d(t)$ to get the time domain residuals $r(t;\theta_{GW})$.
+Under the MESA assumption, we describe the PSD of the residuals as in Eq.~\eqref{eq:MESApsd}, where the order $p$ is fixed by the Monte Carlo sampler.
+Having computed the PSD, we Fourier transform the residuals and compute the likelihood.
+Our sampler is a nested sampling algorithm\cite{}[raynest] and the specific inference model is implemented as part of \texttt{granite}\cite{}[in preparation], a dedicated inference model for ground based interferometric detectors.
+Needless to say, despite the implementation of the fast Burg algorithm in \texttt{memspectrum}, the inference with noise marginalisation is vastly more computational expensive than the standard analysis.
+As our analysis template, we adopt the machine learning based MLGW model\cite{}[MLGW], an aligned spin model trained on TEOBResumS\cite{}, that has been shown to perform well on LVK detected events\cite{}[MLGW].
+\begin{figure}
+	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
+The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
+The confidence interval is computed on  $100$ independent random realization of the predictions.
+	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
+	}
+	\label{fig:gw150914_intrinsic}
+	\includegraphics{Images/gw150914/intrinsic.pdf}
+\end{figure}
+\begin{figure}
+	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
+The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
+The confidence interval is computed on  $100$ independent random realization of the predictions.
+	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
+	}
+	\label{fig:gw150914_intrinsic}
+	\includegraphics{Images/gw150914/extrinsic.pdf}
+\end{figure}
+\begin{figure}
+	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
+The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
+The confidence interval is computed on  $100$ independent random realization of the predictions.
+	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
+	}
+	\label{fig:gw150914_intrinsic}
+	\includegraphics{Images/gw150914/waveforms.pdf}
+\end{figure}
+\begin{figure}
+	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
+The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
+The confidence interval is computed on  $100$ independent random realization of the predictions.
+	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
+	}
+	\label{fig:gw150914_intrinsic}
+	\includegraphics{Images/gw150914/posterior_order_H1.pdf}
 \end{figure}
@@
 \begin{figure}
-	\caption{Difference between the historical temperature $T$ and the forecast temperature $T_{forecast}$ for two years of data.
-The red shadow denotes the $90\%$ confidence interval of the predictions.
-The model is trained with CAT loss function.
-The discrepancy between predicted and actual values is always everywhere included in the $90\%$ confidence interval.}
-	\label{fig:temp_forecast}
-	\includegraphics{Images/climate_plots/forecast_accuracy.pdf}
+	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
+The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
+The confidence interval is computed on  $100$ independent random realization of the predictions.
+	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
+	}
+	\label{fig:gw150914_intrinsic}
+	\includegraphics{Images/gw150914/posterior_order_L1.pdf}
 \end{figure}
+
+\begin{figure}
+	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
+The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
+The confidence interval is computed on  $100$ independent random realization of the predictions.
+	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
+	}
+	\label{fig:gw150914_intrinsic}
+	\includegraphics{Images/gw150914/noise_curves.pdf}
+\end{figure}
+
+In Figs.~\ref{fig:gw150914_intrinsic}, ~\ref{fig:gw150914_extrinsic} and \ref{fig:gw150914_wf} we show the posteriors for the set of intrinsic parameters, extrinsic parameters and reconstructed waveform from our analysis.
+In addition, Fig.~ \ref{fig:gw150914_orders} shows the posterior distribution on the orders $p_{H}$ and $p_{L}$ for the AR processes describing the residuals in the LIGO-Hanford and LIGO-Livingston detectors.
+The inferred PSDs are instead shown in Fig.~\ref{fig:gw150914_psds}.
 

### 2023-07-12 paper/mesa.tex
added the figures and more text

@@
 \subsection{Marginalising over the noise PSD in GW parameter estimation: the case for GW150914}
-As an application of MESA, and its implementation in \texttt{memspectrum}, we analyse GW150914\cite{} using a Bayesian framework that allows for the marginalisation of the order  $p$ of the AR($p$) process representing the noise data stream.
+Since the order of the AR process is not yet determined by the theory, up to this point, we opted for an in-depth investigation of several proposals in the literature and found that different loss functions are required for different situations, with the FPE loss function being the most indicated to deal with gravitational wave data.
+Along these lines, we directly compared the PSDs computed with MESA with the canonical Welch's algorithm.
+As outlined in Sec.~\ref{sec:Welch_comparison}, MESA provides PSD estimates with smaller variance and better accuracy than Welch algorithm.
+The use of MESA is particularly useful for short time series samples, where Welch's method is outperformed in both precision and confidence.
+
+This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD.
+Hence MESA could be employed for {\it online} PSD estimation during a Bayesian parameter estimation exercise.
+%a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
+%\sschmidt{There might be a misunderstanding: I guess the referee is talking about taking into account the PSD errors, while we are talking about dropping the stationariety assumption...
+The misleading word is `marginalize', I think}
+This is possible whenever the log-likelihood of the model for a time series takes the form
+\begin{equation}
+	\log\mathcal{L}(d_t | \theta) \propto  -\frac{1}{2} \frac{|\tilde{d}(f)- \tilde{x}(f;\theta)|^2}{S(f)} - \frac{1}{2}\log\left[\int  df\,S(f)\right]
+\end{equation}
+where $\tilde{\hphantom{a}}$ denotes the Fourier transform and the signal model $x(t;\theta)$, dependent on some parameter $\theta$, is a prediction for a deterministic signal buried in the observed time series $d(t)$.
+In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source: this scheme assumes {\it stationary data} on a long timescale (longer than the data under study) and it might not reflect the structure of the noise in the analyzed time slice.
+Some alternatives exist for dropping this assumption and modifying the likelihood accordingly \citep{R_ver_2010, R_ver_2011, Edwards_2020, Chatziioannou_2021}.
+MESA can add to those an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on the residual $d_t-x_t$.
+In this way, the PSD would depend also on $\theta$ and would effectively model the residuals.
+For this method, we would only need to assume the stationariety of the residual on the batch to analyse\footnote{
+Unlike several works in GW data analysis \citep{Biscoveanu_2020, talbot2020gravitationalwave}, this approach does not address the issue of including the PSD uncertainties in the posterior.
+Instead, it focuses on making minimal assumption about the nature of the data under study.
+}, making a lighter assumption on the nature of the data. Furthermore, this method would get, as a bonus, a posterior distribution for the PSD of the analyzed data.
+Preliminary studies suggest that this is possible~\citep{martini_thesis}.
+Other studies have been done in this direction, mostly using a parametric model for the PSD \citep{Littenberg_2013, Edwards_2015, lalinference}.
+
+Hence, as an application of MESA, and its implementation in \texttt{memspectrum}, we analyse GW150914\cite{} using a Bayesian framework that allows for the marginalisation of the order  $p$ of the AR($p$) process representing the noise data stream.
 The inference algorithm proceeds as usual, see \cite{}[veitch and vecchio, ashton et al, etc etc], with a modification to the likelihood calculation.
@@
 As our analysis template, we adopt the machine learning based MLGW model\cite{}[MLGW], an aligned spin model trained on TEOBResumS\cite{}, that has been shown to perform well on LVK detected events\cite{}[MLGW].
+
 \begin{figure}
-	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
+	\caption{Posterior distributions for $\mathcal{M}$, $q$ and spin parameters $\chi_1$ and $\chi_2$. The recovered posteriors are very similar to the LVK released ones.}
+	\label{fig:gw150914_intrinsic}
+	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/intrinsic.pdf}
+\end{figure}
+\begin{figure}
+	\caption{Posterior distributions for the sky position, (log) luminosity distance, inclination angle, polarisation angle $\psi$, reference phase $\phi_{ref}$ and geocentric coalescence time.
+Also for the extrinsic parameters, we see broad consistency with the LVK released ones.
+However, we notice that the posterior on the declination angle $\delta$ is mirrored compared to the released sky map.
+Nonetheless, there is substantial overlap.
+Similarly, the PSD marginalised posterior for the inclination angle shows a slight preference for a face-on configuration, rather than a face-off.}
+	\label{fig:gw150914_extrinsic}
+	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/extrinsic.pdf}
+\end{figure}
+\begin{figure}
+	\caption{Whitened reconstructed waveforms from our analysis for the Hanford detector (top panel) and the Livingston detector (bottom panel).
+The shaded turquoise area indicates the 90\% credible region.}
+	\label{fig:gw150914_waveforms}
+	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/waveforms.pdf}
+\end{figure}
+%\begin{figure}
+%	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
 The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
 The confidence interval is computed on  $100$ independent random realization of the predictions.
-	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
-	}
-	\label{fig:gw150914_intrinsic}
-	\includegraphics{Images/gw150914/intrinsic.pdf}
-\end{figure}
-\begin{figure}
-	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
+%	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
+%	}
+%	\label{fig:gw150914_intrinsic}
+%	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/posterior_order_H1.pdf}
+%\end{figure}
+%
+%\begin{figure}
+%	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
 The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
 The confidence interval is computed on  $100$ independent random realization of the predictions.
-	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
-	}
-	\label{fig:gw150914_intrinsic}
-	\includegraphics{Images/gw150914/extrinsic.pdf}
-\end{figure}
+%	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
+%	}
+%	\label{fig:gw150914_intrinsic}
+%	\includegraphics{Images/gw150914/posterior_order_L1.pdf}
+%\end{figure}
+
 \begin{figure}
-	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
-The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
-The confidence interval is computed on  $100$ independent random realization of the predictions.
-	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
-	}
-	\label{fig:gw150914_intrinsic}
-	\includegraphics{Images/gw150914/waveforms.pdf}
-\end{figure}
-\begin{figure}
-	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
-The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
-The confidence interval is computed on  $100$ independent random realization of the predictions.
-	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
-	}
-	\label{fig:gw150914_intrinsic}
-	\includegraphics{Images/gw150914/posterior_order_H1.pdf}
-\end{figure}
-
-\begin{figure}
-	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
-The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
-The confidence interval is computed on  $100$ independent random realization of the predictions.
-	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
-	}
-	\label{fig:gw150914_intrinsic}
-	\includegraphics{Images/gw150914/posterior_order_L1.pdf}
-\end{figure}
-
-\begin{figure}
-	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
-The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
-The confidence interval is computed on  $100$ independent random realization of the predictions.
-	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
-	}
-	\label{fig:gw150914_intrinsic}
-	\includegraphics{Images/gw150914/noise_curves.pdf}
-\end{figure}
-
-In Figs.~\ref{fig:gw150914_intrinsic}, ~\ref{fig:gw150914_extrinsic} and \ref{fig:gw150914_wf} we show the posteriors for the set of intrinsic parameters, extrinsic parameters and reconstructed waveform from our analysis.
-In addition, Fig.~ \ref{fig:gw150914_orders} shows the posterior distribution on the orders $p_{H}$ and $p_{L}$ for the AR processes describing the residuals in the LIGO-Hanford and LIGO-Livingston detectors.
+	\caption{Posterior PSDs for the Hanford (orange line) and Livingstone (blue line) as inferred by our analysis.}
+	\label{fig:gw150914_psd}
+	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/noise_curves.pdf}
+\end{figure}
+
+In Figs.~\ref{fig:gw150914_intrinsic}, ~\ref{fig:gw150914_extrinsic} and \ref{fig:gw150914_waveforms} we show the posteriors for the set of intrinsic parameters, extrinsic parameters and reconstructed waveform from our analysis.
 The inferred PSDs are instead shown in Fig.~\ref{fig:gw150914_psds}.
+%In addition, Fig.~ \ref{fig:gw150914_orders} shows the posterior distribution on the orders $p_{H}$ and $p_{L}$ for the AR processes describing the residuals in the LIGO-Hanford and LIGO-Livingston detectors.
 

### 2023-07-25 paper/mesa.tex
Removed garbage from tex & made new structure

@@
 % comments command
-\newcommand{\wdp}[1]{{\textcolor{magenta}{\texttt{WDP: #1}} }}
-\newcommand{\amartini}[1]{{\textcolor{blue}{\texttt{AM: #1}} }}
-\newcommand{\sschmidt}[1]{{\textcolor{red}{\texttt{SS: #1}} }}
+\newcommand{\wdp}[1]{{\textcolor{magenta}{\texttt{walter: #1}} }}
+\newcommand{\amartini}[1]{{\textcolor{blue}{\texttt{alessandro: #1}} }}
+\newcommand{\sschmidt}[1]{{\textcolor{purple}{\texttt{stefano: #1}} }}
+\newcommand{\todo}[1]{{\textcolor{red}{\texttt{TODO: #1}} }}
 
@@
 \section{Introduction}
-
+\todo{Check again intro \& abstract}
 The study of the properties of stochastic processes is a crucial task in many fields of physics, astronomy, quantitative biology, as well as engineering and finance.
@@
 
-After a short introduction to Jaynes' MAXENT (sec.~\ref{sec:MAXENT}), we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form (sec.~\ref{sec:MESA}).
-Next, we will discuss an interesting link between Burg's method and autoregressive processes (sec.~\ref{sec:autoregr}) and in sec.~\ref{sec:forecasting} we will use such link for straightforwardly forecasting a time series.
+After a short introduction to Jaynes' MAXENT (Sec.~\ref{sec:MAXENT}), we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form (Sec.~\ref{sec:MESA}).
+Next, we will discuss an interesting link between Burg's method and autoregressive processes (Sec.~\ref{sec:autoregr}) and in Sec.~\ref{sec:forecasting} we will use such link for straightforwardly forecasting a time series.
 
@@
 
+\subsection{Whitening} \label{sec:whitening}
+\todo{Write here about our whitening procedure with the convolution}
+
 \section{Validation of the model}\label{sec:validation}
-It is now clear that MESA provides a recursive formula for computing the coefficients $a_k$ in Eq.~(\ref{eq:MESApsd}). The number $M$ of such coefficients is
-equivalent to the maximum order of the autocorrelation $\bar{r}_m$ considered.
+It is now clear that MESA provides a recursive formula for computing the coefficients $a_k$ in Eq.~(\ref{eq:MESApsd}).
+The number $M$ of such coefficients is equivalent to the maximum order of the autocorrelation $\bar{r}_m$ considered.
 In an ideal scenario, this would be equal to the number of points the autocorrelation is computed at (equivalent to the length of data considered).
@@
 The remainder of this section is devoted to an extensive study on how to make such choice.
-In Section~\ref{sec:optimizers}, we are going to define three different \textit{loss functions} to measure how well the 
+In Section~\ref{sec:optimizers}, we are going to define two different \textit{loss functions} to measure how well the 
 algorithm is able to reproduce a known PSD.
@@
 by measuring the loss function and pick, among the orders the one that yields better results.
-Second, the performance of different loss functions will be assessed by performing the spectral analysis on time series with an analytical Gaussian PSD, sec.~\ref{sec:arp_validation}.
-In a last subsection~\ref{sec:LIGO_validation}, the same analysis is performed on synthetic data generated from the analytical spectrum released by the LIGO-Virgo collaboration.
+The performance of the different losses will be assessed by answering to two questions: (i) how well the AR order is recovered and (ii) how well the measured PSD is able to whiten the input time series.
+This will be discussed Sec.~\ref{sec:arp_validation} and Sec.~\ref{sec:whitening_validation}.
 
@@
 We then need a more sophisticated method for computing the right value for $m$.
-Various loss functions to assess the algorithm performance have been proposed in literature.
 We summarise them below:
@@
 
-\item \textbf{Criterion Autoregressive Transfer function (CAT)}
-This second loss function has been proposed by Parzen and studied in detail by~\citet{bhansali1986}. 
-It is based on the assumption that the observed process is an infinite order autoregressive process 
-\begin{equation} 
-x_t = \sum_{i = 1}^{\infty}a_{i}x_{t-i} + \nu_t
-\end{equation} 
-and tries to select the order $m$ as the best finite-order approximation for the observed process.
-Being N the number of samples it has the property that $m \to \infty$ as $N \to \infty$.
-Since any real-valued stochastic process can be written as a ARMA(p,q) process (Wold theorem) i.e.
-an $AR(\infty)$ process, this is a physically significant property.
-The loss function has the functional form: 
-\begin{equation}
-    \mathcal{L}_{\rm CAT}(m) = \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m},
-\end{equation}
-and the so-chosen order is found to be asymptotically unbiased for $P_N$.
-
-\item \textbf{Optimum Bayes Decision rule} 
-The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\citep{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
-time series for the process that has to be described as an $AR(m)$ process, with $m$ to be determined.
-The OBD is obtained choosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$.
-Each
-hypothesis is uniquely determined from both the length and the values of the filter under study.
-Choosing a Likelihood of the form Eq.~\eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypotheses, Rao has shown that choosing the minimum for $-log(P(H_i \vert x(t))$, i.e.
-maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of:
-
-   \begin{align}
-        \mathcal{L}_{\rm OBD}&(m) = (N - m - 2) \log(P_m) \nonumber\\
-        &+ m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2\,.
-    \end{align}
+\item \textbf{Variance Maximum (VM)}
+\todo{WRITEME}
 \end{itemize}
@@
 
-\subsection{Choice of the loss function: Gaussian PSD} \label{sec:arp_validation}
-We test the performance of the three loss functions on a random time series generated with a known power spectral density.
-In this first experiment, we take the PSD to be a Gaussian distribution with mean $\mu = 2.5$ and standard deviation $\sigma = 0.5$, where the units are arbitrary.
-The time series are generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$ and performing an inverse Fast Fourier Transform on the frequency series.
-For this investigation, we generate a dataset of $N=1000$ time series of $3000$ points each.
-
-We then apply Burg's method choosing in turn each of the loss functions on the ensemble of simulated time series, 
-thus obtaining an ensemble of PSD estimates. Through these ensembles, we characterize statistically the performance of each loss function.
-The disagreement between the PSD $S_i(f)$ estimated from the $i$-th simulated time series and the target PSD $S(f)$ is measured via 
-the \textit{frequency-averaged relative error} $r_i$:
-\begin{equation}\label{eq:freq_error}
-	r_i = \frac{1}{N_f}\sum_{f_j=\{0, \hdots , Ny\}} \frac{|S_i(f_j) - S(f_j)|}{S(f_j)}
-\end{equation}
-where $Ny$ is the Nyquist frequency of the time series and $N_f$ is the number of the discrete frequencies the PSD is evaluated at.
-
-For each loss function, we compute the ensemble-averaged PSD (as well as the $90\%$ confidence level) and 
-the \textit{ensemble-averaged relative error}:
-\begin{equation}\label{eq:ensemble_averaged_error}
-	r(f) = \frac{1}{N} \sum_i \frac{|S_i(f) - S(f)|}{S(f)} \;.
-\end{equation}
-For each loss function, Figs.~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} show the averaged PSD, 
-as well as the ensemble-averaged relative error from Eq.~\eqref{eq:ensemble_averaged_error}.
-Furthermore, Fig.~\ref{fig:optcomparison} displays the relation between $r_i$ and the autoregressive 
-order $m_i$ chosen for each independently drawn time series.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/FPE_spectrum_estim.pdf}
-	\caption{In the top panel, we show the average spectrum for $\mathcal{L}_{\rm FPE}(m)$ with 90\% confidence regions (purple shaded region), together with the median estimation (red line) and the target PSD (black dashed line).
-In the bottom panel, we display the ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error}.
-	The average is computed with $1000$ realization of a $3000$ points long time series.}
-	\label{fig:FPEmean}
-\end{figure}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/OBD_spectrum_estim.pdf}
-	\caption{In the top panel, we show the average spectrum for $\mathcal{L}_{\rm OBD}(m)$ with 90\% confidence regions (purple shaded region), together with the median estimation (red line) and the target PSD (black dashed line).
-In the bottom panel, we display the ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error}.
-The average is computed with $1000$ realization of a $3000$ points long time series.}
-	\label{fig:OBDmean}
-\end{figure}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/CAT_spectrum_estim.pdf}
-	\caption{In the top panel, we show the average spectrum for $\mathcal{L}_{\rm CAT}(m)$ with 90\% confidence regions (purple shaded region), together with the median estimation (red line) and the target PSD (black dashed line).
-In the bottom panel, we display the ensemble-averaged relative error eq.~\eqref{eq:ensemble_averaged_error}.
-The average is computed with $1000$ realization of a $3000$ points long time series.}
-	\label{fig:CATmean}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/error_length_contour.pdf}
-	\caption{Frequency-averaged relative error $r$ eq.~\eqref{eq:freq_error} vs the length $m$ of autoregressive process, for each of the $1000$ independent realizations of the synthetic time series with a gaussian PSD.
-Different colours refers to different choices for the loss function: in red $\mathcal{L}_{\rm FPE}(m)$, in green $\mathcal{L}_{\rm OBD}(m)$ and in black $\mathcal{L}_{\rm CAT}(m)$.
-The top and right histograms show the marginal distributions.}
-	\label{fig:optcomparison}
-\end{figure}
-
-\paragraph{Final Prediction Error (FPE)}
-The results of our investigation on the performance of FPE are shown in Fig.~\ref{fig:FPEmean}.
-Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum; 
-however, we note that the reconstruction is not very accurate in the low frequency region. 
-Furthermore, the $90\%$ credible region is very small: this means that if we randomly take two of the reconstructed spectra, we expect their differences to be statistically small.
-These facts are evidence for FPE to be a reliable loss function.
-
-By looking at Fig.~\ref{fig:optcomparison} (red series), we note that the AR orders obtained with 
-FPE are clustered in a very small region. This is also a desirable property: FPE, in fact, provides a stable 
-estimation of the AR order, which does not affect much the reconstruction error.
-We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length $m$ is clustered in a region where there is no dependence of error on $m$.
-\paragraph{Optimum Bayes Decision Rule (OBD)}
-The second loss function we consider is OBD and the results are summarized in Fig.~\ref{fig:OBDmean}.
-As for the FPE case, they show a good agreement between the average over the reconstructions and the true spectrum.
-Qualitatively the same behaviour of FPE is observed: a good quality reconstruction at the intermediate and high frequencies with a narrow $90\%$ confidence level as well as a degrading performance at low frequencies.
-However, when looking at the error, the disagreement of this method is found to be larger compared to FPE: 
-in the worst case, the error can be as large as a factor of $2$ compared to FPE.
-
-The error as a function of the AR length (green series in Fig.~\ref{fig:optcomparison}) clusters in a small region, indicating the stability of the reconstructed process order $m$.
-We note that on average, OBD tends to choose smaller values of $m$ with respect to FPE.
-
-\paragraph{Criterion Autoregressive Transfer function (CAT)}
-The performance of CAT is shown in Fig.~\ref{fig:CATmean}. At a first glance, CAT differs substantially from the other two loss functions considered.
-The ensemble-average PSD matches very well the underlying ``true" PSD. This is also true even in the low frequency region, 
-where both OBD and FPE showed poor performance.
-However, the variance of the reconstructed spectrum is quite large (much larger than for FPE and OBD), and the relative error is quite high, $\sim 10\%$ and it is approximately constant over all the frequencies.
-
-The reason for this behaviour becomes clear from Fig.~\ref{fig:optcomparison} (black series): CAT does not converge 
-to any specific value for the order of the AR process. The estimated $m$ spans a large range of values, hence the large variance observed.
-Fig.~\ref{fig:optcomparison} also shows a strong dependence of the error on the estimated length of the filter.
-The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like outliers in different realizations of the time-series, but this is also responsible for an increased variance in the estimate, by introducing spurious peaks in the reproduction.
-
-When averaging the different PSD estimates, the noise in each spectrum cancels, as expected from the Gaussian nature of the AR process. 
-This implies, in a sense, that each estimate of the spectrum is independent of any other, as suggested by the huge variance in the residuals.
-This lack of stability is not a good property for the estimation of the PSD from a single realization of the time-series, however, thanks to the averaging out of errors, this estimator seems optimal in the case of repeatable experiments and ensemble-averages.
-
-\paragraph{Final remarks on the choice of the loss function}
-
-In our analysis, the FPE and OBD loss functions are found to behave similarly while CAT shows fairly different properties.
-CAT provides an accurate average spectrum over all the frequencies at the price of a large variance; in turn OBD and FPE 
-provide a poorer average in the low frequency tails, however they also display a  smaller variance, with FPE showing the lowest.
-
-The poor low-frequency reconstruction from OBD and FPE might be due to the fact that the first tends to select shorter filters, whereas long filters are required to model low frequency correlation.
-This seems to be confirmed by looking at fig.~\ref{fig:optcomparison}.
-OBD, which select the shortest filters, can provide an error as large as $300 \%$ at the extrema, as compared with $85\%$ error of FPE.
-In turn, CAT is $5\%$ accurate in these regions.
-
-However, we also note that the when inferring PSDs from a single time-series realization, FPE provide the lowest averaged error over all frequencies, while CAT can reach errors $5$ times larger (see again Fig.~\ref{fig:optcomparison}).
-Hence, while CAT is the loss function that minimizes errors in the low-frequency end of the spectrum, FPE obtains the best overall accuracy.
-
-The conclusion is clear: even if in some cases CAT is more accurate when taking the average over several realizations of the underlying process, 
-FPE guarantees that the single estimation is more faithful.
-As in any common situation we cannot perform such averaging over different realizations of the same time series, we must prefer FPE over CAT (let alone OBD, which even though qualitatively similar to FPE has worse performance).
-However, if we indeed can measure the PSD by averaging over different time series, using CAT as a loss function is the best choice.
-In this sense we retain CAT to provide the best, and most similar in spirit, alternative to the commonly employed Welch estimation method whenever ensemble-averages are needed and justified.
-
-\subsection{Choice of the loss function: LIGO Spectrum} \label{sec:LIGO_validation}
-We continue our characterization of the various loss functions considered in this work, by investigating the 
-reconstruction of a specific, known, power spectrum: that is the Advanced LIGO \citep{LIGO_2015, Virgo_2014, KAGRA_2012, KAGRA_2013} design sensitivity theoretical spectral curve \citep{LVK_future_2013, LIGO_PSDs}.
-For this analysis, we generate $N = 500$ time series of $40960$ points each, sampled with a sampling rate of $\SI{2048}{Hz}$, hence we fix the duration
-of the time-series to $\SI{20}{s}$. The chosen length is convenient to capture fairly accurately the low-frequency features of the LIGO PSD.
-We report our findings in Figs.~\ref{fig:ligospectrum}~and~\ref{fig:LigoOrderError}.
-
-\begin{figure}
-    \centering
-     \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/compare_estimates.pdf}
-      \caption{Average spectrum for the three different loss functions as compared with the ``true" PSD. The average is computed over $500$ realization of a $40960$ points long time series.}
-       \label{fig:ligospectrum}
-\end{figure}
-
-Fig.~\ref{fig:ligospectrum} shows the simulated spectrum (dashed line) and the ensemble-averaged reconstructed PSD adopting the FPE (green line), OBD (red line) and CAT (black line).
-In all cases, the spectrum is well reconstructed, but with a fairly distinct behaviour at low frequency, where CAT -- as in the Gaussian case -- better captures and resolves the distinct spectral feature at $\sim \SI{17}{Hz}$.
-In Fig.~\ref{fig:ligoPeaks}, we report the reconstructed spectra around the two peaks at $\sim \SI{17}{Hz}$ and $\sim \SI{438}{Hz}$ (left and right panel respectively).
-
-Fig.~\ref{fig:LigoOrderError} shows the distribution of recovered AR orders $m$ against the relative frequency-averaged error. The behaviour
-of the three loss functions is very similar to what found in the Gaussian PSD case (compare it with fig.~\ref{fig:optcomparison}): ODB infers the smallest orders and gives average errors around $20$\%, FPE consistently estimates orders of a few hundreds and shows the smallest errors $\sim 15$ \% while CAT does not show any preference towards any AR order and displays wildly varying errors.
-Yet again, when the PSD is averaged over multiple realization of the time, CAT is able to capture the spectrum very precisely.
-In fact, even in presence of very sharp spectral features, CAT reconstruction seems to be almost perfectly coincident with each of them.
-Hence, also the study of simulated LIGO data seems to indicate that whenever and wherever ensemble-averaged PSDs are necessary, CAT is the optimal choice of loss function.
-However, on a single time-series realization, FPE is the more robust choice.
-
-Let us summarize some key general conclusions:
-\begin{itemize}
-	\item there are no reasons to prefer OBD over CAT or FPE;
-	\item if we have one single realization for the process, we recommend the use of FPE, that would get the best resolution possible.
-In this situation, CAT would provide spurious and unreliable results, with large error;
-	\item in the case of several realizations of the same process, CAT ensemble-average properties provide very a precise spectral estimation.\end{itemize}
-
-Therefore, the choice of loss function, at least in between CAT and FPE, depends on the problem one is attempting to solve.
-\begin{figure}
-    \centering
-    \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/error_length_contour.pdf}
-    \caption{For each of the $500$ independent realization of the time series, we plot the relative error $r$ (as in eq.~\eqref{eq:freq_error}) against the length $m$ of autoregressive process.
-The time series are randomly drawn with a the analytical LIGO PSD in fig.`\ref{fig:ligospectrum}.
-Different colors refers to different choices for the loss function.
-Histograms for the distribution od the individual quantities are also represented.}
-    \label{fig:LigoOrderError}
-\end{figure}
-
-\begin{figure}[t]
-        \includegraphics[width = \linewidth]{Images/optimisers_comparison/ligo/compare_estimates_peaks.pdf}
-        \caption{Details of peaks of the spectrum and their reconstruction with every optimizer}
-        \label{fig:ligoPeaks}
-\end{figure}
-
-\subsection{How well is the AR order recovered?}
+\subsection{How well is the AR order recovered?}\label{sec:arp_validation}
+\todo{WRITEME}
 
@@
 We now address the issue of how well the AR order (i.e. the number of $a_k$ coefficients employed) is estimated by each loss function.
-This is useful to further characterize the properties of the different loss functions.
-Furthermore, as MESA is supposed to model data as an $AR(p)$ process, it is interesting to quantify its precision on some true autoregressive (and stationary) time series.
 
-We generate $100$ autoregressive processes $AR(p)$ with a random value of $p$, drawn such that $log(p) \sim\mathcal{U}_{[log(2), log(5000)]}$.
-Each coefficient $a_k$ is assigned according to a Dirichlet distribution $\mathrm{Dir}([1,\hdots, 1])$.
-The sign of $a_k$ is assigned randomly according to a binomial distribution.
-We report the result of this investigation in Fig.~\ref{fig:p_vs_ptrue}.
-
-It can be seen that OBD and FPE loss functions show very similar behaviour.
-They are {\it very} reliable in capturing the correct AR order up to a certain threshold ($p\sim 100$ for OBD and $p\sim 200$ for FPE).
-For any AR process of order higher than the threshold, the optimization underestimates badly the actual value of $p$.
-For this reason, FPE and OBD seem reliable only for relatively simple AR processes.
-For more complicated processes, they seem to ``underfit'' the problem (i.e.
-they output a model simpler that those required).
-
-On the other hand, CAT shows a different behaviour.
-The selected AR order $p$ is always close to the maximum possible value, regardless the actual value $p_{true}$ of the underlying process.
-This produces always a model that is more complex than those obtained with FPE and OBD and this can explain the origin of the high variance in the estimation observed in the experiments described in the previous sections.
-Of course, CAT does not perform a good job in reconstructing the AR process.
-However, for high order AR processes, the error introduced by CAT is more tolerable than the error introduced by FPE and OBD.
-This is a good reason to prefer CAT in these situations (see also Sec.~\ref{sec:temp_timeseries} for another example of this effect).
+\subsection{How well can MESA whiten the data?}\label{sec:whitening_validation}
+\todo{WRITEME}
 
 \section{Comparison with Welch method}\label{sec:Welch_comparison}
+\todo{Check again the section}
 We perform a {\it qualitative} comparison between the performance of the MESA and of the standard Welch algorithm.
@@
 We set the total observation time $T = 1, 5, 10, 100, 1000 \SI{}{s}$
-For the MESA algorithm, we choose the FPE loss function. For the Welch algorithm, we employ a Tukey window with a parameter $\alpha$ 
+For the MESA algorithm, we choose the ???? loss function. For the Welch algorithm, we employ a Tukey window with a parameter $\alpha$ 
 equal to 0.4, an overlap fraction of $1/2$ for the segments and a length of 
@@
 
-\section{Applications} \label{sec:applications}
-%\subsection{Temperature Time Series} \label{sec:temp_timeseries}
-%As a further example of the breadth of applicability of MESA, we applied our implementation to atmospheric temperature time series. 
-%The reason for this choice is twofold: i) atmospheric temperature time series present a variety of overlapping periodicities most of which are 
-%known; ii) it provides a stress test for the time series forecast analysis.  As dataset, we  used the historical reanalysis data 
-%from the ``ERA5-Land hourly data from 1981 to present" dataset, downloaded from the Climate Data Store~\citep{ERA5_dataset}.
-%The data consist in temperatures taken at coordinates $\textrm{N}\;\ang{45;5;} \;\; \textrm{E}\;\ang{9;1;}}$, corresponding about to the city of Milan, Italy.
-The temperatures are given on an hourly cadence for almost 31 years from 31st December 1989 to 30th November 2020.
-%
-%Fig.~\ref{fig:temp_spectrum}, shows the MESA spectrum inferred from the data. As a comparison, we adopt both the FPE and the CAT loss functions.
-%In agreement with what we found in previous sections, the FPE PSD is more regular compared to the one from CAT. Both spectra show a
-%peak at the frequency $f_D = \SI{1}{day^{-1}}$ corresponding to one day, corresponding to the day-night cycle: this is expected.
-%Higher order harmonics (corresponding to integer multiples of $f_D$) are also visible up to the Nyquist frequency ($f_D = \SI{0.5}{hour^{-1}}$): they correspond to signals that preserve the 1 day period while, at the same time, capturing the complex variability of the daily temperature cycle throughout the year.
-%
-%Looking at low frequencies, FPE does not capture the yearly variability at $f_{yr} = \SI{1}{yr^{-1}}$.
-On the other hand, the peak is captured well by the CAT loss function.
-In analogy as what observed for the peak at $f_D$, in the CAT spectrum we observe a higher order harmonic at a frequency $2 \cdot f_{yr}$.
-As in the previous case, this is required to better model the temperature variation in the year.
-%
-%The failure of FPE in capturing the low frequency trend can be understood by looking at Fig.~\ref{fig:p_vs_ptrue}.
-It is shown that CAT systematically select very long filters (i.e.
-large $p_{CAT}$ values for the autoregressive orders), whereas FPE tends to select smaller values for the order $p_{FPE}$ of the autoregressive process, often underestimating the actual value.
-%In this example, we have $p_{CAT} \sim  36500 \sim  \SI{4}{yr}$, whereas $p_{FPE} \sim  1300 \sim  \SI{2}{months}$.
-%The autoregressive process selected by FPE, hence cannot produce a meaningful prediction on the timescale longer than a few weeks and for this reason it is unable to capture the low frequency peak at $f_{yr}$.
-In other words, the model is too simple to model \textit{both} the behaviour at high and low frequencies (separated by approximately 3 orders of magnitude).
-%On the other hand, the model chosen by CAT is much more complex and is able to model also the high frequency behaviour, at the expense of making a more noisy estimation.
-%The ``actual" length of the autoregressive process should be closer to the choice of CAT.
-%
-%We now assess the accuracy of the forecasting of new observations of the time series. 
-%Based on actual data, we try to predict future values as described in sec.~\ref{sec:forecasting} (of course the MESA is performed with CAT loss function).
-We produce $N =100$ independent predictions and we compute the median as well as the $90\%$ confidence interval.
-This is compared with the actual measured temperature values.
-The predictions span a two years range of time.
-We report our results in Fig.~\ref{fig:temp_spectrum}.
-%
-%We note the observed difference is always well included in the $90\%$ confidence interval: the forecasting predictions seem reliable.
-%On the other hand, the confidence interval is pretty large (almost as large as $\SI{15}{K}$), thus making ``easy" for the actual data to fit the predictions.
-%Indeed, the prediction model, while suitable for spectral estimation, is nothing more than a linear regression (plus noise term).
-Such a simple model hardly catches the complex trend in the variability of the temperature daily trend during a year.
-For more precise predictions, probably one should consider nonlinear regression model, tapping into the wealth of non linear predictors offered by the field of Deep Learning.
-%
-%\begin{figure}
-%	\caption{Spectrum for the historical temperature time series.
-The two lines refers to different loss functions (CAT and FPE); the inset shows the harmonics of the fundamental frequency of a day.
-Two vertical lines are drawn in correspondence to the frequency $f_D$ of a day and $f_{yr}$ of a year.}
-%	\label{fig:temp_spectrum}
-%	\includegraphics{Images/climate_plots/temp_spectrum.pdf}
-%\end{figure}
-%
-%\begin{figure}
-%	\caption{Difference between the historical temperature $T$ and the forecast temperature $T_{forecast}$ for two years of data.
-The red shadow denotes the $90\%$ confidence interval of the predictions.
-The model is trained with CAT loss function.
-The discrepancy between predicted and actual values is always everywhere included in the $90\%$ confidence interval.}
-%	\label{fig:temp_forecast}
-%	\includegraphics{Images/climate_plots/forecast_accuracy.pdf}
-%\end{figure}
-
-\subsection{Marginalising over the noise PSD in GW parameter estimation: the case for GW150914}
+\section{An application: marginalising over the noise PSD in GW parameter estimation} \label{sec:applications}
+%\subsection{Marginalising over the noise PSD in GW parameter estimation: the case for GW150914}
 Since the order of the AR process is not yet determined by the theory, up to this point, we opted for an in-depth investigation of several proposals in the literature and found that different loss functions are required for different situations, with the FPE loss function being the most indicated to deal with gravitational wave data.
@@
 %a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
-%\sschmidt{There might be a misunderstanding: I guess the referee is talking about taking into account the PSD errors, while we are talking about dropping the stationariety assumption...
+%\todo{There might be a misunderstanding: I guess the referee is talking about taking into account the PSD errors, while we are talking about dropping the stationariety assumption...
 The misleading word is `marginalize', I think}
@@
 
-\subsection{Forecasting the LIGO strain} \label{sec:LIGO_forecasting}
-As a last example of an application of MESA, we return to the study of the strain produced by the LIGO-Virgo interferometers 
-and we forecast the future observations.
-A seminal work on the application of autoregressive models to Virgo data is presented in \citet{Cuoco_2001}, where an $AR(p)$ model is trained on the data for the purpose of estimating the PSD and to create a whitener filter.
-
-We focus on the public data released by the LIGO/Virgo collaboration~\citep{Abbott_2021}.
-We reconstruct the PSD both assuming the CAT and FPE loss functions on $\SI{1000}{s}$ of data from the Livingston observatory
-starting from GPS time $1164603392$. The data are sampled at $\SI{4096}{Hz}$.
-We then forecast the following $\SI{100}{s}$ of observations with the model optimized with CAT\footnote{
-The reason for this choice is that CAT, despite showing higher variance and a worse PSD estimation accuracy, estimates sistematically high autoregressive orders.
-We believe that this feature provides a more reliable forecast, as more points are used for predictions.
-In practice, we found that CAT and FPE behave very similarly.}.
-The results are shown in fig.~\ref{fig:LIGO_forecast}.
-The prediction is always in the $90\%$ confidence interval. The confidence interval, however, is very broad and the prediction is not very accurate (being the same order of magnitude of the strain).
-By looking at the standard deviation $\sigma$ of the predictions (bottom panel in Fig.~\ref{fig:LIGO_forecast}), we note that it increases very quickly in the first $\SI{0.5}{s}$.
-The order of the autoregressive process selected is $p_{CAT} = 57766 \sim \SI{14}{s}$, which is much larger than the region in which $\sigma$ is small $\sim \SI{0.2}{s}$.
-The implication is that the series is very difficult to predict: the knowledge of past observations is not very helpful to predict future observations.
-Although FPE selects an autoregressive order $p_{FPE} = 29924 \sim \SI{7}{s}$ smaller than CAT,  the forecasted time series behaves very similarly.
-
-Despite poor predictions on long timescales, the AR process obtained with MESA still can be useful on short timescales.
-Indeed, a precise prediction of the strain time series can be beneficial in the detection of anomalies in the data and, eventually, their removal.
-Indeed loud anomalies in the data, called {\it glitches}, pose a major challenge to the ability of detecting signals and intensive work has been done to mitigate the disruption in sensitivity they cause \citep{Nuttall_2015, Noise_char_2016, Zevin_2017} and to develop effective subraction techniques \citep{Pankow_2018, zackay2019detecting}.
-The predictions can form an expected baseline for the strain; any anomaly (i.e. a glitch or even a transient of physical origin) can show up as a large departure from expected trend.
-Moreover, if a glitch is detected, its shape can be estimated (as well as the confidence level) by subtracting the expected signal with the actual signal.
-This can work: a typical glitch can last as long as $\SI{0.2}{s}$, close to the time scale over which the forecasting is reliable. Future works will explore this exicting opportunity.
-
-As noted above, for pushing further the prediction performance from our simple linear predictor, more sophisticated forecasting methods are available in the Machine Learning literature (see e.g.
-\citep{LSTM_article, oord2016wavenet, lea2016tcn}) and they can be trained for precision forecasting.
-However, a Maximum-Entropy trained AR process can be a simple baseline for comparing such more advanced methods and might suffice for many purposes.
-This opens a promising path in GW data analysis and other fields of physics might also take advantage from this.
-
-\begin{figure}
-	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
-The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
-The confidence interval is computed on  $100$ independent random realization of the predictions.
-	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
-	}
-	\label{fig:LIGO_forecast}
-	\includegraphics{Images/forecast_LIGO/forecast_accuracy.pdf}
-\end{figure}
-
 \section{Final remarks and future prospects} \label{sec:conclusion}
+\todo{The conclusions will need to be rearrenged completely}
 We presented a case study of the application of Maximum Entropy principle to the realm of spectral estimation.
@@
 
-%Several approaches have been developed to try to study the validity of the assumption or even to drop it \cite{R_ver_2010, R_ver_2011, Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave, Edwards_2020, Edwards_2015} \sschmidt{All these citations seems a bit too much for me...}.
+%Several approaches have been developed to try to study the validity of the assumption or even to drop it \cite{R_ver_2010, R_ver_2011, Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave, Edwards_2020, Edwards_2015}.
 %Besides them, MESA can offer an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on $d_t-x_t$.

### 2023-08-10 paper/mesa.tex

@@
 \subsection{Whitening} \label{sec:whitening}
-\todo{Write here about our whitening procedure with the convolution}
+\todo{Do we need this paragraph?}
+
+The theory of the AR processes can be also applied to the problem of whitening a time series.
+Given a time series, $x_t$, the whitening operation produces another timeseries $x^W_t$ such that:
+%
+\begin{equation}\label{eq:whitening_definition}
+	x^W_t = \mathcal{F}^{-1}\left[ \frac{\tilde{x}(f)}{\sqrt{S(f)}} \right]
+\end{equation}
+%
+where $\mathcal{F}^{-1}$ denotes the inverse Fourier transform of a frequency series.
+If $x_t$ is a realization of gaussian noise (see Eq.~\eqref{eq:Whittle}) with PSD $S(F)$, the whitened timeseries $x^W_t$ is just white noise (i.e. uncorrelated samples from a normal gaussian).
+
+From Eq.~\eqref{eq:AR_p}, remembering that $b_i = - a_i$, it's straightforward to derive an expression for the whitened timeseries $x^W_t$:
+\begin{equation}
+	x^W_t = \frac{1}{\sqrt{P_N}} \sum_{i=0}^p a_i x_{t-i}
+\end{equation}
+This amounts to a convolution of the timeseries  $x_t$ with the kernel $(1, a_1, \hdots, a_p)$, plus a variance rescaling.
+Performing a convolution is an appealing alternative to evaluating Eq.~\eqref{eq:whitening_definition} directly.
 

### 2023-08-11 paper/mesa.tex
Added PSD comparison plot and writtern comparison on paper

@@
 \end{pmatrix}
-
 }.
@@
 with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.
-Minimizing FPE is equivalent to minimizing the quantity: 
+Asymptoticall,y minimizing FPE is equivalent to minimizing the quantity: 
 \begin{equation}
@@
 \item \textbf{Variance Maximum (VM)}
-\todo{WRITEME}
+\todo{WRITEME - Page250 book}
+This second criterion was introduced by \todo{[cite Jekins and Watts]} and is based on a similar assumptions to FPE. The quantity to be minimised is
+\begin{equation}
+VM(m) = \frac{1}{N - 2m}\sum_{t=m}^N\left(x_t - \sum_{i=1}^m a_i x_{t-i} \right)^{2}
+\end{equation}
+In this case, the normalising factor takes into account the k degrees of freedom necessary to estimate the forward prediction error filter $a_k$.\\
+The package implementation of VM loss function takes advantage of a recursive re-writing of the above formula, as in equation \todo{[EQ] of [CITE E. CUOCO]}  \\ \\
+
+Several other criteria are available in the literature \todo{[cite CAT, OBD etc?]} and some are implemented in the memspectrum package.
+We don't report them in this paper since they didn't show any additional merit with respect to the aforementioned loss functions \\
 \end{itemize}
@@
 The following sections will be devoted to the study of the statistical properties of the loss functions introduced above: we need to understand which choice provides the best quality in the reproduction of some known power spectral densities.
-
-\subsection{How well is the AR order recovered?}\label{sec:arp_validation}
-\todo{WRITEME}
+In the following paragraph, we will discuss three different comparison (one qualitative and two quantitative) of the two proposed loss functions, o
+\subsection{How well the PSD is reconstructed?}\label{sec:psd_validation}
 
@@
 	\centering
-	\includegraphics[width = \linewidth]{Images/arp_errors/scatter_deltap_ptrue}
+	\includegraphics[width = \linewidth]{Images/Plots/comparison_loss}
+	\label{fig:psd_comparison}
+	\caption{Comparison of the ensemble average PSD estimate}
+\end{figure}
+In our initial qualitative comparison, depicted in Figure \ref{fig:psd_comparison}, we juxtapose the reconstruction of a known a-priori Power Spectral Density (PSD) with those obtained using the two distinct loss functions.
+The red line in the plot represents our chosen reference PSD, derived from the estimated PSD for the GW150914 event.\\
+To conduct this analysis, we generated 1000 noise time-series whose power spectral densities matches the reference PSD by construction \todoc{[insert citation]}.
+The sampling rate and observation time were fixed at $dt = 2048 Hz$ and $T = 5s$, respectively.
+For each noise realization, we employed both the FPE and the VM loss functions to estimate the PSD.
+Ultimately, we compared the reference PSD against the ensemble average of these two estimation methods.\\
+The FPE-derived estimate, represented by the dashed line, effectively identifies and reconstructs peaks across both high and low frequency ranges with commendable accuracy.
+However, as illustrated in the inset plots, FPE struggles when confronted with structured peaks—those containing subordinate modes.
+In such cases, FPE accurately captures the primary mode but overlooks the subsidiary peaks.
+\\
+On the other hand, the VM estimate, depicted as a dot-dashed line, excels in reconstructing both dominant and subordinate modes with remarkable precision.
+VM appears to prioritize comprehensive mode reconstruction, while FPE emphasize an accurate reconstruction of major modes while potentially neglecting more intricate sub-peaks.
+
+\subsection{How well is the AR order recovered?}\label{sec:arp_validation}
+
+\begin{figure}
+	\centering
+	\includegraphics[width = \linewidth]{Images/Plots/ar_order}
 	\caption{Reconstructed value for the autoregressive order plotted against the true value of the autoregressive order.
@@
 
-We now address the issue of how well the AR order (i.e. the number of $a_k$ coefficients employed) is estimated by each loss function.
+Moving to our second comparison, we now focus on another crucial aspect: how accurately each loss function estimates the Autoregressive (AR) order, which represents the number of employed $a_k$ coefficients.
+\\
+Here, the memspectrum package proves quite useful.
+It allows us to assign a specific order to the reconstructed autoregressive filter and use the resulting coefficients to forecast time series.
+With these tools in hand, we embarked on an extensive task: generating various time series, each with a different autoregressive order ranging from m = 0 to m = 4000.\\
+To ensure reliability, we created 30 distinct time series for each autoregressive order.
+This approach lets us compute both the mean and variance, giving us insights into the accuracy of each loss function's order estimation.
+This analysis provides valuable information about how well each method performs in estimating the Autoregressive order across a broad spectrum of scenarios.
+\\
+The results are reported in Figure \ref{fig:p_vs_ptrue}.
+The injected autoregressive order's true value is depicted by the red line.
+The estimations yielded by the two loss functions are illustrated alongside, accompanied by error bars indicating one standard deviation.
+\\
+The plot reveals two distinct regions: one with "short" autoregressive orders (m = 0 to around m = 1600) and another with "long" autoregressive orders (starting from m = 1600).\\ 
+In the first region, both loss functions provide comparable results that generally match the actual autoregressive order.
+FPE performs particularly well, offering estimates close to the injected order and with minimal error bars.
+VM performs slightly worse than FPE in this range, overestimating complexity and showing larger error bars.\\
+Moving into the second region (m > 1600), a shift in performance becomes apparent.
+FPE's estimates tend to stabilize at a certain autoregressive value.
+However, as the injected model becomes more complex beyond this point, FPE's accuracy in recovering the true order diminishes, and its variance increases.
+In contrast, VM performs better in this range, closely following the actual behavior and consistently recovering the true order within one standard deviation.
+To conclude, VM appears to prioritize complexity in its approach. In contrast, FPE seems to lean toward synthesis, emphasizing accurate reconstruction of not too complex models.
 

### 2023-08-12 paper/mesa.tex
Written comparison section

@@
 \subsection{Whitening} \label{sec:whitening}
-\todo{Do we need this paragraph?}
+\todo{Do we need this paragraph? -> YES! (Ale)}
 
@@
 It allows us to assign a specific order to the reconstructed autoregressive filter and use the resulting coefficients to forecast time series.
-With these tools in hand, we embarked on an extensive task: generating various time series, each with a different autoregressive order ranging from m = 0 to m = 4000.\\
+With these tools in hand, we generated various time series, each with a different autoregressive order ranging from m = 0 to m = 4000.\\
 To ensure reliability, we created 30 distinct time series for each autoregressive order.
@@
 \todo{WRITEME}
+\begin{figure}
+	\includegraphics[width = \linewidth]{Images/Plots/P_values_comparison}
+	\caption{Histogram of the P-values obtained with a Kolmogorv Smirnov test on the whitened time series, against a univariate, 0 mean normal distirbuition}
+	\label{fig:whitening}
+\end{figure}
+In section \ref{sec:whitening}, we showed how autoregressive coefficients and noise variance estimate $P$ can jointly be used to create a whitening filter.
+To complete our investigation, we compare how well these whitening filters work when obtained from the two different loss functions we've been studying.
+\\
+For this test, we employed the same set of time-series data as described in paragraph \ref{sec:psd_validation}.
+Each time series underwent the whitening process using the autoregressive filter derived from its corresponding loss function.
+We then evaluated the resulting whitened time series against a zero-mean, univariate normal distribution using the Kolmogorov-Smirnov test.
+\\
+The results are reported as an histogram for the obtained p-values in Figure \ref{fig:whitening} together with the chosen critical region of $p < 0.1$, representing a $90\%$  confidence level.
+In this region, there is no statistical difference between the two.
+Infact, the total number of counts in this bin are respectively $c_{VM} = 47 \pm 7$ and $c_{FPE} = 42 \pm 6$, affirming the absence of a pronounced discrepancy between the two.
+In essence, this final examination underscores a shared proficiency in whitening between the two loss functions, showing that very long filters are not needed to obtain a comparlable result in whitening.
+Both methods showcase comparable results for whitening scopes.\\
+
+From our previous discussions, it's evident that both FPE and VM have their own strengths, and the choice between them greatly depends on the specific analysis requirements.
+In our analysis, VM tends to provide more accurate PSD estimates and often results in longer autoregressive filters.
+However, in cases where the underlying model is simple, there is a risk of VM overestimating complexity and generating patterns that don't truly reflect the data.\\
+On the other hand, FPE is a good option for reconstructing processes without introducing unnecessary complexity.
+However, it might underestimate the complexity of the data, particularly in scenarios involving secondary peaks or in the low-frequency region.\\
+
+Lastly, it's worth noting that FPE holds the advantage of lower numerical complexity due to its straightforward calculations involving simple arithmetic.
+In contrast, VM requires more complex computations, dealing with arrays that might be very long depending on the analysed data.
 

### 2023-09-13 dev/autoregressive_plots.py
Added 0-phase whitening and dev codes for plots

@@ -6,10 +6,10 @@ import sys
 sys.path.insert(0,'..')
 import memspectrum
 
-
+m = memspectrum.MESA()
 N_process = 100 #10 simulated autoregressive processes
 N_data = 50000 #number of timesteps for each autoregressive process
-p_max = 5000
+p_max = 7000 #has to be < 2N_data / log(2N_data) (log is natural log)
 
 save_dir = 'arp_data_dirichlet/'
 plot_dir = '../paper/Images/arp_errors/'
@@ -21,17 +21,18 @@ plot = True
 true_arp_list = []
 rec_MESA_list = []
 data_list = []
-loss_functions = ['FPE', 'CAT', 'OBD']
+loss_functions = ['FPE', 'VM']
 
 for i in range(N_process):
 
 	if load:
 		true_arp_list.append(memspectrum.MESA(save_dir+'arp_model_{}'.format(i)))
 		#data_list.append(np.loadtxt(save_dir+'data_{}'.format(i)))
-		rec_MESA_list.append( (memspectrum.MESA(save_dir+'MESA_model_{}_{}'.format(loss_functions[0],i)), 
+        
+		rec_MESA_list.append( \
+           (memspectrum.MESA(save_dir+'MESA_model_{}_{}'.format(loss_functions[0],i)), 
 			memspectrum.MESA(save_dir+'MESA_model_{}_{}'.format(loss_functions[1],i)),
-			memspectrum.MESA(save_dir+'MESA_model_{}_{}'.format(loss_functions[2],i)) )
-			)
+            ))
 	else:
 		true_arp_list.append(memspectrum.MESA())
 			#getting p
@@ -79,14 +80,14 @@ for i in range(N_process):
 	#doing plots
 if plot:
 	N_p_plot = 1 #N of series to be shown in the err vs p plot
-	colors = {'FPE': 'r', 'CAT': 'k', 'OBD': 'green'}
+	colors = {'FPE': 'k', 'VM': 'dimgrey'}
 	fig_p = init_plotting()
 	ax_p = fig_p.gca()
-	
+
 	fig_scatter = init_plotting()
 	ax_scatter = fig_scatter.gca()
-	
-		#computing erorrs
+	markers = {'FPE': 'x', 'VM': 'o'}
+		#computing errors
 	p_rec = np.zeros((N_process,3))
 	p_true = np.zeros((N_process,))
[…]

### 2023-09-13 dev/optimisers_comparison.py

@@ -7,6 +7,7 @@ Created on Fri Feb 19 10:17:55 2021
 """
 import sys
 sys.path.insert(0, '..')
+sys.path.insert(0, '../memspectrum')
 from memspectrum import MESA 
 from GenerateTimeSeries import generate_data 
 import numpy as np
@@ -15,227 +16,140 @@ import matplotlib.pyplot as plt
 from style_sheet import init_plotting
 from scipy.interpolate import interp1d 
 import matplotlib.ticker as ticker
+import tqdm
+import time
+import os 
 
-def relative_error(real, estimate): 
-    return np.abs(real - estimate) / real 
 
-PSD = 'ligo' #normal or ligo 
+#Decide if loading (load = True) or simulating (load = False) 
+load_noise = False 
+load_estimates = True 
+simulate_noise = False
 save = False
-savedir = '../paper/Images/optimisers_comparison/' + PSD + '/'
-simulate_data = False 
-methods = ['FPE', 'CAT', 'OBD']
-M = MESA() 
-init_plotting()
-plt.close() 
-colors = {'FPE': 'r', 'CAT': 'k', 'OBD': 'green'}
+plot = True 
+save_plots = False 
 
-print("Dealing with ",PSD)
-
-#Initialize single spectrum dictionaries 
-spectra, optimisers, orders, errors = {}, {}, {}, {} 
-for method in methods: 
-    spectra[method], optimisers[method], orders[method], errors[method] = [], [], [], []
-
-#ensemble dictionaries 
-median, p5, p95, ensemble_error = {}, {}, {}, {}
-
-#Generating noise 
-if PSD.lower() == 'normal': 
-        number_of_simulations, dt, number_of_points = 1000, 1./10, 3000 
-        f = np.linspace(0, .5 / dt, number_of_points) 
-        spectrum = scipy.stats.norm.pdf(f, 2.5, .5)
-
-elif PSD.lower() == 'ligo':  
-        number_of_simulations, dt, number_of_points = 500, 1./2048, 40960
-        f, spectrum = np.loadtxt('LIGO-P1200087-v18-AdV_DESIGN_psd.dat', unpack = True)
-        
-T = number_of_points * dt 
-
-if simulate_data: 
-    for i in range(number_of_simulations):
-        if i % 5 ==0: print("Simulating time series {}/{}".format(i, number_of_simulations))
-        time, time_series, frequency, frequency_series, psd =\
[…]

### 2023-09-13 memspectrum/memspectrum.py

@@ -1053,7 +1053,7 @@ class MESA(object):
 		else:
 			c = np.correlate(self.a_k, self.a_k, 'full')
 			c /= c.max() 
-			white_data = np.convolve(c, data, 'same')
+			white_data = np.convolve(c, data, 'same')/np.sqrt(self.P)
             
 		if trim is None: trim = self.get_p()
 		if trim: white_data = white_data[trim:-trim]

### 2023-09-13 paper/mesa.tex
Changed plots in paper

@@
 	\centering
-	\includegraphics[width = \linewidth]{Images/Plots/comparison_loss}
+	\includegraphics[width = \linewidth]{Images/Plots/comparison_loss2}
 	\label{fig:psd_comparison}
@@
 	\centering
-	\includegraphics[width = \linewidth]{Images/Plots/ar_order}
+	\includegraphics[width = \linewidth]{Images/Plots/ar_order2}
 	\caption{Reconstructed value for the autoregressive order plotted against the true value of the autoregressive order.
-	The reconstructed autoregressive orders are computed from a time series randomly drawn with an $AR(p)$ model, with the three different loss functions under investigation.
+	The reconstructed autoregressive orders are computed from a time series randomly drawn with an $AR(p)$ model, with the two different loss functions under investigation.
 	}

### 2023-09-25 paper/mesa.tex
Completed loss functions comparison

@@
 \subsection{Whitening} \label{sec:whitening}
-\todo{Do we need this paragraph? -> YES! (Ale)}
 
@@
 \item \textbf{Variance Maximum (VM)}
-\todo{WRITEME - Page250 book}
-This second criterion was introduced by \todo{[cite Jekins and Watts]} and is based on a similar assumptions to FPE. The quantity to be minimised is
+This second criterion \cite{kay1988modern} is based on a similar assumptions to FPE.
+It minimises the actual value of the least squares (insted of relying to asymptotical behaviour).
+t the normalising factor takes into account the k degrees of freedom necessary to estimate the forward prediction error filter $a_k$.\\ The quantity to be minimised is
 \begin{equation}
@@
 \end{equation}
-In this case, the normalising factor takes into account the k degrees of freedom necessary to estimate the forward prediction error filter $a_k$.\\
-The package implementation of VM loss function takes advantage of a recursive re-writing of the above formula, as in equation \todo{[EQ] of [CITE E. CUOCO]}  \\ \\
+The package implementation of VM loss function takes advantage of a recursive re-writing of the above formula, as in equations (27) and (28) of \citet{Cuoco_2001}.  \\ \\
 
-Several other criteria are available in the literature \todo{[cite CAT, OBD etc?]} and some are implemented in the memspectrum package.
+Several other criteria are available in the literature \citet{doi:10.1029/WR018i004p01097} \citet{10.1214/aos/1176349858} and some are implemented in the memspectrum package.
 We don't report them in this paper since they didn't show any additional merit with respect to the aforementioned loss functions \\
@@
 The red line in the plot represents our chosen reference PSD, derived from the estimated PSD for the GW150914 event.\\
-To conduct this analysis, we generated 1000 noise time-series whose power spectral densities matches the reference PSD by construction \todoc{[insert citation]}.
+To conduct this analysis, we generated 1000 noise time-series whose power spectral densities matches the reference PSD by construction \citet{noiseGen}.
 The sampling rate and observation time were fixed at $dt = 2048 Hz$ and $T = 5s$, respectively.
@@
 \subsection{How well can MESA whiten the data?}\label{sec:whitening_validation}
-\todo{WRITEME}
 \begin{figure}
@@
 We set the total observation time $T = 1, 5, 10, 100, 1000 \SI{}{s}$
-For the MESA algorithm, we choose the ???? loss function. For the Welch algorithm, we employ a Tukey window with a parameter $\alpha$ 
+For the MESA algorithm, we choose the VM loss function. For the Welch algorithm, we employ a Tukey window with a parameter $\alpha$ 
 equal to 0.4, an overlap fraction of $1/2$ for the segments and a length of 

### 2023-10-11 paper/mesa.tex
Inserted footnote on Burg and FastBurg

@@
 The vector obtained as $(1, a_1, \dots, a_N)$ is also known as the \textit{prediction error filter}.
-The coefficients $a_s (s > 0)$, together with an overall multiplicative scale factor $P_N$, are to be determined by an iterative process (called Burg's algorithm).
+The coefficients $a_s (s > 0)$, together with an overall multiplicative scale factor $P_N$, are to be determined by an iterative process (called Burg's algorithm)\footnote{There are two implementations of Burg's algorithm available in the memspectrum package: 'Standard' and 'Fast.' 'Standard' is slower but more stable for solving Levinson recursion, while 'Fast' is quicker but less stable.
+On simulated stationary data, both versions typically yield similar results.
+However, when working with real gravitational waves data, we found that the 'Fast' implementation introduces noise into the PSD estimate without improving accuracy.
+Therefore, it's advisable to use the 'Standard' implementation whenever possible.
+Keep in mind that if instability arises, a message will appear, however, it doesn't guarantee identical results between the two configurations if the message is not displayed.}.
+\\
+
 The number $N$ of such coefficients is a choice that shall be made by the user and indeed it is the only hyperparameter that needs to be tuned.

### 2023-10-13 paper/mesa.tex
Adding Fast vs Burg in text

--- 2023-09-25 paper/mesa.tex
@@
 The vector obtained as $(1, a_1, \dots, a_N)$ is also known as the \textit{prediction error filter}.
-The coefficients $a_s (s > 0)$, together with an overall multiplicative scale factor $P_N$, are to be determined by an iterative process (called Burg's algorithm).
+The coefficients $a_s (s > 0)$, together with an overall multiplicative scale factor $P_N$, are to be determined by an iterative process (called Burg's algorithm)
+At least, two implementations of Burg's algorithm are available in the literature, labeled as `Standard' \cite{} and `Fast' \cite{} in the \texttt{memspectrum} package.
+The `Standard' method is slower but more stable, while `Fast' trades stability for speed.
+On simulated stationary data, both versions typically yield similar results, while our tests with real gravitational waves data seems to indicate that the `Fast' implementation introduces noise into the PSD estimate with no accuracy improvement.\footnote{For this reason, it is advisable to use the `Standard' implementation whenever possible.
+In most case of numerical instability in the `Fast' method, \texttt{memspectrum} will send a warning to user.}.
+
 The number $N$ of such coefficients is a choice that shall be made by the user and indeed it is the only hyperparameter that needs to be tuned.

### 2023-12-20 paper/mesa.tex
edits to the text

@@
 
-\section{An application: marginalising over the noise PSD in GW parameter estimation} \label{sec:applications}
-%\subsection{Marginalising over the noise PSD in GW parameter estimation: the case for GW150914}
-Since the order of the AR process is not yet determined by the theory, up to this point, we opted for an in-depth investigation of several proposals in the literature and found that different loss functions are required for different situations, with the FPE loss function being the most indicated to deal with gravitational wave data.
+\section{Marginalisation over the noise distribution: application to GW parameter estimation} \label{sec:applications}
+
+%Since the order of the AR process is not yet determined by the theory, an alternative route to choosing a loss function specific for each situation is to \emph{marginalise} over the order of the AR process whilst exploring the parameter space the signal hypothesis.
+
+Define the data hypothesis $D$ as the statement that the data $D = S + N$ with $S$ and $N$ some deterministic signal and some noise hypotheses.
+Typically, in this formulation one is choosing both a functional form for the signal of interest $S\equiv ``h(t;\theta)''$ and some parametric form $f(t)$ for the noise distribution $N\equiv ``n(t) \sim f(t)''$.
+Some well established math then leads to the usual Bayesian framework for parameter estimation, see \cite{lalinference} for an application to gravitational wave physics.
+This procedure is very robust as long as the choice of noise distribution is indeed representative of the underlying process.
+Let us relax the $N$ hypothesis by defining a \emph{residuals} hypothesis $R$ as $R = D - S$.
+This might seem a very trivial statement, but it has a non trivial application: given  $d(t) = h(t;\theta) + n(t)$ where $h(t;\theta)$ is our signal model, defined by a set of parameters $\theta$,  the residuals $r(t) \equiv d(t) - h(t;\theta)$.
+Formally, no reference to the noise process is present anymore.
+Under MAXENT, we can model $r(t) \sim \mathrm{AR}(k)$ with $k$ the \emph{unknown} order of the process to be inferred from the residuals, either via one of the aforementioned loss functions or even by marginalising over it while exploring the signal space.
+Moreover, we can \emph{always} write $p(r(t)|N\,I)$ as in Eq.~(\ref{eq:Whittle}) once we know $k$, with the PSD given in Eq.~(\ref{eq:ARspectrum}), whatever the noise process actually is.
+In other words, we care only about maximising the information entropy in the distribution of the residuals.
+
+%Since the order of the AR process is not yet determined by the theory, up to this point, we opted for an in-depth investigation of several proposals in the literature and found that different loss functions are required for different situations, with the FPE loss function being the most indicated to deal with gravitational wave data.
 Along these lines, we directly compared the PSDs computed with MESA with the canonical Welch's algorithm.
 As outlined in Sec.~\ref{sec:Welch_comparison}, MESA provides PSD estimates with smaller variance and better accuracy than Welch algorithm.
-The use of MESA is particularly useful for short time series samples, where Welch's method is outperformed in both precision and confidence.
+%The use of MESA is particularly useful for short time series samples, where Welch's method is outperformed in both precision and confidence.
 
-This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD.
+%This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD.
 Hence MESA could be employed for {\it online} PSD estimation during a Bayesian parameter estimation exercise.
-%a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
-%\todo{There might be a misunderstanding: I guess the referee is talking about taking into account the PSD errors, while we are talking about dropping the stationariety assumption...
+%%a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
+%%\todo{There might be a misunderstanding: I guess the referee is talking about taking into account the PSD errors, while we are talking about dropping the stationariety assumption...
 The misleading word is `marginalize', I think}
-This is possible whenever the log-likelihood of the model for a time series takes the form
-\begin{equation}
-	\log\mathcal{L}(d_t | \theta) \propto  -\frac{1}{2} \frac{|\tilde{d}(f)- \tilde{x}(f;\theta)|^2}{S(f)} - \frac{1}{2}\log\left[\int  df\,S(f)\right]
-\end{equation}
-where $\tilde{\hphantom{a}}$ denotes the Fourier transform and the signal model $x(t;\theta)$, dependent on some parameter $\theta$, is a prediction for a deterministic signal buried in the observed time series $d(t)$.
-In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source: this scheme assumes {\it stationary data} on a long timescale (longer than the data under study) and it might not reflect the structure of the noise in the analyzed time slice.
-Some alternatives exist for dropping this assumption and modifying the likelihood accordingly \citep{R_ver_2010, R_ver_2011, Edwards_2020, Chatziioannou_2021}.
-MESA can add to those an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on the residual $d_t-x_t$.
-In this way, the PSD would depend also on $\theta$ and would effectively model the residuals.
+%This is possible whenever the log-likelihood of the model for a time series takes the form
+%\begin{equation}
+%	\log\mathcal{L}(d_t | \theta) \propto  -\frac{1}{2} \frac{|\tilde{d}(f)- \tilde{x}(f;\theta)|^2}{S(f)} - \frac{1}{2}\log\left[\int  df\,S(f)\right]
+%\end{equation}
+%where $\tilde{\hphantom{a}}$ denotes the Fourier transform and the signal model $x(t;\theta)$, dependent on some parameter $\theta$, is a prediction for a deterministic signal buried in the observed time series $d(t)$.
+%In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source: this scheme assumes {\it stationary data} on a long timescale (longer than the data under study) and it might not reflect the structure of the noise in the analyzed time slice.
+%Some alternatives exist for dropping this assumption and modifying the likelihood accordingly \citep{R_ver_2010, R_ver_2011, Edwards_2020, Chatziioannou_2021}.
+%MESA can add to those an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on the residual $d_t-x_t$.
+%In this way, the PSD would depend also on $\theta$ and would effectively model the residuals.
 For this method, we would only need to assume the stationariety of the residual on the batch to analyse\footnote{
-Unlike several works in GW data analysis \citep{Biscoveanu_2020, talbot2020gravitationalwave}, this approach does not address the issue of including the PSD uncertainties in the posterior.
+%Unlike several works in GW data analysis \citep{Biscoveanu_2020, talbot2020gravitationalwave}, this approach does not address the issue of including the PSD uncertainties in the posterior.
 Instead, it focuses on making minimal assumption about the nature of the data under study.
-}, making a lighter assumption on the nature of the data. Furthermore, this method would get, as a bonus, a posterior distribution for the PSD of the analyzed data.
-Preliminary studies suggest that this is possible~\citep{martini_thesis}.
+%}, making a lighter assumption on the nature of the data. Furthermore, this method would get, as a bonus, a posterior distribution for the PSD of the analyzed data.
+%Preliminary studies suggest that this is possible~\citep{martini_thesis}.
 Other studies have been done in this direction, mostly using a parametric model for the PSD \citep{Littenberg_2013, Edwards_2015, lalinference}.
 
-Hence, as an application of MESA, and its implementation in \texttt{memspectrum}, we analyse GW150914\cite{} using a Bayesian framework that allows for the marginalisation of the order  $p$ of the AR($p$) process representing the noise data stream.
-The inference algorithm proceeds as usual, see \cite{}[veitch and vecchio, ashton et al, etc etc], with a modification to the likelihood calculation.
-The likelihood function we employ is given in Eq.~\eqref{eq:Whittle} and its calculation proceeds as follows: for all proposed Monte Carlo moves $\theta\equiv\theta_{GW}\cup p$\footnote{We indicate the set of all GW parameters (component masses, spins, luminosity distance, etc.) with $\theta_{GW}$.}, for each detector included in the analysis, we compute the detector response to the template $h(t;\theta_{GW})$ in the time-domain, subtract it from the data $d(t)$ to get the time domain residuals $r(t;\theta_{GW})$.
-Under the MESA assumption, we describe the PSD of the residuals as in Eq.~\eqref{eq:MESApsd}, where the order $p$ is fixed by the Monte Carlo sampler.
-Having computed the PSD, we Fourier transform the residuals and compute the likelihood.
-Our sampler is a nested sampling algorithm\cite{}[raynest] and the specific inference model is implemented as part of \texttt{granite}\cite{}[in preparation], a dedicated inference model for ground based interferometric detectors.
-Needless to say, despite the implementation of the fast Burg algorithm in \texttt{memspectrum}, the inference with noise marginalisation is vastly more computational expensive than the standard analysis.
-As our analysis template, we adopt the machine learning based MLGW model\cite{}[MLGW], an aligned spin model trained on TEOBResumS\cite{}, that has been shown to perform well on LVK detected events\cite{}[MLGW].
+Hence, as an application of MESA, and its implementation in \texttt{memspectrum}, we analyse GW150914\cite{} using a Bayesian framework that allows for the marginalisation of the order  $k$ of the AR($k$) process representing the residuals data stream.
+Although the inference is essentially unchanged compared to the standard case, see \cite{lalinference}, there are some substantial modification to the likelihood construction.
+Since MESA is applicable to time-domain data, all calculations prior to the Fourier transform must be performed in time domain, thus increasing the computational cost by a non-negligible amount.
+We shall refer the time-of-arrival parameter $t_c$ of the GW to the geocenter.
+At each iteration of the inference algorithm, we sample a vector $\theta\equiv\theta_{GW}\cup k$\footnote{We indicate the set of all GW parameters (component masses, spins, luminosity distance, etc.) with $\theta_{GW}$.}.
+For each inteferometer $j$, therefore, we need to compute a time-delay $\Delta t_j$ to compute the antenna response functions $F_{j,+}(t+\Delta t_j),F_{j,\times}(t+\Delta t_j)$ as well as the correct time-shift for the GW template
+\begin{align}
+h_j(t) =\sum_{p = +,\times}F_{j,p}(t+\Delta t_j)h_{p}(t+\Delta t_j;\theta_{GW})
+\end{align}
+that we use to compute the time-domain residuals $r_j(t) = d_j(t) - h_j(t)$.
+We apply \texttt{memspectrum} to $r_j(t)$ with the \emph{fixed} value of $k$ and calculate the detector likelihood for $\tilde{r}_j$ using Eq.~\eqref{eq:Whittle} and PSD as in Eq.~\eqref{eq:MESApsd}.
+The coherent likelihood is then given by the product of the individual likelihoods.
+As our analysis template, we adopt the fast machine learning based MLGW model\cite{}[MLGW], an aligned spin model trained on TEOBResumS\cite{}, that has been shown to perform well on LVK events detected during O1 and O2\cite{}[MLGW].
+Our sampler is a nested sampling algorithm\cite{}[raynest] and the specific inference model is implemented as part of \texttt{granite}\cite{}[in preparation], a dedicated inference model for ground-based interferometric detectors.
+LVK samples from https://zenodo.org/records/6513631
 

### 2023-12-20 paper/mesa.tex
almost done with edits. done with plots

@@
 The coherent likelihood is then given by the product of the individual likelihoods.
-As our analysis template, we adopt the fast machine learning based MLGW model\cite{}[MLGW], an aligned spin model trained on TEOBResumS\cite{}, that has been shown to perform well on LVK events detected during O1 and O2\cite{}[MLGW].
-Our sampler is a nested sampling algorithm\cite{}[raynest] and the specific inference model is implemented as part of \texttt{granite}\cite{}[in preparation], a dedicated inference model for ground-based interferometric detectors.
-LVK samples from https://zenodo.org/records/6513631
+As our analysis template, we adopt the fast machine learning based MLGW model\cite{MLGW}, an aligned spin model trained on TEOBResumS\cite{teobresums}, that has been shown to perform well on LVK events detected during O1 and O2\cite{MLGW}.
+Our sampler is a nested sampling algorithm\cite{cpnest} and the specific inference model is implemented as part of \texttt{granite}, a dedicated inference model for ground-based interferometric detectors.
+We compare our results with the combined posterior samples available from GWOSC\cite{GWOSC} and available at \texttt{https://zenodo.org/records/6513631}.
 
 \begin{figure}
-	\caption{Posterior distributions for $\mathcal{M}$, $q$ and spin parameters $\chi_1$ and $\chi_2$. The recovered posteriors are very similar to the LVK released ones.}
-	\label{fig:gw150914_intrinsic}
-	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/intrinsic.pdf}
+	\caption{Posterior samples for $\mathcal{M}$ and mass ratio $q$ from the LVK(blue) and using \texttt{memspectrum}(red).
+The samples are largely consistent among the two models, with the MESA model providing a more conservative estimate.}
+	\label{fig:gw150914_masses}
+	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/masses.pdf}
 \end{figure}
 \begin{figure}
-	\caption{Posterior distributions for the sky position, (log) luminosity distance, inclination angle, polarisation angle $\psi$, reference phase $\phi_{ref}$ and geocentric coalescence time.
-Also for the extrinsic parameters, we see broad consistency with the LVK released ones.
-However, we notice that the posterior on the declination angle $\delta$ is mirrored compared to the released sky map.
-Nonetheless, there is substantial overlap.
-Similarly, the PSD marginalised posterior for the inclination angle shows a slight preference for a face-on configuration, rather than a face-off.}
-	\label{fig:gw150914_extrinsic}
-	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/extrinsic.pdf}
+	\caption{Posterior samples for the sky position angles from the LVK(blue) and using \texttt{memspectrum}(red).
+The samples are largely consistent among the two models, with the MESA model providing a more conservative estimate.}
+	\label{fig:gw150914_sky}
+	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/sky_position.pdf}
 \end{figure}
 \begin{figure}
-	\caption{Whitened reconstructed waveforms from our analysis for the Hanford detector (top panel) and the Livingston detector (bottom panel).
+	\caption{Whitened reconstructed waveforms and data from our analysis for the Hanford detector (top panel) and the Livingston detector (bottom panel).
 The shaded turquoise area indicates the 90\% credible region.}
@@
 \end{figure}
-%\begin{figure}
-%	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
-The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
-The confidence interval is computed on  $100$ independent random realization of the predictions.
-%	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
-%	}
-%	\label{fig:gw150914_intrinsic}
-%	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/posterior_order_H1.pdf}
-%\end{figure}
-%
-%\begin{figure}
-%	\caption{In the top panel, we show the difference between the forecasted strain values (CAT model) and the actual values observed at Livingston interferometer (blue).
-The red shadow denotes the $90\%$ confidence interval subtracted with the actual time series values.
-The confidence interval is computed on  $100$ independent random realization of the predictions.
-%	The bottom panel displays the standard deviation of the predictions with both CAT and FPE models. The inset shows a detail of the first $\SI{0.5}{s}$ of plot in the top panel.
-%	}
-%	\label{fig:gw150914_intrinsic}
-%	\includegraphics{Images/gw150914/posterior_order_L1.pdf}
-%\end{figure}
+\begin{figure}
+	\caption{Posterior distributions for AR process orders in the Hanford(red) and Livingston(blue).}
+	\label{fig:gw150914_orders}
+	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/orders.pdf}
+\end{figure}
 
 \begin{figure}
-	\caption{Posterior PSDs for the Hanford (orange line) and Livingstone (blue line) as inferred by our analysis.}
-	\label{fig:gw150914_psd}
-	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/noise_curves.pdf}
+	\caption{\emph{Top panel}: Posterior PSDs for the Hanford (orange line) and Livingstone (blue line) as inferred by our analysis.
+\emph{Bottom panel}: Relative uncertainty around the median for both the Hanford(blue) and the Livingston(orange) PSDs.}
+	\label{fig:gw150914_psds}
+	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/posterior_psds.pdf}
 \end{figure}
 
-In Figs.~\ref{fig:gw150914_intrinsic}, ~\ref{fig:gw150914_extrinsic} and \ref{fig:gw150914_waveforms} we show the posteriors for the set of intrinsic parameters, extrinsic parameters and reconstructed waveform from our analysis.
-The inferred PSDs are instead shown in Fig.~\ref{fig:gw150914_psds}.
-%In addition, Fig.~ \ref{fig:gw150914_orders} shows the posterior distribution on the orders $p_{H}$ and $p_{L}$ for the AR processes describing the residuals in the LIGO-Hanford and LIGO-Livingston detectors.
+In Figs.~\ref{fig:gw150914_masses}, ~\ref{fig:gw150914_sky} and \ref{fig:gw150914_waveforms} we show the posteriors for the set of intrinsic parameters, extrinsic parameters and reconstructed waveform from our analysis.
+Our results can be summarised as follows: our posterior samples are in general consistent with what has been released by the LVK, however our credible regions tend to be larger.
+This is expected since our likelihood includes additional uncertainty due to the explicit sampling over the process order, hence the PSD.
+For the particular 4 seconds of data, sampled at 4096\,Hz, the recovered orders are $k_{H1} = 1107_{-5}^{+9}$ and $k_{L1} = 1146_{-8}^{+8}$, Fig.~\ref{fig:gw150914_orders}.
+The corresponding PSDs and uncertainties are shown in Fig.~\ref{fig:gw150914_psds}.
 

### 2023-12-20 setup.py

@@ -1,5 +1,6 @@
 from setuptools import setup
 import os
+import sys
 
 try:
 	from sphinx.setup_command import BuildDoc

### 2023-12-21 examples/doc_examples.py
Check of the examples

@@ -1,5 +1,5 @@
 """
-All the examples in the docs...
+All the examples in the docs. To run at once...
 """
 #Getting the data
 from memspectrum.GenerateTimeSeries import generate_data

### 2023-12-21 examples/generate_white_noise.py

@@ -23,7 +23,7 @@ import warnings
 
 parser = argparse.ArgumentParser(__doc__)
 parser.add_argument(
-	"--p", default = 200, type = int, 
+	"--p", default = 300, type = int, 
 	help="Autoregressive order of the final timeseries")
 parser.add_argument(
 	"--srate", default = 4096, type = float, 

### 2023-12-21 examples/gwstrain.py

@@ -1,9 +1,7 @@
-try:
-	import sys
-	sys.path.insert(0,'..')
-	from memspectrum import MESA
-except:
-	from memspectrum import MESA
+"""
+This script will take some data from the Virgo interferometer and will compute the PSD. It will also sample from the distribution of future observations and do some forecast on the future values.
+"""
+from memspectrum import MESA
 
 import numpy as np
 import matplotlib.pyplot as plt

### 2023-12-21 examples/sound_MESA.py

@@ -6,14 +6,8 @@ import numpy as np
 import scipy.io.wavfile
 import matplotlib.pyplot as plt
 
-try:
-    import sys
-    sys.path.insert(0,'..')
-    from memspectrum import *
-    import GenerateTimeSeries
-except:
-    from memspectrum import *
-    import memspectrum.GenerateTimeSeries as GenerateTimeSeries
+from memspectrum import MESA
+import memspectrum.GenerateTimeSeries as GenerateTimeSeries
 
 t = 14. #seconds of data
 

### 2023-12-21 examples/sunspots.py

@@ -1,9 +1,8 @@
-try:
-	import sys
-	sys.path.insert(0,'..')
-	from memspectrum import *
-except:
-	from memspectrum import *
+"""
+Script to compute the power spectral density of the measured timeseries of the number of sunspots measured.
+"""
+
+from memspectrum import MESA
 
 import numpy as np
 import matplotlib.pyplot as plt

### 2023-12-21 examples/doc_examples.py
Added initial plot in the docs

@@ -1,6 +1,38 @@
 """
 All the examples in the docs. To run at once...
 """
+
+#Intro plot - temperature timeseries
+import memspectrum
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.ticker
+
+timeseries = np.loadtxt('data/temperature_data_@1h_len5000.dat')
+dt, T = 1, len(timeseries) #Time step, Lenght of the timeseries
+N_tstep = 100
+f = np.arange(1/(10*24), 0.5, 1/T)
+f_values = [1/(7*24), 1/24, 1/12, 1/8, 0.25, 0.5, 1]
+f_names = ['1/week', '1/day', '2/day', '3/day', '1/4hours', '1/2hours','1/hour']
+
+M = memspectrum.MESA()
+M.solve(timeseries) #perform the analysis on the given time series (a real/complex np.array)
+PSD = M.spectrum(dt, f) #evaluate the PSD on the given frequency grid
+future_timeseries = M.forecast(timeseries[:1000], N_tstep) #forecast from the time series
+
+plt.figure(figsize = (3.54*2, 3.54))
+plt.loglog(f,PSD)
+plt.xlabel('Frequency (1/hour)')
+for f_val in f_values:
+	plt.axvline(f_val, ls = '--', c= 'k')
+ax = plt.gca()
+
+ax.xaxis.set_major_locator(matplotlib.ticker.FixedLocator((f_values)))
+ax.xaxis.set_major_formatter(matplotlib.ticker.FixedFormatter((f_names)))
+plt.xticks(rotation=45)
+plt.tight_layout()
+plt.savefig('../docs/img/temperature_plot_intro.png', dpi = 200)
+
 #Getting the data
 from memspectrum.GenerateTimeSeries import generate_data
 import numpy as np

### 2023-12-21 memspectrum/__init__.py

@@ -15,7 +15,7 @@ For more information, take a look at:
 
 Basic usage (to compute the spectrum of a given time series):
 
-::
+.. code-block:: python
 
 	import memspectrum
 
@@ -23,6 +23,15 @@ Basic usage (to compute the spectrum of a given time series):
 	M.solve(time_series) #perform the analysis on the given time series (a real/complex np.array)
 	M.spectrum(dt,f) #evaluate the PSD on the given frequency grid
 	M.forecast(data, N_tstep) #forecast from the time series
+
+With the proper input, this will produce this nice plot:
+
+.. image:: img/temperature_plot_intro.png
+
+If you're curious, this is the Power Spectral Density of the historical series of temperature measured around the city of Milan (Italy) with 1 hour rate. Data are taken from `Copernicus <https://cds.climate.copernicus.eu/#!/home>`_. You can clearly see that the temperatures are very nicely correlated with themselfs on a day timescale. This amounts to a large peak at a frequency of 1 per hour and its multiples.
+
+You can find this and many other examples around this documentation. For some example code, you can check the `example <https://github.com/martini-alessandro/Maximum-Entropy-Spectrum/tree/main/examples>`_ folder of this repo.
+
 
 """
 

### 2024-02-12 paper/mesa.tex
final paper draft

@@
 	\author{Alessandro \surname{Martini}\inst{\ref{inst1}}\thanks{martini.alessandr@gmail.com} \and 
-	Stefano \surname{Schmidt}\inst{\ref{inst2}}\thanks{s.schmidt@uu.nl} \and Walter \surname{Del Pozzo}\inst{\ref{inst1}}\thanks{walter.delpozzo@unipi.it}}
+	Stefano \surname{Schmidt}\inst{\ref{inst2}}\thanks{s.schmidt@uu.nl} \and Gregory \surname{Ashton}\inst{\ref{inst1}}\thanks{walter.delpozzo@unipi.it} \and Walter \surname{Del Pozzo}\inst{\ref{inst1}}\thanks{walter.delpozzo@unipi.it}}
 
@@
 The Maximum Entropy Spectral Analysis (MESA) method, developed by Burg, provides a powerful tool to perform spectral estimation of a time-series.
-The method relies on a Jaynes' maximum entropy principle and provides the means of inferring the spectrum of a stochastic process in terms of the coefficients of some autoregressive process AR($p$) of order $p$.
+The method relies on Jaynes' maximum entropy principle and provides the means of inferring the spectrum of a stochastic process in terms of the coefficients of some autoregressive process AR($p$) of order $p$.
 A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
-We provide a ready-to-use implementation of the algorithm in the form of a python package \texttt{memspectrum}.
-We characterize our implementation by performing a power spectral density analysis on synthetic data (with known power spectral density) and we compare different criteria for stopping the recursion.
+We provide a ready-to-use implementation of the algorithm in the form of a python package \texttt{memspectrum} and we characterize our implementation by performing a power spectral density analysis on synthetic data (with known power spectral density) and by comparing different criteria for stopping the recursion.
 Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using synthetic data generated from the released spectrum by the LIGO-Virgo-Kagra collaboration.
@@
 \section{Introduction}
-\todo{Check again intro \& abstract}
-The study of the properties of stochastic processes is a crucial task in many fields of physics, astronomy, quantitative biology, as well as engineering and finance.
-Among those, a special place is occupied by the so-called \textit{wide-sense} stationary processes.
-These are stochastic processes that display an invariance of their statistical properties, such as their two-point autocovariance function, with respect to time translation.
-If $x(t)$ is a wide-sense stationary process, it is completely determined by the knowledge of the autocorrelation function
+
+The problem of inferring the morphology and the defining parameters of deterministic signals superimposed to stochastic processes is one of the most wide spread and interesting problems in several areas of human activities.
+Whenever some form of model for the signal we are looking for is available, the problem is typically solved via the Wiener filter, defined as the whitening filter that maximises the signal-to-noise ratio, i.e.
+the relative power of the (known) signal over the power of the (known) underlying stochastic process.
+Hence, signal detection and characterisation requires accurate knowledge of i) the shape of the signal we are looking for and ii) the statistical properties of the stochastic process.
+The construction of signal models is typically driven either by physical or by mathematical arguments hence, although extremely difficult in general, it is doable.
+On the other hand, stochastic process models can be extremely difficult to construct, both for practical and theoretical reasons.
+A stochastic process is fully described by the knowledge of the probability distribution governing its realisations -- the ``paths'' of the random variable under scrutiny -- over the entire time axis, from $t=-\infty$ to $t=\infty$.
+Clearly this is not possible in practice.
+Therefore modelling a stochastic process either relies on modelling of the underlying physical processes, thus falling back onto the deterministic case, or on modelling the mathematical and statistical properties of the process, and potentially infer them from the process realisations.
+The study of the properties of stochastic processes is thus a crucial task in many fields of physics, astronomy, quantitative biology, as well as engineering and finance.
+Among the classes of stochastic processes, a key role is played by \textit{wide-sense} stationary processes.
+These are stochastic processes that display an invariance of their statistical properties, such as their two-point autocovariance function, with respect to translation of the independent variable, usually the time $t$.
+If $x(t)$ is a wide-sense stationary process, it is in fact completely determined by the knowledge of the (many-points) autocorrelation functions.
+In practice, one often has easy access to the two-point correlation function
 \begin{equation}
@@
 \end{equation}
-or, equivalently, by the knowledge of their \emph{power spectral density} (PSD) $S(f)$. Thanks to the Wiener-Khinchin theorem the two are related by a Fourier transform: 
+or, equivalently, to the process \emph{power spectral density} (PSD) $S(f)$.
+Thanks to the Wiener-Khinchin theorem, in wide-sense stationary processes, the two are in fact related by a Fourier transform:
 \begin{equation}\label{eq:psd-autocorrelation}
@@
 \end{equation}
-In some literature, especially in the context of gravitational waves physics, e.g. ~\citep{Finn_1992}, the PSD is introduced as  
+In the context of gravitational waves physics, e.g. ~\citep{Finn_1992}, the PSD is introduced as  
 \begin{equation}\label{eq:psd-f-definition}
@@
 
-An ubiquitous method for such computation is due to \citet{Welch1967} and it is based on Eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
-The PSD is obtained by slicing the observed realization $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
+An ubiquitous method for such a computation is due to \citet{Welch1967} and it is based on Eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
+The PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
 This approach is equivalent~\citep{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
@@
 Despite being effective, this method is problem dependent, since it needs to make definite assumptions on the shape of the PSD.
-Moreover, it can be computationally expensive and it does not come with an handy implementation available to the public.
-For all the above reasons, we did not consider such method in our work.
+Moreover, it can be computationally expensive and it does not come with a handy implementation available to the public.
+For all the above reasons, we did not consider such methods in our work.
 
-An appealing alternative, based on the Maximum Entropy principle \citep{JaynesArticle,jaynes2003ptl, Jaynes_MAXENT}, has been put forward by~\citet{burg1975maximum}.
+An appealing alternative, based on the Maximum Entropy principle \citep{JaynesArticle,jaynes2003ptl, Jaynes_MAXENT}, has been derived by~\citet{burg1975maximum}.
 Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the algorithm parameters, since it provides an iterative closed form expression for the spectrum of a stochastic stationary time series.
@@
 Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many undesired aspects of the Welch's algorithm (or other similar methods).
-To facilitate this goal, we present and describe a new code, \texttt{memspectrum}, that provides a robust and easy-to-use python implementation of the algorithm\footnote{
-It is available at link: \url{https://pypi.org/project/memspectrum/}.
-}.
+To facilitate this goal, we based this study on \texttt{memspectrum}, a freely available, robust and easy-to-use python implementation of the algorithm described below\footnote{
+It is available at link: \url{https://pypi.org/project/memspectrum/}.}.
 We provide a thorough assessment of the performance of our code and we validate our results performing a number of tests on simulated and real data.
@@
 The Maximum Entropy principle (MAXENT) is among the most important results in probability theory.
-It provides a way to uniquely assign probabilities to a phenomenon in a way that best represent our state of knowledge, while being non committal with unavailable information.
+It provides a way to uniquely assign probabilities to a phenomenon in a way that best represent our state of knowledge, while being non-committal to unavailable information.
 Its domain of application turned out to be wider than expected.
@@
 
-After a short introduction to Jaynes' MAXENT (Sec.~\ref{sec:MAXENT}), we will develop in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can be expressed in an analytical closed form (Sec.~\ref{sec:MESA}).
-Next, we will discuss an interesting link between Burg's method and autoregressive processes (Sec.~\ref{sec:autoregr}) and in Sec.~\ref{sec:forecasting} we will use such link for straightforwardly forecasting a time series.
+After a short introduction to Jaynes' MAXENT (Sec.~\ref{sec:MAXENT}), we will review in detail Burg's technique of Maximum Entropy Spectral Analysis (MESA) and show that the estimate can always be expressed in an analytical closed form (Sec.~\ref{sec:MESA}).
+Next, we will discuss the interesting link between Burg's method and autoregressive processes (Sec.~\ref{sec:autoregr}) and in Sec.~\ref{sec:forecasting} we will use such link to forecast a time series.
 
@@
 
-Before introducing MAXENT principle, we will define through some simple examples the two core concepts of the problem and the role they play: the `evidence' and the `information'.
-Let us start with the `information' (or entropy): it is a measure of the degree of uncertainty on the outcomes of some experiment and specifies the length of the message necessary to provide a full description of the system under study.
+Before introducing MAXENT principle, we will define, via some simple examples, the two core concepts of the problem and the roles they play in deductive inference: the `evidence' and the `information'.
+Let us start with the `information' (or information entropy): it is a measure of the degree of uncertainty on the outcomes of some experiment and specifies the length of the message necessary to provide a full description of the system under study.
 For instance, consider a perfectly sinusoidal signal: knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information content.
-Even less information is required if we are studying a system whose outcome is certain (has probability $p = 1$), as in this case, a communication is not even needed.
+Even less information is required if we are studying a system whose outcome is certain (the outcome is known with probability $p = 1$), as in this case, a communication is not even needed.
 \citet{Shannon} proposed the quantity
@@
 to measure the quantity of information brought by an outcome $x$ with probability $p(x)$.
-It is additive quantity as well as monotonically decreasing as a function of $p \in [0, 1]$: the more uncertain the outcome, the higher the information it brings.
+It is additive quantity as well as a monotonically decreasing function of $p \in [0, 1]$: the more uncertain the outcome, the higher the information it brings.
 
@@
 The information content will be very low.
-On the other hand, most unpredictable experiment happens when 
+On the other hand, most unpredictable situation happens when 
 \begin{equation}\nonumber
@@
 this describes a situation of `maximum ignorance' and the information content of such system must be high.
-Any generalization of eq.~\eqref{eq:information}, must then have its maximum when $ P_1 = P_2$.
+Any generalization of Eq.~\eqref{eq:information}, must then have its maximum when $ P_1 = P_2$.
 For $N$ events, the system with the highest possible information content is when:
@@
 \begin{equation}\label{eq:entropy}
-    H[p_1, \dots, p_N] = - \sum_{i = 1}^N p_i\log{p_i},
+    H[p_1, \dots, p_N] = - \sum_{i = 1}^N p_i\log{p_i}\,,
 \end{equation}
-The equation can be interpreted as the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
+which can be interpreted as the `expected information' brought by an experiment with $N$ possible outcomes each with its own probability $p_i$.
 In the continuous case:
@@
 \end{equation}
-We call the functional $H$ (information) entropy\footnote{In defining the information entropy as in Eq.~(\ref{eq:entropy_continuos}) we are implicitly assuming a uniform measure over the parameter space}.
+We call the functional $H$ (information) entropy\footnote{In defining the information entropy as in Eq.~\eqref{eq:entropy_continuos}, we are implicitly assuming a uniform measure over the parameter space.
+In case of a non-uniform measure $m(x)$, the definition generalises to $ H[p(x)] = - \int p(x)\ln \frac{p(x)}{m(x)} dx$.}.
 
-We now turn to the core of our problem: how do we make a probability assignment for a set of events, that keeps into account our knowledge of the system and, at the same time, it is non committal towards unavailable knowledge?
-The knowledge at our disposal about the system under study is what we call `evidence' and any probability assignment must agree with it.
-In the above cases, our knowledge on the system is only the total number $N$ of different outcomes -- this is a minimal requirement. Of course, more complex evidence constraints can be applied.
+We now turn to the core of our problem: how can we assign probabilities to a set of events keeping into account our knowledge of the system and, at the same time, it is non-committal towards unavailable knowledge?
+The ``knowledge'' at our disposal about the system under investigation is what we define `evidence' and any probability assignment is given such evidence, in agreement with \citet{Cox} construction of probability.
+In the case above, our knowledge on the system is only the total number $N$ of different outcomes -- this is a minimal requirement.
+Of course, more complex evidence constraints can be applied.
 
@@
 }.
-In other words, the amount of `information' introduced by the probability assignment should be as high as possible.
-
-MAXENT puts the aforementioned principle into a more precise mathematical context by stating that probabilities should be assigned by maximizing uncertainty (information entropy) using evidence as constraint.
+In other words, the information entropy content introduced by the probability assignment should be as large as possible, in accordance with the available evidence.
+MAXENT formalises this reasoning by stating that probabilities should be assigned by maximizing uncertainty (information entropy) using evidence as constraint.
 This defines a variational problem, where the information entropy functional $H\left[p_1, \dots, p_N\right]$, defined in eq.~\eqref{eq:entropy}, has to be maximized.
 
-The maximisation of entropy, supplemented by evidence in the form of constraints to which the sought-for probability distribution must obey,
-gives rise to several of the most common probability distributions commonly employed in statistics.
+The maximisation of entropy, supplemented by evidence in the form of constraints to which the sought-for probability distribution must obey, gives rise to several of the most common probability distributions commonly employed in statistics.
 For instance, whenever the only constraint available is the normalization of the probability distribution (i.e.
 no evidence is available), the entropy is maximised by the uniform distribution.
-If we have evidence to constraint the expected value, the information entropy is maximised by the exponential distribution
+If we have evidence to constraint the expected value, the information entropy is maximised by the exponential distribution.
 
-Of particular importance is the case in which, in addition to the mean, also the variance is known: MAXENT leads to the Gaussian distribution. 
+Of particular relevance for our purposes is the case in which, in addition to the mean, also the variance is known: MAXENT leads to the Gaussian distribution. 
 This derivation is particularly interesting from the foundational point of view, since it provides a deeper insight into the ubiquitous Gaussian distribution.
-Indeed, it is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy.
-For this reason, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process.
-In some sense, we can interpret the central limit theorem as the natural `statistical' evolution toward a configuration that maximizes entropy.
+Indeed, it is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy for a fixed mean and variance: from the MAXENT principle, it is the correct probability distribution to assign if mean and covariance are the only quantities that fully define our process.
+In some sense, we can interpret the central limit theorem as the natural `statistical' evolution toward a configuration that maximizes entropy in repeated experiments.
 
-For this work, we are particularly interested in the multi-dimensional case.
+For this work, we are especially interested in the multi-dimensional case.
 Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realization of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of autocovariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution~\citep{gregory_2005}:
@@
 where the matrix $S_{ij} = S_i \delta_{ij}$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
-Many readers will recognize the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\citep{prob_information_theory}
+Many readers will recognise the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\citep{prob_information_theory}
 and of gravitational waves data analysis, \citep[e.g.]{Finn_1992, FINDCHIRP}.
@@
 
-In principle, if the autocorrelation was known exactly (i.e. at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD 
-would reduce to a single Fourier transform.
-However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process, hence such computation
-is impossible.
+In principle, if the autocorrelation was known exactly (i.e.
+at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD would reduce to a single Fourier transform.
+However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process, hence such computation is impossible.
 Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases as $\sigma \sim 1/\sqrt{N - k}$\footnote{
-This is clearly understood: when computing the autocorrelation at order $k$, only $N-k$ examples of the product $x_t x_{t+k}$ are available and the variance of the average value goes as the inverse of the square root of the points considered.
+This is easily understood: when computing the autocorrelation at order $k$, only $N-k$ examples of the product $x_t x_{t+k}$ are available and the variance of the average value goes as the inverse of the square root of the points considered.
 }, so that only few values for the autocorrelation function can actually be computed reliably.
@@
 The latter assumes that the unknown values of the autocorrelation are $0$.
-Clearly, this assumption is unjustified and MAXENT is a good way to drop this assumption.}.
+Clearly, this assumption is unjustified and MAXENT is a good way to relax this assumption.}.
 
@@
 Before doing so, there is a technicality to solve: the definition of entropy depends on a probability distribution, not on the PSD.
-It can be shown, ~\citep[e.g.]{AblesMESA, Bartlett}, that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of the filtering a white noise process using a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
+It can be shown ~\citep[e.g.]{AblesMESA, Bartlett} that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of the filtering a white noise process using a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
 A filter with transfer function $T(f)$ takes in input a time series $x_t$ and outputs a times series $y_t$ such that:
@@
 where $\Delta t$ is sampling rate and $Ny \equiv \frac{1}{2 \Delta t}$  is the Nyquist frequency.
-Thus maximising Eq.~\eqref{eq:EntropyGain} is equivalent to maximizing eq.~\eqref{eq:entropy_continuos}.
+Thus maximising Eq.~\eqref{eq:EntropyGain} is equivalent to maximizing Eq.~\eqref{eq:entropy_continuos}.
 
@@
 
-Thus, by maximizing Eq.~\eqref{eq:EntropyGain} with constraints in Eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
-This approach on PSD computation provides a result consistent with the empirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
+Thus, by maximizing Eq.~\eqref{eq:EntropyGain} with constraints in Eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given a time series sample.
+This approach on PSD computation provides a result consistent with the empirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorrelation at large time lags.
 
-Most importantly, the variational problem admits a closed-form analytical expression for $S(f)$.
+Remarkably, the variational problem admits a closed-form analytical expression for $S(f)$.
 The expression was first found by ~\citet{burg1975maximum}:
 \begin{equation}\label{eq:MESApsd}
-    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^{N} a_s z^s\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}, 
+    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^{N} a_s z^s\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}\,,
 \end{equation}
@@
 The coefficients $a_s (s > 0)$, together with an overall multiplicative scale factor $P_N$, are to be determined by an iterative process (called Burg's algorithm)
-At least, two implementations of Burg's algorithm are available in the literature, labeled as `Standard' \cite{} and `Fast' \cite{} in the \texttt{memspectrum} package.
+At least, two implementations of Burg's algorithm are available in the literature, labeled as `Standard' and `Fast' in the \texttt{memspectrum} package.
 The `Standard' method is slower but more stable, while `Fast' trades stability for speed.
@@
 The number $N$ of such coefficients is a choice that shall be made by the user and indeed it is the only hyperparameter that needs to be tuned.
-The details of the derivation and the actual form for the coefficients $a_s$ can be found in appendix \ref{sec:appendix}.
+The details of the derivation and the actual form for the coefficients $a_s$ can be found in Appendix \ref{sec:appendix}.
 
@@
 where $b_1, \ldots, b_p$ are real coefficients and $\nu_t$ is white noise with a given variance $\sigma^2$.
-Thus, an AR($p$) process models the dependence of the value of the process at time $t$ on every past $p$ observations, 
+Thus, an AR($p$) process models the dependence of the value of the process at time $t$ from the last $p$ observations, 
 thus being potentially able to model complex autocorrelation structures within observations.
 
-Thanks to Wold's theorem~\citep{Wold_theorem}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general; it turns out that the maximum entropy principle provides a representation of the time series as an $AR(p)$ process and Burg's algorithm computes the autoregressive coefficients that are suitable to the available data.
+Thanks to Wold's theorem~\citep{Wold_theorem}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general; it turns out that the maximum entropy principle provides a representation of the time series as an $AR(p)$ process and Burg's algorithm computes the corresponding autoregressive coefficients that are suitable to model the available data.
 
@@
 \begin{align}
-    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t.
+    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t\,.
 \end{align}
@@
 \begin{equation}
-    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{n = 1}^p b_n z^n \right)}
+    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{n = 1}^p b_n z^n \right)}\,.
 \end{equation}
@@
 The theory of the AR processes can be also applied to the problem of whitening a time series.
-Given a time series, $x_t$, the whitening operation produces another timeseries $x^W_t$ such that:
+Given a time series, $x_t$, the whitening operation produces another time series $x^W_t$ such that:
 %
@@
 where $\mathcal{F}^{-1}$ denotes the inverse Fourier transform of a frequency series.
-If $x_t$ is a realization of gaussian noise (see Eq.~\eqref{eq:Whittle}) with PSD $S(F)$, the whitened timeseries $x^W_t$ is just white noise (i.e. uncorrelated samples from a normal gaussian).
+If $x_t$ is a realization of gaussian noise (see Eq.~\eqref{eq:Whittle}) with PSD $S(F)$, the whitened time series $x^W_t$ is just white noise (i.e. uncorrelated samples from a normal gaussian).
 
-From Eq.~\eqref{eq:AR_p}, remembering that $b_i = - a_i$, it's straightforward to derive an expression for the whitened timeseries $x^W_t$:
+From Eq.~\eqref{eq:AR_p}, remembering that $b_i = - a_i$, it's straightforward to derive an expression for the whitened time series $x^W_t$:
 \begin{equation}
@@
 \end{equation}
-This amounts to a convolution of the timeseries  $x_t$ with the kernel $(1, a_1, \hdots, a_p)$, plus a variance rescaling.
+This amounts to a convolution of the time series  $x_t$ with the kernel $(1, a_1, \hdots, a_p)$, plus a variance rescaling.
 Performing a convolution is an appealing alternative to evaluating Eq.~\eqref{eq:whitening_definition} directly.
@@
 \section{Validation of the model}\label{sec:validation}
-It is now clear that MESA provides a recursive formula for computing the coefficients $a_k$ in Eq.~(\ref{eq:MESApsd}).
+
+MESA provides a recursive formula for computing the coefficients $a_k$ in Eq.~(\ref{eq:MESApsd}).
 The number $M$ of such coefficients is equivalent to the maximum order of the autocorrelation $\bar{r}_m$ considered.
@@
 with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.
-Asymptoticall,y minimizing FPE is equivalent to minimizing the quantity: 
+Asymptotically minimizing FPE is equivalent to minimizing the quantity: 
 \begin{equation}
@@
 This second criterion \cite{kay1988modern} is based on a similar assumptions to FPE.
-It minimises the actual value of the least squares (insted of relying to asymptotical behaviour).
+It minimises the actual value of the least squares (instead of relying to asymptotical behaviour).
 t the normalising factor takes into account the k degrees of freedom necessary to estimate the forward prediction error filter $a_k$.\\ The quantity to be minimised is
@@
 	\caption{Whitened reconstructed waveforms and data from our analysis for the Hanford detector (top panel) and the Livingston detector (bottom panel).
-The shaded turquoise area indicates the 90\% credible region.}
+The shaded turquoise area indicates the 90\% credible region over the waveforms space while the purple contours indicate the 90\% credible regions over the whitened data.}
 	\label{fig:gw150914_waveforms}
@@
 The corresponding PSDs and uncertainties are shown in Fig.~\ref{fig:gw150914_psds}.
+The full joint posterior distribution recovered when marginalising over the AR order is shown in Appendix~\ref{sec:appendix_b}.
 
-\section{Final remarks and future prospects} \label{sec:conclusion}
-\todo{The conclusions will need to be rearrenged completely}
+\section{Summary and discussion} \label{sec:conclusion}
+
 We presented a case study of the application of Maximum Entropy principle to the realm of spectral estimation.
@@
 This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD.
-Hence MESA could be employed for {\it online} PSD estimation during a Bayesian parameter estimation exercise.
-%a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
-%\todo{There might be a misunderstanding: I guess the referee is talking about taking into account the PSD errors, while we are talking about dropping the stationariety assumption...
-The misleading word is `marginalize', I think}
-This is possible whenever the log-likelihood of the model for a time series takes the form
-\begin{equation}
-	\log\mathcal{L}(d_t | \theta) \propto  -\frac{1}{2} \frac{|\tilde{d}(f)- \tilde{x}(f;\theta)|^2}{S(f)} - \frac{1}{2}\log\left[\int  df\,S(f)\right]
-\end{equation}
-where $\tilde{\hphantom{a}}$ denotes the Fourier transform and the signal model $x(t;\theta)$, dependent on some parameter $\theta$, is a prediction for a deterministic signal buried in the observed time series $d(t)$.
-In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source: this scheme assumes {\it stationary data} on a long timescale (longer than the data under study) and it might not reflect the structure of the noise in the analyzed time slice.
-Some alternatives exist for dropping this assumption and modifying the likelihood accordingly \citep{R_ver_2010, R_ver_2011, Edwards_2020, Chatziioannou_2021}.
-MESA can add to those an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on the residual $d_t-x_t$.
-In this way, the PSD would depend also on $\theta$ and would effectively model the residuals.
-For this method, we would only need to assume the stationariety of the residual on the batch to analyse\footnote{
-Unlike several works in GW data analysis \citep{Biscoveanu_2020, talbot2020gravitationalwave}, this approach does not address the issue of including the PSD uncertainties in the posterior.
-Instead, it focuses on making minimal assumption about the nature of the data under study.
-}, making a lighter assumption on the nature of the data. Furthermore, this method would get, as a bonus, a posterior distribution for the PSD of the analyzed data.
-Preliminary studies suggest that this is possible~\citep{martini_thesis}.
-Other studies have been done in this direction, mostly using a parametric model for the PSD \citep{Littenberg_2013, Edwards_2015, lalinference}.
-
-%Several approaches have been developed to try to study the validity of the assumption or even to drop it \cite{R_ver_2010, R_ver_2011, Littenberg_2013, Biscoveanu_2020, talbot2020gravitationalwave, Edwards_2020, Edwards_2015}.
-%Besides them, MESA can offer an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on $d_t-x_t$.
-%In this way, the PSD would depend also on $\theta$ and would effectively model the residuals. Preliminary studies suggest that this is possible~\cite{martini_thesis}.
+By applying MESA to 4 seconds of data in correspondence to GW150914, we demonstrated that it is possible to simultaneously estimate the signal and noise parameters, hence effectively marginalise over the noise PSD, without the need to
+\begin{itemize}
+\item assume a specific functional form for the PSD;
+\item estimate the PSD in an off-source segment of data.
+\end{itemize}
+Both items are of particular interest for several reasons that we shall discuss in what follows.
+Several proposals exist in literature attempt to marginalise over the PSD, mostly using a parametric model for the PSD \citep{Littenberg_2013, Edwards_2015, lalinference}.
+MAXENT fixes the functional form for us exploiting the correspondence with AR processes, providing a one-parameter family of models that are particularly easy to sample, thus grounding the noise properties marginalisation in solid theoretical foundations and in an easy-to-use numerical implementation.
+The latter point is also particularly relevant, especially in the context of future GW detectors.
+Future detectors are in fact expected to be operating in the signal dominated regime, with several sources -- potentially from different classes -- constantly present within the detectors' data streams.
+In these cases the common procedure of estimating the PSD from off-sources segments is bound to fail and or provide biases inferences.
+MAXENT and MESA model and are relevant \emph{only for the segment of data under consideration}, and make no assumptions over what is not part of the analysis.
+We believe, and we will show in a future study, that using MESA is the only way of computing single-source posteriors whenever multiple sources are overlapping.
+This is possible since, in our formulation, everything that we did not label as \emph{signal} will be part of the residuals, over which we apply MESA.
 
@@
 Another faster recursion method is available in \citet{Vos} and it is also available in \texttt{memspectrum}.
+
+\section{Full posterior distribution for GW150914} \label{sec:appendix_b}
+\begin{figure}
+	\caption{Full posterior distribution for GW150914 when marginalising over the AR process orders.}
+	\label{fig:gw150914_full_pe}
+	\includegraphics[width=\textwidth,keepaspectratio]{Images/gw150914/mega_plot.pdf}
+\end{figure}
 \end{appendix}

### 2024-02-12 paper/mesa.tex
added and corrected affiliations

@@
 	\title{Maximum Entropy Spectral Analysis: a case study}
-	\author{Alessandro \surname{Martini}\inst{\ref{inst1}}\thanks{martini.alessandr@gmail.com} \and 
-	Stefano \surname{Schmidt}\inst{\ref{inst2}}\thanks{s.schmidt@uu.nl} \and Gregory \surname{Ashton}\inst{\ref{inst1}}\thanks{walter.delpozzo@unipi.it} \and Walter \surname{Del Pozzo}\inst{\ref{inst1}}\thanks{walter.delpozzo@unipi.it}}
+	\author{Alessandro \surname{Martini}\inst{1,2,3}\thanks{alessandro.martini-1@unitn.it} \and 
+	Stefano \surname{Schmidt}\inst{4,5}\thanks{s.schmidt@uu.nl} \and Gregory \surname{Ashton}\inst{6} \and Walter \surname{Del Pozzo}\inst{1}\thanks{walter.delpozzo@unipi.it}}
 
-        \institute{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy\label{inst1} \and Institute for Gravitational and Subatomic Physics (GRASP),Utrecht University, Princetonplein 1, 3584 CC, Utrecht, The Netherlands\label{inst2}}
+        \institute{Dipartimento di Fisica  Università di Pisa, and INFN Sezione di Pisa, Pisa I-56127,Italy
+        		      \and Università di Trento, Dipartimento di Fisica, I-38123 Povo, Trento, Italy
+		      \and INFN, Trento Institute for Fundamental Physics and Applications, I-38123 Povo, Trento, Italy
+        		      \and Institute for Gravitational and Subatomic Physics (GRASP),Utrecht University, Princetonplein 1, 3584 CC, Utrecht, The Netherlands
+		      \and Nikhef, Science Park 105, 1098 XG, Amsterdam, The Netherlands
+		      \and Royal Holloway University of London, London, United Kingdom}
 
@@
 \section{Full posterior distribution for GW150914} \label{sec:appendix_b}
-\begin{figure}
+\begin{figure*}
 	\caption{Full posterior distribution for GW150914 when marginalising over the AR process orders.}
@@
 	\includegraphics[width=\textwidth,keepaspectratio]{Images/gw150914/mega_plot.pdf}
-\end{figure}
+\end{figure*}
 \end{appendix}

### 2024-02-12 paper/mesa.tex
revised title
pruned all errors

@@
 	\sisetup{output-decimal-marker={.}}
-\newtheorem{theorem}{Theorem}
 \usepackage{float}
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE
-	\title{Maximum Entropy Spectral Analysis: a case study}
-	\author{Alessandro \surname{Martini}\inst{1,2,3}\thanks{alessandro.martini-1@unitn.it} \and 
-	Stefano \surname{Schmidt}\inst{4,5}\thanks{s.schmidt@uu.nl} \and Gregory \surname{Ashton}\inst{6} \and Walter \surname{Del Pozzo}\inst{1}\thanks{walter.delpozzo@unipi.it}}
+	\title{Maximum Entropy Spectral Analysis: an application to gravitational waves data analysis}
+	\author{Alessandro~{Martini}\inst{1,2,3}\thanks{alessandro.martini-1@unitn.it} \and 
+	Stefano~{Schmidt}\inst{4,5}\thanks{s.schmidt@uu.nl} \and Gregory~{Ashton}\inst{6} \and Walter~{Del Pozzo}\inst{1}\thanks{walter.delpozzo@unipi.it}}
 
@@
 We remind the reader that a Toeplitz matrix is a matrix in the form:
-\begin{pmatrix}
+$\begin{pmatrix}
 	a_0 & a_1 & a_2 & \ldots & \ldots& \ldots &a_n\\
@@
 	a_{-n} & \ldots & \ldots & \ldots& a_{-2} & a_{-1} & a_0
-\end{pmatrix}
+\end{pmatrix}$
 }.
@@
 \section{Comparison with Welch method}\label{sec:Welch_comparison}
-\todo{Check again the section}
+
 We perform a {\it qualitative} comparison between the performance of the MESA and of the standard Welch algorithm.
@@
 In other words, we care only about maximising the information entropy in the distribution of the residuals.
-
-%Since the order of the AR process is not yet determined by the theory, up to this point, we opted for an in-depth investigation of several proposals in the literature and found that different loss functions are required for different situations, with the FPE loss function being the most indicated to deal with gravitational wave data.
-Along these lines, we directly compared the PSDs computed with MESA with the canonical Welch's algorithm.
-As outlined in Sec.~\ref{sec:Welch_comparison}, MESA provides PSD estimates with smaller variance and better accuracy than Welch algorithm.
-%The use of MESA is particularly useful for short time series samples, where Welch's method is outperformed in both precision and confidence.
-
-%This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD.
-Hence MESA could be employed for {\it online} PSD estimation during a Bayesian parameter estimation exercise.
-%%a simultaneous inference of the PSD and of the gravitational wave signal during a Bayesian parameter estimation exercise, effectively marginalizing over the detector noise PSD.
-%%\todo{There might be a misunderstanding: I guess the referee is talking about taking into account the PSD errors, while we are talking about dropping the stationariety assumption...
-The misleading word is `marginalize', I think}
-%This is possible whenever the log-likelihood of the model for a time series takes the form
-%\begin{equation}
-%	\log\mathcal{L}(d_t | \theta) \propto  -\frac{1}{2} \frac{|\tilde{d}(f)- \tilde{x}(f;\theta)|^2}{S(f)} - \frac{1}{2}\log\left[\int  df\,S(f)\right]
-%\end{equation}
-%where $\tilde{\hphantom{a}}$ denotes the Fourier transform and the signal model $x(t;\theta)$, dependent on some parameter $\theta$, is a prediction for a deterministic signal buried in the observed time series $d(t)$.
-%In GW data analysis, a typical approach is to estimate the PSD offline on a large batch of data and off-source: this scheme assumes {\it stationary data} on a long timescale (longer than the data under study) and it might not reflect the structure of the noise in the analyzed time slice.
-%Some alternatives exist for dropping this assumption and modifying the likelihood accordingly \citep{R_ver_2010, R_ver_2011, Edwards_2020, Chatziioannou_2021}.
-%MESA can add to those an elegant way out: at each evaluation of the likelihood, a new spectral analysis is performed by computing the PSD on the residual $d_t-x_t$.
-%In this way, the PSD would depend also on $\theta$ and would effectively model the residuals.
-For this method, we would only need to assume the stationariety of the residual on the batch to analyse\footnote{
-%Unlike several works in GW data analysis \citep{Biscoveanu_2020, talbot2020gravitationalwave}, this approach does not address the issue of including the PSD uncertainties in the posterior.
-Instead, it focuses on making minimal assumption about the nature of the data under study.
-%}, making a lighter assumption on the nature of the data. Furthermore, this method would get, as a bonus, a posterior distribution for the PSD of the analyzed data.
-%Preliminary studies suggest that this is possible~\citep{martini_thesis}.
-Other studies have been done in this direction, mostly using a parametric model for the PSD \citep{Littenberg_2013, Edwards_2015, lalinference}.
 

### 2024-03-13 memspectrum/memspectrum.py
Added a Loss function

@@ -57,6 +57,8 @@ class loss_function:
 	def __call__(self, *args): #Stefano: what are args? We should specify them and call them by name...
 		if self.method == 'FPE':
 			return self._FPE(args[0], args[2], args[3])
+		if self.method == 'MDL':
+			return self._FPE(args[0], args[2], args[3])
 		elif self.method == 'CAT':
 			return self._CAT(args[0], args[2], args[3])
 		elif self.method == 'CAT_OLD':
@@ -110,6 +112,28 @@ class loss_function:
 		"""
 		return P[-1] * (N + m + 1) / (N - m - 1)
 	
+	def _MDL(self, P, N, m):
+		"""
+		Implements Minimum Description Length to estimate the recursive 
+		order 
+		
+		Parameters
+		----------
+		P : float
+			The estimate of the variance for the white noise component.
+		N : 'np.int'
+			The length of the dataset.
+		m : 'np.int'
+			The recursive order.
+
+		Returns
+		-------
+		float
+			The value of MDL loss function.
+
+		"""
+		return N * np.log(P[-1]) + m * np.log(N)
+    
 	def _AIC(self, P, N, m):
 		"""
 		Implements Akaike information criterion to estimate the recursive 

### 2024-03-13 paper/mesa.tex
Implemented some of Greg's comments on paper

@@
 We provide a ready-to-use implementation of the algorithm in the form of a python package \texttt{memspectrum} and we characterize our implementation by performing a power spectral density analysis on synthetic data (with known power spectral density) and by comparing different criteria for stopping the recursion.
-Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using synthetic data generated from the released spectrum by the LIGO-Virgo-Kagra collaboration.
+Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using synthetic data generated from the \textcolor{blue}{GW150914 strain} spectrum released  by the LIGO-Virgo-Kagra collaboration.
 We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
-systematically lower variance and bias. This is particularly manifest in the case of a little number of data points, making Burg's method most suitable to work in this regime. 
+systematically lower variance and bias, crucial for a better posterior estimate.
+This is particularly manifest in the case of a small ($O(5000)$) number of data points, making Burg's method most suitable to work in this regime
 We conclude our investigation by utilising MESA, and its particularly easy parametrisation where the only free parameter is the order $p$ of the AR process, to marginalise over the interferometers noise PSD in conjunction with inferring the parameters of GW150914.}
@@
 Clearly this is not possible in practice.
-Therefore modelling a stochastic process either relies on modelling of the underlying physical processes, thus falling back onto the deterministic case, or on modelling the mathematical and statistical properties of the process, and potentially infer them from the process realisations.
+Therefore modelling a stochastic process either relies on modelling of the underlying physical processes, thus falling back onto the deterministic case, or on modelling the mathematical and statistical properties of the process, and potentially infering them from the process realisations.
 The study of the properties of stochastic processes is thus a crucial task in many fields of physics, astronomy, quantitative biology, as well as engineering and finance.
@@
 These are stochastic processes that display an invariance of their statistical properties, such as their two-point autocovariance function, with respect to translation of the independent variable, usually the time $t$.
-If $x(t)$ is a wide-sense stationary process, it is in fact completely determined by the knowledge of the (many-points) autocorrelation functions.
+If $x(t)$ is a wide-sense stationary process, its statistical properties are completely determined by the knowledge of the (many-points) autocorrelation functions.
 In practice, one often has easy access to the two-point correlation function
@@
 without highlighting its connection with the time structure of the process itself, thus masking some important properties that will be explored further in what follows.
-The latter definition in~\eqref{eq:psd-f-definition} gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.
+The latter definition in Eq.
+~\eqref{eq:psd-f-definition} gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.
 
-An ubiquitous method for such a computation is due to \citet{Welch1967} and it is based on Eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
+An ubiquitous method for such a computation is due to \citet{Welch1967} and it is based on Eqs. (\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
 The PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
@@
 
-A alternative approach providing a smooth PSD estimation, is to adopt a parametric model for the PSD and to fit its parameters to the data with a Markov Chain Monte Carlo \citep[e.g.]{Cornish_2015, Littenberg_2015}.
+A alternative approach providing a smooth PSD estimation, is to adopt a parametric model for the PSD and to fit its parameters to the data with a Reversible Jump Markov Chain Monte Carlo \citep[e.g.]{Cornish_2015, Littenberg_2015}.
 Despite being effective, this method is problem dependent, since it needs to make definite assumptions on the shape of the PSD.
@@
 In this paper, we discuss the details of the Maximum entropy principle, its application to the problem of PSD estimation with Burg's algorithm and the link between Burg's algorithm and autoregressive process.
-Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many undesired aspects of the Welch's algorithm (or other similar methods).
+Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out of the many undesired aspects of the Welch's algorithm (or other similar methods).
 To facilitate this goal, we based this study on \texttt{memspectrum}, a freely available, robust and easy-to-use python implementation of the algorithm described below\footnote{
@@
 
-Before introducing MAXENT principle, we will define, via some simple examples, the two core concepts of the problem and the roles they play in deductive inference: the `evidence' and the `information'.
+Before introducing the MAXENT principle, we will define, via some simple examples, the two core concepts of the problem and the roles they play in deductive inference: the `evidence' and the `information'.
 Let us start with the `information' (or information entropy): it is a measure of the degree of uncertainty on the outcomes of some experiment and specifies the length of the message necessary to provide a full description of the system under study.
-For instance, consider a perfectly sinusoidal signal: knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information content.
-Even less information is required if we are studying a system whose outcome is certain (the outcome is known with probability $p = 1$), as in this case, a communication is not even needed.
+No information is brought if we are studying a system whose outcome is certain (the outcome is known with probability $p = 1$), as in this case, a communication is not even needed.
 \citet{Shannon} proposed the quantity
@@
 \end{equation}
-We call the functional $H$ (information) entropy\footnote{In defining the information entropy as in Eq.~\eqref{eq:entropy_continuos}, we are implicitly assuming a uniform measure over the parameter space.
+We call the functional $H$ information entropy\footnote{In defining the information entropy as in Eq.~\eqref{eq:entropy_continuos}, we are implicitly assuming a uniform measure over the parameter space.
 In case of a non-uniform measure $m(x)$, the definition generalises to $ H[p(x)] = - \int p(x)\ln \frac{p(x)}{m(x)} dx$.}.
 
-We now turn to the core of our problem: how can we assign probabilities to a set of events keeping into account our knowledge of the system and, at the same time, it is non-committal towards unavailable knowledge?
+We now turn to the core of our problem: how can we assign probabilities to a set of events keeping into account our knowledge of the system and, at the same time, ensure it is non-committal towards unavailable knowledge?
 The ``knowledge'' at our disposal about the system under investigation is what we define `evidence' and any probability assignment is given such evidence, in agreement with \citet{Cox} construction of probability.
@@
 In other words, the information entropy content introduced by the probability assignment should be as large as possible, in accordance with the available evidence.
-MAXENT formalises this reasoning by stating that probabilities should be assigned by maximizing uncertainty (information entropy) using evidence as constraint.
-This defines a variational problem, where the information entropy functional $H\left[p_1, \dots, p_N\right]$, defined in eq.~\eqref{eq:entropy}, has to be maximized.
+MAXENT formalises this reasoning by stating that probabilities should be assigned by maximizing uncertainty (information entropy) using evidence as a constraint.
+This defines a variational problem, where the information entropy functional $H\left[p_1, \dots, p_N\right]$, defined in Eq. ~\eqref{eq:entropy}, has to be maximized.
 
-The maximisation of entropy, supplemented by evidence in the form of constraints to which the sought-for probability distribution must obey, gives rise to several of the most common probability distributions commonly employed in statistics.
-For instance, whenever the only constraint available is the normalization of the probability distribution (i.e.
+The maximisation of the entropy, supplemented by evidence in the form of constraints to which the sought-for probability distribution must obey, gives rise to several of the most common probability distributions commonly employed in statistics.
+\textcolor{red}{For instance, whenever the only constraint available is the normalization of the probability distribution (i.e.
 no evidence is available), the entropy is maximised by the uniform distribution.
-If we have evidence to constraint the expected value, the information entropy is maximised by the exponential distribution.
+If we have evidence to constraint the expected value, the information entropy is maximised by the exponential distribution.}
 
@@
 This derivation is particularly interesting from the foundational point of view, since it provides a deeper insight into the ubiquitous Gaussian distribution.
-Indeed, it is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy for a fixed mean and variance: from the MAXENT principle, it is the correct probability distribution to assign if mean and covariance are the only quantities that fully define our process.
+Indeed, it is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy for a fixed mean and variance: from the MAXENT principle, it is the correct probability distribution to assign if the mean and covariance are the only quantities that fully define our process.
 In some sense, we can interpret the central limit theorem as the natural `statistical' evolution toward a configuration that maximizes entropy in repeated experiments.
@@
 where the matrix $S_{ij} = S_i \delta_{ij}$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
-Many readers will recognise the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method\citep{prob_information_theory}
+Many readers will recognise the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method \citep{prob_information_theory}
 and of gravitational waves data analysis, \citep[e.g.]{Finn_1992, FINDCHIRP}.
@@
 In principle, if the autocorrelation was known exactly (i.e.
-at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD would reduce to a single Fourier transform.
+at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD would reduce to a single Fourier transform (i.e.
+Eq.~\eqref{eq:psd-autocorrelation}).
 However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process, hence such computation is impossible.
+Also, the single periodogram is not a consisten estimator for the power spectral density, since its variance doesn't decrease with the sample size N.
 Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases as $\sigma \sim 1/\sqrt{N - k}$\footnote{
@@
 }, so that only few values for the autocorrelation function can actually be computed reliably.
-This bring us the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function?
+This bring us to the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function?
 MAXENT can guide us in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method.
@@
 The `Standard' method is slower but more stable, while `Fast' trades stability for speed.
-On simulated stationary data, both versions typically yield similar results, while our tests with real gravitational waves data seems to indicate that the `Fast' implementation introduces noise into the PSD estimate with no accuracy improvement.\footnote{For this reason, it is advisable to use the `Standard' implementation whenever possible.
+On simulated stationary data, both versions typically yield similar results, while our tests with real gravitational waves data seems to indicate that the `Fast' implementation introduces noise into the PSD estimate.\footnote{For this reason, it is advisable to use the `Standard' implementation whenever possible.
 In most case of numerical instability in the `Fast' method, \texttt{memspectrum} will send a warning to user.}.
@@
 \item \textbf{Variance Maximum (VM)}
-This second criterion \cite{kay1988modern} is based on a similar assumptions to FPE.
+This second criterion (\cite{kay1988modern}) is based on a similar assumptions to FPE.
 It minimises the actual value of the least squares (instead of relying to asymptotical behaviour).
@@
 \end{equation}
-The package implementation of VM loss function takes advantage of a recursive re-writing of the above formula, as in equations (27) and (28) of \citet{Cuoco_2001}.  \\ \\
+The package implementation of VM loss function takes advantage of a recursive re-writing of the above formula, as in Eqs (27) and (28) of \citet{Cuoco_2001}.  \\ \\
 
-Several other criteria are available in the literature \citet{doi:10.1029/WR018i004p01097} \citet{10.1214/aos/1176349858} and some are implemented in the memspectrum package.
+Several other criteria are available in the literature \citet{doi:10.1029/WR018i004p01097} \citet{bhansali1986} and some are implemented in the memspectrum package.
 We don't report them in this paper since they didn't show any additional merit with respect to the aforementioned loss functions \\
@@
 In the following paragraph, we will discuss three different comparison (one qualitative and two quantitative) of the two proposed loss functions, o
-\subsection{How well the PSD is reconstructed?}\label{sec:psd_validation}
+\subsection{How accurate are the reconstructed PSDs?}\label{sec:psd_validation}
 
@@
 \end{figure}
-In our initial qualitative comparison, depicted in Figure \ref{fig:psd_comparison}, we juxtapose the reconstruction of a known a-priori Power Spectral Density (PSD) with those obtained using the two distinct loss functions.
+In our initial qualitative comparison, depicted in Figure \ref{fig:psd_comparison}, we juxtapose the reconstruction of a known a-priori Power Spectral Density with those obtained using the two distinct loss functions.
 The red line in the plot represents our chosen reference PSD, derived from the estimated PSD for the GW150914 event.\\
 To conduct this analysis, we generated 1000 noise time-series whose power spectral densities matches the reference PSD by construction \citet{noiseGen}.
-The sampling rate and observation time were fixed at $dt = 2048 Hz$ and $T = 5s$, respectively.
+The sampling rate and observation time were fixed at $dt = 2048$Hz and $T = 5$s, respectively.
 For each noise realization, we employed both the FPE and the VM loss functions to estimate the PSD.
@@
 \end{figure}
-In section \ref{sec:whitening}, we showed how autoregressive coefficients and noise variance estimate $P$ can jointly be used to create a whitening filter.
+In Section \ref{sec:whitening}, we showed how autoregressive coefficients and noise variance estimate $P$ can jointly be used to create a whitening filter, as in Eq.
+(\ref{eq:whitening_definition}) and .
 To complete our investigation, we compare how well these whitening filters work when obtained from the two different loss functions we've been studying.
 \\
-For this test, we employed the same set of time-series data as described in paragraph \ref{sec:psd_validation}.
+For this test, we employed the same set of time-series data as described in Section \ref{sec:psd_validation}.
 Each time series underwent the whitening process using the autoregressive filter derived from its corresponding loss function.
@@
 Although similar studies can be drawn from any other PSD, in this section we focus on a single PSD and we try to generalize some observations that we make.
-We decide to use the analytical PSD computed for LIGO Handford interferometer, released together with the GWCT-1 catalog~\citep{GWTC1, PSD_release}, and computed with BayesLine package \citep{Cornish_2015, Littenberg_2015, Cornish_2020, Chatziioannou_2019}.
+We decide to use the analytical PSD computed for the LIGO Handford interferometer, released together with the GWCT-1 catalog~\citep{GWTC1, PSD_release}, and computed with the BayesLine package \citep{Cornish_2015, Littenberg_2015, Cornish_2020, Chatziioannou_2019}.
 
@@
 We set the total observation time $T = 1, 5, 10, 100, 1000 \SI{}{s}$
-For the MESA algorithm, we choose the VM loss function. For the Welch algorithm, we employ a Tukey window with a parameter $\alpha$ 
-equal to 0.4, an overlap fraction of $1/2$ for the segments and a length of 
+For the MESA algorithm, we choose the VM loss function. For the Welch algorithm, we employ a Tukey window with the shape parameter $\alpha$ 
+equal to 0.4 (see scipy documentation), an overlap fraction of $1/2$ for the segments and a length of 
 segments $L = 512, 1024, 2048, 8192, 32768$ points, depending on the observation time.
@@
 
-We also note that MESA converges to the underlying spectrum much faster than Welch's method, providing a better estimate even in the case of short time series.
+We also note that MESA \textcolor{red}{converges (Seems unclear that we are reffering to fig 4, 5?} to the underlying spectrum much faster than Welch's method, providing a better estimate even in the case of short time series.
 Although observed at every frequency, this behaviour is more evident in the low frequency region.
@@
 
-Another important element is the noise of the spectral estimation: we find that the PSD estimation provided by the Welch's method is more noisy (i.e.
+Another important element is the noise of the spectral estimation: we find that the PSD estimation provided by the Welch's method is noisier (i.e.
 has a large number of spurious peaks) compared to the PSD measured with MESA and FPE loss function.
@@
 
-Define the data hypothesis $D$ as the statement that the data $D = S + N$ with $S$ and $N$ some deterministic signal and some noise hypotheses.
+\textcolor{red}{Define} the data hypothesis $D$ as the statement that the data $D = S + N$ with $S$ and $N$ some deterministic signal and some noise hypotheses.
 Typically, in this formulation one is choosing both a functional form for the signal of interest $S\equiv ``h(t;\theta)''$ and some parametric form $f(t)$ for the noise distribution $N\equiv ``n(t) \sim f(t)''$.
@@
 The coherent likelihood is then given by the product of the individual likelihoods.
-As our analysis template, we adopt the fast machine learning based MLGW model\cite{MLGW}, an aligned spin model trained on TEOBResumS\cite{teobresums}, that has been shown to perform well on LVK events detected during O1 and O2\cite{MLGW}.
-Our sampler is a nested sampling algorithm\cite{cpnest} and the specific inference model is implemented as part of \texttt{granite}, a dedicated inference model for ground-based interferometric detectors.
-We compare our results with the combined posterior samples available from GWOSC\cite{GWOSC} and available at \texttt{https://zenodo.org/records/6513631}.
+As our analysis template, we adopt the fast machine learning based MLGW model \cite{MLGW}, an aligned spin model trained on TEOBResumS \cite{teobresums}, that has been shown to perform well on LVK events detected during O1 and O2 \cite{MLGW}.
+Our sampler is a nested sampling algorithm \cite{john_veitch_2020_4109277} and the specific inference model is implemented as part of \texttt{granite}\textcolor{blue}{[?
+add ref]}, a dedicated inference model for ground-based interferometric detectors.
+We compare our results with the combined posterior samples available from GWOSC \cite{GWOSC} and available at \texttt{https://zenodo.org/records/6513631}.
 
 \begin{figure}
-	\caption{Posterior samples for $\mathcal{M}$ and mass ratio $q$ from the LVK(blue) and using \texttt{memspectrum}(red).
+	\caption{Posterior samples for $\mathcal{M}$ and mass ratio $q$ from the LVK (blue) and using \texttt{memspectrum} (red).
 The samples are largely consistent among the two models, with the MESA model providing a more conservative estimate.}
@@
 \begin{figure}
-	\caption{Posterior distributions for AR process orders in the Hanford(red) and Livingston(blue).}
+	\caption{Posterior distributions for AR process orders in the Hanford (red) and Livingston (blue).}
 	\label{fig:gw150914_orders}
@@
 The superior nature of maximum entropy methods, and in particular of Burg's method, is exemplified by the closed form estimate of the power spectral density and by the theoretical bridge between spectral analysis and AR processes.
-Moreover, the method presents, in our view, two main advantages when compared with more traditional ones; first there is no need to choose an arbitrary window function to correct the data and, second it provides as straightforward way to compute predictions given past observations.
+Moreover, the method presents, in our view, two main advantages when compared with more traditional ones; first there is no need to choose an arbitrary window function to correct the data and, second it provides as straightforward way to compute \textcolor{blue}{predictions given past observations.
+Greg asks how can be useful.
+We comment it later on so I would not add on this}.
 Accompanying this work, we provide a publicly available \texttt{Python} implementation, called \texttt{memspectrum}, that we used to perform the numerical studies presented in this work..
@@
 As outlined in Sec.~\ref{sec:Welch_comparison}, MESA provides PSD estimates with smaller variance and better accuracy than Welch algorithm.
-The use of MESA is particularly useful for short time series samples, where Welch's method is outperformed in both precision and confidence.
+The use of MESA is particularly useful for short time series samples (comment not clear to me), where Welch's method is outperformed in both precision and confidence.
 
 This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD.
-By applying MESA to 4 seconds of data in correspondence to GW150914, we demonstrated that it is possible to simultaneously estimate the signal and noise parameters, hence effectively marginalise over the noise PSD, without the need to
+\textcolor{red}{By applying MESA to 4 seconds of data in correspondence to GW150914}, we demonstrated that it is possible to simultaneously estimate the signal and noise parameters, hence effectively marginalise over the noise PSD, without the need to
 \begin{itemize}
@@
 Both items are of particular interest for several reasons that we shall discuss in what follows.
-Several proposals exist in literature attempt to marginalise over the PSD, mostly using a parametric model for the PSD \citep{Littenberg_2013, Edwards_2015, lalinference}.
+Several proposals exist in the literature attempt to marginalise over the PSD, mostly using a parametric model for the PSD \citep{Littenberg_2013, Edwards_2015, lalinference}.
 MAXENT fixes the functional form for us exploiting the correspondence with AR processes, providing a one-parameter family of models that are particularly easy to sample, thus grounding the noise properties marginalisation in solid theoretical foundations and in an easy-to-use numerical implementation.
@@
 MAXENT and MESA model and are relevant \emph{only for the segment of data under consideration}, and make no assumptions over what is not part of the analysis.
-We believe, and we will show in a future study, that using MESA is the only way of computing single-source posteriors whenever multiple sources are overlapping.
+We believe, and we will show in a future study, that using MESA is the \textcolor{blue}{only way (greg says is not True.
+remove "only way?")} of computing single-source posteriors whenever multiple sources are overlapping.
 This is possible since, in our formulation, everything that we did not label as \emph{signal} will be part of the residuals, over which we apply MESA.
@@
 This fact is remarkable and can be used in time series analysis for several purposes.
-As discussed in Sec.~\ref{sec:LIGO_forecasting}, an anomaly detection pipeline could be built using the forecasts of MESA: the predictions can form a baseline to compare the actual observations with.
+As an example, an anomaly detection pipeline could be built using the forecasts of MESA: the predictions can form a baseline to compare the actual observations with.
 Whenever the observed data are outside the expectations, an anomaly detection can be claimed.
@@
 
+\section{Computational Time for both MESA methods and Welch}
+In this appendix we shortly introduce the computational times required by the MESA method (considering both the standard and Fast implementation) and the Welch method.
+They are just inserted to give an idea of what the time differences between the methods are.
+These are obtained via the python \%timeit special function, run on a personal machine
+\begin{table}
+\begin{tabular}{ |p{2cm}||p{3cm}|p{3cm}|p{3cm}|  }
+ \hline
+ \multicolumn{4}{|c|}{Computational times} \\
+ \hline
+ Batch Length& MESA std 2 Code &MESA Fast&Welch\\
+
+ \hline
+ 1s   &  22 ms ± 1.22 ms   &$19.6$ ms ± $620$ µs&  335 µs ± 9.24 µ\\
+ 5s &   158 ms ± 21.7 ms   & 42.4 ms ± 353 µs   &839 µs ± 4.61 µs \\
+ 10s & 187 ms ± 11.6 ms & 51.5 ms ± 3.67 ms&  1.74 ms ± 64.3 µs \\
+ 100s    & 1.96 s ± 338 ms  &205 ms ± 5.09 ms&  18.8 ms ± 140 µs\\
+ 1000s&  17.1 s ± 605 ms  &1.33 s ± 17.4 ms&235 ms ± 3.69 ms\\
+
+ \hline
+\end{tabular}
+\caption{Comparison of the computational times for the estimate of the power spectral densities with our implementation of MESA (both standard and Fast implementations) and Welch's method}
+\label{tb1:ComputationalTimes}
+\end{table}
+
 \section{Full posterior distribution for GW150914} \label{sec:appendix_b}

### 2024-04-03 paper/mesa.tex
Updated paper and bibliography

@@
 We provide a ready-to-use implementation of the algorithm in the form of a python package \texttt{memspectrum} and we characterize our implementation by performing a power spectral density analysis on synthetic data (with known power spectral density) and by comparing different criteria for stopping the recursion.
-Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using synthetic data generated from the \textcolor{blue}{GW150914 strain} spectrum released  by the LIGO-Virgo-Kagra collaboration.
+Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using synthetic data generated from the GW150914 strain spectrum released  by the LIGO-Virgo-Kagra collaboration.
 We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
-systematically lower variance and bias, crucial for a better posterior estimate.
-This is particularly manifest in the case of a small ($O(5000)$) number of data points, making Burg's method most suitable to work in this regime
+systematically lower variance and bias.
+This is particularly manifest in the case of a small ($O(5000)$) number of data points, making Burg's method most suitable to work in this regime.
+Since this is close to the typical length of analysed gravitational waves data, improving the estimate of the PSD in this regime leads to more reliable posterior profiles for the system under study.
 We conclude our investigation by utilising MESA, and its particularly easy parametrisation where the only free parameter is the order $p$ of the AR process, to marginalise over the interferometers noise PSD in conjunction with inferring the parameters of GW150914.}
@@
 Let us start with the `information' (or information entropy): it is a measure of the degree of uncertainty on the outcomes of some experiment and specifies the length of the message necessary to provide a full description of the system under study.
-No information is brought if we are studying a system whose outcome is certain (the outcome is known with probability $p = 1$), as in this case, a communication is not even needed.
+As an example, no information is brought if we are studying a system whose outcome is certain (the outcome is known with probability $p = 1$), as in this case, a communication is not even needed.
 \citet{Shannon} proposed the quantity
@@
 The maximisation of the entropy, supplemented by evidence in the form of constraints to which the sought-for probability distribution must obey, gives rise to several of the most common probability distributions commonly employed in statistics.
-\textcolor{red}{For instance, whenever the only constraint available is the normalization of the probability distribution (i.e.
+In the cases of interest, evidence is used to constraint, via Lagrange Multipliers, the momenta of the probability distribuiton we are seeking to evaluate.
+For instance, whenever the only constraint available is the normalization of the probability distribution (i.e.
 no evidence is available), the entropy is maximised by the uniform distribution.
-If we have evidence to constraint the expected value, the information entropy is maximised by the exponential distribution.}
+If we have evidence to constraint the expected value, the information entropy is maximised by the exponential distribution.
 
@@
 Eq.~\eqref{eq:psd-autocorrelation}).
-However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process, hence such computation is impossible.
-Also, the single periodogram is not a consisten estimator for the power spectral density, since its variance doesn't decrease with the sample size N.
+However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process.
+In such cases, the single periodogram is not a consistent estimator for the power spectral density, since its variance doesn't decrease when the sample size increases.
 Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases as $\sigma \sim 1/\sqrt{N - k}$\footnote{
@@
 The `Standard' method is slower but more stable, while `Fast' trades stability for speed.
-On simulated stationary data, both versions typically yield similar results, while our tests with real gravitational waves data seems to indicate that the `Fast' implementation introduces noise into the PSD estimate.\footnote{For this reason, it is advisable to use the `Standard' implementation whenever possible.
+On simulated stationary data, both versions typically yield similar results, while our tests with real gravitational waves data seems to indicate that the `Fast' implementation introduces noise into the PSD estimate.
+\footnote{For this reason, it is advisable to use the `Standard' implementation whenever possible.
 In most case of numerical instability in the `Fast' method, \texttt{memspectrum} will send a warning to user.}.
+A comparison of the computational times for Standard MESA implementation and Fast implementation (together with Welch's) is porvided in appendix \ref{sec:computationaltimes}.
 
@@
 For the Welch algorithm, we use the standard implementation provided by the python library \texttt{scipy}~\citep{numpy,scipy}.
-The results from both methods are summarized in  Figs.~\ref{fig:MESA_LIGO_data} and~\ref{fig:welch_LIGO_data} respectively.
+The results from both methods are summarized in  Figures ~\ref{fig:MESA_LIGO_data} and~\ref{fig:welch_LIGO_data} respectively.
 
@@
 
-We also note that MESA \textcolor{red}{converges (Seems unclear that we are reffering to fig 4, 5?} to the underlying spectrum much faster than Welch's method, providing a better estimate even in the case of short time series.
+We also note that MESA converges (Figures ~\ref{fig:MESA_LIGO_data} and~\ref{fig:welch_LIGO_data}) to the underlying spectrum much faster than Welch's method, providing a better estimate even in the case of short time series.
 Although observed at every frequency, this behaviour is more evident in the low frequency region.
@@
 As outlined in Sec.~\ref{sec:Welch_comparison}, MESA provides PSD estimates with smaller variance and better accuracy than Welch algorithm.
-The use of MESA is particularly useful for short time series samples (comment not clear to me), where Welch's method is outperformed in both precision and confidence.
+The use of MESA is particularly useful for short time series samples, where Welch's method is outperformed in both precision and confidence.
+As an examples, Figures ~\ref{fig:MESA_LIGO_data} and~\ref{fig:welch_LIGO_data} illustrate that MESA's performance over a 10-second interval is more closely aligned with Welch's performance over a 100-second interval than Welch's performance over a 10-second interval alone.
 
@@
 MAXENT and MESA model and are relevant \emph{only for the segment of data under consideration}, and make no assumptions over what is not part of the analysis.
-We believe, and we will show in a future study, that using MESA is the \textcolor{blue}{only way (greg says is not True.
-remove "only way?")} of computing single-source posteriors whenever multiple sources are overlapping.
+We believe, and we will show in a future study, that using MESA can be a natural solution for computing single-source posteriors whenever multiple sources are overlapping.
 This is possible since, in our formulation, everything that we did not label as \emph{signal} will be part of the residuals, over which we apply MESA.
@@
 
-\section{Computational Time for both MESA methods and Welch}
+\section{Computational Time for both MESA methods and Welch}\label{sec:computationaltimes}
 In this appendix we shortly introduce the computational times required by the MESA method (considering both the standard and Fast implementation) and the Welch method.
@@
 These are obtained via the python \%timeit special function, run on a personal machine
-\begin{table}
+\begin{table}[H]
 \begin{tabular}{ |p{2cm}||p{3cm}|p{3cm}|p{3cm}|  }

### 2024-07-01 paper/mesa.tex
Modified paper for EPJ

@@
-\documentclass{aa}
-\bibpunct{(}{)}{;}{a}{}{,}
+\documentclass[epj,nopacs]{svjour}
 \usepackage[varg]{txfonts}
+\let\iint\relax
+\let\iiint\relax
+\let\iiiint\relax
+\let\idotsint\relax
 \usepackage{amsmath,amssymb}
+\usepackage{amssymb}
 \usepackage{amsfonts}
@@
 \usepackage{ragged2e}
-\usepackage{cleveref}
+%\usepackage{cleveref}
 \usepackage{blindtext}
@@
 \usepackage{float}
+%\usepackage{natbib}
 
 	%some math symbols
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
+%\newcommand{\R}{\mathbb{R}}
+%\newcommand{\N}{\mathbb{N}}
 \DeclareMathOperator{\sign}{sign}
-\renewcommand{\d}[1]{\ensuremath{\operatorname{d}\!{#1}}}
+%\renewcommand{\d}[1]{\ensuremath{\operatorname{d}\!{#1}}}
 %argmin and argmax
@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+\usepackage{cite}
 
 % comments command
-\newcommand{\wdp}[1]{{\textcolor{magenta}{\texttt{walter: #1}} }}
-\newcommand{\amartini}[1]{{\textcolor{blue}{\texttt{alessandro: #1}} }}
-\newcommand{\sschmidt}[1]{{\textcolor{purple}{\texttt{stefano: #1}} }}
-\newcommand{\todo}[1]{{\textcolor{red}{\texttt{TODO: #1}} }}
+%\newcommand{\wdp}[1]{{\textcolor{magenta}{\texttt{walter: #1}} }}
+%\newcommand{\amartini}[1]{{\textcolor{blue}{\texttt{alessandro: #1}} }}
+%\newcommand{\sschmidt}[1]{{\textcolor{purple}{\texttt{stefano: #1}} }}
+%\newcommand{\todo}[1]{{\textcolor{red}{\texttt{TODO: #1}} }}
 
@@
 	\author{Alessandro~{Martini}\inst{1,2,3}\thanks{alessandro.martini-1@unitn.it} \and 
-	Stefano~{Schmidt}\inst{4,5}\thanks{s.schmidt@uu.nl} \and Gregory~{Ashton}\inst{6} \and Walter~{Del Pozzo}\inst{1}\thanks{walter.delpozzo@unipi.it}}
+	Stefano~{Schmidt}\inst{4,5}\thanks{s.schmidt@uu.nl} \and Gregory~{Ashton}\inst{6} \thanks{Gregory.Ashton@rhul.ac.uk}\and Walter~{Del Pozzo}\inst{1}\thanks{walter.delpozzo@unipi.it}}
 
@@
 \abstract{
-The Maximum Entropy Spectral Analysis (MESA) method, developed by Burg, provides a powerful tool to perform spectral estimation of a time-series.
+The Maximum Entropy Spectral Analysis (MESA) method, developed by Burg, offers a powerful tool for spectral estimation of a time-series.
+It relies on Jaynes’ maximum entropy principle, allowing the spectrum of a stochastic process to be inferred using the coefficients of an autoregressive process AR($p$) of order $p$.
+A closed-form recursive solution provides estimates for both the autoregressive coefficients and the order $p$ of the process.
+We provide a ready-to-use implementation of this algorithm in a Python package called \texttt{memspectrum}, characterized through power spectral density (PSD) analysis on synthetic data with known PSD and comparisons of different criteria for stopping the recursion.
+Additionally, we compare the performance of our implementation with the ubiquitous Welch algorithm, using synthetic data generated from the GW150914 strain spectrum released by the LIGO-Virgo-Kagra collaboration.
+Our findings indicate that Burg’s method provides PSD estimates with systematically lower variance and bias.
+This is particularly manifest in the case of a small (O($5000$)) number of data points, making Burg’s method most suitable to work in this regime.
+Since this is close to the typical length of analysed gravitational waves data, improving
+the estimate of the PSD in this regime leads to more reliable posterior profiles for the system under study. We conclude
+our investigation by utilising MESA, and its particularly easy parametrisation where the only free parameter is the order
+p of the AR process, to marginalise over the interferometers noise PSD in conjunction with inferring the parameters of
+GW150914 
+}
+
+%  \abstract{
+%The Maximum Entropy Spectral Analysis (MESA) method, developed by Burg, provides a powerful tool to perform spectral estimation of a time-series.
 The method relies on Jaynes' maximum entropy principle and provides the means of inferring the spectrum of a stochastic process in terms of the coefficients of some autoregressive process AR($p$) of order $p$.
-A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
-We provide a ready-to-use implementation of the algorithm in the form of a python package \texttt{memspectrum} and we characterize our implementation by performing a power spectral density analysis on synthetic data (with known power spectral density) and by comparing different criteria for stopping the recursion.
+%A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
+%We provide a ready-to-use implementation of the algorithm in the form of a python package \texttt{memspectrum} and we characterize our implementation by performing a power spectral density analysis on synthetic data (with known power spectral density) and by comparing different criteria for stopping the recursion.
 Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using synthetic data generated from the GW150914 strain spectrum released  by the LIGO-Virgo-Kagra collaboration.
-We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
-systematically lower variance and bias.
+%We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
+%systematically lower variance and bias.
 This is particularly manifest in the case of a small ($O(5000)$) number of data points, making Burg's method most suitable to work in this regime.
 Since this is close to the typical length of analysed gravitational waves data, improving the estimate of the PSD in this regime leads to more reliable posterior profiles for the system under study.
-We conclude our investigation by utilising MESA, and its particularly easy parametrisation where the only free parameter is the order $p$ of the AR process, to marginalise over the interferometers noise PSD in conjunction with inferring the parameters of GW150914.}
+%We conclude our investigation by utilising MESA, and its particularly easy parametrisation where the only free parameter is the order $p$ of the AR process, to marginalise over the interferometers noise PSD in conjunction with inferring the parameters of GW150914.}
 
-\titlerunning{MESA: a case study}
-\authorrunning{Martini, Schmidt, Del Pozzo}
+\titlerunning{MESA: an application to gravitational waves data analysis}
+\authorrunning{Martini, Schmidt, Ashton, Del Pozzo}
 
@@
 \end{equation}
-In the context of gravitational waves physics, e.g. ~\citep{Finn_1992}, the PSD is introduced as  
+In the context of gravitational waves physics, e.g. \cite{Finn_1992}, the PSD is introduced as  
 \begin{equation}\label{eq:psd-f-definition}
@@
 
-An ubiquitous method for such a computation is due to \citet{Welch1967} and it is based on Eqs. (\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
+An ubiquitous method for such a computation is due to \cite{Welch1967} and it is based on Eqs. (\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
 The PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
-This approach is equivalent~\citep{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
+This approach is equivalent~\cite{Lomb}\cite{Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
 \begin{equation}
@@
 
-A alternative approach providing a smooth PSD estimation, is to adopt a parametric model for the PSD and to fit its parameters to the data with a Reversible Jump Markov Chain Monte Carlo \citep[e.g.]{Cornish_2015, Littenberg_2015}.
+A alternative approach providing a smooth PSD estimation, is to adopt a parametric model for the PSD and to fit its parameters to the data with a Reversible Jump Markov Chain Monte Carlo \cite{Cornish_2015,Littenberg_2015}.
 Despite being effective, this method is problem dependent, since it needs to make definite assumptions on the shape of the PSD.
@@
 
-An appealing alternative, based on the Maximum Entropy principle \citep{JaynesArticle,jaynes2003ptl, Jaynes_MAXENT}, has been derived by~\citet{burg1975maximum}.
+An appealing alternative, based on the Maximum Entropy principle \cite{JaynesArticle, jaynes2003ptl, Jaynes_MAXENT}, has been derived by~\cite{burg1975maximum}.
 Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the algorithm parameters, since it provides an iterative closed form expression for the spectrum of a stochastic stationary time series.
@@
 Its domain of application turned out to be wider than expected.
-In fact, thanks to~\citet{burg1975maximum}, this method has also been applied to perform high quality computation of power spectral densities of time series.
+In fact, thanks to~\cite{burg1975maximum}, this method has also been applied to perform high quality computation of power spectral densities of time series.
 
@@
 As an example, no information is brought if we are studying a system whose outcome is certain (the outcome is known with probability $p = 1$), as in this case, a communication is not even needed.
-\citet{Shannon} proposed the quantity
+\cite{Shannon} proposed the quantity
 \begin{equation}\label{eq:information}
@@
 
-\citet{Shannon} showed that the only functional form satisfying continuity with respect to its parameters, additivity and that has a maximum for equal probability events is:
+\cite{Shannon} showed that the only functional form satisfying continuity with respect to its parameters, additivity and that has a maximum for equal probability events is:
 \begin{equation}\label{eq:entropy}
@@
 We now turn to the core of our problem: how can we assign probabilities to a set of events keeping into account our knowledge of the system and, at the same time, ensure it is non-committal towards unavailable knowledge?
-The ``knowledge'' at our disposal about the system under investigation is what we define `evidence' and any probability assignment is given such evidence, in agreement with \citet{Cox} construction of probability.
+The ``knowledge'' at our disposal about the system under investigation is what we define `evidence' and any probability assignment is given such evidence, in agreement with \cite{Cox} construction of probability.
 In the case above, our knowledge on the system is only the total number $N$ of different outcomes -- this is a minimal requirement.
@@
 It is very common that the constraints provided by the evidence are not enough for setting the probabilities for each event: in this case, it is reasonable to assume that the probability assignment should make the experiment as unpredictable as possible\footnote{
-In~\citet{Jaynes_MAXENT} this statement is made more precise and justified more thoroughly, with arguments based on combinatorial analysis.
+In~\cite{Jaynes_MAXENT} this statement is made more precise and justified more thoroughly, with arguments based on combinatorial analysis.
 }.
@@
 For this work, we are especially interested in the multi-dimensional case.
-Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realization of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of autocovariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution~\citep{gregory_2005}:
+Suppose we have a vector of measurements $(x(t_1),\ldots,x(t_n)) = (x_1, \ldots, x_n)$ that we conveniently express as a single realization of an unknown stochastic process $x(t)$ and we have information about the expectation value of the process $\mu(t)$ and on the matrix of autocovariances $C_{ij} \equiv C(t_i,t_j)$, then the MAXENT distribution is the $n$-dimensional multivariate Gaussian distribution~\cite{gregory_2005}:
 \begin{align}
@@
 }.
-Toeplitz matrices are asymptotically equivalent to circulant matrices and thus diagonalized by the discrete Fourier transform base~\citep{Gray}.
+Toeplitz matrices are asymptotically equivalent to circulant matrices and thus diagonalized by the discrete Fourier transform base~\cite{Gray}.
 %R. M. Gray, Foundations and Trends⃝R in Communications and Information Theory 2, 155 (2005), URL https://doi.org/10.1561/0100000006.
@@
 where the matrix $S_{ij} = S_i \delta_{ij}$ is an $n\times n$ diagonal matrix whose elements are the PSD $S(f)$ calculated at frequency $f_i$.
-Many readers will recognise the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method \citep{prob_information_theory}
-and of gravitational waves data analysis, \citep[e.g.]{Finn_1992, FINDCHIRP}.
+Many readers will recognise the familiar form of the Whittle likelihood that stands at the basis of the \emph{matched filter} method \cite{prob_information_theory}
+and of gravitational waves data analysis, \cite[e.g.]{Finn_1992,FINDCHIRP}.
 Thanks to MAXENT, the problem of defining the probability distribution describing a wide-sense stationary process is thus 
@@
 Before doing so, there is a technicality to solve: the definition of entropy depends on a probability distribution, not on the PSD.
-It can be shown ~\citep[e.g.]{AblesMESA, Bartlett} that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of the filtering a white noise process using a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
+It can be shown ~\cite[e.g.]{AblesMESA,Bartlett} that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of the filtering a white noise process using a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
 A filter with transfer function $T(f)$ takes in input a time series $x_t$ and outputs a times series $y_t$ such that:
@@
 Remarkably, the variational problem admits a closed-form analytical expression for $S(f)$.
-The expression was first found by ~\citet{burg1975maximum}:
+The expression was first found by ~\cite{burg1975maximum}:
 \begin{equation}\label{eq:MESApsd}
@@
 The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the study
-of autoregressive processes (AR)~\citep{doi:10.1029/RG013i001p00183}.
+of autoregressive processes (AR)~\cite{doi:10.1029/RG013i001p00183}.
 An autoregressive stationary process of order $p$, AR($p$), is a time series whose values satisfy the following expression: 
@@
 
-Thanks to Wold's theorem~\citep{Wold_theorem}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general; it turns out that the maximum entropy principle provides a representation of the time series as an $AR(p)$ process and Burg's algorithm computes the corresponding autoregressive coefficients that are suitable to model the available data.
+Thanks to Wold's theorem~\cite{Wold_theorem}, every stationary time series can be represented as an autoregressive process: this ensures that maximum entropy estimation is faithful and general; it turns out that the maximum entropy principle provides a representation of the time series as an $AR(p)$ process and Burg's algorithm computes the corresponding autoregressive coefficients that are suitable to model the available data.
 
@@
 
-Guided from numerical experiments, an indication on the upper bound to the autoregressive order $M_{max}$ is~\citep{doi:10.1190/1.1440902}:
+Guided from numerical experiments, an indication on the upper bound to the autoregressive order $M_{max}$ is~\cite{doi:10.1190/1.1440902}:
 \begin{equation}\label{eq:MMAx}
@@
 \item \textbf{Final prediction Error} 
-The first criterion is due to ~\citet{Akaike1970StatisticalPI}. It was proposed that $m$ should be chosen as the 
+The first criterion is due to ~\cite{Akaike1970StatisticalPI}. It was proposed that $m$ should be chosen as the 
 length that minimizes the error when the filter is used as a predictor, the \emph{final prediction} error (FPE): 
@@
 \item \textbf{Variance Maximum (VM)}
-This second criterion (\cite{kay1988modern}) is based on a similar assumptions to FPE.
+This second criterion \cite{kay1988modern} is based on a similar assumptions to FPE.
 It minimises the actual value of the least squares (instead of relying to asymptotical behaviour).
@@
 \end{equation}
-The package implementation of VM loss function takes advantage of a recursive re-writing of the above formula, as in Eqs (27) and (28) of \citet{Cuoco_2001}.  \\ \\
+The package implementation of VM loss function takes advantage of a recursive re-writing of the above formula, as in Eqs (27) and (28) of \cite{Cuoco_2001}.  \\ \\
 
-Several other criteria are available in the literature \citet{doi:10.1029/WR018i004p01097} \citet{bhansali1986} and some are implemented in the memspectrum package.
+Several other criteria are available in the literature \cite{doi:10.1029/WR018i004p01097,bhansali1986} and some are implemented in the memspectrum package.
 We don't report them in this paper since they didn't show any additional merit with respect to the aforementioned loss functions \\
@@
 
-Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson recursion~\citep{doi:10.1002/sapm1946251261} until $M_{max}$, as given in Eq.~\eqref{eq:MMAx}, iterations are reached.
+Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson recursion~\cite{doi:10.1002/sapm1946251261} until $M_{max}$, as given in Eq.~\eqref{eq:MMAx}, iterations are reached.
 Then, the order $m$ is selected to be the one that minimizes the specified loss function.
@@
 The red line in the plot represents our chosen reference PSD, derived from the estimated PSD for the GW150914 event.\\
-To conduct this analysis, we generated 1000 noise time-series whose power spectral densities matches the reference PSD by construction \citet{noiseGen}.
+To conduct this analysis, we generated 1000 noise time-series whose power spectral densities matches the reference PSD by construction \cite{noiseGen}.
 The sampling rate and observation time were fixed at $dt = 2048$Hz and $T = 5$s, respectively.
@@
 VM performs slightly worse than FPE in this range, overestimating complexity and showing larger error bars.\\
-Moving into the second region (m > 1600), a shift in performance becomes apparent.
+Moving into the second region ($m > 1600$), a shift in performance becomes apparent.
 FPE's estimates tend to stabilize at a certain autoregressive value.
@@
 Although similar studies can be drawn from any other PSD, in this section we focus on a single PSD and we try to generalize some observations that we make.
-We decide to use the analytical PSD computed for the LIGO Handford interferometer, released together with the GWCT-1 catalog~\citep{GWTC1, PSD_release}, and computed with the BayesLine package \citep{Cornish_2015, Littenberg_2015, Cornish_2020, Chatziioannou_2019}.
+We decide to use the analytical PSD computed for the LIGO Handford interferometer, released together with the GWCT-1 catalog~\cite{GWTC1,PSD_release}, and computed with the BayesLine package \cite{Cornish_2015,Littenberg_2015,Cornish_2020,Chatziioannou_2019}.
 
@@
 In all cases, the sampling rate is set to $\SI{4096}{Hz}$.
-For the Welch algorithm, we use the standard implementation provided by the python library \texttt{scipy}~\citep{numpy,scipy}.
+For the Welch algorithm, we use the standard implementation provided by the python library \texttt{scipy}~\cite{numpy,scipy}.
 The results from both methods are summarized in  Figures ~\ref{fig:MESA_LIGO_data} and~\ref{fig:welch_LIGO_data} respectively.
@@
 
-\textcolor{red}{Define} the data hypothesis $D$ as the statement that the data $D = S + N$ with $S$ and $N$ some deterministic signal and some noise hypotheses.
+Define the data hypothesis $D$ as the statement that the data $D = S + N$ with $S$ and $N$ some deterministic signal and some noise hypotheses.
 Typically, in this formulation one is choosing both a functional form for the signal of interest $S\equiv ``h(t;\theta)''$ and some parametric form $f(t)$ for the noise distribution $N\equiv ``n(t) \sim f(t)''$.
@@
 As our analysis template, we adopt the fast machine learning based MLGW model \cite{MLGW}, an aligned spin model trained on TEOBResumS \cite{teobresums}, that has been shown to perform well on LVK events detected during O1 and O2 \cite{MLGW}.
-Our sampler is a nested sampling algorithm \cite{john_veitch_2020_4109277} and the specific inference model is implemented as part of \texttt{granite}\textcolor{blue}{[?
-add ref]}, a dedicated inference model for ground-based interferometric detectors.
-We compare our results with the combined posterior samples available from GWOSC \cite{GWOSC} and available at \texttt{https://zenodo.org/records/6513631}.
+Our sampler is a nested sampling algorithm \cite{john_veitch_2020_4109277} and the specific inference model is implemented as part of \texttt{granite}, a dedicated inference model for ground-based interferometric detectors.
+We compare our results with the combined posterior samples available from GWOSC \cite{GWOSC} and available at \linebreak
+\texttt{https://zenodo.org/records/6513631}.
 
@@
 	\label{fig:gw150914_masses}
-	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/masses.pdf}
+	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/masses.png}
 \end{figure}
@@
 	\label{fig:gw150914_sky}
-	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/sky_position.pdf}
+	\includegraphics[width=0.5\textwidth,keepaspectratio]{Images/gw150914/sky_position.png}
 \end{figure}
@@
 The superior nature of maximum entropy methods, and in particular of Burg's method, is exemplified by the closed form estimate of the power spectral density and by the theoretical bridge between spectral analysis and AR processes.
-Moreover, the method presents, in our view, two main advantages when compared with more traditional ones; first there is no need to choose an arbitrary window function to correct the data and, second it provides as straightforward way to compute \textcolor{blue}{predictions given past observations.
-Greg asks how can be useful.
-We comment it later on so I would not add on this}.
+Moreover, the method presents, in our view, two main advantages when compared with more traditional ones; first there is no need to choose an arbitrary window function to correct the data and, second it provides as straightforward way to compute predictions given past observations.
 Accompanying this work, we provide a publicly available \texttt{Python} implementation, called \texttt{memspectrum}, that we used to perform the numerical studies presented in this work..
@@
 This observation suggests a promising avenue to pursue in future developments of gravitational waves data analysis: for short time series, comparable with the length of binary black hole systems as observed by LIGO, Virgo and KAGRA, the computational cost of MESA is moderate and the inferred PSD is an accurate representation of the true underlying PSD.
-\textcolor{red}{By applying MESA to 4 seconds of data in correspondence to GW150914}, we demonstrated that it is possible to simultaneously estimate the signal and noise parameters, hence effectively marginalise over the noise PSD, without the need to
+By applying MESA to 4 seconds of data in correspondence to GW150914, we demonstrated that it is possible to simultaneously estimate the signal and noise parameters, hence effectively marginalise over the noise PSD, without the need to
 \begin{itemize}
@@
 Both items are of particular interest for several reasons that we shall discuss in what follows.
-Several proposals exist in the literature attempt to marginalise over the PSD, mostly using a parametric model for the PSD \citep{Littenberg_2013, Edwards_2015, lalinference}.
+Several proposals exist in the literature attempt to marginalise over the PSD, mostly using a parametric model for the PSD \cite{Littenberg_2013,Edwards_2015, lalinference}.
 MAXENT fixes the functional form for us exploiting the correspondence with AR processes, providing a one-parameter family of models that are particularly easy to sample, thus grounding the noise properties marginalisation in solid theoretical foundations and in an easy-to-use numerical implementation.
@@
 At the same time, several instruments present gaps in their data stream, for instance LISA is expected to show such gaps~(e.g.
-\citet{lisa_gaps} and references therein), MESA forecasting capabilities could be used to fill those gaps with predicted data from past observations.
+\cite{lisa_gaps} and references therein), MESA forecasting capabilities could be used to fill those gaps with predicted data from past observations.
 In conclusion, we reiterate that MESA is a theoretically sound, computationally feasible and reliable way of studying the properties of stochastic processes and we hope that the investigations presented in this work will further stimulate developments and applications of this method.
@@
 
-	\bibliographystyle{aa}
-	\bibliography{Bibliography.bib}
+\bibliographystyle{spphys2}
+\bibliography{Bibliography} % The name of your .bib file without the extension
 
@@
 \subsection{MESA solution}\label{sec:MESA_solution}
-We derive the expression for the MAXENT spectral estimator following the approach proposed by \citet{burg1975maximum}.
+We derive the expression for the MAXENT spectral estimator following the approach proposed by \cite{burg1975maximum}.
 Unlike the standard approach, we do not enforce the constraints in Eq.~\eqref{eq:MaxConstraint} with the standard Lagrange Multipliers approach.
@@
 It is easy to show that if $z_0$ is a root for the polynomial $(z_0^*)^{-1}$ is also a root: for every root laying outside the unit circle there will be another root inside of it and vice-versa.
-These properties allow us to rewrite the Fourier expansion \eqref{eq:zExp} as \citep{1975STIN...7714318B}:
+These properties allow us to rewrite the Fourier expansion \eqref{eq:zExp} as \cite{1975STIN...7714318B}:
 \begin{equation}\label{eq:MESApsd_appendix}
@@
 The solution of the Eqs.~(\ref{eq:errorFilter1}-\ref{eq:errorFilter2}) fully determines the functional form of the power spectral density estimator \eqref{eq:MESApsd_appendix}.
-The method for solving the equations is called the Levinson-Durbin recursion \citep{doi:10.1002/sapm1946251261} and it is described in the following.
+The method for solving the equations is called the Levinson-Durbin recursion \cite{doi:10.1002/sapm1946251261} and it is described in the following.
 For each order $N$ of the iteration we define the quantities:
@@
 Burg's method for spectral analysis is solved via Levinson is implemented in the released \texttt{memspectrum} package.
-Another faster recursion method is available in \citet{Vos} and it is also available in \texttt{memspectrum}.
+Another faster recursion method is available in \cite{Vos} and it is also available in \texttt{memspectrum}.
 
@@
  \hline
- Batch Length& MESA std 2 Code &MESA Fast&Welch\\
+ Batch Length& MESA std &MESA Fast&Welch\\
 

### 2024-07-01 paper/mesa.tex
Added bibliographic reference

@@
 
-Hence, as an application of MESA, and its implementation in \texttt{memspectrum}, we analyse GW150914\cite{} using a Bayesian framework that allows for the marginalisation of the order  $k$ of the AR($k$) process representing the residuals data stream.
+Hence, as an application of MESA, and its implementation in \texttt{memspectrum}, we analyse GW150914\cite{gw150914} using a Bayesian framework that allows for the marginalisation of the order  $k$ of the AR($k$) process representing the residuals data stream.
 Although the inference is essentially unchanged compared to the standard case, see \cite{lalinference}, there are some substantial modification to the likelihood construction.

### 2024-07-01 paper/mesa.tex
Changed tabular style

@@
 \begin{table}[H]
-\begin{tabular}{ |p{2cm}||p{3cm}|p{3cm}|p{3cm}|  }
+\begin{tabular}{ |p{1.5cm}||p{2cm}|p{2cm}|p{2cm}|  }
  \hline
@@
  \hline
- 1s   &  22 ms ± 1.22 ms   &$19.6$ ms ± $620$ µs&  335 µs ± 9.24 µ\\
- 5s &   158 ms ± 21.7 ms   & 42.4 ms ± 353 µs   &839 µs ± 4.61 µs \\
- 10s & 187 ms ± 11.6 ms & 51.5 ms ± 3.67 ms&  1.74 ms ± 64.3 µs \\
- 100s    & 1.96 s ± 338 ms  &205 ms ± 5.09 ms&  18.8 ms ± 140 µs\\
- 1000s&  17.1 s ± 605 ms  &1.33 s ± 17.4 ms&235 ms ± 3.69 ms\\
+ 1s   &  22  ± 1.22 ms   &$19.6$ ± $0.62$ ms &  335 ± 9.24 µs\\
+ 5s &   158  ± 21.7 ms   & 42.4 ± 0.35 ms   &839 ± 4.61 µs \\
+ 10s & 187  ± 11.6 ms & 51.5 ± 3.67 ms&  1.74  ± 0.06 µs \\
+ 100s    & 1.96 ± 0.34 ms  &205  ± 5.09 &  18.8 ± 0.14 ms\\
+ 1000s&  17.1  ± 0.61 ms  &1.33  ± 0.02 ms &235 ± 3.69 ms\\
 

### 2024-08-10 paper/mesa.tex
Implemented some reviewer corrections

@@
 
-An ubiquitous method for such a computation is due to \cite{Welch1967} and it is based on Eqs. (\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
+An ubiquitous method for such a computation is due to Welch \cite{Welch1967} and it is based on Eqs. (\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
 The PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
@@
 
-An appealing alternative, based on the Maximum Entropy principle \cite{JaynesArticle, jaynes2003ptl, Jaynes_MAXENT}, has been derived by~\cite{burg1975maximum}.
+An appealing alternative, based on the Maximum Entropy principle \cite{JaynesArticle, jaynes2003ptl, Jaynes_MAXENT}, has been derived by Burg ~\cite{burg1975maximum}.
 Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the algorithm parameters, since it provides an iterative closed form expression for the spectrum of a stochastic stationary time series.
@@
 As an example, no information is brought if we are studying a system whose outcome is certain (the outcome is known with probability $p = 1$), as in this case, a communication is not even needed.
-\cite{Shannon} proposed the quantity
+Shannon \cite{Shannon}  proposed the quantity
 \begin{equation}\label{eq:information}
@@
 
-\cite{Shannon} showed that the only functional form satisfying continuity with respect to its parameters, additivity and that has a maximum for equal probability events is:
+Shannon \cite{Shannon} showed that the only functional form satisfying continuity with respect to its parameters, additivity and that has a maximum for equal probability events is:
 \begin{equation}\label{eq:entropy}
@@
 We now turn to the core of our problem: how can we assign probabilities to a set of events keeping into account our knowledge of the system and, at the same time, ensure it is non-committal towards unavailable knowledge?
-The ``knowledge'' at our disposal about the system under investigation is what we define `evidence' and any probability assignment is given such evidence, in agreement with \cite{Cox} construction of probability.
+The ``knowledge'' at our disposal about the system under investigation is what we define `evidence' and any probability assignment is given such evidence, in agreement with Cox \cite{Cox} construction of probability.
 In the case above, our knowledge on the system is only the total number $N$ of different outcomes -- this is a minimal requirement.
@@
 This second criterion \cite{kay1988modern} is based on a similar assumptions to FPE.
-It minimises the actual value of the least squares (instead of relying to asymptotical behaviour).
-t the normalising factor takes into account the k degrees of freedom necessary to estimate the forward prediction error filter $a_k$.\\ The quantity to be minimised is
+It minimises the actual value of the least squares (instead of relying to asymptotical behaviour), using a normalising factor that takes into account the m degrees of freedom necessary to estimate the forward prediction error filter $a_k$.\\ The quantity to be minimised is
 \begin{equation}
@@
 
-Several other criteria are available in the literature \cite{doi:10.1029/WR018i004p01097,bhansali1986} and some are implemented in the memspectrum package.
+Several other criteria are available in the literature \cite{doi:10.1029/WR018i004p01097,bhansali1986} and some are implemented in the \texttt{memspectrum} package.
 We don't report them in this paper since they didn't show any additional merit with respect to the aforementioned loss functions \\
@@
 The following sections will be devoted to the study of the statistical properties of the loss functions introduced above: we need to understand which choice provides the best quality in the reproduction of some known power spectral densities.
-In the following paragraph, we will discuss three different comparison (one qualitative and two quantitative) of the two proposed loss functions, o
+In the following paragraph, we will discuss three different comparison (one qualitative and two quantitative) of the two proposed loss functions.
 \subsection{How accurate are the reconstructed PSDs?}\label{sec:psd_validation}
@@
 	\includegraphics[width = \linewidth]{Images/Plots/comparison_loss2}
+	\caption{Comparison of the ensemble average PSD estimate}
 	\label{fig:psd_comparison}
-	\caption{Comparison of the ensemble average PSD estimate}
 \end{figure}
@@
 To conduct this analysis, we generated 1000 noise time-series whose power spectral densities matches the reference PSD by construction \cite{noiseGen}.
-The sampling rate and observation time were fixed at $dt = 2048$Hz and $T = 5$s, respectively.
+The sampling rate and observation time were fixed at $df = 2048$Hz and $T = 5$s, respectively.
 For each noise realization, we employed both the FPE and the VM loss functions to estimate the PSD.
@@
 \\
-Here, the memspectrum package proves quite useful.
+Here, the \texttt{memspectrum} package proves quite useful.
 It allows us to assign a specific order to the reconstructed autoregressive filter and use the resulting coefficients to forecast time series.
@@
 Although similar studies can be drawn from any other PSD, in this section we focus on a single PSD and we try to generalize some observations that we make.
-We decide to use the analytical PSD computed for the LIGO Handford interferometer, released together with the GWCT-1 catalog~\cite{GWTC1,PSD_release}, and computed with the BayesLine package \cite{Cornish_2015,Littenberg_2015,Cornish_2020,Chatziioannou_2019}.
+We decide to use the analytical PSD computed for the LIGO Handford interferometer, released together with the GWTC-1 catalog~\cite{GWTC1,PSD_release}, and computed with the BayesLine package \cite{Cornish_2015,Littenberg_2015,Cornish_2020,Chatziioannou_2019}.
 

### 2024-08-21 paper/mesa.tex
Implemented Referee Answers to the paper

@@
 Despite being effective, this method is problem dependent, since it needs to make definite assumptions on the shape of the PSD.
-Moreover, it can be computationally expensive and it does not come with a handy implementation available to the public.
+Moreover, it can be computationally expensive.
 For all the above reasons, we did not consider such methods in our work.
@@
 In most case of numerical instability in the `Fast' method, \texttt{memspectrum} will send a warning to user.}.
-A comparison of the computational times for Standard MESA implementation and Fast implementation (together with Welch's) is porvided in appendix \ref{sec:computationaltimes}.
+A comparison of the computational times for Standard MESA implementation and Fast implementation (together with Welch's) is provided in appendix \ref{sec:computationaltimes}.
 
@@
 	\centering
-	\includegraphics[width = \linewidth]{Images/Plots/comparison_loss2}
+	\includegraphics[width = \linewidth]{Images/Plots/comparison_loss1}
 	\caption{Comparison of the ensemble average PSD estimate}
@@
 In our initial qualitative comparison, depicted in Figure \ref{fig:psd_comparison}, we juxtapose the reconstruction of a known a-priori Power Spectral Density with those obtained using the two distinct loss functions.
-The red line in the plot represents our chosen reference PSD, derived from the estimated PSD for the GW150914 event.\\
+The black, dotted line in the plot represents our chosen reference PSD, released together with the GWTC-1 catalog~\cite{GWTC1,PSD_release}, and computed for the LIGO Handford with the BayesLine package \cite{Cornish_2015,Littenberg_2015,Cornish_2020,Chatziioannou_2019}.
+The PSD is the median over the reconstructed posterior distribution for the GW150914 event.
+\\
 To conduct this analysis, we generated 1000 noise time-series whose power spectral densities matches the reference PSD by construction \cite{noiseGen}.
@@
 Ultimately, we compared the reference PSD against the ensemble average of these two estimation methods.\\
-The FPE-derived estimate, represented by the dashed line, effectively identifies and reconstructs peaks across both high and low frequency ranges with commendable accuracy.
+The FPE-derived estimate, represented by the red line line, effectively identifies and reconstructs peaks across both high and low frequency ranges with commendable accuracy.
 However, as illustrated in the inset plots, FPE struggles when confronted with structured peaks—those containing subordinate modes.
@@
 \\
-On the other hand, the VM estimate, depicted as a dot-dashed line, excels in reconstructing both dominant and subordinate modes with remarkable precision.
+On the other hand, the VM estimate, depicted as the continuous green line, excels in reconstructing both dominant and subordinate modes with remarkable precision.
 VM appears to prioritize comprehensive mode reconstruction, while FPE emphasize an accurate reconstruction of major modes while potentially neglecting more intricate sub-peaks.
-
+Additional figures of merit are inserted in appendix \ref{sec:AdditionalPlots} 
 \subsection{How well is the AR order recovered?}\label{sec:arp_validation}
@@
 	\centering
-	\includegraphics[width = \linewidth]{Images/Plots/ar_order2}
+	\includegraphics[width = \linewidth]{Images/Plots/ar_order_new}
 	\caption{Reconstructed value for the autoregressive order plotted against the true value of the autoregressive order.
@@
 \begin{figure}
-	\includegraphics[width = \linewidth]{Images/Plots/P_values_comparison}
-	\caption{Histogram of the P-values obtained with a Kolmogorv Smirnov test on the whitened time series, against a univariate, 0 mean normal distirbuition}
+	\includegraphics[width = \linewidth]{Images/Plots/Anderson_pvalues}
+	\caption{Histogram of the P-values obtained with an Anderson Darling test on the whitened time series, against a univariate, 0 mean normal distirbuition}
 	\label{fig:whitening}
@@
 Each time series underwent the whitening process using the autoregressive filter derived from its corresponding loss function.
-We then evaluated the resulting whitened time series against a zero-mean, univariate normal distribution using the Kolmogorov-Smirnov test.
+We then evaluated the resulting whitened time series against a zero-mean, univariate normal distribution using the Anderson-Darling test.
 \\
@@
 In this region, there is no statistical difference between the two.
-Infact, the total number of counts in this bin are respectively $c_{VM} = 47 \pm 7$ and $c_{FPE} = 42 \pm 6$, affirming the absence of a pronounced discrepancy between the two.
+Infact, the total number of counts $c$ in this bin are respectively $c_{VM} = 100 \pm 10$ and $c_{FPE} = 89 \pm 9$, affirming the absence of a pronounced discrepancy between the two.
 In essence, this final examination underscores a shared proficiency in whitening between the two loss functions, showing that very long filters are not needed to obtain a comparlable result in whitening.
@@
 Although similar studies can be drawn from any other PSD, in this section we focus on a single PSD and we try to generalize some observations that we make.
-We decide to use the analytical PSD computed for the LIGO Handford interferometer, released together with the GWTC-1 catalog~\cite{GWTC1,PSD_release}, and computed with the BayesLine package \cite{Cornish_2015,Littenberg_2015,Cornish_2020,Chatziioannou_2019}.
+We used the same reference PSD used for the comparison of the two losses in the previous section \cite{Cornish_2015,Littenberg_2015,Cornish_2020,Chatziioannou_2019}.
 
@@
 We vary the length of the data used for the estimation: this is also useful to assess how the computation depends on the data available.
-We set the total observation time $T = 1, 5, 10, 100, 1000 \SI{}{s}$
+We set the total observation time $T = 1, 5, 8, 10, 100, 1000 \SI{}{s}$.
+The observation time of $8s$ is inserted since it is the observation time over which the reference PSD is computed.
 For the MESA algorithm, we choose the VM loss function. For the Welch algorithm, we employ a Tukey window with the shape parameter $\alpha$ 
@@
 Although observed at every frequency, this behaviour is more evident in the low frequency region.
-An accurate profile reconstruction can be obtained with MESA using a 5 seconds-strain only, while Welch method requires at least 10 seconds of data to obtain a comparable profile.
+An accurate profile reconstruction can be obtained with MESA using a 5 seconds-strain only,  while Welch method requires at least 10 seconds of data to obtain a comparable profile".
+Such a difference persists when the 8 seconds data strain is considered.
 Furthermore, MESA is able to model all the details of the peak at around $\sim \SI{40}{Hz}$ (even with $T=\SI{100}{s}$), while the Welch's algorithm fails to do so even with an observation time of $T=\SI{1000}{s}$.
@@
 Another faster recursion method is available in \cite{Vos} and it is also available in \texttt{memspectrum}.
+\section{Additional Plots for validation}\label{sec:AdditionalPlots}
+In this section we report additional plots for the Validation.
+Figure \ref{fig:NormalisedDifference} represents the difference between the estimated PSD and the reference PSD, normalised by the latter.
+Figures (\ref{fig:allFPE}) and (\ref{fig:allVM}) show the reference PSD (in red) against all the 1000 reconstruction obtained with FPE and VM respectively.
+\begin{figure}[H]
+	\caption{Normalised difference of reconstructed PSD with respect to the reference PSD}
+	\label{fig:NormalisedDifference}
+	\includegraphics[width=\linewidth,keepaspectratio]{Images/Plots/Normalised_residuals}
+\end{figure}
+\begin{figure}[H]
+	\caption{The 1000 reconstruction of the reference PSD obtained with the FPE loss}
+	\label{fig:allFPE}
+	\includegraphics[width=\linewidth,keepaspectratio]{Images/Plots/All_FPEs}
+\end{figure}
+\begin{figure}[H]
+	\caption{The 1000 reconstruction of the reference PSD obtained with the VM loss}
+	\label{fig:allVM}
+	\includegraphics[width=\linewidth,keepaspectratio]{Images/Plots/ALL_VMs}
+\end{figure}
 

### 2024-08-21 dev/compare_LVC_data.py

@@ -1,5 +1,6 @@
 import sys
 sys.path.insert(0,'..')
+sys.path.insert(0,'../memspectrum')
 from memspectrum import MESA
 from welch import *
 import numpy as np
@@ -9,22 +10,22 @@ from style_sheet import *
 import pandas as pd
 from scipy.interpolate import interp1d
 
-plot = False
+plot = True
 plot_same = True
 compute = False
 generate_fake_data = False
 use_fake_data = True
 
-plot_welch = True #if plot_same is true, this decides whether to plot welch method or MESA
+plot_welch = False #if plot_same is true, this decides whether to plot welch method or MESA
 
 	#folder to save the plot at
 save_folder = '../paper/Images/comparison_LVC_data/'
 true_PSD_file = 'GWTC1_GW150914_PSDs.dat'
 #save_folder = 'comparison_LVC_data' #if on the cluster
-srate = 4096.
+srate = 2048.
 
-T_list = [1, 5, 10,100, 1000] #list of times to make the comparison at
-seglen = [512, 1024, 2048, 8192, 32768]
+T_list = [1, 5, 8, 10, 100, 1000] #list of times to make the comparison at
+seglen = [512, 1024, 2048, 2048, 8192, 32768]
 
 if generate_fake_data:
 	PSD = np.loadtxt(true_PSD_file)
@@ -51,7 +52,8 @@ if compute:
 		data_T = data[-int(srate*T)-4096*5:-4096*5]
 		
 		M = MESA()
-		M.solve(data_T, early_stop = True, method = 'Standard')
+		M.solve(data_T, early_stop = True, optimisation_method = 'VM', method = 'Fast')
+		print(M.get_p())
 		print("\tDone MESA")
 		freqs, PSD_Welch = psd(data_T, srate, seglen[i]/float(srate),
 			window_function  = None,
@@ -72,8 +74,8 @@ if plot_same:
 	fig.set_size_inches(3.4,3.4*2)
 	ax = fig.gca()
 
-	ax.set_xlabel(r"$f(Hz)$")
-	ax.set_ylabel(r"$PSD (a.u.)$") #No ylabel
+	ax.set_xlabel(r"$f(Hz)$", size = 8)
[…]

### 2024-08-21 dev/optimisers_comparison.py

@@ -25,10 +25,12 @@ import os
 load_noise = False 
 load_estimates = True 
 simulate_noise = False
-save = False
+save = True
 plot = True 
-save_plots = False 
+save_plots = True 
 
+import os 
+os.chdir('/home/alessandro/Documenti/PhD/Research/MESA/Maximum-Entropy-Spectrum/dev/')
 #Initialise directories 
 load_dir = 'psd_comparison/LIGO_FPE_VM/'
 save_dir = 'psd_comparison/LIGO_FPE_VM/'
@@ -38,19 +40,20 @@ if save:
         raise FileExistsError("Saving directory doesn't exist. Change saving directory and try again")
 noise_file = 'psd_comparison/noise.npy'
 psd_file = 'psd_comparison/psd.txt'
-plot_savedir = None
+plot_savedir = save_dir
 
 #Initialise analysis methods and MESA object
-methods = ['FPE', 'VM']
+methods = ['FPE', 'VM'][::-1]
 M = MESA() 
 
 #Initialise plot infos 
-colors = {'FPE': 'k', 'VM': 'k'}
+#colors = {'FPE': 'r', 'VM': 'g'}
+colors = {'FPE': 'r', 'VM': 'green'}
 fmts = {'FPE': '.', 'VM': 'o'}
-linestyles = {'FPE': '--', 'VM': 'dashdot'}
-alphas = {'FPE': .9, 'VM': 1}
+linestyles = {'FPE': '-', 'VM': '-'}
+alphas = {'FPE': .9, 'VM': .8}
 errbar_colors = {'FPE': 'gray', 'VM': 'dimgray'}
-lws = {'FPE': 1.3, 'VM': 1.3}
+lws = {'FPE': 1.2, 'VM': 2.5}
 errbar_alphas = {'FPE': 0, 'VM': 0}
 
 #Initialise simulation infos 
@@ -113,43 +116,94 @@ if plot:
     median = {}
     lower_percentiles, lp = {}, 5
     upper_percentiles, up = {}, 95
+    
     for method in methods: 
-        mean[method] = spectra[method].mean(0)
+        mean[method] = np.mean(spectra[method], axis = 0)
         median[method] = np.median(spectra[method], axis = 0)
         lower_percentiles[method], upper_percentiles[method] =\
             np.percentile(spectra[method], (lp, up), axis = 0)
             
-    fig, ax = plt.subplots() 
+    fig, ax = plt.subplots(figsize = (10,5)) 
+
+    
+  
[…]

### 2024-08-21 dev/welch.py

@@ -1,5 +1,6 @@
 # -*- coding: utf-8 -*-
-from scipy.signal import tukey, welch
+from scipy.signal import welch
+from scipy.signal.windows import tukey
 
 def psd(strain,
         sampling_rate,

### 2024-09-06 paper/mesa.tex
added the referee requests

@@
 Let us relax the $N$ hypothesis by defining a \emph{residuals} hypothesis $R$ as $R = D - S$.
-This might seem a very trivial statement, but it has a non trivial application: given  $d(t) = h(t;\theta) + n(t)$ where $h(t;\theta)$ is our signal model, defined by a set of parameters $\theta$,  the residuals $r(t) \equiv d(t) - h(t;\theta)$.
+This might seem a very trivial statement, but it has a non-trivial application: given  $d(t) = h(t;\theta) + n(t)$ where $h(t;\theta)$ is our signal model, defined by a set of parameters $\theta$,  the residuals $r(t) \equiv d(t) - h(t;\theta)$.
 Formally, no reference to the noise process is present anymore.
@@
 Our sampler is a nested sampling algorithm \cite{john_veitch_2020_4109277} and the specific inference model is implemented as part of \texttt{granite}, a dedicated inference model for ground-based interferometric detectors.
-We compare our results with the combined posterior samples available from GWOSC \cite{GWOSC} and available at \linebreak
+We compare our results with the combined posterior samples\footnote{In particular, we compare against the posterior samples given in the file \texttt{IGWN-GWTC2p1-v2-GW150914_095045_PEDataRelease_mixed_nocosmo.h5}.} available from GWOSC \cite{GWOSC} and available at \linebreak
 \texttt{https://zenodo.org/records/6513631}.
+As a comparison, we also performed an analysis using a fixed PSD estimated off-source with MESA, performing an averaging procedure analogous to what if performed in the Welch method.
 
 \begin{figure}
-	\caption{Posterior samples for $\mathcal{M}$ and mass ratio $q$ from the LVK (blue) and using \texttt{memspectrum} (red).
-The samples are largely consistent among the two models, with the MESA model providing a more conservative estimate.}
+	\caption{Posterior samples for $\mathcal{M}$ and mass ratio $q$ from the LVK (blue), using \texttt{memspectrum} to estimate with a fixed PSD (green) and using \texttt{memspectrum} to marginalise over the PSD (red).
+The samples are largely consistent among the models, with the MESA model providing a more conservative estimate.}
 	\label{fig:gw150914_masses}
@@
 \begin{figure}
-	\caption{Posterior samples for the sky position angles from the LVK(blue) and using \texttt{memspectrum}(red).
-The samples are largely consistent among the two models, with the MESA model providing a more conservative estimate.}
+	\caption{Posterior samples for $\mathcal{M}$ and mass ratio $q$ from the LVK (blue), using \texttt{memspectrum} to estimate with a fixed PSD (green) and using \texttt{memspectrum} to marginalise over the PSD (red).
+The samples are largely consistent among the models, with the MESA model providing a more conservative estimate.}
 	\label{fig:gw150914_sky}
@@
 
-In Figs.~\ref{fig:gw150914_masses}, ~\ref{fig:gw150914_sky} and \ref{fig:gw150914_waveforms} we show the posteriors for the set of intrinsic parameters, extrinsic parameters and reconstructed waveform from our analysis.
-Our results can be summarised as follows: our posterior samples are in general consistent with what has been released by the LVK, however our credible regions tend to be larger.
+In Figs.~\ref{fig:gw150914_masses}, ~\ref{fig:gw150914_sky} and \ref{fig:gw150914_waveforms} we show the posteriors for the set of intrinsic parameters, extrinsic parameters and reconstructed waveform from our analyses.
+The posterior samples in red show the results coming from marginalising over the PSD, the ones in green show the results from our fixed PSD analysis, while the ones in blue are the samples release by the LVK.
+While the fixed PSD run show posteriors that are largely consistent with the LVK results, the posteriors from the marginalisation run are still in general consistent with what has been released by the LVK, but showing wider credible regions.
 This is expected since our likelihood includes additional uncertainty due to the explicit sampling over the process order, hence the PSD.
@@
 The full joint posterior distribution recovered when marginalising over the AR order is shown in Appendix~\ref{sec:appendix_b}.
+
+We conclude this section with a comparison with Ref.~\cite{Biscoveanu}, where a similar attempt at characterising the effect of marginalising over the PSD was made.
+The aforementioned work adopted a very different methodology compared to ours, the author use independent draws from the posterior over the PSD as inferred by BayesLine and repeat a standard frequency domain analysis for several signals.
+The results reported in Ref.~\cite{Biscoveanu} show a similar trend to ours: they observe a general widening of the sky position posteriors, and little effect (few \%) on the intrinsic parameters width, while we observe a substantial increase in the width of the chirp mass and mass ratio posteriors.
+We believe this difference to be due both to the very different model we employed as well as on the small (200) number of samples from the PSD they used.
+The estimated uncertainty over the PSD is however consistent, of the order of (5-10\%).
 
@@
 \section{Full posterior distribution for GW150914} \label{sec:appendix_b}
+
+Here we report the full posterior distribution obtained from our PSD marginalisation model.
+The prior posterior distributions assumed are identical to what presented in Ref.~\cite{GWTC21} for all overlapping parameters.
+
 \begin{figure*}
